### 2023-02-07 Code/Times.Rmd
Copied from other repo

---
title: "Thesis"
author: "Owen Fiore"
date: "10/20/2022"
output: pdf_document
---

Reading in the Data
```{r}
times <- read.csv("../Data/ThesisData.csv")
View(times)
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
```

#DataSets
```{r}
men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
women_finals <- subset(times, Type == "F" & Gender == "F" & Time != 'DQ')

men_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
women_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "F" & Time != 'DQ')
```

#Models
```{r}
library(lme4)
model1 <- lmer(RxnTime ~ (1|Year), data = men_finals)
#finals.log <- lmer(log(RxnTime) ~ (1|Year), data = men_finals)
model0 <- lm(log(RxnTime) ~ 1, data = men_finals)
finals.gamma <- glmer(RxnTime ~ (1 | Year), data = men_finals, family = Gamma)
fit2 <- glmer(RxnTime ~ (1 | Year), data = men_finals, family = Gamma(link = "log"))
fit0<- glm(RxnTime ~ 1, data = men_finals, family = Gamma(link = "log"))
```

```{r}
model0 <- glm(log(RxnTime) ~ 1, data = men_pooled)
model1 <- lmer(log(RxnTime) ~ (1|Year), data = men_pooled)

#anova(model0, model1)
AIC(model0)
AIC(model1)
anova(fit0, fit2)
#AIC(fit0)
#AIC(fit2)
```

```{r}
qqnorm(resid(model0), main="Linear model")
qqnorm(resid(model1), main = "Log model")
qq(resid(finals.gamma),main="Gamma model no log")
qqnorm(resid(fit2), main = "Gamma model with log")
qqnorm(resid(fit0), main = "Gamma model with log without year effect")
#anova(finals.lme, finals.log, finals.gamma, fit2)
anova(model0, model1)
```

#Matched pairs t test

t test comparing 2019 and 2022
```{r}
t.test(times$RxnTime[times$Year==2019],times$RxnTime[times$Year==2022])
```

Creates a Boxplot comparing 2019 and 2022
```{r}
#boxplot(times$RxnTime[times$Year==2019],times$RxnTime[times$Year==2022])
men_final <- subset(times, Gender == "M" & Type == "F")
women_final <- subset(times, Gender == "F" & Type == "F")
plot(RxnTime ~ Year, data = men_final, xlab = "Years")
plot(RxnTime ~ Year, data = women_final, xlab = "Years")

```

Creates a new data set of just finals and looks at the data from that
```{r}
library(ggplot2)
finals <- subset(times, Type == "F")
boxplot(RxnTime ~ Year, data = men_pooled, xlab = "Year", ylab = "Reaction Time in seconds", Main = "Men's Semifinal and Final Reaction Time Data")
```

```{r}
library(ggplot2)
plot(aggregate(men_pooled$RxnTime, list(men_pooled$Year), FUN = mean), xlab = "Year", ylab = "Reaction time in seconds", main = "Mean Reaction Time By Year")
```
```{r}
library(lme4)
finals.log <- lmer(log(RxnTime) ~ (1|Year), data = men_final)
#pdf(file = "Skew.pdf")
hist(resid(finals.log), xlab = "Residuals of Linear mixed effects model", main = "Right skew when plotting the residuals of the linear mixed effects model")
#dev.off()
```

```{r}
library(dplyr)
finals.aov <- aov(RxnTime ~ as.factor(Year), data = finals)
summary(finals.aov)
tuk.final <- TukeyHSD(finals.aov, conf.level = .95)
tkf <- as.data.frame(tuk.final[1])
tkf %>%filter(as.factor.Year..p.adj < 0.05)
#plot(tuk.final)
```

```{r}
residuals <- resid(finals.aov)
plot(finals$RxnTime, residuals, xlab="RxnTime", ylab="Residuals")
abline(0,0)
```

```{r}
library(lme4)
finals.log <- lmer(log(RxnTime) ~ (1|Year), data = men_final)
finals.lme <- lmer(RxnTime ~ (1|Year), data = men_final)
summary(finals.lme)
qqnorm(resid(finals.lme))
hist(resid(finals.log))
hist(resid(finals.lme))
qqnorm(resid(finals.log))
pnorm(0.1, mean = .148883, sd = sqrt(8.648*10^(-5)+3.719*10^(-4)))
qnorm(0.01,mean = .148883, sd = sqrt(8.648*10^(-5)+3.719*10^(-4)))
```

```{r}
library(hglm)
gamma.GLMM <- hglm(fixed = RxnTime ~ 1, random =   ~ 1|Year , data = finals,
                     family = Gamma(link = "log"), rand.family = gaussian(link = "identity"))
summary(gamma.GLMM)
plot(gamma.GLMM)
resid(gamma.GLMM)
```

```{r}
library(hglm)
gamma.conj <- hglm(fixed = RxnTime ~ Year, random =  ~ 1 | Year, data = finals,
                     family = Gamma(link = "inverse"), rand.family = inverse.gamma(link= "inverse"))
summary(gamma.conj)
plot(gamma.conj)

```

```{r}
library(hglm)
gamma.GLMM <- hglm(fixed = RxnTime ~ Year, random =  ~ 1 | Year, data = finals,
                     family = Gamma(link = "log"), rand.family = Gamma(link= "log"))
plot(gamma.GLMM, dev = "pdf", name = "First Gamma Model")
summary(gamma.GLMM)

```

```{r}
library(zeligverse)
z.out <- zelig(formula= ~ 1 + tag(RxnTime | Year),
data= finals, model="gamma.mixed")
```

```{}
library(lme4)
semi <- subset(times, Type == "S")
finals.lme <- lmer(RxnTime ~ (1|Year), data = semi)
summary(finals.lme)
qqnorm(resid(finals.lme))
hist(resid(finals.lme))
qqnorm(ranef(finals.lme))
```
```{}
library(lme4)
heat <- subset(times, Type == "H")
finals.lme <- lmer(RxnTime ~ (1|Year), data = heat)
summary(finals.lme)
qqnorm(resid(finals.lme))
hist(resid(finals.lme))
qqnorm(ranef(finals.lme))
str(ranef(finals.lme))
hist(ranef(finals.lme)$Year[,1])
qqnorm(ranef(finals.lme)$Year[,1])
qqline(ranef(finals.lme)$Year[,1])
```
```{r}
pnorm(0.1, mean = .157750, sd = sqrt(.0001228+0.0005937))
qnorm(0.01, mean = .157750, sd = sqrt(.0001228+0.0005937))
```

```{r}
my_fit <- lm(RxnTime ~ 1 , data = finals)
summary(my_fit)
```

```{r}
anova(finals.lme, my_fit)
```

```{r}
old <- subset(times, Year < 2020)
old.lme <- lmer(RxnTime ~ (1|Year), data = old)
summary(old.lme)
old_fit <- lm(RxnTime ~ 1 , data = old)
anova(old.lme, old_fit)

```

```{r}
model_coefs <- coef(old.lme)$Year

model_coefs
```

Plotting LME
```{r}
library(ggplot2)
model_coef_plot <- ggplot(data = old, 
       mapping = aes(x = Time, 
                     y = RxnTime, 
                     colour = Year)
       ) +
  scale_y_continuous(limits = c(180, 1020)) +
  scale_x_continuous(breaks = seq(1:10) - 1) +
  theme(legend.position = "top")

model_coef_plot
```

t test comparing men and women times in 2022
```{r}
men <- subset(times, Gender == "M")
women <- subset(times, Gender == "F")
t.test(men$RxnTime, women$RxnTime)

```

### 2023-02-07 Code/lme4.R

@@ -1,0 +1,160 @@
+library(lme4)
+library(ggplot2)
+
+times <- read.csv("../Data/ThesisData.csv")
+colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
+
+## only use Finals, Male, not Disqualified runs, before 2022
+men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
+## boxplot(RxnTime ~ Year, data = finals, xlab = "Years")
+
+library(lme4)
+fit1 <- glmer(RxnTime ~ (1 | Year), data = men_finals, family = Gamma)
+fit2 <- glmer(RxnTime ~ (1 | Year), data = men_finals, family = Gamma(link = "log"))
+fit0 <- glm(RxnTime ~ 1, data = men_finals, family = Gamma(link = "log"))
+
+qqnorm(resid(fit2))
+
+## qqnorm(ranef(fit2)$Year[,1])
+## qqline(ranef(fit2)$Year[,1])
+
+
+## Fitted results
+sfit0 <- summary(fit0)
+pgamma(0.1, shape = 1 / sfit0$dispersion, scale = exp(coef(fit0)) * sfit0$dispersion)
+
+
+simfit <- function(fit2, n = 12000000) {
+  vc <- as.data.frame(VarCorr(fit2))
+  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
+  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
+  mu <- exp(fixef(fit2) + z)
+  shape <- 1 / vc$vcov[2]
+  x <- rgamma(n, shape = shape, scale = mu / shape)
+  x
+}
+
+x <- simfit(fit2)
+mean(x < 0.1)
+quantile(x, prob = c(.00001, .0001, .0005, .001))
+## compare with and without random effect
+plot(density(x)) # with random effect
+curve(dgamma(x, shape = 1 / sfit0$dispersion,
+             scale = exp(coef(fit0)) * sfit0$dispersion),
+      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
+
+## Does it make sense to pool the genders?
+women_finals <- subset(times, Type == "F" & Gender == "F" & Time != 'DQ')
+fit1 <- glmer(RxnTime ~ (1 | Year), data = women_finals, family = Gamma)
[…]

### 2023-02-07 Manuscript/manuscript.tex
Copied from other repo

\documentclass[12pt, letterpaper, titlepage]{article}

\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{cleveref}
\hypersetup{colorlinks = true, linkcolor = blue, citecolor=blue, urlcolor = blue}
\usepackage{natbib}
\usepackage{float}
\usepackage{setspace}
\usepackage{pdfpages}
\usepackage[pagewise]{lineno}
\usepackage{mwe}
%\linenumbers*[1]
% %% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
 \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
 \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
 \renewenvironment{#1}%
 {\linenomath\csname old#1\endcsname}%
 {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
 \patchAmsMathEnvironmentForLineno{#1}%
 \patchAmsMathEnvironmentForLineno{#1*}}%

\AtBeginDocument{%
 \patchBothAmsMathEnvironmentsForLineno{equation}%
 \patchBothAmsMathEnvironmentsForLineno{align}%
 \patchBothAmsMathEnvironmentsForLineno{flalign}%
 \patchBothAmsMathEnvironmentsForLineno{alignat}%
 \patchBothAmsMathEnvironmentsForLineno{gather}%
 \patchBothAmsMathEnvironmentsForLineno{multline}%
}

% control floats
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

\newcommand{\jy}[1]{\textcolor{blue}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{red}{EDS: (#1)}}

\title{Was Devon Allen Unjustly Disqualified at the 2022 World Track and Field Championships?}

\author{Owen Fiore\\
%   \href{mailto:owen.fiore@uconn.edu}
% {\nolinkurl{owen.fiore@uconn.edu}}\\
  Elizabeth Schifano\\
  Jun Yan\\[1ex]
  Department of Statistics, University of Connecticut\\
}
\date{}

\begin{document}
\maketitle

\doublespace

\begin{abstract}
  At the 2022 World Track Championships in Eugene Oregon, Devon Allen was disqualified in the
  Men's 110 meter hurdle final after registering a reaction time of 0.099 seconds, 0.001 
  seconds faster than what is allowed.  Following the games, bloggers on the running website 
  LetsRun scrutinized the results and performed non-statistical analysis of the data \citep{Johnson}
  They found that the reaction time data from the 2022 World Track Championships seemed to be faster 
  compared to the other datasets they looked at, but did not perform any meaningful statistical 
  analysis of the data.  Devon Allen had previously ran the third fastest time in history when 
  he ran 12.84 earlier in 2022 and was considered to be one of the favorites to win \citep{Preview}.  His 
  disqualification thus prompted questions into whether the timing equipment used at the World
  Track Championships was accurate and whether or not Allen actually reacted as fast as the 
  timing equipment stated he did.

%\bigskip
\noindent\sc{Keywords}:
Devon Allen, Reaction Time, Seiko, 2022 World Track and Field Championships

\end{abstract}

\section{Introduction}
\label{sec:intro}

Other scholarly papers have discussed track and field reaction times but not in the
context of whether the minimum reaction time barrier is too low or whether or not timing
systems were accurate.  Outside of the blog posts from LetsRun.com, there have not been
any published papers regarding Devon Allen's disqualification in particular.  This is not
neccessarily suprising given that the World Track and Field Championships occurred only 
four months ago.

The contribution of this paper is to question whether or not Seiko's timing data
was accurate and thus if Devon Allen was disqualified more so because of faulty
equipment rather than reacting illegally fast.  There is little at this point
that could be done to rectify any potential mis-qualification other than to try
to prevent it from happening again.  THis matter needs to be addressed because Devon Allen's 
disqulification could be a problem that could continue to re-appear.  If athletes
are reacting fast enough that the timing system thinks that they are reacting
illegally fast, then World Athletics needs to address this and possibly change
the minimum reaction time allowed.  Currently the barrier is 0.10 seconds, any time
faster than this would result in a disqualification in the athlete who reacted too
quickly and the race is restarted \citep{Seiko-Timing}.  Since 2011, any athletes who false 
start are not given a second chance, they are removed from competition \citep{False-Start}.  
This can have massive implications as events like the Outdoor World Track and Field Championship 
only occur typically once every two years.  It is an opportunity for athletes to earn endorsement
deals and prove to sponsors such as Nike and Adidas that
they deserve a contract.  Thus the significance of a disqualification, especially in the finals 
of a major event such as the World Track and Field Championships, are very
consequential.

\section{Background}
\label{sec:Background}
Since 1985 Seiko Holding Corporations has served every World Athletics Championship
as the official timer \citep{Seiko}.  Since 1985, the technology and the ability
to accurately predict measurements: long jump distances, false starts, total time,
reaction time, etc. have all dramatically improved.  Seiko did not start tracking
reaction time as an official measurement until 1999 in the men's and women's 100 meter
hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
technology to the World Track and Field Championships, the highest stakes in the world
of running outside of the Olympics.

%Citation needed in this section
Seiko's technology for detecting reaction times relies on the pressure that athletes
exert on the plate when they push off.  Their systems measure the time differential
between when the "gun" goes off to start the race and when the pressure changes 
\citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
false start as it is considered that no athlete can react so quickly \citep{Seiko-Timing}.  
Thus, a reaction time of for example 0.05, suggests that the athlete predicted the gun and it
was luck that caused their abnormally high time.  At the World Championship level,
World Athletics and Seiko want to remove that element of luck and thus impose the 0.1
second barrier.

In 2013, Seiko upgraded their timing equipment for sprinting events (100 meter hurdles
falls under this catergory) \citep{WorldAthletics_2013}.  Since 1999, the highest reaction 
times in the men's 100 meter hurdle were recorded in 2013.  That is not to say that the higher 
times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
to 2022, Seiko again upgraded its technology but for its jump management system.

Something that is extremely important to note in the Data is the inconsistency of World Athletics 
in terms of how they penalize false starts.  From 2007-2009, the IAAF (the former name for World 
Athletics, and the governing body for the World Track and Field Champions) instituted a rule change
that allowed one warning false start before a sprinter was disqualified \citep{False-Start}  This 
essentially gave sprinters a second chance and reduced the penalty for a false start.  There were 
18 male and 7 female false starts at the 2007 World Championships, 18 male and 7 female starts at 
the 2009 World Championships. Starting in 2011, the rule was scrapped, and the old policy which 
automatically disqualifies runners who false start was reinsituted.  The effect? At the 2011 World
Championship there were 6 male disqualifications and 4 female disqualifications \citep{False-Start}. 
By returning to the harsher policy and cracking down on false starts, World Athletics had reduced the
number of false starts by two thirds. This was desirable for World Athletics as false starts can 
make already long track meets more tedious, both for athletes and viewers watching the 
television broadcast.\citep{Pilianidis}

\section{Data}
\label{sec:Data}
Data was copied from the World Athletics website and pasted into an excel
spreadsheet. The data covers the men's 110 meter hurdles and the women's 100
meter hurdles from 1999 to 2022.  The variable of prediction interest is reaction
time, and the predicting variables are year, stage of competition (heats, 
semifinal, final), total time, and gender.  The data can be found in the appendix:.  
Factors such as name, country, and identification number were
not considered important for the purposes of this research besides the times for
Devon Allen and the other United States athletes mentioned above.  The variable "Stage"
refers to whether the observation occurred during the "heats" (Preliminary round) which
is denoted by "H" in the data, "S" for "semifinals" (Of which there are either 2 or 3 
heats each year), and "F" for finals. The variable "Gender" has levels "M" and "F" 
which stand for "male" and "female" respectively.  The variable "Batch" takes a 
numerical value and every heat of every round from 1999 to 2022 was assigned it's
own number.  This was added because of suspected issues that reaction time may be
correlated with other runners reacting quickly (one runner reacting quickly may
result in other runners reacting quickly).  There is no statistical signficance to
how the batches were labeled, they were ordered chronologically by meet but
reverse chronologically by year (Batch 1 is a preliminary Heat of 2022 and Batch
111 corresponds to the 1999 finals). The data that looked at in the paper
is best visualized using a scatterplot and box plot. Figure~\ref{fig:DataBoxPlot}
below shows the box plot data.
\begin{figure}[h]
  \centering 
  \includegraphics[scale = 0.5]{DataBoxPlot}
  \label{fig:DataBoxPlot}
\end{figure}

\begin{figure}[h]
  \centering 
  \includegraphics[scale = 0.5]{DataScatterPlot}
  \label{fig:DataScatterPlot}
\end{figure}

One of the most interesting things found was that the reaction times from the USA
athletes in the World Championship 110 meter hurdles, was that they were significantly
faster than at the USA Track Championships.  From June 23-26 2022, USA held it's Track
and Field Championship to decide who to send to the World Championships at Hayward 
Field, the exact same venue as used for the World Track and Field Championship a month 
later.  Thus, a baseline is able to be established for the four USA athletes who 
competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
Grant Holloway, and Devon Allen. Every athlete reacted faster in all of the World 
Track and Field Championship races compared to the USA Track and Field 
Championship races. Devon Allen raced three times at the USA Track and
Field Championship and three times at the World Track and Field  Championship. 
His reaction times were significantly faster in the World Track and
Field Championship compared to the USA Track and Field Championship as were his
teamates.  This comparison suggests that the timing system used between the two
meets may have been meaningfully different to the point where Devon Allen was
disqualified more because of faulty equipment rather than because he reacted
too quickly.  This table below highlights the difference in reaction times for
Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. "USA" denotes
the USA Track and Field Championships, "World" denotes the World Track and Field
Championships, "N/A" was used for any athlete that did not compete and thus
did not register a reaction time.  All numbers listed in the table are the reaction
time for the respective athletes in seconds.

\begin{center}
  \begin{tabular}{||c | c c c | c c c||} 
   \hline
   Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
   \hline\hline
   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
   \hline
   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
   \hline
   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
   \hline
   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
   \hline
  \end{tabular}
  \end{center}

\section{Methods}
\label{sec:methods}

The main statistical model used in the paper is a gamma mixed effects model, which is
in the family of Generalized Linear Mixed Model (GLMM). The GLMM comes from the R package 
lme4, which hosts a variety of GLMM functions \citep{Rpkg:lme4}.  A GLMM is a type of model
that supports inputs that are from different distributions.  The model that we use is for the
Gamma distribution for the fixed effects part of the model and a random effects part.  In
the model used in the paper, the fixed effect comes in the form of an intercept, and then
either a year effect, batch effect, or both are used.  Initially a linear mixed effects
model was attempted to be implemented, but due to a right hand skew in the reaction times, 
a gamma distribution seemed better equipped to model the data (The rightskew is not 
particularly suprising as it is not possible to register a reaction time below 0.0 
seconds, resulting in a lower bound but no upper bound for outliers). In the code, the 
link parameter under Gamma() was changed to be "log" as logarithmic functions are better 
equipped to deal with high outliers.  The graph below shows the skewed outliers: ~\ref{fig:Skew}.
\begin{figure}[h]
  \centering 
  \includegraphics[scale = 0.6]{Skew}
  \label{fig:Skew}
\end{figure}

Once the Gamma mixed effects model was chosen, it needed to be decided which combination
of effects to use.  It was possible to include a year effect, a batch effect, or a year
and batch effect.  Density plots were created that simulated the probability of a mean
density less than 0.1, which is the current threshold for a reaction time
disqualification.  The density plots are shown below.

\begin{figure}[h]
  \centering
  \begin{minipage}{0.45\textwidth}
      \centering
      \includegraphics[width=0.9\textwidth]{Men_Density_Year} % first figure itself
      \caption{Density plot using pooled men's data and the year effect}
  \end{minipage}\hfill
  \begin{minipage}{0.45\textwidth}
      \centering
      \includegraphics[width=0.9\textwidth]{Women_Density_Year} % second figure itself
      \caption{Density plot using pooled women's data and the year effect}
  \end{minipage}
\end{figure}

\begin{figure}[h]
  \centering
  \begin{minipage}{0.45\textwidth}
      \centering
      \includegraphics[width=0.9\textwidth]{Men_Density_Batch} % first figure itself
      \caption{Density plot using pooled men's data and the batch effect}
  \end{minipage}\hfill
  \begin{minipage}{0.45\textwidth}
      \centering
      \includegraphics[width=0.9\textwidth]{Men_Density_Both} % second figure itself
      \caption{Density plot using pooled men's data and both year and batch effect}
  \end{minipage}
\end{figure}

There are several different data sets that could be used and explored in determining 
whether or not Devon Allen was unjustly disqualified.  One of the best examples of this
is the women's 100 meter hurdles. Although the distance of the race is 10 meters shorter
than the men's race, the reaction time data is still important and can be used to
essentially double the data size.  However, one of the largest issues with this is that on average, men are able to react quicker than women.  A study from the 2008 Beijing Olympics found that the p value of comparing the mean fastest reaction time between men and women was below 0.001 and the differnce is in part because of the greater force threshold men are able to exert on the starting blocks \citep{Lipps}. The data from including both men and women could follow a bimodal distribution which would be difficult to model and not helpful for determining if Devon Allen's reaction time was the result of a faulty system or if it is an argument
for changing the 0.10 second minimum barrier that currently exists.  Thus it was decided
to look at men's and women's data separately.  Another decision that was made was to look
at the data from the semifinals and finals together but from the heats or preliminary rounds.  The reason for this was because it was reasoned that in the finals athletes may
be a little more anxious, more "jumpy" and that under stress and they would react faster.
Additionally, exceptionally strong atheletes may not race as hard in the early heats in
order to reduce the risk for injury and to save themselves for the later stages of
competition that matter more. While the preliminary rounds also provide a bulk of the total observations, they are not as relevant to Allen's disqualification, which occured during the finals.  Thus, the decision was made to look at data from the men's semifinals
and finals in a "pooled" data set.
%%%Include picture

\section{Conclusion}
\label{sec:conclusion}
It is worth noting that in the semi-finals of the World Track and
Field Championships; Allen's reaction time was 0.101 which is only 0.001 above
the legal limit.  What this suggests is that Allen may have strong reflexes
and be able to react extremely well to the sound of the gun.  It seems unlikely that
Allen would be able to correctly predict the gun with such precision two times in a row,
which is the exact reason that the 0.10 second reaction time disqualification barrier was
imposed in the first place.  But Allen showed that at the Championship he was able to
react extremely fast two times, which may suggest skill rather than luck. Devon Allen is a very good athlete, as evidenced by his football career at the University of Oregon
and making it onto the Philadelphia Eagles practice squad this past year \citep{Hurley}.
Thus the 0.099 reaction time may not have been a product of Devon Allen
predicting the start of the race but rather a combination of a quick reaction 
and a possibly faulty sensor.

While this sample is too small in order
to perform a meaningful matched pairs t test, or other similar method, it is apparent
that Allen, Cunningham, and Holloway's times are faster at the World Track and Field
Championship compared to the United States.  There does not appear to be a particular
reason for this, as at the World Track and Field Championship, no two United States
athetes were in the same batch in any race.

An anova test that compared the mixed effects model for only a batch effect with the
model that included both a batch and year effect produced a p value of 0.01059, meaning
that the model with both mixed effects was stronger, but not significantly so.  However,
when looking at the density plots, the probability of observing a reaction time of below
0.1 was found to be approximately zero when looking at the model that inclued both effects,
but the probability was 0.0066 when only the batch effect was included.  Based on the density
plot and how close the both effects model was to being statistically signficant at an alpha
value of 0.05, the model with both the batch and year effect was chosen to be analyzed further.
The random effects: Batch and Year both had relatively small variances: 0.003455 and 0.001714
respectively.  For the fixed effect the estimate was -1.87968 and the p value was less than
$2\cdot10{-16}$, which is extremely small.  The model can reparameterize the gamma distribution,
which is calculated by $e^{-1.87968}$ which is equal to 0.152.  The variance of the gamma
distribution is 0.0329 from the R output.  Thus a gamma distribution can be created from
this, and the probability of a reaction time below 0.1 is essentially zero.

The model accuracy improves when certain batches are removed from the data.  Batch 41 in
particular has extremely high reaction times.  This outlier can likely be explained by a
few different causes.  Batch 41 is a semifinals heat of the 2013 World Championship.  2013 had the highest average reaction time from the pooled data set.  2013 was also the second world championship after the rule change that imposed the one strike policy on
disqualifications. 2011 and 2013 had the highest average reaction time in the pooled
data and were the two years after the rule change.  What this suggests is that there could have been a psychological effect of the stricter policy that resulted in athletes being more cautious and reacting slower than they had in the previous two years under the more lax guidelines from 2007 to 2009 in order to avoid a false start and thus a disqualification.
In Batch 41 particularly two runners did not finish, and two runners reacted extremely slow: 0.327 and 0.397 (Remember that Devon Allen's reaction time was
0.099 seconds in the 2022 Championship final and 0.101 in the 2022 semifinal).  Removing this batch and others could greatly improve the strength of the model.  However, while
it is a high outlier, there is not sufficient information available as to why the
reaction times are so high, and no basis for why it should be improved other than because
it is an outlier.

\begin{figure}[h]
  \centering 
  \includegraphics[scale = 0.6]{TBY}
  \label{fig:TBY}
\end{figure}

The figure below plots the average reaction time by year for the pooled men's data since
1999.  The circle in the bottom right is for 2022, which is signifcantly lower than the
data points for the previous three years, and is substially lower than the data points
for 2011 and 2013.  This graph supports the conclusion that the timing system from 2022
had some sort of error that caused the reaction times to be so much lower.  Assuming there
was not an issue with the system, the large variation is confusing because there is little explanation
as to why there is so much of a drop off in reaction time from 2013 to 2015 or the spike
from 2001 to 2003.

Devon Allen's disqulification at the World Track and Field Championships was
possibly the result of both a faulty timing equipment and Devon Allen
reacting extremely quickly to the start.  The code and data from R showed that
the 2022 World Track and Field Championships were a low outlier compared to
every other year in terms of average reaction time.  Addtionally, it was shown
through a gamma mixed effects model that there are significant year and batch
effects that greatly impact reaction time.  However, this practically does not make much sense
as there is no reason for so much random variation without much of a trend.
There has not been a consistent decrease or increase since 1999, much of the data
for reaction times has been random and unpredictable.  Thus while it seems easy
to conclude that Devon Allen was wrongly disqualified, that may not neccessarily
be the case.  If there are issues with reaction time such as athletes approaching the 0.10 
second barrier again at the next World Track Championship (August 2023), then World Athletics 
needs to adjust that threshold and allow for faster reaction times so that athletes with superb 
reflexes are not penalized like how Allen was in 2022.  World Athletics and Seiko 
may want to do further testing of their own to determine what is a human's physically
fastest reaction time, and if they choose to do so, they should start with Devon
Allen.

\section{Appendix}
\label{sec:appendix}
%Put downloadable data spreadsheet
Here is World Athletics website with the data: \url{https://www.worldathletics.org/results/world-athletics-championships}.
Here is the data for the USA Track and Field Championships: \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}

\bibliographystyle{chicago}
\bibliography{citations.bib}

\end{document}

### 2023-02-07 Proposal/Proposal.tex
Copied from other repo

\documentclass[12pt]{article}

%% preamble: Keep it clean; only include those you need
\usepackage{amsmath}
\usepackage[margin = 1in]{geometry}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{natbib}

% highlighting hyper links
\usepackage[colorlinks=true, citecolor=blue]{hyperref}

\title{Was Devon Allen Unjustly Disqualified at the World Track and Field Championships in 2022?}

% Add author information below. Communicating author is indicated by an asterisk, the affiliation is shown by superscripted lower case letter if several affiliations need to be noted.
\author{
Owen Fiore
}

\pagestyle{empty}

\begin{document}
\maketitle

\section{Introduction}

At the 2022 World Track Championships in Eugene Oregon, Devon Allen was disqualified in the
Men’s 110 meter hurdle final after registering a reaction time of 0.099 seconds, 0.001 
seconds faster than what is allowed.  Following the games, bloggers on the running website 
LetsRun scrutinized the results and performed non-statistical analysis of the data \citep{Johnson}
They found that the reaction time data from the 2022 World Track Championships seemed to be faster 
compared to the other datasets they looked at, but did not perform any meaningful statistical 
analysis of the data.  Devon Allen had previously ran the third fastest time in history when 
he ran 12.84 earlier in 2022 and was considered to be one of the favorites to win \citep{Preview}.  His 
disqualification thus prompted questions into whether the timing equipment used at the World
Track Championships was accurate and whether or not Allen actually reacted as fast as the 
timing equipment stated he did.

\section{Specific Aims}
I plan to investigate whether the timing equipment used at the Track and Field World Championships
 produced reaction times different than what times typically are.  Since 1985, Seiko has timed 
 every World Championships with 2022 being no exception\citep{Seiko}.  Thus it would be surprising to discover
 if the timing equipment was producing faster than actual times as the same company was responsible
 for timing the Championships.  By comparing the timing data from 2022 to other World Championship 
 meets, it will be established whether 2022 was an outlier and whether Devon Allen was wrongly 
 disqualified.

\section{Data}

The data set I will be looking at comes from the World Athletics Website which has recorded data
for reaction times in the Men’s 110 meter hurdle since 1999. There are over 800 observations for
reaction times in the heats, semi-finals, and finals for the hurdles since 1999 and there were
67 in 2022.  I intend to look at variables such as year, stage of competition (heats vs finals),
reaction time, total time, and if possible attempt to find environmental data looking at 
temperature, humidity, and elevation.  The data can be found here: \citet{WAData}

\section{Methods}
I plan to use Tukey’s Honestly Significant Difference test to compare pairs of data by year
going back to 1999 to see if any other World Track Championship had significantly different 
results.  Using linear prediction regression in R, I intend to look at whether for each 
championship whether the reaction times got faster in each successive stage of competition.
By comparing data from the heats to the semifinals to the finals, I hope to get a better 
understanding of any confounding variables.  There is also data available for the USA Track
Championships which occurred in June of 2022.  By comparing runners who ran in both the USATF
Championships and the World Championships, a matched pairs t test can be conducted.  Lastly, I
will use a linear regression model that looks at reaction time and the following independent
variables: Year, final time, stage of competition, and any data I can find on the weather at the
meet.

\section{Discussion}
I hope to find data that shows that the 2022 World Championships reaction times were significantly
lower than any other World Championship and that the most significant predictor in determining
reaction time is the year the data was taken.  I have already looked at a very brief sample set
comparing the data from 2019 to 2022 and found that the 2022 reaction times were statistically
significantly faster than in 2019.  If my investigation results in evidence that suggests that
Devon Allen was disqualified due to faulty data, and my work is published I hope it changes the
IAAF ruling of immediately disqualifying all of those who false start, especially under such a
small margin.  If the data is not what I expect, then I do not anticipate much change to occur as
a result.

\bibliographystyle{chicago}
\bibliography{citations.bib}

\end{document}

### 2023-02-07 Code/Models.Rmd
Various models

---
title: "Models"
author: "Owen Fiore"
date: "2023-02-07"
output: pdf_document
---

```{r}
times <- read.csv("../Data/ThesisData.csv")
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
```

```{r}
men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
men_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
```

```{r}
old_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ' & Year < 2022)
old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Year < 2022)
```

##Models
```{r}
men_finals_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_finals)
men_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_pooled)
old_finals_yeareffect <- lmer(RxnTime ~ (1|Year), data = old_finals)
old_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = old_pooled)
AIC(men_finals_yeareffect)
AIC(old_finals_yeareffect)
AIC(men_pooled_yeareffect)
AIC(old_pooled_yeareffect)
```

```{}
Thus this seems to support that the 2022 data should be included and that the semifinals and finals should be pooled together.
```

```{r}
trial <- subset(times, Gender == "M" & Time != 'DQ')
x <- lmer(RxnTime ~ (1|Year), data = trial)
AIC(x)
```

```{}
This shows that increasing the data set increases AIC, unsure what we want to use
```

```{r}
#men_finals_randeffect <- lmer(RxnTime ~ 1, data = men_pooled)
men_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_pooled)
men_pooled_randeffect <- glm(RxnTime ~ 1, data = men_pooled)

anova(men_pooled_yeareffect, men_pooled_randeffect)

```
```{}
As the year effect model is signficant at all alpha levels and the AIC is lower, the year effect is significant and should be considered in the model.  Now we compare the log model to the year effect model
```

```{r}
men_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_pooled)
men_pooled_log_year <- lmer(log(RxnTime) ~ (1|Year), data = men_pooled)
men_pooled_log <- glm(log(RxnTime) ~ 1, data = men_pooled)
anova(men_pooled_yeareffect, men_pooled_log_year, men_pooled_log)
```
```{}
The year effect model is better than both the log random effect model and the log year effect model.
```

```{r}
library(lme4)
men_pooled_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma)
men_pooled_log_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
men_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_pooled)

#Compare how much the link=log matters
anova(men_pooled_gamma_year, men_pooled_log_gamma_year)

#Compare gamma and linear
anova(men_pooled_gamma_year, men_pooled_yeareffect)

```
```{}
Whether or not the model contains "(link = "log")" seems to have very little effect on the strength of the model.  The gamma model seems to perform better than the linear model based on AIC, but due to the gamma model being more difficult to interpret, the linear mixed effects model seems sufficient.
```

##Simulations
```{r}
summary(men_pooled_gamma_year)
qqnorm(resid(men_pooled_yeareffect))
qqline(resid(men_pooled_yeareffect))
qqGamma(resid(men_pooled_gamma_year))
```

```{r}
library(qpToolkit)
qqGamma(resid(men_pooled_gamma_year))
```

```{r}
library(DHARMa)
men_pooled_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
summary(men_pooled_gamma_year)
```
```{r}
simout_gamma <- simulateResiduals(men_pooled_gamma_year, n = 1000)
simout_linear <- simulateResiduals(men_pooled_yeareffect, n = 1000)
plot(simout_gamma)
plot(simout_linear)
#summary(simout)
#str(simout)

```

### 2023-02-07 Code/Models.Rmd
Included clusrank comparisons

@@
 ```
+
+##2019 vs 2022 Comparisons
+```{r}
+comp <- read.csv("../Data/2019vs2022.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(comp) <- c("Athlete", "Country", "NineteenH", "NineteenS", "NineteenF", "TwentyTwoH", "TwentyTwoS", "TwentyTwoF")
+#View(comp)
+```
+
+```{r}
+library(clusrank)
+t.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
+clusWilcox.test(comp$NineteenH, cluster=comp$TwentyTwoH, paired = TRUE, alternative = "greater")
+wilcox.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
+```
+
+```{r}
+t.test(comp$NineteenS, comp$TwentyTwoS, paired = TRUE)
+clusWilcox.test(comp$NineteenS, cluster=comp$TwentyTwoS, paired = TRUE, alternative = "greater")
+wilcox.test(comp$NineteenS, comp$TwentyTwoS, paired = TRUE)
+```
+
+```{}
+A matched pairs t test, clusrank package Wilcox test, and a standard Wilcox test all showed that the heat means differ by a significant amount.
+And two of the three semifinals means had an alpha less than 0.01.
+```
+
+```{r}
+nations <- read.csv("../Data/NatVSWorld.csv")
+colnames(nations) <- c("Athlete", "Country", "NatH", "NatS", "NatF", "WorldH", "WorldS", "WorldF")
+```
+
+```{r}
+library(clusrank)
+t.test(nations$NatH, nations$WorldH, paired = TRUE)
+clusWilcox.test(nations$NatH, cluster=nations$WorldH, paired = TRUE, alternative = "greater")
+wilcox.test(nations$NatH, nations$WorldH, paired = TRUE)
+```
+```{}
+These p values all significant at alpha = 0.05, but not at alpha = 0.01
+```
+
+```{r}
+library(clusrank)
+t.test(nations$NatS, nations$WorldS, paired = TRUE)
+clusWilcox.test(nations$NatS, cluster=nations$WorldS, paired = TRUE, alternative = "greater")
+wilcox.test(nations$NatS, nations$WorldS, paired = TRUE)
+```
+```{}
+These p values are all significant at alpha = 0.01
+```

### 2023-02-07 Code/Models.Rmd
Added simulations

@@
 
+```{r}
+library(lme4)
+men_pooled_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
+baseline <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
+
+simfit <- function(men_pooled_gamma_year, n = 12000000) {
+  vc <- as.data.frame(VarCorr(men_pooled_gamma_year))
+  ## mu <- exp(coef(men_pooled_gamma_year)$Year[,]) ## using fitted random effects only
+  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
+  mu <- exp(fixef(men_pooled_gamma_year) + z)
+  shape <- 1 / vc$vcov[2]
+  x <- rgamma(n, shape = shape, scale = mu / shape)
+  x
+}
+sbaseline <- summary(baseline)
+x <- simfit(men_pooled_gamma_year)
+mean(x < 0.1)
+quantile(x, prob = c(.00001, .0001, .0005, .001))
+## compare with and without random effect
+plot(density(x)) # with random effect
+curve(dgamma(x, shape = 1 / sbaseline$dispersion,
+             scale = exp(coef(baseline)) * sbaseline$dispersion),
+      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
+```
+
 ##2019 vs 2022 Comparisons

### 2023-02-13 Code/Models.Rmd
Added code

--- 2023-02-07 Code/Models.Rmd
@@
 AIC(old_pooled_yeareffect)
+```
+
+```{r}
+men_finals_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_finals)
+men_finals_one <- glm(RxnTime ~ 1, data = men_finals)
+anova(men_finals_one, men_finals_yeareffect)
+AIC(men_finals_one)
+AIC(men_finals_yeareffect)
+summary(men_finals_yeareffect)
+```
+```{r}
+logLik(men_finals_yeareffect)
+logLik(men_finals_one)
+```
+```{}
+For finals only data, the year effect is not helpful
+```
+
+```{r}
+men_finals_yeareffect_log <- lmer(log(RxnTime) ~ (1|Year), data = men_finals)
+men_finals_one_log <- glm(log(RxnTime) ~ 1, data = men_finals)
+AIC(men_finals_yeareffect_log)
+AIC(men_finals_one_log)
+```
+```{r}
+men_final_yeareffect_gamma <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma)
+men_final_one_gamma <- glm(RxnTime ~ 1, data = men_finals, family = Gamma)
+AIC(men_final_yeareffect_gamma)
+AIC(men_final_one_gamma)
+logLik(men_final_yeareffect_gamma)
+logLik(men_final_one_gamma)
+```
+```{}
+Should not fit with normal, gamma is better and shows that the year effect is helpful.
+Can try this again without last years data.
+```
+
+```{r}
+old_final_yeareffect_gamma <- glmer(RxnTime ~ (1|Year), data = old_finals, family = Gamma)
+old_final_one_gamma <- glm(RxnTime ~ 1, data = old_finals, family = Gamma)
+AIC(old_final_yeareffect_gamma)
+AIC(old_final_one_gamma)
+logLik(old_final_yeareffect_gamma)
+logLik(old_final_one_gamma)
+summary(old_final_yeareffect_gamma)
+summary(men_final_yeareffect_gamma)
+```
+
+```{}
+Simulate to see what prob of a time less than 0.1 second rxn time.
 ```
@@
 
+```{r}
+library(lme4)
+men_pooled_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
+baseline <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
+
+simfit <- function(men_pooled_gamma_year, n = 12000000) {
+  vc <- as.data.frame(VarCorr(men_pooled_gamma_year))
+  ## mu <- exp(coef(men_pooled_gamma_year)$Year[,]) ## using fitted random effects only
+  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
+  mu <- exp(fixef(men_pooled_gamma_year) + z)
+  shape <- 1 / vc$vcov[2]
+  x <- rgamma(n, shape = shape, scale = mu / shape)
+  x
+}
+sbaseline <- summary(baseline)
+x <- simfit(men_pooled_gamma_year)
+mean(x < 0.1)
+quantile(x, prob = c(.00001, .0001, .0005, .001))
+## compare with and without random effect
+plot(density(x)) # with random effect
+curve(dgamma(x, shape = 1 / sbaseline$dispersion,
+             scale = exp(coef(baseline)) * sbaseline$dispersion),
+      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
+```
+```{}
+Every 1000 racers, we expect 7 to have a reaction time below 0.1 seconds.  If we want 1 out of every 1000
+```
+
 ##2019 vs 2022 Comparisons
@@
 wilcox.test(comp$NineteenS, comp$TwentyTwoS, paired = TRUE)
+```
+```{}
+Cluster by 2019 and 2022
+Columns: Athlete, RxnTime, 2019/2022
 ```

### 2023-02-14 Code/Models.Rmd
Fixed clusrank

--- 2023-02-07 Code/Models.Rmd
@@
 
-##Models
 ```{r}
-men_finals_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_finals)
-men_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_pooled)
-old_finals_yeareffect <- lmer(RxnTime ~ (1|Year), data = old_finals)
-old_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = old_pooled)
-AIC(men_finals_yeareffect)
-AIC(old_finals_yeareffect)
-AIC(men_pooled_yeareffect)
-AIC(old_pooled_yeareffect)
-```
+library(lme4)
+men_finals_one <- glm(RxnTime ~ 1, data = men_finals)
+men_finals_year <- lmer(RxnTime ~ (1|Year), data = men_finals)
+men_finals_one_gamma <- glm(RxnTime ~ 1, data = men_finals, family = Gamma)
+men_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma)
+men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, family = Gamma(link = "log"))
+men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma(link = "log"))
 
-```{}
-Thus this seems to support that the 2022 data should be included and that the semifinals and finals should be pooled together.
-```
-
-```{r}
-trial <- subset(times, Gender == "M" & Time != 'DQ')
-x <- lmer(RxnTime ~ (1|Year), data = trial)
-AIC(x)
-```
-
-```{}
-This shows that increasing the data set increases AIC, unsure what we want to use
-```
-
-```{r}
-#men_finals_randeffect <- lmer(RxnTime ~ 1, data = men_pooled)
-men_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_pooled)
-men_pooled_randeffect <- glm(RxnTime ~ 1, data = men_pooled)
-
-anova(men_pooled_yeareffect, men_pooled_randeffect)
-
+paste("Linear without year effect has AIC of", AIC(men_finals_one), "and log-lik of", logLik(men_finals_one))
+paste("Linear with year effect has AIC of", AIC(men_finals_year), "and log-lik of", logLik(men_finals_year))
+paste("Gamma without year effect has AIC of", AIC(men_finals_one_gamma), "and log-lik of", logLik(men_finals_one_gamma))
+paste("Gamma with year effect has AIC of", AIC(men_finals_year_gamma), "and log-lik of", logLik(men_finals_year_gamma))
+paste("Gamma and log without year effect has AIC of", AIC(men_finals_one_gamma_log), "and log-lik of", logLik(men_finals_one_gamma_log))
+paste("Gamma and log with year effect has AIC of", AIC(men_finals_year_gamma_log), "and log-lik of", logLik(men_finals_year_gamma_log))
 ```
 ```{}
-As the year effect model is signficant at all alpha levels and the AIC is lower, the year effect is significant and should be considered in the model.
-Now we compare the log model to the year effect model
-```
-
-```{r}
-men_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_pooled)
-men_pooled_log_year <- lmer(log(RxnTime) ~ (1|Year), data = men_pooled)
-men_pooled_log <- glm(log(RxnTime) ~ 1, data = men_pooled)
-anova(men_pooled_yeareffect, men_pooled_log_year, men_pooled_log)
-```
-```{}
-The year effect model is better than both the log random effect model and the log year effect model.
+The best selected model is men_finals_year_gamma as that has the smallest AIC: -491 and largest log-lik of 248.5
 ```
@@
 library(lme4)
-men_pooled_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma)
-men_pooled_log_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
-men_pooled_yeareffect <- lmer(RxnTime ~ (1|Year), data = men_pooled)
+old_finals_one <- glm(RxnTime ~ 1, data = old_finals)
+old_finals_year <- lmer(RxnTime ~ (1|Year), data = old_finals)
+old_finals_one_gamma <- glm(RxnTime ~ 1, data = old_finals, family = Gamma)
+old_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = old_finals, family = Gamma)
+old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = old_finals, family = Gamma(link = "log"))
+old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = old_finals, family = Gamma(link = "log"))
 
-#Compare how much the link=log matters
-anova(men_pooled_gamma_year, men_pooled_log_gamma_year)
-
-#Compare gamma and linear
-anova(men_pooled_gamma_year, men_pooled_yeareffect)
-
+paste("Linear without year effect has AIC of", AIC(old_finals_one), "and log-lik of", logLik(old_finals_one))
+paste("Linear with year effect has AIC of", AIC(old_finals_year), "and log-lik of", logLik(old_finals_year))
+paste("Gamma without year effect has AIC of", AIC(old_finals_one_gamma), "and log-lik of", logLik(old_finals_one_gamma))
+paste("Gamma with year effect has AIC of", AIC(old_finals_year_gamma), "and log-lik of", logLik(old_finals_year_gamma))
+paste("Gamma and log without year effect has AIC of", AIC(old_finals_one_gamma_log), "and log-lik of", logLik(old_finals_one_gamma_log))
+paste("Gamma and log with year effect has AIC of", AIC(old_finals_year_gamma_log), "and log-lik of", logLik(old_finals_year_gamma_log))
 ```
 ```{}
-Whether or not the model contains "(link = "log")" seems to have very little effect on the strength of the model.
-The gamma model seems to perform better than the linear model based on AIC, but due to the gamma model being more difficult to interpret, the linear mixed effects model seems sufficient.
+The best selected model is old_finals_year_gamma as that has the smallest AIC: -465 and largest log-lik of 235.5.
+The models for the older data does not seem to fit the data as good as for the full data set.
+```
+
+```{r}
+library(lme4)
+men_pooled_one <- glm(RxnTime ~ 1, data = men_pooled)
+men_pooled_year <- lmer(RxnTime ~ (1|Year), data = men_pooled)
+men_pooled_one_gamma <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma)
+men_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma)
+men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
+men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma(link = "log"))
+
+paste("Linear without year effect has AIC of", AIC(men_pooled_one), "and log-lik of", logLik(men_pooled_one))
+paste("Linear with year effect has AIC of", AIC(men_pooled_year), "and log-lik of", logLik(men_pooled_year))
+paste("Gamma without year effect has AIC of", AIC(men_pooled_one_gamma), "and log-lik of", logLik(men_pooled_one_gamma))
+paste("Gamma with year effect has AIC of", AIC(men_pooled_year_gamma), "and log-lik of", logLik(men_pooled_year_gamma))
+paste("Gamma and log without year effect has AIC of", AIC(men_pooled_one_gamma_log), "and log-lik of", logLik(men_pooled_one_gamma_log))
+paste("Gamma and log with year effect has AIC of", AIC(men_pooled_year_gamma_log), "and log-lik of", logLik(men_pooled_year_gamma_log))
+```
+```{}
+The best selected model is men_pooled_year_gamma_log as it had the smallest AIC and the largest log-lik
+```
+
+```{r}
+library(lme4)
+old_pooled_one <- glm(RxnTime ~ 1, data = old_pooled)
+old_pooled_year <- lmer(RxnTime ~ (1|Year), data = old_pooled)
+old_pooled_one_gamma <- glm(RxnTime ~ 1, data = old_pooled, family = Gamma)
+old_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = old_pooled, family = Gamma)
+old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = old_pooled, family = Gamma(link = "log"))
+old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = old_pooled, family = Gamma(link = "log"))
+
+paste("Linear without year effect has AIC of", AIC(old_pooled_one), "and log-lik of", logLik(old_pooled_one))
+paste("Linear with year effect has AIC of", AIC(old_pooled_year), "and log-lik of", logLik(old_pooled_year))
+paste("Gamma without year effect has AIC of", AIC(old_pooled_one_gamma), "and log-lik of", logLik(old_pooled_one_gamma))
+paste("Gamma with year effect has AIC of", AIC(old_pooled_year_gamma), "and log-lik of", logLik(old_pooled_year_gamma))
+paste("Gamma and log without year effect has AIC of", AIC(old_pooled_one_gamma_log), "and log-lik of", logLik(old_pooled_one_gamma_log))
+paste("Gamma and log with year effect has AIC of", AIC(old_pooled_year_gamma_log), "and log-lik of", logLik(old_pooled_year_gamma_log))
+```
+```{}
+The best model is old_pooled_year_gamma as that has the smallest AIC: -1687.5 and the largest log-lik: 846.8
+```
+
+```{}
+Simulate to see what prob of a time less than 0.1 second rxn time.
 ```
@@
 ##Simulations
+
 ```{r}
-summary(men_pooled_gamma_year)
-qqnorm(resid(men_pooled_yeareffect))
-qqline(resid(men_pooled_yeareffect))
-qqGamma(resid(men_pooled_gamma_year))
+library(lme4)
+men_pooled_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
+baseline <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
+
+simfit <- function(men_pooled_gamma_year, n = 12000000) {
+  vc <- as.data.frame(VarCorr(men_pooled_gamma_year))
+  ## mu <- exp(coef(men_pooled_gamma_year)$Year[,]) ## using fitted random effects only
+  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
+  mu <- exp(fixef(men_pooled_gamma_year) + z)
+  shape <- 1 / vc$vcov[2]
+  x <- rgamma(n, shape = shape, scale = mu / shape)
+  x
+}
+sbaseline <- summary(baseline)
+print(sbaseline)
+x <- simfit(men_pooled_gamma_year)
+mean(x < 0.1)
+quantile(x, prob = c(.00001, .0001, .0005, .001))
+## compare with and without random effect
+plot(density(x)) # with random effect
+curve(dgamma(x, shape = 1 / sbaseline$dispersion,
+             scale = exp(coef(baseline)) * sbaseline$dispersion),
+      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
 ```
@@
 ```{r}
-library(qpToolkit)
-qqGamma(resid(men_pooled_gamma_year))
+simulation <- function(model, base){
+  simfit <- function(model, n = 12000000) {
+    vc <- as.data.frame(VarCorr(model))
+    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
+    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
+    mu <- exp(fixef(model) + z)
+    shape <- 1 / vc$vcov[2]
+    x <- rgamma(n, shape = shape, scale = mu / shape)
+    x
+  }
+  sbase <- summary(base)
+  #print(sbase)
+  x <- simfit(model)
+  print(mean(x < 0.1))
+  print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
+  ## compare with and without random effect
+  plot(density(x)) # with random effect
+  curve(dgamma(x, shape = 1 / sbase$dispersion, 
+               scale = exp(coef(base)) * sbase$dispersion), 
+        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
+}
+simulation(old_pooled_year_gamma_log, old_pooled_one_gamma_log)
 ```
 
-```{r}
-library(DHARMa)
-men_pooled_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
-summary(men_pooled_gamma_year)
-```
-```{r}
-simout_gamma <- simulateResiduals(men_pooled_gamma_year, n = 1000)
-simout_linear <- simulateResiduals(men_pooled_yeareffect, n = 1000)
-plot(simout_gamma)
-plot(simout_linear)
-#summary(simout)
-#str(simout)
-
+```{}
+Every 1000 racers, we expect 7 to have a reaction time below 0.1 seconds.  If we want 1 out of every 1000
 ```
@@
 ```{r}
-comp <- read.csv("../Data/2019vs2022.csv")
+comp <- read.csv("../Data/Clusrank2019vs2022.csv")
 #"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
-colnames(comp) <- c("Athlete", "Country", "NineteenH", "NineteenS", "NineteenF", "TwentyTwoH", "TwentyTwoS", "TwentyTwoF")
+colnames(comp) <- c("Time", "Group", "Cluster")
 #View(comp)
@@
 library(clusrank)
-t.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
-clusWilcox.test(comp$NineteenH, cluster=comp$TwentyTwoH, paired = TRUE, alternative = "greater")
-wilcox.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
+#t.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
+clusWilcox.test(comp$Time, group=comp$Group, cluster=comp$Cluster, paired = TRUE, alternative = "greater", method = "ds")
+clusWilcox.test(comp$Time, group=comp$Cluster, cluster = comp$Group, paired = TRUE, alternative = "greater", method = "ds")
+#wilcox.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
+```
+
+```{}
+Cluster by 2019 and 2022
+Columns: Athlete, RxnTime, 2019/2022
 ```
@@
 ```{r}
-t.test(comp$NineteenS, comp$TwentyTwoS, paired = TRUE)
-clusWilcox.test(comp$NineteenS, cluster=comp$TwentyTwoS, paired = TRUE, alternative = "greater")
-wilcox.test(comp$NineteenS, comp$TwentyTwoS, paired = TRUE)
-```
-
-```{}
-A matched pairs t test, clusrank package Wilcox test, and a standard Wilcox test all showed that the heat means differ by a significant amount.
-And two of the three semifinals means had an alpha less than 0.01.
-```
-
-```{r}
-nations <- read.csv("../Data/NatVSWorld.csv")
-colnames(nations) <- c("Athlete", "Country", "NatH", "NatS", "NatF", "WorldH", "WorldS", "WorldF")
+nations <- read.csv("../Data/ClusrankNatvsWorld.csv")
+colnames(nations) <- c("Time", "Group", "Cluster")
 ```
@@
 library(clusrank)
-t.test(nations$NatH, nations$WorldH, paired = TRUE)
-clusWilcox.test(nations$NatH, cluster=nations$WorldH, paired = TRUE, alternative = "greater")
-wilcox.test(nations$NatH, nations$WorldH, paired = TRUE)
+#t.test(nations$NatH, nations$WorldH, paired = TRUE)
+clusWilcox.test(nations$Time, group=nations$Group, cluster=nations$Cluster,
+                paired = TRUE, alternative = "greater", method = "ds")
+clusWilcox.test(nations$Time, group=nations$Cluster, cluster = nations$Group, 
+                paired = TRUE, alternative = "greater", method = "ds")
+#wilcox.test(nations$NatH, nations$WorldH, paired = TRUE)
 ```
-```{}
-These p values all significant at alpha = 0.05, but not at alpha = 0.01
-```
-
-```{r}
-library(clusrank)
-t.test(nations$NatS, nations$WorldS, paired = TRUE)
-clusWilcox.test(nations$NatS, cluster=nations$WorldS, paired = TRUE, alternative = "greater")
-wilcox.test(nations$NatS, nations$WorldS, paired = TRUE)
-```
-```{}
-These p values are all significant at alpha = 0.01
-```

### 2023-02-14 Manuscript/manuscript.tex
Added table

@@
 
+\section{Simulations}
+\label{sec:simulation}
+
+A simulation was performed on the gamma model with the year effect and link set
+to "log". At a size of n=12,000,000, the calculated probability of a reaction
+time being below 0.1 seconds was as follows for the various different data sets:
+
+\begin{center}
+  \begin{tabular}{||c|c||} 
+   \hline
+   Data Set & P RxnTime 0.1 \\ 
+   \hline\hline
+   all finals gamma log & 0.000856 \\ 
+   \hline
+   old finals gamma log & 0.000425 \\
+   \hline
+   all pooled gamma log & 0.00684 \\
+   \hline
+   old pooled gamma log & 0.00569 \\
+  \end{tabular}
+  \end{center}
+
+Prior to the 2022 World Championship, there was a 0.0425\% chance (based on the
+simulated results) of observing a reaction time below 0.01 seconds in the men's
+finals.  That means that we would expect a time below the threshold to occur
+once every 2,352 starts.  Considering there are roughly 8 men's finals reaction
+times every 2 years (The World Championship is biennial), every 588 years there
+should be a time as extreme as Allen's.  This seems extremely unlikely and thus
+it seems like there may be an additonal explanantion for the fast reaction times.
+
+If IAAF wanted the probability for a false start to be once every 100 years,
+then they should change the false start barrier to
+
 \section{Appendix}

### 2023-02-14 Code/Models.Rmd
Updated from meeting on 2/14

@@
 sbaseline <- summary(baseline)
-print(sbaseline)
+#print(sbaseline)
 x <- simfit(men_pooled_gamma_year)
@@
 clusWilcox.test(comp$Time, group=comp$Group, cluster=comp$Cluster, paired = TRUE, alternative = "greater", method = "ds")
-clusWilcox.test(comp$Time, group=comp$Cluster, cluster = comp$Group, paired = TRUE, alternative = "greater", method = "ds")
+clusWilcox.test(comp$Time, group=comp$Cluster, cluster = comp$Group, paired = TRUE, alternative = "greater", method = "rgl")
+clusWilcox.test(comp$Time, cluster = comp$Group, paired = TRUE, alternative = "greater", method = "ds")
 #wilcox.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)

### 2023-02-21 Code/Models.Rmd
Fixed clusrank tests

@@
 ```{r}
+set.seed(1)
 simulation <- function(model, base){
@@
 }
-simulation(old_pooled_year_gamma_log, old_pooled_one_gamma_log)
+simulation(men_finals_year_gamma_log, men_finals_one_gamma_log)
 ```
@@
 #"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
-colnames(comp) <- c("Time", "Group", "Cluster")
+colnames(comp) <- c("Time", "Athlete", "Year")
 #View(comp)
@@
 #t.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
-clusWilcox.test(comp$Time, group=comp$Group, cluster=comp$Cluster, paired = TRUE, alternative = "greater", method = "ds")
-clusWilcox.test(comp$Time, group=comp$Cluster, cluster = comp$Group, paired = TRUE, alternative = "greater", method = "rgl")
-clusWilcox.test(comp$Time, cluster = comp$Group, paired = TRUE, alternative = "greater", method = "ds")
+#clusWilcox.test(comp$Time, group=comp$Group, cluster=comp$Cluster, paired = TRUE, alternative = "greater", method = "ds")
+clusWilcox.test(Time, group=Year, cluster = Athlete, paired = TRUE, alternative = "greater", method = "ds", data = comp)
+clusWilcox.test(Time, cluster = Athlete, paired = TRUE, alternative = "greater", method = "rgl", data = comp)
 #wilcox.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
@@
 #t.test(nations$NatH, nations$WorldH, paired = TRUE)
-clusWilcox.test(nations$Time, group=nations$Group, cluster=nations$Cluster,
-                paired = TRUE, alternative = "greater", method = "ds")
+#clusWilcox.test(nations$Time, group=nations$Group, cluster=nations$Cluster,
+                #paired = TRUE, alternative = "greater", method = "ds")
 clusWilcox.test(nations$Time, group=nations$Cluster, cluster = nations$Group, 

### 2023-02-21 Manuscript/manuscript.tex
Added lots of tables

@@
 
-\section{Methods}
+\section{Methods and Models}
 \label{sec:methods}
+
+All Men final's Data
+\begin{center}
+  \begin{tabular}{|c | c | c | c | c | c | c |} 
+   \hline\hline
+   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   \hline
+   linear one & -472.7 & 238.3 & 0.1484 & 0.0019 & N/A & N/A \\ 
+   \hline
+   linear year & -468.7 & 237.4 & 0.1481 & 0.0030 & 7.762e-05 & 2.457e-04 \\
+   \hline
+   men finals one gamma & -477.3 & 240.7 & 6.738 & 0.0844 & N/A & N/A \\
+   \hline
+   men finals year gamma & -491.1 & 248.5 & 6.8121 & 0.1895 & 0.1148 & 0.0112 \\
+   \hline
+   men finals one gamma log & -477.3 & 240.7 & -1.908 & 0.0125 & N/A & N/A \\
+   \hline
+   men finals year gamma log & -491.0 & 248.5 & -1.914 & 0.0280 & 0.0025 & 0.0112 \\ [0.5ex]
+   \hline
+  \end{tabular}
+\end{center}
+
+Pre-2022 final's data
+\begin{center}
+  \begin{tabular}{|c | c | c | c | c | c | c |} 
+   \hline\hline
+   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   \hline
+   linear one & -450.7 & 227.4 & 0.1495 & 0.0019 & N/A & N/A \\
+   \hline
+   linear year & -444.1 & 225.0 & 0.1496 & 0.0028 & 5.691e-05 & 2.470e-04 \\ 
+   \hline
+   gamma one & -455.8 & 229.9 & 6.687 & 0.0834 & N/A & N/A \\
+   \hline
+   gamma year & -465.1 & 235.5 & 6.714 & 0.1674 & 0.0859 & 0.0109 \\
+   \hline
+   gamma one log & -455.8 & 229.9 & -1.900 & 0.0125 & N/A & N/A \\
+   \hline
+   gamma year log & -465.0 & 235.5 & -1.901 & 0.0251 & 0.0019 & 0.0109 \\ [0.5ex]
+   \hline
+  \end{tabular}
+  \end{center}
+
+All Men's semi finals and finals data
+\begin{center}
+  \begin{tabular}{|c | c | c | c | c | c | c |} 
+   \hline
+   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   \hline\hline
+   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
+   \hline
+   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
+   \hline
+   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
+   \hline
+   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
+   \hline
+   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
+   \hline
+   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
+   \hline
+  \end{tabular}
+  \end{center}
+
+Pre-2022 Men's semi finals and finals data
+\begin{center}
+  \begin{tabular}{|c | c | c | c | c | c | c |} 
+   \hline
+   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   \hline\hline
+   linear one & -1539 & 771.5 & 0.1544 & 0.0015 & N/A & N/A \\
+   \hline
+   linear year & -1564 & 785.4 & 0.1562 & 0.0037 & 1.280e-04 & 6.273e-04 \\ 
+   \hline
+   gamma one & -1626 & 814.9 & 6.475 & 0.0608 & N/A & N/A \\
+   \hline
+   gamma year & -1688 & 846.8 & 6.426 & 0.1923 & 0.0989 & 0.0243 \\
+   \hline
+   gamma one log & -1626 & 814.9 & -1.868 & 0.0094 & N/A & N/A \\
+   \hline
+   gamma year log & -1687 & 846.6 & -1.853 & 0.0305 & 0.0024 & 0.0243 \\ [0.5ex]
+   \hline
+  \end{tabular}
+  \end{center}
+
+Sorted By Model instead:
+Linear Model with no year effect
+\begin{center}
+  \begin{tabular}{|c | c | c | c | c | c | c |} 
+   \hline
+   Dataset & AIC & Log Likelihood & Int effect est & Int se \\ [0.5ex] 
+   \hline\hline
+   All Finals & -472.7 & 238.3 & 0.1484 & 0.0019 \\
+   \hline
+   Older Finals & -450.7 & 227.4 & 0.1495 & 0.0019 \\ 
+   \hline
+   All Pooled & -1667 & 835.5 & 0.1529 & 0.0014 \\
+   \hline
+   Older Pooled & -1539 & 771.5 & 0.1544 & 0.0015 \\
+   \hline
+  \end{tabular}
+  \end{center}
+
+Linear model with year effect
+\begin{center}
+  \begin{tabular}{|c | c | c | c | c | c | c |} 
+   \hline
+   Dataset & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   \hline\hline
+   All Finals & -468.7 & 237.4 & 0.1481 & 0.0030 & 7.762e-05 & 2.457e-04 \\
+   \hline
+   Older Finals & -444.1 & 225.0 & 0.1496 & 0.0028 & 5.691e-05 & 2.470e-04 \\ 
+   \hline
+   All Pooled & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\
+   \hline
+   Older Pooled & -1564 & 785.4 & 0.1562 & 0.0037 & 1.280e-04 & 6.273e-04 \\
+   \hline
+  \end{tabular}
+  \end{center}
+
+Gamma model with year effect
+\begin{center}
+  \begin{tabular}{|c | c | c | c | c | c | c |} 
+   \hline
+   Dataset & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   \hline\hline
+   All Finals & -491.1 & 248.5 & 6.8121 & 0.1895 & 0.1148 & 0.0112 \\
+   \hline
+   Older Finals & -465.1 & 235.5 & 6.714 & 0.1674 & 0.0859 & 0.0109 \\ 
+   \hline
+   All Pooled & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
+   \hline
+   Older Pooled & -1688 & 846.8 & 6.426 & 0.1923 & 0.0989 & 0.0243 \\
+   \hline
+  \end{tabular}
+  \end{center}
+
+Gamma model with year effect and link=log
+\begin{center}
+  \begin{tabular}{|c | c | c | c | c | c | c |} 
+   \hline
+   Dataset & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   \hline\hline
+   All Finals & -491.0 & 248.5 & -1.914 & 0.0280 & 0.0025 & 0.0112 \\
+   \hline
+   Older Finals & -465.0 & 235.5 & -1.901 & 0.0251 & 0.0019 & 0.0109 \\ 
+   \hline
+   All Pooled & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\
+   \hline
+   Older Pooled & -1687 & 846.6 & -1.853 & 0.0305 & 0.0024 & 0.0243 \\
+   \hline
+  \end{tabular}
+  \end{center}
+
+When determining the most effective model to use, several differnt examples were
+explored across the 4 different data sets: a linear model without a year effect,
+linear model with year effect, a gamma model without a year effect, a gamma model
+with a year effect, a gamma model without a year effect and link set to log instead
+of the default which is
 
@@
    \hline\hline
-   all finals gamma log & 0.000856 \\ 
+   all finals gamma log & 0.000863 \\ 
    \hline
-   old finals gamma log & 0.000425 \\
+   old finals gamma log & 0.000429 \\
    \hline
-   all pooled gamma log & 0.00684 \\
+   all pooled gamma log & 0.006847 \\
    \hline
-   old pooled gamma log & 0.00569 \\
+   old pooled gamma log & 0.005749 \\
   \end{tabular}

### 2023-02-21 Code/Models.Rmd
Small changes

@@
 
+```{}
+One thing that we discovered is that including group=Year does not effect the the z score or p value.  Changing the method from
+rgl to ds does effect z score.
+```
+
 ```{r}
 library(clusrank)
-#t.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
-#clusWilcox.test(comp$Time, group=comp$Group, cluster=comp$Cluster, paired = TRUE, alternative = "greater", method = "ds")
-clusWilcox.test(Time, group=Year, cluster = Athlete, paired = TRUE, alternative = "greater", method = "ds", data = comp)
-clusWilcox.test(Time, cluster = Athlete, paired = TRUE, alternative = "greater", method = "rgl", data = comp)
-#wilcox.test(comp$NineteenH, comp$TwentyTwoH, paired = TRUE)
-```
-
-```{}
-Cluster by 2019 and 2022
-Columns: Athlete, RxnTime, 2019/2022
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, paired = TRUE, alternative = "greater", method = "ds")
+clusWilcox.test(Time, cluster = Athlete, data = comp, paired = TRUE, alternative = "greater", method = "rgl")
 ```
@@
 nations <- read.csv("../Data/ClusrankNatvsWorld.csv")
-colnames(nations) <- c("Time", "Group", "Cluster")
+colnames(nations) <- c("Time", "Athlete", "Year")
 ```
@@
 library(clusrank)
-#t.test(nations$NatH, nations$WorldH, paired = TRUE)
-#clusWilcox.test(nations$Time, group=nations$Group, cluster=nations$Cluster,
-                #paired = TRUE, alternative = "greater", method = "ds")
-clusWilcox.test(nations$Time, group=nations$Cluster, cluster = nations$Group, 
-                paired = TRUE, alternative = "greater", method = "ds")
-#wilcox.test(nations$NatH, nations$WorldH, paired = TRUE)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, paired = TRUE, alternative = "greater", method = "ds")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, paired = TRUE, alternative = "greater", method = "rgl")
 ```
+
+```{r}
+In both examples, rgl produced a smaller p value
+```

### 2023-02-21 Manuscript/manuscript.tex
Progress to conclusion and interpretation

@@
 with a year effect, a gamma model without a year effect and link set to log instead
-of the default which is
-
-The main statistical model used in the paper is a gamma mixed effects model, which is
-in the family of Generalized Linear Mixed Model (GLMM). The GLMM comes from the R package 
-lme4, which hosts a variety of GLMM functions \citep{Rpkg:lme4}.  A GLMM is a type of model
+of the default which is gaussian.  The R package lme4 was very helpful with this
+as it supported the Generalized Linear Mixed Model (GLMM) which allowed both
+linear and gamma mized effects models to be built \citep{Rpkg:lme4}.A GLMM is a type of model
 that supports inputs that are from different distributions.  The model that we use is for the
@@
 density less than 0.1, which is the current threshold for a reaction time
-disqualification.  The density plots are shown below.
+disqualification.  The density plots are shown below.  The batch effect is very
+difficult to explain or interpret as it means that some heats of races were either
+faster or slower and that everyone in the batch reacted in a similar way.  Although
+it is possible that one athlete reacting faster could lead to others reacting
+similarly, this does not make much practical sense as the difference between
+heats that had fast and slow reaction times are all within 0.12 seconds of each
+other.  It is much easier however to explain the year effect.  The technology
+that World Athletics has used has differed since 1999 and theoretically the
+reaction time data from 2022 should be the best as the timing system should be
+state of the art.  Any other weather or climate related factors such as humidity,
+precipitation, elevation, etc would all be negated when looking at the batch
+effect, but are all factors that could potentially impact the year effect.
 
@@
    \hline\hline
-   all finals gamma log & 0.000863 \\ 
+   all finals gamma log & 0.000864 \\ 
    \hline

### 2023-02-28 Code/Women.Rmd
Women's models and clusrank

---
title: "Women"
author: "Owen Fiore"
date: "2023-02-28"
output: pdf_document
---

```{r}
times <- read.csv("../Data/ThesisData.csv")
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
```

```{r}
women_finals <- subset(times, Type == "F" & Gender == "F" & Time != 'DQ') # & Year < 2022)
women_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "F" & Time != 'DQ')
```

```{r}
women_old_finals <- subset(times, Type == "F" & Gender == "F" & Time != 'DQ' & Year < 2022)
women_old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "F" & Time != 'DQ' & Year < 2022)
```

```{r}
library(lme4)
women_finals_one <- glm(RxnTime ~ 1, data = women_finals)
women_finals_year <- lmer(RxnTime ~ (1|Year), data = women_finals)
women_finals_one_gamma <- glm(RxnTime ~ 1, data = women_finals, family = Gamma)
women_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = women_finals, family = Gamma)
women_finals_one_gamma_log <- glm(RxnTime ~ 1, data = women_finals, family = Gamma(link = "log"))
women_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = women_finals, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(women_finals_one), "and log-lik of", logLik(women_finals_one))
paste("Linear with year effect has AIC of", AIC(women_finals_year), "and log-lik of", logLik(women_finals_year))
paste("Gamma without year effect has AIC of", AIC(women_finals_one_gamma), "and log-lik of", logLik(women_finals_one_gamma))
paste("Gamma with year effect has AIC of", AIC(women_finals_year_gamma), "and log-lik of", logLik(women_finals_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(women_finals_one_gamma_log), "and log-lik of", logLik(women_finals_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(women_finals_year_gamma_log), "and log-lik of", logLik(women_finals_year_gamma_log))
```

```{r}
library(lme4)
women_old_finals_one <- glm(RxnTime ~ 1, data = women_old_finals)
women_old_finals_year <- lmer(RxnTime ~ (1|Year), data = women_old_finals)
women_old_finals_one_gamma <- glm(RxnTime ~ 1, data = women_old_finals, family = Gamma)
women_old_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = women_old_finals, family = Gamma)
women_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = women_old_finals, family = Gamma(link = "log"))
women_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = women_old_finals, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(women_old_finals_one), "and log-lik of", logLik(women_old_finals_one))
paste("Linear with year effect has AIC of", AIC(women_old_finals_year), "and log-lik of", logLik(women_old_finals_year))
paste("Gamma without year effect has AIC of", AIC(women_old_finals_one_gamma), "and log-lik of", logLik(women_old_finals_one_gamma))
paste("Gamma with year effect has AIC of", AIC(women_old_finals_year_gamma), "and log-lik of", logLik(women_old_finals_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(women_old_finals_one_gamma_log), "and log-lik of", logLik(women_old_finals_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(women_old_finals_year_gamma_log), "and log-lik of", logLik(women_old_finals_year_gamma_log))
```

```{r}
library(lme4)
women_pooled_one <- glm(RxnTime ~ 1, data = women_pooled)
women_pooled_year <- lmer(RxnTime ~ (1|Year), data = women_pooled)
women_pooled_one_gamma <- glm(RxnTime ~ 1, data = women_pooled, family = Gamma)
women_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = women_pooled, family = Gamma)
women_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = women_pooled, family = Gamma(link = "log"))
women_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = women_pooled, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(women_pooled_one), "and log-lik of", logLik(women_pooled_one))
paste("Linear with year effect has AIC of", AIC(women_pooled_year), "and log-lik of", logLik(women_pooled_year))
paste("Gamma without year effect has AIC of", AIC(women_pooled_one_gamma), "and log-lik of", logLik(women_pooled_one_gamma))
paste("Gamma with year effect has AIC of", AIC(women_pooled_year_gamma), "and log-lik of", logLik(women_pooled_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(women_pooled_one_gamma_log), "and log-lik of", logLik(women_pooled_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(women_pooled_year_gamma_log), "and log-lik of", logLik(women_pooled_year_gamma_log))
```

```{r}
library(lme4)
women_old_pooled_one <- glm(RxnTime ~ 1, data = women_old_pooled)
women_old_pooled_year <- lmer(RxnTime ~ (1|Year), data = women_old_pooled)
women_old_pooled_one_gamma <- glm(RxnTime ~ 1, data = women_old_pooled, family = Gamma)
women_old_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = women_old_pooled, family = Gamma)
women_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = women_old_pooled, family = Gamma(link = "log"))
women_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = women_old_pooled, family = Gamma(link = "log"))

paste("Linear without year effect has AIC of", AIC(women_old_pooled_one), "and log-lik of", logLik(women_old_pooled_one))
paste("Linear with year effect has AIC of", AIC(women_old_pooled_year), "and log-lik of", logLik(women_old_pooled_year))
paste("Gamma without year effect has AIC of", AIC(women_old_pooled_one_gamma), "and log-lik of", logLik(women_old_pooled_one_gamma))
paste("Gamma with year effect has AIC of", AIC(women_old_pooled_year_gamma), "and log-lik of", logLik(women_old_pooled_year_gamma))
paste("Gamma and log without year effect has AIC of", AIC(women_old_pooled_one_gamma_log), "and log-lik of", logLik(women_old_pooled_one_gamma_log))
paste("Gamma and log with year effect has AIC of", AIC(women_old_pooled_year_gamma_log), "and log-lik of", logLik(women_old_pooled_year_gamma_log))
```

```{r}
boxplot(RxnTime ~ Year, data=women_finals)
```
```{r}
women_nations <- read.csv("../Data/WomenClusrank2019vs2022.csv")
colnames(women_nations) <- c("Time", "Athlete", "Year")
```

```{r}
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_nations, paired = TRUE, alternative = "greater", method = "ds")
clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_nations, paired = TRUE, alternative = "greater", method = "rgl")
```

### 2023-02-28 Manuscript/manuscript.tex
Added more sections and standard error

@@
 any published papers regarding Devon Allen's disqualification in particular.  This is not
-neccessarily suprising given that the World Track and Field Championships occurred only 
-four months ago.
+neccessarily suprising given that his disqualification occurred only last August.
 
@@
 that could be done to rectify any potential mis-qualification other than to try
-to prevent it from happening again.  THis matter needs to be addressed because Devon Allen's 
+to prevent it from happening again.  This matter needs to be addressed because Devon Allen's 
 disqulification could be a problem that could continue to re-appear.  If athletes
@@
 
-\section{Methods and Models}
-\label{sec:methods}
-
 All Men final's Data
@@
 \begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c |} 
+  \begin{tabular}{|c | c | c | c | c | c | c | c |} 
    \hline
-   Dataset & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   Dataset & AIC & Year se & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
    \hline\hline
-   All Finals & -468.7 & 237.4 & 0.1481 & 0.0030 & 7.762e-05 & 2.457e-04 \\
+   All Finals & -468.7 & 0.0088 & 237.4 & 0.1481 & 0.0030 & 7.762e-05 & 2.457e-04 \\
    \hline
-   Older Finals & -444.1 & 225.0 & 0.1496 & 0.0028 & 5.691e-05 & 2.470e-04 \\ 
+   Older Finals & -444.1 & 0.0075 & 225.0 & 0.1496 & 0.0028 & 5.691e-05 & 2.470e-04 \\ 
    \hline
-   All Pooled & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\
+   All Pooled & -1713 & 0.131 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\
    \hline
-   Older Pooled & -1564 & 785.4 & 0.1562 & 0.0037 & 1.280e-04 & 6.273e-04 \\
+   Older Pooled & -1564 & 0.113 & 785.4 & 0.1562 & 0.0037 & 1.280e-04 & 6.273e-04 \\
    \hline
@@
 \begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c |} 
+  \begin{tabular}{|c | c | c | c | c | c | c | c |} 
    \hline
-   Dataset & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   Dataset & AIC & Year se & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
    \hline\hline
-   All Finals & -491.1 & 248.5 & 6.8121 & 0.1895 & 0.1148 & 0.0112 \\
+   All Finals & -491.1 & 0.3388 & 248.5 & 6.8121 & 0.1895 & 0.1148 & 0.0112 \\
    \hline
-   Older Finals & -465.1 & 235.5 & 6.714 & 0.1674 & 0.0859 & 0.0109 \\ 
+   Older Finals & -465.1 & 0.2930 & 235.5 & 6.714 & 0.1674 & 0.0859 & 0.0109 \\ 
    \hline
-   All Pooled & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
+   All Pooled & -1848 & 0.3612 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
    \hline
-   Older Pooled & -1688 & 846.8 & 6.426 & 0.1923 & 0.0989 & 0.0243 \\
+   Older Pooled & -1688 & 0.3145 & 846.8 & 6.426 & 0.1923 & 0.0989 & 0.0243 \\
    \hline
@@
 \begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c |} 
+  \begin{tabular}{|c | c | c | c | c | c | c | c |} 
    \hline
-   Dataset & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   Dataset & AIC & Year se & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
    \hline\hline
-   All Finals & -491.0 & 248.5 & -1.914 & 0.0280 & 0.0025 & 0.0112 \\
+   All Finals & -491.0 & 0.0504 & 248.5 & -1.914 & 0.0280 & 0.0025 & 0.0112 \\
    \hline
-   Older Finals & -465.0 & 235.5 & -1.901 & 0.0251 & 0.0019 & 0.0109 \\ 
+   Older Finals & -465.0 & 0.0439 & 235.5 & -1.901 & 0.0251 & 0.0019 & 0.0109 \\ 
    \hline
-   All Pooled & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\
+   All Pooled & -1848 & 0.0554 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\
    \hline
-   Older Pooled & -1687 & 846.6 & -1.853 & 0.0305 & 0.0024 & 0.0243 \\
+   Older Pooled & -1687 & 0.0495 & 846.6 & -1.853 & 0.0305 & 0.0024 & 0.0243 \\
    \hline
@@
 
+\section{GLMM Gamma Model}
 When determining the most effective model to use, several differnt examples were
@@
 
+In terms of the data that was used, it is worth noting that the standard error
+of the intercept was higher for all 6 comparisons of the data: finals data and 
+pooled data for the linear model, gamma model, gamma and log model for when
+2022 was included versus when it was not.  What this shows is that the addition
+of 2022 to the linear model increased the model's variation and that the addition
+of 2022 resulted in a worse model.  Although this is more of a qualitative analysis,
+this is further evidence that 2022 was a significant year in the role it played
+in altering the gamma model.
+
 \begin{figure}[h]
@@
 \end{figure}
+
+\section{2019 and National Data}
+This problem of whether or not the sensor was faulty can be further explored in
+two ways. First we can look at the reaction times of athletes who competed at the
+World Track and Field Championship in both 2019 and 2022 and compare the times,
+and secondly we can look at how athletes performed compared to how they did at
+their national/qualifiers meet.  We would expect that athletes of this caliber,
+who are world class athletes to not improve significantly from one year to the
+next, but that is not neccessarily the results that were found.
 

### 2023-03-02 Code/Models.Rmd
Not sure what I changed but small

@@
 ```{r}
-old_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ' & Year < 2022)
-old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Year < 2022)
+men_old_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ' & Year < 2022)
+men_old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Year < 2022)
 ```
@@
 ```
+
 ```{}
@@
 library(lme4)
-old_finals_one <- glm(RxnTime ~ 1, data = old_finals)
-old_finals_year <- lmer(RxnTime ~ (1|Year), data = old_finals)
-old_finals_one_gamma <- glm(RxnTime ~ 1, data = old_finals, family = Gamma)
-old_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = old_finals, family = Gamma)
-old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = old_finals, family = Gamma(link = "log"))
-old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = old_finals, family = Gamma(link = "log"))
+men_old_finals_one <- glm(RxnTime ~ 1, data = men_old_finals)
+men_old_finals_year <- lmer(RxnTime ~ (1|Year), data = men_old_finals)
+men_old_finals_one_gamma <- glm(RxnTime ~ 1, data = men_old_finals, family = Gamma)
+men_old_finals_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma)
+men_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_finals, family = Gamma(link = "log"))
+men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "log"))
 
-paste("Linear without year effect has AIC of", AIC(old_finals_one), "and log-lik of", logLik(old_finals_one))
-paste("Linear with year effect has AIC of", AIC(old_finals_year), "and log-lik of", logLik(old_finals_year))
-paste("Gamma without year effect has AIC of", AIC(old_finals_one_gamma), "and log-lik of", logLik(old_finals_one_gamma))
-paste("Gamma with year effect has AIC of", AIC(old_finals_year_gamma), "and log-lik of", logLik(old_finals_year_gamma))
-paste("Gamma and log without year effect has AIC of", AIC(old_finals_one_gamma_log), "and log-lik of", logLik(old_finals_one_gamma_log))
-paste("Gamma and log with year effect has AIC of", AIC(old_finals_year_gamma_log), "and log-lik of", logLik(old_finals_year_gamma_log))
+paste("Linear without year effect has AIC of", AIC(men_old_finals_one), "and log-lik of", logLik(men_old_finals_one))
+paste("Linear with year effect has AIC of", AIC(men_old_finals_year), "and log-lik of", logLik(men_old_finals_year))
+paste("Gamma without year effect has AIC of", AIC(men_old_finals_one_gamma), "and log-lik of", logLik(men_old_finals_one_gamma))
+paste("Gamma with year effect has AIC of", AIC(men_old_finals_year_gamma), "and log-lik of", logLik(men_old_finals_year_gamma))
+paste("Gamma and log without year effect has AIC of", AIC(men_old_finals_one_gamma_log), "and log-lik of", logLik(men_old_finals_one_gamma_log))
+paste("Gamma and log with year effect has AIC of", AIC(men_old_finals_year_gamma_log), "and log-lik of", logLik(men_old_finals_year_gamma_log))
 ```
 ```{}
-The best selected model is old_finals_year_gamma as that has the smallest AIC: -465 and largest log-lik of 235.5.
+The best selected model is men_old_finals_year_gamma as that has the smallest AIC: -465 and largest log-lik of 235.5.
 The models for the older data does not seem to fit the data as good as for the full data set.
@@
 library(lme4)
-old_pooled_one <- glm(RxnTime ~ 1, data = old_pooled)
-old_pooled_year <- lmer(RxnTime ~ (1|Year), data = old_pooled)
-old_pooled_one_gamma <- glm(RxnTime ~ 1, data = old_pooled, family = Gamma)
-old_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = old_pooled, family = Gamma)
-old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = old_pooled, family = Gamma(link = "log"))
-old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = old_pooled, family = Gamma(link = "log"))
+men_old_pooled_one <- glm(RxnTime ~ 1, data = men_old_pooled)
+men_old_pooled_year <- lmer(RxnTime ~ (1|Year), data = men_old_pooled)
+men_old_pooled_one_gamma <- glm(RxnTime ~ 1, data = men_old_pooled, family = Gamma)
+men_old_pooled_year_gamma <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma)
+men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, family = Gamma(link = "log"))
+men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "log"))
 
-paste("Linear without year effect has AIC of", AIC(old_pooled_one), "and log-lik of", logLik(old_pooled_one))
-paste("Linear with year effect has AIC of", AIC(old_pooled_year), "and log-lik of", logLik(old_pooled_year))
-paste("Gamma without year effect has AIC of", AIC(old_pooled_one_gamma), "and log-lik of", logLik(old_pooled_one_gamma))
-paste("Gamma with year effect has AIC of", AIC(old_pooled_year_gamma), "and log-lik of", logLik(old_pooled_year_gamma))
-paste("Gamma and log without year effect has AIC of", AIC(old_pooled_one_gamma_log), "and log-lik of", logLik(old_pooled_one_gamma_log))
-paste("Gamma and log with year effect has AIC of", AIC(old_pooled_year_gamma_log), "and log-lik of", logLik(old_pooled_year_gamma_log))
+paste("Linear without year effect has AIC of", AIC(men_old_pooled_one), "and log-lik of", logLik(men_old_pooled_one))
+paste("Linear with year effect has AIC of", AIC(men_old_pooled_year), "and log-lik of", logLik(men_old_pooled_year))
+paste("Gamma without year effect has AIC of", AIC(men_old_pooled_one_gamma), "and log-lik of", logLik(men_old_pooled_one_gamma))
+paste("Gamma with year effect has AIC of", AIC(men_old_pooled_year_gamma), "and log-lik of", logLik(men_old_pooled_year_gamma))
+paste("Gamma and log without year effect has AIC of", AIC(men_old_pooled_one_gamma_log), "and log-lik of", logLik(men_old_pooled_one_gamma_log))
+paste("Gamma and log with year effect has AIC of", AIC(men_old_pooled_year_gamma_log), "and log-lik of", logLik(men_old_pooled_year_gamma_log))
 ```
 ```{}
-The best model is old_pooled_year_gamma as that has the smallest AIC: -1687.5 and the largest log-lik: 846.8
+The best model is men_old_pooled_year_gamma as that has the smallest AIC: -1687.5 and the largest log-lik: 846.8
 ```
@@
 }
-simulation(men_finals_year_gamma_log, men_finals_one_gamma_log)
+simulation(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log)
 ```
@@
 clusWilcox.test(Time, cluster = Athlete, data = comp, paired = TRUE, alternative = "greater", method = "rgl")
+#pairwise.clusWilcox.test(comp$Time, group= comp$Year, cluster = comp$Athlete, alternative = "greater", method = "rgl")
 ```
@@
 
-```{r}
+```{}
 In both examples, rgl produced a smaller p value
 ```
+
+```{r}
+mini <- min(aggregate(times$RxnTime, list(times$Batch), FUN =mean)$x)
+maxi <- max(aggregate(times$RxnTime, list(times$Batch), FUN =mean)$x)
+cat(maxi - mini)
+```

### 2023-03-02 Manuscript/manuscript.tex
Started analysis of clusrank code

@@
 \section{2019 and National Data}
-This problem of whether or not the sensor was faulty can be further explored in
+This problem of whether the sensor was faulty can be further explored in
 two ways. First we can look at the reaction times of athletes who competed at the
@@
 their national/qualifiers meet.  We would expect that athletes of this caliber,
-who are world class athletes to not improve significantly from one year to the
-next, but that is not neccessarily the results that were found.
+who are world-class athletes, to not improve significantly from one year to the
+next, but that is not necessarily the results that were found.  The R package
+"clusrank" was used to perform a clustered Wilcoxon signed-rank test.  This was
+chosen over a typical t test so as to avoid any assumptions about the distribution
+of population means.  The clusrank package allows clusters to be formed, which
+in this case are for each athlete.  The data containing each athlete's name,
+nationality, and reaction times for the 2019 and 2022 World Championship can be
+found in the csv file labeled "2019vs2022".  This data was transformed to "Clusrank2019vs2022"
+so that it better fits the clusrank package.  Each athlete was given a non-meaningful
+ID (An ID of 1 does not indicate that this athlete is better or worse than any
+other athlete).  Additionally, the "Year" column is used to determine what year
+the time was registered with "1" indicating 2019 and "2" indicating 2022.
+
+Using the "clusWilcox.test" function, a paired and clustered Wilcoxon signed-rank
+test was performed wtih the cluster set to the be the Athlete and group set to Year.
+Two different methods were used: "rgl" which is (Prof Yan should help with this)
+Paired was set to be "true" to perform a matched pairs test, and the alternative
+hypothesis was set to be "greater" as we want to know if the 2019 reaction times
+were larger than the 2022 times by a significant margin.
+
+When the clusrank analysis was performed, the "ds" method produced a z score of
+3.5271 and p value of 0.0002101 and the "rgl" method produced a z score of 3.665
+and a p value of 0.0001237.  These values are for a one-sided test and show that
+the mean reaction time on average in 2019 was larger than the mean reaction time
+in 2022 for athletes who competed in both championships.
+
+Another important data set that was investigated was the 2022 national competition
+data.  In order to send between 1-4 athletes to the World Championships, almost
+every athlete had to qualify by proving they are elite at the national level.
+Thus, data was compiled on athletes who competed at the 2022 World Track and 
+Field Championships on what these athletes also ran at their country's national
+championship.  If there was any argument that athletes have improved their reaction
+time from 2019 to 2022, it should now be considered nearly null, as the national
+competitions took place between June and July 2022 and the World Championships
+were in August.  The equipment used to measure reaction time cannot be guaranteed
+to be the same from national to international competition, but the time span between
+meets has been reduced considerably.  Repeating the data collection steps and
+again performing a clustered analysis for a matched pairs Wilcoxon signed-rank test
+results in two new z scores and p values.  The "ds" method produced a z score of
+3.6755 and p value of 0.0001187 and the "rgl" method produced a z score of 3.7618
+and a p value of 8.434e-05. These results are highly significant and show that
+the times at the national stage of competition were significantly higher than
+the reaction times at the World Championships.
 
@@
 0.099 seconds in the 2022 Championship final and 0.101 in the 2022 semifinal).  Removing this batch and others could greatly improve the strength of the model.  However, while
-it is a high outlier, there is not sufficient information available as to why the
+it is a high outlier, there is insufficient information available as to why the
 reaction times are so high, and no basis for why it should be improved other than because
@@
 The figure below plots the average reaction time by year for the pooled men's data since
-1999.  The circle in the bottom right is for 2022, which is signifcantly lower than the
+1999.  The circle in the bottom right is for 2022, which is significantly lower than the
 data points for the previous three years, and is substially lower than the data points
@@
 was not an issue with the system, the large variation is confusing because there is little explanation
-as to why there is so much of a drop off in reaction time from 2013 to 2015 or the spike
+as to why there is so much of a dropoff in reaction time from 2013 to 2015 or the spike
 from 2001 to 2003.
 
-Devon Allen's disqulification at the World Track and Field Championships was
+Devon Allen's disqualification at the World Track and Field Championships was
 possibly the result of both a faulty timing equipment and Devon Allen
@@
 the 2022 World Track and Field Championships were a low outlier compared to
-every other year in terms of average reaction time.  Addtionally, it was shown
+every other year in terms of average reaction time.  Additionally, it was shown
 through a gamma mixed effects model that there are significant year and batch
@@
 for reaction times has been random and unpredictable.  Thus while it seems easy
-to conclude that Devon Allen was wrongly disqualified, that may not neccessarily
+to conclude that Devon Allen was wrongly disqualified, that may not necessarily
 be the case.  If there are issues with reaction time such as athletes approaching the 0.10 

### 2023-03-21 Code/Models.Rmd
Updated rank sum test

@@
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, paired = TRUE, alternative = "greater", method = "ds")
-clusWilcox.test(Time, cluster = Athlete, data = comp, paired = TRUE, alternative = "greater", method = "rgl")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, alternative = "greater", method = "ds")
+#clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, alternative = "greater", method = "rgl")
 #pairwise.clusWilcox.test(comp$Time, group= comp$Year, cluster = comp$Athlete, alternative = "greater", method = "rgl")
@@
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, paired = TRUE, alternative = "greater", method = "ds")
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, paired = TRUE, alternative = "greater", method = "rgl")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, alternative = "greater", method = "ds")
+#clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, paired = TRUE, alternative = "greater", method = "rgl")
+
 ```
@@
 ```
+
+```{r}
+library(xtable)
+xtable(men_pooled_year_gamma_log)
+```
+```{r}
+library(memisc)
+mtable(men_pooled_year_gamma_log)
+```

### 2023-03-21 Manuscript/manuscript.tex
Worked on every section

@@
 \usepackage{mwe}
+\usepackage{comment}
 %\linenumbers*[1]
@@
 equipment rather than reacting illegally fast.  There is little at this point
-that could be done to rectify any potential mis-qualification other than to try
+that could be done to rectify any potential mis-disqualification other than to try
 to prevent it from happening again.  This matter needs to be addressed because Devon Allen's 
@@
 
-\section{Background}
-\label{sec:Background}
 Since 1985 Seiko Holding Corporations has served every World Athletics Championship
@@
 the 2009 World Championships. Starting in 2011, the rule was scrapped, and the old policy which 
-automatically disqualifies runners who false start was reinsituted.  The effect? At the 2011 World
+automatically disqualifies runners who false start was reinstituted.  The effect? At the 2011 World
 Championship there were 6 male disqualifications and 4 female disqualifications \citep{False-Start}. 
@@
 make already long track meets more tedious, both for athletes and viewers watching the 
-television broadcast.\citep{Pilianidis}
+television broadcast.  Since 2011, IAAF has instituted the one false start rule.\citep{Pilianidis}
 
-\section{Data}
-\label{sec:Data}
-Data was copied from the World Athletics website and pasted into an excel
+\section{GLMM Model}
+\label{sec:GLMM}
+Data was copied from the World Athletics website and pasted into an Excel
 spreadsheet. The data covers the men's 110 meter hurdles and the women's 100
@@
 
+\iffalse
 \begin{center}
@@
   \end{center}
+\fi
+
+AIC with and without random effect, Gamma and log model on all semifinal and finals data
+\begin{center}
+  \begin{tabular}{|c | c|}
+    \hline
+    Random Effect? & AIC \\
+    No & -1756 \\
+    Yes & -1848 \\
+    \hline
+  \end{tabular}
+\end{center}
+
+Addition of 2022 on Gamma and log model on semifinal and finals data
+\begin{center}
+  \begin{tabular}{|c | c|}
+    \hline
+    Dataset & Year Effect Standard Error \\
+    1999-2019 & 0.0495 \\
+    1999-2022 & 0.0554 \\
+    \hline
+  \end{tabular}
+\end{center}
+
+All summary statistics for Gamma and log model on all semifinal and finals data
+\begin{center}
+  \begin{tabular}{| c | c | c | c | c |}
+    \hline
+    AIC & BIC & logLik & Random Effect standard dev & Residual standard dev \\
+    -1847 & -1836 & 927.0 & 0.0031 & 0.0235 \\
+    \hline
+  \end{tabular}
+\end{center}
 
 \section{GLMM Gamma Model}
-When determining the most effective model to use, several differnt examples were
+When determining the most effective model to use, several examples were
 explored across the 4 different data sets: a linear model without a year effect,
@@
 with a year effect, a gamma model without a year effect and link set to log instead
-of the default which is gaussian.  The R package lme4 was very helpful with this
+of the default which is Gaussian.  The R package lme4 was very helpful with this
 as it supported the Generalized Linear Mixed Model (GLMM) which allowed both
-linear and gamma mized effects models to be built \citep{Rpkg:lme4}.A GLMM is a type of model
+linear and gamma mixed effects models to be built \citep{Rpkg:lme4}. A GLMM is a type of model
 that supports inputs that are from different distributions.  The model that we use is for the
@@
 the model used in the paper, the fixed effect comes in the form of an intercept, and then
-either a year effect, batch effect, or both are used.  Initially a linear mixed effects
-model was attempted to be implemented, but due to a right hand skew in the reaction times, 
-a gamma distribution seemed better equipped to model the data (The rightskew is not 
-particularly suprising as it is not possible to register a reaction time below 0.0 
-seconds, resulting in a lower bound but no upper bound for outliers). In the code, the 
+there is a random effect which in this case we will call the venue effect.
+
+In the code, the 
 link parameter under Gamma() was changed to be "log" as logarithmic functions are better 
@@
 difficult to explain or interpret as it means that some heats of races were either
-faster or slower and that everyone in the batch reacted in a similar way.  Although
+faster or slower and that everyone in the batch reacted similarly.  Although
 it is possible that one athlete reacting faster could lead to others reacting
@@
 heats that had fast and slow reaction times are all within 0.12 seconds of each
-other.  It is much easier however to explain the year effect.  The technology
+other.  It is much easier however to explain the venue effect.  The technology
 that World Athletics has used has differed since 1999 and theoretically the
@@
 state of the art.  Any other weather or climate related factors such as humidity,
-precipitation, elevation, etc would all be negated when looking at the batch
-effect, but are all factors that could potentially impact the year effect.
+precipitation, elevation, etc. would all be negated when looking at the batch
+effect, but are all factors that could potentially impact the venue effect.
 
@@
 pooled data for the linear model, gamma model, gamma and log model for when
-2022 was included versus when it was not.  What this shows is that the addition
+2022 was included versus when it was not.  This shows that the addition
 of 2022 to the linear model increased the model's variation and that the addition
@@
 
-\section{2019 and National Data}
+\section{Further Analysis}
 This problem of whether the sensor was faulty can be further explored in
@@
 next, but that is not necessarily the results that were found.  The R package
-"clusrank" was used to perform a clustered Wilcoxon signed-rank test.  This was
+"clusrank" was used to perform a clustered Wilcoxon rank sum test.  This was
 chosen over a typical t test so as to avoid any assumptions about the distribution
@@
 
-Using the "clusWilcox.test" function, a paired and clustered Wilcoxon signed-rank
-test was performed wtih the cluster set to the be the Athlete and group set to Year.
-Two different methods were used: "rgl" which is (Prof Yan should help with this)
-Paired was set to be "true" to perform a matched pairs test, and the alternative
-hypothesis was set to be "greater" as we want to know if the 2019 reaction times
-were larger than the 2022 times by a significant margin.
+%Citations needed
+Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
+test was performed wtih the cluster set to be the Athlete and group set to Year.
+The method was set to be "ds" which is used for the Data and Satten method, which
+is preferable when the cluster sizes are informative.  We do not care that some
+athletes made it further in either competition than another or that it should
+not matter how elite an athlete is at the 110 meter hurdles once we get to the
+national international competition level.  Cluster sizes vary from 2 observations
+to 6.
 
-When the clusrank analysis was performed, the "ds" method produced a z score of
-3.5271 and p value of 0.0002101 and the "rgl" method produced a z score of 3.665
-and a p value of 0.0001237.  These values are for a one-sided test and show that
+When the clusrank rank sum test was performed, the "ds" method produced a z score of
+2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
 the mean reaction time on average in 2019 was larger than the mean reaction time
-in 2022 for athletes who competed in both championships.
+in 2022 for athletes who competed in both championships at an alpha level of 0.01.
+This provides good evidence that athletes who competed in both 2019 and 2022 got
+faster during that timespan.
 
@@
 competitions took place between June and July 2022 and the World Championships
-were in August.  The equipment used to measure reaction time cannot be guaranteed
+were in August.  Thus it should be expected that elite athletes such as these
+will not get reasonably faster in such a short time span.
+
+The equipment used to measure reaction time cannot be guaranteed
 to be the same from national to international competition, but the time span between
 meets has been reduced considerably.  Repeating the data collection steps and
-again performing a clustered analysis for a matched pairs Wilcoxon signed-rank test
+again performing a clustered analysis for a Wilcoxon ranked sum test
 results in two new z scores and p values.  The "ds" method produced a z score of
-3.6755 and p value of 0.0001187 and the "rgl" method produced a z score of 3.7618
-and a p value of 8.434e-05. These results are highly significant and show that
+3.6069 and p value of 0.000155.  This result is highly significant and shows that
 the times at the national stage of competition were significantly higher than
@@
 
-There are several different data sets that could be used and explored in determining 
-whether or not Devon Allen was unjustly disqualified.  One of the best examples of this
-is the women's 100 meter hurdles. Although the distance of the race is 10 meters shorter
-than the men's race, the reaction time data is still important and can be used to
-essentially double the data size.
-However, one of the largest issues with this is that on average, men are able to react quicker than women.
-A study from the 2008 Beijing Olympics found that the p value of comparing the mean fastest reaction time between men and women was below 0.001 and the differnce is in part because of the greater force threshold men are able to exert on the starting blocks \citep{Lipps}.
-The data from including both men and women could follow a bimodal distribution which would be difficult to model and not helpful for determining if Devon Allen's reaction time was the result of a faulty system or if it is an argument
-for changing the 0.10 second minimum barrier that currently exists.  Thus it was decided
-to look at men's and women's data separately.  Another decision that was made was to look
-at the data from the semifinals and finals together but from the heats or preliminary rounds.  The reason for this was because it was reasoned that in the finals athletes may
-be a little more anxious, more "jumpy" and that under stress and they would react faster.
-Additionally, exceptionally strong atheletes may not race as hard in the early heats in
-order to reduce the risk for injury and to save themselves for the later stages of
-competition that matter more.
-While the preliminary rounds also provide a bulk of the total observations, they are not as relevant to Allen's disqualification, which occured during the finals.
-Thus, the decision was made to look at data from the men's semifinals
-and finals in a "pooled" data set.
-%%%Include picture
+There is not a reasonable explanation for this.  It is inconceivable that so many
+athletes would improve such a significant amount in such a small amount of time.
+6 countries are represented in this comparison, which drastically decreases the
+chances of this issue being caused by one specific country.  The United States
+athlete's improvement in times was discussed earlier in the paper, but this data
+%Use a reference in this part
+shows that on average there was improvement for British, Polish, French, Brazilian,
+and Spanish athletes.
 
@@
 
-While this sample is too small in order
-to perform a meaningful matched pairs t test, or other similar method, it is apparent
-that Allen, Cunningham, and Holloway's times are faster at the World Track and Field
-Championship compared to the United States.  There does not appear to be a particular
-reason for this, as at the World Track and Field Championship, no two United States
-athetes were in the same batch in any race.
-
-An anova test that compared the mixed effects model for only a batch effect with the
-model that included both a batch and year effect produced a p value of 0.01059, meaning
-that the model with both mixed effects was stronger, but not significantly so.  However,
-when looking at the density plots, the probability of observing a reaction time of below
-0.1 was found to be approximately zero when looking at the model that inclued both effects,
-but the probability was 0.0066 when only the batch effect was included.  Based on the density
-plot and how close the both effects model was to being statistically signficant at an alpha
-value of 0.05, the model with both the batch and year effect was chosen to be analyzed further.
-The random effects: Batch and Year both had relatively small variances: 0.003455 and 0.001714
-respectively.  For the fixed effect the estimate was -1.87968 and the p value was less than
-$2\cdot10{-16}$, which is extremely small.  The model can reparameterize the gamma distribution,
-which is calculated by $e^{-1.87968}$ which is equal to 0.152.  The variance of the gamma
-distribution is 0.0329 from the R output.  Thus a gamma distribution can be created from
-this, and the probability of a reaction time below 0.1 is essentially zero.
-
-The model accuracy improves when certain batches are removed from the data.  Batch 41 in
-particular has extremely high reaction times.  This outlier can likely be explained by a
-few different causes.
-Batch 41 is a semifinals heat of the 2013 World Championship.
-2013 had the highest average reaction time from the pooled data set.
-2013 was also the second world championship after the rule change that imposed the one strike policy on
-disqualifications. 2011 and 2013 had the highest average reaction time in the pooled
-data and were the two years after the rule change.
-What this suggests is that there could have been a psychological effect of the stricter policy that resulted in athletes being more cautious and reacting slower than they had in the previous two years under the more lax guidelines from 2007 to 2009 in order to avoid a false start and thus a disqualification.
-In Batch 41 particularly two runners did not finish, and two runners reacted extremely slow: 0.327 and 0.397 (Remember that Devon Allen's reaction time was
-0.099 seconds in the 2022 Championship final and 0.101 in the 2022 semifinal).  Removing this batch and others could greatly improve the strength of the model.  However, while
-it is a high outlier, there is insufficient information available as to why the
-reaction times are so high, and no basis for why it should be improved other than because
-it is an outlier.
-
 \begin{figure}[h]
@@
 as to why there is so much of a dropoff in reaction time from 2013 to 2015 or the spike
-from 2001 to 2003.
+from 2001 to 2003.  The variation from year to year is slightly concerning in the
+sense that if there is a strong year effect then there is strong variation between
+years.  This is not desirable and what this could possibly suggest is that
+there is a lot of randomness in how athletes react or that there is at least one
+other factor that affects reaction time. If the timing systems are succeptible to
+external factors that affect how athletes react, then this is something that 
+IAAF may want to consider when they select cities to host an important event 
+such as the Track and Field World Championship. Additionally, Seiko may want to 
+do more research on their timing systems to see if climate factors can impact
+the precision of the accuracy of the equipment.
 
@@
 Allen.
-
-\section{Simulations}
-\label{sec:simulation}
 
@@
 times every 2 years (The World Championship is biennial), every 588 years there
-should be a time as extreme as Allen's.  This seems extremely unlikely and thus
-it seems like there may be an additonal explanantion for the fast reaction times.
-
-If IAAF wanted the probability for a false start to be once every 100 years,
-then they should change the false start barrier to
+should be a time as extreme as Allen's.  This seems extremely unlikely, and thus
+it seems like there may be an additional explanation for the fast reaction times.
 

### 2023-03-21 Manuscript/manuscript.tex
Small changes

@@
 athletes in the World Championship 110 meter hurdles, was that they were significantly
-faster than at the USA Track Championships.  From June 23-26 2022, USA held it's Track
+faster than at the USA Track Championships.  From June 23-26 2022, USA held its Track
 and Field Championship to decide who to send to the World Championships at Hayward 
@@
 as it supported the Generalized Linear Mixed Model (GLMM) which allowed both
-linear and gamma mixed effects models to be built \citep{Rpkg:lme4}. A GLMM is a type of model
+linear models with various distributions to be modeled\citep{Rpkg:lme4}. A GLMM is a type of model
 that supports inputs that are from different distributions.  The model that we use is for the
@@
 the model used in the paper, the fixed effect comes in the form of an intercept, and then
-there is a random effect which in this case we will call the venue effect.
+there is a random effect which in this case we will call the venue effect.  The
+venue effect is responsible for the variation between years and tries to explain
+why some years have athletes with faster reaction times than others.
 
@@
 \end{figure}
+
+The effect of adding link = "log" is that
 
@@
 the times at the national stage of competition were significantly higher than
-the reaction times at the World Championships.
+the reaction times at the World Championships even for an alpha level of 0.001.
 
@@
 
-Devon Allen's disqualification at the World Track and Field Championships was
-possibly the result of both a faulty timing equipment and Devon Allen
-reacting extremely quickly to the start.  The code and data from R showed that
-the 2022 World Track and Field Championships were a low outlier compared to
-every other year in terms of average reaction time.  Additionally, it was shown
-through a gamma mixed effects model that there are significant year and batch
-effects that greatly impact reaction time.  However, this practically does not make much sense
-as there is no reason for so much random variation without much of a trend.
-There has not been a consistent decrease or increase since 1999, much of the data
-for reaction times has been random and unpredictable.  Thus while it seems easy
-to conclude that Devon Allen was wrongly disqualified, that may not necessarily
-be the case.  If there are issues with reaction time such as athletes approaching the 0.10 
-second barrier again at the next World Track Championship (August 2023), then World Athletics 
-needs to adjust that threshold and allow for faster reaction times so that athletes with superb 
-reflexes are not penalized like how Allen was in 2022.  World Athletics and Seiko 
-may want to do further testing of their own to determine what is a human's physically
-fastest reaction time, and if they choose to do so, they should start with Devon
-Allen.
-
-A simulation was performed on the gamma model with the year effect and link set
-to "log". At a size of n=12,000,000, the calculated probability of a reaction
+To try and approximate the distribution of the reaction times, R code was written
+to try and estimate the probabilites of observing a reaction time as extreme as
+0.1 seconds.At a size of n=12,000,000, the calculated probability of a reaction
 time being below 0.1 seconds was as follows for the various different data sets:
@@
 
+It is interesting that the probability of observing a reaction time below 0.1
+seconds increases drastically for the finals data and increases by a smaller margin
+for the pooled data set after 2022 is included in the data set.  By including 2022
+in the data, the chances of observing an extreme time are increased, showing that
+2022 is a high outlier for reaction times compared to other years.
+
+Devon Allen's disqualification at the World Track and Field Championships was
+possibly the result of both a faulty timing equipment and Devon Allen
+reacting extremely quickly to the start.  The code and data from R showed that
+the 2022 World Track and Field Championships were a low outlier compared to
+every other year in terms of average reaction time.  Additionally, it was shown
+through a gamma mixed effects model that there are significant year and batch
+effects that greatly impact reaction time.  However, this practically does not make much sense
+as there is no reason for so much random variation without much of a trend.
+There has not been a consistent decrease or increase since 1999, much of the data
+for reaction times has been random and unpredictable.  Thus while it seems easy
+to conclude that Devon Allen was wrongly disqualified, that may not necessarily
+be the case.  If there are issues with reaction time such as athletes approaching the 0.10 
+second barrier again at the next World Track Championship (August 2023), then World Athletics 
+needs to adjust that threshold and allow for faster reaction times so that athletes with superb 
+reflexes are not penalized like how Allen was in 2022.  World Athletics and Seiko 
+may want to do further testing of their own to determine what is a human's physically
+fastest reaction time, and if they choose to do so, they should start with Devon
+Allen.
+
 \section{Appendix}

### 2023-03-21 Manuscript/manuscript.tex
removed figure plots, please email me if this does not compile correctly

@@
 below shows the box plot data.
-\begin{figure}[h]
-  \centering 
-  \includegraphics[scale = 0.5]{DataBoxPlot}
-  \label{fig:DataBoxPlot}
-\end{figure}
-
-\begin{figure}[h]
-  \centering 
-  \includegraphics[scale = 0.5]{DataScatterPlot}
-  \label{fig:DataScatterPlot}
-\end{figure}
 
@@
 
-\iffalse
 \begin{center}
@@
   \end{center}
-\fi
 
@@
 equipped to deal with high outliers.  The graph below shows the skewed outliers: ~\ref{fig:Skew}.
-\begin{figure}[h]
-  \centering 
-  \includegraphics[scale = 0.6]{Skew}
-  \label{fig:Skew}
-\end{figure}
 
@@
 in altering the gamma model.
-
-\begin{figure}[h]
-  \centering
-  \begin{minipage}{0.45\textwidth}
-      \centering
-      \includegraphics[width=0.9\textwidth]{Men_Density_Year} % first figure itself
-      \caption{Density plot using pooled men's data and the year effect}
-  \end{minipage}\hfill
-  \begin{minipage}{0.45\textwidth}
-      \centering
-      \includegraphics[width=0.9\textwidth]{Women_Density_Year} % second figure itself
-      \caption{Density plot using pooled women's data and the year effect}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-  \begin{minipage}{0.45\textwidth}
-      \centering
-      \includegraphics[width=0.9\textwidth]{Men_Density_Batch} % first figure itself
-      \caption{Density plot using pooled men's data and the batch effect}
-  \end{minipage}\hfill
-  \begin{minipage}{0.45\textwidth}
-      \centering
-      \includegraphics[width=0.9\textwidth]{Men_Density_Both} % second figure itself
-      \caption{Density plot using pooled men's data and both year and batch effect}
-  \end{minipage}
-\end{figure}
 
@@
 and a possibly faulty sensor.
-
-\begin{figure}[h]
-  \centering 
-  \includegraphics[scale = 0.6]{TBY}
-  \label{fig:TBY}
-\end{figure}
 

### 2023-03-21 Manuscript/manuscript.tex
comments on styles

@@
 
+\jy{Avoid references in abstract.}
+
+\jy{The bib file needs quality control. The bib keys need to have a consistent
+  style. I suggest using the keys from google scholar.}
+
+\jy{Practice my writing tips in the stat-wriging notes: chapter 2 on
+  latex. E.g., no vertical/horizontal lines in professional tables. Use the
+  tables in my template.}
+
 \section{Introduction}
 \label{sec:intro}
+
+\jy{Not a good opening. Could start by telling the story, recapping the major
+  events and discussions on the internet. Then dig into the relevant
+  literature. Finally outline the contributions of this work.
+}
 

### 2023-03-25 Manuscript/manuscript.tex
added paper structure

@@
 
-\doublespace
-
 \begin{abstract}
@@
 
+  \jy{Avoid references in abstract.}
+
+  \jy{This is only the background. An abstract also needs to summarize the contributions.}
 %\bigskip
@@
 
-\jy{Avoid references in abstract.}
+\doublespace
 
@@
 
+\jy{Turn on column numbers in your editor to keep line width under 80.}
+
 \section{Introduction}
@@
 }
+
+\jy{use the following structure and flow:
+  1) A paragraph to tell the story;
+  2) A paragraph on informal statistical analyses on the internet/social media;
+  3) A paragraph on scholarly articles on reaction time in the literature;
+  4) A paragraph on what's new: methods; results.
+  5) A paragraph of roadmap.}
 
@@
 television broadcast.  Since 2011, IAAF has instituted the one false start rule.\citep{Pilianidis}
+
+\section{Data} \label{sec:data}
+
+\jy{Describe you data types; give details on sources; introduce notations for
+  both types of data
+  present exploratory plots.}
+
+\section{Methods} \label{sec:meth}
+
+\jy{set up the models/methods using the notations from last section; cite the
+  right references.}
+
+\subsection{GLMM}
+
+\jy{explain the gamma model in details}
+
+\subsection{Rank-based comparison for Clustered}
+
+\section{Results}
+
+\jy{Presents and discuss the results here}
+
+\section{Concluding Remarks}
+
+\jy{summarize your findings; discuss limitations and possible future works.}
 

### 2023-03-25 Manuscript/manuscript.tex
Edits to introduction

@@
   seconds faster than what is allowed.  Following the games, bloggers on the running website 
-  LetsRun scrutinized the results and performed non-statistical analysis of the data \citep{Johnson}
+  LetsRun scrutinized the results and performed non-statistical analysis of the data.
   They found that the reaction time data from the 2022 World Track Championships seemed to be faster 
@@
   analysis of the data.  Devon Allen had previously ran the third fastest time in history when 
-  he ran 12.84 earlier in 2022 and was considered to be one of the favorites to win \citep{Preview}.  His 
+  he ran 12.84 earlier in 2022 and was considered to be one of the favorites to win.  His 
   disqualification thus prompted questions into whether the timing equipment used at the World
@@
   timing equipment stated he did.
+
+  This paper questions the reaction time disqualification barrier, which is
+  currently 0.1 seconds to determine whether this is a reasonable threshold as
+  well as look at data that calls into question whether the timing equipment
+  used at the 2022 Track and Field World Championships was faulty.
+  There is little at this point that could be done to rectify any potential 
+  mis-disqualification other than to try to prevent it from happening again.  
+  This matter needs to be addressed because Devon Allen's disqulification could 
+  be a problem that could continue to re-appear.  If athletes
+  are reacting fast enough that the timing system thinks that they are reacting
+  illegally fast, then World Athletics needs to address this and possibly change
+  the minimum reaction time allowed.
 
@@
 
+After coming in third in the 110 meter hurdles at the United States Track and 
+Field Championships in Eugene Oregon at the end of June, Allen was expected to 
+be a contender for the World Championships, which were scheduled for two months 
+later, also in Eugene. At the World Championships, Allen made it through the 
+heats and semifinals to reach the finals while posting reaction times of 0.123
+in the heats and 0.101 in the semifinal.  Allen, who attended the University
+of Oregon (Location of US World Championship and the World Championship) was
+considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
+in early June; only 0.04 seconds away from the world record \citep{Preview}.
+Thus his disqualification by reacting in 0.099 seconds instead of 0.01 was met
+with shock from both Allen and fans alike.  The crowd erupted with jeers when
+they found out that Allen was disqualified, angry at the result.  Allen himself
+appeared frustrated and angered with the result. Had he reacted just 0.001 
+seconds slower he would have been allowed to compete and been spared waiting
+another two years until the next World Track and Field Championship.
+%Do I need to cite a video?
+
 Other scholarly papers have discussed track and field reaction times but not in the
@@
 any published papers regarding Devon Allen's disqualification in particular.  This is not
-neccessarily suprising given that his disqualification occurred only last August.
+necessarily suprising given that his disqualification occurred only last August.
+Letsrun is a website that is part message board and part news.  The message board,
+which functions similarly to Reddit is very active during important running
+events, which include during the finals of major events such as
+the Olympics and the World Track and Field Championships.  However, it is the
+news section of the website that generated articles such as "Was Devon Allen
+Screwed? There's At Least a 99\% That He Was" and "The Data Keeps Pouring In and
+It Continues to Look Bad For World Athletics and Great For Devon Allen".
+%Should I include how I read LetsRun and talk about any personal experiences
+%I have using it?
 
@@
 was accurate and thus if Devon Allen was disqualified more so because of faulty
-equipment rather than reacting illegally fast.  There is little at this point
-that could be done to rectify any potential mis-disqualification other than to try
-to prevent it from happening again.  This matter needs to be addressed because Devon Allen's 
-disqulification could be a problem that could continue to re-appear.  If athletes
-are reacting fast enough that the timing system thinks that they are reacting
-illegally fast, then World Athletics needs to address this and possibly change
-the minimum reaction time allowed.  Currently the barrier is 0.10 seconds, any time
+equipment rather than reacting illegally fast.  Currently the barrier is 0.10 seconds, any time
 faster than this would result in a disqualification in the athlete who reacted too

### 2023-03-27 Manuscript/manuscript.tex
Added scholarly paper works

@@
 
+1)
 After coming in third in the 110 meter hurdles at the United States Track and 
@@
 
+2) and 3)
 Other scholarly papers have discussed track and field reaction times but not in the
-context of whether the minimum reaction time barrier is too low or whether or not timing
-systems were accurate.  Outside of the blog posts from LetsRun.com, there have not been
+context of whether the minimum reaction time barrier is too low or whether timing
+systems were accurate.  Outside the blog posts from LetsRun.com, there have not been
 any published papers regarding Devon Allen's disqualification in particular.  This is not
-necessarily suprising given that his disqualification occurred only last August.
+necessarily surprising given that his disqualification occurred only last August.
 Letsrun is a website that is part message board and part news.  The message board,
@@
 Screwed? There's At Least a 99\% That He Was" and "The Data Keeps Pouring In and
-It Continues to Look Bad For World Athletics and Great For Devon Allen".
+It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
+Johnson, who created and runs the Letsrun website argued that there was an error
+with the timing equipment at the 2022 Track and Field World Championship and
+cited graphs, statistics such as median reaction time, and a reference to a study
+published in cooperation with IAAF in 2009 questioning the validity of the 0.1 
+second reaction barrier.  The paper analyzed sprinters reaction times
+and the biological processes that went into reacting to the starting gun and
+concluded that athletes can react in as low as 0.08 seconds. It even recommended
+to IAAF that the disqualification barrier be decreased to allow sprinters to
+react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
+Additionally, it was recommmended to IAAF that high speed cameras be used to
+change the reaction time criteria from pushing off on the block to body movement.
+This could provide a more accurate representation of reaction time, and thus the
+possibility of a false disqualification would be lowered \citep{Komi}.
+Despite the study, IAAF has not changed the reaction barrier since it was
+implemented at the 1999 World Track and Field Championships.  It is confusing
+as to why IAAF would commission a study only to disregard the results and not
+make the necessary changes.  It is unknown how many athletes since 2009 could
+have potentially benefited from this, but it is probable that there are others
+besides Allen who were disqualified.
+
 %Should I include how I read LetsRun and talk about any personal experiences
 %I have using it?
+
+4)
+This paper will try to model the data using a 
+When determining the most effective model to use, several examples were
+explored across the 4 different data sets: a linear model without a year effect,
+linear model with year effect, a gamma model without a year effect, a gamma model
+with a year effect, a gamma model without a year effect and link set to log instead
+of the default which is Gaussian.  The R package lme4 was very helpful with this
+as it supported the Generalized Linear Mixed Model (GLMM) which allowed both
+linear models with various distributions to be modeled\citep{Rpkg:lme4}. A GLMM is a type of model
+that supports inputs that are from different distributions.  The model that we use is for the
+Gamma distribution for the fixed effects part of the model and a random effects part.  In
+the model used in the paper, the fixed effect comes in the form of an intercept, and then
+there is a random effect which in this case we will call the venue effect.  The
+venue effect is responsible for the variation between years and tries to explain
+why some years have athletes with faster reaction times than others.
+
+In the code, the 
+link parameter under Gamma() was changed to be "log" as logarithmic functions are better 
+equipped to deal with high outliers.
 
@@
 \section{GLMM Gamma Model}
-When determining the most effective model to use, several examples were
-explored across the 4 different data sets: a linear model without a year effect,
-linear model with year effect, a gamma model without a year effect, a gamma model
-with a year effect, a gamma model without a year effect and link set to log instead
-of the default which is Gaussian.  The R package lme4 was very helpful with this
-as it supported the Generalized Linear Mixed Model (GLMM) which allowed both
-linear models with various distributions to be modeled\citep{Rpkg:lme4}. A GLMM is a type of model
-that supports inputs that are from different distributions.  The model that we use is for the
-Gamma distribution for the fixed effects part of the model and a random effects part.  In
-the model used in the paper, the fixed effect comes in the form of an intercept, and then
-there is a random effect which in this case we will call the venue effect.  The
-venue effect is responsible for the variation between years and tries to explain
-why some years have athletes with faster reaction times than others.
-
-In the code, the 
-link parameter under Gamma() was changed to be "log" as logarithmic functions are better 
-equipped to deal with high outliers.  The graph below shows the skewed outliers: ~\ref{fig:Skew}.
-
-The effect of adding link = "log" is that
 

### 2023-03-27 Manuscript/manuscript.tex
Added more to intro

@@
 react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
-Additionally, it was recommmended to IAAF that high speed cameras be used to
+Additionally, it was recommended to IAAF that high speed cameras be used to
 change the reaction time criteria from pushing off on the block to body movement.
@@
 4)
-This paper will try to model the data using a 
-When determining the most effective model to use, several examples were
-explored across the 4 different data sets: a linear model without a year effect,
-linear model with year effect, a gamma model without a year effect, a gamma model
-with a year effect, a gamma model without a year effect and link set to log instead
+One of the first objectives of the paper is to try to model the data, which is
+the reaction times of every non-disqualification since the 1999 World Championships. 
+When determining the most effective model to use, several variations were
+explored across the 4 different data sets: a linear model without a venue effect,
+linear model with venue effect, a gamma model without a venue effect, a gamma model
+with a venue effect, a gamma model without a venue effect and link set to log instead
 of the default which is Gaussian.  The R package lme4 was very helpful with this
-as it supported the Generalized Linear Mixed Model (GLMM) which allowed both
-linear models with various distributions to be modeled\citep{Rpkg:lme4}. A GLMM is a type of model
-that supports inputs that are from different distributions.  The model that we use is for the
-Gamma distribution for the fixed effects part of the model and a random effects part.  In
-the model used in the paper, the fixed effect comes in the form of an intercept, and then
-there is a random effect which in this case we will call the venue effect.  The
-venue effect is responsible for the variation between years and tries to explain
-why some years have athletes with faster reaction times than others.
+as it supported the Generalized Linear Mixed Model (GLMM) which allowed 
+linear models with various distributions to be modeled\citep{Rpkg:lme4}.  The
+model we use is a liner model with a fixed effect and random effect and follows
+the gamma distribution. The fixed effect comes in the form of an intercept, and the
+random effect will be called the venue effect and is reflective of how significant
+the changes in the reaction time is for each venue, i.e. the venue effect is 
+responsible for the variation between years.
 
-In the code, the 
-link parameter under Gamma() was changed to be "log" as logarithmic functions are better 
-equipped to deal with high outliers.
+Another objective of this paper is to compare data for athletes who have competed
+at additional competitions to see how their reaction times compare from previous
+competitions to this one. First we can look at the reaction times of athletes who competed at the
+World Track and Field Championship in both 2019 and 2022 and compare the times,
+and secondly we can look at how athletes performed compared to how they did at
+their national/qualifiers meet.  We would expect that athletes of this caliber,
+who are world-class athletes, to not improve significantly from one year to the
+next.  In order to reduce any chances of athletes improving reaction time, we
+can also look at data from qualifying meets.  Almost every athlete had to qualify 
+by proving they are elite at the national level, usually by coming in at least
+fourth. Thus, data was compiled on athletes who competed
+at the 2022 World Track and Field Championships on what these athletes also ran 
+at their country's national championship. If there was any argument that athletes
+have improved their reaction time from 2019 to 2022, it should now be considered
+nearly null, as the national competitions took place between June and July 2022 
+and the World Championships were in August.  Thus, it should be expected that 
+elite athletes such as these will not get reasonably faster in such a short time
+span. The equipment used to measure reaction time cannot be guaranteed to be
+the same from national to international competition, but the time span between
+meets has been reduced considerably.
+
+5)
 
@@
 
+\section{Data} \label{sec:data}
+
+\jy{Describe you data types; give details on sources; introduce notations for
+  both types of data
+  present exploratory plots.}
 Something that is extremely important to note in the Data is the inconsistency of World Athletics 
@@
 make already long track meets more tedious, both for athletes and viewers watching the 
-television broadcast.  Since 2011, IAAF has instituted the one false start rule.\citep{Pilianidis}
-
-\section{Data} \label{sec:data}
-
-\jy{Describe you data types; give details on sources; introduce notations for
-  both types of data
-  present exploratory plots.}
+television broadcast.  Since 2011, IAAF has instituted the one false start rule
+\citep{Pilianidis}.
 
@@
 Sorted By Model instead:
-Linear Model with no year effect
+Linear Model with no venue effect
 \begin{center}
@@
 
-Linear model with year effect
+Linear model with venue effect
 \begin{center}
@@
 
-Gamma model with year effect
+Gamma model with venue effect
 \begin{center}
@@
 
-Gamma model with year effect and link=log
+Gamma model with venue effect and link=log
 \begin{center}
@@
     \hline
-    Dataset & Year Effect Standard Error \\
+    Dataset & venue effect Standard Error \\
     1999-2019 & 0.0495 \\
@@
 Once the Gamma mixed effects model was chosen, it needed to be decided which combination
-of effects to use.  It was possible to include a year effect, a batch effect, or a year
+of effects to use.  It was possible to include a venue effect, a batch effect, or a year
 and batch effect.  Density plots were created that simulated the probability of a mean
@@
 \section{Further Analysis}
-This problem of whether the sensor was faulty can be further explored in
-two ways. First we can look at the reaction times of athletes who competed at the
-World Track and Field Championship in both 2019 and 2022 and compare the times,
-and secondly we can look at how athletes performed compared to how they did at
-their national/qualifiers meet.  We would expect that athletes of this caliber,
-who are world-class athletes, to not improve significantly from one year to the
-next, but that is not necessarily the results that were found.  The R package
+, but that is not necessarily the results that were found.  The R package
 "clusrank" was used to perform a clustered Wilcoxon rank sum test.  This was
@@
 
-Another important data set that was investigated was the 2022 national competition
-data.  In order to send between 1-4 athletes to the World Championships, almost
-every athlete had to qualify by proving they are elite at the national level.
-Thus, data was compiled on athletes who competed at the 2022 World Track and 
-Field Championships on what these athletes also ran at their country's national
-championship.  If there was any argument that athletes have improved their reaction
-time from 2019 to 2022, it should now be considered nearly null, as the national
-competitions took place between June and July 2022 and the World Championships
-were in August.  Thus it should be expected that elite athletes such as these
-will not get reasonably faster in such a short time span.
-
-The equipment used to measure reaction time cannot be guaranteed
-to be the same from national to international competition, but the time span between
-meets has been reduced considerably.  Repeating the data collection steps and
+  Repeating the data collection steps and
 again performing a clustered analysis for a Wilcoxon ranked sum test
@@
 from 2001 to 2003.  The variation from year to year is slightly concerning in the
-sense that if there is a strong year effect then there is strong variation between
+sense that if there is a strong venue effect then there is strong variation between
 years.  This is not desirable and what this could possibly suggest is that

### 2023-03-28 Manuscript/manuscript.tex
Major changes to methods and results sections

@@
   present exploratory plots.}
+
 Something that is extremely important to note in the Data is the inconsistency of World Athletics 
@@
 
-\section{Methods} \label{sec:meth}
-
-\jy{set up the models/methods using the notations from last section; cite the
-  right references.}
-
-\subsection{GLMM}
-
-\jy{explain the gamma model in details}
-
-\subsection{Rank-based comparison for Clustered}
-
-\section{Results}
-
-\jy{Presents and discuss the results here}
-
-\section{Concluding Remarks}
-
-\jy{summarize your findings; discuss limitations and possible future works.}
-
-\section{GLMM Model}
-\label{sec:GLMM}
 Data was copied from the World Athletics website and pasted into an Excel
@@
 111 corresponds to the 1999 finals). The data that looked at in the paper
-is best visualized using a scatterplot and box plot. Figure~\ref{fig:DataBoxPlot}
-below shows the box plot data.
+is best visualized using a box plot. Figure~\ref{fig:DataBoxPlot}.  Included are
+box plots of the reaction times of the men's semi finals and finals pooled
+together and just the finals.
 
-One of the most interesting things found was that the reaction times from the USA
-athletes in the World Championship 110 meter hurdles, was that they were significantly
-faster than at the USA Track Championships.  From June 23-26 2022, USA held its Track
+\begin{figure}[h]
+  \centering 
+  \includegraphics{DataBoxPlot}
+  \label{fig:DataBoxPlot}
+\end{figure}
+
+\begin{figure}[h]
+  \centering 
+  \includegraphics{BoxplotFinals}
+  \label{fig:BoxplotFinals}
+\end{figure}
+
+\section{Methods} \label{sec:meth}
+
+\jy{set up the models/methods using the notations from last section; cite the
+  right references.}
+
+\subsection{GLMM}
+A gamma mixed linear effects model was chosen as the best model to model the data.
+Many models were tried and compared to determine what the strongest model.  For
+a fixed data set such as the men's pooled data including 2022, the models with
+the venue effect included performed better than the models without the venue
+effect for each set of comparable models; as indicated by the lower AIC and
+higher log likelihood.  Thus, the venue effect is significant and should be
+included in the model.  It can also be established that a gamma distribution
+better models the data than linear data from comparing the linear model with
+venue effect and them gamma model with a venue effect.  The final decision that
+was made in model selection was to include link = log in the model.  Despite
+%Does this part need to be changed?
+doing very little to alter the AIC or log likelihood, the log scale was included
+in order to try and better deal with high outliers. %Include Residuals plot?
+
+All Men's semi finals and finals data
+\begin{center}
+  \begin{tabular}{|c  c  c  c  c  c  c |}
+   \hline 
+   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   \hline\hline
+   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
+   \hline
+   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
+   \hline
+   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
+   \hline
+   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
+   \hline
+   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
+   \hline
+   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
+   \hline
+  \end{tabular}
+  \end{center}
+
+Once the Gamma mixed effects model was chosen, it needed to be decided which combination
+of effects to use.  It was possible to include a venue effect, a batch effect, or a year
+and batch effect.  Density plots were created that simulated the probability of a mean
+density less than 0.1, which is the current threshold for a reaction time
+disqualification.  The density plots are shown below.  The batch effect is very
+difficult to explain or interpret as it means that some heats of races were either
+faster or slower and that everyone in the batch reacted similarly.  Although
+it is possible that one athlete reacting faster could lead to others reacting
+similarly, this does not make much practical sense as the difference between
+heats that had fast and slow reaction times are all within 0.12 seconds of each
+other.  It is much easier however to explain the venue effect.  The technology
+that World Athletics has used has differed since 1999 and theoretically the
+reaction time data from 2022 should be the best as the timing system should be
+state of the art.  Any other weather or climate related factors such as humidity,
+precipitation, elevation, etc. would all be negated when looking at the batch
+effect, but are all factors that could potentially impact the venue effect.
+
+\subsection{Rank-based comparison for Clustered}
+This section was motivated from a qualitative analysis that started by looking
+at how United States athletes reacted at the 2022 United States Track and Field
+Championships. From June 23-26 2022, USA held its Track
 and Field Championship to decide who to send to the World Championships at Hayward 
@@
 
+This dataset is quite small to perform any formal statistical analysis on, but
+the data was expanded to include data for athletes who competed at national
+competitions (many of which were held from May-July of 2022).  Once the data
+was collected, each athlete was designated to be a cluster and the stage of
+competition (national or international) was specified to be the group.  Next,
+a Clustered Wilcoxon rank sum test was performed using the clusrank R package
+\citep{Clusrank}.  Using a clustered rank sum test was chosen over a typical t 
+test so as to avoid any assumptions about the distribution of population means.
+
+%Citations needed
+Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
+test was performed wtih the cluster set to be the Athlete and group set to Year.
+The method was set to be "ds" which is used for the Data and Satten method, which
+is preferable when the cluster sizes are informative \citep{Clusrank}.  We do not care that some
+athletes made it further in either competition than another or that it should
+not matter how elite an athlete is at the 110 meter hurdles once we get to the
+national international competition level.  Cluster sizes vary from 2 observations
+to 6.
+
+When the clusrank rank sum test was performed, the "ds" method produced a z score of
+2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
+the mean reaction time on average in 2019 was larger than the mean reaction time
+in 2022 for athletes who competed in both championships at an alpha level of 0.01.
+This provides good evidence that athletes who competed in both 2019 and 2022 got
+faster during that timespan.
+
+Repeating the data collection steps and
+again performing a clustered analysis for a Wilcoxon ranked sum test
+results in two new z scores and p values.  The "ds" method produced a z score of
+3.6069 and p value of 0.000155.  This result is highly significant and shows that
+the times at the national stage of competition were significantly higher than
+the reaction times at the World Championships even for an alpha level of 0.001.
+
+\section{Results}
+\jy{summarize your findings; discuss limitations and possible future works.}
+
+To try and approximate the distribution of the reaction times, R code was written
+to try and estimate the probabilities of observing a reaction time as extreme as
+0.1 seconds. At a size of n=12,000,000, the calculated probability of a reaction
+time being below 0.1 seconds was as follows for the various different data sets:
+
+\begin{center}
+  \begin{tabular}{|c c|} 
+   \hline
+   Data Set & Probability RxnTime < 0.1 \\ 
+   \hline\hline
+   all finals gamma log & 0.000864 \\ 
+   \hline
+   old finals gamma log & 0.000429 \\
+   \hline
+   all pooled gamma log & 0.006847 \\
+   \hline
+   old pooled gamma log & 0.005749 \\
+  \end{tabular}
+  \end{center}
+
+Prior to the 2022 World Championship, there was a 0.0425\% chance (based on the
+simulated results) of observing a reaction time below 0.01 seconds in the men's
+finals.  That means that we would expect a time below the threshold to occur
+once every 2,352 starts.  Considering there are roughly 8 men's finals reaction
+times every 2 years (The World Championship is biennial), every 588 years there
+should be a time as extreme as Allen's.  This seems extremely unlikely, and thus
+it seems like there may be an additional explanation for the fast reaction times.
+
+It is interesting that the probability of observing a reaction time below 0.1
+seconds increases drastically for the finals data and increases by a smaller margin
+for the pooled data set after 2022 is included in the data set.  By including 2022
+in the data, the chances of observing an extreme time are increased, showing that
+2022 is a high outlier for reaction times compared to other years.
+
+It is worth noting that changing the reaction time barrier from 0.1 seconds to
+0.08 seconds drastically reduces the chances of observing a reaction time that
+would break the barrier.  For the dataset of all finals data, the probability
+associated with observing a reaction time below 0.08 seconds was $6.67\cdot10{-7}$.
+If IAAF were to change the reaction time the way that \citet{Komi}'s paper
+suggests, the probability would be so much lower that athletes would likely not
+have a case to make when arguing against a disqualification.
+
+The p values produced by the clustered rank sum tests were 0.000155 for the
+same athletes competing in the same year, and 0.001464 for the same athletes in
+different years.  To put it simply, there is not a reasonable explanation for 
+this.  
+It is inconceivable that so many athletes would improve such a significant
+amount in such a small amount of time. 6 countries are represented in this 
+comparison, which drastically decreases the chances of this issue being caused by
+one specific country.  The United States athlete's improvement in times was
+discussed earlier in the paper, but this data shows that on average there was
+%Use a reference in this part
+improvement for British, Polish, French, Brazilian, and Spanish athletes.
+
+It is worth noting that in the semi-finals of the World Track and
+Field Championships; Allen's reaction time was 0.101 which is only 0.001 above
+the legal limit.  What this suggests is that Allen may have strong reflexes
+and be able to react extremely well to the sound of the gun.  It seems unlikely that
+Allen would be able to correctly predict the gun with such precision two times in a row,
+which is the exact reason that the 0.10 second reaction time disqualification barrier was
+imposed in the first place.  But Allen showed that at the Championship he was able to
+react extremely fast two times, which may suggest skill rather than luck. Devon Allen is a very good athlete, as evidenced by his football career at the University of Oregon
+and making it onto the Philadelphia Eagles practice squad this past year \citep{Hurley}.
+Thus the 0.099 reaction time may not have been a product of Devon Allen
+predicting the start of the race but rather a combination of a quick reaction 
+and a possibly faulty sensor.
+
+Devon Allen's disqualification at the World Track and Field Championships was
+possibly the result of both a faulty timing equipment and Devon Allen
+reacting extremely quickly to the start.  The code and data from R showed that
+the 2022 World Track and Field Championships were a low outlier compared to
+every other year in terms of average reaction time.  Additionally, it was shown
+through a gamma mixed effects model that there are significant year and batch
+effects that greatly impact reaction time.  However, this practically does not make much sense
+as there is no reason for so much random variation without much of a trend.
+There has not been a consistent decrease or increase since 1999, much of the data
+for reaction times has been random and unpredictable.  Thus while it seems easy
+to conclude that Devon Allen was wrongly disqualified, that may not necessarily
+be the case.  If there are issues with reaction time such as athletes approaching the 0.10 
+second barrier again at the next World Track Championship (August 2023), then World Athletics 
+needs to adjust that threshold and allow for faster reaction times so that athletes with superb 
+reflexes are not penalized like how Allen was in 2022.  World Athletics and Seiko 
+may want to do further testing of their own to determine what is a human's physically
+fastest reaction time, and if they choose to do so, they should start with Devon
+Allen.
+
+\section{Appendix}
+\label{sec:appendix}
+%Put downloadable data spreadsheet
+Here is World Athletics website with the data: \url{https://www.worldathletics.org/results/world-athletics-championships}.
+Here is the data for the USA Track and Field Championships: \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
+
+Here are important statistics grouped by both model and data set:
 All Men final's Data
@@
 
-All Men's semi finals and finals data
+All Men's semifinal and finals data
 \begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c |} 
+  \begin{tabular}{||c | c c c | c c c||} 
    \hline
-   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
    \hline\hline
-   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
+   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
    \hline
-   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
+   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
    \hline
-   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
+   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
    \hline
-   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
-   \hline
-   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
-   \hline
-   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
+   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
    \hline
@@
 
-AIC with and without random effect, Gamma and log model on all semifinal and finals data
-\begin{center}
-  \begin{tabular}{|c | c|}
-    \hline
-    Random Effect? & AIC \\
-    No & -1756 \\
-    Yes & -1848 \\
-    \hline
-  \end{tabular}
-\end{center}
-
-Addition of 2022 on Gamma and log model on semifinal and finals data
-\begin{center}
-  \begin{tabular}{|c | c|}
-    \hline
-    Dataset & venue effect Standard Error \\
-    1999-2019 & 0.0495 \\
-    1999-2022 & 0.0554 \\
-    \hline
-  \end{tabular}
-\end{center}
-
-All summary statistics for Gamma and log model on all semifinal and finals data
-\begin{center}
-  \begin{tabular}{| c | c | c | c | c |}
-    \hline
-    AIC & BIC & logLik & Random Effect standard dev & Residual standard dev \\
-    -1847 & -1836 & 927.0 & 0.0031 & 0.0235 \\
-    \hline
-  \end{tabular}
-\end{center}
-
-\section{GLMM Gamma Model}
-
-Once the Gamma mixed effects model was chosen, it needed to be decided which combination
-of effects to use.  It was possible to include a venue effect, a batch effect, or a year
-and batch effect.  Density plots were created that simulated the probability of a mean
-density less than 0.1, which is the current threshold for a reaction time
-disqualification.  The density plots are shown below.  The batch effect is very
-difficult to explain or interpret as it means that some heats of races were either
-faster or slower and that everyone in the batch reacted similarly.  Although
-it is possible that one athlete reacting faster could lead to others reacting
-similarly, this does not make much practical sense as the difference between
-heats that had fast and slow reaction times are all within 0.12 seconds of each
-other.  It is much easier however to explain the venue effect.  The technology
-that World Athletics has used has differed since 1999 and theoretically the
-reaction time data from 2022 should be the best as the timing system should be
-state of the art.  Any other weather or climate related factors such as humidity,
-precipitation, elevation, etc. would all be negated when looking at the batch
-effect, but are all factors that could potentially impact the venue effect.
-
-In terms of the data that was used, it is worth noting that the standard error
-of the intercept was higher for all 6 comparisons of the data: finals data and 
-pooled data for the linear model, gamma model, gamma and log model for when
-2022 was included versus when it was not.  This shows that the addition
-of 2022 to the linear model increased the model's variation and that the addition
-of 2022 resulted in a worse model.  Although this is more of a qualitative analysis,
-this is further evidence that 2022 was a significant year in the role it played
-in altering the gamma model.
-
-\section{Further Analysis}
-, but that is not necessarily the results that were found.  The R package
-"clusrank" was used to perform a clustered Wilcoxon rank sum test.  This was
-chosen over a typical t test so as to avoid any assumptions about the distribution
-of population means.  The clusrank package allows clusters to be formed, which
-in this case are for each athlete.  The data containing each athlete's name,
-nationality, and reaction times for the 2019 and 2022 World Championship can be
-found in the csv file labeled "2019vs2022".  This data was transformed to "Clusrank2019vs2022"
-so that it better fits the clusrank package.  Each athlete was given a non-meaningful
-ID (An ID of 1 does not indicate that this athlete is better or worse than any
-other athlete).  Additionally, the "Year" column is used to determine what year
-the time was registered with "1" indicating 2019 and "2" indicating 2022.
-
-%Citations needed
-Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
-test was performed wtih the cluster set to be the Athlete and group set to Year.
-The method was set to be "ds" which is used for the Data and Satten method, which
-is preferable when the cluster sizes are informative.  We do not care that some
-athletes made it further in either competition than another or that it should
-not matter how elite an athlete is at the 110 meter hurdles once we get to the
-national international competition level.  Cluster sizes vary from 2 observations
-to 6.
-
-When the clusrank rank sum test was performed, the "ds" method produced a z score of
-2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
-the mean reaction time on average in 2019 was larger than the mean reaction time
-in 2022 for athletes who competed in both championships at an alpha level of 0.01.
-This provides good evidence that athletes who competed in both 2019 and 2022 got
-faster during that timespan.
-
-  Repeating the data collection steps and
-again performing a clustered analysis for a Wilcoxon ranked sum test
-results in two new z scores and p values.  The "ds" method produced a z score of
-3.6069 and p value of 0.000155.  This result is highly significant and shows that
-the times at the national stage of competition were significantly higher than
-the reaction times at the World Championships even for an alpha level of 0.001.
-
-There is not a reasonable explanation for this.  It is inconceivable that so many
-athletes would improve such a significant amount in such a small amount of time.
-6 countries are represented in this comparison, which drastically decreases the
-chances of this issue being caused by one specific country.  The United States
-athlete's improvement in times was discussed earlier in the paper, but this data
-%Use a reference in this part
-shows that on average there was improvement for British, Polish, French, Brazilian,
-and Spanish athletes.
-
-\section{Conclusion}
-\label{sec:conclusion}
-It is worth noting that in the semi-finals of the World Track and
-Field Championships; Allen's reaction time was 0.101 which is only 0.001 above
-the legal limit.  What this suggests is that Allen may have strong reflexes
-and be able to react extremely well to the sound of the gun.  It seems unlikely that
-Allen would be able to correctly predict the gun with such precision two times in a row,
-which is the exact reason that the 0.10 second reaction time disqualification barrier was
-imposed in the first place.  But Allen showed that at the Championship he was able to
-react extremely fast two times, which may suggest skill rather than luck. Devon Allen is a very good athlete, as evidenced by his football career at the University of Oregon
-and making it onto the Philadelphia Eagles practice squad this past year \citep{Hurley}.
-Thus the 0.099 reaction time may not have been a product of Devon Allen
-predicting the start of the race but rather a combination of a quick reaction 
-and a possibly faulty sensor.
-
-The figure below plots the average reaction time by year for the pooled men's data since
-1999.  The circle in the bottom right is for 2022, which is significantly lower than the
-data points for the previous three years, and is substially lower than the data points
-for 2011 and 2013.  This graph supports the conclusion that the timing system from 2022
-had some sort of error that caused the reaction times to be so much lower.  Assuming there
-was not an issue with the system, the large variation is confusing because there is little explanation
-as to why there is so much of a dropoff in reaction time from 2013 to 2015 or the spike
-from 2001 to 2003.  The variation from year to year is slightly concerning in the
-sense that if there is a strong venue effect then there is strong variation between
-years.  This is not desirable and what this could possibly suggest is that
-there is a lot of randomness in how athletes react or that there is at least one
-other factor that affects reaction time. If the timing systems are succeptible to
-external factors that affect how athletes react, then this is something that 
-IAAF may want to consider when they select cities to host an important event 
-such as the Track and Field World Championship. Additionally, Seiko may want to 
-do more research on their timing systems to see if climate factors can impact
-the precision of the accuracy of the equipment.
-
-To try and approximate the distribution of the reaction times, R code was written
-to try and estimate the probabilites of observing a reaction time as extreme as
-0.1 seconds.At a size of n=12,000,000, the calculated probability of a reaction
-time being below 0.1 seconds was as follows for the various different data sets:
-
-\begin{center}
-  \begin{tabular}{||c|c||} 
-   \hline
-   Data Set & P RxnTime 0.1 \\ 
-   \hline\hline
-   all finals gamma log & 0.000864 \\ 
-   \hline
-   old finals gamma log & 0.000429 \\
-   \hline
-   all pooled gamma log & 0.006847 \\
-   \hline
-   old pooled gamma log & 0.005749 \\
-  \end{tabular}
-  \end{center}
-
-Prior to the 2022 World Championship, there was a 0.0425\% chance (based on the
-simulated results) of observing a reaction time below 0.01 seconds in the men's
-finals.  That means that we would expect a time below the threshold to occur
-once every 2,352 starts.  Considering there are roughly 8 men's finals reaction
-times every 2 years (The World Championship is biennial), every 588 years there
-should be a time as extreme as Allen's.  This seems extremely unlikely, and thus
-it seems like there may be an additional explanation for the fast reaction times.
-
-It is interesting that the probability of observing a reaction time below 0.1
-seconds increases drastically for the finals data and increases by a smaller margin
-for the pooled data set after 2022 is included in the data set.  By including 2022
-in the data, the chances of observing an extreme time are increased, showing that
-2022 is a high outlier for reaction times compared to other years.
-
-Devon Allen's disqualification at the World Track and Field Championships was
-possibly the result of both a faulty timing equipment and Devon Allen
-reacting extremely quickly to the start.  The code and data from R showed that
-the 2022 World Track and Field Championships were a low outlier compared to
-every other year in terms of average reaction time.  Additionally, it was shown
-through a gamma mixed effects model that there are significant year and batch
-effects that greatly impact reaction time.  However, this practically does not make much sense
-as there is no reason for so much random variation without much of a trend.
-There has not been a consistent decrease or increase since 1999, much of the data
-for reaction times has been random and unpredictable.  Thus while it seems easy
-to conclude that Devon Allen was wrongly disqualified, that may not necessarily
-be the case.  If there are issues with reaction time such as athletes approaching the 0.10 
-second barrier again at the next World Track Championship (August 2023), then World Athletics 
-needs to adjust that threshold and allow for faster reaction times so that athletes with superb 
-reflexes are not penalized like how Allen was in 2022.  World Athletics and Seiko 
-may want to do further testing of their own to determine what is a human's physically
-fastest reaction time, and if they choose to do so, they should start with Devon
-Allen.
-
-\section{Appendix}
-\label{sec:appendix}
-%Put downloadable data spreadsheet
-Here is World Athletics website with the data: \url{https://www.worldathletics.org/results/world-athletics-championships}.
-Here is the data for the USA Track and Field Championships: \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
-
 \bibliographystyle{chicago}

### 2023-03-28 Code/Plots.Rmd
R code to create plots used in the manuscript

---
title: "Plots"
author: "Owen Fiore"
date: "2022-12-09"
output: html_document
---

```{r}
times <- read.csv("../Data/ThesisData.csv")
View(times)
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
```

#DataSets
```{r}
men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
women_finals <- subset(times, Type == "F" & Gender == "F" & Time != 'DQ')

men_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
women_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "F" & Time != 'DQ')
```

#Models
```{r}
library(lme4)
finals.lme <- lmer(RxnTime ~ (1|Year), data = men_pooled)
```

###Functions
```{r}
simfit <- function(fit2, n = 12000000) {
  vc <- as.data.frame(VarCorr(fit2))
  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
  mu <- exp(fixef(fit2) + z)
  shape <- 1 / vc$vcov[2]
  x <- rgamma(n, shape = shape, scale = mu / shape)
  x
}

x <- simfit(fit2)
mean(x < 0.1)
quantile(x, prob = c(.00001, .0001, .0005, .001))
## compare with and without random effect
plot(density(x)) # with random effect
curve(dgamma(x, shape = 1 / sfit0$dispersion,
             scale = exp(coef(fit0)) * sfit0$dispersion),
      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
```

#Figures
```{r}
pdf(file = "TBY.pdf")
plot(aggregate(men_pooled$RxnTime, list(men_pooled$Year), FUN = mean), xlab = "Year", ylab = "Reaction time in seconds", main = "Mean Reaction Time By Year")
dev.off()
```

```{r}
pdf(file = "Skew.pdf")
hist(resid(finals.lme), xlab = "Residuals of Linear mixed effects model", main = "Right skew when plotting the residuals of the linear mixed effects model")
dev.off()
```
```{r}
pdf(file = "Gamma_skew.pdf")
dev.off()
```

```{r}
pdf(file = "DataBoxplot.pdf")
boxplot(RxnTime ~ Year, data = men_pooled, xlab = "Year", ylab = "Reaction Time in seconds", main = "Men's Semifinal and Final Reaction Time Boxplot Data")
dev.off()
```

```{r}
pdf(file = "DataScatterplot.pdf")
plot(RxnTime ~ Year, data = men_pooled, xlab = "Year", ylab = "Reaction Time in seconds", main = "Men's Semifinal and Final Reaction Time Scatterplot Data")
dev.off()
```

```{r}
set.seed(1)
library(lme4)
pdf(file = "Men_Density_Year.pdf")
fit2 <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
fit0 <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
sfit0 <- summary(fit0)
simfit <- function(fit2, n = 12000000) {
  vc <- as.data.frame(VarCorr(fit2))
  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
  mu <- exp(fixef(fit2) + z)
  shape <- 1 / vc$vcov[2]
  x <- rgamma(n, shape = shape, scale = mu / shape)
  x
}

x <- simfit(fit2)
mean(x < 0.1)
quantile(x, prob = c(.00001, .0001, .0005, .001))
## compare with and without random effect
plot(density(x), main = "Men's Pooled With Only Year Effect") # with random effect
curve(dgamma(x, shape = 1 / sfit0$dispersion,
             scale = exp(coef(fit0)) * sfit0$dispersion),
      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
legend("right", title = "Probability mean less than", legend = " 0.1 is 0.006847")
dev.off()
```

```{r}
set.seed(1)
library(lme4)
pdf(file = "Men_Density_Batch.pdf")
fit2 <- glmer(RxnTime ~ (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
fit0 <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
sfit0 <- summary(fit0)
simfit <- function(fit2, n = 12000000) {
  vc <- as.data.frame(VarCorr(fit2))
  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
  mu <- exp(fixef(fit2) + z)
  shape <- 1 / vc$vcov[2]
  x <- rgamma(n, shape = shape, scale = mu / shape)
  x
}

x <- simfit(fit2)
mean(x < 0.1)
quantile(x, prob = c(.00001, .0001, .0005, .001))
## compare with and without random effect
plot(density(x), main = "Men's Pooled With Only Batch Effect") # with random effect
curve(dgamma(x, shape = 1 / sfit0$dispersion,
             scale = exp(coef(fit0)) * sfit0$dispersion),
      0.08, 0.22, add = TRUE, col = "red") # fixed effect only
legend("right", title = "Probability mean less than", legend = " 0.1 is 0.0066115")
dev.off()
```

```{r}
set.seed(1)
library(lme4)
pdf(file = "Men_Density_Both.pdf")
fit2 <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
fit0 <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
sfit0 <- summary(fit0)
simfit <- function(fit2, n = 12000000) {
  vc <- as.data.frame(VarCorr(fit2))
  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
  mu <- exp(fixef(fit2) + z)
  shape <- 1 / vc$vcov[2]
  x <- rgamma(n, shape = shape, scale = mu / shape)
  x
}

x <- simfit(fit2)
mean(x < 0.1)
quantile(x, prob = c(.00001, .0001, .0005, .001))
## compare with and without random effect
plot(density(x), main = "Men's Pooled With Year and Batch Effect") # with random effect
curve(dgamma(x, shape = 1 / sfit0$dispersion,
             scale = exp(coef(fit0)) * sfit0$dispersion),
      0.08, 0.22, add = TRUE, col = "orange") # fixed effect only
legend("right", title = "Probability mean less than", legend = " 0.1 is 0")
dev.off()
```
```{r}
set.seed(1)
library(lme4)
pdf(file = "Women_Density_Year.pdf")
fit2 <- glmer(RxnTime ~ (1 | Year), data = women_pooled, family = Gamma(link = "log"))
fit0 <- glm(RxnTime ~ 1, data = women_pooled, family = Gamma(link = "log"))
sfit0 <- summary(fit0)
simfit <- function(fit2, n = 12000000) {
  vc <- as.data.frame(VarCorr(fit2))
  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
  mu <- exp(fixef(fit2) + z)
  shape <- 1 / vc$vcov[2]
  x <- rgamma(n, shape = shape, scale = mu / shape)
  x
}

x <- simfit(fit2)
mean(x < 0.1)
quantile(x, prob = c(.00001, .0001, .0005, .001))
## compare with and without random effect
plot(density(x), main = "Women's Pooled With Only Year Effect") # with random effect
curve(dgamma(x, shape = 1 / sfit0$dispersion,
             scale = exp(coef(fit0)) * sfit0$dispersion),
      0.08, 0.22, add = TRUE, col = "deeppink") # fixed effect only
legend("right", title = "Probability mean less than", legend = " 0.1 is 0.0042296")
dev.off()
```
```{r}
year_effect <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
batch_effect <- glmer(RxnTime ~ (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
both <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
#anova(batch_effect, both)
summary(both)
```

```{r}
t.test(men_pooled$RxnTime[men_pooled$Year==2019],men_pooled$RxnTime[men_pooled$Year==2022])
```

```{r}
pdf(file = "BoxplotFinals.pdf")
boxplot(RxnTime ~ Year, data = men_finals, xlab = "Year", ylab = "Reaction Time in seconds", main = "Men's Final Reaction Time Boxplot Data")
dev.off()
```

### 2023-03-28 Manuscript/manuscript.tex
Meeting notes from 3/28 are in-line in the manuscript

@@
 
+%Abstract should be 200 words
+%Devon Allen was disqualified in 2022... discussions about dq, talk methods and 
+%models, venue effect, and conclusion
 \begin{abstract}
@@
   compared to the other datasets they looked at, but did not perform any meaningful statistical 
-  analysis of the data.  Devon Allen had previously ran the third fastest time in history when 
+  analysis of the data.  Devon Allen had previously run the third-fastest time in history when 
   he ran 12.84 earlier in 2022 and was considered to be one of the favorites to win.  His 
   disqualification thus prompted questions into whether the timing equipment used at the World
-  Track Championships was accurate and whether or not Allen actually reacted as fast as the 
-  timing equipment stated he did.
-
-  This paper questions the reaction time disqualification barrier, which is
+  Track Championships was accurate and whether Allen actually reacted as fast as the 
+  timing equipment stated he did. This paper questions the reaction time disqualification barrier, which is
   currently 0.1 seconds to determine whether this is a reasonable threshold as
@@
   mis-disqualification other than to try to prevent it from happening again.  
-  This matter needs to be addressed because Devon Allen's disqulification could 
+  This matter needs to be addressed because Devon Allen's disqualification could 
   be a problem that could continue to re-appear.  If athletes
@@
 another two years until the next World Track and Field Championship.
-%Do I need to cite a video?
+%Can talk about youtube video and put the link in text here
 
@@
 besides Allen who were disqualified.
+%Reaction time barrier has been examined before: cite reviews
+%Then go into specifics and talk about people who wanted to change barrier
+%General reviews are important
+%People have studied the biological processes, cite a few references
+%How barrier is picked by regulator
+%Read abstract and into and summarize into about one sentence what each paper contributes
+%1 and 2 can be merged but 3 need to stand alone
 
@@
 meets has been reduced considerably.
+%Simplify contribution to one paragraph: collect data and formal statistical
+%analysis such as model used
+%Move a lot of this to methods section
+5)
+%Follow Anthony's work, explicit references as to what each section does
 
-5)
-
+%Seiko section should go in the Results section.  From results it seems reasonable
+%question the data and talk about Seiko
 The contribution of this paper is to question whether or not Seiko's timing data
@@
 
+  %yij denotes ith years and j is observation in year
+  %yij will have clustered i and we use random effect to account for this
+  %When model is presented talk about how I modeled yij
+  %Zij where player is i and cluster, group is year and j
+  %Test hypothesis that reaction times are the same vs that they are not
 Something that is extremely important to note in the Data is the inconsistency of World Athletics 
@@
 together and just the finals.
-
+%Fix quotations
+%Dont need the titles will add captions, use width =5, height = 3 in aspect ratio
 \begin{figure}[h]
@@
 \subsection{GLMM}
+%Include latex math symbols for the equations
 A gamma mixed linear effects model was chosen as the best model to model the data.
@@
 
+%Throughout exploratory analysis, gamma provides better AIC
+%Spend paragraph on after fitting the model how to figure out 0.1 p value
+
+%Table should go to the results section
 All Men's semi finals and finals data
@@
 \subsection{Rank-based comparison for Clustered}
+%This should go to data section
 This section was motivated from a qualitative analysis that started by looking
@@
 \begin{center}
-  \begin{tabular}{||c | c c c | c c c||} 
+  \begin{tabular}{|c c c c c c c|} 
    \hline
@@
 %Citations needed
+%Include math formulas and the test statistic
+%Equation of test statistic is from papers cited by clusrank. Asymptootic dist
+%Was found by __ referecne and the implementation is from the clusrank R package
 Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
-test was performed wtih the cluster set to be the Athlete and group set to Year.
+test was performed with the cluster set to be the Athlete and group set to Year.
 The method was set to be "ds" which is used for the Data and Satten method, which
@@
 
+%This should go in the method section, generate from fitted model explain what
+%The R code does.  Simulate from fitted model. Use distribution of model to 
+%Figure out the p value of reaction time less than 0.1
 To try and approximate the distribution of the reaction times, R code was written
@@
 
+%Section 4.1: Gamma model results show the fitting with and without 2022,
+%talk about p values and thresholds from the simulations (do not include the
+%results in the methods section).  Discussion should show that random effect
+%and standard error increase and that 2022 is special.  Now talk about Seiko
+%timing equipment
+
+%Section 4.2 is motivated by seiko timing and look at same athletes across
+%multiple types of competition: 2019 and national
+
+%Discussion/Conclusion is beyond results.  Limitations, what type of data we
+%hope to have.
+
 The p values produced by the clustered rank sum tests were 0.000155 for the
@@
 imposed in the first place.  But Allen showed that at the Championship he was able to
-react extremely fast two times, which may suggest skill rather than luck. Devon Allen is a very good athlete, as evidenced by his football career at the University of Oregon
+react extremely fast two times, which may suggest skill rather than luck. Devon 
+Allen is a very good athlete, as evidenced by his football career at the University of Oregon
 and making it onto the Philadelphia Eagles practice squad this past year \citep{Hurley}.

### 2023-04-03 Manuscript/manuscript.tex
Abstract now ~190 words, included contributions of this paper

@@
 \begin{abstract}
-  At the 2022 World Track Championships in Eugene Oregon, Devon Allen was disqualified in the
-  Men's 110 meter hurdle final after registering a reaction time of 0.099 seconds, 0.001 
+  Devon Allen was disqualified in the Men's 110 meter hurdle final of the 2022
+  Track and Field World Championships after registering a reaction time of 0.099 seconds, 0.001 
   seconds faster than what is allowed.  Following the games, bloggers on the running website 
@@
   compared to the other datasets they looked at, but did not perform any meaningful statistical 
-  analysis of the data.  Devon Allen had previously run the third-fastest time in history when 
-  he ran 12.84 earlier in 2022 and was considered to be one of the favorites to win.  His 
-  disqualification thus prompted questions into whether the timing equipment used at the World
-  Track Championships was accurate and whether Allen actually reacted as fast as the 
-  timing equipment stated he did. This paper questions the reaction time disqualification barrier, which is
-  currently 0.1 seconds to determine whether this is a reasonable threshold as
-  well as look at data that calls into question whether the timing equipment
-  used at the 2022 Track and Field World Championships was faulty.
-  There is little at this point that could be done to rectify any potential 
-  mis-disqualification other than to try to prevent it from happening again.  
+  analysis of the data.  This paper questions the reaction time disqualification barrier, which is
+  currently 0.1 seconds to determine whether this is a reasonable threshold.
+  We employ a generalized linear mixed model (GLMM) with a random effect term
+  which will be referred to as the venue effect in order to try and model the
+  data.  Additionally, we employ a signed-rank test for clustered data to compare
+  reaction times for the same athletes at different competitions.
   This matter needs to be addressed because Devon Allen's disqualification could 
-  be a problem that could continue to re-appear.  If athletes
-  are reacting fast enough that the timing system thinks that they are reacting
-  illegally fast, then World Athletics needs to address this and possibly change
-  the minimum reaction time allowed.
+  be a problem that could continue to re-appear in future world championships.
 

### 2023-04-03 Manuscript/manuscript.tex
Revised Intro, moved seiko to results

@@
 Thus his disqualification by reacting in 0.099 seconds instead of 0.01 was met
-with shock from both Allen and fans alike.  The crowd erupted with jeers when
-they found out that Allen was disqualified, angry at the result.  Allen himself
-appeared frustrated and angered with the result. Had he reacted just 0.001 
-seconds slower he would have been allowed to compete and been spared waiting
-another two years until the next World Track and Field Championship.
-%Can talk about youtube video and put the link in text here
+with shock from both Allen and fans alike.  NBC, who was broadcasting the
+championships in the United States, later uploaded a video to their Youtube
+channel that show the disqualification and can be found here:
+\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers when
+they found out that Allen was disqualified, angry at the result.  Had Allen 
+reacted just 0.001 seconds slower he would have been allowed to compete and been 
+spared waiting another two years until the next World Track and Field Championship.
 
 2) and 3)
-Other scholarly papers have discussed track and field reaction times but not in the
-context of whether the minimum reaction time barrier is too low or whether timing
-systems were accurate.  Outside the blog posts from LetsRun.com, there have not been
-any published papers regarding Devon Allen's disqualification in particular.  This is not
-necessarily surprising given that his disqualification occurred only last August.
+Outside the blog posts from LetsRun.com, there have not been any published
+papers regarding Devon Allen's disqualification in particular.  This is not
+particularly surprising given that his disqualification occurred only last August.
 Letsrun is a website that is part message board and part news.  The message board,
@@
 with the timing equipment at the 2022 Track and Field World Championship and
-cited graphs, statistics such as median reaction time, and a reference to a study
-published in cooperation with IAAF in 2009 questioning the validity of the 0.1 
-second reaction barrier.  The paper analyzed sprinters reaction times
-and the biological processes that went into reacting to the starting gun and
-concluded that athletes can react in as low as 0.08 seconds. It even recommended
+cited graphs, statistics such as median reaction time, and compared the reaction
+times at the 2022 Track and Field World Championship.
+
+In 2009, the International Association of Athletics Federations (Also referred to
+as IAAF or World Athletics), which is the governing board for international running
+competitions, commissioned a study into changing the reaction time barrier.
+The paper analyzed sprinters reaction times and the biological processes 
+that went into reacting to the starting gun and concluded 
+that athletes can react in as low as 0.08 seconds.  One study by Pain looked at
+reaction times for nine male athletes and analyzed their reaction times.  The
+fastest athlete of the nine, had on average (even with their two fastest time 
+removed) a reaction time of 0.087 seconds and a standard deviation of 0.004
+seconds.  Two other athletes were able to react similarly when under certain
+muscular tightening conditions \citep{pain2007sprint}.
 to IAAF that the disqualification barrier be decreased to allow sprinters to
@@
 possibility of a false disqualification would be lowered \citep{Komi}.  
+  A
+follow-up study from 2012 by Pilianidis, Mantzouranis, and Kasabalis found
+significant differences between reaction times at several world championships
+between 1997 and 2009 for the men's 110 meter hurdles.  Another important
+conclusion was that for reaction times did not decrease over the twelve years
+covered by the research in the study \citep{pilianidis2012start}.  This will
+be important when considering the results in section \ref{sec:Results}.
+
 Despite the study, IAAF has not changed the reaction barrier since it was
@@
 %1 and 2 can be merged but 3 need to stand alone
-
-%Should I include how I read LetsRun and talk about any personal experiences
-%I have using it?
 
@@
 5)
+The rest of this paper is as follows: Section~\ref{sec:Data} describes how data was
+collected and begins to detail the GLMM model which is further developed in the
+Section~\ref{sec:Methods}.  The signed-rank test analysis for the same athletes is
+also included in the Section~\ref{sec:Methods}.  The Section~\ref{sec:Results}
+section provides the probabilities from the methods section and the paper ends 
+by questioning the reaction time barrier.
+
 %Follow Anthony's work, explicit references as to what each section does
@@
 %question the data and talk about Seiko
-The contribution of this paper is to question whether or not Seiko's timing data
-was accurate and thus if Devon Allen was disqualified more so because of faulty
-equipment rather than reacting illegally fast.  Currently the barrier is 0.10 seconds, any time
-faster than this would result in a disqualification in the athlete who reacted too
-quickly and the race is restarted \citep{Seiko-Timing}.  Since 2011, any athletes who false 
-start are not given a second chance, they are removed from competition \citep{False-Start}.  
-This can have massive implications as events like the Outdoor World Track and Field Championship 
-only occur typically once every two years.  It is an opportunity for athletes to earn endorsement
-deals and prove to sponsors such as Nike and Adidas that
-they deserve a contract.  Thus the significance of a disqualification, especially in the finals 
-of a major event such as the World Track and Field Championships, are very
-consequential.
-
-Since 1985 Seiko Holding Corporations has served every World Athletics Championship
-as the official timer \citep{Seiko}.  Since 1985, the technology and the ability
-to accurately predict measurements: long jump distances, false starts, total time,
-reaction time, etc. have all dramatically improved.  Seiko did not start tracking
-reaction time as an official measurement until 1999 in the men's and women's 100 meter
-hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
-technology to the World Track and Field Championships, the highest stakes in the world
-of running outside of the Olympics.
-
-%Citation needed in this section
-Seiko's technology for detecting reaction times relies on the pressure that athletes
-exert on the plate when they push off.  Their systems measure the time differential
-between when the "gun" goes off to start the race and when the pressure changes 
-\citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
-false start as it is considered that no athlete can react so quickly \citep{Seiko-Timing}.  
-Thus, a reaction time of for example 0.05, suggests that the athlete predicted the gun and it
-was luck that caused their abnormally high time.  At the World Championship level,
-World Athletics and Seiko want to remove that element of luck and thus impose the 0.1
-second barrier.
-
-In 2013, Seiko upgraded their timing equipment for sprinting events (100 meter hurdles
-falls under this catergory) \citep{WorldAthletics_2013}.  Since 1999, the highest reaction 
-times in the men's 100 meter hurdle were recorded in 2013.  That is not to say that the higher 
-times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
-to 2022, Seiko again upgraded its technology but for its jump management system.
 
@@
 
+Since 1985 Seiko Holding Corporations has served every World Athletics Championship
+as the official timer \citep{Seiko}.  Since 1985, the technology and the ability
+to accurately predict measurements: long jump distances, false starts, total time,
+reaction time, etc. have all dramatically improved.  Seiko did not start tracking
+reaction time as an official measurement until 1999 in the men's and women's 100 meter
+hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
+technology to the World Track and Field Championships, the highest stakes in the world
+of running outside of the Olympics.
+
+%Citation needed in this section
+Seiko's technology for detecting reaction times relies on the pressure that athletes
+exert on the plate when they push off.  Their systems measure the time differential
+between when the "gun" goes off to start the race and when the pressure changes 
+\citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
+false start as it is considered that no athlete can react so quickly \citep{Seiko-Timing}.  
+Thus, a reaction time of for example 0.05, suggests that the athlete predicted the gun and it
+was luck that caused their abnormally high time.  At the World Championship level,
+World Athletics and Seiko want to remove that element of luck and thus impose the 0.1
+second barrier.
+
+In 2013, Seiko upgraded their timing equipment for sprinting events (100 meter hurdles
+falls under this catergory) \citep{WorldAthletics_2013}.  Since 1999, the highest reaction 
+times in the men's 100 meter hurdle were recorded in 2013.  That is not to say that the higher 
+times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
+to 2022, Seiko again upgraded its technology but for its jump management system.
+
 \section{Appendix}

### 2023-04-04 Manuscript/manuscript.tex
Lot of copy and paste

@@
   \centering 
-  \includegraphics{DataBoxPlot}
-  \label{fig:DataBoxPlot}
+  \includegraphics[width=\textwidth,height=4in]{DataBoxPlot}
+  \caption{Boxplot of semifinals and finals data}\label{fig:DataBoxPlot}
 \end{figure}
@@
 
-\section{Methods} \label{sec:meth}
-
-\jy{set up the models/methods using the notations from last section; cite the
-  right references.}
-
-\subsection{GLMM}
-%Include latex math symbols for the equations
-A gamma mixed linear effects model was chosen as the best model to model the data.
-Many models were tried and compared to determine what the strongest model.  For
-a fixed data set such as the men's pooled data including 2022, the models with
-the venue effect included performed better than the models without the venue
-effect for each set of comparable models; as indicated by the lower AIC and
-higher log likelihood.  Thus, the venue effect is significant and should be
-included in the model.  It can also be established that a gamma distribution
-better models the data than linear data from comparing the linear model with
-venue effect and them gamma model with a venue effect.  The final decision that
-was made in model selection was to include link = log in the model.  Despite
-%Does this part need to be changed?
-doing very little to alter the AIC or log likelihood, the log scale was included
-in order to try and better deal with high outliers. %Include Residuals plot?
-
-%Throughout exploratory analysis, gamma provides better AIC
-%Spend paragraph on after fitting the model how to figure out 0.1 p value
-
-%Table should go to the results section
-All Men's semi finals and finals data
-\begin{center}
-  \begin{tabular}{|c  c  c  c  c  c  c |}
-   \hline 
-   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
-   \hline\hline
-   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
-   \hline
-   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
-   \hline
-   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
-   \hline
-   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
-   \hline
-   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
-   \hline
-   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
-   \hline
-  \end{tabular}
-  \end{center}
-
-Once the Gamma mixed effects model was chosen, it needed to be decided which combination
-of effects to use.  It was possible to include a venue effect, a batch effect, or a year
-and batch effect.  Density plots were created that simulated the probability of a mean
-density less than 0.1, which is the current threshold for a reaction time
-disqualification.  The density plots are shown below.  The batch effect is very
-difficult to explain or interpret as it means that some heats of races were either
-faster or slower and that everyone in the batch reacted similarly.  Although
-it is possible that one athlete reacting faster could lead to others reacting
-similarly, this does not make much practical sense as the difference between
-heats that had fast and slow reaction times are all within 0.12 seconds of each
-other.  It is much easier however to explain the venue effect.  The technology
-that World Athletics has used has differed since 1999 and theoretically the
-reaction time data from 2022 should be the best as the timing system should be
-state of the art.  Any other weather or climate related factors such as humidity,
-precipitation, elevation, etc. would all be negated when looking at the batch
-effect, but are all factors that could potentially impact the venue effect.
-
-\subsection{Rank-based comparison for Clustered}
-%This should go to data section
 This section was motivated from a qualitative analysis that started by looking
@@
 was collected, each athlete was designated to be a cluster and the stage of
-competition (national or international) was specified to be the group.  Next,
-a Clustered Wilcoxon rank sum test was performed using the clusrank R package
-\citep{Clusrank}.  Using a clustered rank sum test was chosen over a typical t 
-test so as to avoid any assumptions about the distribution of population means.
+competition (national or international) was specified to be the group.
+
+\section{Methods} \label{sec:meth}
+
+\jy{set up the models/methods using the notations from last section; cite the
+  right references.}
+
+\subsection{GLMM}
+%Include latex math symbols for the equations
+A gamma mixed linear effects model was chosen as the best model to model the data.
+Many models were tried and compared to determine what the strongest model.  For
+a fixed data set such as the men's pooled data including 2022, the models with
+the venue effect included performed better than the models without the venue
+effect for each set of comparable models; as indicated by the lower AIC and
+higher log likelihood.  Thus, the venue effect is significant and should be
+included in the model.  It can also be established that a gamma distribution
+better models the data than linear data from comparing the linear model with
+venue effect and them gamma model with a venue effect.  The final decision that
+was made in model selection was to include link = log in the model.  Despite
+doing very little to alter the AIC or log likelihood, the log scale was included
+in order to try and better deal with high outliers. %Include Residuals plot?
+
+%Throughout exploratory analysis, gamma provides better AIC
+%Spend paragraph on after fitting the model how to figure out 0.1 p value
+
+Once the Gamma mixed effects model was chosen, it needed to be decided which combination
+of effects to use.  It was possible to include a venue effect, a batch effect, or a year
+and batch effect.  Density plots were created that simulated the probability of a mean
+density less than 0.1, which is the current threshold for a reaction time
+disqualification.  The density plots are shown below.  The batch effect is very
+difficult to explain or interpret as it means that some heats of races were either
+faster or slower and that everyone in the batch reacted similarly.  Although
+it is possible that one athlete reacting faster could lead to others reacting
+similarly, this does not make much practical sense as the difference between
+heats that had fast and slow reaction times are all within 0.12 seconds of each
+other.  It is much easier however to explain the venue effect.  The technology
+that World Athletics has used has differed since 1999 and theoretically the
+reaction time data from 2022 should be the best as the timing system should be
+state of the art.  Any other weather or climate related factors such as humidity,
+precipitation, elevation, etc. would all be negated when looking at the batch
+effect, but are all factors that could potentially impact the venue effect.
+
+%T, generate from fitted model explain what
+%The R code does.  Simulate from fitted model. Use distribution of model to 
+%Figure out the p value of reaction time less than 0.1
+To try and approximate the distribution of the reaction times, R code was written
+to try and estimate the probabilities of observing a reaction time as extreme as
+0.1 seconds.  The code re-paramertizes the gamma distribution to be in terms of
+$\mu$ and shape rather than by $\alpha$ and $\beta$.
+
+At a size of n=12,000,000, the calculated probability of a reaction
+time being below 0.1 seconds was as follows for the various different data sets:
+
+\begin{center}
+  \begin{tabular}{|c c|} 
+   \hline
+   Data Set & Probability RxnTime < 0.1 \\ 
+   \hline\hline
+   all finals gamma log & 0.000864 \\ 
+   \hline
+   old finals gamma log & 0.000429 \\
+   \hline
+   all pooled gamma log & 0.006847 \\
+   \hline
+   old pooled gamma log & 0.005749 \\
+  \end{tabular}
+  \end{center}
+
+\subsection{Rank-based comparison for Clustered}
 
@@
 %Was found by __ referecne and the implementation is from the clusrank R package
+
+The test statistic of the DS method of the Wilcoxon rank-sum statistic is:
+$Z = \frac{S - E(S)}{\sqrt{\hat{\text{Var}}(S)}}$ where 
+$S = \mathbb{E}[W^* | X, \delta]$,
+$W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is
+the rank of $X_{i}^{*}$.
+The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
+\citet{datta2005rank} and the R package "clusrank" allows for its implementation.
+\citep{jiang2017wilcoxon}.
+
 Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
 test was performed with the cluster set to be the Athlete and group set to Year.
-The method was set to be "ds" which is used for the Data and Satten method, which
-is preferable when the cluster sizes are informative \citep{Clusrank}.  We do not care that some
+The method was set to be "DS" which is used for the Data and Satten method, which
+is preferable when the cluster sizes are informative \citep{jiang2017wilcoxon}.  We do not care that some
 athletes made it further in either competition than another or that it should
@@
 
-%This should go in the method section, generate from fitted model explain what
-%The R code does.  Simulate from fitted model. Use distribution of model to 
-%Figure out the p value of reaction time less than 0.1
-To try and approximate the distribution of the reaction times, R code was written
-to try and estimate the probabilities of observing a reaction time as extreme as
-0.1 seconds. At a size of n=12,000,000, the calculated probability of a reaction
-time being below 0.1 seconds was as follows for the various different data sets:
-
+%Table should go to the results section
+All Men's semi finals and finals data
 \begin{center}
-  \begin{tabular}{|c c|} 
+  \begin{tabular}{|c  c  c  c  c  c  c |}
    \hline 
-   Data Set & Probability RxnTime < 0.1 \\ 
+   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
    \hline\hline
-   all finals gamma log & 0.000864 \\ 
+   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
    \hline
-   old finals gamma log & 0.000429 \\
+   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
    \hline
-   all pooled gamma log & 0.006847 \\
+   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
    \hline
-   old pooled gamma log & 0.005749 \\
+   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
+   \hline
+   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
+   \hline
+   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
+   \hline
   \end{tabular}
   \end{center}
-
+%Section 4.1: Gamma model results show the fitting with and without 2022,
+%talk about p values and thresholds from the simulations (do not include the
+%results in the methods section). 
+Using the gamma function with link set to log (whose summary statistics can be
+found in the last row of the above table), the probability of the function being
+less than 0.1 seconds can be calculated to see if Allen's disqualification was an anomaly.
 Prior to the 2022 World Championship, there was a 0.0425\% chance (based on the
@@
 
-%Section 4.1: Gamma model results show the fitting with and without 2022,
-%talk about p values and thresholds from the simulations (do not include the
-%results in the methods section).  Discussion should show that random effect
+%Discussion should show that random effect
 %and standard error increase and that 2022 is special.  Now talk about Seiko
 %timing equipment
+
+%Citation needed in this section
+Since 1985 Seiko Holding Corporations has served every World Athletics Championship
+as the official timer \citep{Seiko}.  Since 1985, the technology and the ability
+to accurately predict measurements: long jump distances, false starts, total time,
+reaction time, etc. have all dramatically improved.  Seiko did not start tracking
+reaction time as an official measurement until 1999 in the men's and women's 100 meter
+hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
+technology to the World Track and Field Championships, the highest stakes in the world
+of running outside of the Olympics. Seiko's technology for detecting reaction 
+times relies on the pressure that athletes
+exert on the plate when they push off.  Their systems measure the time differential
+between when the "gun" goes off to start the race and when the pressure changes 
+\citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
+false start as it is considered that no athlete can react so quickly 
+\citep{Seiko-Timing}.  Thus, a reaction time of for example 0.05, suggests that 
+the athlete predicted the gun and it was luck that caused their abnormally high 
+time.  At the World Championship level, World Athletics and Seiko want to remove 
+that element of luck and thus impose the 0.1 second barrier. In 2013, Seiko 
+upgraded their timing equipment for sprinting events (100 meter hurdles falls 
+under this category) \citep{WorldAthletics_2013}.  Since 1999, the highest reaction 
+times in the men's 110-meter hurdle were recorded in 2013.  That is not to say that the higher 
+times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
+to 2022, Seiko again upgraded its technology but for its jump management system.
 
@@
 
-The p values produced by the clustered rank sum tests were 0.000155 for the
-same athletes competing in the same year, and 0.001464 for the same athletes in
-different years.  To put it simply, there is not a reasonable explanation for 
-this.  
-It is inconceivable that so many athletes would improve such a significant
+For the comparison between athletes who competed in 2019 and 2022, the p value
+was 0.001464.  As Seiko timed both events, there should not be a signficant
+difference if the timing equipment in both instances was working correctly.
+Additionally, the p value was 0.000155 for the same athletes competing in the 
+same year  To put it simply, there is not a reasonable explanation for 
+this. It is inconceivable that so many athletes would improve such a significant
 amount in such a small amount of time. 6 countries are represented in this 
@@
 
+%Wrap up and Conclusion
 It is worth noting that in the semi-finals of the World Track and
@@
 needs to adjust that threshold and allow for faster reaction times so that athletes with superb 
-reflexes are not penalized like how Allen was in 2022.  World Athletics and Seiko 
-may want to do further testing of their own to determine what is a human's physically
-fastest reaction time, and if they choose to do so, they should start with Devon
-Allen.
-
-Since 1985 Seiko Holding Corporations has served every World Athletics Championship
-as the official timer \citep{Seiko}.  Since 1985, the technology and the ability
-to accurately predict measurements: long jump distances, false starts, total time,
-reaction time, etc. have all dramatically improved.  Seiko did not start tracking
-reaction time as an official measurement until 1999 in the men's and women's 100 meter
-hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
-technology to the World Track and Field Championships, the highest stakes in the world
-of running outside of the Olympics.
-
-%Citation needed in this section
-Seiko's technology for detecting reaction times relies on the pressure that athletes
-exert on the plate when they push off.  Their systems measure the time differential
-between when the "gun" goes off to start the race and when the pressure changes 
-\citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
-false start as it is considered that no athlete can react so quickly \citep{Seiko-Timing}.  
-Thus, a reaction time of for example 0.05, suggests that the athlete predicted the gun and it
-was luck that caused their abnormally high time.  At the World Championship level,
-World Athletics and Seiko want to remove that element of luck and thus impose the 0.1
-second barrier.
-
-In 2013, Seiko upgraded their timing equipment for sprinting events (100 meter hurdles
-falls under this catergory) \citep{WorldAthletics_2013}.  Since 1999, the highest reaction 
-times in the men's 100 meter hurdle were recorded in 2013.  That is not to say that the higher 
-times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
-to 2022, Seiko again upgraded its technology but for its jump management system.
+reflexes are not penalized like how Allen was in 2022.  World Athletics did not
+do anything about \citet{komi2009iaaf}'s study, but perhaps they should now so
+that the story of Devon Allen is not repeated.
 

### 2023-04-04 Manuscript/manuscript.tex
amsfonts needed for mathbb

@@
 
-\usepackage{amsmath}
+\usepackage{amsmath, amsfonts}
 \usepackage{booktabs}
@@
 
-2) and 3)
+2)
 Outside the blog posts from LetsRun.com, there have not been any published
@@
 
+3)
 In 2009, the International Association of Athletics Federations (Also referred to
@@
 This could provide a more accurate representation of reaction time, and thus the
-possibility of a false disqualification would be lowered \citep{Komi}.  
-  A
-follow-up study from 2012 by Pilianidis, Mantzouranis, and Kasabalis found
+possibility of a false disqualification would be lowered \citep{komi2009iaaf}.  
+A follow-up study from 2012 by Pilianidis, Mantzouranis, and Kasabalis found
 significant differences between reaction times at several world championships
@@
   present exploratory plots.}
+
+  \begin{equation}
+    Y_{ij} = \beta_0 + \beta_1 X_{ij} + \gamma_i + \epsilon_{ij}
+    \end{equation}
+
+    where $Y_{ij}$ is the response variable for observation $j$ in year $i$, 
+    $X_{ij}$ is the predictor variable for observation $j$ in year $i$, 
+    $\beta_0$ is the intercept, $\beta_1$ is the coefficient for $X_{ij}$, 
+    $\gamma_i$ is the random effect term for year $i$, and $\epsilon_{ij}$ 
+    is the error term for observation $j$ in year $i$.
 
@@
 $Z = \frac{S - E(S)}{\sqrt{\hat{\text{Var}}(S)}}$ where 
-$S = \mathbb{E}[W^* | X, \delta]$,
 $W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is

### 2023-04-04 Manuscript/manuscript.tex
fixed two left quotes

@@
 the Olympics and the World Track and Field Championships.  However, it is the
-news section of the website that generated articles such as "Was Devon Allen
-Screwed? There's At Least a 99\% That He Was" and "The Data Keeps Pouring In and
+news section of the website that generated articles such as ``Was Devon Allen
+Screwed? There's At Least a 99\% That He Was" and ``The Data Keeps Pouring In and
 It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert

### 2023-04-10 Manuscript/manuscript.tex
latex code
comments on figures
From last week meeting

@@
 
-\title{Was Devon Allen Unjustly Disqualified at the 2022 World Track and Field Championships?}
+\title{On Devon Allen's Disqualification at the 2022 Track and Field World
+Championship}
 
@@
 
-\section{Data} \label{sec:data}
+\section{Data} \label{sec:Data}
 
@@
 \end{figure}
+\jy{The plot was stretched. Don't. Get their sizes right before generating them.}
 
+\jy{Remove title from the figures. Make the figure captions self-contained.}
 \begin{figure}[h]
@@
   \begin{tabular}{|c c c c c c c|} 
-   \hline
+   \toprule
    Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
-   \hline\hline
+   \midrule
    Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
-   \hline
    Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
-   \hline
    Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
-   \hline
    Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
-   \hline
+   \bottomrule
   \end{tabular}
@@
 
-\section{Methods} \label{sec:meth}
+\section{Methods} \label{sec:Methods}
 

### 2023-04-10 Manuscript/manuscript.tex
Added few references, moved tables to after end document

@@
 
-1)
+%1
 After coming in third in the 110 meter hurdles at the United States Track and 
@@
 
-2)
+%2
 Outside the blog posts from LetsRun.com, there have not been any published
@@
 
-3)
+%3
 In 2009, the International Association of Athletics Federations (Also referred to
@@
 conclusion was that for reaction times did not decrease over the twelve years
-covered by the research in the study \citep{pilianidis2012start}.  This will
-be important when considering the results in section \ref{sec:Results}.
+covered by the research in the study \citep{pilianidis2012start}.  However, not
+all studies have been in favor of a decrease in the reaction time barrier.
+Brosnan in 2016 argues for increasing the reaction time barrier to 0.115 seconds,
+and cited evidence from European Championships and World Championships in his
+article \citep{brosnan2017effects}.  A study looking at the 2008 Olympics
+likewise supported the conclusion that athletes could not react in as little as
+0.1 seconds \citep{lipps2011implications}.
+%About timing
+It has been concluded that
 
-Despite the study, IAAF has not changed the reaction barrier since it was
-implemented at the 1999 World Track and Field Championships.  It is confusing
-as to why IAAF would commission a study only to disregard the results and not
-make the necessary changes.  It is unknown how many athletes since 2009 could
-have potentially benefited from this, but it is probable that there are others
-besides Allen who were disqualified.
 %Reaction time barrier has been examined before: cite reviews
@@
 
-4)
+%4
 One of the first objectives of the paper is to try to model the data, which is
@@
 responsible for the variation between years. 
-
 Another objective of this paper is to compare data for athletes who have competed
@@
 %Move a lot of this to methods section
-5)
+
+%5
 The rest of this paper is as follows: Section~\ref{sec:Data} describes how data was
@@
 111 corresponds to the 1999 finals). The data that looked at in the paper
-is best visualized using a box plot. Figure~\ref{fig:DataBoxPlot}.  Included are
+is best visualized using a box plot as show in Figure~\ref{fig:Boxplots}.  Included are
 box plots of the reaction times of the men's semi finals and finals pooled
-together and just the finals.
+together and just the finals.  It is clear that the 2022 boxplot for both the
+men's finals and semifinals data is lower (indicating a faster reaction time)
+than in previous years.  The median for the men's pooled data in 2022 was 0.129
+and the median for the men's final data in 2022 was 0.126.  For comparison,
+Brosnan found reaction times of 0.156 seconds for men when looking at data from
+1999 to 2014 \citep{brosnan2017effects}.
 %Fix quotations
 %Dont need the titles will add captions, use width =5, height = 3 in aspect ratio
+
 \begin{figure}[h]
   \centering
-  \includegraphics[width=\textwidth,height=4in]{DataBoxPlot}
-  \caption{Boxplot of semifinals and finals data}\label{fig:DataBoxPlot}
-\end{figure}
-\jy{The plot was stretched. Don't. Get their sizes right before generating them.}
-
-\jy{Remove title from the figures. Make the figure captions self-contained.}
-\begin{figure}[h]
-  \centering 
-  \includegraphics{BoxplotFinals}
-  \label{fig:BoxplotFinals}
+  \includegraphics{Finals_Pooled_Boxplot}
+  \label{fig:Boxplots}
+  \caption{The reaction times from 2019 to 2022 for the men's final (top) and
+  the men's semifinals and finals (bottom)}
 \end{figure}
@@
 
+\bibliographystyle{chicago}
+\bibliography{citations.bib}
+
+\end{document}
+
 Here are important statistics grouped by both model and data set:
@@
   \end{center}
-
-\bibliographystyle{chicago}
-\bibliography{citations.bib}
-
-\end{document}

### 2023-04-11 Manuscript/manuscript.tex
Overhauled entire document

@@
 spared waiting another two years until the next World Track and Field Championship.
+For track and field athletes, they usually only get the chance to prove 
+themselves to sponsors every two years, meaning that races such as the finals
+of the World Championship is an impactful way to earn a living. An 
+investigation into reaction times is also important because reaction times have
+been shown to affect the overall time of the race \citep{delalija2008reaction}.
+Thus, faster reaction times lead to faster overall times; which is the goal of
+every sprinter who steps onto the track.
 
@@
 seconds.  Two other athletes were able to react similarly when under certain
-muscular tightening conditions \citep{pain2007sprint}.
-to IAAF that the disqualification barrier be decreased to allow sprinters to
+muscular tightening conditions \citep{pain2007sprint}. Another study commissioned
+by IAAF concluded that the disqualification barrier be decreased to allow sprinters to
 react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
@@
 change the reaction time criteria from pushing off on the block to body movement.
-This could provide a more accurate representation of reaction time, and thus the
-possibility of a false disqualification would be lowered \citep{komi2009iaaf}.  
-A follow-up study from 2012 by Pilianidis, Mantzouranis, and Kasabalis found
+\citep{komi2009iaaf} This could provide a more accurate representation of 
+reaction time, and thus the possibility of a false disqualification would be lowered.
+The possibility of using a video-based medium has been proved to be a reliable
+measure of reaction time \citep{mudric2015evaluation}.
+A 2012 follow-up study to Komi's by Pilianidis found
 significant differences between reaction times at several world championships
@@
 conclusion was that for reaction times did not decrease over the twelve years
-covered by the research in the study \citep{pilianidis2012start}.  However, not
-all studies have been in favor of a decrease in the reaction time barrier.
-Brosnan in 2016 argues for increasing the reaction time barrier to 0.115 seconds,
-and cited evidence from European Championships and World Championships in his
-article \citep{brosnan2017effects}.  A study looking at the 2008 Olympics
-likewise supported the conclusion that athletes could not react in as little as
-0.1 seconds \citep{lipps2011implications}.
-%About timing
-It has been concluded that
+covered by the research in the study \citep{pilianidis2012start}.  This contradicts
+what was found in a 2021 study that concluded the year of competition has a
+significant effect on reaction times \citep{zhang2021correlation}.  Zhang's
+analysis looked at data from the 2011-2019 World Championships across multiple
+sprint events and concluded that from 2011 to 2019, athletes reaction times got
+faster. However, not  all studies have been in favor of a decrease in the
+reaction time barrier. Brosnan in 2016 argues for increasing the reaction time 
+barrier to 0.115 seconds, and cited evidence from European Championships and 
+World Championships in his article \citep{brosnan2017effects}.  A study looking 
+at the 2008 Olympics likewise supported the conclusion that athletes could not
+ react in as little as 0.1 seconds \citep{lipps2011implications}.
+It is worth noting that one study looking at data from various IAAF meets found
+that the fastest reaction times for men were found between the ages of 26 and 29
+\citep{tonnessen2013reaction}.  This is notable because Allen was 27 years old
+at the 2022 Track and Field World Championship and in his physical peak.
 
-%Reaction time barrier has been examined before: cite reviews
-%Then go into specifics and talk about people who wanted to change barrier
-%General reviews are important
-%People have studied the biological processes, cite a few references
 %How barrier is picked by regulator
-%Read abstract and into and summarize into about one sentence what each paper contributes
-%1 and 2 can be merged but 3 need to stand alone
 
@@
 the reaction times of every non-disqualification since the 1999 World Championships. 
-When determining the most effective model to use, several variations were
-explored across the 4 different data sets: a linear model without a venue effect,
-linear model with venue effect, a gamma model without a venue effect, a gamma model
-with a venue effect, a gamma model without a venue effect and link set to log instead
-of the default which is Gaussian.  The R package lme4 was very helpful with this
-as it supported the Generalized Linear Mixed Model (GLMM) which allowed 
-linear models with various distributions to be modeled\citep{Rpkg:lme4}.  The
-model we use is a liner model with a fixed effect and random effect and follows
-the gamma distribution. The fixed effect comes in the form of an intercept, and the
-random effect will be called the venue effect and is reflective of how significant
-the changes in the reaction time is for each venue, i.e. the venue effect is 
-responsible for the variation between years. 
+The model we use is a linear model, following a gamma distribution with a random
+effect, which will be referred to as the year effect.  The year effect is
+reflective of how significant the changes in the reaction time is for each venue,
+i.e. the venue effect is  responsible for the variation between years.  
 Another objective of this paper is to compare data for athletes who have competed
 at additional competitions to see how their reaction times compare from previous
-competitions to this one. First we can look at the reaction times of athletes who competed at the
-World Track and Field Championship in both 2019 and 2022 and compare the times,
-and secondly we can look at how athletes performed compared to how they did at
-their national/qualifiers meet.  We would expect that athletes of this caliber,
-who are world-class athletes, to not improve significantly from one year to the
-next.  In order to reduce any chances of athletes improving reaction time, we
-can also look at data from qualifying meets.  Almost every athlete had to qualify 
-by proving they are elite at the national level, usually by coming in at least
-fourth. Thus, data was compiled on athletes who competed
+competitions to this one. First we can look at the reaction times of athletes who
+competed at the World Track and Field Championship in both 2019 and 2022 and
+compare the times, and secondly we can look at how athletes performed compared
+to how they did at their national/qualifiers meet. In order to reduce any chances
+of athletes improving reaction time, we can also look at data from qualifying meets.
+Almost every athlete had to qualify by proving they are elite at the national level, 
+usually by coming in at least fourth. Thus, data was compiled on athletes who competed
 at the 2022 World Track and Field Championships on what these athletes also ran 
@@
 nearly null, as the national competitions took place between June and July 2022 
-and the World Championships were in August.  Thus, it should be expected that 
-elite athletes such as these will not get reasonably faster in such a short time
-span. The equipment used to measure reaction time cannot be guaranteed to be
-the same from national to international competition, but the time span between
-meets has been reduced considerably.
+and the World Championships were in August. The equipment used to measure reaction
+time cannot be guaranteed to be the same from national to international
+competition, but the time span between meets has been reduced considerably.
+
 %Simplify contribution to one paragraph: collect data and formal statistical
@@
 
-%Follow Anthony's work, explicit references as to what each section does
-
-%Seiko section should go in the Results section.  From results it seems reasonable
-%question the data and talk about Seiko
-
 \section{Data} \label{sec:Data}
@@
 
-  \begin{equation}
-    Y_{ij} = \beta_0 + \beta_1 X_{ij} + \gamma_i + \epsilon_{ij}
-  \end{equation}
-
-    where $Y_{ij}$ is the response variable for observation $j$ in year $i$, 
-    $X_{ij}$ is the predictor variable for observation $j$ in year $i$, 
-    $\beta_0$ is the intercept, $\beta_1$ is the coefficient for $X_{ij}$, 
-    $\gamma_i$ is the random effect term for year $i$, and $\epsilon_{ij}$ 
-    is the error term for observation $j$ in year $i$.
-
-  %yij denotes ith years and j is observation in year
-  %yij will have clustered i and we use random effect to account for this
-  %When model is presented talk about how I modeled yij
-  %Zij where player is i and cluster, group is year and j
-  %Test hypothesis that reaction times are the same vs that they are not
-Something that is extremely important to note in the Data is the inconsistency of World Athletics 
-in terms of how they penalize false starts.  From 2007-2009, the IAAF (the former name for World 
-Athletics, and the governing body for the World Track and Field Champions) instituted a rule change
-that allowed one warning false start before a sprinter was disqualified \citep{False-Start}  This 
-essentially gave sprinters a second chance and reduced the penalty for a false start.  There were 
-18 male and 7 female false starts at the 2007 World Championships, 18 male and 7 female starts at 
-the 2009 World Championships. Starting in 2011, the rule was scrapped, and the old policy which 
-automatically disqualifies runners who false start was reinstituted.  The effect? At the 2011 World
-Championship there were 6 male disqualifications and 4 female disqualifications \citep{False-Start}. 
-By returning to the harsher policy and cracking down on false starts, World Athletics had reduced the
-number of false starts by two thirds. This was desirable for World Athletics as false starts can 
+Something that is extremely important to note in the Data is the inconsistency 
+of World Athletics in terms of how they penalize false starts.  From 2007-2009,
+the IAAF (the former name for World Athletics, and the governing body for the 
+World Track and Field Champions) instituted a rule change that allowed one 
+warning false start before a sprinter was disqualified \citep{False-Start}  This 
+essentially gave sprinters a second chance and reduced the penalty for a false 
+start.  There were 18 male and 7 female false starts at the 2007 World 
+Championships, 18 male and 7 female starts at the 2009 World Championships. 
+Starting in 2011, the rule was scrapped, and the old policy which automatically 
+disqualifies runners who false start was reinstituted.  The effect? At the 2011 
+World Championship there were 6 male disqualifications and 4 female 
+disqualifications \citep{False-Start}. By returning to the harsher policy and 
+cracking down on false starts, World Athletics had reduced the number of 
+false starts by two thirds. One study that looked at different false start
+rules that IAAF has imposed from 1997 to 2009 and found that the reaction 
+times when there was less leniency improved by a statistically significant amount
+\citep{haugen2013effect}. This was desirable for World Athletics as false starts can 
 make already long track meets more tedious, both for athletes and viewers watching the 
 television broadcast.  Since 2011, IAAF has instituted the one false start rule
-\citep{Pilianidis}.
+\citep{Pilianidis}.  However, we chose not to include women's data in this
+paper due to the numerous papers citing difference in reaction times in men and
+women's reaction times: \citet{lipps2011implications}, \citet{babicc2009reaction}
+\citet{panoutsakopoulos2020gender}.
 
@@
 time, and the predicting variables are year, stage of competition (heats, 
-semifinal, final), total time, and gender.  The data can be found in the appendix:.  
+semifinal, final), total time, and gender.  The data can be found in the
+appendix~\ref{sec:Appendix} 
 Factors such as name, country, and identification number were
@@
 correlated with other runners reacting quickly (one runner reacting quickly may
-result in other runners reacting quickly).  There is no statistical signficance to
+result in other runners reacting quickly).  There is no statistical significance to
 how the batches were labeled, they were ordered chronologically by meet but
 reverse chronologically by year (Batch 1 is a preliminary Heat of 2022 and Batch
-111 corresponds to the 1999 finals). The data that looked at in the paper
-is best visualized using a box plot as show in Figure~\ref{fig:Boxplots}.  Included are
-box plots of the reaction times of the men's semi finals and finals pooled
-together and just the finals.  It is clear that the 2022 boxplot for both the
+111 corresponds to the 1999 finals). Included are
+box plots of the reaction times of the men's finals and the men's semifinals and
+finals pooled together.  We will refer to the latter as the "pooled data"
+throughout the paper.  The heats or preliminary rounds were not studied because
+athletes react faster as the competition progresses.  Athletes react faster in
+the finals than the semifinals which are faster than the heats 
+\ref{zhang2021correlation}.  If we want to determine if Allen's reaction time
+is reasonable, then we thus need to look at the men's final data, with the men's
+pooled data being included in order to increase the sample size.  In 2022, there
+were only five data points due to two disqualifications and one athlete who 
+qualified but did not compete.
+The data that looked at in the paper is best visualized using a box plot as show 
+in Figure~\ref{fig:Boxplots}.  It is clear that the 2022 box plot for both the
 men's finals and semifinals data is lower (indicating a faster reaction time)
@@
 
-\begin{figure}[h]
+\begin{figure}[tbp]
   \centering
@@
 competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
-Grant Holloway, and Devon Allen. Every athlete reacted faster in all of the World 
+Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
 Track and Field Championship races compared to the USA Track and Field 
@@
 Field Championship compared to the USA Track and Field Championship as were his
-teamates.  This comparison suggests that the timing system used between the two
+teammates.  This comparison suggests that the timing system used between the two
 meets may have been meaningfully different to the point where Devon Allen was
@@
 
-%Throughout exploratory analysis, gamma provides better AIC
-%Spend paragraph on after fitting the model how to figure out 0.1 p value
-
 Once the Gamma mixed effects model was chosen, it needed to be decided which combination
-of effects to use.  It was possible to include a venue effect, a batch effect, or a year
-and batch effect.  Density plots were created that simulated the probability of a mean
-density less than 0.1, which is the current threshold for a reaction time
-disqualification.  The density plots are shown below.  The batch effect is very
+of effects to use.  It was possible to include a venue effect, a batch effect, 
+or a year and batch effect.  The batch effect is very
 difficult to explain or interpret as it means that some heats of races were either
@@
 effect, but are all factors that could potentially impact the venue effect.
+Thus, we decided to inclue only the year effect in the GLMM, which is presented
+as:
+
+\begin{equation}
+  \label{eq:GLMM}
+  Y_{ij} = \beta_0 + \gamma_i + \epsilon_{ij}
+\end{equation}
+
+Where $Y_{ij}$ is the response variable for observation $j$ in year $i$, , 
+$\beta_0$ is the intercept, $\gamma_i$ is the random effect term for year $i$, 
+and $\epsilon_{ij}$ is the error term for observation $j$ in year $i$.
+
+%yij denotes ith years and j is observation in year
+%yij will have clustered i and we use random effect to account for this
+%When model is presented talk about how I modeled yij
+%Zij where player is i and cluster, group is year and j
+%Test hypothesis that reaction times are the same vs that they are not
+
+%Throughout exploratory analysis, gamma provides better AIC
+%Spend paragraph on after fitting the model how to figure out 0.1 p value
+As the GLMM model has mixed form, the best way to try to gauge the probability
+of observing a reaction time is to simulate the distribution of times using a 
+large sample.  In order to approximate the distribution, R code was written to 
+try and estimate the probabilities of observing a reaction time as extreme as
+0.1 seconds.  The code re-paramertizes the gamma distribution to be in terms of
+$\mu$ and shape rather than by $\alpha$ and $\beta$.  From the code, it is
+possible to calculate the probability of observing a reaction time faster than
+a given time.  We want the probability of a reaction time being less than 0.1
+seconds in order to gauge if that is a reasonable disqualification barrier.
 
@@
 %Figure out the p value of reaction time less than 0.1
-To try and approximate the distribution of the reaction times, R code was written
-to try and estimate the probabilities of observing a reaction time as extreme as
-0.1 seconds.  The code re-paramertizes the gamma distribution to be in terms of
-$\mu$ and shape rather than by $\alpha$ and $\beta$.
-
-At a size of n=12,000,000, the calculated probability of a reaction
-time being below 0.1 seconds was as follows for the various different data sets:
-
-\begin{center}
-  \begin{tabular}{|c c|} 
-   \hline
-   Data Set & Probability RxnTime < 0.1 \\ 
-   \hline\hline
-   all finals gamma log & 0.000864 \\ 
-   \hline
-   old finals gamma log & 0.000429 \\
-   \hline
-   all pooled gamma log & 0.006847 \\
-   \hline
-   old pooled gamma log & 0.005749 \\
-  \end{tabular}
-  \end{center}
 
@@
 
-\section{Results}
+\section{Results} \label{sec:Results}
 \jy{summarize your findings; discuss limitations and possible future works.}
 
-%Table should go to the results section
-All Men's semi finals and finals data
-\begin{center}
-  \begin{tabular}{|c  c  c  c  c  c  c |}
-   \hline 
-   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
-   \hline\hline
-   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
-   \hline
-   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
-   \hline
-   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
-   \hline
-   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
-   \hline
-   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
-   \hline
-   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
-   \hline
-  \end{tabular}
-  \end{center}
+\subsection{GLMM Results}
 %Section 4.1: Gamma model results show the fitting with and without 2022,
@@
 %results in the methods section). 
+At a size of n=10,000,000, the calculated probability of a reaction
+time being below 0.1 seconds was as follows for the various different data sets:
+
+\begin{center}
+  \begin{tabular}{|c c|} 
+   \hline
+   Data Set & Probability RxnTime < 0.1 \\ 
+   \hline\hline
+   all finals gamma log & 0.000856 \\ 
+   \hline
+   old finals gamma log & 0.000439 \\
+   \hline
+   all pooled gamma log & 0.006865 \\
+   \hline
+   old pooled gamma log & 0.005761 \\
+  \end{tabular}
+  \end{center}
+
 Using the gamma function with link set to log (whose summary statistics can be
@@
 less than 0.1 seconds can be calculated to see if Allen's disqualification was an anomaly.
-Prior to the 2022 World Championship, there was a 0.0425\% chance (based on the
+Prior to the 2022 World Championship, there was a 0.0439\% chance (based on the
 simulated results) of observing a reaction time below 0.01 seconds in the men's
 finals.  That means that we would expect a time below the threshold to occur
-once every 2,352 starts.  Considering there are roughly 8 men's finals reaction
-times every 2 years (The World Championship is biennial), every 588 years there
+once every 2,277 starts.  Considering there are roughly 8 men's finals reaction
+times every 2 years (The World Championship is biennial), every 569 years there
 should be a time as extreme as Allen's.  This seems extremely unlikely, and thus
@@
 would break the barrier.  For the dataset of all finals data, the probability
-associated with observing a reaction time below 0.08 seconds was $6.67\cdot10{-7}$.
+associated with observing a reaction time below 0.08 seconds was $6\cdot10{-7}$.
 If IAAF were to change the reaction time the way that \citet{Komi}'s paper
@@
 
+\subsection{Rank Comparison Results}
 %Section 4.2 is motivated by seiko timing and look at same athletes across
@@
 
-%Wrap up and Conclusion
+\section{Discussion}\label{sec:Discussion}
+
+One conclusion from \citet{zhang2021correlation}'s analysis was that athletes
+reaction times improved from 2011 to 2019.  However this study does not consider 
+data from 1999 to 2009, and over this timespan athletes get slower and then
+faster.  A boxplot~\ref{fig:Boxplot} shows that for the men's pooled data, the 
+year with the fastest median reaction time before 2022 was 1999.  Thus, this 
+reject's Zhang's conclusion that athletes are getting faster because they were 
+fast over twenty years ago.  Instead, a more appropriate conclusion may be that
+there is a lot of variability in reaction times as evidenced by the two fastest
+years occurring 23 years apart.
+
 It is worth noting that in the semi-finals of the World Track and
@@
 \section{Appendix}
-\label{sec:appendix}
+\label{sec:Appendix}
 %Put downloadable data spreadsheet
@@
 Here is the data for the USA Track and Field Championships: \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
+
+For the pooled data set
+\begin{center}
+  \begin{tabular}{|c  c  c  c  c  c  c |}
+   \hline 
+   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+   \hline\hline
+   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
+   \hline
+   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
+   \hline
+   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
+   \hline
+   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
+   \hline
+   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
+   \hline
+   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
+   \hline
+  \end{tabular}
+  \end{center}
 

### 2023-04-11 Code/Models.Rmd
Changed n to be 10 million

@@
   x <- rgamma(n, shape = shape, scale = mu / shape)
-  x
 }
@@
 simulation <- function(model, base){
-  simfit <- function(model, n = 12000000) {
+    simfit <- function(model, n = 10000000) {
     vc <- as.data.frame(VarCorr(model))
@@
 }
-simulation(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log)
+simulation(old_pooled_year_gamma_log, old_pooled_one_gamma_log)
 ```

### 2023-04-11 Code/Plots.Rmd
ggplot boxplot

@@
 sfit0 <- summary(fit0)
-simfit <- function(fit2, n = 12000000) {
+simfit <- function(fit2, n = 10000000) {
   vc <- as.data.frame(VarCorr(fit2))
@@
 ```
+
 ```{r}
@@
 ```{r}
+opar <- par()
+par(mfrow = c(1,2))
 pdf(file = "BoxplotFinals.pdf")
-boxplot(RxnTime ~ Year, data = men_finals, xlab = "Year", ylab = "Reaction Time in seconds", main = "Men's Final Reaction Time Boxplot Data")
+boxplot(RxnTime ~ Year, data = men_finals, xlab = "Year", ylab = "Reaction Time in seconds")
+boxplot(RxnTime ~ Year, data = men_finals, xlab = "Year", ylab = "Reaction Time in seconds")
 dev.off()
+par(opar)
 ```
+
+```{r}
+#ggplot(data = win, aes(x = strength_diff)) + geom_boxplot() + theme_minimal()
+ggplot(data = men_finals, aes(x=Year, y = RxnTime)) + geom_boxplot() + theme_minimal()
+```
+
+```{r}
+library(gridExtra)
+library(ggplot2)
+final <- ggplot(men_finals, aes(x = Year, y = RxnTime, group = Year)) +
+  geom_boxplot(fill = "lightblue", color = "darkblue") +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.1, 0.225),
+                            breaks = seq(0.1, 0.225, 0.025)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2023),
+                                              breaks = seq(1997, 2022, 2))
+
+pooled <- ggplot(men_pooled, aes(x = Year, y = RxnTime, group = Year)) +
+  geom_boxplot(fill = "orange1", color = "orangered4") +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.1, 0.225),
+                            breaks = seq(0.1, 0.225, 0.025)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2023),
+                                              breaks = seq(1997, 2022, 2))
+both <- grid.arrange(final,pooled, ncol=1)
+ggsave(filename="Finals_Pooled_Boxplot.pdf", path = "../Manuscript", plot=both, height = 6, width = 6)
+```

### 2023-04-11 Code/Models.Rmd
Minor change

@@
   x <- simfit(model)
-  print(mean(x < 0.1))
+  print(mean(x < 0.08))
   print(quantile(x, prob = c(.00001, .0001, .0005, .001)))

### 2023-04-11 Manuscript/manuscript.tex
Additional citation

@@
 at the 2008 Olympics likewise supported the conclusion that athletes could not
- react in as little as 0.1 seconds \citep{lipps2011implications}.
-It is worth noting that one study looking at data from various IAAF meets found
-that the fastest reaction times for men were found between the ages of 26 and 29
-\citep{tonnessen2013reaction}.  This is notable because Allen was 27 years old
-at the 2022 Track and Field World Championship and in his physical peak.
+react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
+concluded that visual reaction times to the starting gun may be faster by 0.007
+seconds than the IAAF method, which is from force on block starts 
+\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
+only by 0.01 seconds. It is also worth noting that one study looking at data 
+from various IAAF meets found that the fastest reaction times for men were found
+between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
+because Allen was 27 years old at the 2022 Track and Field World Championship 
+and in his physical peak.
 

### 2023-04-11 Manuscript/manuscript.tex
added comments about womens data

@@
 section provides the probabilities from the methods section and the paper ends 
-by questioning the reaction time barrier.
+by questioning the reaction time barrier.  Lastly, the section~\ref{sec:Discussion}
+discusses the impact and limitations of the paper.
 
@@
 reject's Zhang's conclusion that athletes are getting faster because they were 
-fast over twenty years ago.  Instead, a more appropriate conclusion may be that
-there is a lot of variability in reaction times as evidenced by the two fastest
-years occurring 23 years apart.
+fast over twenty years ago. One possible explanation for why 2011 had such
+high variation was because of the IAAF rule discussed earlier that increased the
+penalty of a false start to an instant disqualification \citep{False-Start}.
+Athletes may have been overly cautious to false start and may have consciously
+or sub-consciously reacted slower as a result.  Instead, a more appropriate 
+conclusion may be that there is a lot of variability in reaction times and that
+the context behind the races matters.
 
@@
 that the story of Devon Allen is not repeated.
+
+This paper does not consider data from the women's 100 meter hurdle, and although
+that was initially because of concerns over reaction time differences in men and
+women, exploratory analysis was performed and showed that 2022 was at best only
+slightly faster relative to other championships going back to 2022.  Likewise,
+the rank based comparisons results in p values that were not significant.  The
+lack of evidence from looking at the women's data makes it harder to conclude 
+that the sensors used at the 2022 championships were faulty, although it is
+unknown if the sensors for the men's 110 meter hurdle and women's 100 meter
+hurdle were the same.
 

### 2023-04-11 Code/Women.Rmd
added extra boxplot

@@
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_nations, paired = TRUE, alternative = "greater", method = "ds")
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_nations, paired = TRUE, alternative = "greater", method = "rgl")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_nations, alternative = "greater", method = "ds")
+#clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_nations, alternative = "greater", method = "rgl")
+#clusWilcox.test(Time ~ Year + cluster(Athlete), data = women_nations, alternative = "greater", method = "ds")
 ```

### 2023-04-12 Manuscript/manuscript.tex
Worked on stuff from 4/11 meeting notes

@@
 \noindent\sc{Keywords}:
-Devon Allen, Reaction Time, Seiko, 2022 World Track and Field Championships
+Reaction Time, Seiko,
 
@@
 with shock from both Allen and fans alike.  NBC, who was broadcasting the
-championships in the United States, later uploaded a video to their Youtube
+championships in the United States, later uploaded a video to their YouTube
 channel that show the disqualification and can be found here:
@@
 Section~\ref{sec:Methods}.  The signed-rank test analysis for the same athletes is
-also included in the Section~\ref{sec:Methods}.  The Section~\ref{sec:Results}
+also included in Section~\ref{sec:Methods}.  Section~\ref{sec:Results}
 section provides the probabilities from the methods section and the paper ends 
-by questioning the reaction time barrier.  Lastly, the section~\ref{sec:Discussion}
+by questioning the reaction time barrier.  Lastly, section~\ref{sec:Discussion}
 discusses the impact and limitations of the paper.
@@
 World Track and Field Champions) instituted a rule change that allowed one 
-warning false start before a sprinter was disqualified \citep{False-Start}  This 
+warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}  This 
 essentially gave sprinters a second chance and reduced the penalty for a false 
@@
 World Championship there were 6 male disqualifications and 4 female 
-disqualifications \citep{False-Start}. By returning to the harsher policy and 
+disqualifications \citep{iaaf2009falsestart}. By returning to the harsher policy and 
 cracking down on false starts, World Athletics had reduced the number of 
@@
 television broadcast.  Since 2011, IAAF has instituted the one false start rule
-\citep{Pilianidis}.  However, we chose not to include women's data in this
+\citep{iaaf2009falsestart}.  However, we chose not to include women's data in this
 paper due to the numerous papers citing difference in reaction times in men and
-women's reaction times: \citet{lipps2011implications}, \citet{babicc2009reaction}
-\citet{panoutsakopoulos2020gender}.
+women's reaction times: \citep{lipps2011implications, babicc2009reaction,
+panoutsakopoulos2020gender}
 
-Data was copied from the World Athletics website and pasted into an Excel
-spreadsheet. The data covers the men's 110 meter hurdles and the women's 100
-meter hurdles from 1999 to 2022.  The variable of prediction interest is reaction
-time, and the predicting variables are year, stage of competition (heats, 
-semifinal, final), total time, and gender.  The data can be found in the
-appendix~\ref{sec:Appendix} 
+\subsection{GLMM}
+Data was taken from the World Athletics and covers the men's 110 meter hurdles 
+and the women's 100 meter hurdles from 1999 to 2022.  The variable of prediction 
+interest is reaction time, and the predicting variables are year, stage of 
+competition (heats, semifinal, final), total time, and gender.  
+The data can be found in the appendix~\ref{sec:Appendix} 
 Factors such as name, country, and identification number were
@@
 the finals than the semifinals which are faster than the heats 
-\ref{zhang2021correlation}.  If we want to determine if Allen's reaction time
+\citep{zhang2021correlation}.  If we want to determine if Allen's reaction time
 is reasonable, then we thus need to look at the men's final data, with the men's
@@
   \caption{The reaction times from 2019 to 2022 for the men's final (top) and
-  the men's semifinals and finals (bottom)}
+  the men's semifinals and finals (bottom).}
 \end{figure}
 
+\subsection{Rank based Comparisons}
 This section was motivated from a qualitative analysis that started by looking
@@
 
-When the clusrank rank sum test was performed, the "ds" method produced a z score of
-2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
-the mean reaction time on average in 2019 was larger than the mean reaction time
-in 2022 for athletes who competed in both championships at an alpha level of 0.01.
-This provides good evidence that athletes who competed in both 2019 and 2022 got
-faster during that timespan.
-
-Repeating the data collection steps and
-again performing a clustered analysis for a Wilcoxon ranked sum test
-results in two new z scores and p values.  The "ds" method produced a z score of
-3.6069 and p value of 0.000155.  This result is highly significant and shows that
-the times at the national stage of competition were significantly higher than
-the reaction times at the World Championships even for an alpha level of 0.001.
-
 \section{Results} \label{sec:Results}
@@
 would break the barrier.  For the dataset of all finals data, the probability
-associated with observing a reaction time below 0.08 seconds was $6\cdot10{-7}$.
-If IAAF were to change the reaction time the way that \citet{Komi}'s paper
+associated with observing a reaction time below 0.08 seconds was $6\cdot10^{-7}$.
+If IAAF were to change the reaction time the way that \citet{komi2009iaaf}'s paper
 suggests, the probability would be so much lower that athletes would likely not
@@
 
-For the comparison between athletes who competed in 2019 and 2022, the p value
-was 0.001464.  As Seiko timed both events, there should not be a signficant
+When the clusrank rank sum test was performed, the "ds" method produced a z score of
+2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
+the mean reaction time on average in 2019 was larger than the mean reaction time
+in 2022 for athletes who competed in both championships at an alpha level of 0.01.
+This provides good evidence that athletes who competed in both 2019 and 2022 got
+faster during that timespan.
+
+Repeating the data collection steps and
+again performing a clustered analysis for a Wilcoxon ranked sum test
+results in two new z scores and p values.  The "ds" method produced a z score of
+3.6069 and p value of 0.000155.  This result is highly significant and shows that
+the times at the national stage of competition were significantly higher than
+the reaction times at the World Championships even for an alpha level of 0.001.
+
+As Seiko timed both events, there should not be a signficant
 difference if the timing equipment in both instances was working correctly.
-Additionally, the p value was 0.000155 for the same athletes competing in the 
-same year  To put it simply, there is not a reasonable explanation for 
+  To put it simply, there is not a reasonable explanation for 
 this. It is inconceivable that so many athletes would improve such a significant
@@
 data from 1999 to 2009, and over this timespan athletes get slower and then
-faster.  A boxplot~\ref{fig:Boxplot} shows that for the men's pooled data, the 
+faster.  A boxplot~\ref{fig:Boxplots} shows that for the men's pooled data, the 
 year with the fastest median reaction time before 2022 was 1999.  Thus, this 
@@
 high variation was because of the IAAF rule discussed earlier that increased the
-penalty of a false start to an instant disqualification \citep{False-Start}.
+penalty of a false start to an instant disqualification \citep{iaaf2009falsestart}.
 Athletes may have been overly cautious to false start and may have consciously

### 2023-04-14 Manuscript/manuscript.tex
comments/suggestons

@@
 
-  \jy{Avoid references in abstract.}
-
-  \jy{This is only the background. An abstract also needs to summarize the contributions.}
-%\bigskip
-\noindent\sc{Keywords}:
+\bigskip
+\noindent{\sc Keywords}:
 Reaction Time, Seiko,
@@
 
-\jy{Not a good opening. Could start by telling the story, recapping the major
-  events and discussions on the internet. Then dig into the relevant
-  literature. Finally outline the contributions of this work.
-}
-
 \jy{use the following structure and flow:
@@
 
+\jy{could condense a bit}
 %1
@@
 %2
-Outside the blog posts from LetsRun.com, there have not been any published
+Outside the blog posts from \url{www.LetsRun.com}, there have not been any published
 papers regarding Devon Allen's disqualification in particular.  This is not
@@
 
+\jy{break into two paragraphs}
 %3
@@
 as IAAF or World Athletics), which is the governing board for international running
-competitions, commissioned a study into changing the reaction time barrier.
+competitions, commissioned a study into changing the reaction time barrier \citep{reference?}.
 The paper analyzed sprinters reaction times and the biological processes 
 that went into reacting to the starting gun and concluded 
-that athletes can react in as low as 0.08 seconds.  One study by Pain looked at
+that athletes can react in as low as 0.08 seconds.   \citet{pain2007sprint} looked at
 reaction times for nine male athletes and analyzed their reaction times.  The
@@
 seconds.  Two other athletes were able to react similarly when under certain
-muscular tightening conditions \citep{pain2007sprint}. Another study commissioned
+muscular tightening conditions. Another study commissioned
+\jy{Is this the same study at the beginning of this paragraph?}
 by IAAF concluded that the disqualification barrier be decreased to allow sprinters to
@@
 change the reaction time criteria from pushing off on the block to body movement.
-\citep{komi2009iaaf} This could provide a more accurate representation of 
+\citep{komi2009iaaf}. This could provide a more accurate representation of 
 reaction time, and thus the possibility of a false disqualification would be lowered.
@@
 significant effect on reaction times \citep{zhang2021correlation}.  Zhang's
+\jy{Not Zhang's. There are coauthors. Consider using citet when appropriate in this paragraph.}
 analysis looked at data from the 2011-2019 World Championships across multiple
@@
 %4
+\jy{I think you have one objective but two methods to approach it.}
 One of the first objectives of the paper is to try to model the data, which is
@@
 
-\subsection{GLMM}
+\subsection{World Championship}
 Data was taken from the World Athletics and covers the men's 110 meter hurdles 
@@
 not considered important for the purposes of this research besides the times for
-Devon Allen and the other United States athletes mentioned above.  The variable "Stage"
-refers to whether the observation occurred during the "heats" (Preliminary round) which
-is denoted by "H" in the data, "S" for "semifinals" (Of which there are either 2 or 3 
-heats each year), and "F" for finals. The variable "Gender" has levels "M" and "F" 
-which stand for "male" and "female" respectively.  The variable "Batch" takes a 
+Devon Allen and the other United States athletes mentioned above.  The variable ``Stage"
+refers to whether the observation occurred during the ``heats" (Preliminary round) which
+is denoted by ``H" in the data, ``S" for ``semifinals" (Of which there are either 2 or 3 
+heats each year), and ``F" for finals. The variable ``Gender" has levels ``M" and ``F" 
+which stand for ``male" and ``female" respectively.  The variable ``Batch" takes a 
 numerical value and every heat of every round from 1999 to 2022 was assigned it's
@@
 box plots of the reaction times of the men's finals and the men's semifinals and
-finals pooled together.  We will refer to the latter as the "pooled data"
+finals pooled together.  We will refer to the latter as the ``pooled data"
 throughout the paper.  The heats or preliminary rounds were not studied because
@@
 
-\subsection{Rank based Comparisons}
+\subsection{Beyond World Championship}
 This section was motivated from a qualitative analysis that started by looking
@@
 disqualified more because of faulty equipment rather than because he reacted
+\jy{cross-reference to the table by its label}
 too quickly.  This table below highlights the difference in reaction times for
-Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. "USA" denotes
-the USA Track and Field Championships, "World" denotes the World Track and Field
-Championships, "N/A" was used for any athlete that did not compete and thus
+Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
+the USA Track and Field Championships, ``World" denotes the World Track and Field
+Championships, ``N/A" was used for any athlete that did not compete and thus
 did not register a reaction time.  All numbers listed in the table are the reaction
@@
 \begin{center}
-  \begin{tabular}{|c c c c c c c|} 
+  \begin{tabular}{c c c c c c c} 
    \toprule
@@
 
-\subsection{Rank-based comparison for Clustered}
+\subsection{Rank-based Comparison}
 
@@
 
+\jy{Undefined notations}
 The test statistic of the DS method of the Wilcoxon rank-sum statistic is:
@@
 The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
-\citet{datta2005rank} and the R package "clusrank" allows for its implementation.
+\citet{datta2005rank} and the R package ``clusrank" allows for its implementation.
 \citep{jiang2017wilcoxon}.
 
-Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
-test was performed with the cluster set to be the Athlete and group set to Year.
-The method was set to be "DS" which is used for the Data and Satten method, which
-is preferable when the cluster sizes are informative \citep{jiang2017wilcoxon}.  We do not care that some
+Using the \texttt{clusWilcox.test()} function, a Wilcoxon rank sum
+test was performed with the cluster set to be the Athlete and group set to Year
+using the DS method. The DS method is preferable when the cluster sizes are
+informative \citep{jiang2017wilcoxon}. \jy{explain. More data means faster
+  reaction time?}
+We do not care that some
 athletes made it further in either competition than another or that it should
@@
 \section{Results} \label{sec:Results}
-\jy{summarize your findings; discuss limitations and possible future works.}
 
-\subsection{GLMM Results}
+\subsection{GLMM}
 %Section 4.1: Gamma model results show the fitting with and without 2022,
@@
 %results in the methods section).
-At a size of n=10,000,000, the calculated probability of a reaction
+
+\jy{present the estimated parameters first before deriving the p-value using the
+  estimated parameters. The difference in variance of the random effects using
+  different data motivated the rank-based study.}
+
+At a size of $n=10,000,000$, the calculated probability of a reaction
 time being below 0.1 seconds was as follows for the various different data sets:
-
+\jy{label the table and reference to it; clean it up}
 \begin{center}
@@
 exert on the plate when they push off.  Their systems measure the time differential
-between when the "gun" goes off to start the race and when the pressure changes 
+between when the ``gun" goes off to start the race and when the pressure changes 
 \citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
@@
 
-\subsection{Rank Comparison Results}
+\subsection{Rank-based Comparison}
 %Section 4.2 is motivated by seiko timing and look at same athletes across
@@
 
-When the clusrank rank sum test was performed, the "ds" method produced a z score of
+When the clusrank rank sum test was performed, the ``ds" method produced a z score of
 2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
@@
 again performing a clustered analysis for a Wilcoxon ranked sum test
-results in two new z scores and p values.  The "ds" method produced a z score of
+results in two new z scores and p values.  The ``ds" method produced a z score of
 3.6069 and p value of 0.000155.  This result is highly significant and shows that

### 2023-04-18 Manuscript/manuscript.tex
Went over most suggestions

@@
 considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
-in early June; only 0.04 seconds away from the world record \citep{Preview}.
+in early June; only 0.04 seconds away from the world record \citep{wa2022preview}.
 Thus his disqualification by reacting in 0.099 seconds instead of 0.01 was met
@@
 spared waiting another two years until the next World Track and Field Championship.
-For track and field athletes, they usually only get the chance to prove 
-themselves to sponsors every two years, meaning that races such as the finals
-of the World Championship is an impactful way to earn a living. An 
-investigation into reaction times is also important because reaction times have
+An investigation into reaction times is also important because reaction times have
 been shown to affect the overall time of the race \citep{delalija2008reaction}.
@@
 cited graphs, statistics such as median reaction time, and compared the reaction
-times at the 2022 Track and Field World Championship.
+times at the 2022 Track and Field World Championship \citep{johnson2022devon}.
 
@@
 %3
-In 2009, the International Association of Athletics Federations (Also referred to
-as IAAF or World Athletics), which is the governing board for international running
-competitions, commissioned a study into changing the reaction time barrier \citep{reference?}.
-The paper analyzed sprinters reaction times and the biological processes 
-that went into reacting to the starting gun and concluded 
-that athletes can react in as low as 0.08 seconds.   \citet{pain2007sprint} looked at
+Other studies have looked at minimum reaction times such as \citet{pain2007sprint}  which looked at
 reaction times for nine male athletes and analyzed their reaction times.  The
@@
 seconds.  Two other athletes were able to react similarly when under certain
-muscular tightening conditions. Another study commissioned
-\jy{Is this the same study at the beginning of this paragraph?}
-by IAAF concluded that the disqualification barrier be decreased to allow sprinters to
+muscular tightening conditions.  In 2009, IAAF commissioned a study to examine
+the disqualification barrier. It was concluded that the disqualification 
+barrier be decreased to allow sprinters to
 react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
@@
 measure of reaction time \citep{mudric2015evaluation}.
-A 2012 follow-up study to Komi's by Pilianidis found
+
+\jy{Not Zhang's. There are coauthors. Consider using citet when appropriate in this paragraph.}
+
+A 2012 follow-up study to Komi's by \citet*{pilianidis2012start} found
 significant differences between reaction times at several world championships
@@
 what was found in a 2021 study that concluded the year of competition has a
-significant effect on reaction times \citep{zhang2021correlation}.  Zhang's
-\jy{Not Zhang's. There are coauthors. Consider using citet when appropriate in this paragraph.}
+significant effect on reaction times \citep*{zhang2021correlation}.  Their
 analysis looked at data from the 2011-2019 World Championships across multiple
-sprint events and concluded that from 2011 to 2019, athletes reaction times got
+sprint events and concluded that from 2011 to 2019, reaction times got
 faster. However, not  all studies have been in favor of a decrease in the
@@
   present exploratory plots.}
-
-Something that is extremely important to note in the Data is the inconsistency 
-of World Athletics in terms of how they penalize false starts.  From 2007-2009,
-the IAAF (the former name for World Athletics, and the governing body for the 
-World Track and Field Champions) instituted a rule change that allowed one 
-warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}  This 
-essentially gave sprinters a second chance and reduced the penalty for a false 
-start.  There were 18 male and 7 female false starts at the 2007 World 
-Championships, 18 male and 7 female starts at the 2009 World Championships. 
-Starting in 2011, the rule was scrapped, and the old policy which automatically 
-disqualifies runners who false start was reinstituted.  The effect? At the 2011 
-World Championship there were 6 male disqualifications and 4 female 
-disqualifications \citep{iaaf2009falsestart}. By returning to the harsher policy and 
-cracking down on false starts, World Athletics had reduced the number of 
-false starts by two thirds. One study that looked at different false start
-rules that IAAF has imposed from 1997 to 2009 and found that the reaction 
-times when there was less leniency improved by a statistically significant amount
-\citep{haugen2013effect}. This was desirable for World Athletics as false starts can 
-make already long track meets more tedious, both for athletes and viewers watching the 
-television broadcast.  Since 2011, IAAF has instituted the one false start rule
-\citep{iaaf2009falsestart}.  However, we chose not to include women's data in this
-paper due to the numerous papers citing difference in reaction times in men and
-women's reaction times: \citep{lipps2011implications, babicc2009reaction,
-panoutsakopoulos2020gender}
+There are two types of data that is explored throughout this paper: data from
+1999-2022 for every world championship, and data from athletes who competed
+at both the 2019 and 2022 World Championships and 2022 national competitions
+and the 2022 World Championship.  The first data set is used for the GLMM portion
+of the paper, while the second data set is used for the rank based comparisons.
 
@@
 competition (heats, semifinal, final), total time, and gender.  
-The data can be found in the appendix~\ref{sec:Appendix} 
-Factors such as name, country, and identification number were
-not considered important for the purposes of this research besides the times for
-Devon Allen and the other United States athletes mentioned above.  The variable ``Stage"
+The data can be found in the appendix~\ref{sec:Appendix}. Factors such as name, 
+country, and identification number were not considered important for the 
+purposes of this research besides the times for Devon Allen and the other 
+United States athletes mentioned above.  The variable ``Stage"
 refers to whether the observation occurred during the ``heats" (Preliminary round) which
@@
 1999 to 2014 \citep{brosnan2017effects}.
+
+Something that is extremely important to note in the Data is the inconsistency 
+of World Athletics in terms of how they penalize false starts.  From 2007-2009,
+the IAAF (the former name for World Athletics, and the governing body for the 
+World Track and Field Champions) instituted a rule change that allowed one 
+warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}  This 
+essentially gave sprinters a second chance and reduced the penalty for a false 
+start.  There were 18 male and 7 female false starts at the 2007 World 
+Championships, 18 male and 7 female starts at the 2009 World Championships. 
+Starting in 2011, the rule was scrapped, and the old policy which automatically 
+disqualifies runners who false start was reinstituted.  The effect? At the 2011 
+World Championship there were 6 male disqualifications and 4 female 
+disqualifications \citep{iaaf2009falsestart}. By returning to the harsher policy and 
+cracking down on false starts, World Athletics had reduced the number of 
+false starts by two thirds. One study that looked at different false start
+rules that IAAF has imposed from 1997 to 2009 and found that the reaction 
+times when there was less leniency improved by a statistically significant amount
+\citep{haugen2013effect}. This was desirable for World Athletics as false starts can 
+make already long track meets more tedious, both for athletes and viewers watching the 
+television broadcast.  Since 2011, IAAF has instituted the one false start rule
+\citep{iaaf2009falsestart}.  However, we chose not to include women's data in this
+paper due to the numerous papers citing difference in reaction times in men and
+women's reaction times: \citep{lipps2011implications, babicc2009reaction,
+panoutsakopoulos2020gender}
 %Fix quotations
@@
 \jy{cross-reference to the table by its label}
-too quickly.  This table below highlights the difference in reaction times for
+too quickly.  Table~\ref{USAvsWorld} below highlights the difference in reaction times for
 Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
@@
    \bottomrule
+   \label{fig:USAvsWorld}
   \end{tabular}
@@
 %Include latex math symbols for the equations
+The R package \citet{lme4} allows various linear mixed effects models to be
+created, and was the primary modeling package used for our data analysis.
 A gamma mixed linear effects model was chosen as the best model to model the data.
@@
 \begin{center}
-  \begin{tabular}{|c c|} 
-   \hline
+  \begin{tabular}{c c} 
+   \toprule
    Data Set & Probability RxnTime < 0.1 \\ 
-   \hline\hline
+   \midrule
    all finals gamma log & 0.000856 \\ 
-   \hline
    old finals gamma log & 0.000439 \\
-   \hline
    all pooled gamma log & 0.006865 \\
-   \hline
    old pooled gamma log & 0.005761 \\
+   \bottomrule
   \end{tabular}
@@
 Since 1985 Seiko Holding Corporations has served every World Athletics Championship
-as the official timer \citep{Seiko}.  Since 1985, the technology and the ability
+as the official timer \citep{wa2022seiko}.  Since 1985, the technology and the ability
 to accurately predict measurements: long jump distances, false starts, total time,
@@
 between when the ``gun" goes off to start the race and when the pressure changes 
-\citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
+\citep{wa2022seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
 false start as it is considered that no athlete can react so quickly 
@@
 upgraded their timing equipment for sprinting events (100 meter hurdles falls 
-under this category) \citep{WorldAthletics_2013}.  Since 1999, the highest reaction 
+under this category) \citep{wa2013backtage}.  Since 1999, the highest reaction 
 times in the men's 110-meter hurdle were recorded in 2013.  That is not to say that the higher 
@@
 Allen is a very good athlete, as evidenced by his football career at the University of Oregon
-and making it onto the Philadelphia Eagles practice squad this past year \citep{Hurley}.
-Thus the 0.099 reaction time may not have been a product of Devon Allen
-predicting the start of the race but rather a combination of a quick reaction 
-and a possibly faulty sensor.
+and making it onto the Philadelphia Eagles practice squad this past year 
+\citep{hurley2022eagles}. Thus, the 0.099 reaction time may not have been 
+a product of Devon Allen predicting the start of the race but rather a 
+combination of a quick reaction and a possibly faulty sensor.
 

### 2023-04-18 Manuscript/manuscript.tex
Cleaned up some tables

@@
 The model we use is a linear model, following a gamma distribution with a random
-effect, which will be referred to as the year effect.  The year effect is
+effect, which will be referred to as the venue effect.  The venue effect is
 reflective of how significant the changes in the reaction time is for each venue,
@@
 \jy{cross-reference to the table by its label}
-too quickly.  Table~\ref{USAvsWorld} below highlights the difference in reaction times for
+too quickly.  Table~\ref{fig:USAvsWorld} below highlights the difference in reaction times for
 Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
@@
 %Include latex math symbols for the equations
-The R package \citet{lme4} allows various linear mixed effects models to be
-created, and was the primary modeling package used for our data analysis.
+The R package lme3 allows various linear mixed effects models to be created, and 
+was the primary modeling package used for our data analysis \citep{lme4}.
 A gamma mixed linear effects model was chosen as the best model to model the data.
@@
 effect, but are all factors that could potentially impact the venue effect.
-Thus, we decided to inclue only the year effect in the GLMM, which is presented
+Thus, we decided to inclue only the venue effect in the GLMM, which is presented
 as:
@@
 
-\subsection{GLMM}
+\subsection{GLMM} \label{subsec:Results_GLMM}
 %Section 4.1: Gamma model results show the fitting with and without 2022,
@@
   different data motivated the rank-based study.}
+
+As shown in table~\ref{fig:Gamma_parameters}, the standard error of the year
+effect increases for both the men's final and men's pooled data sets when
+including 2022.  The men's final data venue effect error increased from 0.439
+to 0.504 and from 0.0495 to 0.0554 for the men's pooled data.  This demonstrates
+that the inclusion of 2022 increased the randomness of the data and motivates
+the rank-based study~\ref{subsec:Results_Rank}.
+
+  Gamma model with venue effect and link=log
+  \begin{center}
+    \begin{tabular}{c  c  c  c  c  c  c } 
+     \toprule
+     Dataset & Venue standard error & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+     \midrule
+     Older Finals & 0.0439 & -1.901 & 0.0251 & 0.0019 & 0.0109 \\
+     All Finals & 0.0504 & -1.914 & 0.0280 & 0.0025 & 0.0112 \\
+     Older Pooled & 0.0495 & -1.853 & 0.0305 & 0.0024 & 0.0243 \\
+     All Pooled & 0.0554 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\
+     \bottomrule
+     \label{fig:Gamma_parameters}
+    \end{tabular}
+    \end{center}
 
@@
    \bottomrule
+   \label{fig:Sim_probability}
   \end{tabular}
@@
 
-Using the gamma function with link set to log (whose summary statistics can be
-found in the last row of the above table), the probability of the function being
+Using the gamma function with link set to log, the probability of the function being
 less than 0.1 seconds can be calculated to see if Allen's disqualification was an anomaly.
@@
 simulated results) of observing a reaction time below 0.01 seconds in the men's
-finals.  That means that we would expect a time below the threshold to occur
+finals as shown in table~\ref{fig:Sim_probability}.
+That means that we would expect a time below the threshold to occur
 once every 2,277 starts.  Considering there are roughly 8 men's finals reaction
@@
 
-\subsection{Rank-based Comparison}
+\subsection{Rank-based Comparison} \label{subsec:Results_Rank}
 %Section 4.2 is motivated by seiko timing and look at same athletes across
@@
 
-As Seiko timed both events, there should not be a signficant
+As Seiko timed both events, there should not be a significant
 difference if the timing equipment in both instances was working correctly.

### 2023-04-18 Manuscript/manuscript.tex
minor changes

@@
   1) A paragraph to tell the story;
-  2) A paragraph on informal statistical analyses on the internet/social media;
+  2) A paragraph on informal statistical analysis on the internet/social media;
   3) A paragraph on scholarly articles on reaction time in the literature;
@@
 what was found in a 2021 study that concluded the year of competition has a
-significant effect on reaction times \citep*{zhang2021correlation}.  Their
+significant effect on reaction times \citep{zhang2021correlation}.  Their
 analysis looked at data from the 2011-2019 World Championships across multiple
@@
 faster. However, not  all studies have been in favor of a decrease in the
-reaction time barrier. Brosnan in 2016 argues for increasing the reaction time 
-barrier to 0.115 seconds, and cited evidence from European Championships and 
-World Championships in his article \citep{brosnan2017effects}.  A study looking 
+reaction time barrier. \citet*{brosnan2017effects} in 2016 argue for increasing 
+the reaction time barrier to 0.115 seconds, and cited evidence from European 
+Championships and World Championships in their article.  A study looking 
 at the 2008 Olympics likewise supported the conclusion that athletes could not
@@
 \jy{I think you have one objective but two methods to approach it.}
-One of the first objectives of the paper is to try to model the data, which is
+The objective to this paper is to analyze the reaction times at the 2022
+World Track and Field Championship to check for any abnormalities, and we approach
+this two different ways.  Our first method is to try to model the data, which is
 the reaction times of every non-disqualification since the 1999 World Championships. 
@@
 i.e. the venue effect is  responsible for the variation between years.  
-Another objective of this paper is to compare data for athletes who have competed
+The second method of this paper is to compare data for athletes who have competed
 at additional competitions to see how their reaction times compare from previous
@@
    \toprule
-   Data Set & Probability RxnTime < 0.1 \\ 
+   Data Set & Probability RxnTime $<$ 0.1 \\ 
    \midrule

### 2023-04-18 Manuscript/manuscript.tex
Updates from 4/18 meeting

@@
 as:
-
 \begin{equation}
   \label{eq:GLMM}
-  Y_{ij} = \beta_0 + \gamma_i + \epsilon_{ij}
+  Y_{ij} = \beta_0 + \gamma_i + \epsilon_{ij},
 \end{equation}
-
-Where $Y_{ij}$ is the response variable for observation $j$ in year $i$, , 
+where $Y_{ij}$ is the response variable for observation $j$ in year $i$, 
 $\beta_0$ is the intercept, $\gamma_i$ is the random effect term for year $i$, 

### 2023-04-20 Manuscript/manuscript.tex
edits/comments section 1-2

@@
 
-\jy{The bib file needs quality control. The bib keys need to have a consistent
-  style. I suggest using the keys from google scholar.}
+\jy{The bib file needs quality control. Practice my writing tips in the
+  stat-wriging notes: chapter 2 on latex/bibtex.}
 
-\jy{Practice my writing tips in the stat-wriging notes: chapter 2 on
-  latex. E.g., no vertical/horizontal lines in professional tables. Use the
-  tables in my template.}
+\jy{Go over the reference section to see what need to be changed in the bib source}
 
@@
 
-\jy{use the following structure and flow:
-  1) A paragraph to tell the story;
-  2) A paragraph on informal statistical analysis on the internet/social media;
-  3) A paragraph on scholarly articles on reaction time in the literature;
-  4) A paragraph on what's new: methods; results.
-  5) A paragraph of roadmap.}
-
-\jy{could condense a bit}
-%1
 After coming in third in the 110 meter hurdles at the United States Track and 
-Field Championships in Eugene Oregon at the end of June, Allen was expected to 
+Field Championships in Eugene Oregon at the end of June, Devin Allen was expected to
 be a contender for the World Championships, which were scheduled for two months 
@@
 
-%2
-Outside the blog posts from \url{www.LetsRun.com}, there have not been any published
-papers regarding Devon Allen's disqualification in particular.  This is not
-particularly surprising given that his disqualification occurred only last August.
-Letsrun is a website that is part message board and part news.  The message board,
+Allen's disqualification has been discussed extensively in online communities
+such as \url{www.LetsRun.com}, a website that is part message board and part
+news. The message board,
 which functions similarly to Reddit is very active during important running
@@
 the Olympics and the World Track and Field Championships.  However, it is the
+\jy{cite these blog articles formally}
 news section of the website that generated articles such as ``Was Devon Allen
@@
 
-\jy{break into two paragraphs}
-%3
-Other studies have looked at minimum reaction times such as \citet{pain2007sprint}  which looked at
+Reaction time in sprint and hurdle events has been studied by many authors.
+\citet{pain2007sprint} looked at
 reaction times for nine male athletes and analyzed their reaction times.  The
@@
 seconds.  Two other athletes were able to react similarly when under certain
-muscular tightening conditions.  In 2009, IAAF commissioned a study to examine
-the disqualification barrier. It was concluded that the disqualification 
+muscular tightening conditions.  In a study commissioned by IAAF,
+\jy{define acronym at first occurrence}
+\citep{komi2009iaaf} recommended that the disqualification 
 barrier be decreased to allow sprinters to
@@
 Additionally, it was recommended to IAAF that high speed cameras be used to
-change the reaction time criteria from pushing off on the block to body movement.
-\citep{komi2009iaaf}. This could provide a more accurate representation of 
+change the reaction time criteria from pushing off on the block to body
+movement. This could provide a more accurate representation of 
 reaction time, and thus the possibility of a false disqualification would be lowered.
@@
 
-\jy{Not Zhang's. There are coauthors. Consider using citet when appropriate in this paragraph.}
-
-A 2012 follow-up study to Komi's by \citet*{pilianidis2012start} found
+\jy{Could be condensed}
+\citet{pilianidis2012start} found
 significant differences between reaction times at several world championships
@@
 
-%How barrier is picked by regulator
-
-%4
-\jy{I think you have one objective but two methods to approach it.}
+\jy{language needs to be formalized and condensed.}
 The objective to this paper is to analyze the reaction times at the 2022
@@
 the reaction times of every non-disqualification since the 1999 World Championships. 
-The model we use is a linear model, following a gamma distribution with a random
-effect, which will be referred to as the venue effect.  The venue effect is
-reflective of how significant the changes in the reaction time is for each venue,
-i.e. the venue effect is  responsible for the variation between years.  
-The second method of this paper is to compare data for athletes who have competed
+The model we use is a gamma linear mixed-effects model with a random effect
+accounting for the venue (or year) effect. 
+The second method is to compare data for athletes who have competed
 at additional competitions to see how their reaction times compare from previous
@@
 
-%Simplify contribution to one paragraph: collect data and formal statistical
-%analysis such as model used
-%Move a lot of this to methods section
-
-%5
 The rest of this paper is as follows: Section~\ref{sec:Data} describes how data was
-collected and begins to detail the GLMM model which is further developed in the
+collected and begins to detail the generalized linear mixed-effect (GLMM)
+model which is further developed in the
 Section~\ref{sec:Methods}.  The signed-rank test analysis for the same athletes is
@@
 
-\jy{Describe you data types; give details on sources; introduce notations for
-  both types of data
-  present exploratory plots.}
 There are two types of data that is explored throughout this paper: data from
@@
 \subsection{World Championship}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics{Finals_Pooled_Boxplot}
+  \caption{The reaction times from 2019 to 2022 for the men's final (top) and
+  the men's semifinals and finals (bottom).}
+  \label{fig:Boxplots}
+\end{figure}
+
 Data was taken from the World Athletics and covers the men's 110 meter hurdles 
@@
 
-Something that is extremely important to note in the Data is the inconsistency 
-of World Athletics in terms of how they penalize false starts.  From 2007-2009,
+Something that is extremely important to note in the data is the inconsistency 
+of World Athletics in terms of how they penalize false starts.  From 2007--2009,
 the IAAF (the former name for World Athletics, and the governing body for the 
@@
 panoutsakopoulos2020gender}
-%Fix quotations
-%Dont need the titles will add captions, use width =5, height = 3 in aspect ratio
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics{Finals_Pooled_Boxplot}
-  \label{fig:Boxplots}
-  \caption{The reaction times from 2019 to 2022 for the men's final (top) and
-  the men's semifinals and finals (bottom).}
-\end{figure}
 
@@
 
-\jy{set up the models/methods using the notations from last section; cite the
-  right references.}
+The World Championship data and the data beyond the World Champisonship were
+analysed with GLMM and rank-based approaches, respectively.
 
 \subsection{GLMM}
-%Include latex math symbols for the equations
-The R package lme3 allows various linear mixed effects models to be created, and 
+Based on an exploratory analysis, the GLMM model that we chose was a gamma
+mixed-effect model for the reaction times from the World Championship. The
+normal error model on the log-scale was found to be inferior to the gamma model.
+Let $Y_{ij}$ be the reaction time of observation~$j$ in year~$i$.
+Conditional on a venue effect $z_i$ of year~i,  $Y_{ij}$ is modeled by a
+gamma distribution with mean
+\[
+\mu_{i} = \alpha + z_i,
+\]
+and variance
+\[
+\sigma_i
+\]
+...
+\jy{need to add assumptions ...}
+
+The R package \texttt{lme4} allows various linear mixed effects models to be created, and 
 was the primary modeling package used for our data analysis \citep{lme4}.
@@
 effect, but are all factors that could potentially impact the venue effect.
-Thus, we decided to inclue only the venue effect in the GLMM, which is presented
-as:
-\begin{equation}
-  \label{eq:GLMM}
-  Y_{ij} = \beta_0 + \gamma_i + \epsilon_{ij},
-\end{equation}
-where $Y_{ij}$ is the response variable for observation $j$ in year $i$, 
-$\beta_0$ is the intercept, $\gamma_i$ is the random effect term for year $i$, 
-and $\epsilon_{ij}$ is the error term for observation $j$ in year $i$.
-
-%yij denotes ith years and j is observation in year
-%yij will have clustered i and we use random effect to account for this
-%When model is presented talk about how I modeled yij
-%Zij where player is i and cluster, group is year and j
-%Test hypothesis that reaction times are the same vs that they are not
 

### 2023-04-20 Manuscript/manuscript.tex
Updated tables, added some stuff based of Tuesday meeting notes

--- 2023-04-18 Manuscript/manuscript.tex
@@
 the USA Track and Field Championships, ``World" denotes the World Track and Field
-Championships, ``N/A" was used for any athlete that did not compete and thus
-did not register a reaction time.  All numbers listed in the table are the reaction
+Championships,   All numbers listed in the table are the reaction
 time for the respective athletes in seconds.
 
+\begin{table}
 \begin{center}
@@
    \bottomrule
+  \end{tabular}
   \label{fig:USAvsWorld}
-  \end{tabular}
+  \caption{Rection times for four United States Athletes at the 2022 USA Track
+  and Field World Championships and the 2022 World Track and Field Championship.
+  ``N/A" was used for any athlete that did not compete and thus did not register 
+  a reaction time.}
   \end{center}
+\end{table}
 
@@
 $W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is
-the rank of $X_{i}^{*}$.
+the rank of $X_{i}^{*}$. S is given as $S = \mathbb{E}[W^* | X, \delta]$ and
+$E(S) = E(W^*) = \frac{1}{2} \sum_{i=1}^{N} \frac{n_i}{N}$.
 The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
@@
 \citep{jiang2017wilcoxon}.
+---Please see manuscript around line 464 for comments regarding this section---
+
+%I included additional statistics from the paper, I am unsure of how in depth
+%This explanation needs to be
+
+%$F_{i}(x) = \frac{1}{n_i} \sum_{k=1}^{n_i} \mathbb{I}(X_{i,k} \leq x)$
+
+%$F_{i}(x-) = \frac{1}{n_i} \sum_{k=1}^{n_i} \mathbb{I}(X_{i,k} < x)$
+
+%$S = \frac{1}{N + 1} \sum_{i=1}^{N}  \sum_{k=1}^{n_i} \frac{\delta_{ik}}{n_i} \left[ 1 + \frac{1}{2} \sum_{j \neq i} \frac{1}{n_j}
+%\left( F_j(X_{ik}) + F_j(X_{ik-}) \right) \right]$
 
@@
 using the DS method. The DS method is preferable when the cluster sizes are
-informative \citep{jiang2017wilcoxon}. \jy{explain. More data means faster
-  reaction time?}
-We do not care that some
-athletes made it further in either competition than another or that it should
-not matter how elite an athlete is at the 110 meter hurdles once we get to the
-national international competition level.  Cluster sizes vary from 2 observations
-to 6.
+informative \citep{jiang2017wilcoxon}.  In this case, athletes who have more
+data are generally thought to be faster, as they have made it further in the
+competition.  Clusters range from two to six, with two indicating that at each
+competition they only made it to the heats, but a cluster size of six indicates
+that they made the finals in both competitions.  While the number of times
+athletes compete is based off their overall time not their reaction time, the
+two are correlated \citep{delalija2008reaction}.
 
@@
 
-As shown in table~\ref{fig:Gamma_parameters}, the standard error of the year
+As shown in table~\ref{tab:Gamma_parameters}, the standard error of the year
 effect increases for both the men's final and men's pooled data sets when
@@
 
-  Gamma model with venue effect and link=log
-  \begin{center}
-    \begin{tabular}{c  c  c  c  c  c  c } 
+\begin{table}
+  \centering
+  \begin{tabular}{c c c c c c}
       \toprule
-     Dataset & Venue standard error & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
+      Dataset & Venue standard error & Int effect est & Int se & Int var & Resid Var \\
       \midrule
-     Older Finals & 0.0439 & -1.901 & 0.0251 & 0.0019 & 0.0109 \\
-     All Finals & 0.0504 & -1.914 & 0.0280 & 0.0025 & 0.0112 \\
-     Older Pooled & 0.0495 & -1.853 & 0.0305 & 0.0024 & 0.0243 \\
-     All Pooled & 0.0554 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\
+      Older Finals & 0.0439 & $-1.901$ & 0.0251 & 0.0019 & 0.0109 \\
+      All Finals & 0.0504 & $-1.914$ & 0.0280 & 0.0025 & 0.0112 \\
+      Older Pooled & 0.0495 & $-1.853$ & 0.0305 & 0.0024 & 0.0243 \\
+      All Pooled & 0.0554 & $-1.869$ & 0.0348 & 0.0031 & 0.0235 \\
       \bottomrule
-     \label{fig:Gamma_parameters}
   \end{tabular}
-    \end{center}
+  \caption{Summary statistics for the GLMM on various data sets}
+  \label{tab:Gamma_parameters}
+\end{table}
 
@@
 \jy{label the table and reference to it; clean it up}
-\begin{center}
-  \begin{tabular}{c c} 
+\begin{table}
+  \centering
+  \begin{tabular}{c c c c} 
    \toprule
-   Data Set & Probability RxnTime $<$ 0.1 \\ 
+   Data Set & P(RxnTime $<$ 0.1) & P(RxnTime $<$ 0.09) & P(RxnTime $<$ 0.08) \\ 
    \midrule
-   all finals gamma log & 0.000856 \\ 
-   old finals gamma log & 0.000439 \\
-   all pooled gamma log & 0.006865 \\
-   old pooled gamma log & 0.005761 \\
+   all finals gamma log & $8.56\cdot10^{-4}$ & $4.1\cdot10^{-5}$ & $6.0\cdot10^{-7}$ \\ 
+   old finals gamma log & $4.39\cdot10^{-4}$ & $1.63\cdot10^{-5}$ & $2.0\cdot10^{-5}$ \\
+   all pooled gamma log & $6.86\cdot10^{-3}$ & $1.21\cdot10^{-3}$ & $1.3\cdot10^{-4}$ \\
+   old pooled gamma log & $5.76\cdot10^{-3}$ & $9.93\cdot10^{-4}$ & $1.0\cdot10^{-4}$\\
    \bottomrule
-   \label{fig:Sim_probability}
   \end{tabular}
-  \end{center}
+  \caption{Probabilities of observing reaction times less than 0.1, 0.09, and
+  0.08 for various data sets}
+  \label{tab:Sim_probability}
+\end{table}
 
@@
 simulated results) of observing a reaction time below 0.01 seconds in the men's
-finals as shown in table~\ref{fig:Sim_probability}.
+finals as shown in table~\ref{tab:Sim_probability}.
 That means that we would expect a time below the threshold to occur
@@
 
-It is worth noting that changing the reaction time barrier from 0.1 seconds to
+Table ~\ref{tab:Sim_probability} shows that changing the reaction time barrier from 0.1 seconds to
 0.08 seconds drastically reduces the chances of observing a reaction time that

### 2023-04-21 Manuscript/manuscript.tex
glmm method

@@
 \subsection{GLMM}
-Based on an exploratory analysis, the GLMM model that we chose was a gamma
-mixed-effect model for the reaction times from the World Championship. The
-normal error model on the log-scale was found to be inferior to the gamma model.
+Based on an exploratory analysis using the GLMM model for the reaction times
+from the World Championship, the final model that we chose was a gamma
+mixed-effect model. As suggested by the log-likelihood and the Akaike Information
+Criterion (AIC), the normal error model on the log-scale was found to be
+much inferior to the gamma model which better captures the skewness of the
+data. A random effect for the venue or year was also found to be necessary.
 Let $Y_{ij}$ be the reaction time of observation~$j$ in year~$i$.
-Conditional on a venue effect $z_i$ of year~i,  $Y_{ij}$ is modeled by a
-gamma distribution with mean
+Conditional on a venue effect $z_i$ of year~i,  $Y_{ij}$ is modeled by 
+gamma distribution. In the classic generlized linear model (GLM) sense,
+$Y_{ij}$ has conditional mean
 \[
-\mu_{i} = \alpha + z_i,
+g\{E(Y_{ij} | z_i)\} = \alpha + z_i,
 \]
-and variance
-\[
-\sigma_i
-\]
-...
-\jy{need to add assumptions ...}
+and dispersion parameter $\phi$, where $g$ is a known link function and
+$z_i$ is a normally distributed random effect with mean zero and
+variance~$\sigma^2$. The link function $g$ can be the logarithm function to
+ensure the positivity of the conditional mean.
+The gamma mixed-effect model can be fit with the \texttt{glmer()} function in R
+package \texttt{lme4} \citep{lme4}.
 
-The R package \texttt{lme4} allows various linear mixed effects models to be created, and 
-was the primary modeling package used for our data analysis \citep{lme4}.
-A gamma mixed linear effects model was chosen as the best model to model the data.
-Many models were tried and compared to determine what the strongest model.  For
-a fixed data set such as the men's pooled data including 2022, the models with
-the venue effect included performed better than the models without the venue
-effect for each set of comparable models; as indicated by the lower AIC and
-higher log likelihood.  Thus, the venue effect is significant and should be
-included in the model.  It can also be established that a gamma distribution
-better models the data than linear data from comparing the linear model with
-venue effect and them gamma model with a venue effect.  The final decision that
-was made in model selection was to include link = log in the model.  Despite
-doing very little to alter the AIC or log likelihood, the log scale was included
-in order to try and better deal with high outliers. %Include Residuals plot?
+To better understand this model, we can identify the gamma model in terms of the
+commonly used shape and scale parameters. For a gamma distribution with
+shape~$a$ and scale~$b$, we have mean $\mu = ab$ and variance $v = ab^2$. The
+variance as a function of the mean $\mu$ in the GLM sense is
+$v(\mu) = \mu^2 / a$, which implies that the dispersion parameter
+$\phi = 1 / a$. Therefore, for the gamma mixed-effect model, the conditional
+gamma distribution of the reaction time $Y_{ij}$ given the venue effect $z_i$
+has shape $a = 1 / \phi$ and scale $b = g^{-1}(\alpha + z_i) \phi$. The marginal
+distribution of $Y_{ij}$ is a scale-mixture of gamma distributions, which can be
+easily simulated from once the parameters are estimated. A large number of
+random numbers generated from the fitted mixture distribution can be used to
+approimate the probability of observing a reaction time faster than any given
+threshold.  We want the probability of a reaction time being less than 0.1
+seconds in order to gauge if that is a reasonable disqualification barrier.
 
-Once the Gamma mixed effects model was chosen, it needed to be decided which combination
-of effects to use.  It was possible to include a venue effect, a batch effect, 
-or a year and batch effect.  The batch effect is very
+As for the random effect, it was possible to include as well a race effect.
+From our exploration, however, the race effect is very
 difficult to explain or interpret as it means that some heats of races were either
-faster or slower and that everyone in the batch reacted similarly.  Although
+faster or slower and that everyone in the race reacted similarly.  Although
 it is possible that one athlete reacting faster could lead to others reacting
@@
 state of the art.  Any other weather or climate related factors such as humidity,
-precipitation, elevation, etc. would all be negated when looking at the batch
+precipitation, elevation, etc. would all be negated when looking at the race
 effect, but are all factors that could potentially impact the venue effect.
 
-%Throughout exploratory analysis, gamma provides better AIC
-%Spend paragraph on after fitting the model how to figure out 0.1 p value
-As the GLMM model has mixed form, the best way to try to gauge the probability
-of observing a reaction time is to simulate the distribution of times using a 
-large sample.  In order to approximate the distribution, R code was written to 
-try and estimate the probabilities of observing a reaction time as extreme as
-0.1 seconds.  The code re-paramertizes the gamma distribution to be in terms of
-$\mu$ and shape rather than by $\alpha$ and $\beta$.  From the code, it is
-possible to calculate the probability of observing a reaction time faster than
-a given time.  We want the probability of a reaction time being less than 0.1
-seconds in order to gauge if that is a reasonable disqualification barrier.
-
-%T, generate from fitted model explain what
-%The R code does.  Simulate from fitted model. Use distribution of model to 
-%Figure out the p value of reaction time less than 0.1
-
 \subsection{Rank-based Comparison}
-
-%Citations needed
-%Include math formulas and the test statistic
-%Equation of test statistic is from papers cited by clusrank. Asymptootic dist
-%Was found by __ referecne and the implementation is from the clusrank R package
 
@@
 every other year in terms of average reaction time.  Additionally, it was shown
-through a gamma mixed effects model that there are significant year and batch
+through a gamma mixed effects model that there are significant year and race
 effects that greatly impact reaction time.  However, this practically does not make much sense

### 2023-04-21 Manuscript/manuscript.tex
clusrank ds

@@
 \jy{cross-reference to the table by its label}
-too quickly.  Table~\ref{fig:USAvsWorld} below highlights the difference in reaction times for
+too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction times for
 Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
@@
 \end{table}
+
+\jy{You have two such datasets in the analysis. Use one paragraph for each.}
 
@@
 
-\jy{Undefined notations}
-The test statistic of the DS method of the Wilcoxon rank-sum statistic is:
-$Z = \frac{S - E(S)}{\sqrt{\hat{\text{Var}}(S)}}$ where 
-$W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is
-the rank of $X_{i}^{*}$. S is given as $S = \mathbb{E}[W^* | X, \delta]$ and
-$E(S) = E(W^*) = \frac{1}{2} \sum_{i=1}^{N} \frac{n_i}{N}$.
-The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
-\citet{datta2005rank} and the R package ``clusrank" allows for its implementation.
+To test the conjecture that the 2022 World Champion timing device may have led
+to faster recorded reaction times, we compare the reaction times of the same
+athletes who have attended both the 2022 World Champion and other competitions.
+In this setting, we have clustered data with subunit grouping. In particular,
+each athlete is a cluster and the multiple reaction times from the same athlete
+can be from either the 2022 World Championship or otherwise.
+Let $X_{ij}$ be the $j$th reaction time of athelete~$i$, $i = 1, \ldots, n$,
+$j = 1, \ldots, m_i$ where $m_i$ is the number of observations from
+athelete~$i$. Let $\delta_{ij}$ be the group indicator of $X_{ij}$; $\delta_{ij}
+= 1$ if $X_{IJ}$ is in group~1 and $\delta_{ij} = 0$ otherwise. Atheletes are
+assumed to be independent, while subunit observations from the same athelete are
+not. The null hypothesis $H_0$ to be tested is that there is no difference
+between the two groups; i.e., the distribution of $X_{ij}$ remains the same
+regardless of the group indicator $\delta_{ij}$.
+
+\citet{datta2005rank} proposed an extension of the Wilcoxon rank-sum test to
+clustered data with subnit-level grouping. The test is designed based on a
+within-cluster resampling principle. Consider randomly picking one observation
+from each cluster to form a pseudo-sample. Let $X_i^*$ be a random pick from the
+$i$th cluster in the pseudo-sample and $\delta_i^*$ its group indicator. The
+Wilcoxon rank-sum statistic for the pseudo-sample is
+\[
+W^* = \frac{1}{n + 1} + \sum_{i=1}^{n} \delta_{i}^{*} R_{i}^{*},
+\]
+where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the psudo-sample.
+The test statistic $S$ is the average of $W^*$ averaged over all possible
+pseudo-samples conditioning on the observed data and group indicators.
+The mean and variance of $S$ under $H_0$ can be derived so that $S$ can be
+standardized to form a statistic $Z$ which follows standard normal distribution
+asymptotically \citep[p.910]{datta2005rank}. An implementation of this method is
+available from the \texttt{clusWilcox.test()} function with
+\texttt{method = 'ds'} from R package \texttt{clusrank}
 \citep{jiang2017wilcoxon}.
----Please see manuscript around line 464 for comments regarding this section---
-
-%I included additional statistics from the paper, I am unsure of how in depth
-%This explanation needs to be
-
-%$F_{i}(x) = \frac{1}{n_i} \sum_{k=1}^{n_i} \mathbb{I}(X_{i,k} \leq x)$
-
-%$F_{i}(x-) = \frac{1}{n_i} \sum_{k=1}^{n_i} \mathbb{I}(X_{i,k} < x)$
-
-%$S = \frac{1}{N + 1} \sum_{i=1}^{N}  \sum_{k=1}^{n_i} \frac{\delta_{ik}}{n_i} \left[ 1 + \frac{1}{2} \sum_{j \neq i} \frac{1}{n_j}
-%\left( F_j(X_{ik}) + F_j(X_{ik-}) \right) \right]$
-
-Using the \texttt{clusWilcox.test()} function, a Wilcoxon rank sum
-test was performed with the cluster set to be the Athlete and group set to Year
-using the DS method. The DS method is preferable when the cluster sizes are
-informative \citep{jiang2017wilcoxon}.  In this case, athletes who have more
-data are generally thought to be faster, as they have made it further in the
-competition.  Clusters range from two to six, with two indicating that at each
-competition they only made it to the heats, but a cluster size of six indicates
-that they made the finals in both competitions.  While the number of times
-athletes compete is based off their overall time not their reaction time, the
-two are correlated \citep{delalija2008reaction}.
+%% In this case, athletes who have more
+%% data are generally thought to be faster, as they have made it further in the
+%% competition.  Clusters range from two to six, with two indicating that at each
+%% competition they only made it to the heats, but a cluster size of six indicates
+%% that they made the finals in both competitions.  While the number of times
+%% athletes compete is based off their overall time not their reaction time, the
+%% two are correlated \citep{delalija2008reaction}.
 
@@
 
+\jy{Go over the references to see clues to clean in the bib source}
 \bibliographystyle{chicago}

### 2023-04-21 Manuscript/manuscript.tex
comments

@@
 \subsection{GLMM} \label{subsec:Results_GLMM}
-%Section 4.1: Gamma model results show the fitting with and without 2022,
-%talk about p values and thresholds from the simulations (do not include the
-%results in the methods section).
 
-\jy{present the estimated parameters first before deriving the p-value using the
-  estimated parameters. The difference in variance of the random effects using
-  different data motivated the rank-based study.}
+\jy{motivate first why you want to run the analysis with and without the 2022 data.}
 
@@
 
+\jy{Close captions with a period.}
 \begin{table}
@@
    \midrule
-   all finals gamma log & $8.56\cdot10^{-4}$ & $4.1\cdot10^{-5}$ & $6.0\cdot10^{-7}$ \\ 
-   old finals gamma log & $4.39\cdot10^{-4}$ & $1.63\cdot10^{-5}$ & $2.0\cdot10^{-5}$ \\
-   all pooled gamma log & $6.86\cdot10^{-3}$ & $1.21\cdot10^{-3}$ & $1.3\cdot10^{-4}$ \\
-   old pooled gamma log & $5.76\cdot10^{-3}$ & $9.93\cdot10^{-4}$ & $1.0\cdot10^{-4}$\\
+   all finals & $8.56\cdot10^{-4}$ & $4.1\cdot10^{-5}$ & $6.0\cdot10^{-7}$ \\ 
+   old finals & $4.39\cdot10^{-4}$ & $1.63\cdot10^{-5}$ & $2.0\cdot10^{-5}$ \\
+   all pooled & $6.86\cdot10^{-3}$ & $1.21\cdot10^{-3}$ & $1.3\cdot10^{-4}$ \\
+   old pooled & $5.76\cdot10^{-3}$ & $9.93\cdot10^{-4}$ & $1.0\cdot10^{-4}$\\
    \bottomrule
@@
 \subsection{Rank-based Comparison} \label{subsec:Results_Rank}
-%Section 4.2 is motivated by seiko timing and look at same athletes across
-%multiple types of competition: 2019 and national
 
-%Discussion/Conclusion is beyond results.  Limitations, what type of data we
-%hope to have.
+\jy{motivate first; describe the data (what are the grouping variables in the
+  two data sets? use one paragraph for each.}
 
@@
 discussed earlier in the paper, but this data shows that on average there was
-%Use a reference in this part
+\jy{any reference?}
 improvement for British, Polish, French, Brazilian, and Spanish athletes.
@@
 %Put downloadable data spreadsheet
-Here is World Athletics website with the data: \url{https://www.worldathletics.org/results/world-athletics-championships}.
+The World Championship data is available from the World Athletics website:
+\url{https://www.worldathletics.org/results/world-athletics-championships}.
+
+\jy{You have countries beyond USA}
 Here is the data for the USA Track and Field Championships: \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
-
-For the pooled data set
-\begin{center}
-  \begin{tabular}{|c  c  c  c  c  c  c |}
-   \hline 
-   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
-   \hline\hline
-   linear one & -1667 & 835.5 & 0.1529 & 0.0014 & N/A & N/A \\
-   \hline
-   linear year & -1713 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\ 
-   \hline
-   gamma one & -1756 & 880.2 & 6.554 & 0.0598 & N/A & N/A \\
-   \hline
-   gamma year & -1848 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
-   \hline
-   gamma one log & -1756 & 880.2 & -1.880 & 0.0091 & N/A & N/A \\
-   \hline
-   gamma year log & -1848 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\ [0.5ex]
-   \hline
-  \end{tabular}
-  \end{center}
 

### 2023-04-24 Manuscript/manuscript.tex
just started reading intro, have not finished

--- 2023-04-21 Manuscript/manuscript.tex
@@
 
-\title{On Devon Allen's Disqualification at the 2022 Track and Field World
-Championship}
+\title{Was Devon Allen Unjustly Disqualified at the 2022 World Track and Field Championships?}
 
@@
 Devon Allen was disqualified in the Men's 110 meter hurdle final of the 2022
-  Track and Field World Championships after registering a reaction time of 0.099 seconds, 0.001 
-  seconds faster than what is allowed.  Following the games, bloggers on the running website 
-  LetsRun scrutinized the results and performed non-statistical analysis of the data.
-  They found that the reaction time data from the 2022 World Track Championships seemed to be faster 
-  compared to the other datasets they looked at, but did not perform any meaningful statistical 
-  analysis of the data.  This paper questions the reaction time disqualification barrier, which is
-  currently 0.1 seconds to determine whether this is a reasonable threshold.
-  We employ a generalized linear mixed model (GLMM) with a random effect term
-  which will be referred to as the venue effect in order to try and model the
-  data.  Additionally, we employ a signed-rank test for clustered data to compare
-  reaction times for the same athletes at different competitions.
-  This matter needs to be addressed because Devon Allen's disqualification could 
-  be a problem that could continue to re-appear in future world championships.
+Track and Field World Championships after registering a reaction time of 0.099 
+seconds, 0.001 seconds faster than what is allowed. Following the games, 
+bloggers on the running website LetsRun scrutinized the results %and performed 
+%non-statistical analysis of the data. They 
+and found that the reaction time data from the 2022 World Track Championships 
+seemed to be generally faster compared to the other datasets they looked at, but 
+they did not perform any formal %meaningful 
+statistical analysis of the data. This paper questions the reaction time 
+disqualification barrier, which is currently 0.1 seconds, to determine whether 
+this is a reasonable threshold. We employ a generalized linear mixed model 
+(GLMM) with a random %effect term which will be referred to as the 
+venue effect in order to %try and 
+model the data. Additionally, we employ a signed-rank test for clustered data to 
+compare reaction times for the same athletes at different competitions. This 
+matter needs to be addressed because %Devon Allen's 
+disqualification based on allowable reaction time 
+%could be a 
+%problem that could 
+will likely continue to be an issue in future world championships.
 
-\bigskip
-\noindent{\sc Keywords}:
-Reaction Time, Seiko,
+\jy{Avoid references in abstract.}
+
+\jy{This is only the background. An abstract also needs to summarize the 
+contributions.}
+%\bigskip
+\eds{removed Devon Allen from keywords because his name is already in title
+and replaced with others}
+\noindent\sc{Keywords}:
+False Start, Hurdles, Reaction Time, Seiko, 2022 World Track and Field Championships
 
@@
 
-\jy{The bib file needs quality control. Practice my writing tips in the
-  stat-wriging notes: chapter 2 on latex/bibtex.}
+\jy{The bib file needs quality control. The bib keys need to have a consistent
+  style. I suggest using the keys from google scholar.}
 
-\jy{Go over the reference section to see what need to be changed in the bib source}
+\jy{Practice my writing tips in the stat-wriging notes: chapter 2 on
+  latex. E.g., no vertical/horizontal lines in professional tables. Use the
+  tables in my template.}
 
@@
 
+\jy{Not a good opening. Could start by telling the story, recapping the major
+  events and discussions on the internet. Then dig into the relevant
+  literature. Finally outline the contributions of this work.
+}
+
+\jy{use the following structure and flow:
+  1) A paragraph to tell the story;
+  2) A paragraph on informal statistical analyses on the internet/social media;
+  3) A paragraph on scholarly articles on reaction time in the literature;
+  4) A paragraph on what's new: methods; results.
+  5) A paragraph of roadmap.}
+
+%1)
 After coming in third in the 110 meter hurdles at the United States Track and 
-Field Championships in Eugene Oregon at the end of June, Devin Allen was expected to
-be a contender for the World Championships, which were scheduled for two months 
-later, also in Eugene. At the World Championships, Allen made it through the 
+Field Championships in Eugene, Oregon at the end of June, Devon Allen was 
+expected to 
+be a contender for the 2022 World Track and Field Championships, which were also 
+scheduled in Eugene 
+two months later. At the World Championships, Allen made it through the 
 heats and semifinals to reach the finals while posting reaction times of 0.123
-in the heats and 0.101 in the semifinal.  Allen, who attended the University
-of Oregon (Location of US World Championship and the World Championship) was
+in the heats \eds{how many heats are there? Likely more than one, so I think you
+should include all of them, or an average} and 0.101 in the semifinal.  
+Allen, who attended the University
+of Oregon (the location of US World Championship \eds{US World?} and the World 
+Championship) was
 considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
-in early June; only 0.04 seconds away from the world record \citep{wa2022preview}.
+in early June; only 0.04 seconds away from the world record \citep{Preview}.
 Thus his disqualification by reacting in 0.099 seconds instead of 0.01 was met
 with shock from both Allen and fans alike.  NBC, who was broadcasting the
-championships in the United States, later uploaded a video to their YouTube
-channel that show the disqualification and can be found here:
-\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers when
-they found out that Allen was disqualified, angry at the result.  Had Allen 
+championships in the United States, later uploaded a video to their Youtube
+channel that showed the disqualification and can be found here:
+\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers 
+when they found out that Allen was disqualified, angry at the result.  Had Allen 
 reacted just 0.001 seconds slower he would have been allowed to compete and been 
 spared waiting another two years until the next World Track and Field Championship.
-An investigation into reaction times is also important because reaction times have
-been shown to affect the overall time of the race \citep{delalija2008reaction}.
-Thus, faster reaction times lead to faster overall times; which is the goal of
-every sprinter who steps onto the track.
 
-Allen's disqualification has been discussed extensively in online communities
-such as \url{www.LetsRun.com}, a website that is part message board and part
-news. The message board,
-which functions similarly to Reddit is very active during important running
-events, which include during the finals of major events such as
+%2)
+Outside the blog posts from LetsRun.com, there have not been any published
+papers specifically regarding Devon Allen's disqualification.  This is not
+particularly surprising given that his disqualification occurred only last August.
+LetsRun is a website that is part message board and part news. The message board,
+which functions similarly to Reddit, is very active during important running
+events, including during the finals of major events such as
 the Olympics and the World Track and Field Championships.  However, it is the
-\jy{cite these blog articles formally}
 news section of the website that generated articles such as ``Was Devon Allen
-Screwed? There's At Least a 99\% That He Was" and ``The Data Keeps Pouring In and
+Screwed? There's At Least a 99\% Chance That He Was" and ``The Data Keeps Pouring 
+In and
 It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
-Johnson, who created and runs the Letsrun website argued that there was an error
+Johnson, who created and runs the LetsRun website argued that there was an error
 with the timing equipment at the 2022 Track and Field World Championship and
-cited graphs, statistics such as median reaction time, and compared the reaction
-times at the 2022 Track and Field World Championship \citep{johnson2022devon}.
+cited graphs, simple statistics such as median reaction time, and compared the 
+reaction times at the 2022 Track and Field World Championship. 
+\eds{compared the reaction times at the 2022 Track and Field World Championship 
+to what? Also restate here that no formal statistical analyses have been done.}
 
-Reaction time in sprint and hurdle events has been studied by many authors.
-\citet{pain2007sprint} looked at
+%3)
+The International Association of Athletics Federations (also referred to
+as IAAF or World Athletics), which is the governing board for international running
+competitions, commissioned a study in 2009 regarding the reaction time 
+barrier and whether the threshold should be lowered. 
+\eds{verify previous change is correct} 
+The paper analyzed sprinters reaction times and the biological processes 
+that went into reacting to the starting gun and concluded 
+that athletes can react in as low as 0.08 seconds.  
+\eds{what paper are you referring to in the previous sentence?  Please cite.} 
+One study by Pain looked at
 reaction times for nine male athletes and analyzed their reaction times.  The
-fastest athlete of the nine, had on average (even with their two fastest time 
+fastest athlete of the nine, had on average (even with their two fastest times 
 removed) a reaction time of 0.087 seconds and a standard deviation of 0.004
 seconds.  Two other athletes were able to react similarly when under certain
-muscular tightening conditions.  In a study commissioned by IAAF,
-\jy{define acronym at first occurrence}
-\citep{komi2009iaaf} recommended that the disqualification 
-barrier be decreased to allow sprinters to
+muscular tightening conditions \citep{pain2007sprint}. 
+\eds{It was recommended (?sentence was previously missing the opening words)}
+to IAAF that the disqualification barrier be decreased to allow sprinters to
 react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
 Additionally, it was recommended to IAAF that high speed cameras be used to
-change the reaction time criteria from pushing off on the block to body
-movement. This could provide a more accurate representation of 
-reaction time, and thus the possibility of a false disqualification would be lowered.
-The possibility of using a video-based medium has been proved to be a reliable
-measure of reaction time \citep{mudric2015evaluation}.
-
-\jy{Could be condensed}
-\citet{pilianidis2012start} found
+change the reaction time criteria from pushing off on the block to body movement.
+This could provide a more accurate representation of reaction time, and thus the
+possibility of a false disqualification would be lowered \citep{komi2009iaaf}.  
+A follow-up study from 2012 by Pilianidis, Mantzouranis, and Kasabalis found
 significant differences between reaction times at several world championships
 between 1997 and 2009 for the men's 110 meter hurdles.  Another important
-conclusion was that for reaction times did not decrease over the twelve years
-covered by the research in the study \citep{pilianidis2012start}.  This contradicts
-what was found in a 2021 study that concluded the year of competition has a
-significant effect on reaction times \citep{zhang2021correlation}.  Their
-analysis looked at data from the 2011-2019 World Championships across multiple
-sprint events and concluded that from 2011 to 2019, reaction times got
-faster. However, not  all studies have been in favor of a decrease in the
-reaction time barrier. \citet*{brosnan2017effects} in 2016 argue for increasing 
-the reaction time barrier to 0.115 seconds, and cited evidence from European 
-Championships and World Championships in their article.  A study looking 
-at the 2008 Olympics likewise supported the conclusion that athletes could not
-react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
-concluded that visual reaction times to the starting gun may be faster by 0.007
-seconds than the IAAF method, which is from force on block starts 
-\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
-only by 0.01 seconds. It is also worth noting that one study looking at data 
-from various IAAF meets found that the fastest reaction times for men were found
-between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
-because Allen was 27 years old at the 2022 Track and Field World Championship 
-and in his physical peak.
+conclusion was that reaction times did not decrease over the twelve years
+covered by the research in the study \citep{pilianidis2012start}.  This will
+be important when considering the results in section \ref{sec:Results}.
 
-\jy{language needs to be formalized and condensed.}
-The objective to this paper is to analyze the reaction times at the 2022
-World Track and Field Championship to check for any abnormalities, and we approach
-this two different ways.  Our first method is to try to model the data, which is
+Despite the study, IAAF has not changed the reaction barrier since it was
+implemented at the 1999 World Track and Field Championships.  It is confusing
+as to why IAAF would commission a study only to disregard the results and not
+make the necessary changes.  It is unknown how many athletes since 2009 could
+have potentially benefited from this, but it is probable that there are others
+besides Allen who were disqualified.
+%Reaction time barrier has been examined before: cite reviews
+%Then go into specifics and talk about people who wanted to change barrier
+%General reviews are important
+%People have studied the biological processes, cite a few references
+%How barrier is picked by regulator
+%Read abstract and into and summarize into about one sentence what each paper contributes
+%1 and 2 can be merged but 3 need to stand alone
+
+4)
+One of the first objectives of the paper is to try to model the data, which is
 the reaction times of every non-disqualification since the 1999 World Championships. 
-The model we use is a gamma linear mixed-effects model with a random effect
-accounting for the venue (or year) effect. 
-The second method is to compare data for athletes who have competed
+When determining the most effective model to use, several variations were
+explored across the 4 different data sets: a linear model without a venue effect,
+linear model with venue effect, a gamma model without a venue effect, a gamma model
+with a venue effect, a gamma model without a venue effect and link set to log instead
+of the default which is Gaussian.  The R package lme4 was very helpful with this
+as it supported the Generalized Linear Mixed Model (GLMM) which allowed 
+linear models with various distributions to be modeled\citep{Rpkg:lme4}.  The
+model we use is a liner model with a fixed effect and random effect and follows
+the gamma distribution. The fixed effect comes in the form of an intercept, and the
+random effect will be called the venue effect and is reflective of how significant
+the changes in the reaction time is for each venue, i.e. the venue effect is 
+responsible for the variation between years.
+
+Another objective of this paper is to compare data for athletes who have competed
 at additional competitions to see how their reaction times compare from previous
-competitions to this one. First we can look at the reaction times of athletes who
-competed at the World Track and Field Championship in both 2019 and 2022 and
-compare the times, and secondly we can look at how athletes performed compared
-to how they did at their national/qualifiers meet. In order to reduce any chances
-of athletes improving reaction time, we can also look at data from qualifying meets.
-Almost every athlete had to qualify by proving they are elite at the national level, 
-usually by coming in at least fourth. Thus, data was compiled on athletes who competed
+competitions to this one. First we can look at the reaction times of athletes who competed at the
+World Track and Field Championship in both 2019 and 2022 and compare the times,
+and secondly we can look at how athletes performed compared to how they did at
+their national/qualifiers meet.  We would expect that athletes of this caliber,
+who are world-class athletes, to not improve significantly from one year to the
+next.  In order to reduce any chances of athletes improving reaction time, we
+can also look at data from qualifying meets.  Almost every athlete had to qualify 
+by proving they are elite at the national level, usually by coming in at least
+fourth. Thus, data was compiled on athletes who competed
 at the 2022 World Track and Field Championships on what these athletes also ran 
@@
 nearly null, as the national competitions took place between June and July 2022 
-and the World Championships were in August. The equipment used to measure reaction
-time cannot be guaranteed to be the same from national to international
-competition, but the time span between meets has been reduced considerably.
+and the World Championships were in August.  Thus, it should be expected that 
+elite athletes such as these will not get reasonably faster in such a short time
+span. The equipment used to measure reaction time cannot be guaranteed to be
+the same from national to international competition, but the time span between
+meets has been reduced considerably.
+%Simplify contribution to one paragraph: collect data and formal statistical
+%analysis such as model used
+%Move a lot of this to methods section
+5)
+The rest of this paper is as follows: Section~\ref{sec:Data} describes how data was
+collected and begins to detail the GLMM model which is further developed in the
+Section~\ref{sec:Methods}.  The signed-rank test analysis for the same athletes is
+also included in the Section~\ref{sec:Methods}.  The Section~\ref{sec:Results}
+section provides the probabilities from the methods section and the paper ends 
+by questioning the reaction time barrier.
 
-The rest of this paper is as follows: Section~\ref{sec:Data} describes how data was
-collected and begins to detail the generalized linear mixed-effect (GLMM)
-model which is further developed in the
-Section~\ref{sec:Methods}.  The signed-rank test analysis for the same athletes is
-also included in Section~\ref{sec:Methods}.  Section~\ref{sec:Results}
-section provides the probabilities from the methods section and the paper ends 
-by questioning the reaction time barrier.  Lastly, section~\ref{sec:Discussion}
-discusses the impact and limitations of the paper.
+%Follow Anthony's work, explicit references as to what each section does
 
-\section{Data} \label{sec:Data}
+%Seiko section should go in the Results section.  From results it seems reasonable
+%question the data and talk about Seiko
 
-There are two types of data that is explored throughout this paper: data from
-1999-2022 for every world championship, and data from athletes who competed
-at both the 2019 and 2022 World Championships and 2022 national competitions
-and the 2022 World Championship.  The first data set is used for the GLMM portion
-of the paper, while the second data set is used for the rank based comparisons.
+\section{Data} \label{sec:data}
 
-\subsection{World Championship}
+\jy{Describe you data types; give details on sources; introduce notations for
+  both types of data
+  present exploratory plots.}
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics{Finals_Pooled_Boxplot}
-  \caption{The reaction times from 2019 to 2022 for the men's final (top) and
-  the men's semifinals and finals (bottom).}
-  \label{fig:Boxplots}
-\end{figure}
+  \begin{equation}
+    Y_{ij} = \beta_0 + \beta_1 X_{ij} + \gamma_i + \epsilon_{ij}
+    \end{equation}
 
-Data was taken from the World Athletics and covers the men's 110 meter hurdles 
-and the women's 100 meter hurdles from 1999 to 2022.  The variable of prediction 
-interest is reaction time, and the predicting variables are year, stage of 
-competition (heats, semifinal, final), total time, and gender.  
-The data can be found in the appendix~\ref{sec:Appendix}. Factors such as name, 
-country, and identification number were not considered important for the 
-purposes of this research besides the times for Devon Allen and the other 
-United States athletes mentioned above.  The variable ``Stage"
-refers to whether the observation occurred during the ``heats" (Preliminary round) which
-is denoted by ``H" in the data, ``S" for ``semifinals" (Of which there are either 2 or 3 
-heats each year), and ``F" for finals. The variable ``Gender" has levels ``M" and ``F" 
-which stand for ``male" and ``female" respectively.  The variable ``Batch" takes a 
+    where $Y_{ij}$ is the response variable for observation $j$ in year $i$, 
+    $X_{ij}$ is the predictor variable for observation $j$ in year $i$, 
+    $\beta_0$ is the intercept, $\beta_1$ is the coefficient for $X_{ij}$, 
+    $\gamma_i$ is the random effect term for year $i$, and $\epsilon_{ij}$ 
+    is the error term for observation $j$ in year $i$.
+
+  %yij denotes ith years and j is observation in year
+  %yij will have clustered i and we use random effect to account for this
+  %When model is presented talk about how I modeled yij
+  %Zij where player is i and cluster, group is year and j
+  %Test hypothesis that reaction times are the same vs that they are not
+Something that is extremely important to note in the Data is the inconsistency of World Athletics 
+in terms of how they penalize false starts.  From 2007-2009, the IAAF (the former name for World 
+Athletics, and the governing body for the World Track and Field Champions) instituted a rule change
+that allowed one warning false start before a sprinter was disqualified \citep{False-Start}  This 
+essentially gave sprinters a second chance and reduced the penalty for a false start.  There were 
+18 male and 7 female false starts at the 2007 World Championships, 18 male and 7 female starts at 
+the 2009 World Championships. Starting in 2011, the rule was scrapped, and the old policy which 
+automatically disqualifies runners who false start was reinstituted.  The effect? At the 2011 World
+Championship there were 6 male disqualifications and 4 female disqualifications \citep{False-Start}. 
+By returning to the harsher policy and cracking down on false starts, World Athletics had reduced the
+number of false starts by two thirds. This was desirable for World Athletics as false starts can 
+make already long track meets more tedious, both for athletes and viewers watching the 
+television broadcast.  Since 2011, IAAF has instituted the one false start rule
+\citep{Pilianidis}.
+
+Data was copied from the World Athletics website and pasted into an Excel
+spreadsheet. The data covers the men's 110 meter hurdles and the women's 100
+meter hurdles from 1999 to 2022.  The variable of prediction interest is reaction
+time, and the predicting variables are year, stage of competition (heats, 
+semifinal, final), total time, and gender.  The data can be found in the appendix:.  
+Factors such as name, country, and identification number were
+not considered important for the purposes of this research besides the times for
+Devon Allen and the other United States athletes mentioned above.  The variable "Stage"
+refers to whether the observation occurred during the "heats" (Preliminary round) which
+is denoted by "H" in the data, "S" for "semifinals" (Of which there are either 2 or 3 
+heats each year), and "F" for finals. The variable "Gender" has levels "M" and "F" 
+which stand for "male" and "female" respectively.  The variable "Batch" takes a 
 numerical value and every heat of every round from 1999 to 2022 was assigned it's
@@
 correlated with other runners reacting quickly (one runner reacting quickly may
-result in other runners reacting quickly).  There is no statistical significance to
+result in other runners reacting quickly).  There is no statistical signficance to
 how the batches were labeled, they were ordered chronologically by meet but
 reverse chronologically by year (Batch 1 is a preliminary Heat of 2022 and Batch
-111 corresponds to the 1999 finals). Included are
-box plots of the reaction times of the men's finals and the men's semifinals and
-finals pooled together.  We will refer to the latter as the ``pooled data"
-throughout the paper.  The heats or preliminary rounds were not studied because
-athletes react faster as the competition progresses.  Athletes react faster in
-the finals than the semifinals which are faster than the heats 
-\citep{zhang2021correlation}.  If we want to determine if Allen's reaction time
-is reasonable, then we thus need to look at the men's final data, with the men's
-pooled data being included in order to increase the sample size.  In 2022, there
-were only five data points due to two disqualifications and one athlete who 
-qualified but did not compete.
-The data that looked at in the paper is best visualized using a box plot as show 
-in Figure~\ref{fig:Boxplots}.  It is clear that the 2022 box plot for both the
-men's finals and semifinals data is lower (indicating a faster reaction time)
-than in previous years.  The median for the men's pooled data in 2022 was 0.129
-and the median for the men's final data in 2022 was 0.126.  For comparison,
-Brosnan found reaction times of 0.156 seconds for men when looking at data from
-1999 to 2014 \citep{brosnan2017effects}.
+111 corresponds to the 1999 finals). The data that looked at in the paper
+is best visualized using a box plot. Figure~\ref{fig:DataBoxPlot}.  Included are
+box plots of the reaction times of the men's semi finals and finals pooled
+together and just the finals.
+%Fix quotations
+%Dont need the titles will add captions, use width =5, height = 3 in aspect ratio
+\begin{figure}[h]
+  \centering 
+  \includegraphics[width=\textwidth,height=4in]{DataBoxPlot}
+  \caption{Boxplot of semifinals and finals data}\label{fig:DataBoxPlot}
+\end{figure}
+\jy{The plot was stretched. Don't. Get their sizes right before generating them.}
 
-Something that is extremely important to note in the data is the inconsistency 
-of World Athletics in terms of how they penalize false starts.  From 2007--2009,
-the IAAF (the former name for World Athletics, and the governing body for the 
-World Track and Field Champions) instituted a rule change that allowed one 
-warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}  This 
-essentially gave sprinters a second chance and reduced the penalty for a false 
-start.  There were 18 male and 7 female false starts at the 2007 World 
-Championships, 18 male and 7 female starts at the 2009 World Championships. 
-Starting in 2011, the rule was scrapped, and the old policy which automatically 
-disqualifies runners who false start was reinstituted.  The effect? At the 2011 
-World Championship there were 6 male disqualifications and 4 female 
-disqualifications \citep{iaaf2009falsestart}. By returning to the harsher policy and 
-cracking down on false starts, World Athletics had reduced the number of 
-false starts by two thirds. One study that looked at different false start
-rules that IAAF has imposed from 1997 to 2009 and found that the reaction 
-times when there was less leniency improved by a statistically significant amount
-\citep{haugen2013effect}. This was desirable for World Athletics as false starts can 
-make already long track meets more tedious, both for athletes and viewers watching the 
-television broadcast.  Since 2011, IAAF has instituted the one false start rule
-\citep{iaaf2009falsestart}.  However, we chose not to include women's data in this
-paper due to the numerous papers citing difference in reaction times in men and
-women's reaction times: \citep{lipps2011implications, babicc2009reaction,
-panoutsakopoulos2020gender}
+\jy{Remove title from the figures. Make the figure captions self-contained.}
+\begin{figure}[h]
+  \centering 
+  \includegraphics{BoxplotFinals}
+  \label{fig:BoxplotFinals}
+\end{figure}
 
-\subsection{Beyond World Championship}
 This section was motivated from a qualitative analysis that started by looking
@@
 competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
-Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
+Grant Holloway, and Devon Allen. Every athlete reacted faster in all of the World 
 Track and Field Championship races compared to the USA Track and Field 
@@
 Field Championship compared to the USA Track and Field Championship as were his
-teammates.  This comparison suggests that the timing system used between the two
+teamates.  This comparison suggests that the timing system used between the two
 meets may have been meaningfully different to the point where Devon Allen was
 disqualified more because of faulty equipment rather than because he reacted
-\jy{cross-reference to the table by its label}
-too quickly.  Table~\ref{fig:USAvsWorld} below highlights the difference in reaction times for
-Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
-the USA Track and Field Championships, ``World" denotes the World Track and Field
-Championships,   All numbers listed in the table are the reaction
+too quickly.  This table below highlights the difference in reaction times for
+Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. "USA" denotes
+the USA Track and Field Championships, "World" denotes the World Track and Field
+Championships, "N/A" was used for any athlete that did not compete and thus
+did not register a reaction time.  All numbers listed in the table are the reaction
 time for the respective athletes in seconds.
 
-\begin{table}
 \begin{center}
-  \begin{tabular}{c c c c c c c} 
-   \toprule
+  \begin{tabular}{|c c c c c c c|} 
+   \hline
    Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
-   \midrule
+   \hline\hline
    Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
+   \hline
    Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
+   \hline
    Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
+   \hline
    Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
-   \bottomrule
+   \hline
   \end{tabular}
-  \label{fig:USAvsWorld}
-  \caption{Rection times for four United States Athletes at the 2022 USA Track
-  and Field World Championships and the 2022 World Track and Field Championship.
-  ``N/A" was used for any athlete that did not compete and thus did not register 
-  a reaction time.}
   \end{center}
-\end{table}
 
@@
 
-\section{Methods} \label{sec:Methods}
+\section{Methods} \label{sec:meth}
 
-The World Championship data and the data beyond the World Champisonship were
-analysed with GLMM and rank-based approaches, respectively.
+\jy{set up the models/methods using the notations from last section; cite the
+  right references.}
 
 \subsection{GLMM}
-Based on an exploratory analysis using the GLMM model for the reaction times
-from the World Championship, the final model that we chose was a gamma
-mixed-effect model. As suggested by the log-likelihood and the Akaike Information
-Criterion (AIC), the normal error model on the log-scale was found to be
-much inferior to the gamma model which better captures the skewness of the
-data. A random effect for the venue or year was also found to be necessary.
-Let $Y_{ij}$ be the reaction time of observation~$j$ in year~$i$.
-Conditional on a venue effect $z_i$ of year~i,  $Y_{ij}$ is modeled by 
-gamma distribution. In the classic generlized linear model (GLM) sense,
-$Y_{ij}$ has conditional mean
-\[
-g\{E(Y_{ij} | z_i)\} = \alpha + z_i,
-\]
-and dispersion parameter $\phi$, where $g$ is a known link function and
-$z_i$ is a normally distributed random effect with mean zero and
-variance~$\sigma^2$. The link function $g$ can be the logarithm function to
-ensure the positivity of the conditional mean.
-The gamma mixed-effect model can be fit with the \texttt{glmer()} function in R
-package \texttt{lme4} \citep{lme4}.
+%Include latex math symbols for the equations
+A gamma mixed linear effects model was chosen as the best model to model the data.
+Many models were tried and compared to determine what the strongest model.  For
+a fixed data set such as the men's pooled data including 2022, the models with
+the venue effect included performed better than the models without the venue
+effect for each set of comparable models; as indicated by the lower AIC and
+higher log likelihood.  Thus, the venue effect is significant and should be
+included in the model.  It can also be established that a gamma distribution
+better models the data than linear data from comparing the linear model with
+venue effect and them gamma model with a venue effect.  The final decision that
+was made in model selection was to include link = log in the model.  Despite
+doing very little to alter the AIC or log likelihood, the log scale was included
+in order to try and better deal with high outliers. %Include Residuals plot?
 
-To better understand this model, we can identify the gamma model in terms of the
-commonly used shape and scale parameters. For a gamma distribution with
-shape~$a$ and scale~$b$, we have mean $\mu = ab$ and variance $v = ab^2$. The
-variance as a function of the mean $\mu$ in the GLM sense is
-$v(\mu) = \mu^2 / a$, which implies that the dispersion parameter
-$\phi = 1 / a$. Therefore, for the gamma mixed-effect model, the conditional
-gamma distribution of the reaction time $Y_{ij}$ given the venue effect $z_i$
-has shape $a = 1 / \phi$ and scale $b = g^{-1}(\alpha + z_i) \phi$. The marginal
-distribution of $Y_{ij}$ is a scale-mixture of gamma distributions, which can be
-easily simulated from once the parameters are estimated. A large number of
-random numbers generated from the fitted mixture distribution can be used to
-approimate the probability of observing a reaction time faster than any given
-threshold.  We want the probability of a reaction time being less than 0.1
-seconds in order to gauge if that is a reasonable disqualification barrier.
+%Throughout exploratory analysis, gamma provides better AIC
+%Spend paragraph on after fitting the model how to figure out 0.1 p value
 
-As for the random effect, it was possible to include as well a race effect.
-From our exploration, however, the race effect is very
+Once the Gamma mixed effects model was chosen, it needed to be decided which combination
+of effects to use.  It was possible to include a venue effect, a batch effect, or a year
+and batch effect.  Density plots were created that simulated the probability of a mean
+density less than 0.1, which is the current threshold for a reaction time
+disqualification.  The density plots are shown below.  The batch effect is very
 difficult to explain or interpret as it means that some heats of races were either
-faster or slower and that everyone in the race reacted similarly.  Although
+faster or slower and that everyone in the batch reacted similarly.  Although
 it is possible that one athlete reacting faster could lead to others reacting
@@
 state of the art.  Any other weather or climate related factors such as humidity,
-precipitation, elevation, etc. would all be negated when looking at the race
+precipitation, elevation, etc. would all be negated when looking at the batch
 effect, but are all factors that could potentially impact the venue effect.
 
-\subsection{Rank-based Comparison}
+%T, generate from fitted model explain what
+%The R code does.  Simulate from fitted model. Use distribution of model to 
+%Figure out the p value of reaction time less than 0.1
+To try and approximate the distribution of the reaction times, R code was written
+to try and estimate the probabilities of observing a reaction time as extreme as
+0.1 seconds.  The code re-paramertizes the gamma distribution to be in terms of
+$\mu$ and shape rather than by $\alpha$ and $\beta$.
 
-\jy{Undefined notations}
+At a size of n=12,000,000, the calculated probability of a reaction
+time being below 0.1 seconds was as follows for the various different data sets:
+
+\begin{center}
+  \begin{tabular}{|c c|} 
+   \hline
+   Data Set & Probability RxnTime < 0.1 \\ 
+   \hline\hline
+   all finals gamma log & 0.000864 \\ 
+   \hline
+   old finals gamma log & 0.000429 \\
+   \hline
+   all pooled gamma log & 0.006847 \\
+   \hline
+   old pooled gamma log & 0.005749 \\
+  \end{tabular}
+  \end{center}
+
+\subsection{Rank-based comparison for Clustered}
+
+%Citations needed
+%Include math formulas and the test statistic
+%Equation of test statistic is from papers cited by clusrank. Asymptootic dist
+%Was found by __ referecne and the implementation is from the clusrank R package
+
 The test statistic of the DS method of the Wilcoxon rank-sum statistic is:
@@
 $W^* = \frac{1}{N} + \sum_{i=1}^{N} \delta_{i}^{*} R_{i}^{*}$, $R_{i}^{*}$ is
-the rank of $X_{i}^{*}$. S is given as $S = \mathbb{E}[W^* | X, \delta]$ and
-$E(S) = E(W^*) = \frac{1}{2} \sum_{i=1}^{N} \frac{n_i}{N}$.
+the rank of $X_{i}^{*}$.
 The asymptomatic distribution of the Wilcoxon rank-sum statistic was found by
-\citet{datta2005rank} and the R package ``clusrank" allows for its implementation.
+\citet{datta2005rank} and the R package "clusrank" allows for its implementation.
 \citep{jiang2017wilcoxon}.
----Please see manuscript around line 464 for comments regarding this section---
 
-%I included additional statistics from the paper, I am unsure of how in depth
-%This explanation needs to be
+Using the "clusWilcox.test" function, a grouped and clustered Wilcoxon rank sum
+test was performed with the cluster set to be the Athlete and group set to Year.
+The method was set to be "DS" which is used for the Data and Satten method, which
+is preferable when the cluster sizes are informative \citep{jiang2017wilcoxon}.  We do not care that some
+athletes made it further in either competition than another or that it should
+not matter how elite an athlete is at the 110 meter hurdles once we get to the
+national international competition level.  Cluster sizes vary from 2 observations
+to 6.
 
-%$F_{i}(x) = \frac{1}{n_i} \sum_{k=1}^{n_i} \mathbb{I}(X_{i,k} \leq x)$
-
-%$F_{i}(x-) = \frac{1}{n_i} \sum_{k=1}^{n_i} \mathbb{I}(X_{i,k} < x)$
-
-%$S = \frac{1}{N + 1} \sum_{i=1}^{N}  \sum_{k=1}^{n_i} \frac{\delta_{ik}}{n_i} \left[ 1 + \frac{1}{2} \sum_{j \neq i} \frac{1}{n_j}
-%\left( F_j(X_{ik}) + F_j(X_{ik-}) \right) \right]$
-
-Using the \texttt{clusWilcox.test()} function, a Wilcoxon rank sum
-test was performed with the cluster set to be the Athlete and group set to Year
-using the DS method. The DS method is preferable when the cluster sizes are
-informative \citep{jiang2017wilcoxon}.  In this case, athletes who have more
-data are generally thought to be faster, as they have made it further in the
-competition.  Clusters range from two to six, with two indicating that at each
-competition they only made it to the heats, but a cluster size of six indicates
-that they made the finals in both competitions.  While the number of times
-athletes compete is based off their overall time not their reaction time, the
-two are correlated \citep{delalija2008reaction}.
-
-\section{Results} \label{sec:Results}
-
-\subsection{GLMM} \label{subsec:Results_GLMM}
-%Section 4.1: Gamma model results show the fitting with and without 2022,
-%talk about p values and thresholds from the simulations (do not include the
-%results in the methods section).
-
-\jy{present the estimated parameters first before deriving the p-value using the
-  estimated parameters. The difference in variance of the random effects using
-  different data motivated the rank-based study.}
-
-As shown in table~\ref{tab:Gamma_parameters}, the standard error of the year
-effect increases for both the men's final and men's pooled data sets when
-including 2022.  The men's final data venue effect error increased from 0.439
-to 0.504 and from 0.0495 to 0.0554 for the men's pooled data.  This demonstrates
-that the inclusion of 2022 increased the randomness of the data and motivates
-the rank-based study~\ref{subsec:Results_Rank}.
-
-\begin{table}
-  \centering
-  \begin{tabular}{c c c c c c}
-      \toprule
-      Dataset & Venue standard error & Int effect est & Int se & Int var & Resid Var \\
-      \midrule
-      Older Finals & 0.0439 & $-1.901$ & 0.0251 & 0.0019 & 0.0109 \\
-      All Finals & 0.0504 & $-1.914$ & 0.0280 & 0.0025 & 0.0112 \\
-      Older Pooled & 0.0495 & $-1.853$ & 0.0305 & 0.0024 & 0.0243 \\
-      All Pooled & 0.0554 & $-1.869$ & 0.0348 & 0.0031 & 0.0235 \\
-      \bottomrule
-  \end{tabular}
-  \caption{Summary statistics for the GLMM on various data sets}
-  \label{tab:Gamma_parameters}
-\end{table}
-
-At a size of $n=10,000,000$, the calculated probability of a reaction
-time being below 0.1 seconds was as follows for the various different data sets:
-\jy{label the table and reference to it; clean it up}
-\begin{table}
-  \centering
-  \begin{tabular}{c c c c} 
-   \toprule
-   Data Set & P(RxnTime $<$ 0.1) & P(RxnTime $<$ 0.09) & P(RxnTime $<$ 0.08) \\ 
-   \midrule
-   all finals gamma log & $8.56\cdot10^{-4}$ & $4.1\cdot10^{-5}$ & $6.0\cdot10^{-7}$ \\ 
-   old finals gamma log & $4.39\cdot10^{-4}$ & $1.63\cdot10^{-5}$ & $2.0\cdot10^{-5}$ \\
-   all pooled gamma log & $6.86\cdot10^{-3}$ & $1.21\cdot10^{-3}$ & $1.3\cdot10^{-4}$ \\
-   old pooled gamma log & $5.76\cdot10^{-3}$ & $9.93\cdot10^{-4}$ & $1.0\cdot10^{-4}$\\
-   \bottomrule
-  \end{tabular}
-  \caption{Probabilities of observing reaction times less than 0.1, 0.09, and
-  0.08 for various data sets}
-  \label{tab:Sim_probability}
-\end{table}
-
-Using the gamma function with link set to log, the probability of the function being
-less than 0.1 seconds can be calculated to see if Allen's disqualification was an anomaly.
-Prior to the 2022 World Championship, there was a 0.0439\% chance (based on the
-simulated results) of observing a reaction time below 0.01 seconds in the men's
-finals as shown in table~\ref{tab:Sim_probability}.
-That means that we would expect a time below the threshold to occur
-once every 2,277 starts.  Considering there are roughly 8 men's finals reaction
-times every 2 years (The World Championship is biennial), every 569 years there
-should be a time as extreme as Allen's.  This seems extremely unlikely, and thus
-it seems like there may be an additional explanation for the fast reaction times.
-
-It is interesting that the probability of observing a reaction time below 0.1
-seconds increases drastically for the finals data and increases by a smaller margin
-for the pooled data set after 2022 is included in the data set.  By including 2022
-in the data, the chances of observing an extreme time are increased, showing that
-2022 is a high outlier for reaction times compared to other years.
-
-Table ~\ref{tab:Sim_probability} shows that changing the reaction time barrier from 0.1 seconds to
-0.08 seconds drastically reduces the chances of observing a reaction time that
-would break the barrier.  For the dataset of all finals data, the probability
-associated with observing a reaction time below 0.08 seconds was $6\cdot10^{-7}$.
-If IAAF were to change the reaction time the way that \citet{komi2009iaaf}'s paper
-suggests, the probability would be so much lower that athletes would likely not
-have a case to make when arguing against a disqualification.
-
-%Discussion should show that random effect
-%and standard error increase and that 2022 is special.  Now talk about Seiko
-%timing equipment
-
-%Citation needed in this section
-Since 1985 Seiko Holding Corporations has served every World Athletics Championship
-as the official timer \citep{wa2022seiko}.  Since 1985, the technology and the ability
-to accurately predict measurements: long jump distances, false starts, total time,
-reaction time, etc. have all dramatically improved.  Seiko did not start tracking
-reaction time as an official measurement until 1999 in the men's and women's 100 meter
-hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
-technology to the World Track and Field Championships, the highest stakes in the world
-of running outside of the Olympics. Seiko's technology for detecting reaction 
-times relies on the pressure that athletes
-exert on the plate when they push off.  Their systems measure the time differential
-between when the ``gun" goes off to start the race and when the pressure changes 
-\citep{wa2022seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
-false start as it is considered that no athlete can react so quickly 
-\citep{Seiko-Timing}.  Thus, a reaction time of for example 0.05, suggests that 
-the athlete predicted the gun and it was luck that caused their abnormally high 
-time.  At the World Championship level, World Athletics and Seiko want to remove 
-that element of luck and thus impose the 0.1 second barrier. In 2013, Seiko 
-upgraded their timing equipment for sprinting events (100 meter hurdles falls 
-under this category) \citep{wa2013backtage}.  Since 1999, the highest reaction 
-times in the men's 110-meter hurdle were recorded in 2013.  That is not to say that the higher 
-times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
-to 2022, Seiko again upgraded its technology but for its jump management system.
-
-\subsection{Rank-based Comparison} \label{subsec:Results_Rank}
-%Section 4.2 is motivated by seiko timing and look at same athletes across
-%multiple types of competition: 2019 and national
-
-%Discussion/Conclusion is beyond results.  Limitations, what type of data we
-%hope to have.
-
-When the clusrank rank sum test was performed, the ``ds" method produced a z score of
+When the clusrank rank sum test was performed, the "ds" method produced a z score of
 2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
@@
 again performing a clustered analysis for a Wilcoxon ranked sum test
-results in two new z scores and p values.  The ``ds" method produced a z score of
+results in two new z scores and p values.  The "ds" method produced a z score of
 3.6069 and p value of 0.000155.  This result is highly significant and shows that
@@
 
-As Seiko timed both events, there should not be a significant
-difference if the timing equipment in both instances was working correctly.
-  To put it simply, there is not a reasonable explanation for 
-this. It is inconceivable that so many athletes would improve such a significant
-amount in such a small amount of time. 6 countries are represented in this 
-comparison, which drastically decreases the chances of this issue being caused by
-one specific country.  The United States athlete's improvement in times was
-discussed earlier in the paper, but this data shows that on average there was
-%Use a reference in this part
-improvement for British, Polish, French, Brazilian, and Spanish athletes.
+\section{Results}
+\jy{summarize your findings; discuss limitations and possible future works.}
 
-\section{Discussion}\label{sec:Discussion}
-
-One conclusion from \citet{zhang2021correlation}'s analysis was that athletes
-reaction times improved from 2011 to 2019.  However this study does not consider 
-data from 1999 to 2009, and over this timespan athletes get slower and then
-faster.  A boxplot~\ref{fig:Boxplots} shows that for the men's pooled data, the 
-year with the fastest median reaction time before 2022 was 1999.  Thus, this 
-reject's Zhang's conclusion that athletes are getting faster because they were 
-fast over twenty years ago. One possible explanation for why 2011 had such
-high variation was because of the IAAF rule discussed earlier that increased the
-penalty of a false start to an instant disqualification \citep{iaaf2009falsestart}.
-Athletes may have been overly cautious to false start and may have consciously
-or sub-consciously reacted slower as a result.  Instead, a more appropriate 
-conclusion may be that there is a lot of variability in reaction times and that
-the context behind the races matters.
-
-It is worth noting that in the semi-finals of the World Track and
-Field Championships; Allen's reaction time was 0.101 which is only 0.001 above
-the legal limit.  What this suggests is that Allen may have strong reflexes
-and be able to react extremely well to the sound of the gun.  It seems unlikely that
-Allen would be able to correctly predict the gun with such precision two times in a row,
-which is the exact reason that the 0.10 second reaction time disqualification barrier was
-imposed in the first place.  But Allen showed that at the Championship he was able to
-react extremely fast two times, which may suggest skill rather than luck. Devon 
-Allen is a very good athlete, as evidenced by his football career at the University of Oregon
-and making it onto the Philadelphia Eagles practice squad this past year 
-\citep{hurley2022eagles}. Thus, the 0.099 reaction time may not have been 
-a product of Devon Allen predicting the start of the race but rather a 
-combination of a quick reaction and a possibly faulty sensor.
-
-Devon Allen's disqualification at the World Track and Field Championships was
-possibly the result of both a faulty timing equipment and Devon Allen
-reacting extremely quickly to the start.  The code and data from R showed that
-the 2022 World Track and Field Championships were a low outlier compared to
-every other year in terms of average reaction time.  Additionally, it was shown
-through a gamma mixed effects model that there are significant year and race
-effects that greatly impact reaction time.  However, this practically does not make much sense
-as there is no reason for so much random variation without much of a trend.
-There has not been a consistent decrease or increase since 1999, much of the data
-for reaction times has been random and unpredictable.  Thus while it seems easy
-to conclude that Devon Allen was wrongly disqualified, that may not necessarily
-be the case.  If there are issues with reaction time such as athletes approaching the 0.10 
-second barrier again at the next World Track Championship (August 2023), then World Athletics 
-needs to adjust that threshold and allow for faster reaction times so that athletes with superb 
-reflexes are not penalized like how Allen was in 2022.  World Athletics did not
-do anything about \citet{komi2009iaaf}'s study, but perhaps they should now so
-that the story of Devon Allen is not repeated.
-
-This paper does not consider data from the women's 100 meter hurdle, and although
-that was initially because of concerns over reaction time differences in men and
-women, exploratory analysis was performed and showed that 2022 was at best only
-slightly faster relative to other championships going back to 2022.  Likewise,
-the rank based comparisons results in p values that were not significant.  The
-lack of evidence from looking at the women's data makes it harder to conclude 
-that the sensors used at the 2022 championships were faulty, although it is
-unknown if the sensors for the men's 110 meter hurdle and women's 100 meter
-hurdle were the same.
-
-\section{Appendix}
-\label{sec:Appendix}
-%Put downloadable data spreadsheet
-Here is World Athletics website with the data: \url{https://www.worldathletics.org/results/world-athletics-championships}.
-Here is the data for the USA Track and Field Championships: \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
-
-For the pooled data set
+%Table should go to the results section
+All Men's semi finals and finals data
 \begin{center}
@@
   \end{center}
+%Section 4.1: Gamma model results show the fitting with and without 2022,
+%talk about p values and thresholds from the simulations (do not include the
+%results in the methods section). 
+Using the gamma function with link set to log (whose summary statistics can be
+found in the last row of the above table), the probability of the function being
+less than 0.1 seconds can be calculated to see if Allen's disqualification was an anomaly.
+Prior to the 2022 World Championship, there was a 0.0425\% chance (based on the
+simulated results) of observing a reaction time below 0.01 seconds in the men's
+finals.  That means that we would expect a time below the threshold to occur
+once every 2,352 starts.  Considering there are roughly 8 men's finals reaction
+times every 2 years (The World Championship is biennial), every 588 years there
+should be a time as extreme as Allen's.  This seems extremely unlikely, and thus
+it seems like there may be an additional explanation for the fast reaction times.
 
-\bibliographystyle{chicago}
-\bibliography{citations.bib}
+It is interesting that the probability of observing a reaction time below 0.1
+seconds increases drastically for the finals data and increases by a smaller margin
+for the pooled data set after 2022 is included in the data set.  By including 2022
+in the data, the chances of observing an extreme time are increased, showing that
+2022 is a high outlier for reaction times compared to other years.
 
-\end{document}
+It is worth noting that changing the reaction time barrier from 0.1 seconds to
+0.08 seconds drastically reduces the chances of observing a reaction time that
+would break the barrier.  For the dataset of all finals data, the probability
+associated with observing a reaction time below 0.08 seconds was $6.67\cdot10{-7}$.
+If IAAF were to change the reaction time the way that \citet{Komi}'s paper
+suggests, the probability would be so much lower that athletes would likely not
+have a case to make when arguing against a disqualification.
+
+%Discussion should show that random effect
+%and standard error increase and that 2022 is special.  Now talk about Seiko
+%timing equipment
+
+%Citation needed in this section
+Since 1985 Seiko Holding Corporations has served every World Athletics Championship
+as the official timer \citep{Seiko}.  Since 1985, the technology and the ability
+to accurately predict measurements: long jump distances, false starts, total time,
+reaction time, etc. have all dramatically improved.  Seiko did not start tracking
+reaction time as an official measurement until 1999 in the men's and women's 100 meter
+hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
+technology to the World Track and Field Championships, the highest stakes in the world
+of running outside of the Olympics. Seiko's technology for detecting reaction 
+times relies on the pressure that athletes
+exert on the plate when they push off.  Their systems measure the time differential
+between when the "gun" goes off to start the race and when the pressure changes 
+\citep{Seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
+false start as it is considered that no athlete can react so quickly 
+\citep{Seiko-Timing}.  Thus, a reaction time of for example 0.05, suggests that 
+the athlete predicted the gun and it was luck that caused their abnormally high 
+time.  At the World Championship level, World Athletics and Seiko want to remove 
+that element of luck and thus impose the 0.1 second barrier. In 2013, Seiko 
+upgraded their timing equipment for sprinting events (100 meter hurdles falls 
+under this category) \citep{WorldAthletics_2013}.  Since 1999, the highest reaction 
+times in the men's 110-meter hurdle were recorded in 2013.  That is not to say that the higher 
+times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
+to 2022, Seiko again upgraded its technology but for its jump management system.
+
+%Section 4.2 is motivated by seiko timing and look at same athletes across
+%multiple types of competition: 2019 and national
+
+%Discussion/Conclusion is beyond results.  Limitations, what type of data we
+%hope to have.
+
+For the comparison between athletes who competed in 2019 and 2022, the p value
+was 0.001464.  As Seiko timed both events, there should not be a signficant
+difference if the timing equipment in both instances was working correctly.
+Additionally, the p value was 0.000155 for the same athletes competing in the 
+same year  To put it simply, there is not a reasonable explanation for 
+this. It is inconceivable that so many athletes would improve such a significant
+amount in such a small amount of time. 6 countries are represented in this 
+comparison, which drastically decreases the chances of this issue being caused by
+one specific country.  The United States athlete's improvement in times was
+discussed earlier in the paper, but this data shows that on average there was
+%Use a reference in this part
+improvement for British, Polish, French, Brazilian, and Spanish athletes.
+
+%Wrap up and Conclusion
+It is worth noting that in the semi-finals of the World Track and
+Field Championships; Allen's reaction time was 0.101 which is only 0.001 above
+the legal limit.  What this suggests is that Allen may have strong reflexes
+and be able to react extremely well to the sound of the gun.  It seems unlikely that
+Allen would be able to correctly predict the gun with such precision two times in a row,
+which is the exact reason that the 0.10 second reaction time disqualification barrier was
+imposed in the first place.  But Allen showed that at the Championship he was able to
+react extremely fast two times, which may suggest skill rather than luck. Devon 
+Allen is a very good athlete, as evidenced by his football career at the University of Oregon
+and making it onto the Philadelphia Eagles practice squad this past year \citep{Hurley}.
+Thus the 0.099 reaction time may not have been a product of Devon Allen
+predicting the start of the race but rather a combination of a quick reaction 
+and a possibly faulty sensor.
+
+Devon Allen's disqualification at the World Track and Field Championships was
+possibly the result of both a faulty timing equipment and Devon Allen
+reacting extremely quickly to the start.  The code and data from R showed that
+the 2022 World Track and Field Championships were a low outlier compared to
+every other year in terms of average reaction time.  Additionally, it was shown
+through a gamma mixed effects model that there are significant year and batch
+effects that greatly impact reaction time.  However, this practically does not make much sense
+as there is no reason for so much random variation without much of a trend.
+There has not been a consistent decrease or increase since 1999, much of the data
+for reaction times has been random and unpredictable.  Thus while it seems easy
+to conclude that Devon Allen was wrongly disqualified, that may not necessarily
+be the case.  If there are issues with reaction time such as athletes approaching the 0.10 
+second barrier again at the next World Track Championship (August 2023), then World Athletics 
+needs to adjust that threshold and allow for faster reaction times so that athletes with superb 
+reflexes are not penalized like how Allen was in 2022.  World Athletics did not
+do anything about \citet{komi2009iaaf}'s study, but perhaps they should now so
+that the story of Devon Allen is not repeated.
+
+\section{Appendix}
+\label{sec:appendix}
+%Put downloadable data spreadsheet
+Here is World Athletics website with the data: \url{https://www.worldathletics.org/results/world-athletics-championships}.
+Here is the data for the USA Track and Field Championships: \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
 
@@
   \end{center}
+
+\bibliographystyle{chicago}
+\bibliography{citations.bib}
+
+\end{document}

### 2023-04-24 Manuscript/manuscript.tex

--- 2023-04-21 Manuscript/manuscript.tex
@@
 
+\jy{Not a good opening. Could start by telling the story, recapping the major
+  events and discussions on the internet. Then dig into the relevant
+  literature. Finally outline the contributions of this work.
+}
+
+\jy{use the following structure and flow:
+  1) A paragraph to tell the story;
+  2) A paragraph on informal statistical analyses on the internet/social media;
+  3) A paragraph on scholarly articles on reaction time in the literature;
+  4) A paragraph on what's new: methods; results.
+  5) A paragraph of roadmap.}
+
+%1)
 After coming in third in the 110 meter hurdles at the United States Track and 
-Field Championships in Eugene Oregon at the end of June, Devin Allen was expected to
-be a contender for the World Championships, which were scheduled for two months 
-later, also in Eugene. At the World Championships, Allen made it through the 
+Field Championships in Eugene, Oregon at the end of June, Devon Allen was 
+expected to 
+be a contender for the 2022 World Track and Field Championships, which were also 
+scheduled in Eugene 
+two months later. At the World Championships, Allen made it through the 
 heats and semifinals to reach the finals while posting reaction times of 0.123
-in the heats and 0.101 in the semifinal.  Allen, who attended the University
-of Oregon (Location of US World Championship and the World Championship) was
+in the heats \eds{how many heats are there? Likely more than one, so I think you
+should include all of them, or an average} and 0.101 in the semifinal.  
+Allen, who attended the University
+of Oregon (the location of US World Championship \eds{US World?} and the World 
+Championship) was
 considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
@@
 with shock from both Allen and fans alike.  NBC, who was broadcasting the
-championships in the United States, later uploaded a video to their YouTube
-channel that show the disqualification and can be found here:
-\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers when
-they found out that Allen was disqualified, angry at the result.  Had Allen 
+championships in the United States, later uploaded a video to their Youtube
+channel that showed the disqualification and can be found here:
+\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers 
+when they found out that Allen was disqualified, angry at the result.  Had Allen 
 reacted just 0.001 seconds slower he would have been allowed to compete and been 
@@
 
+%2)
 Allen's disqualification has been discussed extensively in online communities
@@
 news section of the website that generated articles such as ``Was Devon Allen
-Screwed? There's At Least a 99\% That He Was" and ``The Data Keeps Pouring In and
+Screwed? There's At Least a 99\% Chance That He Was" and ``The Data Keeps Pouring 
+In and
 It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
-Johnson, who created and runs the Letsrun website argued that there was an error
+Johnson, who created and runs the LetsRun website argued that there was an error
 with the timing equipment at the 2022 Track and Field World Championship and
-cited graphs, statistics such as median reaction time, and compared the reaction
-times at the 2022 Track and Field World Championship \citep{johnson2022devon}.
+cited graphs, simple statistics such as median reaction time, and compared the 
+reaction times at the 2022 Track and Field World Championship. 
+\eds{compared the reaction times at the 2022 Track and Field World Championship 
+to what? Also restate here that no formal statistical analyses have been done.}
+
+%3)
+The International Association of Athletics Federations (also referred to
+as IAAF or World Athletics), which is the governing board for international running
+competitions, commissioned a study in 2009 regarding the reaction time 
+barrier and whether the threshold should be lowered. 
+\eds{verify previous change is correct} 
+The paper analyzed sprinters reaction times and the biological processes 
+that went into reacting to the starting gun and concluded 
+that athletes can react in as low as 0.08 seconds.  
+\eds{what paper are you referring to in the previous sentence?  Please cite.} 
+%One study by Pain looked at
+%cited graphs, statistics such as median reaction time, and compared the reaction
+%times at the 2022 Track and Field World Championship \citep{johnson2022devon}.
 
@@
 reaction times for nine male athletes and analyzed their reaction times.  The
-fastest athlete of the nine, had on average (even with their two fastest time 
+fastest athlete of the nine, had on average (even with their two fastest times 
 removed) a reaction time of 0.087 seconds and a standard deviation of 0.004
 seconds.  Two other athletes were able to react similarly when under certain
+muscular tightening conditions \citep{pain2007sprint}. 
+\eds{It was recommended (?sentence was previously missing the opening words)}
+to IAAF that the disqualification barrier be decreased to allow sprinters to
 muscular tightening conditions.  In a study commissioned by IAAF,
@@
 between 1997 and 2009 for the men's 110 meter hurdles.  Another important
+conclusion was that reaction times did not decrease over the twelve years
+covered by the research in the study \citep{pilianidis2012start}.  This will
+be important when considering the results in section \ref{sec:Results}.
+
+Despite the study, IAAF has not changed the reaction barrier since it was
+implemented at the 1999 World Track and Field Championships.  It is confusing
+as to why IAAF would commission a study only to disregard the results and not
+make the necessary changes.  It is unknown how many athletes since 2009 could
+have potentially benefited from this, but it is probable that there are others
+besides Allen who were disqualified.
+%Reaction time barrier has been examined before: cite reviews
+%Then go into specifics and talk about people who wanted to change barrier
+%General reviews are important
+%People have studied the biological processes, cite a few references
+%How barrier is picked by regulator
+%Read abstract and into and summarize into about one sentence what each paper contributes
+%1 and 2 can be merged but 3 need to stand alone
+
+4)
+One of the first objectives of the paper is to try to model the data, which is
+=======
 conclusion was that for reaction times did not decrease over the twelve years
@@
 this two different ways.  Our first method is to try to model the data, which is
+>>>>>>> fa823ff9cc9ce17ef8e00f9181f5f5f4bf2972aa
 the reaction times of every non-disqualification since the 1999 World Championships. 

### 2023-04-24 Manuscript/manuscript.tex
more work on intro

@@
 Devon Allen was disqualified in the Men's 110 meter hurdle final of the 2022
-Track and Field World Championships after registering a reaction time of 0.099 
+World Track and Field Championships after registering a reaction time of 0.099 
 seconds, 0.001 seconds faster than what is allowed. Following the games, 
-bloggers on the running website LetsRun scrutinized the results %and performed 
-%non-statistical analysis of the data. They 
-and found that the reaction time data from the 2022 World Track Championships 
+bloggers on the running website LetsRun %scrutinized the results %and performed 
+%non-statistical analysis of the data. They found 
+concluded that the reaction time data from the 2022 World Championships 
 seemed to be generally faster compared to the other datasets they looked at, but 
 they did not perform any formal %meaningful 
-statistical analysis of the data. This paper questions the reaction time 
+statistical analyses. This paper questions the reaction time 
 disqualification barrier, which is currently 0.1 seconds, to determine whether 
@@
 %problem that could 
-will likely continue to be an issue in future world championships.
+will continue to be an issue in future world championships.
 
@@
 and replaced with others}
+
 \noindent\sc{Keywords}:
@@
 %1)
-After coming in third in the 110 meter hurdles at the United States Track and 
-Field Championships in Eugene, Oregon at the end of June, Devon Allen was 
-expected to 
-be a contender for the 2022 World Track and Field Championships, which were also 
-scheduled in Eugene 
-two months later. At the World Championships, Allen made it through the 
+Devon Allen, \eds{a student (or an alumnus)?} of the University of Oregon in 
+Eugene, Oregon, was
+expected to be a contender for the 2022 World Track and Field Championships 110 
+meter hurdle.  After running the event in 12.84 seconds in early June 2022 
+(only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
+in third in the event at the United States Track and Field 
+Championships also in Eugene at the end of June, Allen was poised to 
+compete in the 2022 World Track and Field Championships again in Eugene in August.
+%After coming in third in the 110 meter hurdles at the United States Track and 
+%Field Championships also in Eugene at the end of June, Allen was poised to 
+%compete in the 2022 World Track and Field Championships also 
+%located at Eugene in August.
+%two months later. 
+At the World Championships, Allen made it through the 
 heats and semifinals to reach the finals while posting reaction times of 0.123
-in the heats \eds{how many heats are there? Likely more than one, so I think you
-should include all of them, or an average} and 0.101 in the semifinal.  
-Allen, who attended the University
-of Oregon (the location of US World Championship \eds{US World?} and the World 
-Championship) was
-considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
-in early June; only 0.04 seconds away from the world record \citep{wa2022preview}.
-Thus his disqualification by reacting in 0.099 seconds instead of 0.01 was met
+in the heats \eds{if more than one heat, include all of them or the average;
+if only one heat, change to singular form} and 0.101 in the semifinal.  
+%Allen, who attended the University of Oregon also in Eugene, 
+%(the location of US World Championship 
+%\eds{US World? remove World?} and the World Championship) 
+%was
+%considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
+%in early June; only 0.04 seconds away from the world record \citep{wa2022preview}.
+%Thus 
+His disqualification by reacting in 0.099 seconds instead of 0.1 was met
 with shock from both Allen and fans alike.  NBC, who was broadcasting the
-championships in the United States, later uploaded a video to their Youtube
-channel that showed the disqualification and can be found here:
+championships in the United States, later uploaded a video to their YouTube
+channel that showed the disqualification; it can be found here:
 \url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers 
@@
 every sprinter who steps onto the track.
+\eds{final two sentence seems out of place here; consider moving elsewhere?}
 
@@
 which functions similarly to Reddit is very active during important running
-events, which include during the finals of major events such as
+events, including during the finals of major events such as
 the Olympics and the World Track and Field Championships.  However, it is the
@@
 cited graphs, simple statistics such as median reaction time, and compared the 
-reaction times at the 2022 Track and Field World Championship. 
-\eds{compared the reaction times at the 2022 Track and Field World Championship 
+reaction times at the 2022 World Track and Field Championship 
+\citep{johnson2022devon}.
+\eds{compared the reaction times at the 2022 World Track and Field  Championship 
 to what? Also restate here that no formal statistical analyses have been done.}
@@
 %3)
-The International Association of Athletics Federations (also referred to
-as IAAF or World Athletics), which is the governing board for international running
-competitions, commissioned a study in 2009 regarding the reaction time 
-barrier and whether the threshold should be lowered. 
-\eds{verify previous change is correct} 
-The paper analyzed sprinters reaction times and the biological processes 
-that went into reacting to the starting gun and concluded 
-that athletes can react in as low as 0.08 seconds.  
-\eds{what paper are you referring to in the previous sentence?  Please cite.} 
+%The International Association of Athletics Federations (also referred to
+%as IAAF or World Athletics), which is the governing board for international running
+%competitions, commissioned a study in 2009 regarding the reaction time 
+%barrier and whether the threshold should be lowered. 
+%The paper analyzed sprinters reaction times and the biological processes 
+%that went into reacting to the starting gun and concluded 
+%that athletes can react in as low as 0.08 seconds.  
 %One study by Pain looked at
-%cited graphs, statistics such as median reaction time, and compared the reaction
-%times at the 2022 Track and Field World Championship \citep{johnson2022devon}.
-
 Reaction time in sprint and hurdle events has been studied by many authors.
-\citet{pain2007sprint} looked at
-reaction times for nine male athletes and analyzed their reaction times.  The
-fastest athlete of the nine, had on average (even with their two fastest times 
-removed) a reaction time of 0.087 seconds and a standard deviation of 0.004
+\citet{pain2007sprint} examined
+reaction times for nine male athletes.  The
+fastest athlete of the nine had, on average (even with their two fastest times 
+removed), a reaction time of 0.087 seconds and a standard deviation of 0.004
 seconds.  Two other athletes were able to react similarly when under certain
 muscular tightening conditions \citep{pain2007sprint}. 
-\eds{It was recommended (?sentence was previously missing the opening words)}
-to IAAF that the disqualification barrier be decreased to allow sprinters to
-muscular tightening conditions.  In a study commissioned by IAAF,
-\jy{define acronym at first occurrence}
-\citep{komi2009iaaf} recommended that the disqualification 
+In a study commissioned by International Association of Athletics Federations 
+(IAAF), \citet{komi2009iaaf} recommended that the disqualification 
 barrier be decreased to allow sprinters to
@@
 change the reaction time criteria from pushing off on the block to body
-movement. This could provide a more accurate representation of 
-reaction time, and thus the possibility of a false disqualification would be lowered.
+movement. This could provide a more accurate representation of reaction time, 
+and thus the possibility of a false disqualification would be lowered.
 The possibility of using a video-based medium has been proved to be a reliable
@@
 conclusion was that reaction times did not decrease over the twelve years
-covered by the research in the study \citep{pilianidis2012start}.  This will
-be important when considering the results in section \ref{sec:Results}.
+covered by the research in the study \citep{pilianidis2012start}. This contradicts
+what was found in a 2021 study that concluded the year of competition has a
+significant effect on reaction times \citep{zhang2021correlation}.  Their
+analysis looked at data from the 2011-2019 World Championships across multiple
+sprint events and concluded that from 2011 to 2019, reaction times got
+faster. However, not all studies have been in favor of a decrease in the
+reaction time barrier. \citet*{brosnan2017effects} argue for increasing 
+the reaction time barrier to 0.115 seconds, and cited evidence from European 
+Championships and World Championships in their article.  A study looking 
+at the 2008 Olympics likewise supported the conclusion that athletes could not
+react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
+concluded that visual reaction times to the starting gun may be faster by 0.007
+seconds than the IAAF method, which is from force on block starts 
+\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
+only by 0.001 seconds. It is also worth noting that one study looking at data 
+from various IAAF meets found that the fastest reaction times for men were found
+between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
+because Allen was 27 years old at the 2022 Track and Field World Championship 
+and in his physical peak.
 
-Despite the study, IAAF has not changed the reaction barrier since it was
+%This will
+%be important when considering the results in section \ref{sec:Results}.
+
+Despite their study, IAAF has not changed the reaction barrier since it was
 implemented at the 1999 World Track and Field Championships.  It is confusing
@@
 
-4)
-One of the first objectives of the paper is to try to model the data, which is
-=======
-conclusion was that for reaction times did not decrease over the twelve years
-covered by the research in the study \citep{pilianidis2012start}.  This contradicts
-what was found in a 2021 study that concluded the year of competition has a
-significant effect on reaction times \citep{zhang2021correlation}.  Their
-analysis looked at data from the 2011-2019 World Championships across multiple
-sprint events and concluded that from 2011 to 2019, reaction times got
-faster. However, not  all studies have been in favor of a decrease in the
-reaction time barrier. \citet*{brosnan2017effects} in 2016 argue for increasing 
-the reaction time barrier to 0.115 seconds, and cited evidence from European 
-Championships and World Championships in their article.  A study looking 
-at the 2008 Olympics likewise supported the conclusion that athletes could not
-react in as little as 0.1 seconds \citep{lipps2011implications}.  Another study
-concluded that visual reaction times to the starting gun may be faster by 0.007
-seconds than the IAAF method, which is from force on block starts 
-\citep{holmes2018method}.  While this may seem trivial, Allen was disqualified
-only by 0.01 seconds. It is also worth noting that one study looking at data 
-from various IAAF meets found that the fastest reaction times for men were found
-between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
-because Allen was 27 years old at the 2022 Track and Field World Championship 
-and in his physical peak.
-
+%4)
 \jy{language needs to be formalized and condensed.}
@@
 this two different ways.  Our first method is to try to model the data, which is
->>>>>>> fa823ff9cc9ce17ef8e00f9181f5f5f4bf2972aa
 the reaction times of every non-disqualification since the 1999 World Championships. 
@@
 accounting for the venue (or year) effect. 
-The second method is to compare data for athletes who have competed
-at additional competitions to see how their reaction times compare from previous
-competitions to this one. First we can look at the reaction times of athletes who
-competed at the World Track and Field Championship in both 2019 and 2022 and
-compare the times, and secondly we can look at how athletes performed compared
-to how they did at their national/qualifiers meet. In order to reduce any chances
-of athletes improving reaction time, we can also look at data from qualifying meets.
-Almost every athlete had to qualify by proving they are elite at the national level, 
-usually by coming in at least fourth. Thus, data was compiled on athletes who competed
-at the 2022 World Track and Field Championships on what these athletes also ran 
-at their country's national championship. If there was any argument that athletes
-have improved their reaction time from 2019 to 2022, it should now be considered
-nearly null, as the national competitions took place between June and July 2022 
-and the World Championships were in August. The equipment used to measure reaction
-time cannot be guaranteed to be the same from national to international
-competition, but the time span between meets has been reduced considerably.
+The second method is to compare reaction times for athletes who have competed
+at additional competitions to the 2022 World Track and Field Championship. 
+%\eds{I recommend moving the rest of this paragraph to Section 2.2 as I don't
+%think it is critical for readers to know all the details here.}
+%First we can compare the reaction times of athletes who
+%competed at the World Track and Field Championship in both 2019 and 2022, and 
+%second we can compare how athletes performed at the 2022 championship relative 
+%to their national/qualifiers meet. In order to reduce any chances
+%of athletes improving reaction time, we can also look at data from qualifying meets.
+%Almost every athlete had to qualify by proving they are elite at the national level, 
+%usually by coming in at least fourth. Thus, data was compiled on athletes who competed
+%at the 2022 World Track and Field Championships on what these athletes also ran 
+%at their country's national championship. If there was any argument that athletes
+%have improved their reaction time from 2019 to 2022, it should now be considered
+%nearly null, as the national competitions took place between June and July 2022 
+%and the World Championships were in August. The equipment used to measure reaction
+%time cannot be guaranteed to be the same from national to international
+%competition, but the time span between meets has been reduced considerably.
 
-The rest of this paper is as follows: Section~\ref{sec:Data} describes how data was
-collected and begins to detail the generalized linear mixed-effect (GLMM)
-model which is further developed in the
-Section~\ref{sec:Methods}.  The signed-rank test analysis for the same athletes is
-also included in Section~\ref{sec:Methods}.  Section~\ref{sec:Results}
-section provides the probabilities from the methods section and the paper ends 
-by questioning the reaction time barrier.  Lastly, section~\ref{sec:Discussion}
+The rest of this paper is organized as follows. Section~\ref{sec:Data} describes 
+how data was collected and begins to detail the generalized linear mixed-effect 
+(GLMM) model that is further developed in the Section~\ref{sec:Methods}. The 
+signed-rank test that analyzes reaction times for the same athletes is
+also described in Section~\ref{sec:Methods}.  Section~\ref{sec:Results}
+ provides the results of the two analyses and ends 
+by questioning the reaction time barrier.  Lastly, Section~\ref{sec:Discussion}
 discusses the impact and limitations of the paper.
+\eds{capitalize first letter when referring to a numbered section, table, figure, 
+etc. ``Section 4 provides...'' vs. ``The next section provides...''}
 
@@
 
-There are two types of data that is explored throughout this paper: data from
+There are two types of data that are explored throughout this paper: data from
 1999-2022 for every world championship, and data from athletes who competed
-at both the 2019 and 2022 World Championships and 2022 national competitions
-and the 2022 World Championship.  The first data set is used for the GLMM portion
-of the paper, while the second data set is used for the rank based comparisons.
+the 2022 World Championships and another competition (either 2019 World 
+Championship or a 2022 national competition held within two months of the  
+2022 World Championships. The first data set is used for the GLMM portion
+of the paper, while the second data set is used for the rank-based comparisons.
 
@@
 \jy{You have two such datasets in the analysis. Use one paragraph for each.}
+
+\eds{unless target journal has specified otherwise, Table captions should go
+above the table. (Figure captions go below the figure)}
 

### 2023-04-25 Manuscript/manuscript.tex
Started to make changes

@@
 \newcommand{\eds}[1]{\textcolor{red}{EDS: (#1)}}
+\newcommand{\of}[1]{\textcolor{violet}{OF: #1}}
 
@@
 %1)
-Devon Allen, \eds{a student (or an alumnus)?} of the University of Oregon in 
+Devon Allen, an alumnus of the University of Oregon in 
 Eugene, Oregon, was
@@
 in the heats \eds{if more than one heat, include all of them or the average;
-if only one heat, change to singular form} and 0.101 in the semifinal.  
+if only one heat, change to singular form}
+\of{There are roughly 5 heats, but each athlete only competes in one heat} 
+and 0.101 in the semifinal.  
 %Allen, who attended the University of Oregon also in Eugene, 
@@
 spared waiting another two years until the next World Track and Field Championship.
-An investigation into reaction times is also important because reaction times have
-been shown to affect the overall time of the race \citep{delalija2008reaction}.
-Thus, faster reaction times lead to faster overall times; which is the goal of
-every sprinter who steps onto the track.
+
 \eds{final two sentence seems out of place here; consider moving elsewhere?}
+\of{Moved to the end of paragraph 4 of introduction}
 
@@
 World Track and Field Championship to check for any abnormalities, and we approach
-this two different ways.  Our first method is to try to model the data, which is
+this two different ways.  First we try to model the data, which is
 the reaction times of every non-disqualification since the 1999 World Championships. 
@@
 at additional competitions to the 2022 World Track and Field Championship. 
+Our investigation into reaction times is important because reaction times have
+been shown to affect the overall time of the race \citep{delalija2008reaction}.
+Thus, faster reaction times lead to faster overall times; which is the goal of
+every sprinter who steps onto the track.
 %\eds{I recommend moving the rest of this paragraph to Section 2.2 as I don't

### 2023-04-25 Manuscript/manuscript.tex
Went over Prof. Schifano's comments

@@
 the Olympics and the World Track and Field Championships.  However, it is the
-\jy{cite these blog articles formally}
-news section of the website that generated articles such as ``Was Devon Allen
-Screwed? There's At Least a 99\% Chance That He Was" and ``The Data Keeps Pouring 
-In and
-It Continues to Look Bad For World Athletics and Great For Devon Allen".  Robert
-Johnson, who created and runs the LetsRun website argued that there was an error
+news section of the website that generated articles such as \citet{johnson2022was}
+and \citet{johnson2022data} from  Robert Johnson, who created and runs 
+the LetsRun website.  He argued that there was an error
 with the timing equipment at the 2022 Track and Field World Championship and
 cited graphs, simple statistics such as median reaction time, and compared the 
-reaction times at the 2022 World Track and Field Championship 
-\citep{johnson2022devon}.
+reaction times at the 2022 World Track and Field Championship to other competitions
+such as the USA Track and Field Championships and other World Championships.
+However, their analysis lacked formal analysis. The alleged 99.95\% chance Allen
+was wrong came from that the lowest average reaction time for the men's 100 meter,
+men's 110 meter hurdle, and women's 100 meters from 2003 to 2022 all occurred at
+the 2022 World Championships and Johnson claimed that the probability of this
+happening was $\frac{1}{13} \cdot \frac{1}{13} \cdot \frac{1}{13} = 0.05\%$
+\citep{johnson2022was}.
 \eds{compared the reaction times at the 2022 World Track and Field  Championship 
 to what? Also restate here that no formal statistical analyses have been done.}
+\of{Cleaned this up a little but still a little messy.}
 
@@
 accounting for the venue (or year) effect. 
-The second method is to compare reaction times for athletes who have competed
-at additional competitions to the 2022 World Track and Field Championship. 
+Then we compare reaction times for athletes who have competed
+at additional competitions to the 2022 World Track and Field Championship.  By
+comparing the same athlete across multiple competitions we are able to see if the
+2022 data differed significantly from other competitions.
 Our investigation into reaction times is important because reaction times have
@@
 how data was collected and begins to detail the generalized linear mixed-effect 
-(GLMM) model that is further developed in the Section~\ref{sec:Methods}. The 
+(GLMM) model that is further developed in Section~\ref{sec:Methods}. The 
 signed-rank test that analyzes reaction times for the same athletes is
@@
 Championship or a 2022 national competition held within two months of the  
-2022 World Championships. The first data set is used for the GLMM portion
+2022 World Championships). The first data set is used for the GLMM portion
 of the paper, while the second data set is used for the rank-based comparisons.
@@
 disqualified more because of faulty equipment rather than because he reacted
-\jy{cross-reference to the table by its label}
 too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction times for
@@
 \begin{center}
+  \caption{Rection times for four United States Athletes at the 2022 USA Track
+  and Field World Championships and the 2022 World Track and Field Championship.
+  ``N/A" was used for any athlete that did not compete and thus did not register 
+  a reaction time.}
   \begin{tabular}{c c c c c c c} 
@@
   \label{fig:USAvsWorld}
-  \caption{Rection times for four United States Athletes at the 2022 USA Track
-  and Field World Championships and the 2022 World Track and Field Championship.
-  ``N/A" was used for any athlete that did not compete and thus did not register 
-  a reaction time.}
+
   \end{center}
@@
 \jy{You have two such datasets in the analysis. Use one paragraph for each.}
+\of{Split it into 2 sections, do you stil want me to do something else?}
 
@@
 above the table. (Figure captions go below the figure)}
+\of{Moved table caption to be above}
 
@@
 \citet{datta2005rank} proposed an extension of the Wilcoxon rank-sum test to
-clustered data with subnit-level grouping. The test is designed based on a
+clustered data with subunit-level grouping. The test is designed based on a
 within-cluster resampling principle. Consider randomly picking one observation
@@
 available from the \texttt{clusWilcox.test()} function with
-\texttt{method = 'ds'} from R package \texttt{clusrank}
+\texttt{method = ``ds'} from R package \texttt{clusrank}
 \citep{jiang2017wilcoxon}.
@@
 %% two are correlated \citep{delalija2008reaction}.
+%% Group is either designated to be 1 or 2, with 1 corresponding to either the
+%% national level data or the previous world championship data and 2 corresponding
+%% to the 2022 data.
 
@@
 \jy{motivate first why you want to run the analysis with and without the 2022 data.}
+
+\of{Add something like this to the start of the paragraph?}
+We may want to consider using all four data sets for this problem because
+comparing all the data to the data excluding 2022 can give us a qualitative
+assessment of how impactful 2022 was.
 
@@
   \centering
+  \caption{Summary statistics for the GLMM on various data sets.}
   \begin{tabular}{c c c c c c}
@@
   \end{tabular}
-  \caption{Summary statistics for the GLMM on various data sets}
   \label{tab:Gamma_parameters}
@@
   \centering
+  \caption{Probabilities of observing reaction times less than 0.1, 0.09, and
+  0.08 for various data sets.}
   \begin{tabular}{c c c c} 
@@
   \end{tabular}
-  \caption{Probabilities of observing reaction times less than 0.1, 0.09, and
-  0.08 for various data sets}
   \label{tab:Sim_probability}
@@
   two data sets? use one paragraph for each.}
+\of{Shouldnt describing the grouping variables go more in the data or methods 
+section?  Currently, there is some writing about this but it is commented out. It
+can be found in section 3.2}
 

### 2023-04-25 Manuscript/manuscript.tex
comments through Section 2

@@
 (only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
-in third in the event at the United States Track and Field 
-Championships also in Eugene at the end of June, Allen was poised to 
-compete in the 2022 World Track and Field Championships again in Eugene in August.
+in third at the United States Track and Field 
+Championships also in Eugene in late June, Allen was poised to compete in the 
+2022 World Track and Field Championships at the same Eugene venue in August.
 %After coming in third in the 110 meter hurdles at the United States Track and 
@@
 %two months later. 
-At the World Championships, Allen made it through the 
-heats and semifinals to reach the finals while posting reaction times of 0.123
-in the heats \eds{if more than one heat, include all of them or the average;
-if only one heat, change to singular form}
-\of{There are roughly 5 heats, but each athlete only competes in one heat} 
-and 0.101 in the semifinal.  
+At the World Championships, Allen made it through the preliminary
+heats and semifinals to reach the final heat while posting reaction times of 
+0.123 in his preliminary heat and 0.101 in his semifinal.  
 %Allen, who attended the University of Oregon also in Eugene, 
@@
 %Thus 
-His disqualification by reacting in 0.099 seconds instead of 0.1 was met
+His disqualification in the final heat by reacting in 0.099 seconds instead of 
+0.1 was met
 with shock from both Allen and fans alike.  NBC, who was broadcasting the
@@
 
-\eds{final two sentence seems out of place here; consider moving elsewhere?}
-\of{Moved to the end of paragraph 4 of introduction}
+%\eds{final two sentence seems out of place here; consider moving elsewhere?}
+%\of{Moved to the end of paragraph 4 of introduction}
 
@@
 news. The message board,
-which functions similarly to Reddit is very active during important running
+which functions similarly to Reddit, is very active during important running
 events, including during the finals of major events such as
 the Olympics and the World Track and Field Championships.  However, it is the
-news section of the website that generated articles such as \citet{johnson2022was}
-and \citet{johnson2022data} from  Robert Johnson, who created and runs 
+news section of the website that generated articles such as \citet{johnson2022data}
+and \citet{johnson2022was} from  Robert Johnson, who created and runs 
 the LetsRun website.  He argued that there was an error
 with the timing equipment at the 2022 Track and Field World Championship and
-cited graphs, simple statistics such as median reaction time, and compared the 
-reaction times at the 2022 World Track and Field Championship to other competitions
+cited graphs, computed simple statistics such as median reaction times, and 
+compared the reaction times at the 2022 World Track and Field Championship to 
+other competitions
 such as the USA Track and Field Championships and other World Championships.
-However, their analysis lacked formal analysis. The alleged 99.95\% chance Allen
-was wrong came from that the lowest average reaction time for the men's 100 meter,
-men's 110 meter hurdle, and women's 100 meters from 2003 to 2022 all occurred at
-the 2022 World Championships and Johnson claimed that the probability of this
+The comparisons lacked formal statistical analysis, however. 
+The alleged 99.95\% chance Allen was wrong \eds{can we use different word 
+instead of wrong?  at fault?  However, should that be 0.05\% chance at fault?}
+came from the fact that the lowest average reaction times for the men's 100 meter,
+men's 110 meter hurdle, and women's 100 meter competitions held between 2003 and 
+2022 each \eds{?} occurred at the 2022 World 
+Championships.  Johnson claimed that the probability of this
 happening was $\frac{1}{13} \cdot \frac{1}{13} \cdot \frac{1}{13} = 0.05\%$
 \citep{johnson2022was}.
-\eds{compared the reaction times at the 2022 World Track and Field  Championship 
-to what? Also restate here that no formal statistical analyses have been done.}
-\of{Cleaned this up a little but still a little messy.}
+%\eds{compared the reaction times at the 2022 World Track and Field  Championship 
+%to what? Also restate here that no formal statistical analyses have been done.}
+%\of{Cleaned this up a little but still a little messy.}
 
@@
 conclusion was that reaction times did not decrease over the twelve years
-covered by the research in the study \citep{pilianidis2012start}. This contradicts
-what was found in a 2021 study that concluded the year of competition has a
-significant effect on reaction times \citep{zhang2021correlation}.  Their
+covered by the research in the study \citep{pilianidis2012start}. This 
+contradicts the results of \citet{zhang2021correlation}, which indicated that 
+year of competition has a significant effect on reaction times.  Their
 analysis looked at data from the 2011-2019 World Championships across multiple
@@
 because Allen was 27 years old at the 2022 Track and Field World Championship 
-and in his physical peak.
+and at his physical peak.
 
@@
 \jy{language needs to be formalized and condensed.}
-The objective to this paper is to analyze the reaction times at the 2022
-World Track and Field Championship to check for any abnormalities, and we approach
-this two different ways.  First we try to model the data, which is
-the reaction times of every non-disqualification since the 1999 World Championships. 
-The model we use is a gamma linear mixed-effects model with a random effect
-accounting for the venue (or year) effect. 
-Then we compare reaction times for athletes who have competed
+The objective to this paper is to analyze the reaction times at the 2022 World 
+Track and Field Championship to check for any abnormalities, and we approach
+this two different ways.  First we model the data, which is
+the reaction times of every non-disqualification since the 1999 World 
+Championships, using a gamma linear mixed-effects model with a random effect 
+accounting for the venue (or year). \eds{are you using covariates?  Data section
+implies that you are, but Results section does not} 
+Based on the fitted model, we can then estimate the probability of a reaction 
+time being below a certain threshold (e.g., 0.1). 
+Second, we compare reaction times for athletes who have competed
 at additional competitions to the 2022 World Track and Field Championship.  By
-comparing the same athlete across multiple competitions we are able to see if the
-2022 data differed significantly from other competitions.
-Our investigation into reaction times is important because reaction times have
+comparing reaction times from the same athlete across multiple competitions we 
+are able to determine if the 2022 data differed significantly from other 
+competitions.
+Our investigation into reaction times is important not only because if its
+implications for future races, but also because reaction times have
 been shown to affect the overall time of the race \citep{delalija2008reaction}.
@@
 (GLMM) model that is further developed in Section~\ref{sec:Methods}. The 
-signed-rank test that analyzes reaction times for the same athletes is
+signed-rank test that analyzes reaction times for the same athletes across
+different competitions is
 also described in Section~\ref{sec:Methods}.  Section~\ref{sec:Results}
@@
 There are two types of data that are explored throughout this paper: data from
-1999-2022 for every world championship, and data from athletes who competed
+1999-2022 for every world championship, and data from athletes who competed in
 the 2022 World Championships and another competition (either 2019 World 
@@
 
-Data was taken from the World Athletics and covers the men's 110 meter hurdles 
-and the women's 100 meter hurdles from 1999 to 2022.  The variable of prediction 
-interest is reaction time, and the predicting variables are year, stage of 
-competition (heats, semifinal, final), total time, and gender.  
+Data was taken from the World Athletics 
+\eds{is there a website you can cite here?}
+and covers the men's 110 meter hurdles 
+and the women's 100 meter hurdles from 1999 to 2022.  The dependent
+variable interest is reaction time, and the independent variables are year, 
+stage of competition, total time, and gender.  \eds{see earlier comment - 
+are you using any of these as covariates?}
 The data can be found in the appendix~\ref{sec:Appendix}. Factors such as name, 
@@
 purposes of this research besides the times for Devon Allen and the other 
-United States athletes mentioned above.  The variable ``Stage"
-refers to whether the observation occurred during the ``heats" (Preliminary round) which
-is denoted by ``H" in the data, ``S" for ``semifinals" (Of which there are either 2 or 3 
-heats each year), and ``F" for finals. The variable ``Gender" has levels ``M" and ``F" 
-which stand for ``male" and ``female" respectively.  The variable ``Batch" takes a 
+United States athletes mentioned above. \eds{not sure which other US athletes 
+you are referring to in the previous sentence} 
+The variable ``Stage" refers to whether the observation occurred during a 
+%heat (preliminary round)  
+preliminary heat denoted by ``H", semifinal heat (of which there are either two
+or three each year) denoted by ``S", or final heat denoted by ``F". The variable 
+``Gender" has levels ``M" and ``F" for ``male" and ``female" respectively.  The
+variable ``Batch" takes a 
 numerical value and every heat of every round from 1999 to 2022 was assigned it's
@@
 correlated with other runners reacting quickly (one runner reacting quickly may
-result in other runners reacting quickly).  There is no statistical significance to
+result in other runners reacting quickly). There is no statistical meaning to
 how the batches were labeled, they were ordered chronologically by meet but
-reverse chronologically by year (Batch 1 is a preliminary Heat of 2022 and Batch
-111 corresponds to the 1999 finals). Included are
-box plots of the reaction times of the men's finals and the men's semifinals and
-finals pooled together.  We will refer to the latter as the ``pooled data"
-throughout the paper.  The heats or preliminary rounds were not studied because
-athletes react faster as the competition progresses.  Athletes react faster in
-the finals than the semifinals which are faster than the heats 
-\citep{zhang2021correlation}.  If we want to determine if Allen's reaction time
-is reasonable, then we thus need to look at the men's final data, with the men's
-pooled data being included in order to increase the sample size.  In 2022, there
+reverse chronologically by year (Batch 1 is a preliminary heat of 2022 and Batch
+111 corresponds to the 1999 finals). %Included are
+%box plots of the reaction times of the men's finals and the men's semifinals and
+%finals pooled together.  
+For analysis purposes, we will often pool reaction times of men's final and 
+men's semifinal heats together and refer to this data as the ``pooled data"
+%We will refer to the latter as the ``pooled data"
+throughout the paper.  %The heats or preliminary rounds were not studied because
+The preliminary heats were not studied because athletes react faster as the 
+competition progresses \citep[e.g.,][]{zhang2021correlation}. % Athletes react 
+%faster in the finals than the semifinals which are faster than the heats 
+%\citep{zhang2021correlation}.  
+\eds{if you are really not using ``H'' data at all, it does not appear necessary
+to spend time defining earlier.  You could just say directly when you define 
+stage that you only consider semifinals and finals and cite the Zhang paper for 
+rationale}
+If we want to determine if Allen's reaction time
+was reasonable, then we thus need to look at the men's final data, with the men's
+pooled data being included in order to increase the sample size.  
+\eds{previous sentence is not clear; pooled data already includes final data}
+In 2022, there
 were only five data points due to two disqualifications and one athlete who 
 qualified but did not compete.
-The data that looked at in the paper is best visualized using a box plot as show 
-in Figure~\ref{fig:Boxplots}.  It is clear that the 2022 box plot for both the
-men's finals and semifinals data is lower (indicating a faster reaction time)
+The data under consideration is best visualized using side-by-side box plots, as
+shown in Figure~\ref{fig:Boxplots}.  It is clear that the 2022 box plot for both 
+the men's finals and semifinals data is lower (indicating faster reaction times)
 than in previous years.  The median for the men's pooled data in 2022 was 0.129
 and the median for the men's final data in 2022 was 0.126.  For comparison,
-Brosnan found reaction times of 0.156 seconds for men when looking at data from
-1999 to 2014 \citep{brosnan2017effects}.
+\citet{brosnan2017effects} found reaction times of 0.156 seconds \eds{mean?  median?} for men when looking at data from 1999 to 2014.
 
+\eds{I wonder if this paragraph would be better placed in the introduction?}
 Something that is extremely important to note in the data is the inconsistency 
@@
 World Track and Field Champions) instituted a rule change that allowed one 
-warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}  This 
+warning false start before a sprinter was disqualified 
+\citep{iaaf2009falsestart}.  This 
 essentially gave sprinters a second chance and reduced the penalty for a false 
 start.  There were 18 male and 7 female false starts at the 2007 World 
-Championships, 18 male and 7 female starts at the 2009 World Championships. 
+Championships, and 18 male and 7 female starts at the 2009 World Championships. 
 Starting in 2011, the rule was scrapped, and the old policy which automatically 
-disqualifies runners who false start was reinstituted.  The effect? At the 2011 
-World Championship there were 6 male disqualifications and 4 female 
-disqualifications \citep{iaaf2009falsestart}. By returning to the harsher policy and 
-cracking down on false starts, World Athletics had reduced the number of 
-false starts by two thirds. One study that looked at different false start
-rules that IAAF has imposed from 1997 to 2009 and found that the reaction 
-times when there was less leniency improved by a statistically significant amount
-\citep{haugen2013effect}. This was desirable for World Athletics as false starts can 
-make already long track meets more tedious, both for athletes and viewers watching the 
-television broadcast.  Since 2011, IAAF has instituted the one false start rule
+disqualifies runners who false start was reinstated. By returning to the harsher 
+policy and cracking down on false starts, World Athletics had reduced the number 
+of false starts by two thirds \eds{for men, not overall} in 2011 (6 male and 4 female disqualifications 
+\citep{iaaf2009falsestart}).
+%The effect? At the 2011 
+%World Championship there were 6 male disqualifications and 4 female 
+%disqualifications \citep{iaaf2009falsestart}.  
+%One study that looked at different false start rules that IAAF has
+%imposed from 1997 to 2009 and found that the reaction times when there was less 
+%leniency improved by a statistically significant amount
+%\citep{haugen2013effect}. 
+\citet{haugen2013effect} examined the effect of different false start
+rules that IAAF has imposed from 1997 to 2009 and found statistically 
+significant improvement in reaction times under harsher rules.
+\eds{is that right?  it seems counterintuitive to me}
+This was desirable for World Athletics as false starts can make already long 
+track meets more tedious, both for athletes and viewers watching the 
+television broadcast.  
+\eds{I am not sure the previous sentence follows from the one preceding it. Why
+ does World Athletics find faster reaction times desireable?  Do you mean that
+they find fewer false starts desirable?}
+Since 2011, IAAF has instituted the one false start rule
 \citep{iaaf2009falsestart}. However, we chose not to include women's data in this
@@
 women's reaction times: \citep{lipps2011implications, babicc2009reaction,
-panoutsakopoulos2020gender}
+panoutsakopoulos2020gender}  \eds{Two comments about the previous sentence: 1. 
+it does not logically follow the previous sentence as there is no mention of 
+gender, and 2. if you are not using women's data, don't spend time talking about 
+it at the beginning of the data section.  Similar to my comment about 
+preliminary heats, you can move this sentence to the beginning of the Data 
+section to provide a rationale for why you do not include data for women.}
 
 \subsection{Beyond World Championship}
-This section was motivated from a qualitative analysis that started by looking
-at how United States athletes reacted at the 2022 United States Track and Field
-Championships. From June 23-26 2022, USA held its Track
-and Field Championship to decide who to send to the World Championships at Hayward 
-Field, the exact same venue as used for the World Track and Field Championship a month 
-later.  Thus, a baseline is able to be established for the four USA athletes who 
+The motivation for looking beyond World Championship data came from an 
+initial qualitative analysis examining how United States athletes reacted at the 
+2022 United States Track and Field Championships.
+\eds{be consistent with how you refer to the US championship:
+``United States Track and Field Championships" and ``USA Track and Field 
+Championships" are both being used; please pick one and use that throughout}
+%This section was motivated from a qualitative analysis that started by looking
+%at how United States athletes reacted at the 2022 United States Track and Field
+%Championships. 
+From June 23-26 2022, the United States held its Track
+and Field Championship at Hayward 
+Field in Eugene, Oregon to decide who to send to the World Championships 
+being held at the same venue in August. %the exact same venue as used 
+%for the World Track and Field Championship a month later.  
+%Thus, a baseline is able to be established for the four USA athletes who 
+Thus, we can establish a baseline for the four USA athletes who 
 competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
@@
 Field Championship compared to the USA Track and Field Championship as were his
-teammates.  This comparison suggests that the timing system used between the two
+teammates. \eds{statistically significantly faster? If you used a statistical 
+test, it should not be described in this section.  If you didn't use a 
+statistical test, you should replace the word `significantly' with something 
+else (substantially?)} 
+This comparison suggests that the timing system used between the two
 meets may have been meaningfully different to the point where Devon Allen was
 disqualified more because of faulty equipment rather than because he reacted
-too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction times for
-Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. ``USA" denotes
-the USA Track and Field Championships, ``World" denotes the World Track and Field
-Championships,   All numbers listed in the table are the reaction
-time for the respective athletes in seconds.
+too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction 
+times for Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. 
+%``USA" denotes
+%the USA Track and Field Championships, ``World" denotes the World Track and Field
+%Championships. All numbers listed in the table are the reaction
+%time for the respective athletes in seconds.
 
@@
 \begin{center}
-  \caption{Rection times for four United States Athletes at the 2022 USA Track
-  and Field World Championships and the 2022 World Track and Field Championship.
+  \caption{Reaction times (in seconds) for four United States Athletes at the 
+	2022 USA Track
+  and Field World Championships (``USA") and the 2022 World Track and Field 
+	Championships (``World"); H denotes preliminary heat, S denotes semifinal, and 
+	F denotes final.
   ``N/A" was used for any athlete that did not compete and thus did not register 
-  a reaction time.}
+  a reaction time. }
   \begin{tabular}{c c c c c c c} 
    \toprule
-   Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
+	 & \multicolumn{3}{c}{USA} & \multicolumn{3}{c}{World} \\
+	\cmidrule(lr){2-4}
+    \cmidrule(lr){5-7}
+   Athlete &  H &  S &  F &  H &  S &  F \\ [0.5ex] 
    \midrule
@@
 
-\eds{unless target journal has specified otherwise, Table captions should go
-above the table. (Figure captions go below the figure)}
-\of{Moved table caption to be above}
-
-This dataset is quite small to perform any formal statistical analysis on, but
-the data was expanded to include data for athletes who competed at national
-competitions (many of which were held from May-July of 2022).  Once the data
-was collected, each athlete was designated to be a cluster and the stage of
-competition (national or international) was specified to be the group.
+This dataset is too small to perform any meaningful statistical analysis, but
+the data was expanded to include reaction times for athletes who competed at 
+national competitions (many of which were held from May-July of 2022). Once the 
+data was collected, each athlete was designated as a cluster and the stage of
+competition (national or international) was specified for each competition 
+within the cluster (athlete).
 
@@
 
-The World Championship data and the data beyond the World Champisonship were
+The World Championship data and the data beyond the World Championship were
 analysed with GLMM and rank-based approaches, respectively.
+\eds{need consistency with ``Championship" or ``Championships - I am not sure 
+which one is correct}
 

### 2023-04-25 Manuscript/manuscript.tex
Updates for tuesday meeting

--- 2023-04-25 Manuscript/manuscript.tex
@@
 \begin{center}
-  \caption{Rection times for four United States Athletes at the 2022 USA Track
+  \caption{Reaction times for four United States Athletes at the 2022 USA Track
   and Field World Championships and the 2022 World Track and Field Championship.
@@
 \jy{You have two such datasets in the analysis. Use one paragraph for each.}
-\of{Split it into 2 sections, do you stil want me to do something else?}
+\of{Split it into 2 sections, do you still want me to do something else?}
 
@@
 random numbers generated from the fitted mixture distribution can be used to
-approimate the probability of observing a reaction time faster than any given
+approximate the probability of observing a reaction time faster than any given
 threshold.  We want the probability of a reaction time being less than 0.1

### 2023-04-26 Manuscript/manuscript.tex
Update after tuesdays meeting
finished first pass of manuscript

@@
 spared waiting another two years until the next World Track and Field Championship.
-
+\eds{you mention in the final section that the next world championship is in 
+2023, so he won't need to wait two years if that is true}
 %\eds{final two sentence seems out of place here; consider moving elsewhere?}
@@
 
-\subsection{World Championship}
+\subsection{World Championship}\label{sec:dataworld}
 
@@
 Data was taken from the World Athletics 
-\eds{is there a website you can cite here?}
+%\eds{is there a website you can cite here?}
 and covers the men's 110 meter hurdles 
-and the women's 100 meter hurdles from 1999 to 2022.  The dependent
-variable interest is reaction time, and the independent variables are year, 
-stage of competition, total time, and gender.  \eds{see earlier comment - 
+and the women's 100 meter hurdles from 1999 to 2022.  \eds{see earlier comment - 
 are you using any of these as covariates?}
-The data can be found in the appendix~\ref{sec:Appendix}. Factors such as name, 
+The data source can be found in the Appendix.%~\ref{sec:Appendix}. 
+Factors such as name, 
 country, and identification number were not considered important for the 
@@
 
-\subsection{Beyond World Championship}
+\subsection{Beyond World Championship}\label{sec:databeyond}
 The motivation for looking beyond World Championship data came from an 
-initial qualitative analysis examining how United States athletes reacted at the 
+initial qualitative \eds{exploratory instead of qualitative?}
+analysis examining how United States athletes reacted at the 
 2022 United States Track and Field Championships.
@@
 competition (national or international) was specified for each competition 
-within the cluster (athlete).
+within the cluster (athlete).  \eds{this is where you can put the relevant 
+information about the other competitions that was commented out in the 
+introduction.}
+\eds{Can you show some descriptive statistics for this data?  Total number of 
+clusters? Average number of international vs national competitions across 
+athletes/clusters?}
 
@@
 
-The World Championship data and the data beyond the World Championship were
-analysed with GLMM and rank-based approaches, respectively.
+The %World Championship data and the data eyond the World Championship 
+data described in Sections~\ref{sec:dataworld} and~\ref{sec:databeyond} were
+analysed with GLMM and rank-based methods, respectively.
 \eds{need consistency with ``Championship" or ``Championships - I am not sure 
@@
 Based on an exploratory analysis using the GLMM model for the reaction times
-from the World Championship, the final model that we chose was a gamma
-mixed-effect model. As suggested by the log-likelihood and the Akaike Information
+from the World Championship, the final model we chose was a gamma mixed-effect 
+model. As suggested by the log-likelihood and the Akaike Information
 Criterion (AIC), the normal error model on the log-scale was found to be
@@
 Let $Y_{ij}$ be the reaction time of observation~$j$ in year~$i$.
+\eds{we know that many of the athletes compete in multiple world championships,
+how is this athlete clustering being accounted for here?}
 Conditional on a venue effect $z_i$ of year~i,  $Y_{ij}$ is modeled by 
-gamma distribution. In the classic generlized linear model (GLM) sense,
+gamma distribution. In the classic generalized linear model (GLM) sense,
 $Y_{ij}$ has conditional mean
@@
 ensure the positivity of the conditional mean.
+\eds{Did you try an identity link also? Lo and Andrews (2015) suggest that the 
+Gamma family with an identity link works well for their reaction-time data.
+It doesn't seem like a good idea in general because it is likely that some 
+negative predicted responses will occur either during model fitting or in the 
+final model, but the internet seems to favor this link so a reviewer may ask 
+about it too.}
 The gamma mixed-effect model can be fit with the \texttt{glmer()} function in R
@@
 variance as a function of the mean $\mu$ in the GLM sense is
-$v(\mu) = \mu^2 / a$, which implies that the dispersion parameter
+$v(\mu) = \mu^2 / a$, which implies that the dispersion parameter is
 $\phi = 1 / a$. Therefore, for the gamma mixed-effect model, the conditional
@@
 approximate the probability of observing a reaction time faster than any given
-threshold.  We want the probability of a reaction time being less than 0.1
-seconds in order to gauge if that is a reasonable disqualification barrier.
-
-As for the random effect, it was possible to include as well a race effect.
-From our exploration, however, the race effect is very
-difficult to explain or interpret as it means that some heats of races were either
+threshold.  We are specifically interested in the probability of a reaction time
+ being less than 0.1 seconds in order to gauge if that is a reasonable 
+disqualification barrier.
+
+As for the random effect, it was possible to include a race 
+\eds{heat? when I first read this, I thought you meant race as in ethnicity} 
+effect, as well.
+From our exploration, however, the race effect is very difficult to explain or 
+interpret as it means that some heats of races were either
 faster or slower and that everyone in the race reacted similarly.  Although
@@
 heats that had fast and slow reaction times are all within 0.12 seconds of each
-other.  It is much easier however to explain the venue effect.  The technology
+other.  It is much easier, however, to explain the venue effect.  The technology
 that World Athletics has used has differed since 1999 and theoretically the
@@
 can be from either the 2022 World Championship or otherwise.
-Let $X_{ij}$ be the $j$th reaction time of athelete~$i$, $i = 1, \ldots, n$,
+Let $X_{ij}$ be the $j$th reaction time of athlete~$i$, $i = 1, \ldots, n$,
 $j = 1, \ldots, m_i$ where $m_i$ is the number of observations from
 athelete~$i$. Let $\delta_{ij}$ be the group indicator of $X_{ij}$; $\delta_{ij}
-= 1$ if $X_{IJ}$ is in group~1 and $\delta_{ij} = 0$ otherwise. Atheletes are
-assumed to be independent, while subunit observations from the same athelete are
+= 1$ if $X_{ij}$ is in group~1 (2022 World Championship) and $\delta_{ij} = 0$ 
+otherwise. Athletes are
+assumed to be independent, while subunit observations from the same athlete are
 not. The null hypothesis $H_0$ to be tested is that there is no difference
@@
 \]
-where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the psudo-sample.
+where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the pseudo-sample.
 The test statistic $S$ is the average of $W^*$ averaged over all possible
@@
 \citep{jiang2017wilcoxon}.
+\eds{are you including preliminary heat data as possible within cluster 
+observations in this analysis?  Based on your 
+discussion in Section 2.1 about how they are not comparable, it may not be fair 
+to include them here?  Or perhaps do the analysis with and without prelim heat
+data to see if it makes a difference?}
+
 %% In this case, athletes who have more
@@
 assessment of how impactful 2022 was.
+\eds{Agree with JY, but you need to describe what these four datasets are.}
+
+Table~\ref{tab:Gamma_parameters} summarizes the results from the gamma model 
+from different data sets.
 
@@
   \centering
-  \caption{Summary statistics for the GLMM on various data sets.}
+  \caption{Summary statistics for the GLMM on various data sets. 
+	\eds{Int=intercept?}}
   \begin{tabular}{c c c c c c}
@@
 
-At a size of $n=10,000,000$, the calculated probability of a reaction
-time being below 0.1 seconds was as follows for the various different data sets:
+\eds{be consistent with the order of the rows between 
+Tables~\ref{tab:Gamma_parameters} and~\ref{tab:Sim_probability}}
+
 \jy{label the table and reference to it; clean it up}
@@
 
-Using the gamma function with link set to log, the probability of the function being
-less than 0.1 seconds can be calculated to see if Allen's disqualification was an anomaly.
-Prior to the 2022 World Championship, there was a 0.0439\% chance (based on the
+Using the gamma function with link set to log, the probability of the reaction 
+time being less than 0.1 seconds can be calculated to see if Allen's 
+disqualification was an anomaly. \eds{explain how simulation was used here.}
+At a size of $n=10,000,000$ \eds{use different letter as $n$ was already used 
+for number of athletes, $B?$}, the calculated probabilities of a reaction time 
+being below 0.1 seconds for the various different data sets are shown in 
+Table~\ref{tab:Sim_probability}. Prior to the 2022 World Championship, there was 
+a 0.0439\% chance (based on the
 simulated results) of observing a reaction time below 0.01 seconds in the men's
-finals as shown in table~\ref{tab:Sim_probability}.
+finals.
 That means that we would expect a time below the threshold to occur
 once every 2,277 starts.  Considering there are roughly 8 men's finals reaction
-times every 2 years (The World Championship is biennial), every 569 years there
+times every 2 years (the World Championship is biennial), every 569 years there
 should be a time as extreme as Allen's. This seems extremely unlikely, and thus
@@
 
-It is interesting that the probability of observing a reaction time below 0.1
-seconds increases drastically for the finals data and increases by a smaller margin
-for the pooled data set after 2022 is included in the data set.  By including 2022
-in the data, the chances of observing an extreme time are increased, showing that
-2022 is a high outlier for reaction times compared to other years.
-
-Table ~\ref{tab:Sim_probability} shows that changing the reaction time barrier from 0.1 seconds to
-0.08 seconds drastically reduces the chances of observing a reaction time that
-would break the barrier.  For the dataset of all finals data, the probability
+When the 2022 data is included in the analysis, it is interesting that the 
+probability of observing a reaction time below 0.1 seconds increases drastically 
+for the finals data and increases by a smaller margin for the pooled data.  By
+including 2022 in the data, the chances of observing an extreme time are 
+increased, suggesting that 2022 is a high outlier for reaction times compared to
+ other years.
+
+Table ~\ref{tab:Sim_probability} also shows that changing the reaction time 
+barrier from 0.1 seconds to 0.08 seconds drastically reduces the chances of 
+observing a reaction time that would break the barrier.  For the dataset of all 
+finals data, the probability
 associated with observing a reaction time below 0.08 seconds was $6\cdot10^{-7}$.
-If IAAF were to change the reaction time the way that \citet{komi2009iaaf}'s paper
+If IAAF \eds{or World Athletics?} were to change the reaction time the way that 
+\citet{komi2009iaaf}
 suggests, the probability would be so much lower that athletes would likely not
@@
 %Citation needed in this section
-Since 1985 Seiko Holding Corporations has served every World Athletics Championship
-as the official timer \citep{wa2022seiko}.  Since 1985, the technology and the ability
-to accurately predict measurements: long jump distances, false starts, total time,
-reaction time, etc. have all dramatically improved.  Seiko did not start tracking
-reaction time as an official measurement until 1999 in the men's and women's 100 meter
-hurdles.  Seiko regularly updates their equipment so that they provide cutting edge
-technology to the World Track and Field Championships, the highest stakes in the world
-of running outside of the Olympics. Seiko's technology for detecting reaction 
-times relies on the pressure that athletes
-exert on the plate when they push off.  Their systems measure the time differential
+Since 1985 Seiko Holding Corporations has served every World Athletics 
+Championship as the official timer \citep{wa2022seiko}.  Since 1985, the 
+technology and the ability
+to accurately predict measurements: long jump distances, false starts, total 
+time, reaction time, etc. have all dramatically improved.  Seiko did not start 
+tracking reaction time as an official measurement until 1999 in the men's and 
+women's 100 meter \eds{110 for men?} hurdles. 
+Seiko regularly updates their equipment so that they provide cutting edge
+technology to the World Track and Field Championships, the highest stakes in the 
+world of running outside of the Olympics. Seiko's technology for detecting 
+reaction times relies on the pressure that athletes exert on the plate when they 
+push off.  Their systems measure the time differential
 between when the ``gun" goes off to start the race and when the pressure changes 
-\citep{wa2022seiko}  If an athlete has a reaction time under 0.10 seconds, it is deemed a 
+\citep{wa2022seiko}.  If an athlete has a reaction time under 0.1 seconds, it 
+is deemed a 
 false start as it is considered that no athlete can react so quickly 
@@
 upgraded their timing equipment for sprinting events (100 meter hurdles falls 
-under this category) \citep{wa2013backtage}.  Since 1999, the highest reaction 
-times in the men's 110-meter hurdle were recorded in 2013.  That is not to say that the higher 
-times were caused by Seiko's equipment, but the two may be related.  It is worth noting that prior
+under this category) \citep{wa2013backtage} \eds{110 for men?}. Since 1999, the
+ highest \eds{highest as in fastest or slowest? Neither seems consistent with 
+the boxplots in Figure 1} reaction 
+times in the men's 110-meter hurdle were recorded in 2013.  That is not to say
+ that the higher 
+times were caused by Seiko's equipment, but the two may be related.  It is worth 
+noting that prior
 to 2022, Seiko again upgraded its technology but for its jump management system.
+\eds{I would argue that this paragraph should go in the discussion as there are 
+no results presented here}
 
@@
 When the clusrank rank sum test was performed, the ``ds" method produced a z score of
-2.9751 and p value of 0.001464.  This value is for a one-sided test and show that
-the mean reaction time on average in 2019 was larger than the mean reaction time
+2.9751 and p value of 0.001464.  \eds{what is the ds method?} 
+This value is for a one-sided test and show that
+the mean reaction time on average in 2019 was larger (slower) than the mean 
+reaction time
 in 2022 for athletes who competed in both championships at an alpha level of 0.01.
 This provides good evidence that athletes who competed in both 2019 and 2022 got
-faster during that timespan.
+faster during that timespan.  \eds{That is not the only explanation}
 
@@
 again performing a clustered analysis for a Wilcoxon ranked sum test
-results in two new z scores and p values.  The ``ds" method produced a z score of
+results in two new z scores and p values. 
+\eds{I am not sure why you are repeating the data collection steps and
+again performing a clustered analysis here.  Also, if there are two new z-scores
+and p values, why are you only reporting one?}
+  The ``ds" method produced a z score of
 3.6069 and p value of 0.000155. This result is highly significant and shows that
@@
 
+\eds{also consider moving the following paragraph to discussion??}
 As Seiko timed both events, there should not be a significant
 difference if the timing equipment in both instances was working correctly.
+\eds{you just mentioned in the last subsection that Seiko upgraded its 
+technology prior to 2022}
   To put it simply, there is not a reasonable explanation for 
 this. It is inconceivable that so many athletes would improve such a significant
-amount in such a small amount of time. 6 countries are represented in this 
+amount in such a small amount of time. Six countries are represented in this 
 comparison, which drastically decreases the chances of this issue being caused by
-one specific country.  The United States athlete's improvement in times was
+one specific country.  The United States athletes' improvement in times was
 discussed earlier in the paper, but this data shows that on average there was
@@
 
-One conclusion from \citet{zhang2021correlation}'s analysis was that athletes
+One conclusion from the analysis in \citet{zhang2021correlation} was that 
+athletes'
 reaction times improved from 2011 to 2019.  However this study does not consider 
 data from 1999 to 2009, and over this timespan athletes get slower and then
-faster.  A boxplot~\ref{fig:Boxplots} shows that for the men's pooled data, the 
+faster.  Figure~\ref{fig:Boxplots} shows that for the men's pooled data, the 
 year with the fastest median reaction time before 2022 was 1999.  Thus, this 
-reject's Zhang's conclusion that athletes are getting faster because they were 
+reject's Zhang et al.'s conclusion that athletes are getting faster because they 
+were 
 fast over twenty years ago. One possible explanation for why 2011 had such
@@
 
-It is worth noting that in the semi-finals of the World Track and
-Field Championships; Allen's reaction time was 0.101 which is only 0.001 above
+It is worth noting that in the semifinals of the 2022 World Track and
+Field Championships, Allen's reaction time was 0.101 which is only 0.001 above
 the legal limit. What this suggests is that Allen may have strong reflexes
-and be able to react extremely well to the sound of the gun.  It seems unlikely that
-Allen would be able to correctly predict the gun with such precision two times in a row,
-which is the exact reason that the 0.10 second reaction time disqualification barrier was
-imposed in the first place.  But Allen showed that at the Championship he was able to
+and be able to react extremely well to the sound of the gun. It seems unlikely 
+that Allen would be able to correctly predict the gun with such precision two 
+times in a row, which is the exact reason that the 0.1 second reaction time 
+disqualification barrier was
+imposed in the first place. Allen showed at the Championship he was able to
 react extremely fast two times, which may suggest skill rather than luck. Devon 
-Allen is a very good athlete, as evidenced by his football career at the University of Oregon
+Allen is a very good athlete, as evidenced by his football career at the 
+University of Oregon
 and making it onto the Philadelphia Eagles practice squad this past year 
@@
 possibly the result of both a faulty timing equipment and Devon Allen
-reacting extremely quickly to the start.  The code and data from R showed that
+reacting extremely quickly to the start.  Our analysis showed that
 the 2022 World Track and Field Championships were a low outlier compared to
@@
 through a gamma mixed effects model that there are significant year and race
-effects that greatly impact reaction time.  However, this practically does not make much sense
+effects \eds{remove race effects?} that greatly impact reaction time.  
+However, this practically does not make much sense
 as there is no reason for so much random variation without much of a trend.
-There has not been a consistent decrease or increase since 1999, much of the data
-for reaction times has been random and unpredictable.  Thus while it seems easy
-to conclude that Devon Allen was wrongly disqualified, that may not necessarily
-be the case.  If there are issues with reaction time such as athletes approaching the 0.10 
-second barrier again at the next World Track Championship (August 2023), then World Athletics 
-needs to adjust that threshold and allow for faster reaction times so that athletes with superb 
-reflexes are not penalized like how Allen was in 2022.  World Athletics did not
-do anything about \citet{komi2009iaaf}'s study, but perhaps they should now so
-that the story of Devon Allen is not repeated.
+There has not been a consistent decrease or increase since 1999, and much of the 
+data
+for reaction times has been random and unpredictable. 
+\eds{I would suggest removing the last two sentences, as they would suggest your
+ modeling with GLMM is not good}.
 
@@
 that was initially because of concerns over reaction time differences in men and
-women, exploratory analysis was performed and showed that 2022 was at best only
-slightly faster relative to other championships going back to 2022.  Likewise,
-the rank based comparisons results in p values that were not significant.  The
-lack of evidence from looking at the women's data makes it harder to conclude 
+women, exploratory analysis showed that 2022 was at best only
+slightly faster relative to other championships going back to 2022 
+\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
+the intro, unless he only considered sprint and not hurdles?} Likewise,
+the rank-based comparisons resulted in p values that were not significant.  The
+lack of evidence from the women's data makes it harder to conclude 
 that the sensors used at the 2022 championships were faulty, although it is
@@
 
-\section{Appendix}
+Thus while it seems easy
+to conclude that Devon Allen was wrongly disqualified, that may not necessarily
+be the case. 
+If there are issues with reaction time such as athletes approaching 
+the 0.1 
+second barrier again at the next World Track Championship (August 2023)
+\eds{check against introduction}, then World Athletics 
+needs to adjust that threshold and allow for faster reaction times so that 
+athletes with superb 
+reflexes are not penalized. %like how Allen was in 2022.  
+World Athletics did not
+do anything in response to the \citet{komi2009iaaf} study, but perhaps they 
+should now so
+that the story of Devon Allen is not repeated.
+
+\section*{Appendix}
 \label{sec:Appendix}
@@
 \jy{You have countries beyond USA}
-Here is the data for the USA Track and Field Championships: \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
+The USA Track and Field Championships data is available at: 
+\url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
 

### 2023-04-27 Manuscript/manuscript.tex
Went over some of Prof Schifano's notes

@@
 reacted just 0.001 seconds slower he would have been allowed to compete and been 
-spared waiting another two years until the next World Track and Field Championship.
+spared waiting until the next World Track and Field Championship.
 \eds{you mention in the final section that the next world championship is in 
 2023, so he won't need to wait two years if that is true}
+\of{Resolved}
 %\eds{final two sentence seems out of place here; consider moving elsewhere?}
@@
 other competitions
-such as the USA Track and Field Championships and other World Championships.
+such as the United States Track and Field Championships and other World Championships.
 The comparisons lacked formal statistical analysis, however. 
-The alleged 99.95\% chance Allen was wrong \eds{can we use different word 
+The alleged 99.95\% chance Allen was wronged \eds{can we use different word 
 instead of wrong?  at fault?  However, should that be 0.05\% chance at fault?}
-came from the fact that the lowest average reaction times for the men's 100 meter,
-men's 110 meter hurdle, and women's 100 meter competitions held between 2003 and 
-2022 each \eds{?} occurred at the 2022 World 
-Championships.  Johnson claimed that the probability of this
-happening was $\frac{1}{13} \cdot \frac{1}{13} \cdot \frac{1}{13} = 0.05\%$
+\of{Changed to "wronged"}
+came from Johnson claiming there is a $(1/13)^3$ that the fastest reaction
+times came during the same championship for the men's 100 meter,
+men's 110 meter hurdle, and women's 100 meter for 13 competitions held between 2003 and 
+2022 \eds{?} \of{Cleaned up but may need additional work}
 \citep{johnson2022was}.
@@
 implies that you are, but Results section does not}
+\of{Cleaned up confusion about covariates, but no none were used in the paper} 
 Based on the fitted model, we can then estimate the probability of a reaction 
@@
 etc. ``Section 4 provides...'' vs. ``The next section provides...''}
+\of{I think capitalization should be all set: all are in form of ``Section i''}
 
@@
 %\eds{is there a website you can cite here?}
-and covers the men's 110 meter hurdles 
-and the women's 100 meter hurdles from 1999 to 2022.  \eds{see earlier comment - 
+and covers the men's 110 meter hurdles from 1999 to 2022; and can be found in
+the Appendix~\ref{sec:Appendix}.   \eds{see earlier comment - 
 are you using any of these as covariates?}
-The data source can be found in the Appendix.%~\ref{sec:Appendix}. 
-Factors such as name, 
-country, and identification number were not considered important for the 
-purposes of this research besides the times for Devon Allen and the other 
-United States athletes mentioned above. \eds{not sure which other US athletes 
-you are referring to in the previous sentence} 
+\of{removed covariate confusion, talk about womens a couple sentences down} 
+We chose not to include women's data in this
+paper due to the numerous papers citing difference in reaction times in men and
+women's reaction times: \citep{lipps2011implications, babicc2009reaction,
+panoutsakopoulos2020gender}
 The variable ``Stage" refers to whether the observation occurred during a 
 %heat (preliminary round)  
-preliminary heat denoted by ``H", semifinal heat (of which there are either two
-or three each year) denoted by ``S", or final heat denoted by ``F". The variable 
+semifinal heat (of which there are either two
+or three each year) denoted by ``S", or final heat denoted by ``F".
+The preliminary heats were not studied because athletes react faster as the 
+competition progresses \citep[e.g.,][]{zhang2021correlation}. The variable 
 ``Gender" has levels ``M" and ``F" for ``male" and ``female" respectively.  The
@@
 throughout the paper.  %The heats or preliminary rounds were not studied because
-The preliminary heats were not studied because athletes react faster as the 
-competition progresses \citep[e.g.,][]{zhang2021correlation}. % Athletes react 
+ % Athletes react 
 %faster in the finals than the semifinals which are faster than the heats 
@@
 rationale}
-If we want to determine if Allen's reaction time
-was reasonable, then we thus need to look at the men's final data, with the men's
-pooled data being included in order to increase the sample size.  
-\eds{previous sentence is not clear; pooled data already includes final data}
-In 2022, there
+\of{Ok I updated this.  Should I delete the batch variable too?  Although I don't
+use it, I do reference it in the methods section}
+The pooled data increases the sample size which is advantageous because in some
+years there are very few finals observations.  For example, in 2022 there
 were only five data points due to two disqualifications and one athlete who 
-qualified but did not compete.
+did not compete.
 The data under consideration is best visualized using side-by-side box plots, as
@@
 and the median for the men's final data in 2022 was 0.126.  For comparison,
-\citet{brosnan2017effects} found reaction times of 0.156 seconds \eds{mean?  median?} for men when looking at data from 1999 to 2014.
+\citet{brosnan2017effects} found median reaction times of 0.156 seconds \eds{mean?  median?} 
+\of{Resolved}
+for men when looking at data from 1999 to 2014.
 
 \eds{I wonder if this paragraph would be better placed in the introduction?}
+\of{I think it could fit in the brief section between the start of the Data section
+and the GLMM subsection in the paragraph starting with "There are two types..."}
 Something that is extremely important to note in the data is the inconsistency 
@@
 World Track and Field Champions) instituted a rule change that allowed one 
-warning false start before a sprinter was disqualified 
-\citep{iaaf2009falsestart}.  This 
-essentially gave sprinters a second chance and reduced the penalty for a false 
-start.  There were 18 male and 7 female false starts at the 2007 World 
+warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}. 
+This was desirable for World Athletics as false starts can make already long 
+track meets more tedious, both for athletes and viewers watching the television 
+broadcast.  There were 18 male and 7 female false starts at the 2007 World 
 Championships, and 18 male and 7 female starts at the 2009 World Championships. 
@@
 disqualifies runners who false start was reinstated. By returning to the harsher 
-policy and cracking down on false starts, World Athletics had reduced the number 
-of false starts by two thirds \eds{for men, not overall} in 2011 (6 male and 4 female disqualifications 
-\citep{iaaf2009falsestart}).
+policy and cracking down on false starts, World Athletics had reduced men's 
+false starts by two thirds \eds{for men, not overall} \of{Better?} in 2011 (6 male and 4 female disqualifications) 
+\citep{iaaf2009falsestart}.
 %The effect? At the 2011 
@@
 rules that IAAF has imposed from 1997 to 2009 and found statistically 
-significant improvement in reaction times under harsher rules.
+significant improvement in reaction times under more lenient rules.
 \eds{is that right?  it seems counterintuitive to me}
-This was desirable for World Athletics as false starts can make already long 
-track meets more tedious, both for athletes and viewers watching the 
-television broadcast.  
+\of{Yes you are correct, I updated the language.}
+
 \eds{I am not sure the previous sentence follows from the one preceding it. Why
- does World Athletics find faster reaction times desireable?  Do you mean that
+ does World Athletics find faster reaction times desirable?  Do you mean that
 they find fewer false starts desirable?}
-Since 2011, IAAF has instituted the one false start rule
-\citep{iaaf2009falsestart}. However, we chose not to include women's data in this
-paper due to the numerous papers citing difference in reaction times in men and
-women's reaction times: \citep{lipps2011implications, babicc2009reaction,
-panoutsakopoulos2020gender}  \eds{Two comments about the previous sentence: 1. 
+
+\eds{Two comments about the previous sentence: 1. 
 it does not logically follow the previous sentence as there is no mention of 
@@
 section to provide a rationale for why you do not include data for women.} 
+\of{Updates: I moved any talk about women's data to start of Data section, I think
+it flows much better now}
 
@@
 The motivation for looking beyond World Championship data came from an 
-initial qualitative \eds{exploratory instead of qualitative?}
+exploratory \eds{exploratory instead of qualitative?} \of{I agree; fixed}
 analysis examining how United States athletes reacted at the 
@@
 \eds{be consistent with how you refer to the US championship:
-``United States Track and Field Championships" and ``USA Track and Field 
+``United States Track and Field Championships" and ``United States Track and Field 
 Championships" are both being used; please pick one and use that throughout}
+\of{Resolved; changed all USA to United States.}
 %This section was motivated from a qualitative analysis that started by looking
@@
 %for the World Track and Field Championship a month later.  
-%Thus, a baseline is able to be established for the four USA athletes who 
-Thus, we can establish a baseline for the four USA athletes who 
+%Thus, a baseline is able to be established for the four United States athletes who 
+Thus, we can establish a baseline for the four United States athletes who 
 competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
 Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
-Track and Field Championship races compared to the USA Track and Field 
-Championship races. Devon Allen raced three times at the USA Track and
+Track and Field Championship races compared to the United States Track and Field 
+Championship races. Devon Allen raced three times at the United States Track and
 Field Championship and three times at the World Track and Field Championship. 
-His reaction times were significantly faster in the World Track and
-Field Championship compared to the USA Track and Field Championship as were his
+His reaction times were substantially faster in the World Track and
+Field Championship compared to the United States Track and Field Championship as were his
 teammates. \eds{statistically significantly faster? If you used a statistical 
@@
 statistical test, you should replace the word `significantly' with something 
-else (substantially?)} 
+else (substantially?)} \of{Replaced signficantly with substantially}
 This comparison suggests that the timing system used between the two
@@
 times for Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts. 
-%``USA" denotes
-%the USA Track and Field Championships, ``World" denotes the World Track and Field
+%``United States" denotes
+%the United States Track and Field Championships, ``World" denotes the World Track and Field
 %Championships. All numbers listed in the table are the reaction
@@
   \caption{Reaction times (in seconds) for four United States Athletes at the 
-	2022 USA Track
-  and Field World Championships (``USA") and the 2022 World Track and Field 
+	2022 United States Track
+  and Field World Championships (``United States") and the 2022 World Track and Field 
 	Championships (``World"); H denotes preliminary heat, S denotes semifinal, and 
@@
    \toprule
-	 & \multicolumn{3}{c}{USA} & \multicolumn{3}{c}{World} \\
+	 & \multicolumn{3}{c}{United States} & \multicolumn{3}{c}{World} \\
 	\cmidrule(lr){2-4}
@@
 \end{table}
+
+\of{Im not sure how to fix this but the spacing between the lines of the
+table is different than the rest of the paper}
 
@@
 distribution of $Y_{ij}$ is a scale-mixture of gamma distributions, which can be
-easily simulated from once the parameters are estimated. A large number of
+easily simulated from once the parameters are estimated. Many
 random numbers generated from the fitted mixture distribution can be used to
@@
 
-As for the random effect, it was possible to include a race 
+As for the random effect, it was possible to include a heat 
 \eds{heat? when I first read this, I thought you meant race as in ethnicity}
+\of{Yes I think heat or batch effect works better, especially now that I have
+removed most talk about heat from the data section} 
 effect, as well.
@@
 that element of luck and thus impose the 0.1 second barrier. In 2013, Seiko 
-upgraded their timing equipment for sprinting events (100 meter hurdles falls 
-under this category) \citep{wa2013backtage} \eds{110 for men?}. Since 1999, the
+upgraded their timing equipment for sprinting events (110 meter hurdles falls 
+under this category) \citep{wa2013backtage} \eds{110 for men?} \of{Fixed}. Since 1999, the
  highest \eds{highest as in fastest or slowest? Neither seems consistent with 
@@
 
-\jy{You have countries beyond USA}
-The USA Track and Field Championships data is available at: 
+\jy{You have countries beyond United States}
+The United States Track and Field Championships data is available at: 
 \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
@@
    \hline
-   Athlete & USA H & USA S & USA F & World H & World S & World F \\ [0.5ex] 
+   Athlete & United States H & United States S & United States F & World H & World S & World F \\ [0.5ex] 
    \hline\hline

### 2023-04-27 Manuscript/manuscript.tex
removed or commented out fixed comments

@@
 
-\title{On Devon Allen's Disqualification at the 2022 Track and Field World
-Championship}
+\title{On Devon Allen's Disqualification at the 2022 World Track and Field 
+Championships}
 
@@
 
-\jy{Avoid references in abstract.}
+%\jy{Avoid references in abstract.}
 
@@
 %\bigskip
-\eds{removed Devon Allen from keywords because his name is already in title
-and replaced with others}
+%\eds{removed Devon Allen from keywords because his name is already in title
+%and replaced with others}
+
+\eds{also removed 2022 World Track and Field Championships from keywords and 
+changed location of World in the title. World Track and Field Championships
+has more Google search results than Track and Field World Championships, so
+I opted to go with this. Also used plural Championships. Feel free to switch 
+order or use singular if one or both of these is 
+preferred for some reason.  }
 
 \noindent\sc{Keywords}:
-False Start, Hurdles, Reaction Time, Seiko, 2022 World Track and Field Championships
+False Start, Hurdles, Reaction Time, Seiko
 
@@
 reacted just 0.001 seconds slower he would have been allowed to compete and been 
-spared waiting until the next World Track and Field Championship.
-\eds{you mention in the final section that the next world championship is in 
-2023, so he won't need to wait two years if that is true}
-\of{Resolved}
-%\eds{final two sentence seems out of place here; consider moving elsewhere?}
-%\of{Moved to the end of paragraph 4 of introduction}
+spared waiting until the next World Track and Field Championships.
 
@@
 the LetsRun website.  He argued that there was an error
-with the timing equipment at the 2022 Track and Field World Championship and
+with the timing equipment at the 2022 World Track and Field Championship and
 cited graphs, computed simple statistics such as median reaction times, and 
@@
 The comparisons lacked formal statistical analysis, however. 
-The alleged 99.95\% chance Allen was wronged \eds{can we use different word 
-instead of wrong?  at fault?  However, should that be 0.05\% chance at fault?}
-\of{Changed to "wronged"}
-came from Johnson claiming there is a $(1/13)^3$ that the fastest reaction
-times came during the same championship for the men's 100 meter,
-men's 110 meter hurdle, and women's 100 meter for 13 competitions held between 2003 and 
-2022 \eds{?} \of{Cleaned up but may need additional work}
+The alleged 99.95\% chance Allen was wronged 
+came from Johnson claiming there is a $(1/13)^3$ chance that the fastest 
+reaction times over the past 13 competitions of the men's 100 meter, men's 110 
+meter hurdle, and women's 100 meter would all come from the same championship 
 \citep{johnson2022was}.
-%\eds{compared the reaction times at the 2022 World Track and Field  Championship 
-%to what? Also restate here that no formal statistical analyses have been done.}
-%\of{Cleaned this up a little but still a little messy.}
+%came from Johnson claiming there is a $(1/13)^3$ chance that the fastest 
+%reaction times came during the same championship for the men's 100 meter,
+%men's 110 meter hurdle, and women's 100 meter for 13 competitions held between 2003 and 2022 
+%\of{Cleaned up but may need additional work}
+\of{Cleaned this up a little but still a little messy.}
+\eds{is this better? (and still correct?)}
 
@@
 between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
-because Allen was 27 years old at the 2022 Track and Field World Championship 
+because Allen was 27 years old at the 2022 World Track and Field Championships 
 and at his physical peak.
@@
 Championships, using a gamma linear mixed-effects model with a random effect 
-accounting for the venue (or year). \eds{are you using covariates?  Data section
-implies that you are, but Results section does not}
-\of{Cleaned up confusion about covariates, but no none were used in the paper} 
+accounting for the venue (or year). 
+%\eds{are you using covariates?  Data section
+%implies that you are, but Results section does not}
+%\of{Cleaned up confusion about covariates, but no none were used in the paper} 
 Based on the fitted model, we can then estimate the probability of a reaction 
@@
 discusses the impact and limitations of the paper.
-\eds{capitalize first letter when referring to a numbered section, table, figure, 
-etc. ``Section 4 provides...'' vs. ``The next section provides...''}
-\of{I think capitalization should be all set: all are in form of ``Section i''}
 
@@
 the 2022 World Championships and another competition (either 2019 World 
-Championship or a 2022 national competition held within two months of the  
-2022 World Championships). The first data set is used for the GLMM portion
-of the paper, while the second data set is used for the rank-based comparisons.
-
-\subsection{World Championship}\label{sec:dataworld}
+Championships or a 2022 national competition typically held within three months 
+of the 2022 World Championships). 
+\eds{is this true? are you still using the 2019 world championships data for the
+rank-based analyses, or only the national competitions?}
+The first data set is used for the GLMM analyses described in 
+Section~\ref{sec:glmm}, while the second collection of data is used for the 
+rank-based comparisons described in Section~\ref{sec:rank}.
+
+\subsection{World Championships}\label{sec:dataworld}
 
@@
   \includegraphics{Finals_Pooled_Boxplot}
-  \caption{The reaction times from 2019 to 2022 for the men's final (top) and
+  \caption{The reaction times from 1999 to 2022 for the men's final (top) and
   the men's semifinals and finals (bottom).}
@@
 %\eds{is there a website you can cite here?}
-and covers the men's 110 meter hurdles from 1999 to 2022; and can be found in
-the Appendix~\ref{sec:Appendix}.   \eds{see earlier comment - 
-are you using any of these as covariates?}
-\of{removed covariate confusion, talk about womens a couple sentences down} 
-We chose not to include women's data in this
-paper due to the numerous papers citing difference in reaction times in men and
-women's reaction times: \citep{lipps2011implications, babicc2009reaction,
-panoutsakopoulos2020gender}
-The variable ``Stage" refers to whether the observation occurred during a 
+and covers the men's 110 meter hurdles from 1999 to 2022; the source can be 
+found in the Appendix.%~\ref{sec:Appendix}.   %\eds{see earlier comment - 
+%are you using any of these as covariates?}
+%\of{removed covariate confusion, talk about womens a couple sentences down} 
+Although women's data is available, we chose not to include women's data in this
+analyses to follow due to the numerous papers citing differences in men and
+women's reaction times \citep{lipps2011implications, babicc2009reaction,
+panoutsakopoulos2020gender}.
+%The variable ``Stage" refers to whether the observation occurred during a 
 %heat (preliminary round)  
-semifinal heat (of which there are either two
-or three each year) denoted by ``S", or final heat denoted by ``F".
-The preliminary heats were not studied because athletes react faster as the 
-competition progresses \citep[e.g.,][]{zhang2021correlation}. The variable 
-``Gender" has levels ``M" and ``F" for ``male" and ``female" respectively.  The
-variable ``Batch" takes a 
-numerical value and every heat of every round from 1999 to 2022 was assigned it's
-own number.  This was added because of suspected issues that reaction time may be
-correlated with other runners reacting quickly (one runner reacting quickly may
-result in other runners reacting quickly). There is no statistical meaning to
-how the batches were labeled, they were ordered chronologically by meet but
-reverse chronologically by year (Batch 1 is a preliminary heat of 2022 and Batch
-111 corresponds to the 1999 finals). %Included are
+%semifinal heat (of which there are either two
+%or three each year) denoted by ``S", or final heat denoted by ``F".
+We focus on the reaction times during the semifinal and final heats only, as 
+reaction times from preliminary heats are often not as fast as those in later 
+heats %were not studied because athletes react faster as the 
+%competition progresses 
+\citep[e.g.,][]{zhang2021correlation}. 
+%The variable 
+%``Gender" has levels ``M" and ``F" for ``male" and ``female" respectively.  
+%The variable ``Batch" takes a numerical value and every heat of every round from 
+%1999 to 2022 was assigned it's
+%own number.  This was added because of suspected issues that reaction time may be
+%correlated with other runners reacting quickly (one runner reacting quickly may
+%result in other runners reacting quickly). There is no statistical meaning to
+%how the batches were labeled, they were ordered chronologically by meet but
+%reverse chronologically by year (Batch 1 is a preliminary heat of 2022 and Batch
+%111 corresponds to the 1999 finals). %Included are
 %box plots of the reaction times of the men's finals and the men's semifinals and
@@
 %\citep{zhang2021correlation}.  
-\eds{if you are really not using ``H'' data at all, it does not appear necessary
-to spend time defining earlier.  You could just say directly when you define 
-stage that you only consider semifinals and finals and cite the Zhang paper for 
-rationale}
-\of{Ok I updated this.  Should I delete the batch variable too?  Although I don't
-use it, I do reference it in the methods section}
+%\eds{if you are really not using ``H'' data at all, it does not appear necessary
+%to spend time defining earlier.  You could just say directly when you define 
+%stage that you only consider semifinals and finals and cite the Zhang paper for 
+%rationale}
+%\of{Ok I updated this.  Should I delete the batch variable too?  Although I don't
+%use it, I do reference it in the methods section}
 The pooled data increases the sample size which is advantageous because in some
-years there are very few finals observations.  For example, in 2022 there
+years there are very few final heat observations.  For example, in 2022 there
 were only five data points due to two disqualifications and one athlete who 
 did not compete.
+
 The data under consideration is best visualized using side-by-side box plots, as
-shown in Figure~\ref{fig:Boxplots}.  It is clear that the 2022 box plot for both 
-the men's finals and semifinals data is lower (indicating faster reaction times)
+shown in Figure~\ref{fig:Boxplots}.  It is clear that the reaction times in the 
+2022 box plot for both 
+the men's finals and men's pooled semifinal and final data are lower (indicating 
+faster reaction times)
 than in previous years.  The median for the men's pooled data in 2022 was 0.129
 and the median for the men's final data in 2022 was 0.126.  For comparison,
-\citet{brosnan2017effects} found median reaction times of 0.156 seconds \eds{mean?  median?} 
-\of{Resolved}
+\citet{brosnan2017effects} found median reaction times of 0.156 seconds 
 for men when looking at data from 1999 to 2014.
@@
 the IAAF (the former name for World Athletics, and the governing body for the 
-World Track and Field Champions) instituted a rule change that allowed one 
+World Track and Field Championships) instituted a rule change that allowed one 
 warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}. 
@@
 policy and cracking down on false starts, World Athletics had reduced men's 
-false starts by two thirds \eds{for men, not overall} \of{Better?} in 2011 (6 male and 4 female disqualifications) 
+false starts by two thirds in 2011 (6 male and 4 female disqualifications) 
 \citep{iaaf2009falsestart}.
@@
 significant improvement in reaction times under more lenient rules.
-\eds{is that right?  it seems counterintuitive to me}
-\of{Yes you are correct, I updated the language.}
-
-\eds{I am not sure the previous sentence follows from the one preceding it. Why
- does World Athletics find faster reaction times desirable?  Do you mean that
-they find fewer false starts desirable?}
-
-\eds{Two comments about the previous sentence: 1. 
-it does not logically follow the previous sentence as there is no mention of 
-gender, and 2. if you are not using women's data, don't spend time talking about 
-it at the beginning of the data section.  Similar to my comment about 
-preliminary heats, you can move this sentence to the beginning of the Data 
-section to provide a rationale for why you do not include data for women.} 
-\of{Updates: I moved any talk about women's data to start of Data section, I think
-it flows much better now}
-
-\subsection{Beyond World Championship}\label{sec:databeyond}
-The motivation for looking beyond World Championship data came from an 
-exploratory \eds{exploratory instead of qualitative?} \of{I agree; fixed}
-analysis examining how United States athletes reacted at the 
+  %\eds{I am not sure the previous sentence follows from the one preceding it. Why
+% does World Athletics find faster reaction times desirable?  Do you mean that
+%they find fewer false starts desirable?}
+
+%\eds{Two comments about the previous sentence: 1. 
+%it does not logically follow the previous sentence as there is no mention of 
+%gender, and 2. if you are not using women's data, don't spend time talking about 
+%it at the beginning of the data section.  Similar to my comment about 
+%preliminary heats, you can move this sentence to the beginning of the Data 
+%section to provide a rationale for why you do not include data for women.} 
+%\of{Updates: I moved any talk about women's data to start of Data section, I think
+%it flows much better now}
+
+\subsection{Beyond World Championships}\label{sec:databeyond}
+The motivation for looking beyond World Championships data came from an 
+exploratory analysis examining how United States athletes reacted at the 
 2022 United States Track and Field Championships.
-\eds{be consistent with how you refer to the US championship:
-``United States Track and Field Championships" and ``United States Track and Field 
-Championships" are both being used; please pick one and use that throughout}
-\of{Resolved; changed all USA to United States.}
-%This section was motivated from a qualitative analysis that started by looking
-%at how United States athletes reacted at the 2022 United States Track and Field
-%Championships. 
+%\eds{be consistent with how you refer to the US championship:
+%``United States Track and Field Championships" and ``United States Track and Field 
+%Championships" are both being used; please pick one and use that throughout}
+%\of{Resolved; changed all USA to United States.}
 From June 23-26 2022, the United States held its Track
-and Field Championship at Hayward 
+and Field Championships at Hayward 
 Field in Eugene, Oregon to decide who to send to the World Championships 
@@
 Championship races. Devon Allen raced three times at the United States Track and
-Field Championship and three times at the World Track and Field Championship. 
+Field Championship and three times at the World Track and Field Championships. 
 His reaction times were substantially faster in the World Track and
-Field Championship compared to the United States Track and Field Championship as were his
-teammates. \eds{statistically significantly faster? If you used a statistical 
-test, it should not be described in this section.  If you didn't use a 
-statistical test, you should replace the word `significantly' with something 
-else (substantially?)} \of{Replaced signficantly with substantially}
+Field Championships compared to the United States Track and Field Championships 
+as were his
+teammates. 
 This comparison suggests that the timing system used between the two
@@
 table is different than the rest of the paper}
+\eds{LaTeX keeps captions singlespaced automatically with the doublespace 
+command; don't worry about it!}
 
@@
 \eds{need consistency with ``Championship" or ``Championships - I am not sure 
-which one is correct}
-
-\subsection{GLMM}
+which one is correct.  I am using Championships (plural) but
+change everything to singular if that is preferred}
+
+\subsection{GLMM}\label{sec:glmm}
 Based on an exploratory analysis using the GLMM model for the reaction times
-from the World Championship, the final model we chose was a gamma mixed-effect 
+from the World Championships, the final model we chose was a gamma mixed-effect 
 model. As suggested by the log-likelihood and the Akaike Information
@@
 about it too.}
+We include a random effect for venue because the technology
+that World Athletics has used has differed since 1999, and theoretically the
+reaction time data from 2022 should be the best as the timing system should be
+state of the art. Additionally, any other weather or climate related factors 
+such as humidity, precipitation, elevation, could be captured by this effect.
 The gamma mixed-effect model can be fit with the \texttt{glmer()} function in R
@@
 
-As for the random effect, it was possible to include a heat 
+Finally, we remark here that it is also possible to include a heat or race 
 \eds{heat? when I first read this, I thought you meant race as in ethnicity}
@@
 removed most talk about heat from the data section} 
-effect, as well.
-From our exploration, however, the race effect is very difficult to explain or 
-interpret as it means that some heats of races were either
-faster or slower and that everyone in the race reacted similarly.  Although
+\eds{I see, I did not understand this to be Batch effect when I read it the 
+first time.  I commented out the Batch text in the Data section - I think it is 
+fine to just discuss it here, as it ultimately was not used.  I will defer to 
+Dr Yan on this, however.}
+effect to account for some heats possibly having faster or slower 
+reaction times for everyone involved in the heat.
+%From our exploration, however, the race effect is very difficult to explain or 
+%interpret as it means that some heats of races were either
+%faster or slower and that everyone in the race reacted similarly.  
+Although
 it is possible that one athlete reacting faster could lead to others reacting
@@
 heats that had fast and slow reaction times are all within 0.12 seconds of each
-other.  It is much easier, however, to explain the venue effect.  The technology
-that World Athletics has used has differed since 1999 and theoretically the
-reaction time data from 2022 should be the best as the timing system should be
-state of the art. Any other weather or climate related factors such as humidity,
-precipitation, elevation, etc. would all be negated when looking at the race
-effect, but are all factors that could potentially impact the venue effect.
-
-\subsection{Rank-based Comparison}
-
-To test the conjecture that the 2022 World Champion timing device may have led
-to faster recorded reaction times, we compare the reaction times of the same
-athletes who have attended both the 2022 World Champion and other competitions.
+other. 
+\eds{Did you try it and find that the effect was not needed?  If yes,
+I would include that here.  That argument 
+would be more compelling for its non-use than its difficulty in interpretation.}
+
+\subsection{Rank-based Comparison}\label{sec:rank}
+
+To test the conjecture that the 2022 World Championships timing device may have 
+led to faster recorded reaction times, we compare the reaction times of the same
+athletes who have attended both the 2022 World Championships and other 
+competitions. 
 In this setting, we have clustered data with subunit grouping. In particular,
 each athlete is a cluster and the multiple reaction times from the same athlete
-can be from either the 2022 World Championship or otherwise.
+can be from either the 2022 World Championships or otherwise.
 Let $X_{ij}$ be the $j$th reaction time of athlete~$i$, $i = 1, \ldots, n$,
@@
 athelete~$i$. Let $\delta_{ij}$ be the group indicator of $X_{ij}$; $\delta_{ij}
-= 1$ if $X_{ij}$ is in group~1 (2022 World Championship) and $\delta_{ij} = 0$ 
+= 1$ if $X_{ij}$ is in group~1 (2022 World Championships) and $\delta_{ij} = 0$ 
 otherwise. Athletes are
@@
 
-As shown in table~\ref{tab:Gamma_parameters}, the standard error of the year
+As shown in Table~\ref{tab:Gamma_parameters}, the standard error of the year
 effect increases for both the men's final and men's pooled data sets when
@@
 that the inclusion of 2022 increased the randomness of the data and motivates
-the rank-based study~\ref{subsec:Results_Rank}.
+the rank-based study (see Section~\ref{subsec:Results_Rank}).
 
@@
 being below 0.1 seconds for the various different data sets are shown in 
-Table~\ref{tab:Sim_probability}. Prior to the 2022 World Championship, there was 
+Table~\ref{tab:Sim_probability}. Prior to the 2022 World Championships, there was 
 a 0.0439\% chance (based on the
@@
 once every 2,277 starts.  Considering there are roughly 8 men's finals reaction
-times every 2 years (the World Championship is biennial), every 569 years there
+times every 2 years (the World Championships are biennial), every 569 years there
 should be a time as extreme as Allen's. This seems extremely unlikely, and thus
@@
 
-Table ~\ref{tab:Sim_probability} also shows that changing the reaction time 
+Table~\ref{tab:Sim_probability} also shows that changing the reaction time 
 barrier from 0.1 seconds to 0.08 seconds drastically reduces the chances of 
@@
 the athlete predicted the gun and it was luck that caused their abnormally high 
-time.  At the World Championship level, World Athletics and Seiko want to remove 
+time.  At the World Championships level, World Athletics and Seiko want to remove 
 that element of luck and thus impose the 0.1 second barrier. In 2013, Seiko 
 upgraded their timing equipment for sprinting events (110 meter hurdles falls 
-under this category) \citep{wa2013backtage} \eds{110 for men?} \of{Fixed}. Since 1999, the
+under this category) \citep{wa2013backtage}. Since 1999, the
  highest \eds{highest as in fastest or slowest? Neither seems consistent with 
@@
 the 0.1 
-second barrier again at the next World Track Championship (August 2023)
-\eds{check against introduction}, then World Athletics 
+second barrier again at the next World Championships (August 2023), then 
+World Athletics 
 needs to adjust that threshold and allow for faster reaction times so that 
@@
 %Put downloadable data spreadsheet
-The World Championship data is available from the World Athletics website:
+The World Championships data are available from the World Athletics website:
 \url{https://www.worldathletics.org/results/world-athletics-championships}.
@@
 \jy{You have countries beyond United States}
-The United States Track and Field Championships data is available at: 
+The United States Track and Field Championships data are available at: 
 \url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}

### 2023-04-28 Manuscript/manuscript.tex
minor comments

@@
 
-%\jy{Avoid references in abstract.}
-
-\jy{This is only the background. An abstract also needs to summarize the 
-contributions.}
-%\bigskip
 %\eds{removed Devon Allen from keywords because his name is already in title
@@
 
-\noindent\sc{Keywords}:
+\bigskip\noindent\sc{Keywords}:
 False Start, Hurdles, Reaction Time, Seiko
@@
 
-\jy{Turn on column numbers in your editor to keep line width under 80.}
-
 \section{Introduction}
@@
 
-\jy{Not a good opening. Could start by telling the story, recapping the major
-  events and discussions on the internet. Then dig into the relevant
-  literature. Finally outline the contributions of this work.
-}
-
-\jy{use the following structure and flow:
-  1) A paragraph to tell the story;
-  2) A paragraph on informal statistical analyses on the internet/social media;
-  3) A paragraph on scholarly articles on reaction time in the literature;
-  4) A paragraph on what's new: methods; results.
-  5) A paragraph of roadmap.}
-
-%1)
 Devon Allen, an alumnus of the University of Oregon in 
@@
 
-\jy{Could be condensed}
 \citet{pilianidis2012start} found
@@
 %4)
-\jy{language needs to be formalized and condensed.}
 The objective to this paper is to analyze the reaction times at the 2022 World 
@@
 
-\jy{You have two such datasets in the analysis. Use one paragraph for each.}
-\of{Split it into 2 sections, do you still want me to do something else?}
-
 This dataset is too small to perform any meaningful statistical analysis, but
@@
   how is this athlete clustering being accounted for here?}
+\jy{We could add an athlete random effect too, if the players are identified in the data.}
 Conditional on a venue effect $z_i$ of year~i,  $Y_{ij}$ is modeled by 

### 2023-05-11 Manuscript/manuscript.tex
Starting to work again

--- 2023-04-27 Manuscript/manuscript.tex
@@
 available from the \texttt{clusWilcox.test()} function with
-\texttt{method = ``ds'} from R package \texttt{clusrank}
+\texttt{method = `ds'} from R package \texttt{clusrank}
 \citep{jiang2017wilcoxon}.
@@
 effects \eds{remove race effects?} that greatly impact reaction time.
-However, this practically does not make much sense
-as there is no reason for so much random variation without much of a trend.
-There has not been a consistent decrease or increase since 1999, and much of the 
-data
-for reaction times has been random and unpredictable. 
+
 \eds{I would suggest removing the last two sentences, as they would suggest your
  modeling with GLMM is not good}.  
+\of{I agree and they have been removed}
 
@@
 women, exploratory analysis showed that 2022 was at best only
-slightly faster relative to other championships going back to 2022 
+slightly faster relative to other championships going back to 2003 
 \eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 

### 2023-05-11 Code/Models.Rmd
Identity vs log comparison is in the bottom 4 chunks.  For each dataset the link=identity was compared with link=log

--- 2023-04-11 Code/Models.Rmd
@@
 
+#Trying out identity link
 ```{r}
-library(xtable)
-xtable(men_pooled_year_gamma_log)
+men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma(link = "log"))
+t1 <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma(link = "identity"))
+paste("Gamma and log with year effect has AIC of", AIC(men_finals_year_gamma_log), "and log-lik of", logLik(men_finals_year_gamma_log))
+paste("Linear without year effect has AIC of", AIC(t1), "and log-lik of", logLik(t1))
 ```
+
 ```{r}
-library(memisc)
-mtable(men_pooled_year_gamma_log)
+men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "log"))
+t2 <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "identity"))
+paste("Linear without year effect has AIC of", AIC(men_old_finals_year_gamma_log), "and log-lik of", logLik(men_old_finals_year_gamma_log))
+paste("Linear without year effect has AIC of", AIC(t2), "and log-lik of", logLik(t2))
 ```
+
+```{r}
+men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma(link = "log"))
+t3 <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma(link = "identity"))
+paste("Gamma and log with year effect has AIC of", AIC(men_pooled_year_gamma_log), "and log-lik of", logLik(men_pooled_year_gamma_log))
+paste("Linear without year effect has AIC of", AIC(t3), "and log-lik of", logLik(t3))
+```
+
+```{r}
+men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "log"))
+t4 <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "identity"))
+paste("Gamma and log with year effect has AIC of", AIC(men_old_pooled_year_gamma_log), "and log-lik of", logLik(men_old_pooled_year_gamma_log))
+paste("Linear without year effect has AIC of", AIC(t4), "and log-lik of", logLik(t4))
+
+```

### 2023-05-11 Manuscript/manuscript.tex
Beginning to go over changes

@@
 final model, but the internet seems to favor this link so a reviewer may ask 
-about it too.}
+about it too.} \of{I have added code at the bottom of Code/Models.Rmd that
+compared the link = "identity" and link="log" for each dataset.  I found that
+the link=log was marginally better.}
 We include a random effect for venue because the technology
@@
 under this category) \citep{wa2013backtage}. Since 1999, the
- highest \eds{highest as in fastest or slowest? Neither seems consistent with 
-the boxplots in Figure 1} reaction 
+slowest \eds{highest as in fastest or slowest? Neither seems consistent with 
+the boxplots in Figure 1} \of{Resolved} reaction 
 times in the men's 110-meter hurdle were recorded in 2013.  That is not to say

### 2023-05-11 Manuscript/manuscript.tex
Have gone through more of Prof Schifano's suggestions

@@
 %Allen, who attended the University of Oregon also in Eugene, 
-%(the location of US World Championship 
+%(the location of US World Championships 
 %\eds{US World? remove World?} and the World Championship) 
@@
 the LetsRun website.  He argued that there was an error
-with the timing equipment at the 2022 World Track and Field Championship and
+with the timing equipment at the 2022 World Track and Field Championships and
 cited graphs, computed simple statistics such as median reaction times, and 
-compared the reaction times at the 2022 World Track and Field Championship to 
+compared the reaction times at the 2022 World Track and Field Championships to 
 other competitions
@@
 reaction times over the past 13 competitions of the men's 100 meter, men's 110 
-meter hurdle, and women's 100 meter would all come from the same championship 
+meter hurdle, and women's 100 meter would all come from the same year. 
 \citep{johnson2022was}.
 %came from Johnson claiming there is a $(1/13)^3$ chance that the fastest 
-%reaction times came during the same championship for the men's 100 meter,
+%reaction times came during the same championships for the men's 100 meter,
 %men's 110 meter hurdle, and women's 100 meter for 13 competitions held between 2003 and 2022 
@@
 \eds{is this better? (and still correct?)}
+\of{Yes, I think it looks good}
 
@@
 The objective to this paper is to analyze the reaction times at the 2022 World 
-Track and Field Championship to check for any abnormalities, and we approach
+Track and Field Championships to check for any abnormalities, and we approach
 this two different ways.  First we model the data, which is
@@
 %First we can compare the reaction times of athletes who
-%competed at the World Track and Field Championship in both 2019 and 2022, and 
-%second we can compare how athletes performed at the 2022 championship relative 
+%competed at the World Track and Field Championships in both 2019 and 2022, and 
+%second we can compare how athletes performed at the 2022 championships relative 
 %to their national/qualifiers meet. In order to reduce any chances
@@
 rank-based analyses, or only the national competitions?}
+\of{Yes the ranked based comparison has 2 subsections: 2022 national vs 2022
+international and 2019 international vs 2022 international.  Should I instead say
+that there are three types of data across 2 sections or something like that?}
 The first data set is used for the GLMM analyses described in 
@@
 %The effect? At the 2011 
-%World Championship there were 6 male disqualifications and 4 female 
+%World Championships there were 6 male disqualifications and 4 female 
 %disqualifications \citep{iaaf2009falsestart}.  
@@
 being held at the same venue in August. %the exact same venue as used 
-%for the World Track and Field Championship a month later.  
+%for the World Track and Field Championships a month later.  
 %Thus, a baseline is able to be established for the four United States athletes who 
@@
 Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
-Track and Field Championship races compared to the United States Track and Field 
-Championship races. Devon Allen raced three times at the United States Track and
-Field Championship and three times at the World Track and Field Championships. 
+Track and Field Championships races compared to the United States Track and Field 
+Championships races. Devon Allen raced three times at the United States Track and
+Field Championships and three times at the World Track and Field Championships. 
 His reaction times were substantially faster in the World Track and
@@
 
-The %World Championship data and the data eyond the World Championship 
+The %World Championships data and the data eyond the World Championships 
 data described in Sections~\ref{sec:dataworld} and~\ref{sec:databeyond} were
@@
 change everything to singular if that is preferred}
+\of{I believe that championships is correct, I have changed all instances of
+championship to championships}
 
@@
 %% Group is either designated to be 1 or 2, with 1 corresponding to either the
-%% national level data or the previous world championship data and 2 corresponding
+%% national level data or the previous world championships data and 2 corresponding
 %% to the 2022 data.
@@
 Since 1985 Seiko Holding Corporations has served every World Athletics 
-Championship as the official timer \citep{wa2022seiko}.  Since 1985, the 
+Championships as the official timer \citep{wa2022seiko}.  Since 1985, the 
 technology and the ability
@@
 disqualification barrier was
-imposed in the first place. Allen showed at the Championship he was able to
-react extremely fast two times, which may suggest skill rather than luck. Devon 
+imposed in the first place. Thus having a quick reaction time multiple times
+may indicate skill rather than luck. Devon 
 Allen is a very good athlete, as evidenced by his football career at the 
@@
 and making it onto the Philadelphia Eagles practice squad this past year 
-\citep{hurley2022eagles}. Thus, the 0.099 reaction time may not have been 
+\citep{hurley2022eagles}. The 0.099 reaction time may not have been 
 a product of Devon Allen predicting the start of the race but rather a 

### 2023-05-12 Manuscript/manuscript.tex
Moved a paragraph commented out in the intro to the end of the data section

@@
 every sprinter who steps onto the track.
-%\eds{I recommend moving the rest of this paragraph to Section 2.2 as I don't
-%think it is critical for readers to know all the details here.}
-%First we can compare the reaction times of athletes who
-%competed at the World Track and Field Championships in both 2019 and 2022, and 
-%second we can compare how athletes performed at the 2022 championships relative 
-%to their national/qualifiers meet. In order to reduce any chances
-%of athletes improving reaction time, we can also look at data from qualifying meets.
-%Almost every athlete had to qualify by proving they are elite at the national level, 
-%usually by coming in at least fourth. Thus, data was compiled on athletes who competed
-%at the 2022 World Track and Field Championships on what these athletes also ran 
-%at their country's national championship. If there was any argument that athletes
-%have improved their reaction time from 2019 to 2022, it should now be considered
-%nearly null, as the national competitions took place between June and July 2022 
-%and the World Championships were in August. The equipment used to measure reaction
-%time cannot be guaranteed to be the same from national to international
-%competition, but the time span between meets has been reduced considerably.
 
@@
 athletes/clusters?}
+\eds{I recommend moving the rest of this paragraph to Section 2.2 as I don't
+think it is critical for readers to know all the details here.}
+\of{Is the one below what you're talking about?}
+First we can compare the reaction times of athletes who
+competed at the World Track and Field Championships in both 2019 and 2022, and 
+second we can compare how athletes performed at the 2022 championships relative 
+to their national/qualifiers meet. In order to reduce any chances
+of athletes improving reaction time, we can also look at data from qualifying meets.
+Almost every athlete had to qualify by proving they are elite at the national level, 
+usually by coming in at least fourth. Thus, data was compiled on athletes who competed
+at the 2022 World Track and Field Championships on what these athletes also ran 
+at their country's national championship. If there was any argument that athletes
+have improved their reaction time from 2019 to 2022, it should now be considered
+nearly null, as the national competitions took place between June and July 2022 
+and the World Championships were in August. The equipment used to measure reaction
+time cannot be guaranteed to be the same from national to international
+competition, but the time span between meets has been reduced considerably.
 

### 2023-05-16 Manuscript/manuscript.tex
cleaning of resolved comments

@@
 \begin{abstract}
-Devon Allen was disqualified in the Men's 110 meter hurdle final of the 2022
+Devon Allen was disqualified in the men's 110 meter hurdle final of the 2022
 World Track and Field Championships after registering a reaction time of 0.099 
@@
 concluded that the reaction time data from the 2022 World Championships 
-seemed to be generally faster compared to the other datasets they looked at, but 
-they did not perform any formal %meaningful 
+seemed to be generally faster compared to the other datasets they considered, 
+but they did not perform any formal %meaningful 
 statistical analyses. This paper questions the reaction time 
@@
 venue effect in order to %try and 
-model the data. Additionally, we employ a signed-rank test for clustered data to 
+model the reaction time data. Additionally, we employ a signed-rank test for 
+clustered data to 
 compare reaction times for the same athletes at different competitions. This 
@@
 
-\eds{also removed 2022 World Track and Field Championships from keywords and 
-changed location of World in the title. World Track and Field Championships
-has more Google search results than Track and Field World Championships, so
-I opted to go with this. Also used plural Championships. Feel free to switch 
-order or use singular if one or both of these is 
-preferred for some reason.  }
+%\eds{also removed 2022 World Track and Field Championships from keywords and 
+%changed location of World in the title. World Track and Field Championships
+%has more Google search results than Track and Field World Championships, so
+%I opted to go with this. Also used plural Championships. Feel free to switch 
+%order or use singular if one or both of these is 
+%preferred for some reason.  }
 
@@
 heats and semifinals to reach the final heat while posting reaction times of 
-0.123 in his preliminary heat and 0.101 in his semifinal.  
+0.123 and 0.101 in his preliminary heat and semifinal, respectively.  
 %Allen, who attended the University of Oregon also in Eugene, 
@@
 reaction times over the past 13 competitions of the men's 100 meter, men's 110 
-meter hurdle, and women's 100 meter would all come from the same year. 
+meter hurdle, and women's 100 meter would all come from the same year 
 \citep{johnson2022was}.
@@
 %\of{Cleaned up but may need additional work}
-\of{Cleaned this up a little but still a little messy.}
-\eds{is this better? (and still correct?)}
-\of{Yes, I think it looks good}
+%\of{Cleaned this up a little but still a little messy.}
+%\eds{is this better? (and still correct?)}
+%\of{Yes, I think it looks good}
 
@@
 react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
+\eds{are hurdlers generally considered a subset of sprinters or are they 
+considered a separate type of athlete?  Specifically did the study above 
+recommend the barrier be lowered for both sprinters and hurdlers?
+You also refer to sprinters elsewhere in the manuscript.  Are those instances
+meant to refer to both sprinters and hurdlers, as well?}
 Additionally, it was recommended to IAAF that high speed cameras be used to
@@
 because Allen was 27 years old at the 2022 World Track and Field Championships 
-and at his physical peak.
+and likely at his physical peak.
 
@@
 Despite their study, IAAF has not changed the reaction barrier since it was
-implemented at the 1999 World Track and Field Championships.  It is confusing
-as to why IAAF would commission a study only to disregard the results and not
-make the necessary changes.  It is unknown how many athletes since 2009 could
+implemented at the 1999 World Track and Field Championships.  It is unclear
+why IAAF would commission a study only to disregard the results and not
+make the recommended changes.  It is unknown how many athletes since 2009 could
 have potentially benefited from this, but it is probable that there are others
@@
 Track and Field Championships to check for any abnormalities, and we approach
-this two different ways.  First we model the data, which is
+this in two different ways.  First we model the data, which are
 the reaction times of every non-disqualification since the 1999 World 
@@
 accounting for the venue (or year). 
-%\eds{are you using covariates?  Data section
-%implies that you are, but Results section does not}
-%\of{Cleaned up confusion about covariates, but no none were used in the paper} 
 Based on the fitted model, we can then estimate the probability of a reaction 
@@
 comparing reaction times from the same athlete across multiple competitions we 
-are able to determine if the 2022 data differed significantly from other 
-competitions.
-Our investigation into reaction times is important not only because if its
+are able to determine if the 2022 reaction times differed significantly from 
+those of other competitions.
+Our investigation into reaction times is important not only because of its
 implications for future races, but also because reaction times have
 been shown to affect the overall time of the race \citep{delalija2008reaction}.
-Thus, faster reaction times lead to faster overall times; which is the goal of
+Thus, faster reaction times lead to faster overall times, which is the goal of
 every sprinter who steps onto the track.
@@
 that there are three types of data across 2 sections or something like that?}
+\eds{I think this is fine here, but it is not at all clear that there are two
+subsections in the rank-based results for 2022 national vs 2022
+international and 2019 international vs 2022 international.}
 The first data set is used for the GLMM analyses described in 
@@
 and covers the men's 110 meter hurdles from 1999 to 2022; the source can be 
-found in the Appendix.%~\ref{sec:Appendix}.   %\eds{see earlier comment - 
+found in the Appendix. %~\ref{sec:Appendix}.   %\eds{see earlier comment - 
 %are you using any of these as covariates?}
 %\of{removed covariate confusion, talk about womens a couple sentences down} 
-Although women's data is available, we chose not to include women's data in this
+Although women's data is available, we chose not to include women's data in the
 analyses to follow due to the numerous papers citing differences in men and
-women's reaction times \citep{lipps2011implications, babicc2009reaction,
-panoutsakopoulos2020gender}.
+women's reaction times \citep[e.g.,][]{lipps2011implications, 
+babicc2009reaction, panoutsakopoulos2020gender}.
 %The variable ``Stage" refers to whether the observation occurred during a 
@@
 
-The data under consideration is best visualized using side-by-side box plots, as
+The data under consideration is best visualized using side-by-side boxplots, as
 shown in Figure~\ref{fig:Boxplots}.  It is clear that the reaction times in the 
-2022 box plot for both 
+2022 boxplot for both 
 the men's finals and men's pooled semifinal and final data are lower (indicating 
@@
 and the GLMM subsection in the paragraph starting with "There are two types..."}
-Something that is extremely important to note in the data is the inconsistency 
-of World Athletics in terms of how they penalize false starts.  From 2007--2009,
+\eds{With the following newly added sentences, I think we can leave this paragraph here.}
+We can also see from Figure~\ref{fig:Boxplots} that the reaction times clearly
+differ from year to year. The venue of the World Championships changes each 
+year, so weather or climate related factors such as humidity, precipitation, elevation, may indeed play a role, but additionally, both the technology and 
+false start penalization rules have changed over the study period. 
+%Something that is extremely important to note in the data is the inconsistency 
+%of World Athletics in terms of how they penalize false starts.  
+From 2007--2009,
 the IAAF (the former name for World Athletics, and the governing body for the 
@@
 warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}. 
-This was desirable for World Athletics as false starts can make already long 
-track meets more tedious, both for athletes and viewers watching the television 
-broadcast.  There were 18 male and 7 female false starts at the 2007 World 
-Championships, and 18 male and 7 female starts at the 2009 World Championships. 
+%This was desirable for World Athletics as false starts can make already long 
+%track meets more tedious, both for athletes and viewers watching the television 
+%broadcast.  
+%\eds{I think we lost something here with all of the edits so I removed the 
+%previous sentence. What was desirable for World Athletics -- the 2007-2009 rule
+% change?  Why was this desirable?  The rule change would seem to make the event
+% longer and not shorter? } 
+There were 18 male and 7 female false starts at the 2007 World 
+Championships, and 18 male and 7 female false starts at the 2009 World 
+Championships. 
 Starting in 2011, the rule was scrapped, and the old policy which automatically 
@@
 significant improvement in reaction times under more lenient rules.
-  %\eds{I am not sure the previous sentence follows from the one preceding it. Why
-% does World Athletics find faster reaction times desirable?  Do you mean that
-%they find fewer false starts desirable?}
-
-%\eds{Two comments about the previous sentence: 1. 
-%it does not logically follow the previous sentence as there is no mention of 
-%gender, and 2. if you are not using women's data, don't spend time talking about 
-%it at the beginning of the data section.  Similar to my comment about 
-%preliminary heats, you can move this sentence to the beginning of the Data 
-%section to provide a rationale for why you do not include data for women.} 
-%\of{Updates: I moved any talk about women's data to start of Data section, I think
-%it flows much better now}
 
@@
 His reaction times were substantially faster in the World Track and
-Field Championships compared to the United States Track and Field Championships 
+Field Championships compared to the United States Track and Field Championships, 
 as were his
@@
 times for Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts.
-%``United States" denotes
-%the United States Track and Field Championships, ``World" denotes the World Track and Field
-%Championships. All numbers listed in the table are the reaction
-%time for the respective athletes in seconds.
 
@@
 
-\of{Im not sure how to fix this but the spacing between the lines of the
-table is different than the rest of the paper}
-\eds{LaTeX keeps captions singlespaced automatically with the doublespace 
-command; don't worry about it!}
+%\of{Im not sure how to fix this but the spacing between the lines of the
+%table is different than the rest of the paper}
+%\eds{LaTeX keeps captions singlespaced automatically with the doublespace 
+%command; don't worry about it!}
 
 This dataset is too small to perform any meaningful statistical analysis, but
-the data was expanded to include reaction times for athletes who competed at 
+the data were expanded to include reaction times for athletes who competed at 
 national competitions (many of which were held from May-July of 2022). Once the 
@@
 athletes/clusters?}
-\eds{I recommend moving the rest of this paragraph to Section 2.2 as I don't
-think it is critical for readers to know all the details here.}
 \of{Is the one below what you're talking about?}
+\eds{Yes, this is the part from the intro I was referring to, but you need to 
+edit this paragraph now so it all works together.}
 First we can compare the reaction times of athletes who
@@
 analysed with GLMM and rank-based methods, respectively.
-\eds{need consistency with ``Championship" or ``Championships - I am not sure 
-which one is correct.  I am using Championships (plural) but
-change everything to singular if that is preferred}
-\of{I believe that championships is correct, I have changed all instances of
-championship to championships}
+%\eds{need consistency with ``Championship" or ``Championships - I am not sure 
+%which one is correct.  I am using Championships (plural) but
+%change everything to singular if that is preferred}
+%\of{I believe that championships is correct, I have changed all instances of
+%championship to championships}
 
@@
 much inferior to the gamma model which better captures the skewness of the
-data. A random effect for the venue or year was also found to be necessary.
+data. A random effect for the venue or year was also found to be necessary, 
+which makes intuitive sense given the discussion in Section~\ref{sec:dataworld}. 
 Let $Y_{ij}$ be the reaction time of observation~$j$ in year~$i$.
@@
 $z_i$ is a normally distributed random effect with mean zero and
-variance~$\sigma^2$. The link function $g$ can be the logarithm function to
-ensure the positivity of the conditional mean.
-\eds{Did you try an identity link also? Lo and Andrews (2015) suggest that the 
-Gamma family with an identity link works well for their reaction-time data.
-It doesn't seem like a good idea in general because it is likely that some 
-negative predicted responses will occur either during model fitting or in the 
-final model, but the internet seems to favor this link so a reviewer may ask 
-about it too.} \of{I have added code at the bottom of Code/Models.Rmd that
-compared the link = "identity" and link="log" for each dataset.  I found that
-the link=log was marginally better.}
-We include a random effect for venue because the technology
-that World Athletics has used has differed since 1999, and theoretically the
-reaction time data from 2022 should be the best as the timing system should be
-state of the art. Additionally, any other weather or climate related factors 
-such as humidity, precipitation, elevation, could be captured by this effect.
+variance~$\sigma^2$. Although \citet{lo2015idlink} suggested that the gamma 
+family with an identity link works well for reaction time data, the logarithm 
+link function worked better for our data and additionally ensures the positivity
+of the conditional mean.
+% The link function $g$ can be the logarithm function to
+% ensure the positivity of the conditional mean.
+%\eds{Did you try an identity link also? Lo and Andrews (2015) suggest that the 
+%Gamma family with an identity link works well for their reaction-time data.
+%It doesn't seem like a good idea in general because it is likely that some 
+%negative predicted responses will occur either during model fitting or in the 
+%final model, but the internet seems to favor this link so a reviewer may ask 
+%about it too.} \of{I have added code at the bottom of Code/Models.Rmd that
+%compared the link = "identity" and link="log" for each dataset.  I found that
+%the link=log was marginally better.}
+%We include a random effect for venue (or year) because the technology
+%that World Athletics has used has differed since 1999, and theoretically the
+%reaction time data from 2022 should be the best as the timing system should be
+%state of the art. Additionally, any other weather or climate related factors 
+%such as humidity, precipitation, elevation, could be captured by this effect.
+%\eds{I moved this info into the Data section.}
 The gamma mixed-effect model can be fit with the \texttt{glmer()} function in R
@@
 
-\eds{I would suggest removing the last two sentences, as they would suggest your
- modeling with GLMM is not good}.  
-\of{I agree and they have been removed}
+%\eds{I would suggest removing the last two sentences, as they would suggest your
+% modeling with GLMM is not good}.  
+%\of{I agree and they have been removed}
 

### 2023-05-19 Manuscript/manuscript.tex
Went through some of Prof Schifanos revisions

@@
 meant to refer to both sprinters and hurdlers, as well?}
+\of{Yes I would say that hurdlers are a subset of sprinter.  The reaction time
+barrier for the men's 100 meters and men's 110 meter hurdle are both 0.1 seconds.
+The paper recommended for the reaction time barrier to be decreased to roughly 0.8
+seconds.  It does not reference any distinction between hurdlers/sprinters, so I
+think they can be grouped together.  In my personal opinion I don't see why the
+reaction time for hurdlers versus 100m dash would differ significantly.  In terms
+of what I wrote in the paper, I want to stay accurate to what is in the study
+but also generalize studies involving reaction times to include hurdlers.  I believe
+many of the studies examining the reaction time barrier from a physiological
+perspective to find the minimum possible reaction time do not deal with hurdles.
+Thus, I don't know if I can change what I wrote to be both accurate and understandable,
+but what do you both think?}
 Additionally, it was recommended to IAAF that high speed cameras be used to
@@
 international and 2019 international vs 2022 international.}
+\of{Currently under the Methods/Results section I have GLMM and rank based comparison
+do you think I need subsections under rank based comparison, or is that too much
+compartmentalization?  Or should I instead say there are 3 types of data that are
+used in two different methods, or something like that?}
 The first data set is used for the GLMM analyses described in 
@@
 \eds{With the following newly added sentences, I think we can leave this paragraph here.}
+\of{I agree I think this works well.}
 We can also see from Figure~\ref{fig:Boxplots} that the reaction times clearly
 differ from year to year. The venue of the World Championships changes each 
-year, so weather or climate related factors such as humidity, precipitation, elevation, may indeed play a role, but additionally, both the technology and 
+year, so weather or climate related factors such as humidity, precipitation, 
+elevation, may indeed play a role, but additionally, both the technology and 
 false start penalization rules have changed over the study period. 
@@
 
-This dataset is too small to perform any meaningful statistical analysis, but
-the data were expanded to include reaction times for athletes who competed at 
-national competitions (many of which were held from May-July of 2022). Once the 
-data was collected, each athlete was designated as a cluster and the stage of
-competition (national or international) was specified for each competition 
-within the cluster (athlete).  \eds{this is where you can put the relevant 
+\eds{this is where you can put the relevant 
 information about the other competitions that was commented out in the 
@@
 edit this paragraph now so it all works together.}
-First we can compare the reaction times of athletes who
+The dataset of the four United States athletes 
+is too small to perform any meaningful statistical analysis, but
+the data was expanded to include reaction times for athletes who competed at 
+national competitions (many of which were held from May-July of 2022). Once the 
+data was collected, each athlete was designated as a cluster and the stage of
+competition (national or international) was specified for each competition 
+within the cluster (athlete). First we can compare the reaction times of athletes who
 competed at the World Track and Field Championships in both 2019 and 2022, and 
-second we can compare how athletes performed at the 2022 championships relative 
+then we can compare how athletes performed at the 2022 championships relative 
 to their national/qualifiers meet. In order to reduce any chances

### 2023-05-21 Manuscript/manuscript.tex
Removed commented out sections, fixed other issues

@@
 seconds, 0.001 seconds faster than what is allowed. Following the games, 
-bloggers on the running website LetsRun %scrutinized the results %and performed 
-%non-statistical analysis of the data. They found 
-concluded that the reaction time data from the 2022 World Championships 
-seemed to be generally faster compared to the other datasets they considered, 
-but they did not perform any formal %meaningful 
+bloggers on the running website LetsRun concluded that the reaction time data 
+from the 2022 World Championships seemed to be generally faster compared to the 
+other datasets they considered, but they did not perform any formal 
 statistical analyses. This paper questions the reaction time 
@@
 this is a reasonable threshold. We employ a generalized linear mixed model 
-(GLMM) with a random %effect term which will be referred to as the 
-venue effect in order to %try and 
-model the reaction time data. Additionally, we employ a signed-rank test for 
-clustered data to 
+(GLMM) with a random venue effect in order to  model the reaction time data. 
+Additionally, we employ a signed-rank test for clustered data to 
 compare reaction times for the same athletes at different competitions. This 
-matter needs to be addressed because %Devon Allen's 
-disqualification based on allowable reaction time 
-%could be a 
-%problem that could 
-will continue to be an issue in future world championships.
-
-%\eds{removed Devon Allen from keywords because his name is already in title
-%and replaced with others}
-
-%\eds{also removed 2022 World Track and Field Championships from keywords and 
-%changed location of World in the title. World Track and Field Championships
-%has more Google search results than Track and Field World Championships, so
-%I opted to go with this. Also used plural Championships. Feel free to switch 
-%order or use singular if one or both of these is 
-%preferred for some reason.  }
+matter needs to be addressed because disqualification based on allowable 
+reaction time will continue to be an issue in future world championships.
 
@@
 2022 World Track and Field Championships at the same Eugene venue in August.
-%After coming in third in the 110 meter hurdles at the United States Track and 
-%Field Championships also in Eugene at the end of June, Allen was poised to 
-%compete in the 2022 World Track and Field Championships also 
-%located at Eugene in August.
-%two months later. 
 At the World Championships, Allen made it through the preliminary
@@
 0.123 and 0.101 in his preliminary heat and semifinal, respectively.  
-%Allen, who attended the University of Oregon also in Eugene, 
-%(the location of US World Championships 
-%\eds{US World? remove World?} and the World Championship) 
-%was
-%considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
-%in early June; only 0.04 seconds away from the world record \citep{wa2022preview}.
-%Thus 
 His disqualification in the final heat by reacting in 0.099 seconds instead of 
@@
 \citep{johnson2022was}.
-%came from Johnson claiming there is a $(1/13)^3$ chance that the fastest 
-%reaction times came during the same championships for the men's 100 meter,
-%men's 110 meter hurdle, and women's 100 meter for 13 competitions held between 2003 and 2022 
-%\of{Cleaned up but may need additional work}
-%\of{Cleaned this up a little but still a little messy.}
-%\eds{is this better? (and still correct?)}
-%\of{Yes, I think it looks good}
-
-%3)
-%The International Association of Athletics Federations (also referred to
-%as IAAF or World Athletics), which is the governing board for international running
-%competitions, commissioned a study in 2009 regarding the reaction time 
-%barrier and whether the threshold should be lowered. 
-%The paper analyzed sprinters reaction times and the biological processes 
-%that went into reacting to the starting gun and concluded 
-%that athletes can react in as low as 0.08 seconds.  
-%One study by Pain looked at
+
 Reaction time in sprint and hurdle events has been studied by many authors.
@@
 
-%This will
-%be important when considering the results in section \ref{sec:Results}.
-
 Despite their study, IAAF has not changed the reaction barrier since it was
@@
 besides Allen who were disqualified.
-%Reaction time barrier has been examined before: cite reviews
-%Then go into specifics and talk about people who wanted to change barrier
-%General reviews are important
-%People have studied the biological processes, cite a few references
-%How barrier is picked by regulator
-%Read abstract and into and summarize into about one sentence what each paper contributes
-%1 and 2 can be merged but 3 need to stand alone
-
-%4)
+
 The objective to this paper is to analyze the reaction times at the 2022 World 
@@
 and covers the men's 110 meter hurdles from 1999 to 2022; the source can be 
-found in the Appendix. %~\ref{sec:Appendix}.   %\eds{see earlier comment - 
-%are you using any of these as covariates?}
-%\of{removed covariate confusion, talk about womens a couple sentences down} 
+found in the Appendix.
 Although women's data is available, we chose not to include women's data in the
@@
 babicc2009reaction, panoutsakopoulos2020gender}.
-%The variable ``Stage" refers to whether the observation occurred during a 
-%heat (preliminary round)  
-%semifinal heat (of which there are either two
-%or three each year) denoted by ``S", or final heat denoted by ``F".
 We focus on the reaction times during the semifinal and final heats only, as 
 reaction times from preliminary heats are often not as fast as those in later 
-heats %were not studied because athletes react faster as the 
-%competition progresses 
-\citep[e.g.,][]{zhang2021correlation}. 
-%The variable 
-%``Gender" has levels ``M" and ``F" for ``male" and ``female" respectively.  
-%The variable ``Batch" takes a numerical value and every heat of every round from 
-%1999 to 2022 was assigned it's
-%own number.  This was added because of suspected issues that reaction time may be
-%correlated with other runners reacting quickly (one runner reacting quickly may
-%result in other runners reacting quickly). There is no statistical meaning to
-%how the batches were labeled, they were ordered chronologically by meet but
-%reverse chronologically by year (Batch 1 is a preliminary heat of 2022 and Batch
-%111 corresponds to the 1999 finals). %Included are
-%box plots of the reaction times of the men's finals and the men's semifinals and
-%finals pooled together.  
+heats \citep[e.g.,][]{zhang2021correlation}.
+
 For analysis purposes, we will often pool reaction times of men's final and 
 men's semifinal heats together and refer to this data as the ``pooled data"
-%We will refer to the latter as the ``pooled data"
-throughout the paper.  %The heats or preliminary rounds were not studied because
- % Athletes react 
-%faster in the finals than the semifinals which are faster than the heats 
-%\citep{zhang2021correlation}.  
-%\eds{if you are really not using ``H'' data at all, it does not appear necessary
-%to spend time defining earlier.  You could just say directly when you define 
-%stage that you only consider semifinals and finals and cite the Zhang paper for 
-%rationale}
+throughout the paper. 
 %\of{Ok I updated this.  Should I delete the batch variable too?  Although I don't
@@
 false start penalization rules have changed over the study period.
-%Something that is extremely important to note in the data is the inconsistency 
-%of World Athletics in terms of how they penalize false starts.  
-From 2007--2009,
-the IAAF (the former name for World Athletics, and the governing body for the 
-World Track and Field Championships) instituted a rule change that allowed one 
-warning false start before a sprinter was disqualified \citep{iaaf2009falsestart}. 
-%This was desirable for World Athletics as false starts can make already long 
-%track meets more tedious, both for athletes and viewers watching the television 
-%broadcast.  
-%\eds{I think we lost something here with all of the edits so I removed the 
-%previous sentence. What was desirable for World Athletics -- the 2007-2009 rule
-% change?  Why was this desirable?  The rule change would seem to make the event
-% longer and not shorter? } 
-There were 18 male and 7 female false starts at the 2007 World 
-Championships, and 18 male and 7 female false starts at the 2009 World 
-Championships. 
-Starting in 2011, the rule was scrapped, and the old policy which automatically 
-disqualifies runners who false start was reinstated. By returning to the harsher 
-policy and cracking down on false starts, World Athletics had reduced men's 
-false starts by two thirds in 2011 (6 male and 4 female disqualifications) 
-\citep{iaaf2009falsestart}.
-%The effect? At the 2011 
-%World Championships there were 6 male disqualifications and 4 female 
-%disqualifications \citep{iaaf2009falsestart}.  
-%One study that looked at different false start rules that IAAF has
-%imposed from 1997 to 2009 and found that the reaction times when there was less 
-%leniency improved by a statistically significant amount
-%\citep{haugen2013effect}. 
-\citet{haugen2013effect} examined the effect of different false start
-rules that IAAF has imposed from 1997 to 2009 and found statistically 
-significant improvement in reaction times under more lenient rules.
+
+From 2007--2009, the IAAF (the former name for World Athletics, and the 
+governing body for the World Track and Field Championships) instituted a rule 
+change that allowed one warning false start before a sprinter was disqualified 
+\citep{iaaf2009falsestart}. There were 18 male and 7 female false starts at the 
+2007 World Championships, and 18 male and 7 female false starts at the 2009 World 
+Championships. Starting in 2011, the rule was scrapped, and the old policy which
+automatically disqualifies runners who false start was reinstated. This was 
+desirable for World Athletics as false starts can make already lengthy track 
+meets more tedious, both for athletes and viewers watching the television 
+broadcast. By returning to the harsher policy and cracking down on false starts,
+World Athletics had reduced men's false starts by two thirds in 2011 (6 male and
+4 female disqualifications) \citep{iaaf2009falsestart}. \citet{haugen2013effect}
+examined the effect of different false start rules that IAAF has imposed from 
+1997 to 2009 and found statistically significant improvement in reaction times 
+under more lenient rules.
 
@@
 2022 United States Track and Field Championships.
-%\eds{be consistent with how you refer to the US championship:
-%``United States Track and Field Championships" and ``United States Track and Field 
-%Championships" are both being used; please pick one and use that throughout}
-%\of{Resolved; changed all USA to United States.}
-From June 23-26 2022, the United States held its Track
-and Field Championships at Hayward 
-Field in Eugene, Oregon to decide who to send to the World Championships 
-being held at the same venue in August. %the exact same venue as used 
-%for the World Track and Field Championships a month later.  
-%Thus, a baseline is able to be established for the four United States athletes who 
-Thus, we can establish a baseline for the four United States athletes who 
+
+From June 23-26 2022, the United States held its Track and Field Championships 
+at Hayward Field in Eugene, Oregon to decide who to send to the World Championships 
+being held at the same venue in August. Thus, we can establish a baseline for the four United States athletes who 
 competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
@@
 
-%\of{Im not sure how to fix this but the spacing between the lines of the
-%table is different than the rest of the paper}
-%\eds{LaTeX keeps captions singlespaced automatically with the doublespace 
-%command; don't worry about it!}
-
 \eds{this is where you can put the relevant 
@@
 edit this paragraph now so it all works together.}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics{2019vs2022BarGraph}
+  \caption{The reaction times from 2019 and 2022}
+  \label{fig:2019vs2022Graph}
+\end{figure}
+
 The dataset of the four United States athletes 
@@
 
-The %World Championships data and the data eyond the World Championships 
-data described in Sections~\ref{sec:dataworld} and~\ref{sec:databeyond} were
-analysed with GLMM and rank-based methods, respectively.
-%\eds{need consistency with ``Championship" or ``Championships - I am not sure 
-%which one is correct.  I am using Championships (plural) but
-%change everything to singular if that is preferred}
-%\of{I believe that championships is correct, I have changed all instances of
-%championship to championships}
+The data described in Sections~\ref{sec:dataworld} and~\ref{sec:databeyond} were
+analyzed with GLMM and rank-based methods, respectively.
 
@@
 link function worked better for our data and additionally ensures the positivity
-of the conditional mean.
-% The link function $g$ can be the logarithm function to
-% ensure the positivity of the conditional mean.
-%\eds{Did you try an identity link also? Lo and Andrews (2015) suggest that the 
-%Gamma family with an identity link works well for their reaction-time data.
-%It doesn't seem like a good idea in general because it is likely that some 
-%negative predicted responses will occur either during model fitting or in the 
-%final model, but the internet seems to favor this link so a reviewer may ask 
-%about it too.} \of{I have added code at the bottom of Code/Models.Rmd that
-%compared the link = "identity" and link="log" for each dataset.  I found that
-%the link=log was marginally better.}
-%We include a random effect for venue (or year) because the technology
-%that World Athletics has used has differed since 1999, and theoretically the
-%reaction time data from 2022 should be the best as the timing system should be
-%state of the art. Additionally, any other weather or climate related factors 
-%such as humidity, precipitation, elevation, could be captured by this effect.
-%\eds{I moved this info into the Data section.}
-The gamma mixed-effect model can be fit with the \texttt{glmer()} function in R
-package \texttt{lme4} \citep{lme4}.
+of the conditional mean. The gamma mixed-effect model can be fit with the 
+\texttt{glmer()} function in R package \texttt{lme4} \citep{lme4}.
 
@@
 reaction times for everyone involved in the heat.
-%From our exploration, however, the race effect is very difficult to explain or 
-%interpret as it means that some heats of races were either
-%faster or slower and that everyone in the race reacted similarly.  
-Although
-it is possible that one athlete reacting faster could lead to others reacting
+
+Although it is possible that one athlete reacting faster could lead to others reacting
 similarly, this does not make much practical sense as the difference between
@@
 data to see if it makes a difference?}
-
-%% In this case, athletes who have more
-%% data are generally thought to be faster, as they have made it further in the
-%% competition.  Clusters range from two to six, with two indicating that at each
-%% competition they only made it to the heats, but a cluster size of six indicates
-%% that they made the finals in both competitions.  While the number of times
-%% athletes compete is based off their overall time not their reaction time, the
-%% two are correlated \citep{delalija2008reaction}.
-%% Group is either designated to be 1 or 2, with 1 corresponding to either the
-%% national level data or the previous world championships data and 2 corresponding
-%% to the 2022 data.
 

### 2023-05-21 Code/RankComparison.Rmd
Beginning of descriptive statistics/graphs for the rank comparison part

---
title: "Descriptive_Stats"
author: "Owen Fiore"
date: "2023-05-21"
output: pdf_document
---

```{r}
inter <- read.csv("../Data/2019vs2022.csv")

#nine is used for 2019 and two is used for 2022.  H, S, F are used to denote the heats, semifinals, and finals.
colnames(inter) <- c("Athlete", "Country", "nineH", "nineS", "nineF", "twoH", "twoS", "twoF")
```

```{r}
my_sums <- colSums(!is.na(inter))
print(my_sums)
barplot(my_sums)
```

```{r}
means <- sapply(inter[,3:ncol(inter)], mean, na.rm = TRUE)
print(means)

```

```{r}
library(ggplot2)
means_df <- data.frame(variable = names(means), mean = means)

# Define the grouping of bars
groupings <- list(c("nineH", "twoH"), c("nineS", "twoS"), c("nineF", "twoF"))

# Convert variable to a factor with desired order
means_df$variable <- factor(means_df$variable, levels = unlist(groupings))

# Create the bar graph
ggplot(means_df, aes(x = variable, y = mean, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("sienna2", "tan", "sienna2", "tan", "sienna2", "tan")) +
  labs(x = "Variable", y = "Mean", title = "Grouped Bar Graph") +
  theme_minimal() +
  theme(legend.position = "none")
```

### 2023-05-21 Code/RankComparison.Rmd
Changed variable names

@@
 #nine is used for 2019 and two is used for 2022.  H, S, F are used to denote the heats, semifinals, and finals.
-colnames(inter) <- c("Athlete", "Country", "nineH", "nineS", "nineF", "twoH", "twoS", "twoF")
+colnames(inter) <- c("Athlete", "Country", "2019H", "2019S", "2019F", "2022H", "2022S", "2022F")
 ```
@@
 # Define the grouping of bars
-groupings <- list(c("nineH", "twoH"), c("nineS", "twoS"), c("nineF", "twoF"))
+groupings <- list(c("2019H", "2022H"), c("2019S", "2022S"), c("2019F", "2022F"))
 
@@
   scale_fill_manual(values = c("sienna2", "tan", "sienna2", "tan", "sienna2", "tan")) +
-  labs(x = "Variable", y = "Mean", title = "Grouped Bar Graph") +
+  labs(y = "Mean Reaction Time", x ="") +
   theme_minimal() +
   theme(legend.position = "none")
+ggsave(filename="2019vs2022BarGraph.pdf", path = "../Manuscript", height=4, width=6)
 ```

### 2023-05-21 Manuscript/manuscript.tex
Minor edits

@@
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics{2019vs2022BarGraph}
-  \caption{The reaction times from 2019 and 2022}
-  \label{fig:2019vs2022Graph}
-\end{figure}
-
 The dataset of the four United States athletes 
@@
 competition, but the time span between meets has been reduced considerably.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics{2019vs2022BarGraph}
+  \caption{The reaction times from 2019 and 2022 World Track and Field
+  Championship. ``H'', ``S'', and``F'' refer to the heats, semifinals, and finals
+  respectively.}
+  \label{fig:2019vs2022Graph}
+\end{figure}
+\of{I included this here for now, but feel free to change the location.  I will
+add references to it and explanation of it along with some descriptive statistics about
+the size and mean/median,std dev, size of the data.}
 

### 2023-05-31 Manuscript/manuscript.tex
Major changes before Tuesday meeting

@@
 were only five data points due to two disqualifications and one athlete who 
-did not compete.
+did not compete.  We also look at data that does not include 2022 with year values
+between 1999 and 2019.  Thus, we have four datasets with their respective naming
+in parentheses: the final's data from 1999 to 2019 (Old Finals), final's data 
+from 2019 to 2022 (All Finals), pooled data from 1999 to 2019 (Old Pooled), and 
+pooled data from 1999 to 2022 (All Pooled).
 
@@
 would be more compelling for its non-use than its difficulty in interpretation.}
+\of{We found that including the race effect resulted in the venue effect not
+being significant, and thus Prof Yan suggested that we don't include it.  I am
+wondering if this is even worth mentioning? Prof Yan, what do you think?}
 
@@
 data to see if it makes a difference?}
+\of{Yes I am including prelminary heat data.  The primary reason is that there
+would be few datapoints otherwise.  I ran the numbes without prelims, the number
+of observations dropped from 65 to 26 and the number of clusters from 15 to 9, the
+p value increased from 0.000155 to 0.007332.
+The calculations are at the bottom of RankComparison.Rmd}
 
@@
 We may want to consider using all four data sets for this problem because
-comparing all the data to the data excluding 2022 can give us a qualitative
-assessment of how impactful 2022 was.
+comparing all the data to the data excluding 2022 can give us an
+assessment of how impactful 2022 was. We hope to see that by including 2022,
+the probability of observing a significantly low reaction time increases. 
 \eds{Agree with JY, but you need to describe what these four datasets are.}
-
+\of{I added a description of the four datasets in the Data paragraph right below
+where it says <Thus, we have four datasets>}
 Table~\ref{tab:Gamma_parameters} summarizes the results from the gamma model 
-from different data sets.
-
-As shown in Table~\ref{tab:Gamma_parameters}, the standard error of the year
+from different data sets and shows the standard error of the year
 effect increases for both the men's final and men's pooled data sets when
@@
   \centering
-  \caption{Summary statistics for the GLMM on various data sets. 
-	\eds{Int=intercept?}}
+  \caption{Summary statistics for the GLMM on various data sets.} 
   \begin{tabular}{c c c c c c}
@@
       \midrule
-      Older Finals & 0.0439 & $-1.901$ & 0.0251 & 0.0019 & 0.0109 \\
+      Old Finals & 0.0439 & $-1.901$ & 0.0251 & 0.0019 & 0.0109 \\
       All Finals & 0.0504 & $-1.914$ & 0.0280 & 0.0025 & 0.0112 \\
-      Older Pooled & 0.0495 & $-1.853$ & 0.0305 & 0.0024 & 0.0243 \\
+      Old Pooled & 0.0495 & $-1.853$ & 0.0305 & 0.0024 & 0.0243 \\
       All Pooled & 0.0554 & $-1.869$ & 0.0348 & 0.0031 & 0.0235 \\
@@
 
-\eds{be consistent with the order of the rows between 
-Tables~\ref{tab:Gamma_parameters} and~\ref{tab:Sim_probability}}
-
 \jy{label the table and reference to it; clean it up}
@@
   \centering
-  \caption{Probabilities of observing reaction times less than 0.1, 0.09, and
+  \caption{Probabilities of observing reaction times less than 0.1, 0.08, and
   0.08 for various data sets.}
@@
    \toprule
-   Data Set & P(RxnTime $<$ 0.1) & P(RxnTime $<$ 0.09) & P(RxnTime $<$ 0.08) \\ 
+   Data Set & P(RxnTime $<$ 0.1) & P(RxnTime $<$ 0.08) \\ 
    \midrule
-   all finals & $8.56\cdot10^{-4}$ & $4.1\cdot10^{-5}$ & $6.0\cdot10^{-7}$ \\ 
-   old finals & $4.39\cdot10^{-4}$ & $1.63\cdot10^{-5}$ & $2.0\cdot10^{-5}$ \\
-   all pooled & $6.86\cdot10^{-3}$ & $1.21\cdot10^{-3}$ & $1.3\cdot10^{-4}$ \\
-   old pooled & $5.76\cdot10^{-3}$ & $9.93\cdot10^{-4}$ & $1.0\cdot10^{-4}$\\
+   Old Finals & $4.39\cdot10^{-4}$ & $2.0\cdot10^{-5}$ \\
+   All Finals & $8.56\cdot10^{-4}$ & $6.0\cdot10^{-7}$ \\
+   Old Pooled & $5.76\cdot10^{-3}$ & $1.0\cdot10^{-4}$ \\ 
+   All Pooled & $6.86\cdot10^{-3}$ & $1.3\cdot10^{-4}$ \\
    \bottomrule
@@
 
+\begin{table}
+  \centering
+  \caption{Suggested reaction time barriers based on probability of observing
+  a time less the barrier. \of{Not sure if this makes sense}}
+  \begin{tabular}{c c c c} 
+   \toprule
+   Data Set & P(RxnTime $<$ x = .001) & P(RxnTime $<$ x = 0.0001) \\ 
+   \midrule
+   Old Finals & $4.39\cdot10^{-4}$ & $2.0\cdot10^{-5}$ \\
+   All Finals & $8.56\cdot10^{-4}$ & $6.0\cdot10^{-7}$ \\
+   Old Pooled & $5.76\cdot10^{-3}$ & $1.0\cdot10^{-4}$ \\ 
+   All Pooled & $6.86\cdot10^{-3}$ & $1.3\cdot10^{-4}$ \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:ExpectedValues}
+\end{table}
+
 Using the gamma function with link set to log, the probability of the reaction 
 time being less than 0.1 seconds can be calculated to see if Allen's 
-disqualification was an anomaly. \eds{explain how simulation was used here.}
-At a size of $n=10,000,000$ \eds{use different letter as $n$ was already used 
-for number of athletes, $B?$}, the calculated probabilities of a reaction time 
+disqualification was an anomaly. 
+\eds{explain how simulation was used here.} \of{The methods section already
+details how the calculations are performed, can you please elaborate what you
+are exactly looking for.}
+Calculations were performed in R using the methods described in 
+section~\ref{sec:glmm} with the use of the \citet{lme4} package.
+
+At a size of $n=10,000,000$, \eds{use different letter as $n$ was already used 
+for number of athletes, $B?$} \of{Could I just say "At a size of 10,000,000"? 
+Otherwise I am fine with using "B" instead of "n"}
+
+The calculated probabilities of a reaction time 
 being below 0.1 seconds for the various different data sets are shown in 

### 2023-05-31 Manuscript/manuscript.tex
minor comments/edits

@@
 
-\bigskip\noindent\sc{Keywords}:
+\bigskip\noindent{\sc Keywords}:
 False Start, Hurdles, Reaction Time, Seiko
@@
 the size and mean/median,std dev, size of the data.}
+\jy{Show dot plots which would give the distribution; without uncertainty
+  measure, the mean does not tell much.}
 
@@
   how is this athlete clustering being accounted for here?}
-\jy{We could add an athlete random effect too, if the players are identified in the data.}
+\jy{We could add an athlete random effect too, if the players are identified in
+  the data.}
+\jy{Owen, do we have athlete id in the data?}
 Conditional on a venue effect $z_i$ of year~i,  $Y_{ij}$ is modeled by 
@@
 wondering if this is even worth mentioning? Prof Yan, what do you think?} 
+\jy{We can revisit the results. It is not surprising that venue effect is no
+  longer significant after the batch effect was included. We could do the same
+  analysis with the batch-random-effect model.}
 
@@
 
-\jy{motivate first; describe the data (what are the grouping variables in the
-  two data sets? use one paragraph for each.}
-\of{Shouldnt describing the grouping variables go more in the data or methods 
-section?  Currently, there is some writing about this but it is commented out. It
-can be found in section 3.2}
+\jy{Remind the readers what the analysis was about: two-group comparison with
+  group being ....}
 

### 2023-05-31 Manuscript/manuscript.tex
Added reaction times table
update before tuesday meeting

@@
   0.08 for various data sets.}
-  \begin{tabular}{c c c c} 
+  \begin{tabular}{c c c} 
    \toprule
@@
   a time less the barrier. \of{Not sure if this makes sense}}
-  \begin{tabular}{c c c c} 
+  \begin{tabular}{c c c} 
    \toprule
@@
    \midrule
-   Old Finals & $4.39\cdot10^{-4}$ & $2.0\cdot10^{-5}$ \\
-   All Finals & $8.56\cdot10^{-4}$ & $6.0\cdot10^{-7}$ \\
-   Old Pooled & $5.76\cdot10^{-3}$ & $1.0\cdot10^{-4}$ \\ 
-   All Pooled & $6.86\cdot10^{-3}$ & $1.3\cdot10^{-4}$ \\
+   Old Finals & $0.103$ & $0.095$ \\
+   All Finals & $0.101$ & $0.093$ \\
+   Old Pooled & $0.090$ & $0.080$ \\ 
+   All Pooled & $0.089$ & $0.079$ \\
    \bottomrule
   \end{tabular}
-  \label{tab:ExpectedValues}
+  \label{tab:Sim_time}
 \end{table}
@@
 
+We can use the same model but alter our interpretation slightly to select a 
+reaction time barrier given the probability of observing a time less than the 
+barrier.  For example, if we want a 0.1\% chance of observing an illegally fast 
+reaction time for the men's pooled dataset that includes 2022, then the reaction
+time barrier should be 0.089 seconds, as shown in Table~\ref{tab:Sim_time}.  We
+can repeat this calculation for each dataset and for various probability levels,
+for example a probability of 0.0001 (one disqualification for every 10,000 starts)
+results in a reaction time barrier of 0.079 for the men's pooled dataset.  
+Tables~\ref{tab:Sim_time} shows that the inclusion of 2022 results in a lower
+reaction time barrier, further showing the claims above that the reaction times
+in 2022 were low.
+
 %Discussion should show that random effect
@@
   group being ....}
+
+The rank based methods described in Section~\ref{sec:rank} are used
+to determine if the reaction times for athletes who had competed at other 
+championships differed significantly from the 2022 World Championships.  We can
+use the 2019 World Championships and the 2022 national-level championships as a 
+precedent to see how athletes performance changed across different championships.
 

### 2023-05-31 Code/Plots.Rmd
Added another plot

@@
                                               breaks = seq(1997, 2022, 2))
-both <- grid.arrange(final,pooled, ncol=1)
-ggsave(filename="Finals_Pooled_Boxplot.pdf", path = "../Manuscript", plot=both, height = 6, width = 6)
+#both <- grid.arrange(final,pooled, ncol=1)
+grid.arrange(final,pooled, ncol=1)
+#ggsave(filename="Finals_Pooled_Boxplot.pdf", path = "../Manuscript", plot=both, height = 6, width = 6)
 ```

### 2023-05-31 Code/Simulation.Rmd
Code for finding reaction time based on probability

---
title: "Simulate"
author: "Owen Fiore"
date: "2023-05-30"
output: pdf_document
---

```{}
The probability of observing a reaction time less than 0.01 is 0.0068649
The probability of observing a reaction time less than 0.095 is 0.0030386
The probability of observing a reaction time less than 0.09 is 0.00121

If we want to pick the barrier based on the reaction time a barrier of 0.08905, 1 in every 1001.603
If we want to pick the barrier based on the reaction time a barrier of 0.089, 1 in every 1012.248
If we want to pick the barrier based on the reaction time a barrier of 0.089057, 1 in every 999.9
If we want to pick the barrier based on the reaction time a barrier of 0.089056, 1 in every 1000.2
```

```{r}
library(lme4)
set.seed(1)
simulation <- function(model, base, RxnTime){
    simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
    }
  sbase <- summary(base)
  #print(sbase)
  x <- simfit(model)
  print(1/mean(x < RxnTime))
  print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
  ## compare with and without random effect
  plot(density(x)) # with random effect
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
}
simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.089)
```

```{r}
library(lme4)
set.seed(1)
simulation <- function(model, base, prob){
    simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
    }
  sbase <- summary(base)
  #print(sbase)
  x <- simfit(model)
  print(mean(x < 0.089))
  #print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
  ## compare with and without random effect
  plot(density(x)) # with random effect
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
}
simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.01)
```

```{r}
library(lme4)
set.seed(1)

simulation <- function(model, base, prob) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    z <- rnorm(n, sd = vc$sdcor[1])
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))

  plot(density(x), main = "Density Plot", xlab = "RxnTime", ylab = "Density")
  curve(dgamma(x, shape = 1 / sbase$dispersion, scale = exp(coef(base)) * sbase$dispersion), 
        from = min(x), to = max(x), add = TRUE, col = "blue")
}

simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.001)
```

```{}
Currently mean(x < 0.089) returns a proportion of x values that are less than 0.089
We instead want to set the proportion to be some set value, ex: 0.001
```

### 2023-05-31 Code/RankComparison.Rmd
Added section about removing heats from ranked based comparison for 2019vs2022

@@
 ```
+
+```{r}
+nations <- read.csv("../Data/ClusrankNatvsWorld.csv")
+colnames(nations) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, alternative = "greater", method = "ds")
+```
+
+```{r}
+nations_noprelim <- read.csv("../Data/Clusrank2019vs2022PrelimsOnly.csv")
+colnames(nations_noprelim) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim, alternative = "greater", method = "ds")
+```

### 2023-05-31 Code/BatchEffect.Rmd
Analsysis of various effects

---
title: "Random_Effects"
author: "Owen Fiore"
date: "2023-05-31"
output: pdf_document
---

```{r}
two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
one_effect <- glmer(RxnTime ~ (1 | Year) , data = men_pooled, family = Gamma(link = "log"))
no_effect <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
batch <- glmer(RxnTime ~ (1 | Batch) , data = men_pooled, family = Gamma(link = "log"))
anova(batch, two_effect)
print("------------------")
anova(one_effect, two_effect)
print("--------------")
par(mfrow=c(1,1))
qqnorm(resid(two_effect))
par(mfrow=c(2,1))
qqnorm(ranef(two_effect)$Year[,1])
qqnorm(ranef(two_effect)$Batch[,1])
qqnorm(ranef(batch)$Batch[,1])
qqnorm(resid(one_effect), main = "Without Batch Effect")
#anova(one_effect, two_effect)
```

### 2023-05-31 Code/BatchEffect.Rmd
Updates from today's meeting

@@
 ```
+```{r}
+nested <- glmer(RxnTime ~ (1|Year) + (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
+anova(nested, one_effect)
+batch_year <- glmer(RxnTime ~ (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
+anova(batch_year, nested)
+ranef(nested)
+```

### 2023-05-31 Code/Simulation.Rmd
Updates from today's meeting

@@
 ```{}
-The probability of observing a reaction time less than 0.01 is 0.0068649
-The probability of observing a reaction time less than 0.095 is 0.0030386
-The probability of observing a reaction time less than 0.09 is 0.00121
 
@@
   x <- simfit(model)
-  print(1/mean(x < RxnTime))
+  print(mean(x < RxnTime))
   print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
@@
   print(quantile(x, probs = prob))
-
-  plot(density(x), main = "Density Plot", xlab = "RxnTime", ylab = "Density")
-  curve(dgamma(x, shape = 1 / sbase$dispersion, scale = exp(coef(base)) * sbase$dispersion), 
-        from = min(x), to = max(x), add = TRUE, col = "blue")
 }
 
-simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.001)
+#simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.001)
 ```
 
-```{}
-Currently mean(x < 0.089) returns a proportion of x values that are less than 0.089
-We instead want to set the proportion to be some set value, ex: 0.001
+```{r}
+set.seed(1)
+simulation(old_finals_year_gamma_log, old_finals_one_gamma_log, 0.0001)
+simulation(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.0001)
+simulation(old_pooled_year_gamma_log, old_pooled_one_gamma_log, 0.0001)
+simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.0001)
 ```

### 2023-06-01 Manuscript/manuscript.tex
removed resolved

@@
 react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
-\eds{are hurdlers generally considered a subset of sprinters or are they 
-considered a separate type of athlete?  Specifically did the study above 
-recommend the barrier be lowered for both sprinters and hurdlers?
-You also refer to sprinters elsewhere in the manuscript.  Are those instances
-meant to refer to both sprinters and hurdlers, as well?}
-\of{Yes I would say that hurdlers are a subset of sprinter.  The reaction time
-barrier for the men's 100 meters and men's 110 meter hurdle are both 0.1 seconds.
-The paper recommended for the reaction time barrier to be decreased to roughly 0.8
-seconds.  It does not reference any distinction between hurdlers/sprinters, so I
-think they can be grouped together.  In my personal opinion I don't see why the
-reaction time for hurdlers versus 100m dash would differ significantly.  In terms
-of what I wrote in the paper, I want to stay accurate to what is in the study
-but also generalize studies involving reaction times to include hurdlers.  I believe
-many of the studies examining the reaction time barrier from a physiological
-perspective to find the minimum possible reaction time do not deal with hurdles.
-Thus, I don't know if I can change what I wrote to be both accurate and understandable,
-but what do you both think?}
 Additionally, it was recommended to IAAF that high speed cameras be used to
@@
 compartmentalization?  Or should I instead say there are 3 types of data that are
-used in two different methods, or something like that?}
+used in two different methods, or something like that?} \eds{I prefer
+subsections within rank-based comparison.}
 The first data set is used for the GLMM analyses described in 
@@
 Data was taken from the World Athletics 
-%\eds{is there a website you can cite here?}
 and covers the men's 110 meter hurdles from 1999 to 2022; the source can be 
@@
 
-\eds{I wonder if this paragraph would be better placed in the introduction?}
-\of{I think it could fit in the brief section between the start of the Data section
-and the GLMM subsection in the paragraph starting with "There are two types..."}
-\eds{With the following newly added sentences, I think we can leave this paragraph here.}
-\of{I agree I think this works well.}
 We can also see from Figure~\ref{fig:Boxplots} that the reaction times clearly
@@
 
-\eds{this is where you can put the relevant 
-information about the other competitions that was commented out in the 
-introduction.}
 \eds{Can you show some descriptive statistics for this data?  Total number of 
@@
 athletes/clusters?}
-\of{Is the one below what you're talking about?}
-\eds{Yes, this is the part from the intro I was referring to, but you need to 
-edit this paragraph now so it all works together.}
 
+\eds{The following paragraph still needs to be smoothed (the paragraph that was 
+here and the paragraph that was in the intro still need to be integrated into 
+one (or two) cohesive paragraph(s) that accurately describe the two types of 
+data that will be used for two rank-based analyses.}
 The dataset of the four United States athletes 
@@
 competition (national or international) was specified for each competition 
-within the cluster (athlete). First we can compare the reaction times of athletes who
+within the cluster (athlete). First we can compare the reaction times of 
+athletes who
 competed at the World Track and Field Championships in both 2019 and 2022, and 
@@
 to their national/qualifiers meet. In order to reduce any chances
-of athletes improving reaction time, we can also look at data from qualifying meets.
-Almost every athlete had to qualify by proving they are elite at the national level, 
-usually by coming in at least fourth. Thus, data was compiled on athletes who competed
-at the 2022 World Track and Field Championships on what these athletes also ran 
-at their country's national championship. If there was any argument that athletes
+of athletes improving reaction time, we can also look at data from qualifying 
+meets. Almost every athlete had to qualify by proving they are elite at the 
+national level, usually by coming in at least fourth. Thus, data was compiled on 
+athletes who competed at the 2022 World Track and Field Championships on what 
+these athletes also ran at their country's national championship. If there was 
+any argument that athletes
 have improved their reaction time from 2019 to 2022, it should now be considered
 nearly null, as the national competitions took place between June and July 2022 
-and the World Championships were in August. The equipment used to measure reaction
-time cannot be guaranteed to be the same from national to international
+and the World Championships were in August. The equipment used to measure 
+reaction time cannot be guaranteed to be the same from national to international
 competition, but the time span between meets has been reduced considerably.
@@
 Let $Y_{ij}$ be the reaction time of observation~$j$ in year~$i$.
-\eds{we know that many of the athletes compete in multiple world championships,
-  how is this athlete clustering being accounted for here?}
-\jy{We could add an athlete random effect too, if the players are identified in
-  the data.}
-\jy{Owen, do we have athlete id in the data?}
+%\eds{we know that many of the athletes compete in multiple world championships,
+%  how is this athlete clustering being accounted for here?}
+%\jy{We could add an athlete random effect too, if the players are identified in
+%  the data.}
+%\jy{Owen, do we have athlete id in the data?}
 Conditional on a venue effect $z_i$ of year~i,  $Y_{ij}$ is modeled by 
@@
 \of{Add something like this to the start of the paragraph?}
-We may want to consider using all four data sets for this problem because
-comparing all the data to the data excluding 2022 can give us an
-assessment of how impactful 2022 was. We hope to see that by including 2022,
+We may want to consider analyzing all four data sets separately because
+comparing the model based on all the data to the model based on data excluding 
+2022 can provide an indication about the impact of the 2022 data.
+We hope to see that by including 2022 data,
 the probability of observing a significantly low reaction time increases. 
-\eds{Agree with JY, but you need to describe what these four datasets are.}
-\of{I added a description of the four datasets in the Data paragraph right below
-where it says <Thus, we have four datasets>}
 Table~\ref{tab:Gamma_parameters} summarizes the results from the gamma model 
@@
 time being less than 0.1 seconds can be calculated to see if Allen's 
-disqualification was an anomaly. 
-\eds{explain how simulation was used here.} \of{The methods section already
-details how the calculations are performed, can you please elaborate what you
-are exactly looking for.}
+disqualification was an anomaly as described in Section~\ref{sec:glmm}.
 Calculations were performed in R using the methods described in 
@@
 for number of athletes, $B?$} \of{Could I just say "At a size of 10,000,000"? 
-Otherwise I am fine with using "B" instead of "n"}
+Otherwise I am fine with using "B" instead of "n"} \eds{we may have lost the 
+rest of this sentence?}
 
@@
 being below 0.1 seconds for the various different data sets are shown in 
-Table~\ref{tab:Sim_probability}. Prior to the 2022 World Championships, there was 
-a 0.0439\% chance (based on the
+Table~\ref{tab:Sim_probability}. Prior to the 2022 World Championships, there 
+was a 0.0439\% chance (based on the
 simulated results) of observing a reaction time below 0.01 seconds in the men's
@@
 under this category) \citep{wa2013backtage}. Since 1999, the
-slowest \eds{highest as in fastest or slowest? Neither seems consistent with 
-the boxplots in Figure 1} \of{Resolved} reaction 
+slowest reaction 
 times in the men's 110-meter hurdle were recorded in 2013.  That is not to say

### 2023-06-05 Code/BatchEffect.Rmd
Added simulation code for 2 random effects under BatchEffect.Rmd

--- 2023-05-31 Code/BatchEffect.Rmd
@@
 ```{r}
+library(lme4)
 two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
@@
 ```
+
+```{r}
+nested <- glmer(RxnTime ~ (1|Year) + (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
+anova(nested, one_effect)
+batch_year <- glmer(RxnTime ~ (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
+anova(batch_year, nested)
+#ranef(nested)
+```
+
+```{r}
+simulation <- function(model, base, RxnTime) {
+  simfit <- function(model, n = 10000000) {
+    vc <- as.data.frame(VarCorr(model))
+    z1 <- rnorm(n, sd = vc$sdcor[1])  # random effects for Year
+    z2 <- rnorm(n, sd = vc$sdcor[2])  # random effects for Batch
+    mu <- exp(fixef(model) + z1 + z2)
+    shape <- 1 / vc$vcov[2]
+    x <- rgamma(n, shape = shape, scale = mu / shape)
+    x
+  }
+
+  sbase <- summary(base)
+  x <- simfit(model)
+  print(mean(x < RxnTime))
+  print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
+  ## compare with and without random effect
+  plot(density(x))  # with random effects
+  curve(dgamma(x, shape = 1 / sbase$dispersion, 
+               scale = exp(coef(base)) * sbase$dispersion), 
+        0.08, 0.22, add = TRUE, col = "blue")  # fixed effect only
+}
+simulation(old_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
+```
+```{}
+All probabilities are for RxnTime=0.1 seconds
+old pooled: 0.0283374 vs 0.0057613
+men_pooled: 0.0313561 vs 0.0068649
+old finals: 0.0056892 vs 0.0004389
+men_finals: 0.0083227 vs 0.000856
+The times on the left are for the 2 batch, times on the right are for the 1 batch
+```

### 2023-06-05 Code/Simulation.Rmd
Added simulation code for 2 random effects under BatchEffect.Rmd

@@
 }
-simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.089)
+simulation(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
 ```

### 2023-06-05 Code/BatchEffect.Rmd
Further updates, previous code was wrong

@@
 ```{r}
-nested <- glmer(RxnTime ~ (1|Year) + (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
+nested <- glmer(RxnTime ~ (1|Year) + (1|Batch), data=men_pooled, family = Gamma(link="log"))
 anova(nested, one_effect)
@@
 ```{r}
+old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_finals, family = Gamma(link = "log"))
+men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_finals, family = Gamma(link = "log"))
+old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
+men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
+
+```
+
+```{r}
 simulation <- function(model, base, RxnTime) {
@@
     vc <- as.data.frame(VarCorr(model))
-    z1 <- rnorm(n, sd = vc$sdcor[1])  # random effects for Year
-    z2 <- rnorm(n, sd = vc$sdcor[2])  # random effects for Batch
+    sd_year <- vc$sdcor[1]
+    sd_year_batch <- vc$sdcor[2]
+
+    z1 <- rnorm(n, sd = sd_year)
+    z2 <- rnorm(n, sd = sd_year_batch)
+
+    shape <- 1 / sqrt(sd_year^2 + sd_year_batch^2)
     mu <- exp(fixef(model) + z1 + z2)
-    shape <- 1 / vc$vcov[2]
+
     x <- rgamma(n, shape = shape, scale = mu / shape)
@@
   print(mean(x < RxnTime))
-  print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
+  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
   ## compare with and without random effect
@@
 }
-simulation(old_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
+simulation(old_finals_two_effect, old_finals_one_gamma_log, 0.1)
+simulation(men_finals_two_effect, men_finals_one_gamma_log, 0.1)
+simulation(old_pooled_two_effect, old_pooled_one_gamma_log, 0.1)
+simulation(men_pooled_two_effect, men_pooled_one_gamma_log, 0.1)
 ```
+
 ```{}
 All probabilities are for RxnTime=0.1 seconds
-old pooled: 0.0283374 vs 0.0057613
-men_pooled: 0.0313561 vs 0.0068649
-old finals: 0.0056892 vs 0.0004389
-men_finals: 0.0083227 vs 0.000856
+old finals: 0.0452759 vs 0.0004389
+men_finals: 0.0647663 vs 0.000856
+old pooled: 0.0779887 vs 0.0057613
+men_pooled: 0.0897533 vs 0.0068649
 The times on the left are for the 2 batch, times on the right are for the 1 batch

### 2023-06-07 Code/BatchEffect.Rmd
comments and correction to simulation code

@@
 ---
+
+Clearn it up: two_effect and nested are the same model. Don't confuse the readers.
+
+Keep the code styles (80 characters linewidth; etc.)
+
+Add text between the code chunks to document what you are doing.
 
@@
 
-    shape <- 1 / sqrt(sd_year^2 + sd_year_batch^2)
+    shape <- 1 / sigma(model)^2
     mu <- exp(fixef(model) + z1 + z2)

### 2023-06-07 Manuscript/manuscript.tex
comments and correction to simulation code

@@
 championships in the United States, later uploaded a video to their YouTube
-channel that showed the disqualification; it can be found here:
-\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers 
+channel that showed the disqualification
+(\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}).
+The crowd erupted with jeers
 when they found out that Allen was disqualified, angry at the result.  Had Allen 
@@
 \citep{johnson2022was}.
+
+\jy{Use double blank lines to separate paragraphs in tex source}
 
@@
 
+\jy{This paragraph reads disconnected. Should it be merged with paragraph~3
+  where the study was reviewed?}
 Despite their study, IAAF has not changed the reaction barrier since it was

### 2023-06-26 Code/Plots.Rmd
Small changes

@@
 ```
+
+```{r}
+library(ggplot2)
+means_df <- data.frame(variable = names(means), mean = means)
+
+# Define the grouping of bars
+groupings <- list(c("2019H", "2022H"), c("2019S", "2022S"), c("2019F", "2022F"))
+
+# Convert variable to a factor with desired order
+means_df$variable <- factor(means_df$variable, levels = unlist(groupings))
+
+# Create the bar graph
+ggplot(means_df, aes(x = variable, y = mean, color = variable)) +
+  geom_bar(size = 3) +
+  labs(y = "Mean Reaction Time", x = "") +
+  scale_color_manual(values = c("sienna2", "tan", "sienna2", "tan", "sienna2", "tan")) +
+  theme_minimal() +
+  theme(legend.position = "none")
+#ggsave(filename = "2019vs2022ScatterPlot.pdf", path = "../Manuscript", height = 4, width = 6)
+```
+
+```{r}
+comp <- read.csv("../Data/2019vs2022.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(comp) <- c("Athlete", "Country", "2019 Heats", "2019 Semifinals", "2019 Finals", "2022 Heats", "2022 Semifinals", "2022 Finals")
+```
+
+```{r}
+comp <- comp[, -c(1,2)]
+```
+```{r}
+library(tidyr)
+library(ggplot2)
+df_long <- tidyr::pivot_longer(comp, cols = everything(), names_to = "Column", values_to = "Time")
+
+order <- c("2019 Heats", "2019 Semifinals", "2019 Finals", "2022 Heats", "2022 Semifinals", "2022 Finals")
+#df_long$Event <- factor(df_long$Event, levels = order)
+```
+```{r}
+comp_plot <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Time") +
+  ylim(0.08, 0.22)
+  theme_minimal()
+ggsave(filename="Scatterplot2019vs2022.pdf", path = "../Manuscript", plot=comp_plot, height = 4, width = 6)
+```

### 2023-06-27 Code/BatchEffect.Rmd
Made code more readable, added comments, fixed sigma issue

@@
 two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
-one_effect <- glmer(RxnTime ~ (1 | Year) , data = men_pooled, family = Gamma(link = "log"))
+year_effect <- glmer(RxnTime ~ (1 | Year) , data = men_pooled, family = Gamma(link = "log"))
 no_effect <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
-batch <- glmer(RxnTime ~ (1 | Batch) , data = men_pooled, family = Gamma(link = "log"))
-anova(batch, two_effect)
-print("------------------")
-anova(one_effect, two_effect)
-print("--------------")
-par(mfrow=c(1,1))
-qqnorm(resid(two_effect))
-par(mfrow=c(2,1))
-qqnorm(ranef(two_effect)$Year[,1])
-qqnorm(ranef(two_effect)$Batch[,1])
-qqnorm(ranef(batch)$Batch[,1])
-qqnorm(resid(one_effect), main = "Without Batch Effect")
-#anova(one_effect, two_effect)
+batch_effect <- glmer(RxnTime ~ (1 | Batch) , data = men_pooled, family = Gamma(link = "log"))
+```
+
+```{}
+The code chunk above creates 4 models: no random effects, a year effect only, 
+batch effect only, and both year and batch effect.
 ```
@@
 ```{r}
-nested <- glmer(RxnTime ~ (1|Year) + (1|Batch), data=men_pooled, family = Gamma(link="log"))
-anova(nested, one_effect)
-batch_year <- glmer(RxnTime ~ (1|Batch:Year), data=men_pooled, family = Gamma(link="log"))
-anova(batch_year, nested)
-#ranef(nested)
+anova(batch, two_effect)
+```
+```{}
+This anova test determined that the two_effect model is an improvement over the
+batch effect model. The test statistic is significant at alpha = 0.05 but not at
+alpha = 0.01.
+```
+
+```{r}
+anova(one_effect, two_effect)
+```
+```{}
+In this anova test, it is clear that the two random effect model is superior to
+having only the year effect.
 ```
@@
 
+```{}
+This section fits a two random effect model for each dataset.
+```
+
 ```{r}
-simulation <- function(model, base, RxnTime) {
+set.seed(1)
+library(lme4)
+simulation_two_effect <- function(model, base, RxnTime) {
   simfit <- function(model, n = 10000000) {
@@
 }
-simulation(old_finals_two_effect, old_finals_one_gamma_log, 0.1)
-simulation(men_finals_two_effect, men_finals_one_gamma_log, 0.1)
-simulation(old_pooled_two_effect, old_pooled_one_gamma_log, 0.1)
-simulation(men_pooled_two_effect, men_pooled_one_gamma_log, 0.1)
+simulation_two_effect(old_finals_two_effect, old_finals_one_gamma_log, 0.1)
+simulation_two_effect(men_finals_two_effect, men_finals_one_gamma_log, 0.1)
+simulation_two_effect(old_pooled_two_effect, old_pooled_one_gamma_log, 0.1)
+simulation_two_effect(men_pooled_two_effect, men_pooled_one_gamma_log, 0.1)
+```
+
+```{}
+This code simulates the probability of observing a reaction time less than 0.1
+for the various datasets.
+```
+
+```{r}
+library(lme4)
+set.seed(1)
+simulation_one_effect <- function(model, base, RxnTime){
+    simfit <- function(model, n = 10000000) {
+    vc <- as.data.frame(VarCorr(model))
+    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
+    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
+    mu <- exp(fixef(model) + z)
+    shape <- 1 / vc$vcov[2]
+    x <- rgamma(n, shape = shape, scale = mu / shape)
+    x
+    }
+  sbase <- summary(base)
+  #print(sbase)
+  x <- simfit(model)
+  print(mean(x < RxnTime))
+  #print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
+  ## compare with and without random effect
+  plot(density(x)) # with random effect
+  curve(dgamma(x, shape = 1 / sbase$dispersion, 
+               scale = exp(coef(base)) * sbase$dispersion), 
+        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
+}
+simulation_one_effect(old_finals_year_gamma_log, old_finals_one_gamma_log, 0.1)
+simulation_one_effect(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.1)
+simulation_one_effect(old_pooled_year_gamma_log, old_pooled_one_gamma_log, 0.1)
+simulation_one_effect(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
+```
+
+```{}
+The above code simulates the probability of observing a reaction time less than
+0.1 seconds for the year effect model.
 ```
@@
 All probabilities are for RxnTime=0.1 seconds
-old finals: 0.0452759 vs 0.0004389
-men_finals: 0.0647663 vs 0.000856
-old pooled: 0.0779887 vs 0.0057613
-men_pooled: 0.0897533 vs 0.0068649
+old finals: 0.0004356 vs 0.0004389
+men_finals: 0.0008642 vs 0.0008626
+old pooled: 0.0041495 vs 0.0057287
+men_pooled: 0.005145 vs 0.0068441
 The times on the left are for the 2 batch, times on the right are for the 1 batch
+
+Thus if we want to pick the model with the larger probabilities to show that 
+0.01 is not extreme, the one random effect probabilities are slightly smaller.
 ```

### 2023-06-27 Manuscript/manuscript.tex
added graph
Started responding to prompts and breaking Rank comparison into 2 smaller subsections

@@
 
-\subsection{World Championships}\label{sec:dataworld}
+\subsection{2022 World Championships}\label{sec:dataworld}
 
@@
 
-\subsection{Beyond World Championships}\label{sec:databeyond}
+\subsection{Beyond 2022 World Championships}\label{sec:databeyond}
 The motivation for looking beyond World Championships data came from an 
@@
 
+\subsection{2022 National Competitions}\label{sec:datanational}
 From June 23-26 2022, the United States held its Track and Field Championships 
@@
 Field Championships compared to the United States Track and Field Championships, 
-as were his
-teammates. 
+as were his teammates. 
 This comparison suggests that the timing system used between the two
@@
 \end{table}
-
-\eds{Can you show some descriptive statistics for this data?  Total number of 
-clusters? Average number of international vs national competitions across 
-athletes/clusters?}
-
-\eds{The following paragraph still needs to be smoothed (the paragraph that was 
-here and the paragraph that was in the intro still need to be integrated into 
-one (or two) cohesive paragraph(s) that accurately describe the two types of 
-data that will be used for two rank-based analyses.}
 The dataset of the four United States athletes 
@@
 competition (national or international) was specified for each competition 
-within the cluster (athlete). First we can compare the reaction times of 
+within the cluster (athlete).
+
+\eds{Can you show some descriptive statistics for this data?  Total number of 
+clusters? Average number of international vs national competitions across 
+athletes/clusters?}
+\eds{The following paragraph still needs to be smoothed (the paragraph that was 
+here and the paragraph that was in the intro still need to be integrated into 
+one (or two) cohesive paragraph(s) that accurately describe the two types of 
+data that will be used for two rank-based analyses.}
+
+\subsection{2019 World Track and Field Championships}
+The 2019 World Track and Field Championship data is included in the larger dataset
+described in section~\ref{sec:databeyond}, but we can also use it similarly to
+how we used national level in section~\ref{sec:datanational}.  We can compare
+the times of athletes who participated in the 2019 and 2022 World Track and Field
+Championships to see if there were lower times in 2022.  While it is theoretically
+possible that athletes got faster between 2019 and 2022, it is important to
+remember that these are some of the fastest sprinters on the planet, and thus
+improving on elite times is not easy.  Nonetheless, this is a worthy inclusion
+to the data because Seiko was the timing company responsible for both events,
+the World Championships have the biggest stakes, and because
+
+First we can compare the reaction times of 
 athletes who
@@
   \centering
-  \includegraphics{2019vs2022BarGraph}
-  \caption{The reaction times from 2019 and 2022 World Track and Field
-  Championship. ``H'', ``S'', and``F'' refer to the heats, semifinals, and finals
-  respectively.}
-  \label{fig:2019vs2022Graph}
+  \includegraphics{Scatterplot2019vs2022}
+  \caption{The reaction times from athletes who competed at least once in both 
+  the 2019 and 2022 World Track and Field Championships}
+  \label{Scatterplot2019vs2022}
 \end{figure}

### 2023-06-28 Manuscript/manuscript.tex
small changes

@@
 \end{table}
+
 The dataset of the four United States athletes 
@@
 improving on elite times is not easy.  Nonetheless, this is a worthy inclusion
-to the data because Seiko was the timing company responsible for both events,
-the World Championships have the biggest stakes, and because
-
-First we can compare the reaction times of 
-athletes who
-competed at the World Track and Field Championships in both 2019 and 2022, and 
-then we can compare how athletes performed at the 2022 championships relative 
-to their national/qualifiers meet. In order to reduce any chances
-of athletes improving reaction time, we can also look at data from qualifying 
-meets. Almost every athlete had to qualify by proving they are elite at the 
-national level, usually by coming in at least fourth. Thus, data was compiled on 
-athletes who competed at the 2022 World Track and Field Championships on what 
-these athletes also ran at their country's national championship. If there was 
-any argument that athletes
-have improved their reaction time from 2019 to 2022, it should now be considered
-nearly null, as the national competitions took place between June and July 2022 
-and the World Championships were in August. The equipment used to measure 
-reaction time cannot be guaranteed to be the same from national to international
-competition, but the time span between meets has been reduced considerably.
+to the data because Seiko was the timing company responsible for both events and
+because the level of competition is the extremely elite in both cases.
 

### 2023-06-28 Manuscript/manuscript.tex
more small changes to end of Data section

@@
   the 2019 and 2022 World Track and Field Championships}
-  \label{Scatterplot2019vs2022}
+  \label{fig:Scatterplot2019vs2022}
 \end{figure}
@@
   measure, the mean does not tell much.}
+
+Figure~\ref{fig:Scatterplot2019vs2022} shows the distribution of times from the
+2019 and 2022 World Championships.  This data was only for athletes who competed
+in at least one race at each championship, however it is possible for athletes
+to have competed in one race in 2019 and two races in 2022.  The 2022 World
+Championship Heats and Semifinals both had on average lower reaction times than
+their respective 2019 races, and the 2022 finals were substantially faster.  Also
+worth noting is that the fastest reaction time in the 2022 Finals and 2022
+Semifinals both came from Devon Allen, but the difference between 0.101 and 0.009
+seconds was the difference in what caused him to be disqualified.
 

### 2023-07-05 Code/BatchEffect.Rmd
Clarified models

@@
 ```{r}
-library(lme4)
-two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
-year_effect <- glmer(RxnTime ~ (1 | Year) , data = men_pooled, family = Gamma(link = "log"))
-no_effect <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
-batch_effect <- glmer(RxnTime ~ (1 | Batch) , data = men_pooled, family = Gamma(link = "log"))
+times <- read.csv("../Data/ThesisData.csv")
+colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
+men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
+men_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
+men_old_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ' & Year < 2022)
+men_old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Year < 2022)
 ```
-
 ```{}
-The code chunk above creates 4 models: no random effects, a year effect only, 
-batch effect only, and both year and batch effect.
+This loads in the datasets
 ```
@@
 ```{r}
-anova(batch, two_effect)
+library(lme4)
+men_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_finals, family = Gamma(link = "log"))
+men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "log"))
+
+men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, family = Gamma(link = "log"))
+men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma(link = "log"))
+
+men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, family = Gamma(link = "log"))
+men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "log"))
+
+men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
+men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma(link = "log"))
+
+```
+```{}
+This creates the year effect and the no year effect models
+```
+
+```{r}
+men_old_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = old_finals, family = Gamma(link = "log"))
+men_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = men_finals, family = Gamma(link = "log"))
+men_old_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
+men_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
+```
+```{}
+This creates the batch effect only models
+```
+
+```{r}
+men_old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_finals, family = Gamma(link = "log"))
+men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_finals, family = Gamma(link = "log"))
+men_old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
+men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
+```
+```{}
+This creates the year and batch effect models
+```
+
+```{r}
+anova(men_pooled_batch, men_pooled_two_effect)
 ```
@@
 ```{r}
-anova(one_effect, two_effect)
+anova(men_pooled_year_gamma_log, men_pooled_two_effect)
 ```
@@
 having only the year effect.
-```
-
-```{r}
-old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_finals, family = Gamma(link = "log"))
-men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_finals, family = Gamma(link = "log"))
-old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
-men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
-
-```
-
-```{}
-This section fits a two random effect model for each dataset.
 ```

### 2023-07-05 Code/BatchEffect.Rmd
Fixed line width issues

@@
 colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
-men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
-men_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
-men_old_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ' & Year < 2022)
-men_old_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Year < 2022)
+men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ')
+men_pooled <- subset(times, (Type == "F" | Type == "S") & 
+                       Gender == "M" & Time != 'DQ')
+men_old_finals <- subset(times, Type == "F" & Gender == "M" & 
+                           Time != 'DQ' & Year < 2022)
+men_old_pooled <- subset(times, (Type == "F" | Type == "S") & 
+                           Gender == "M" & Time != 'DQ' & Year < 2022)
 ```
@@
 library(lme4)
-men_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_finals, family = Gamma(link = "log"))
-men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, family = Gamma(link = "log"))
+men_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_finals, 
+                                    family = Gamma(link = "log"))
+men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, 
+                                       family = Gamma(link = "log"))
 
-men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, family = Gamma(link = "log"))
-men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, family = Gamma(link = "log"))
+men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, 
+                                family = Gamma(link = "log"))
+men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, 
+                                   family = Gamma(link = "log"))
 
-men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, family = Gamma(link = "log"))
-men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, family = Gamma(link = "log"))
+men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, 
+                                    family = Gamma(link = "log"))
+men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, 
+                                       family = Gamma(link = "log"))
 
-men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
-men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, family = Gamma(link = "log"))
+men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, 
+                                family = Gamma(link = "log"))
+men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, 
+                                   family = Gamma(link = "log"))
 
@@
 ```{r}
-men_old_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = old_finals, family = Gamma(link = "log"))
-men_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = men_finals, family = Gamma(link = "log"))
-men_old_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
-men_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
+men_old_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = old_finals, 
+                             family = Gamma(link = "log"))
+men_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = men_finals, 
+                         family = Gamma(link = "log"))
+men_old_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = old_pooled, 
+                             family = Gamma(link = "log"))
+men_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, 
+                         family = Gamma(link = "log"))
 ```
@@
 ```{r}
-men_old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_finals, family = Gamma(link = "log"))
-men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_finals, family = Gamma(link = "log"))
-men_old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = old_pooled, family = Gamma(link = "log"))
-men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
+men_old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+                                   data = old_finals, family = Gamma(link = "log"))
+men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+                              data = men_finals, family = Gamma(link = "log"))
+men_old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+                                   data = old_pooled, family = Gamma(link = "log"))
+men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+                               data = men_pooled, family = Gamma(link = "log"))
 ```

### 2023-07-05 Code/Simulation.Rmd
Something from meeting

@@
 ```{}
-
 If we want to pick the barrier based on the reaction time a barrier of 0.08905, 1 in every 1001.603
@@
 
+Change the name
 ```{r}

### 2023-07-09 Code/BatchEffect.Rmd
Renamed several variables and fixed the commenting issues

@@
 ```
-```{}
-This loads in the datasets
-```
+This loads in the 4 datasets from the csv file.
 
@@
 ```
-```{}
-This creates the year effect and the no year effect models
+This creates the year effect and the no effect models.  The models without a year
+effect end in "one_gamma_log" as in men_old_finals_one_gamma_log.
+
+```{r}
+men_old_finals_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = old_finals, 
+                             family = Gamma(link = "log"))
+men_finals_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_finals, 
+                         family = Gamma(link = "log"))
+men_old_pooled_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = old_pooled, 
+                             family = Gamma(link = "log"))
+men_pooled_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, 
+                         family = Gamma(link = "log"))
 ```
 
-```{r}
-men_old_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = old_finals, 
-                             family = Gamma(link = "log"))
-men_finals_batch <-glmer(RxnTime ~ (1 | Batch), data = men_finals, 
-                         family = Gamma(link = "log"))
-men_old_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = old_pooled, 
-                             family = Gamma(link = "log"))
-men_pooled_batch <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, 
-                         family = Gamma(link = "log"))
-```
-```{}
-This creates the batch effect only models
-```
+This creates the models that have the batch effect instead of the year effect.
 
 ```{r}
-men_old_finals_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+men_old_finals_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                                    data = old_finals, family = Gamma(link = "log"))
-men_finals_two_effect<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+men_finals_two_effect_gamma_log<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                               data = men_finals, family = Gamma(link = "log"))
-men_old_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+men_old_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                                    data = old_pooled, family = Gamma(link = "log"))
-men_pooled_two_effect <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+men_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                                data = men_pooled, family = Gamma(link = "log"))
 ```
-```{}
-This creates the year and batch effect models
-```
+This creates the models with both a year and batch effect.
 
@@
 ```
-```{}
-This anova test determined that the two_effect model is an improvement over the
-batch effect model. The test statistic is significant at alpha = 0.05 but not at
-alpha = 0.01.
-```
+
+This anova test determined that the two effect model is an improvement over the
+batch effect model. The test statistic is 0.01059 which is significant at alpha 
+= 0.05 but not at alpha = 0.01.
 
@@
 ```
-```{}
-In this anova test, it is clear that the two random effect model is superior to
-having only the year effect.
-```
+This anova test determined that the two effect model is a significant improvement
+over the year effect model.
 
@@
 ```
-
-```{}
 This code simulates the probability of observing a reaction time less than 0.1
-for the various datasets.
-```
+for the various data sets.  The function takes in two models and a reaction time,
+it is important that the two models that are compared are from the same data set;
+we want to compare old_finals_two_effect and old_finals_one_gamma_log.
 

### 2023-07-11 Code/BatchEffect.Rmd
Fixed issues and added good comments

@@
 
+The code below loads in the 4 datasets from the csv file.
 ```{r}
@@
 ```
-This loads in the 4 datasets from the csv file.
 
+This code below creates the base models without any random effects.  These are used
+as a baseline to compare the year effect models to in the simulation code below.
 ```{r}
@@
                                     family = Gamma(link = "log"))
+men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, 
+                                family = Gamma(link = "log"))
+men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, 
+                                    family = Gamma(link = "log"))
+men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, 
+                                family = Gamma(link = "log"))
+```
+
+This code creates the year effect only models.
+```{r}
+library(lme4)
+
 men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, 
-                                       family = Gamma(link = "log"))
-
-men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, 
                                        family = Gamma(link = "log"))
@@
                                    family = Gamma(link = "log"))
-
-men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, 
-                                    family = Gamma(link = "log"))
 men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, 
-                                       family = Gamma(link = "log"))
-
-men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, 
                                        family = Gamma(link = "log"))
@@
 ```
-This creates the year effect and the no effect models.  The models without a year
-effect end in "one_gamma_log" as in men_old_finals_one_gamma_log.
 
+This creates the batch effect only models.
 ```{r}
+library(lme4)
 men_old_finals_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = old_finals, 
@@
 
-This creates the models that have the batch effect instead of the year effect.
-
+This creates the models with a year effect and batch effect.
 ```{r}
+library(lme4)
 men_old_finals_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
@@
 ```
-This creates the models with both a year and batch effect.
 
@@
 ```
-
 This anova test determined that the two effect model is an improvement over the
@@
 
+This code simulates the probability of observing a reaction time less than 0.1
+for the various data sets.  The function takes in two models and a reaction time,
+it is important that the two models that are compared are from the same data set;
+we want to compare old_finals_two_effect and old_finals_one_gamma_log.
 ```{r}
@@
 library(lme4)
-simulation_two_effect <- function(model, base, RxnTime) {
+sim_two_effect_time <- function(model, base, RxnTime) {
   simfit <- function(model, n = 10000000) {
@@
 }
-simulation_two_effect(old_finals_two_effect, old_finals_one_gamma_log, 0.1)
-simulation_two_effect(men_finals_two_effect, men_finals_one_gamma_log, 0.1)
-simulation_two_effect(old_pooled_two_effect, old_pooled_one_gamma_log, 0.1)
-simulation_two_effect(men_pooled_two_effect, men_pooled_one_gamma_log, 0.1)
+sim_two_effect_time(old_finals_two_effect, old_finals_one_gamma_log, 0.1)
+sim_two_effect_time(men_finals_two_effect, men_finals_one_gamma_log, 0.1)
+sim_two_effect_time(old_pooled_two_effect, old_pooled_one_gamma_log, 0.1)
+sim_two_effect_time(men_pooled_two_effect, men_pooled_one_gamma_log, 0.1)
 ```
-This code simulates the probability of observing a reaction time less than 0.1
-for the various data sets.  The function takes in two models and a reaction time,
-it is important that the two models that are compared are from the same data set;
-we want to compare old_finals_two_effect and old_finals_one_gamma_log.
 
+sim_two_effect_prob is a slight modification of the earlier function that instead
+takes as input a probability rather than a time.  The output is a reaction time 
+barrier that allows a reaction time below the barrier one in every 1/(input probability)
+times. Decreasing the probability will impose a stricter barrier as it is less
+probable to observe a reaction time of 0.08 rather than 0.09.  Below are the
+recommended reaction time barriers to disqualify on average 1 in every 1,000 starts.
+```{r}
+set.seed(1)
+library(lme4)
+sim_two_effect_prob <- function(model, base, prob) {
+  simfit <- function(model, n = 10000000) {
+    vc <- as.data.frame(VarCorr(model))
+    sd_year <- vc$sdcor[1]
+    sd_year_batch <- vc$sdcor[2]
+
+    z1 <- rnorm(n, sd = sd_year)
+    z2 <- rnorm(n, sd = sd_year_batch)
+
+    shape <- 1 / sigma(model)^2
+    mu <- exp(fixef(model) + z1 + z2)
+
+    x <- rgamma(n, shape = shape, scale = mu / shape)
+    x
+  }
+
+  sbase <- summary(base)
+  x <- simfit(model)
+  print(quantile(x, probs = prob))
+}
+```
+
+This runs the sim_two_effect_prob function at a probability of 1 in every 1,000
+sprinters.
+```{r}
+sim_two_effect_prob(old_finals_two_effect, old_finals_one_gamma_log, 0.001)
+sim_two_effect_prob(men_finals_two_effect, men_finals_one_gamma_log, 0.001)
+sim_two_effect_prob(old_pooled_two_effect, old_pooled_one_gamma_log, 0.001)
+sim_two_effect_prob(men_pooled_two_effect, men_pooled_one_gamma_log, 0.001)
+```
+
+The code below repeats the previous section but for an average of 1 in every
+10,000 sprinters.
+```{r}
+sim_two_effect_prob(old_finals_two_effect, old_finals_one_gamma_log, 0.0001)
+sim_two_effect_prob(men_finals_two_effect, men_finals_one_gamma_log, 0.0001)
+sim_two_effect_prob(old_pooled_two_effect, old_pooled_one_gamma_log, 0.0001)
+sim_two_effect_prob(men_pooled_two_effect, men_pooled_one_gamma_log, 0.0001)
+```
+
+This next code section replicates the code above but the various simulation
+functions are meant for one random effect instead of two.
 ```{r}
@@
 set.seed(1)
-simulation_one_effect <- function(model, base, RxnTime){
+sim_one_effect_time <- function(model, base, RxnTime){
     simfit <- function(model, n = 10000000) {
@@
   print(mean(x < RxnTime))
-  #print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
   ## compare with and without random effect
@@
 }
-simulation_one_effect(old_finals_year_gamma_log, old_finals_one_gamma_log, 0.1)
-simulation_one_effect(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.1)
-simulation_one_effect(old_pooled_year_gamma_log, old_pooled_one_gamma_log, 0.1)
-simulation_one_effect(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
 ```
 
-```{}
-The above code simulates the probability of observing a reaction time less than
-0.1 seconds for the year effect model.
+```{r}
+sim_one_effect_time(old_finals_year_gamma_log, old_finals_one_gamma_log, 0.1)
+sim_one_effect_time(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.1)
+sim_one_effect_time(old_pooled_year_gamma_log, old_pooled_one_gamma_log, 0.1)
+sim_one_effect_time(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
 ```
 
-```{}
-All probabilities are for RxnTime=0.1 seconds
-old finals: 0.0004356 vs 0.0004389
-men_finals: 0.0008642 vs 0.0008626
-old pooled: 0.0041495 vs 0.0057287
-men_pooled: 0.005145 vs 0.0068441
-The times on the left are for the 2 batch, times on the right are for the 1 batch
+```{r}
+library(lme4)
+set.seed(1)
+sim_one_effect_prob <- function(model, base, prob) {
+  simfit <- function(model, n = 10000000) {
+    vc <- as.data.frame(VarCorr(model))
+    z <- rnorm(n, sd = vc$sdcor[1])
+    mu <- exp(fixef(model) + z)
+    shape <- 1 / vc$vcov[2]
+    x <- rgamma(n, shape = shape, scale = mu / shape)
+    x
+  }
 
-Thus if we want to pick the model with the larger probabilities to show that 
-0.01 is not extreme, the one random effect probabilities are slightly smaller.
+  sbase <- summary(base)
+  x <- simfit(model)
+  print(quantile(x, probs = prob))
+}
 ```
+
+```{r}
+sim_one_effect_prob(old_finals_year_gamma_log, old_finals_one_gamma_log, 0.001)
+sim_one_effect_prob(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.001)
+sim_one_effect_prob(old_pooled_year_gamma_log, old_pooled_one_gamma_log, 0.001)
+sim_one_effect_prob(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.001)
+```
+
+```{r}
+sim_one_effect_prob(old_finals_year_gamma_log, old_finals_one_gamma_log, 0.0001)
+sim_one_effect_prob(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.0001)
+sim_one_effect_prob(old_pooled_year_gamma_log, old_pooled_one_gamma_log, 0.0001)
+sim_one_effect_prob(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.0001)
+```

### 2023-07-12 Manuscript/manuscript.tex
Minor edits
Went over some of Prof Schifano's notes, still need to work on Results section

@@
 babicc2009reaction, panoutsakopoulos2020gender}.
-We focus on the reaction times during the semifinal and final heats only, as 
-reaction times from preliminary heats are often not as fast as those in later 
-heats \citep[e.g.,][]{zhang2021correlation}.
+For the creation of the GLMM model we focus on the reaction times during the 
+semifinal and final heats only, as reaction times from preliminary heats are 
+often not as fast as those in later heats \citep[e.g.,][]{zhang2021correlation}.
 
@@
 
-\subsection{2022 National Competitions}\label{sec:datanational}
+\subsubsection{2022 National Competitions}\label{sec:datanational}
 From June 23-26 2022, the United States held its Track and Field Championships 
@@
 competition (national or international) was specified for each competition 
-within the cluster (athlete).
+within the cluster (athlete).  We chose to include preliminary rounds in this
+section as well as in \ref{sec:data2019} in order to increase the size of our
+data.  Without data from the preliminary rounds, the number of observations
+dropped from 65 to 26 and the number of athletes included in the analysis
+decreased from 15 to 9.
 
@@
 data that will be used for two rank-based analyses.}
+\of{Does this work?}
 
-\subsection{2019 World Track and Field Championships}
+\subsubsection{2019 World Track and Field Championships}\label{sec:data2019}
 The 2019 World Track and Field Championship data is included in the larger dataset
-described in section~\ref{sec:databeyond}, but we can also use it similarly to
+described in section~\ref{sec:dataworld}, but we can also use it similarly to
 how we used national level in section~\ref{sec:datanational}.  We can compare
@@
 \jy{Close captions with a period.}
+\of{Resolved}
 \begin{table}
@@
 time, reaction time, etc. have all dramatically improved.  Seiko did not start 
-tracking reaction time as an official measurement until 1999 in the men's and 
-women's 100 meter \eds{110 for men?} hurdles. 
+tracking reaction time as an official measurement until 1999 in the men's 110 meter
+hurdles and women's 100 meter hurdles. 
 Seiko regularly updates their equipment so that they provide cutting edge
@@
   group being ....}
-
+\subsubsection{2019 World Track and Field Championships} \label{sec:results2019vs2022}
 The rank based methods described in Section~\ref{sec:rank} are used
@@
 faster during that timespan.  \eds{That is not the only explanation}
+\subsection{2022 National Championships} \label{sec:resultsnational}
 
-Repeating the data collection steps and
-again performing a clustered analysis for a Wilcoxon ranked sum test
-results in two new z scores and p values. 
-\eds{I am not sure why you are repeating the data collection steps and
-again performing a clustered analysis here.  Also, if there are two new z-scores
-and p values, why are you only reporting one?}
   The ``ds" method produced a z score of

### 2023-07-12 Manuscript/manuscript.tex
a few edits and responses to Qs

@@
 From June 23-26 2022, the United States held its Track and Field Championships 
-at Hayward Field in Eugene, Oregon to decide who to send to the World Championships 
-being held at the same venue in August. Thus, we can establish a baseline for the four United States athletes who 
-competed in at least one round of both events: Trey Cunningham, Daniel Roberts, 
-Grant Holloway, and Devon Allen. Every athlete reacted faster in all the World 
+at Hayward Field in Eugene, Oregon to decide who to send to the World 
+Championships being held at the same venue in August. Thus, we can establish a 
+baseline for the four United States athletes who competed in at least one round 
+of both events: Trey Cunningham, Daniel Roberts, 
+Grant Holloway, and Devon Allen. Table~\ref{fig:USAvsWorld} highlights the 
+difference in reaction times for these four athletes in the two events. 
+Every athlete reacted faster in all the World 
 Track and Field Championships races compared to the United States Track and Field 
@@
 disqualified more because of faulty equipment rather than because he reacted
-too quickly.  Table~\ref{fig:USAvsWorld} highlights the difference in reaction 
-times for Devon Allen, Trey Cunningham, Grant Holloway, and Daniel Roberts.
+too quickly.
 
@@
 \of{Does this work?}
+\eds{Yes, but you still need figure or descriptive statistics for this data.}
 
@@
 Figure~\ref{fig:Scatterplot2019vs2022} shows the distribution of times from the
-2019 and 2022 World Championships.  This data was only for athletes who competed
+2019 and 2022 World Championships.  
+\eds{Please reorganize the labels so they are either in one of the following 
+orders:
+2019 Heat, 2019 Semi, 2019 final, 2022 Heat, 2022 Semi, 2022 final  OR
+2019 Heat, 2022 Heat, 2019 Semi, 2022 Semi, 2019 final, 2022 final.  
+(I am not sure which is better, but perhaps the latter?)}
+This data was only for athletes who competed
 in at least one race at each championship, however it is possible for athletes
@@
 Semifinals both came from Devon Allen, but the difference between 0.101 and 0.009
-seconds was the difference in what caused him to be disqualified.
+seconds was the difference in what caused him to be disqualified. \eds{check 
+these numbers? 0.099?}
 
@@
 
+	The rank based methods described in Section~\ref{sec:rank} are used
+to determine if the reaction times for athletes who had competed at other 
+championships differed significantly from the 2022 World Championships. 
 \jy{Remind the readers what the analysis was about: two-group comparison with
   group being ....}
+
 \subsubsection{2019 World Track and Field Championships} \label{sec:results2019vs2022}
-The rank based methods described in Section~\ref{sec:rank} are used
-to determine if the reaction times for athletes who had competed at other 
-championships differed significantly from the 2022 World Championships.  We can
+ We can
 use the 2019 World Championships and the 2022 national-level championships as a 
@@
 faster during that timespan.  \eds{That is not the only explanation}
-\subsection{2022 National Championships} \label{sec:resultsnational}
+\subsubsection{2022 National Championships} \label{sec:resultsnational}
 

### 2023-07-13 Code/GLMM.Rmd
Should be reproducible now

---
title: "Random_Effects"
author: "Owen Fiore"
date: "2023-05-31"
output: pdf_document
---

Professors Yan and Schifano: to see the probabilities of the three different models
please look at the results from code chunks 7, 12, 13.  You will need to run
code from other chunks to generate the models and for the simulation code, but
these three will produce probabilities of a reaction time less than 0.1 seconds.

The code below loads in the 4 datasets from the csv file.
```{r}
#1
times <- read.csv("../Data/ThesisData.csv")
colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ')
men_pooled <- subset(times, (Type == "F" | Type == "S") & 
                       Gender == "M" & Time != 'DQ')
men_old_finals <- subset(times, Type == "F" & Gender == "M" & 
                           Time != 'DQ' & Year < 2022)
men_old_pooled <- subset(times, (Type == "F" | Type == "S") & 
                           Gender == "M" & Time != 'DQ' & Year < 2022)
```

This code below creates the base models without any random effects.  These are used
as a baseline to compare the year effect models to in the simulation code below.
```{r}
#2
library(lme4)
men_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_finals, 
                                    family = Gamma(link = "log"))
men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, 
                                family = Gamma(link = "log"))
men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, 
                                    family = Gamma(link = "log"))
men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, 
                                family = Gamma(link = "log"))
```

This code creates the year effect only models.
```{r}
#3
library(lme4)

men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, 
                                       family = Gamma(link = "log"))
men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, 
                                   family = Gamma(link = "log"))
men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, 
                                       family = Gamma(link = "log"))
men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, 
                                   family = Gamma(link = "log"))

```

This creates the batch effect only models.
```{r}
#4
library(lme4)
men_old_finals_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_old_finals, 
                             family = Gamma(link = "log"))
men_finals_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_finals, 
                         family = Gamma(link = "log"))
men_old_pooled_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_old_pooled, 
                             family = Gamma(link = "log"))
men_pooled_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, 
                         family = Gamma(link = "log"))
```

This creates the models with a year effect and batch effect.
```{r}
#5
library(lme4)
men_old_finals_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                                   data = men_old_finals, family = Gamma(link = "log"))
men_finals_two_effect_gamma_log<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                              data = men_finals, family = Gamma(link = "log"))
men_old_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                                   data = men_old_pooled, family = Gamma(link = "log"))
men_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
                               data = men_pooled, family = Gamma(link = "log"))
```

```{r}
anova(men_pooled_batch_gamma_log, men_pooled_two_effect_gamma_log)
```
This anova test determined that the two effect model is an improvement over the
batch effect model. The test statistic is 0.01059 which is significant at alpha 
= 0.05 but not at alpha = 0.01.

```{r}
anova(men_pooled_year_gamma_log, men_pooled_two_effect_gamma_log)
```
This anova test determined that the two effect model is a significant improvement
over the year effect model.

This code simulates the probability of observing a reaction time less than 0.1
for the various data sets.  The function takes in two models and a reaction time,
it is important that the two models that are compared are from the same data set;
we want to compare old_finals_two_effect and old_finals_one_gamma_log.
```{r}
#6
set.seed(1)
library(lme4)
sim_two_effect_time <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_year <- vc$sdcor[1]
    sd_year_batch <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_year)
    z2 <- rnorm(n, sd = sd_year_batch)

    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue")  # fixed effect only
}
```

This next chunk of code runs the simulation code for each dataset comparing the
two effect model to the base model
```{r}
#7
sim_two_effect_time(men_old_finals_two_effect_gamma_log, men_old_finals_one_gamma_log, 0.1)
sim_two_effect_time(men_finals_two_effect_gamma_log, men_finals_one_gamma_log, 0.1)
sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.1)
sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.1)
```

sim_two_effect_prob is a slight modification of the earlier function that instead
takes as input a probability rather than a time.  The output is a reaction time 
barrier that allows a reaction time below the barrier one in every 1/(input probability)
times. Decreasing the probability will impose a stricter barrier as it is less
probable to observe a reaction time of 0.08 rather than 0.09.  Below are the
recommended reaction time barriers to disqualify on average 1 in every 1,000 starts.
```{r}
#8
set.seed(1)
library(lme4)
sim_two_effect_prob <- function(model, base, prob) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_year <- vc$sdcor[1]
    sd_year_batch <- vc$sdcor[2]

    z1 <- rnorm(n, sd = sd_year)
    z2 <- rnorm(n, sd = sd_year_batch)

    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))
}
```

This runs the sim_two_effect_prob function at a probability of 1 in every 1,000
sprinters.
```{r}
#9
sim_two_effect_prob(men_old_finals_two_effect_gamma_log, men_old_finals_one_gamma_log, 0.001)
sim_two_effect_prob(men_finals_two_effect_gamma_log, men_finals_one_gamma_log, 0.001)
sim_two_effect_prob(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.001)
sim_two_effect_prob(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.001)
```

The code below repeats the previous section but for an average of 1 in every
10,000 sprinters.
```{r}
#10
sim_two_effect_prob(men_old_finals_two_effect_gamma_log, men_old_finals_one_gamma_log, 0.0001)
sim_two_effect_prob(men_finals_two_effect_gamma_log, men_finals_one_gamma_log, 0.0001)
sim_two_effect_prob(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.0001)
sim_two_effect_prob(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.0001)
```

This next code section replicates the code above but the various simulation
functions are meant for one random effect instead of two.
```{r}
#11
library(lme4)
set.seed(1)
sim_one_effect_time <- function(model, base, RxnTime){
    simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
    }
  sbase <- summary(base)
  #print(sbase)
  x <- simfit(model)
  print(mean(x < RxnTime))
  ## compare with and without random effect
  plot(density(x)) # with random effect
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
}
```

```{r}
#12
sim_one_effect_time(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log, 0.1)
sim_one_effect_time(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.1)
sim_one_effect_time(men_old_pooled_year_gamma_log, men_old_pooled_one_gamma_log, 0.1)
sim_one_effect_time(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
```

```{r}
#13
sim_one_effect_time(men_old_finals_batch_gamma_log, men_old_finals_one_gamma_log, 0.1)
sim_one_effect_time(men_finals_batch_gamma_log, men_finals_one_gamma_log, 0.1)
sim_one_effect_time(men_old_pooled_batch_gamma_log, men_old_pooled_one_gamma_log, 0.1)
sim_one_effect_time(men_pooled_batch_gamma_log, men_pooled_one_gamma_log, 0.1)
```

```{r}
#14
library(lme4)
set.seed(1)
sim_one_effect_prob <- function(model, base, prob) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    z <- rnorm(n, sd = vc$sdcor[1])
    mu <- exp(fixef(model) + z)
    shape <- 1 / vc$vcov[2]
    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(quantile(x, probs = prob))
}
```

```{r}
#15
sim_one_effect_prob(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log, 0.001)
sim_one_effect_prob(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.001)
sim_one_effect_prob(men_old_pooled_year_gamma_log, men_old_pooled_one_gamma_log, 0.001)
sim_one_effect_prob(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.001)
```

```{r}
#16
sim_one_effect_prob(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log, 0.0001)
sim_one_effect_prob(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.0001)
sim_one_effect_prob(men_old_pooled_year_gamma_log, men_old_pooled_one_gamma_log, 0.0001)
sim_one_effect_prob(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.0001)
```

### 2023-07-17 Manuscript/manuscript.tex
Fixed 1 thing

@@
 worth noting is that the fastest reaction time in the 2022 Finals and 2022
-Semifinals both came from Devon Allen, but the difference between 0.101 and 0.009
+Semifinals both came from Devon Allen, but the difference between 0.101 and 0.099
 seconds was the difference in what caused him to be disqualified. \eds{check 
-these numbers? 0.099?}
+these numbers? 0.099?}\of{resolved}
 

### 2023-07-18 Manuscript/manuscript.tex
Updated tables and moved Seiko to Discussion, still need to make it fit

@@
    \midrule
-   Old Finals & $4.39\cdot10^{-4}$ & $2.0\cdot10^{-5}$ \\
-   All Finals & $8.56\cdot10^{-4}$ & $6.0\cdot10^{-7}$ \\
-   Old Pooled & $5.76\cdot10^{-3}$ & $1.0\cdot10^{-4}$ \\ 
-   All Pooled & $6.86\cdot10^{-3}$ & $1.3\cdot10^{-4}$ \\
+   Old Finals & $4.36\cdot10^{-4}$ & $3.0\cdot10^{-7}$ \\
+   All Finals & $8.64\cdot10^{-4}$ & $6.0\cdot10^{-7}$ \\
+   Old Pooled & $4.15\cdot10^{-3}$ & $4.67\cdot10^{-5}$ \\ 
+   All Pooled & $5.15\cdot10^{-3}$ & $6.01\cdot10^{-5}$ \\
    \bottomrule
@@
    All Finals & $0.101$ & $0.093$ \\
-   Old Pooled & $0.090$ & $0.080$ \\ 
-   All Pooled & $0.089$ & $0.079$ \\
+   Old Pooled & $0.093$ & $0.083$ \\ 
+   All Pooled & $0.091$ & $0.082$ \\
    \bottomrule
@@
 
-%Citation needed in this section
+\subsection{Rank-based Comparison} \label{subsec:Results_Rank}
+
+The rank based methods described in Section~\ref{sec:rank} are used
+to determine if the reaction times for athletes who had competed at other 
+championships differed significantly from the 2022 World Championships.  This is
+a two group comparison performed two times: once comparing 2019 World
+Championships to 2022 World Championships and can be found in 
+section~\ref{sec:results2019vs2022} and secondly comparing 2022 national
+championships to the 2022 World Championships and can be found in 
+~section\ref{sec:resultsnational}.
+\jy{Remind the readers what the analysis was about: two-group comparison with
+  group being ....}
+
+\subsubsection{2019 World Track and Field Championships} \label{sec:results2019vs2022}
+The p value associated with the two group comparison for 2019 and 2022 was 0.001464.  
+\eds{what is the ds method?} \of{Its described in the methods section but I dont
+think its no longer needed now that we are no longer using the other grouping
+method (I believe it was called something like the rosner method)}
+This value is for a one-sided test and show that
+the mean reaction time on average in 2019 was larger (slower) than the mean 
+reaction time
+in 2022 for athletes who competed in both championships at an alpha level of 0.01.
+This provides good evidence that athletes who competed in both 2019 and 2022 got
+faster during that timespan.  \eds{That is not the only explanation}
+\subsubsection{2022 National Championships} \label{sec:resultsnational}
+
+  The ``ds" method produced a z score of
+3.6069 and p value of 0.000155. This result is highly significant and shows that
+the times at the national stage of competition were significantly higher than
+the reaction times at the World Championships even for an alpha level of 0.001.
+
+\eds{also consider moving the following paragraph to discussion??}
+As Seiko timed both events, there should not be a significant
+difference if the timing equipment in both instances was working correctly.
+\eds{you just mentioned in the last subsection that Seiko upgraded its 
+technology prior to 2022}
+  To put it simply, there is not a reasonable explanation for 
+this. It is inconceivable that so many athletes would improve such a significant
+amount in such a small amount of time. Six countries are represented in this 
+comparison, which drastically decreases the chances of this issue being caused by
+one specific country.  The United States athletes' improvement in times was
+discussed earlier in the paper, but this data shows that on average there was
+\jy{any reference?}
+improvement for British, Polish, French, Brazilian, and Spanish athletes.
+
+\section{Discussion}\label{sec:Discussion}
+
 Since 1985 Seiko Holding Corporations has served every World Athletics 
@@
 no results presented here}
-
-\subsection{Rank-based Comparison} \label{subsec:Results_Rank}
-
-	The rank based methods described in Section~\ref{sec:rank} are used
-to determine if the reaction times for athletes who had competed at other 
-championships differed significantly from the 2022 World Championships. 
-\jy{Remind the readers what the analysis was about: two-group comparison with
-  group being ....}
-
-\subsubsection{2019 World Track and Field Championships} \label{sec:results2019vs2022}
- We can
-use the 2019 World Championships and the 2022 national-level championships as a 
-precedent to see how athletes performance changed across different championships.
-
-When the clusrank rank sum test was performed, the ``ds" method produced a z score of
-2.9751 and p value of 0.001464.  \eds{what is the ds method?} 
-This value is for a one-sided test and show that
-the mean reaction time on average in 2019 was larger (slower) than the mean 
-reaction time
-in 2022 for athletes who competed in both championships at an alpha level of 0.01.
-This provides good evidence that athletes who competed in both 2019 and 2022 got
-faster during that timespan.  \eds{That is not the only explanation}
-\subsubsection{2022 National Championships} \label{sec:resultsnational}
-
-  The ``ds" method produced a z score of
-3.6069 and p value of 0.000155. This result is highly significant and shows that
-the times at the national stage of competition were significantly higher than
-the reaction times at the World Championships even for an alpha level of 0.001.
-
-\eds{also consider moving the following paragraph to discussion??}
-As Seiko timed both events, there should not be a significant
-difference if the timing equipment in both instances was working correctly.
-\eds{you just mentioned in the last subsection that Seiko upgraded its 
-technology prior to 2022}
-  To put it simply, there is not a reasonable explanation for 
-this. It is inconceivable that so many athletes would improve such a significant
-amount in such a small amount of time. Six countries are represented in this 
-comparison, which drastically decreases the chances of this issue being caused by
-one specific country.  The United States athletes' improvement in times was
-discussed earlier in the paper, but this data shows that on average there was
-\jy{any reference?}
-improvement for British, Polish, French, Brazilian, and Spanish athletes.
-
-\section{Discussion}\label{sec:Discussion}
 

### 2023-07-19 Manuscript/manuscript.tex
Went over many suggestions

@@
   group being ....}
+\of{Better?}
 
 \subsubsection{2019 World Track and Field Championships} \label{sec:results2019vs2022}
-The p value associated with the two group comparison for 2019 and 2022 was 0.001464.  
-\eds{what is the ds method?} \of{Its described in the methods section but I dont
-think its no longer needed now that we are no longer using the other grouping
-method (I believe it was called something like the rosner method)}
-This value is for a one-sided test and show that
-the mean reaction time on average in 2019 was larger (slower) than the mean 
-reaction time
-in 2022 for athletes who competed in both championships at an alpha level of 0.01.
-This provides good evidence that athletes who competed in both 2019 and 2022 got
-faster during that timespan.  \eds{That is not the only explanation}
+The p value from the two group comparison for 2019 and 2022 was 0.001464.  This 
+value is for a one-sided test and show that the mean reaction time on average in
+2019 was larger (slower) than the mean reaction time in 2022 for athletes who 
+competed in both championships at an $\alpha$ level of 0.01. This provides good
+evidence that on average, the reaction times from the 2019 and 2022 World 
+Championships are significantly different.  
+\eds{That is not the only explanation} \of{You are correct so I changed the 
+wording}
+
 \subsubsection{2022 National Championships} \label{sec:resultsnational}
+The p value from the two group comparison between the 2022 national
+championships and World Championships was 0.000155. This value is for a one sided
+test and the result is significant at $\alpha = 0.01$ and shows that
+the times at the World Championships were significantly faster than at the
+national championships.
 
-  The ``ds" method produced a z score of
-3.6069 and p value of 0.000155. This result is highly significant and shows that
-the times at the national stage of competition were significantly higher than
-the reaction times at the World Championships even for an alpha level of 0.001.
+\section{Discussion}\label{sec:Discussion}
+Since 1985 Seiko Holding Corporations has served every World Athletics 
+Championships as the official timer \citep{wa2022seiko}.  Since 1985, the 
+technology and the ability to accurately predict measurements: long jump 
+distances, false starts, total time, reaction time, etc. have all dramatically 
+improved.  Seiko did not start tracking reaction time as an official measurement
+until 1999 in the men's 110 meter hurdles and women's 100 meter hurdles. 
+Seiko regularly updates their equipment so that they provide cutting edge
+technology to the World Track and Field Championships, the highest stakes in the 
+world of running outside of the Olympics. Seiko's technology for detecting 
+reaction times relies on the pressure that athletes exert on the plate when they 
+push off.  Their systems measure the time differential between when the ``gun" 
+goes off to start the race and when the pressure changes \citep{wa2022seiko}.  
+If an athlete has a reaction time under 0.1 seconds, it is deemed a false start 
+as it is considered that no athlete can react so quickly \citep{Seiko-Timing}.  
+Thus, a reaction time of for example 0.05, suggests that the athlete predicted 
+the gun and it was luck that caused their abnormally high time.  At the World 
+Championships level, World Athletics and Seiko want to remove that element of 
+luck and thus impose the 0.1 second barrier. In 2013, Seiko upgraded their 
+timing equipment for sprinting events (110 meter hurdles falls under this 
+category) \citep{wa2013backtage}. Since 1999, the slowest reaction times in the 
+men's 110-meter hurdle were recorded in 2013.  That is not to say that the higher 
+times were caused by Seiko's equipment, but the two may be related.  It is worth 
+noting that prior to 2022, Seiko again upgraded its technology but for its jump 
+management system. 
+\eds{I would argue that this paragraph should go in the discussion as there are 
+no results presented here}
+\of{I agree so I moved it here. Not sure if it belongs at the top of the
+discussion or elsewhere but here it is.}
 
@@
 technology prior to 2022}
+of{Yes I agree}
 To put it simply, there is not a reasonable explanation for 
@@
 
-\section{Discussion}\label{sec:Discussion}
-
-Since 1985 Seiko Holding Corporations has served every World Athletics 
-Championships as the official timer \citep{wa2022seiko}.  Since 1985, the 
-technology and the ability
-to accurately predict measurements: long jump distances, false starts, total 
-time, reaction time, etc. have all dramatically improved.  Seiko did not start 
-tracking reaction time as an official measurement until 1999 in the men's 110 meter
-hurdles and women's 100 meter hurdles. 
-Seiko regularly updates their equipment so that they provide cutting edge
-technology to the World Track and Field Championships, the highest stakes in the 
-world of running outside of the Olympics. Seiko's technology for detecting 
-reaction times relies on the pressure that athletes exert on the plate when they 
-push off.  Their systems measure the time differential
-between when the ``gun" goes off to start the race and when the pressure changes 
-\citep{wa2022seiko}.  If an athlete has a reaction time under 0.1 seconds, it 
-is deemed a 
-false start as it is considered that no athlete can react so quickly 
-\citep{Seiko-Timing}.  Thus, a reaction time of for example 0.05, suggests that 
-the athlete predicted the gun and it was luck that caused their abnormally high 
-time.  At the World Championships level, World Athletics and Seiko want to remove 
-that element of luck and thus impose the 0.1 second barrier. In 2013, Seiko 
-upgraded their timing equipment for sprinting events (110 meter hurdles falls 
-under this category) \citep{wa2013backtage}. Since 1999, the
-slowest reaction 
-times in the men's 110-meter hurdle were recorded in 2013.  That is not to say
- that the higher 
-times were caused by Seiko's equipment, but the two may be related.  It is worth 
-noting that prior
-to 2022, Seiko again upgraded its technology but for its jump management system.
-\eds{I would argue that this paragraph should go in the discussion as there are 
-no results presented here}
-
 One conclusion from the analysis in \citet{zhang2021correlation} was that 
-athletes'
-reaction times improved from 2011 to 2019.  However this study does not consider 
-data from 1999 to 2009, and over this timespan athletes get slower and then
-faster.  Figure~\ref{fig:Boxplots} shows that for the men's pooled data, the 
-year with the fastest median reaction time before 2022 was 1999.  Thus, this 
+athletes' reaction times improved from 2011 to 2019.  However this study does 
+not consider data from 1999 to 2009, and over this timespan athletes get slower 
+and then faster.  Figure~\ref{fig:Boxplots} shows that for the men's pooled data,
+the year with the fastest median reaction time before 2022 was 1999.  Thus, this 
 reject's Zhang et al.'s conclusion that athletes are getting faster because they 
-were 
-fast over twenty years ago. One possible explanation for why 2011 had such
+were fast over twenty years ago. One possible explanation for why 2011 had such
 high variation was because of the IAAF rule discussed earlier that increased the
@@
 through a gamma mixed effects model that there are significant year and race
-effects \eds{remove race effects?} that greatly impact reaction time.
+effects \eds{remove race effects?}\of{I think were going to go with the two
+effect model so no.} that greatly impact reaction time.
 
@@
 
-Thus while it seems easy
-to conclude that Devon Allen was wrongly disqualified, that may not necessarily
-be the case. 
-If there are issues with reaction time such as athletes approaching 
-the 0.1 
-second barrier again at the next World Championships (August 2023), then 
-World Athletics 
-needs to adjust that threshold and allow for faster reaction times so that 
-athletes with superb 
-reflexes are not penalized. %like how Allen was in 2022.  
-World Athletics did not
+It is confusing as to why the reaction time barrier for men's and women's data
+is the same despite numerous articles such as \citep[e.g.,][]{lipps2011implications, 
+babicc2009reaction, panoutsakopoulos2020gender} stating differently.  World
+Athletics is ignoring differences in gender by setting the reaction time barrier
+to be equal. If 0.1 seconds is considered a fair time for women, then 0.1 seconds
+could not be considered fair for men, as the probability of observing reaction
+times lower than 0.1 would be elevated.
+
+If there are issues with reaction time such as athletes approaching the 0.1 
+second barrier again at the next World Championships (August 2023), then World 
+Athletics needs to adjust that threshold and allow for faster reaction times so 
+that athletes with superb reflexes are not penalized.  World Athletics did not
 do anything in response to the \citet{komi2009iaaf} study, but perhaps they 
-should now so
-that the story of Devon Allen is not repeated.
+should now so that the story of Devon Allen is not repeated.
 

### 2023-07-19 Manuscript/manuscript.tex
added table comparing the 3 models and their probabilities for each dataset

@@
 
+\jy{label the table and reference to it; clean it up}
+\begin{table}
+  \centering
+  \caption{Compares the various effect models by dataset, all probabilities are
+  for observing a reaction time less than 0.1 seconds. In general the model with
+  the highest probability (least extreme reaction time) is the Year Effect model.}
+  \begin{tabular}{c c c c} 
+   \toprule
+   Data Set & Year Effect & Batch Effect & Both Effects \\ 
+   \midrule
+   Old Finals & $4.39\cdot10^{-4}$ & $4.24\cdot10^{-4}$ & $4.36\cdot10^{-4}$ \\
+   All Finals & $8.63\cdot10^{-4}$ & $8.65\cdot10^{-4}$ & $8.64\cdot10^{-4}$ \\
+   Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
+   All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:EffectComparison}
+\end{table}
+
 Using the gamma function with link set to log, the probability of the reaction 
@@
 no results presented here}
-\of{I agree so I moved it here. Not sure if it belongs at the top of the
-discussion or elsewhere but here it is.}
+\of{I agree so I moved it here. I think it also could fit into the top of the
+data section as in section~\ref{sec:Data}. Thoughts?}
 

### 2023-07-20 Code/RankComparison.Rmd
Added the exact=True argument below the other tests

@@
 ---
-
-```{r}
-inter <- read.csv("../Data/2019vs2022.csv")
-
-#nine is used for 2019 and two is used for 2022.  H, S, F are used to denote the heats, semifinals, and finals.
-colnames(inter) <- c("Athlete", "Country", "2019H", "2019S", "2019F", "2022H", "2022S", "2022F")
-```
-
-```{r}
-my_sums <- colSums(!is.na(inter))
-print(my_sums)
-barplot(my_sums)
-```
-
-```{r}
-means <- sapply(inter[,3:ncol(inter)], mean, na.rm = TRUE)
-print(means)
-
-```
-
-```{r}
-library(ggplot2)
-means_df <- data.frame(variable = names(means), mean = means)
-
-# Define the grouping of bars
-groupings <- list(c("2019H", "2022H"), c("2019S", "2022S"), c("2019F", "2022F"))
-
-# Convert variable to a factor with desired order
-means_df$variable <- factor(means_df$variable, levels = unlist(groupings))
-
-# Create the bar graph
-ggplot(means_df, aes(x = variable, y = mean, fill = variable)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  scale_fill_manual(values = c("sienna2", "tan", "sienna2", "tan", "sienna2", "tan")) +
-  labs(y = "Mean Reaction Time", x ="") +
-  theme_minimal() +
-  theme(legend.position = "none")
-ggsave(filename="2019vs2022BarGraph.pdf", path = "../Manuscript", height=4, width=6)
-```
 
@@
 clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, alternative = "greater", method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, alternative = "greater", method = "ds", exact = TRUE)
+```
+
+```{r}
+comp <- read.csv("../Data/Clusrank2019vs2022.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(comp) <- c("Time", "Athlete", "Year")
+#View(comp)
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, alternative = "greater", method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, alternative = "greater", method = "ds", exact = TRUE)
 ```

### 2023-07-20 Manuscript/manuscript.tex
Started to clean up the discussion section

@@
 dropped from 65 to 26 and the number of athletes included in the analysis
-decreased from 15 to 9.
+decreased from 15 to 9.  Likewise
 
@@
 \section{Discussion}\label{sec:Discussion}
+One conclusion from the analysis in \citet{zhang2021correlation} was that 
+athletes' reaction times improved from 2011 to 2019.  However this study does 
+not consider data from 1999 to 2009, and over this timespan athletes get slower 
+and then faster.  Figure~\ref{fig:Boxplots} shows that for the men's pooled data,
+the year with the fastest median reaction time before 2022 was 1999.  Thus, this 
+reject's Zhang et al.'s conclusion that athletes are getting faster because they 
+were fast over twenty years ago. One possible explanation for why 2011 had such
+high variation was because of the IAAF rule discussed earlier that increased the
+penalty of a false start to an instant disqualification \citep{iaaf2009falsestart}.
+Athletes may have been overly cautious to false start and may have consciously
+or sub-consciously reacted slower as a result.  Instead, a more appropriate 
+conclusion may be that there is a lot of variability in reaction times and that
+the context of the races matters.
+
+This paper does not consider data from the women's 100 meter hurdle, and although
+that was initially because of concerns over reaction time differences in men and
+women, exploratory analysis showed that 2022 was at best only
+slightly faster relative to other championships going back to 2003 
+\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
+the intro, unless he only considered sprint and not hurdles?} \of{Johnson's
+results considered the mens 100m, mens 110m hurdles, womens 100m but not the 
+womens 100m hurdles.} Likewise, the rank-based comparisons resulted in p values
+that were not significant.  The lack of evidence from the women's data makes it 
+harder to conclude that the sensors used at the 2022 championships were faulty, 
+although it is unknown if the sensors for the men's 110 meter hurdle and women's
+100 meter hurdle were the same.
+
+It is confusing as to why the reaction time barrier for men's and women's data
+is the same despite numerous articles such as \citep[e.g.,][]{lipps2011implications, 
+babicc2009reaction, panoutsakopoulos2020gender} stating differently.  World
+Athletics is ignoring differences in gender by setting the reaction time barrier
+to be equal. If 0.1 seconds is considered a fair time for women, then 0.1 seconds
+could not be considered fair for men, as the probability of observing reaction
+times lower than 0.1 would be elevated.  \citet{brosnan2017effects} advocated
+for gender based reaction time barriers which seems logical if there are biological
+differences in reaction times between men and women.
+
 Since 1985 Seiko Holding Corporations has served every World Athletics 
-Championships as the official timer \citep{wa2022seiko}.  Since 1985, the 
-technology and the ability to accurately predict measurements: long jump 
-distances, false starts, total time, reaction time, etc. have all dramatically 
-improved.  Seiko did not start tracking reaction time as an official measurement
-until 1999 in the men's 110 meter hurdles and women's 100 meter hurdles. 
-Seiko regularly updates their equipment so that they provide cutting edge
-technology to the World Track and Field Championships, the highest stakes in the 
-world of running outside of the Olympics. Seiko's technology for detecting 
+Championships as the official timer \citep{wa2022seiko}.  Reaction times were
+not officially measured until 1999 in the men's 110 meter hurdles and women's
+100 meter hurdles. Seiko's technology for detecting 
 reaction times relies on the pressure that athletes exert on the plate when they 
@@
 the gun and it was luck that caused their abnormally high time.  At the World 
-Championships level, World Athletics and Seiko want to remove that element of 
-luck and thus impose the 0.1 second barrier. In 2013, Seiko upgraded their 
-timing equipment for sprinting events (110 meter hurdles falls under this 
-category) \citep{wa2013backtage}. Since 1999, the slowest reaction times in the 
-men's 110-meter hurdle were recorded in 2013.  That is not to say that the higher 
-times were caused by Seiko's equipment, but the two may be related.  It is worth 
-noting that prior to 2022, Seiko again upgraded its technology but for its jump 
-management system. 
+Championships level, World Athletics want to remove that element of 
+luck and thus impose the 0.1 second barrier.
 \eds{I would argue that this paragraph should go in the discussion as there are 
@@
 
-One conclusion from the analysis in \citet{zhang2021correlation} was that 
-athletes' reaction times improved from 2011 to 2019.  However this study does 
-not consider data from 1999 to 2009, and over this timespan athletes get slower 
-and then faster.  Figure~\ref{fig:Boxplots} shows that for the men's pooled data,
-the year with the fastest median reaction time before 2022 was 1999.  Thus, this 
-reject's Zhang et al.'s conclusion that athletes are getting faster because they 
-were fast over twenty years ago. One possible explanation for why 2011 had such
-high variation was because of the IAAF rule discussed earlier that increased the
-penalty of a false start to an instant disqualification \citep{iaaf2009falsestart}.
-Athletes may have been overly cautious to false start and may have consciously
-or sub-consciously reacted slower as a result.  Instead, a more appropriate 
-conclusion may be that there is a lot of variability in reaction times and that
-the context behind the races matters.
-
 It is worth noting that in the semifinals of the 2022 World Track and
@@
 times in a row, which is the exact reason that the 0.1 second reaction time 
-disqualification barrier was
-imposed in the first place. Thus having a quick reaction time multiple times
-may indicate skill rather than luck. Devon 
-Allen is a very good athlete, as evidenced by his football career at the 
-University of Oregon
-and making it onto the Philadelphia Eagles practice squad this past year 
-\citep{hurley2022eagles}. The 0.099 reaction time may not have been 
-a product of Devon Allen predicting the start of the race but rather a 
-combination of a quick reaction and a possibly faulty sensor.
+disqualification barrier was imposed in the first place. Thus having a quick 
+reaction time repeatedly may indicate skill rather than luck. Devon Allen is
+an exceptional athlete, as evidenced by his football career at the University of
+Oregon and making it onto the Philadelphia Eagles practice squad in 2022
+\citep{hurley2022eagles}.  \citet{komi2009iaaf}'s paper showed that athletes are
+capable of reacting in less than 0.1 seconds, and if this is the case, it is not
+unreasonable to conclude that an outstanding athlete like Allen is capable of
+reacting below the 0.1 second barrier.
+
+The results from the GLMM are described in section~\ref{subsec:Results_GLMM} and
+show that there
 
@@
 
-This paper does not consider data from the women's 100 meter hurdle, and although
-that was initially because of concerns over reaction time differences in men and
-women, exploratory analysis showed that 2022 was at best only
-slightly faster relative to other championships going back to 2003 
-\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
-the intro, unless he only considered sprint and not hurdles?} Likewise,
-the rank-based comparisons resulted in p values that were not significant.  The
-lack of evidence from the women's data makes it harder to conclude 
-that the sensors used at the 2022 championships were faulty, although it is
-unknown if the sensors for the men's 110 meter hurdle and women's 100 meter
-hurdle were the same.
-
-It is confusing as to why the reaction time barrier for men's and women's data
-is the same despite numerous articles such as \citep[e.g.,][]{lipps2011implications, 
-babicc2009reaction, panoutsakopoulos2020gender} stating differently.  World
-Athletics is ignoring differences in gender by setting the reaction time barrier
-to be equal. If 0.1 seconds is considered a fair time for women, then 0.1 seconds
-could not be considered fair for men, as the probability of observing reaction
-times lower than 0.1 would be elevated.
-
 If there are issues with reaction time such as athletes approaching the 0.1 

### 2023-07-23 Code/RankComparison.Rmd
set an example of linewdith 80
updated glmm model and cleaning

@@
 
+Keep the linewidth under 80. Otherwise the code lines are not complete in the
+generated pdf files.
+
 ```{r}
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, alternative = "greater", method = "ds")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations,
+                alternative = "greater", method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, alternative = "greater", method = "ds", exact = TRUE)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations,
+                alternative = "greater", method = "ds",
+                exact = TRUE, B = 10000)
 ```
@@
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, alternative = "greater", method = "ds")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp,
+                alternative = "greater", method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, alternative = "greater", method = "ds", exact = TRUE)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp,
+                alternative = "greater", method = "ds",
+                exact = TRUE, B = 10000)
 ```
@@
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim, alternative = "greater", method = "ds")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,
+                alternative = "greater", method = "ds")
 ```

### 2023-07-23 Manuscript/manuscript.tex
updated glmm model and cleaning

@@
 \usepackage{comment}
-%\linenumbers*[1]
+\linenumbers*[1]
 % %% patches to make lineno work better with amsmath
@@
 0.123 and 0.101 in his preliminary heat and semifinal, respectively.  
-His disqualification in the final heat by reacting in 0.099 seconds instead of 
-0.1 was met
+His disqualification in the final heat by reacting in 0.099 seconds, just 0.001
+seconds faster than the 0.1 threshold set by the International Association of
+Athletics Federations (IAAF), was met
 with shock from both Allen and fans alike.  NBC, who was broadcasting the
@@
 
-%2)
 Allen's disqualification has been discussed extensively in online communities
@@
 
-\jy{Use double blank lines to separate paragraphs in tex source}
+\jy{Use double blank lines to separate paragraphs in tex source. Acknowledge if you copy.}
 
 Reaction time in sprint and hurdle events has been studied by many authors.
-\citet{pain2007sprint} examined
-reaction times for nine male athletes.  The
+\citet{pain2007sprint} examined reaction times for nine male athletes.  The
 fastest athlete of the nine had, on average (even with their two fastest times 
@@
 muscular tightening conditions \citep{pain2007sprint}. 
-In a study commissioned by International Association of Athletics Federations 
-(IAAF), \citet{komi2009iaaf} recommended that the disqualification 
+In a study commissioned by IAAF,
+\citet{komi2009iaaf} recommended that the disqualification 
 barrier be decreased to allow sprinters to
 react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
-Additionally, it was recommended to IAAF that high speed cameras be used to
+Unfortunately, IAAF has not changed the reaction barrier since it was
+implemented at the 1999 World Track and Field Championships.
+It is unknown how many athletes since the 2009 study could have potentially
+benefited from this, but it is probable that there are others besides Allen who
+were disqualified. Additionally,
+\citet{komi2009iaaf} recommended that high speed cameras be used to
 change the reaction time criteria from pushing off on the block to body
@@
 
-\jy{This paragraph reads disconnected. Should it be merged with paragraph~3
-  where the study was reviewed?}
-Despite their study, IAAF has not changed the reaction barrier since it was
-implemented at the 1999 World Track and Field Championships.  It is unclear
-why IAAF would commission a study only to disregard the results and not
-make the recommended changes.  It is unknown how many athletes since 2009 could
-have potentially benefited from this, but it is probable that there are others
-besides Allen who were disqualified.
-
 The objective to this paper is to analyze the reaction times at the 2022 World 
@@
 Championships, using a gamma linear mixed-effects model with a random effect 
-accounting for the venue (or year). 
+accounting for the venue (or year) and a second random effect accounting for the
+heat effect within each year. \jy{do a thorough check to make sure this is updated.}
 Based on the fitted model, we can then estimate the probability of a reaction 
@@
 
+\jy{give specifics of your sample size}
+
 \begin{figure}[tbp]
@@
 the size and mean/median,std dev, size of the data.}
-\jy{Show dot plots which would give the distribution; without uncertainty
-  measure, the mean does not tell much.}
 
@@
 much inferior to the gamma model which better captures the skewness of the
-data. A random effect for the venue or year was also found to be necessary, 
-which makes intuitive sense given the discussion in Section~\ref{sec:dataworld}. 
-Let $Y_{ij}$ be the reaction time of observation~$j$ in year~$i$.
-%\eds{we know that many of the athletes compete in multiple world championships,
-%  how is this athlete clustering being accounted for here?}
-%\jy{We could add an athlete random effect too, if the players are identified in
-%  the data.}
-%\jy{Owen, do we have athlete id in the data?}
-Conditional on a venue effect $z_i$ of year~i,  $Y_{ij}$ is modeled by 
+data. Two random effects were found useful: one captures the venue effect in
+each year and the other captures the heat effect within each year.
+Let $Y_{ijk}$ be the reaction time of observation~$k$ in heat~$j$ in year~$i$.
+Conditional on a venue effect $v_i$ of year~$i$ and a heat effect
+$h_{i/j}$ nested within each year~$i$, $Y_{ijk}$ is modeled by a
 gamma distribution. In the classic generalized linear model (GLM) sense,
-$Y_{ij}$ has conditional mean
+$Y_{ijk}$ has conditional mean $E(Y_{ijk} | v_i, h_{i/j}) > 0$ and a
+dispersion parameter~$\phi > 0$. The conditional mean is characterized by
 \[
-g\{E(Y_{ij} | z_i)\} = \alpha + z_i,
+g\{E(Y_{ijk} | v_i, h_{i/j})\} = \alpha + v_i + h_{i/j},
 \]
-and dispersion parameter $\phi$, where $g$ is a known link function and
-$z_i$ is a normally distributed random effect with mean zero and
-variance~$\sigma^2$. Although \citet{lo2015idlink} suggested that the gamma 
-family with an identity link works well for reaction time data, the logarithm 
+where $g$ is a known link function,
+$\gamma_i$ is normally distributed with mean zero and
+variance~$\sigma_v^2$, and $\delta_{i/j}$ is normally distributed with mean
+zero and variance~$\sigma_h^2$.
+Although \citet{lo2015idlink} suggested that the gamma 
+family with an identity link works well for reaction time in cognitive
+psychological research, the logarithm
 link function worked better for our data and additionally ensures the positivity
@@
 commonly used shape and scale parameters. For a gamma distribution with
-shape~$a$ and scale~$b$, we have mean $\mu = ab$ and variance $v = ab^2$. The
-variance as a function of the mean $\mu$ in the GLM sense is
-$v(\mu) = \mu^2 / a$, which implies that the dispersion parameter is
+shape~$a$ and scale~$b$, we have mean $\mu = ab$ and variance $\nu = ab^2$. The
+variance as a function of the mean~$\mu$ in the GLM sense is
+$\text{var}(\mu) = \mu^2 / a$, which implies that the dispersion parameter is
 $\phi = 1 / a$. Therefore, for the gamma mixed-effect model, the conditional
-gamma distribution of the reaction time $Y_{ij}$ given the venue effect $z_i$
-has shape $a = 1 / \phi$ and scale $b = g^{-1}(\alpha + z_i) \phi$. The marginal
+gamma distribution of the reaction time $Y_{ij}$ given the venue
+effect~$v_i$ and heat effect~$h_{i/j}$
+has shape $a = 1 / \phi$ and scale
+$b = g^{-1}(\alpha + v_i + h_{i/j}) \phi$. The marginal
 distribution of $Y_{ij}$ is a scale-mixture of gamma distributions, which can be
@@
 disqualification barrier.
-
-Finally, we remark here that it is also possible to include a heat or race 
-\eds{heat? when I first read this, I thought you meant race as in ethnicity}
-\of{Yes I think heat or batch effect works better, especially now that I have
-removed most talk about heat from the data section} 
-\eds{I see, I did not understand this to be Batch effect when I read it the 
-first time.  I commented out the Batch text in the Data section - I think it is 
-fine to just discuss it here, as it ultimately was not used.  I will defer to 
-Dr Yan on this, however.}
-effect to account for some heats possibly having faster or slower 
-reaction times for everyone involved in the heat.
-
-Although it is possible that one athlete reacting faster could lead to others reacting
-similarly, this does not make much practical sense as the difference between
-heats that had fast and slow reaction times are all within 0.12 seconds of each
-other. 
-\eds{Did you try it and find that the effect was not needed?  If yes,
-I would include that here.  That argument 
-would be more compelling for its non-use than its difficulty in interpretation.}
-\of{We found that including the race effect resulted in the venue effect not
-being significant, and thus Prof Yan suggested that we don't include it.  I am
-wondering if this is even worth mentioning? Prof Yan, what do you think?} 
-\jy{We can revisit the results. It is not surprising that venue effect is no
-  longer significant after the batch effect was included. We could do the same
-  analysis with the batch-random-effect model.}
 
@@
 standardized to form a statistic $Z$ which follows standard normal distribution
-asymptotically \citep[p.910]{datta2005rank}. An implementation of this method is
+asymptotically \citep[p.910]{datta2005rank}.
+
+With our small sample size, the asymptotic normal distribution may not be
+reliable, so we use random permutation to obtain the p-value of the observed
+test statistics $W^*$. This method is
 available from the \texttt{clusWilcox.test()} function with
-\texttt{method = `ds'} from R package \texttt{clusrank}
-\citep{jiang2017wilcoxon}.
-\eds{are you including preliminary heat data as possible within cluster 
-observations in this analysis?  Based on your 
-discussion in Section 2.1 about how they are not comparable, it may not be fair 
-to include them here?  Or perhaps do the analysis with and without prelim heat
-data to see if it makes a difference?}
-\of{Yes I am including prelminary heat data.  The primary reason is that there
-would be few datapoints otherwise.  I ran the numbes without prelims, the number
-of observations dropped from 65 to 26 and the number of clusters from 15 to 9, the
-p value increased from 0.000155 to 0.007332.
-The calculations are at the bottom of RankComparison.Rmd}
+\texttt{method = `ds'} and \texttt{exact = TRUE} from R package
+\texttt{clusrank} \citep{jiang2020wilcoxon}.
 
@@
 
-\jy{motivate first why you want to run the analysis with and without the 2022 data.}
+\jy{Use the pooled data; present the analysis with 2022 first; then suggest that
+  is worth looking at the results without 2022.}
+
+\jy{Write the story up as if you are going to submit. Then I'll review it.}
 
@@
 
-\jy{Close captions with a period.}
-\of{Resolved}
 \begin{table}
@@
 
-\jy{label the table and reference to it; clean it up}
 \begin{table}
@@
 Championships to 2022 World Championships and can be found in 
-section~\ref{sec:results2019vs2022} and secondly comparing 2022 national
+Section~\ref{sec:results2019vs2022} and secondly comparing 2022 national
 championships to the 2022 World Championships and can be found in
-~section\ref{sec:resultsnational}.
-\jy{Remind the readers what the analysis was about: two-group comparison with
-  group being ....}
-\of{Better?}
+Section~\ref{sec:resultsnational}.
 
-\subsubsection{2019 World Track and Field Championships} \label{sec:results2019vs2022}
+\jy{use two-sided}
+
+\jy{include sample sizes; permutation number (B);}
+
 The p value from the two group comparison for 2019 and 2022 was 0.001464.  This 
@@
 Championships are significantly different.  
-\eds{That is not the only explanation} \of{You are correct so I changed the 
-wording}
+\jy{elaborate on the inprepretation; what we have rejected;}
 
-\subsubsection{2022 National Championships} \label{sec:resultsnational}
 The p value from the two group comparison between the 2022 national
@@
 \section{Discussion}\label{sec:Discussion}
+
+\jy{condense; could use chatGPT to improve your writing.}
+
 One conclusion from the analysis in \citet{zhang2021correlation} was that 

### 2023-07-26 Code/GLMM.Rmd
Changed blue curve and smoothed exact=True

@@
                scale = exp(coef(base)) * sbase$dispersion), 
-        0.08, 0.22, add = TRUE, col = "blue")  # fixed effect only
+        0.08, 0.25, add = TRUE, col = "blue")  # fixed effect only
 }
@@
 sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.1)
+```
+```{r}
+sim_two_effect_time(men_old_finals_two_effect_gamma_log, men_old_finals_one_gamma_log, 0.08)
+sim_two_effect_time(men_finals_two_effect_gamma_log, men_finals_one_gamma_log, 0.08)
+sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.08)
+sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.08)
 ```
@@
                scale = exp(coef(base)) * sbase$dispersion), 
-        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
+        0.08, 0.25, add = TRUE, col = "blue") # fixed effect only
 }

### 2023-07-26 Code/RankComparison.Rmd
Changed blue curve and smoothed exact=True

@@
 ```
+```{r}
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, method = "ds",
+                exact = TRUE, B = 10000)
+```
 
@@
 clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp,
-                alternative = "greater", method = "ds")
+                method = "ds")
 
 clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp,
-                alternative = "greater", method = "ds",
+                method = "ds",
                 exact = TRUE, B = 10000)

### 2023-07-26 Manuscript/manuscript.tex
Responded to some comments

@@
 other datasets they considered, but they did not perform any formal 
-statistical analyses. This paper questions the reaction time 
+statistical analysis. This paper questions the reaction time 
 disqualification barrier, which is currently 0.1 seconds, to determine whether 
@@
 the reaction times of every non-disqualification since the 1999 World 
-Championships, using a gamma linear mixed-effects model with a random effect 
-accounting for the venue (or year) and a second random effect accounting for the
-heat effect within each year. \jy{do a thorough check to make sure this is updated.}
+Championships, using a gamma linear mixed-effects model with two random effects 
+one effect for the year and a second effect for the
+heat within each year. \jy{do a thorough check to make sure this is updated.}
+\of{This should be better now.}
 Based on the fitted model, we can then estimate the probability of a reaction 
@@
   \centering
-  \caption{Probabilities of observing reaction times less than 0.1, 0.08, and
-  0.08 for various data sets.}
+  \caption{Probabilities of observing reaction times less than 0.1 and 0.08
+  for various data sets.}
   \begin{tabular}{c c c} 
@@
    \midrule
-   Old Finals & $4.36\cdot10^{-4}$ & $3.0\cdot10^{-7}$ \\
-   All Finals & $8.64\cdot10^{-4}$ & $6.0\cdot10^{-7}$ \\
    Old Pooled & $4.15\cdot10^{-3}$ & $4.67\cdot10^{-5}$ \\ 
@@
    \midrule
-   Old Finals & $0.103$ & $0.095$ \\
-   All Finals & $0.101$ & $0.093$ \\
    Old Pooled & $0.093$ & $0.083$ \\ 
@@
 
-\begin{table}
-  \centering
-  \caption{Compares the various effect models by dataset, all probabilities are
-  for observing a reaction time less than 0.1 seconds. In general the model with
-  the highest probability (least extreme reaction time) is the Year Effect model.}
-  \begin{tabular}{c c c c} 
-   \toprule
-   Data Set & Year Effect & Batch Effect & Both Effects \\ 
-   \midrule
-   Old Finals & $4.39\cdot10^{-4}$ & $4.24\cdot10^{-4}$ & $4.36\cdot10^{-4}$ \\
-   All Finals & $8.63\cdot10^{-4}$ & $8.65\cdot10^{-4}$ & $8.64\cdot10^{-4}$ \\
-   Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
-   All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
-   \bottomrule
-  \end{tabular}
-  \label{tab:EffectComparison}
-\end{table}
+%\begin{table}
+  %\centering
+  %\caption{Compares the various effect models by dataset, all probabilities are
+  %for observing a reaction time less than 0.1 seconds. In general the model with
+  %the highest probability (least extreme reaction time) is the Year Effect model.}
+  %\begin{tabular}{c c c c} 
+  % \toprule
+  % Data Set & Year Effect & Batch Effect & Both Effects \\ 
+   %\midrule
+   %Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
+   %All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
+   %\bottomrule
+  %\end{tabular}
+  %\label{tab:EffectComparison}
+%\end{table}
 

### 2023-07-26 Manuscript/manuscript.tex
added table
Not compiling correctly

@@
 
+\begin{table}
+  \centering
+  \caption{}
+  \begin{tabular}{c c c c c} 
+   \toprule
+   Comparison & $exact = True$ & $exact \neq True$ & Sample size & Clusters/Athletes \\ 
+   \midrule
+   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10{-2}$ & 65 & 15 \\ 
+   International & $2.2\cdot10^{-16}$ & $3.1\cdot10^{-4}$ & 61 & 15 \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:Clusrankresults}
+\end{table}
 The p value from the two group comparison for 2019 and 2022 was 0.001464.  This 
@@
 reaction time repeatedly may indicate skill rather than luck. Devon Allen is
-an exceptional athlete, as evidenced by his football career at the University of
-Oregon and making it onto the Philadelphia Eagles practice squad in 2022
-\citep{hurley2022eagles}.  \citet{komi2009iaaf}'s paper showed that athletes are
+an exceptional athlete, as evidenced by his football career outside of running.  
+\citet{komi2009iaaf}'s paper showed that athletes are
 capable of reacting in less than 0.1 seconds, and if this is the case, it is not

### 2023-07-26 Manuscript/manuscript.tex
Added table back

@@
 
-%\begin{table}
-  %\centering
-  %\caption{Compares the various effect models by dataset, all probabilities are
-  %for observing a reaction time less than 0.1 seconds. In general the model with
-  %the highest probability (least extreme reaction time) is the Year Effect model.}
-  %\begin{tabular}{c c c c} 
-  % \toprule
-  % Data Set & Year Effect & Batch Effect & Both Effects \\ 
-   %\midrule
-   %Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
-   %All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
-   %\bottomrule
-  %\end{tabular}
-  %\label{tab:EffectComparison}
-%\end{table}
+\begin{table}
+  \centering
+  \caption{Compares the various effect models by dataset, all probabilities are
+  for observing a reaction time less than 0.1 seconds. In general the model with
+  the highest probability (least extreme reaction time) is the Year Effect model.}
+  \begin{tabular}{c c c c} 
+   \toprule
+   Data Set & Year Effect & Batch Effect & Both Effects \\ 
+   \midrule
+   Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
+   All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:EffectComparison}
+\end{table}
 
@@
    2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10{-2}$ & 65 & 15 \\ 
-   International & $2.2\cdot10^{-16}$ & $3.1\cdot10^{-4}$ & 61 & 15 \\
+   Nat vs Inter & $2.2\cdot10^{-16}$ & $3.1\cdot10^{-4}$ & 61 & 15 \\
    \bottomrule
@@
 \end{table}
-The p value from the two group comparison for 2019 and 2022 was 0.001464.  This 
-value is for a one-sided test and show that the mean reaction time on average in
-2019 was larger (slower) than the mean reaction time in 2022 for athletes who 
-competed in both championships at an $\alpha$ level of 0.01. This provides good
-evidence that on average, the reaction times from the 2019 and 2022 World 
-Championships are significantly different.  
-\jy{elaborate on the inprepretation; what we have rejected;}
 
-The p value from the two group comparison between the 2022 national
-championships and World Championships was 0.000155. This value is for a one sided
-test and the result is significant at $\alpha = 0.01$ and shows that
-the times at the World Championships were significantly faster than at the
-national championships.
+Table~\ref{tab:Clusrankresults} shows the results generated from four tests in R.
+We use a \texttt{clusWilcox.test()} and set the permutation number (B) to be
+1000 as the sample size in this analysis is small.  The sample size indicates
+the total number of observations while clusters represnts the number of athletes
+in each dataset.  It is worth noting that it is entirely coincidental that there
+are fiftenn athletes in sets.
 

### 2023-07-30 Code/RankComparison.Rmd
Added explanations to code

@@
 
+The code below reads in the data comparing the 2022 National Competitions to the 2022
+World Championships.
 ```{r}
-nations <- read.csv("../Data/ClusrankNatvsWorld.csv")
-colnames(nations) <- c("Time", "Athlete", "Year")
+national <- read.csv("../Data/ClusrankNatvsWorld.csv")
+colnames(national) <- c("Time", "Athlete", "Stage")
 ```
 
-Keep the linewidth under 80. Otherwise the code lines are not complete in the
-generated pdf files.
+The code below performs two clustered Wilcox tests, the first one without using
+exact = True, and the second one with exact = True and 10,000 random
+permutations. The tests compare the times and group by the Stage of the
+competition which in the data is coded as 1 if it was a national championship
+and 2 if it was the world championships. The tests cluster by athlete, of which
+for each dataset an integer from 1-15 is used (the values are not significant).
+For more information about the "ds" method, please look at the Rank-based
+Comparison subsection of the Methods section of the paper or at this link here:
+https://www.jstatsoft.org/article/view/v096i06.  The second test also includes
+exact = TRUE which is used when sample sizes are small (as they are in this
+case), additionally the number of random permutations (B) is set to be 10,000.
 
@@
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations,
-                alternative = "greater", method = "ds")
+clusWilcox.test(Time, group=Stage, cluster = Athlete, data = national,
+                method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations,
-                alternative = "greater", method = "ds",
-                exact = TRUE, B = 10000)
-```
-```{r}
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, method = "ds",
+clusWilcox.test(Time, group=Stage, cluster = Athlete, data = national,
+                method = "ds",
                 exact = TRUE, B = 10000)
@@
 
+The code below reads in the data comparing the 2019 and 2022 World Championships.
 ```{r}
@@
 
+The tests below are very similar to the ones above, the only difference being
+grouping by Year instead of Stage (1 if 2019 and 2 if 2022).
 ```{r}
@@
 
+This is seeing what would happen if we remove prelims, the sample sizes in the
+test become very small: 26 observations and 9 clusters.
 ```{r}
@@
 clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,
-                alternative = "greater", method = "ds")
+                method = "ds")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,
+                method = "ds", 
+                exact = TRUE, B = 10000)
 ```

### 2023-08-03 Manuscript/manuscript.tex
Overhealued GLMM results

@@
 \jy{Use double blank lines to separate paragraphs in tex source. Acknowledge if you copy.}
+\of{Resolved}
 
@@
 within the cluster (athlete).  We chose to include preliminary rounds in this
-section as well as in \ref{sec:data2019} in order to increase the size of our
+section as well as in section~\ref{sec:data2019} in order to increase the size of our
 data.  Without data from the preliminary rounds, the number of observations
 dropped from 65 to 26 and the number of athletes included in the analysis
-decreased from 15 to 9.  Likewise
+decreased from 15 to 9.
 
@@
 
-\of{Add something like this to the start of the paragraph?}
-We may want to consider analyzing all four data sets separately because
-comparing the model based on all the data to the model based on data excluding 
-2022 can provide an indication about the impact of the 2022 data.
-We hope to see that by including 2022 data,
-the probability of observing a significantly low reaction time increases. 
-Table~\ref{tab:Gamma_parameters} summarizes the results from the gamma model 
-from different data sets and shows the standard error of the year
+A summary of the two effect model is shown in table~\ref{tab:Gamma_parameters}
+for the pooled datasets and shows the standard error of the year
 effect increases for both the men's final and men's pooled data sets when
-including 2022.  The men's final data venue effect error increased from 0.439
-to 0.504 and from 0.0495 to 0.0554 for the men's pooled data.  This demonstrates
-that the inclusion of 2022 increased the randomness of the data and motivates
-the rank-based study (see Section~\ref{subsec:Results_Rank}).
+including 2022. We choose to only look at the pooled datasets, as R displayed
+warnings when creating the finals only models
+\of{Is this section okay? Please let me know what I should rework}
 
@@
   \centering
-  \caption{Summary statistics for the GLMM on various data sets.} 
-  \begin{tabular}{c c c c c c}
+  \caption{Summary statistics for the two effect GLMM on various data sets. SE
+  is short for standard error.} 
+  \begin{tabular}{c c c c c}
       \toprule
-      Dataset & Venue standard error & Int effect est & Int se & Int var & Resid Var \\
+      Dataset & Batch SE & Year SE & Intercept & Residual SE \\
       \midrule
-      Old Finals & 0.0439 & $-1.901$ & 0.0251 & 0.0019 & 0.0109 \\
-      All Finals & 0.0504 & $-1.914$ & 0.0280 & 0.0025 & 0.0112 \\
-      Old Pooled & 0.0495 & $-1.853$ & 0.0305 & 0.0024 & 0.0243 \\
-      All Pooled & 0.0554 & $-1.869$ & 0.0348 & 0.0031 & 0.0235 \\
+      Old Pooled & 0.060 & 0.034 & -1.865 & 0.139 \\
+      All Pooled  & 0.059 & 0.041 & -1.880 & 0.014 \\
       \bottomrule
@@
 \end{table}
+
+Two points of this section are to determine how extreme a reaction time below
+0.1 seconds is, and to determine how the exclusion of 2022 could alter results.
+Using the gamma function with link set to log, the probability of the reaction 
+time being less than 0.1 seconds can be calculated in R using the methods 
+described in section~\ref{sec:glmm}. A simulation size of $10,000,000$ was 
+chosen to try and estimate the probability of observing a reaction time below 
+two thresholds: 0.1 and 0.08. When looked at the All Pooled datset, a probability
+of $5.15\cdot10^{-3}$ corresponds to roughly 1 in every 195 starts
+being below 0.1 seconds.  When we exclude 2022, the probability of observing a
+reaction time below 0.1 seconds decreased to $4.15\cdot10^{-3}$, or 1 in every
+240 starts.
+\of{I feel like with the two effect model the probability change is not as
+different as it was previously}
+
+Table~\ref{tab:Sim_probability} shows that changing the reaction time 
+barrier from 0.1 seconds to 0.08 seconds drastically reduces the chances of 
+observing a reaction time that would break the barrier. If World Athletics were
+to change the reaction time the way that \citet{komi2009iaaf} suggests, the 
+probability would be so much lower that athletes would likely not
+have a case to make when arguing against a disqualification.
 
@@
 
+We can use the same model but alter our interpretation slightly to select a 
+reaction time barrier given the probability of observing a time less than the 
+barrier.  For example, if we want a 0.1\% chance of observing an illegally fast 
+reaction time for the men's pooled dataset that includes 2022, then the reaction
+time barrier should be 0.089 seconds, as shown in Table~\ref{tab:Sim_time}.  We
+can repeat this calculation for each dataset and for various probability levels,
+for example a probability of 0.0001 (one disqualification for every 10,000 starts)
+results in a reaction time barrier of 0.079 for the men's pooled dataset.  
+Tables~\ref{tab:Sim_time} shows that the inclusion of 2022 results in a lower
+reaction time barrier, further showing the claims above that the reaction times
+in 2022 were low.
+
 \begin{table}
@@
 
+\of{Should I report table~\ref{tab:EffectComparison}}
 \begin{table}
@@
 \end{table}
-
-Using the gamma function with link set to log, the probability of the reaction 
-time being less than 0.1 seconds can be calculated to see if Allen's 
-disqualification was an anomaly as described in Section~\ref{sec:glmm}.
-Calculations were performed in R using the methods described in 
-section~\ref{sec:glmm} with the use of the \citet{lme4} package.
-
-At a size of $n=10,000,000$, \eds{use different letter as $n$ was already used 
-for number of athletes, $B?$} \of{Could I just say "At a size of 10,000,000"? 
-Otherwise I am fine with using "B" instead of "n"} \eds{we may have lost the 
-rest of this sentence?}
-
-The calculated probabilities of a reaction time 
-being below 0.1 seconds for the various different data sets are shown in 
-Table~\ref{tab:Sim_probability}. Prior to the 2022 World Championships, there 
-was a 0.0439\% chance (based on the
-simulated results) of observing a reaction time below 0.01 seconds in the men's
-finals.
-That means that we would expect a time below the threshold to occur
-once every 2,277 starts.  Considering there are roughly 8 men's finals reaction
-times every 2 years (the World Championships are biennial), every 569 years there
-should be a time as extreme as Allen's. This seems extremely unlikely, and thus
-it seems like there may be an additional explanation for the fast reaction times.
-
-When the 2022 data is included in the analysis, it is interesting that the 
-probability of observing a reaction time below 0.1 seconds increases drastically 
-for the finals data and increases by a smaller margin for the pooled data.  By
-including 2022 in the data, the chances of observing an extreme time are 
-increased, suggesting that 2022 is a high outlier for reaction times compared to
- other years.
-
-Table~\ref{tab:Sim_probability} also shows that changing the reaction time 
-barrier from 0.1 seconds to 0.08 seconds drastically reduces the chances of 
-observing a reaction time that would break the barrier.  For the dataset of all 
-finals data, the probability
-associated with observing a reaction time below 0.08 seconds was $6\cdot10^{-7}$.
-If IAAF \eds{or World Athletics?} were to change the reaction time the way that 
-\citet{komi2009iaaf}
-suggests, the probability would be so much lower that athletes would likely not
-have a case to make when arguing against a disqualification.
-
-We can use the same model but alter our interpretation slightly to select a 
-reaction time barrier given the probability of observing a time less than the 
-barrier.  For example, if we want a 0.1\% chance of observing an illegally fast 
-reaction time for the men's pooled dataset that includes 2022, then the reaction
-time barrier should be 0.089 seconds, as shown in Table~\ref{tab:Sim_time}.  We
-can repeat this calculation for each dataset and for various probability levels,
-for example a probability of 0.0001 (one disqualification for every 10,000 starts)
-results in a reaction time barrier of 0.079 for the men's pooled dataset.  
-Tables~\ref{tab:Sim_time} shows that the inclusion of 2022 results in a lower
-reaction time barrier, further showing the claims above that the reaction times
-in 2022 were low.
-
-%Discussion should show that random effect
-%and standard error increase and that 2022 is special.  Now talk about Seiko
-%timing equipment
 
@@
 improvement for British, Polish, French, Brazilian, and Spanish athletes.
-
-It is worth noting that in the semifinals of the 2022 World Track and
-Field Championships, Allen's reaction time was 0.101 which is only 0.001 above
-the legal limit. What this suggests is that Allen may have strong reflexes
-and be able to react extremely well to the sound of the gun. It seems unlikely 
-that Allen would be able to correctly predict the gun with such precision two 
-times in a row, which is the exact reason that the 0.1 second reaction time 
-disqualification barrier was imposed in the first place. Thus having a quick 
-reaction time repeatedly may indicate skill rather than luck. Devon Allen is
-an exceptional athlete, as evidenced by his football career outside of running.  
-\citet{komi2009iaaf}'s paper showed that athletes are
-capable of reacting in less than 0.1 seconds, and if this is the case, it is not
-unreasonable to conclude that an outstanding athlete like Allen is capable of
-reacting below the 0.1 second barrier.
-
-The results from the GLMM are described in section~\ref{subsec:Results_GLMM} and
-show that there
-
-Devon Allen's disqualification at the World Track and Field Championships was
-possibly the result of both a faulty timing equipment and Devon Allen
-reacting extremely quickly to the start.  Our analysis showed that
-the 2022 World Track and Field Championships were a low outlier compared to
-every other year in terms of average reaction time.  Additionally, it was shown
-through a gamma mixed effects model that there are significant year and race
-effects \eds{remove race effects?}\of{I think were going to go with the two
-effect model so no.} that greatly impact reaction time.
 

### 2023-08-03 Code/GLMM.Rmd
added small code

@@
 
+```{r}
+summary(men_pooled_year_gamma_log)
+
+```
+
 This creates the batch effect only models.
@@
                                data = men_pooled, family = Gamma(link = "log"))
+```
+```{r}
+summary(men_pooled_two_effect_gamma_log)
 ```

### 2023-08-03 Manuscript/manuscript.tex
Fixed important issues with bib file and cleaned up manuscript

@@
 
-\subsection{Rank-based Comparison}\label{sec:rank}
+\subsection{Rank-based Comparison}\label{sec:rank_methods}
 
@@
 With our small sample size, the asymptotic normal distribution may not be
-reliable, so we use random permutation to obtain the p-value of the observed
-test statistics $W^*$. This method is
-available from the \texttt{clusWilcox.test()} function with
-\texttt{method = `ds'} and \texttt{exact = TRUE} from R package
+reliable, so we use random permutation, which simulates the observed test 
+statistic: $W^*$ given number of times to obtain the p-value of said test 
+statistics. This method is available from the \texttt{clusWilcox.test()} function
+with \texttt{method = `ds'} and \texttt{exact = TRUE} from R package
 \texttt{clusrank} \citep{jiang2020wilcoxon}.
@@
 
-The rank based methods described in Section~\ref{sec:rank} are used
+The rank based methods described in are used in section~\ref{sec:rank_methods}
 to determine if the reaction times for athletes who had competed at other 
@@
 a two group comparison performed two times: once comparing 2019 World
-Championships to 2022 World Championships and can be found in 
-Section~\ref{sec:results2019vs2022} and secondly comparing 2022 national
-championships to the 2022 World Championships and can be found in
-Section~\ref{sec:resultsnational}.
+Championships to 2022 World Championships and then comparing 2022 national
+championships to the 2022 World Championships.
 
 \jy{use two-sided}
+\of{Resolved}
 
 \jy{include sample sizes; permutation number (B);}
+\of{Resolved}
+
+Table~\ref{tab:Clusrankresults} shows the results generated from four tests in R.
+We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
+\texttt{B = 1000} indicating that the random permutation size was chosen to be 
+1000.  As the sample sizes are small: 65 and 49, we use \texttt{exact = TRUE} as
+that is preferred.  In both cases, the national versus international comparison
+produced stronger p values than the 2019 versus 2022 comparison, with the
+national vs international \texttt{exact = TRUE} test producing a p value of 
+$2.2\cdot10^{-16}$, which is the smallest possible value acheivable in R.  This
+is a highly significant test result and shows that athletes who competed at the
+2022 Track and Field World Championships were on average, significantly faster
+at reacting than they were a few months prior.  For the 2019 versus 2022
+same athlete comparison with \texttt{exact = TRUE} argument, the p value was
+significant at $\alpha = 0.01$, however for when \texttt{exact = FALSE}, the p
+value was higher and only significant at $\alpha = 0.05$. Nonetheless, these
+tests help to show that athletes competing at the 2022 Track and Field World
+Championships had faster reaction times than when they had competed previously.
 
@@
    2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10{-2}$ & 65 & 15 \\ 
-   Nat vs Inter & $2.2\cdot10^{-16}$ & $3.1\cdot10^{-4}$ & 61 & 15 \\
+   Nat vs Inter & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ & 49 & 11 \\
    \bottomrule
   \end{tabular}
+  \caption{2019 vs 2022 compares data from the same athletes who competed at the
+  2019 and 2022 World Track and Field Championships. Nat vs Inter compares data
+  from the same athletes who competed at the 2022 national-level championships
+  and the 2022 World Track and Field Championships}
   \label{tab:Clusrankresults}
 \end{table}
-
-Table~\ref{tab:Clusrankresults} shows the results generated from four tests in R.
-We use a \texttt{clusWilcox.test()} and set the permutation number (B) to be
-1000 as the sample size in this analysis is small.  The sample size indicates
-the total number of observations while clusters represnts the number of athletes
-in each dataset.  It is worth noting that it is entirely coincidental that there
-are fiftenn athletes in sets.
 
@@
 
-One conclusion from the analysis in \citet{zhang2021correlation} was that 
-athletes' reaction times improved from 2011 to 2019.  However this study does 
-not consider data from 1999 to 2009, and over this timespan athletes get slower 
-and then faster.  Figure~\ref{fig:Boxplots} shows that for the men's pooled data,
-the year with the fastest median reaction time before 2022 was 1999.  Thus, this 
-reject's Zhang et al.'s conclusion that athletes are getting faster because they 
-were fast over twenty years ago. One possible explanation for why 2011 had such
-high variation was because of the IAAF rule discussed earlier that increased the
-penalty of a false start to an instant disqualification \citep{iaaf2009falsestart}.
-Athletes may have been overly cautious to false start and may have consciously
-or sub-consciously reacted slower as a result.  Instead, a more appropriate 
-conclusion may be that there is a lot of variability in reaction times and that
-the context of the races matters.
+%One conclusion from the analysis in \citet{zhang2021correlation} was that 
+%athletes' reaction times improved from 2011 to 2019.  However this study does 
+%not consider data from 1999 to 2009, and over this timespan athletes get slower 
+%and then faster.  Figure~\ref{fig:Boxplots} shows that for the men's pooled data,
+%the year with the fastest median reaction time before 2022 was 1999.  Thus, this 
+%reject's Zhang et al.'s conclusion that athletes are getting faster because they 
+%were fast over twenty years ago. One possible explanation for why 2011 had such
+%high variation was because of the IAAF rule discussed earlier that increased the
+%penalty of a false start to an instant disqualification \citep{iaaf2009falsestart}.
+%Athletes may have been overly cautious to false start and may have consciously
+%or sub-consciously reacted slower as a result.  Instead, a more appropriate 
+%conclusion may be that there is a lot of variability in reaction times and that
+%the context of the races matters.
 
@@
 
-Since 1985 Seiko Holding Corporations has served every World Athletics 
-Championships as the official timer \citep{wa2022seiko}.  Reaction times were
-not officially measured until 1999 in the men's 110 meter hurdles and women's
-100 meter hurdles. Seiko's technology for detecting 
-reaction times relies on the pressure that athletes exert on the plate when they 
-push off.  Their systems measure the time differential between when the ``gun" 
-goes off to start the race and when the pressure changes \citep{wa2022seiko}.  
-If an athlete has a reaction time under 0.1 seconds, it is deemed a false start 
-as it is considered that no athlete can react so quickly \citep{Seiko-Timing}.  
-Thus, a reaction time of for example 0.05, suggests that the athlete predicted 
-the gun and it was luck that caused their abnormally high time.  At the World 
-Championships level, World Athletics want to remove that element of 
-luck and thus impose the 0.1 second barrier.
-\eds{I would argue that this paragraph should go in the discussion as there are 
-no results presented here}
-\of{I agree so I moved it here. I think it also could fit into the top of the
-data section as in section~\ref{sec:Data}. Thoughts?}
-
-\eds{also consider moving the following paragraph to discussion??}
-As Seiko timed both events, there should not be a significant
-difference if the timing equipment in both instances was working correctly.
-\eds{you just mentioned in the last subsection that Seiko upgraded its 
-technology prior to 2022}
-of{Yes I agree}
-To put it simply, there is not a reasonable explanation for 
-this. It is inconceivable that so many athletes would improve such a significant
-amount in such a small amount of time. Six countries are represented in this 
-comparison, which drastically decreases the chances of this issue being caused by
-one specific country.  The United States athletes' improvement in times was
-discussed earlier in the paper, but this data shows that on average there was
-\jy{any reference?}
-improvement for British, Polish, French, Brazilian, and Spanish athletes.
+%Since 1985 Seiko Holding Corporations has served every World Athletics 
+%Championships as the official timer \citep{wa2022seiko}.  Reaction times were
+%not officially measured until 1999 in the men's 110 meter hurdles and women's
+%100 meter hurdles. Seiko's technology for detecting 
+%reaction times relies on the pressure that athletes exert on the plate when they 
+%push off.  Their systems measure the time differential between when the ``gun" 
+%goes off to start the race and when the pressure changes \citep{wa2022seiko}.  
+%If an athlete has a reaction time under 0.1 seconds, it is deemed a false start 
+%as it is considered that no athlete can react so quickly \citep{Seiko-Timing}.  
+%Thus, a reaction time of for example 0.05, suggests that the athlete predicted 
+%the gun and it was luck that caused their abnormally high time.  At the World 
+%Championships level, World Athletics want to remove that element of 
+%luck and thus impose the 0.1 second barrier.
+%\eds{I would argue that this paragraph should go in the discussion as there are 
+%no results presented here}
+%\of{I agree so I moved it here. I think it also could fit into the top of the
+%data section as in section~\ref{sec:Data}. Thoughts?}
 
@@
 
-If there are issues with reaction time such as athletes approaching the 0.1 
-second barrier again at the next World Championships (August 2023), then World 
-Athletics needs to adjust that threshold and allow for faster reaction times so 
-that athletes with superb reflexes are not penalized.  World Athletics did not
-do anything in response to the \citet{komi2009iaaf} study, but perhaps they 
-should now so that the story of Devon Allen is not repeated.
+%If there are issues with reaction time such as athletes approaching the 0.1 
+%second barrier again at the next World Championships (August 2023), then World 
+%Athletics needs to adjust that threshold and allow for faster reaction times so 
+%that athletes with superb reflexes are not penalized.  World Athletics did not
+%do anything in response to the \citet{komi2009iaaf} study, but perhaps they 
+%should now so that the story of Devon Allen is not repeated.
 

### 2023-08-04 Manuscript/manuscript.tex
removed addressed comments

@@
 
-\jy{The bib file needs quality control. Practice my writing tips in the
-  stat-wriging notes: chapter 2 on latex/bibtex.}
-
-\jy{Go over the reference section to see what need to be changed in the bib source}
-
 \section{Introduction}
@@
 \citep{johnson2022was}.
-
-\jy{Use double blank lines to separate paragraphs in tex source. Acknowledge if you copy.}
-\of{Resolved}
 
@@
 one effect for the year and a second effect for the
-heat within each year. \jy{do a thorough check to make sure this is updated.}
-\of{This should be better now.}
+heat within each year.
 Based on the fitted model, we can then estimate the probability of a reaction 
@@
 
-\jy{Use the pooled data; present the analysis with 2022 first; then suggest that
-  is worth looking at the results without 2022.}
-
-\jy{Write the story up as if you are going to submit. Then I'll review it.}
-
 A summary of the two effect model is shown in table~\ref{tab:Gamma_parameters}
@@
 \section{Discussion}\label{sec:Discussion}
-
-\jy{condense; could use chatGPT to improve your writing.}
 

### 2023-08-11 Manuscript/manuscript.tex
Added info about the size of each cluster size

@@
   \centering
-  \includegraphics{Finals_Pooled_Boxplot}
+  \includegraphics{Pooled_Boxplot}
   \caption{The reaction times from 1999 to 2022 for the men's final (top) and
@@
 baseline for the four United States athletes who competed in at least one round 
-of both events: Trey Cunningham, Daniel Roberts, 
-Grant Holloway, and Devon Allen. Table~\ref{fig:USAvsWorld} highlights the 
-difference in reaction times for these four athletes in the two events. 
-Every athlete reacted faster in all the World 
-Track and Field Championships races compared to the United States Track and Field 
-Championships races. Devon Allen raced three times at the United States Track and
-Field Championships and three times at the World Track and Field Championships. 
-His reaction times were substantially faster in the World Track and
-Field Championships compared to the United States Track and Field Championships, 
-as were his teammates. 
-This comparison suggests that the timing system used between the two
-meets may have been meaningfully different to the point where Devon Allen was
-disqualified more because of faulty equipment rather than because he reacted
-too quickly.
-
-\begin{table}
-\begin{center}
-  \caption{Reaction times (in seconds) for four United States Athletes at the 
-	2022 United States Track
-  and Field World Championships (``United States") and the 2022 World Track and Field 
-	Championships (``World"); H denotes preliminary heat, S denotes semifinal, and 
-	F denotes final.
-  ``N/A" was used for any athlete that did not compete and thus did not register 
-  a reaction time. }
-  \begin{tabular}{c c c c c c c} 
-   \toprule
-	 & \multicolumn{3}{c}{United States} & \multicolumn{3}{c}{World} \\
-	\cmidrule(lr){2-4}
-    \cmidrule(lr){5-7}
-   Athlete &  H &  S &  F &  H &  S &  F \\ [0.5ex] 
-   \midrule
-   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
-   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
-   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
-   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
-   \bottomrule
-  \end{tabular}
-  \label{fig:USAvsWorld}
-
-\end{center}
-\end{table}
-
-The dataset of the four United States athletes 
-is too small to perform any meaningful statistical analysis, but
-the data was expanded to include reaction times for athletes who competed at 
-national competitions (many of which were held from May-July of 2022). Once the 
-data was collected, each athlete was designated as a cluster and the stage of
-competition (national or international) was specified for each competition 
-within the cluster (athlete).  We chose to include preliminary rounds in this
+of both events: Trey Cunningham, Daniel Roberts, Grant Holloway, and Devon Allen.   
+Every athlete reacted faster in all the World Track and Field Championships 
+races compared to the United States Track and Field Championships races. The 
+dataset of the four United States athletes is too small to perform any 
+meaningful statistical analysis, but the data was expanded to include reaction 
+times for 11 athletes who competed at national competitions (many of which were 
+held from May-July of 2022). This data is difficult to find as it was taken from 
+various websites and was at times difficult to interpret as the data if often 
+recorded in the language of the country. \of{I don't know how to say this}. Each
+athlete was designated to be a cluster and the stage of competition (national or international) was specified for each competition within the cluster (athlete).
+Thus, this is subunit level grouping where we assign the treatment: national or
+international within each cluster.  Cluster sizes range from three to six, with
+one cluster of size three, six clusters of size four, two clusters of size five
+and two clusters of size six.  We chose to include preliminary rounds in this
 section as well as in section~\ref{sec:data2019} in order to increase the size of our
@@
 remember that these are some of the fastest sprinters on the planet, and thus
-improving on elite times is not easy.  Nonetheless, this is a worthy inclusion
-to the data because Seiko was the timing company responsible for both events and
-because the level of competition is the extremely elite in both cases.
+improving on elite times is not easy.  Once again we repeat a similar procedure
+as in section~\ref{sec:datanational}; we assign each athlete to be a cluster and
+within each cluster assign times to be from either the 2019 World Championships
+or the 2022 World Championships.  In this analysis, we look at data from 15
+athletes, where cluster size ranges from two to six.   There were two clusters 
+of size two, three clusters of size three, five clusters of size four, two 
+clusters of size five and three clusters of size six.
 

### 2023-08-15 Manuscript/manuscript.tex
Almost done with final changes

@@
  provides the results of the two analyses and ends 
-by questioning the reaction time barrier.  Lastly, Section~\ref{sec:Discussion}
+by questioning the reaction time barrier.  Lastly, Section~\ref{sec:concludingremark}
 discusses the impact and limitations of the paper.
@@
 With our small sample size, the asymptotic normal distribution may not be
-reliable, so we use random permutation, which simulates the observed test 
+reliable, so we use 1000 random permutations to simulate the observed test 
 statistic: $W^*$ given number of times to obtain the p-value of said test 
@@
 including 2022. We choose to only look at the pooled datasets, as R displayed
-warnings when creating the finals only models
-\of{Is this section okay? Please let me know what I should rework}
+warnings when creating the finals only models.  The two effect model with venue
+and heat effect was the best model, as proven by an ANOVA test and decrease in
+AIC.  For the entire pooled dataset; the AIC for the venue effect only model was
+$-1847.9$, the heat effect only model AIC was $-1894.4$ and the both effect model,
+which we will present throughout this section AIC was $-1898.9$. Thus, the two
+effect model is superior and will be presented throughout the rest of the paper.
+\of{How specific do I need to be in this section? Will people who read this paper
+run the code to see the p value or should I say it directly and how do I say it?}
 
@@
   is short for standard error.} 
-  \begin{tabular}{c c c c c}
+  \begin{tabular}{c c c c c c}
       \toprule
-      Dataset & Batch SE & Year SE & Intercept & Residual SE \\
+      Dataset & AIC & Batch SE & Year SE & Intercept & Residual SE \\
       \midrule
-      Old Pooled & 0.060 & 0.034 & -1.865 & 0.139 \\
-      All Pooled  & 0.059 & 0.041 & -1.880 & 0.014 \\
+      Old Pooled & 0.060 & 0.034 & $-1.865$ & 0.139 \\
+      All Pooled & 0.059 & 0.041 & $-1.880$ & 0.014 \\
       \bottomrule
@@
   for various data sets.}
-  \begin{tabular}{c c c} 
+  \begin{tabular}{c c c c} 
    \toprule
-   Data Set & P(RxnTime $<$ 0.1) & P(RxnTime $<$ 0.08) \\ 
+   Data Set & P(R. Time $<$ 0.08) & P(R. Time $<$ 0.09) & P(R. Time $<$ 0.1)  \\ 
    \midrule
-   Old Pooled & $4.15\cdot10^{-3}$ & $4.67\cdot10^{-5}$ \\ 
-   All Pooled & $5.15\cdot10^{-3}$ & $6.01\cdot10^{-5}$ \\
+   Old Pooled & $4.67\cdot10^{-5}$ & $5.91\cdot10^{-4}$ &  $4.15\cdot10^{-3}$  \\ 
+   All Pooled & $6.01\cdot10^{-5}$ & $7.39\cdot10^{-4}$ & $5.15\cdot10^{-3}$ \\
    \bottomrule
@@
   \caption{Suggested reaction time barriers based on probability of observing
-  a time less the barrier. \of{Not sure if this makes sense}}
+  a time less than the barrier. \of{Not sure if this makes sense}}
   \begin{tabular}{c c c} 
    \toprule
-   Data Set & P(RxnTime $<$ x = .001) & P(RxnTime $<$ x = 0.0001) \\ 
+   Data Set & P(RxnTime $<$ x = $10^{-3}$) & P(RxnTime $<$ x = $10^{-4}$) \\ 
    \midrule
@@
 
-\of{Should I report table~\ref{tab:EffectComparison}}
-\begin{table}
-  \centering
-  \caption{Compares the various effect models by dataset, all probabilities are
-  for observing a reaction time less than 0.1 seconds. In general the model with
-  the highest probability (least extreme reaction time) is the Year Effect model.}
-  \begin{tabular}{c c c c} 
-   \toprule
-   Data Set & Year Effect & Batch Effect & Both Effects \\ 
-   \midrule
-   Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
-   All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
-   \bottomrule
-  \end{tabular}
-  \label{tab:EffectComparison}
-\end{table}
+%\of{Should I report table~\ref{tab:EffectComparison}}
+%\begin{table}
+%  \centering
+%  \caption{Compares the various effect models by dataset, all probabilities are
+%  for observing a reaction time less than 0.1 seconds. In general the model with
+%  the highest probability (least extreme reaction time) is the Year Effect model.}
+%  \begin{tabular}{c c c c} 
+%   \toprule
+%   Data Set & Year Effect & Batch Effect & Both Effects \\ 
+%   \midrule
+%   Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
+%   All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
+%   \bottomrule
+%  \end{tabular}
+%  \label{tab:EffectComparison}
+%\end{table}
 
@@
    \toprule
-   Comparison & $exact = True$ & $exact \neq True$ & Sample size & Clusters/Athletes \\ 
+   Comparison & Permutation & Asymptotic \\ 
    \midrule
-   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10{-2}$ & 65 & 15 \\ 
-   Nat vs Inter & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ & 49 & 11 \\
+   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10{-2}$ \\ 
+   Nat vs Inter & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ \\
    \bottomrule
@@
   from the same athletes who competed at the 2022 national-level championships
-  and the 2022 World Track and Field Championships}
+  and the 2022 World Track and Field Championships.}
   \label{tab:Clusrankresults}
@@
 
-\section{Discussion}\label{sec:Discussion}
+\section{Concluding Remarks}\label{sec:concludingremarks}
 

### 2023-08-15 Code/GLMM.Rmd
added some stuff to be able to view summary statistics

@@
 summary(men_pooled_year_gamma_log)
-
+summary(men_old_pooled_year_gamma_log)
 ```
@@
                          family = Gamma(link = "log"))
+```
+
+```{r}
+summary(men_pooled_batch_gamma_log)
 ```
@@
 ```
+
+```{r}
+sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.09)
+sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.09)
+```
+
 ```{r}

### 2023-08-17 Manuscript/manuscript.tex
minor edits

@@
 
-\subsection{2022 World Championships}\label{sec:dataworld}
+\subsection{World Championships 1999--2022}\label{sec:dataworld}
 
@@
 between 1999 and 2019.  Thus, we have four datasets with their respective naming
+\jy{Clean up}
 in parentheses: the final's data from 1999 to 2019 (Old Finals), final's data 
@@
 and the median for the men's final data in 2022 was 0.126.  For comparison,
-\citet{brosnan2017effects} found median reaction times of 0.156 seconds 
+{brosnan2017effects} found median reaction times of 0.156 seconds 
 for men when looking at data from 1999 to 2014.
-
 We can also see from Figure~\ref{fig:Boxplots} that the reaction times clearly
@@
 
-\subsection{Beyond 2022 World Championships}\label{sec:databeyond}
+\subsection{Comparison Data with 2022 World Championships}\label{sec:databeyond}
 The motivation for looking beyond World Championships data came from an 
@@
 
-\subsection{Rank-based Comparison}\label{sec:rank_methods}
+\subsection{Rank-based Comparison}\label{sec:rank}
 
@@
 
-A summary of the two effect model is shown in table~\ref{tab:Gamma_parameters}
+A summary of the two effect model is shown in Table~\ref{tab:Gamma_parameters}
 for the pooled datasets and shows the standard error of the year
@@
 championships to the 2022 World Championships.
-
-\jy{use two-sided}
-\of{Resolved}
-
-\jy{include sample sizes; permutation number (B);}
-\of{Resolved}
 

### 2023-08-22 Manuscript/manuscript.tex
Started to go through pooled revisions

@@
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics{Pooled_Boxplot}
-  \caption{The reaction times from 1999 to 2022 for the men's final (top) and
-  the men's semifinals and finals (bottom).}
-  \label{fig:Boxplots}
-\end{figure}
-
 Data was taken from the World Athletics 
@@
 For analysis purposes, we will often pool reaction times of men's final and 
-men's semifinal heats together and refer to this data as the ``pooled data"
-throughout the paper. 
-%\of{Ok I updated this.  Should I delete the batch variable too?  Although I don't
-%use it, I do reference it in the methods section}
-The pooled data increases the sample size which is advantageous because in some
-years there are very few final heat observations.  For example, in 2022 there
-were only five data points due to two disqualifications and one athlete who 
-did not compete.  We also look at data that does not include 2022 with year values
-between 1999 and 2019.  Thus, we have four datasets with their respective naming
-\jy{Clean up}
-in parentheses: the final's data from 1999 to 2019 (Old Finals), final's data 
-from 2019 to 2022 (All Finals), pooled data from 1999 to 2019 (Old Pooled), and 
-pooled data from 1999 to 2022 (All Pooled).
+men's semifinal heats together which increases the sample size which is 
+advantageous because in some years there are very few finals observations, such
+as in 2022 there were only five data points due to two disqualifications and one
+athlete who did not compete.  We also look at data can also look at data that
+excludes 2022 to see how our analysis differs based on the exclusion of one
+particular year of interest.
 
-The data under consideration is best visualized using side-by-side boxplots, as
-shown in Figure~\ref{fig:Boxplots}.  It is clear that the reaction times in the 
+The data under consideration is best visualized using using a boxplots, as
+shown in Figure~\ref{fig:Boxplot}.  It is clear that the reaction times in the 
 2022 boxplot for both 
@@
 false start penalization rules have changed over the study period.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics{Boxplot}
+  \caption{The reaction times from 1999 to 2022 for the men's final (top) and
+  the men's semifinals and finals (bottom).}
+  \label{fig:Boxplot}
+\end{figure}
 
@@
       \toprule
-      Dataset & AIC & Batch SE & Year SE & Intercept & Residual SE \\
+      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\delta_{i/j}$ \\
       \midrule

### 2023-08-23 Manuscript/manuscript.tex
Minor change

@@
 spared waiting until the next World Track and Field Championships.
+\eds{He is not competing this year, correct? If yes, we may need to rephrase.}
 
@@
 benefited from this, but it is probable that there are others besides Allen who
-were disqualified. Additionally,
+could have. Additionally,
+%were disqualified. Additionally,
 \citet{komi2009iaaf} recommended that high speed cameras be used to
@@
 every sprinter who steps onto the track.
+\eds{Since the World Championships are happening again now, is data from this 
+year already available? If it is available, we would need to justify its omission
+in the current analysis.
+If it is not available, but becomes available while under review, would we 
+reanalyze the data?  Does it seem likely our conclusions would change based on
+what is going on this year?}
 
@@
  provides the results of the two analyses and ends 
-by questioning the reaction time barrier.  Lastly, Section~\ref{sec:concludingremark}
+by questioning the reaction time barrier.  Lastly, Section~\ref{sec:concludingremarks}
 discusses the impact and limitations of the paper.
@@
 of the 2022 World Championships). 
-\eds{is this true? are you still using the 2019 world championships data for the
-rank-based analyses, or only the national competitions?}
-\of{Yes the ranked based comparison has 2 subsections: 2022 national vs 2022
-international and 2019 international vs 2022 international.  Should I instead say
-that there are three types of data across 2 sections or something like that?}
-\eds{I think this is fine here, but it is not at all clear that there are two
-subsections in the rank-based results for 2022 national vs 2022
-international and 2019 international vs 2022 international.}
-\of{Currently under the Methods/Results section I have GLMM and rank based comparison
-do you think I need subsections under rank based comparison, or is that too much
-compartmentalization?  Or should I instead say there are 3 types of data that are
-used in two different methods, or something like that?} \eds{I prefer
-subsections within rank-based comparison.}
+%\eds{is this true? are you still using the 2019 world championships data for the
+%rank-based analyses, or only the national competitions?}
+%\of{Yes the ranked based comparison has 2 subsections: 2022 national vs 2022
+%international and 2019 international vs 2022 international.  Should I instead say
+%that there are three types of data across 2 sections or something like that?}
+%\eds{I think this is fine here, but it is not at all clear that there are two
+%subsections in the rank-based results for 2022 national vs 2022
+%international and 2019 international vs 2022 international.}
+%\of{Currently under the Methods/Results section I have GLMM and rank based comparison
+%do you think I need subsections under rank based comparison, or is that too much
+%compartmentalization?  Or should I instead say there are 3 types of data that are
+%used in two different methods, or something like that?} \eds{I prefer
+%subsections within rank-based comparison.}
 The first data set is used for the GLMM analyses described in 
@@
 For analysis purposes, we will often pool reaction times of men's final and 
-men's semifinal heats together which increases the sample size which is 
+men's semifinal heats together which increases the sample size. This is 
 advantageous because in some years there are very few finals observations, such
-as in 2022 there were only five data points due to two disqualifications and one
-athlete who did not compete.  We also look at data can also look at data that
+as in 2022, when there were only five data points due to two disqualifications 
+and one
+athlete who did not compete.  We also consider data that
 excludes 2022 to see how our analysis differs based on the exclusion of one
@@
 
-The data under consideration is best visualized using using a boxplots, as
-shown in Figure~\ref{fig:Boxplot}.  It is clear that the reaction times in the 
+The data under consideration is summarized using a boxplots in 
+Figure~\ref{fig:Boxplot}.  It is clear that the reaction times in the 
 2022 boxplot for both 
@@
 and the median for the men's final data in 2022 was 0.126.  For comparison,
-{brosnan2017effects} found median reaction times of 0.156 seconds 
+\cite{brosnan2017effects} found median reaction times of 0.156 seconds 
 for men when looking at data from 1999 to 2014.
-We can also see from Figure~\ref{fig:Boxplots} that the reaction times clearly
+We can also see from Figure~\ref{fig:Boxplot} that the reaction times clearly
 differ from year to year. The venue of the World Championships changes each 
@@
   \caption{The reaction times from 1999 to 2022 for the men's final (top) and
-  the men's semifinals and finals (bottom).}
+  the men's semifinals and finals (bottom). 
+	\eds{there is only one plot here, not two. If you are keeping only one, the
+	text and caption need to be modified accordingly}}
   \label{fig:Boxplot}
@@
 \subsubsection{2022 National Competitions}\label{sec:datanational}
-From June 23-26 2022, the United States held its Track and Field Championships 
+From June 23-26, 2022, the United States held its Track and Field Championships 
 at Hayward Field in Eugene, Oregon to decide who to send to the World 
@@
 meaningful statistical analysis, but the data was expanded to include reaction 
-times for 11 athletes who competed at national competitions (many of which were 
-held from May-July of 2022). This data is difficult to find as it was taken from 
-various websites and was at times difficult to interpret as the data if often 
-recorded in the language of the country. \of{I don't know how to say this}. Each
-athlete was designated to be a cluster and the stage of competition (national or international) was specified for each competition within the cluster (athlete).
-Thus, this is subunit level grouping where we assign the treatment: national or
-international within each cluster.  Cluster sizes range from three to six, with
-one cluster of size three, six clusters of size four, two clusters of size five
+times for 11 more athletes who competed at other national competitions (many of which were 
+held from May-July of 2022). This data was difficult to find, as the data is not
+centrally located but instead found on different national websites (often 
+recorded in the national language without translation, and hence not used). 
+\of{I don't know how to say this}. \eds{is my change correct?} Each athlete was 
+designated as a ``cluster'' and the stage of competition (national or international) 
+was specified for each competition within the cluster (athlete).
+Thus, this is subunit level grouping where we assign the treatment, national or
+international, within each cluster.  Cluster sizes range from three to six, with
+one cluster of size three, six clusters of size four, two clusters of size five,
 and two clusters of size six.  We chose to include preliminary rounds in this
-section as well as in section~\ref{sec:data2019} in order to increase the size of our
-data.  Without data from the preliminary rounds, the number of observations
+section as well as in Section~\ref{sec:data2019} in order to increase the size 
+of our data.  Without data from the preliminary rounds, the number of observations
 dropped from 65 to 26 and the number of athletes included in the analysis
@@
 
-\eds{Can you show some descriptive statistics for this data?  Total number of 
-clusters? Average number of international vs national competitions across 
-athletes/clusters?}
-\eds{The following paragraph still needs to be smoothed (the paragraph that was 
-here and the paragraph that was in the intro still need to be integrated into 
-one (or two) cohesive paragraph(s) that accurately describe the two types of 
-data that will be used for two rank-based analyses.}
-\of{Does this work?}
-\eds{Yes, but you still need figure or descriptive statistics for this data.}
+%\eds{Can you show some descriptive statistics for this data?  Total number of 
+%clusters? Average number of international vs national competitions across 
+%athletes/clusters?}
+%\eds{The following paragraph still needs to be smoothed (the paragraph that was 
+%here and the paragraph that was in the intro still need to be integrated into 
+%one (or two) cohesive paragraph(s) that accurately describe the two types of 
+%data that will be used for two rank-based analyses.}
+%\of{Does this work?}
+%\eds{Yes, but you still need figure or descriptive statistics for this data.}
 
@@
 The 2019 World Track and Field Championship data is included in the larger dataset
-described in section~\ref{sec:dataworld}, but we can also use it similarly to
-how we used national level in section~\ref{sec:datanational}.  We can compare
+described in Section~\ref{sec:dataworld}, but we can also use it similarly to
+how we used national level in Section~\ref{sec:datanational}.  We can compare
 the times of athletes who participated in the 2019 and 2022 World Track and Field
@@
 improving on elite times is not easy.  Once again we repeat a similar procedure
-as in section~\ref{sec:datanational}; we assign each athlete to be a cluster and
-within each cluster assign times to be from either the 2019 World Championships
+as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
+within each cluster label times to be from either the 2019 World Championships
 or the 2022 World Championships.  In this analysis, we look at data from 15
@@
 of size two, three clusters of size three, five clusters of size four, two 
-clusters of size five and three clusters of size six.
+clusters of size five, and three clusters of size six.
 
@@
 Semifinals both came from Devon Allen, but the difference between 0.101 and 0.099
-seconds was the difference in what caused him to be disqualified. \eds{check 
-these numbers? 0.099?}\of{resolved}
+seconds was the difference in what caused him to be disqualified. 
+%\eds{check these numbers? 0.099?}\of{resolved}
 
@@
 $h_{i/j}$ nested within each year~$i$, $Y_{ijk}$ is modeled by a
-gamma distribution. In the classic generalized linear model (GLM) sense,
+gamma distribution. 
+\eds{I was talking to my husband about this. Could the heat effect be justified 
+to capture the variability in the amount of time between the ``set'' and the 
+gun going off (assuming this is different from heat to heat)? E.g., quicker 
+gunshots after the ``set'' could lead to slower reaction times (or the reverse), 
+but this effect would be shared across all athletes in the same heat.}
+In the classic generalized linear model (GLM) sense,
 $Y_{ijk}$ has conditional mean $E(Y_{ijk} | v_i, h_{i/j}) > 0$ and a
@@
 $\phi = 1 / a$. Therefore, for the gamma mixed-effect model, the conditional
-gamma distribution of the reaction time $Y_{ij}$ given the venue
+gamma distribution of the reaction time $Y_{ijk}$ given the venue
 effect~$v_i$ and heat effect~$h_{i/j}$
@@
 $b = g^{-1}(\alpha + v_i + h_{i/j}) \phi$. The marginal
-distribution of $Y_{ij}$ is a scale-mixture of gamma distributions, which can be
+distribution of $Y_{ijk}$ is a scale-mixture of gamma distributions, which can be
 easily simulated from once the parameters are estimated. Many
@@
 statistic: $W^*$ given number of times to obtain the p-value of said test 
-statistics. This method is available from the \texttt{clusWilcox.test()} function
+statistics. 
+\eds{I do not understand what is meant in the text after the colon in the 
+previous sentence.  Please rephrase.}
+This method is available from the \texttt{clusWilcox.test()} function
 with \texttt{method = `ds'} and \texttt{exact = TRUE} from R package
@@
 warnings when creating the finals only models.  The two effect model with venue
-and heat effect was the best model, as proven by an ANOVA test and decrease in
-AIC.  For the entire pooled dataset; the AIC for the venue effect only model was
-$-1847.9$, the heat effect only model AIC was $-1894.4$ and the both effect model,
-which we will present throughout this section AIC was $-1898.9$. Thus, the two
+and heat effect was the best model, as indicated by an ANOVA test 
+\eds{I think it may be better to call this a likelihood ratio test rather than
+an ANOVA test, assuming that is the formal test being used} and decrease in
+AIC.  For the entire pooled dataset, the AIC for the venue effect only model was
+$-1847.9$, the AIC for the heat effect only model $-1894.4$, and the AIC for the 
+two effect model, which we will present throughout this section, was $-1898.9$. 
+Thus, the two
 effect model is superior and will be presented throughout the rest of the paper.
@@
 run the code to see the p value or should I say it directly and how do I say it?}
+\eds{If you are doing a likelihood ratio test, I suggest including the p-value 
+and the degrees of freedom associated with the test.}
 
@@
   \caption{Summary statistics for the two effect GLMM on various data sets. SE
-  is short for standard error.} 
+  is short for standard error. \eds{data in table needs to be revisited and table
+	needs to be referenced within the text. Also, it needs to be clarified 
+	somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}} 
   \begin{tabular}{c c c c c c}
@@
 
-Two points of this section are to determine how extreme a reaction time below
-0.1 seconds is, and to determine how the exclusion of 2022 could alter results.
+Two aims of this section are to determine how extreme a reaction time below
+0.1 seconds is, and to determine how the exclusion of 2022 could alter the results.
 Using the gamma function with link set to log, the probability of the reaction 
 time being less than 0.1 seconds can be calculated in R using the methods 
-described in section~\ref{sec:glmm}. A simulation size of $10,000,000$ was 
-chosen to try and estimate the probability of observing a reaction time below 
-two thresholds: 0.1 and 0.08. When looked at the All Pooled datset, a probability
+described in Section~\ref{sec:glmm}. A simulation size of $10,000,000$ was 
+chosen to estimate the probability of observing a reaction time below 
+two thresholds: 0.1 and 0.08. When examining the All Pooled dataset, a probability
 of $5.15\cdot10^{-3}$ corresponds to roughly 1 in every 195 starts
@@
 results in a reaction time barrier of 0.079 for the men's pooled dataset.  
-Tables~\ref{tab:Sim_time} shows that the inclusion of 2022 results in a lower
+Table~\ref{tab:Sim_time} shows that the inclusion of 2022 results in a lower
 reaction time barrier, further showing the claims above that the reaction times
@@
 
-The rank based methods described in are used in section~\ref{sec:rank_methods}
+The rank based methods described in Section~\ref{sec:rank}
 to determine if the reaction times for athletes who had competed at other 
@@
 \texttt{B = 1000} indicating that the random permutation size was chosen to be 
-1000.  As the sample sizes are small: 65 and 49, we use \texttt{exact = TRUE} as
-that is preferred.  In both cases, the national versus international comparison
-produced stronger p values than the 2019 versus 2022 comparison, with the
+1000, as discussed previously in Section~\ref{sec:rank}.  
+%As the sample sizes are small, 65 and 49, we use \texttt{exact = TRUE} as
+%that is preferred.  
+In both cases, the national versus international comparison
+produced stronger \eds{lower?} p values than the 2019 versus 2022 comparison, with the
 national vs international \texttt{exact = TRUE} test producing a p value of 
-$2.2\cdot10^{-16}$, which is the smallest possible value acheivable in R.  This
+$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  This
 is a highly significant test result and shows that athletes who competed at the
@@
 same athlete comparison with \texttt{exact = TRUE} argument, the p value was
-significant at $\alpha = 0.01$, however for when \texttt{exact = FALSE}, the p
-value was higher and only significant at $\alpha = 0.05$. Nonetheless, these
+significant at $\alpha = 0.01$, however when \texttt{exact = FALSE}, the p
+value was higher and only significant at $\alpha = 0.05$. 
+\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
+Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
+Nonetheless, these
 tests help to show that athletes competing at the 2022 Track and Field World
@@
 slightly faster relative to other championships going back to 2003 
-\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
-the intro, unless he only considered sprint and not hurdles?} \of{Johnson's
-results considered the mens 100m, mens 110m hurdles, womens 100m but not the 
-womens 100m hurdles.} Likewise, the rank-based comparisons resulted in p values
-that were not significant.  The lack of evidence from the women's data makes it 
+%\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
+%the intro, unless he only considered sprint and not hurdles?} \of{Johnson's
+%results considered the mens 100m, mens 110m hurdles, womens 100m but not the 
+%womens 100m hurdles.} 
+Likewise, the rank-based comparisons resulted in p values
+that were not significant. 
+\eds{could that be because of even smaller sample sizes?}
+  The lack of evidence from the women's data makes it 
 harder to conclude that the sensors used at the 2022 championships were faulty, 
@@
 100 meter hurdle were the same.  
+\eds{is any calibration of the sensors needed (e.g. for weight?) that might
+also play a role for women vs men. (My husband asked me about this as a possible 
+reason for the difference in results for men and women, but he did not know
+if any calibration is actually done.)}
 

### 2023-08-29 Code/2023Analysis.Rmd
Exploratory analysis of data including 2023

---
title: '2023'
author: "Owen Fiore"
date: "2023-08-29"
output: pdf_document
---

```{r}
new <- read.csv("../Data/Thesis2023.csv")
colnames(new) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")

men_pooled <- subset(new, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != 'DNF'))
```

```{r}
plot(aggregate(men_pooled$RxnTime, list(men_pooled$Year), FUN = mean), xlab = "Year", ylab = "Reaction time in seconds", main = "Mean Reaction Time By Year")
```
```{r}
library(ggplot2)
ggplot(men_pooled, aes(x = Year, y = RxnTime, group = Year)) +
  geom_boxplot(fill = "orange1", color = "orangered4") +
  labs(x = "Year", y = "Reaction Time (s)") +
       scale_y_continuous(limits = c(0.1, 0.225),
                            breaks = seq(0.1, 0.225, 0.025)) +
  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
                                              breaks = seq(1997, 2024, 2))
```

### 2023-08-30 Code/2023Analysis.Rmd
Added GLMM results for 2023 and added clusrank code

@@
 
-men_pooled <- subset(new, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != 'DNF'))
+men_new_pooled <- subset(new, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != 'DNF'))
 ```
@@
 ```{r}
-plot(aggregate(men_pooled$RxnTime, list(men_pooled$Year), FUN = mean), xlab = "Year", ylab = "Reaction time in seconds", main = "Mean Reaction Time By Year")
-```
-```{r}
 library(ggplot2)
-ggplot(men_pooled, aes(x = Year, y = RxnTime, group = Year)) +
+ggplot(men_new_pooled, aes(x = Year, y = RxnTime, group = Year)) +
   geom_boxplot(fill = "orange1", color = "orangered4") +
@@
 ```
+```{r}
+library(lme4)
+men_new_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_new_pooled, 
+                                family = Gamma(link = "log"))
+
+men_new_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_new_pooled, 
+                                       family = Gamma(link = "log"))
+men_new_pooled_batch_gamma_log <- glmer(RxnTime ~ (1|Batch), data = men_new_pooled, 
+                                       family = Gamma(link = "log"))
+men_new_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+                                   data = men_new_pooled, family = Gamma(link = "log"))
+```
+```{r}
+set.seed(1)
+library(lme4)
+sim_two_effect_time <- function(model, base, RxnTime) {
+  simfit <- function(model, n = 10000000) {
+    vc <- as.data.frame(VarCorr(model))
+    sd_year <- vc$sdcor[1]
+    sd_year_batch <- vc$sdcor[2]
+
+    z1 <- rnorm(n, sd = sd_year)
+    z2 <- rnorm(n, sd = sd_year_batch)
+
+    shape <- 1 / sigma(model)^2
+    mu <- exp(fixef(model) + z1 + z2)
+
+    x <- rgamma(n, shape = shape, scale = mu / shape)
+    x
+  }
+
+  sbase <- summary(base)
+  x <- simfit(model)
+  print(mean(x < RxnTime))
+  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
+  ## compare with and without random effect
+  plot(density(x))  # with random effects
+  curve(dgamma(x, shape = 1 / sbase$dispersion, 
+               scale = exp(coef(base)) * sbase$dispersion), 
+        0.08, 0.25, add = TRUE, col = "blue")  # fixed effect only
+}
+```
+
+```{r}
+sim_two_effect_time(men_new_pooled_two_effect_gamma_log, men_new_pooled_one_gamma_log, 0.1)
+```
+```{r}
+comp <- read.csv("../Data/Clusrank2022vs2023.csv")
+colnames(comp) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp,
+                method = "ds",
+                exact = TRUE, B = 10000)
+```

### 2023-09-06 Manuscript/manuscript.tex
One comment
Started to change language and add 2023 data

@@
 \eds{He is not competing this year, correct? If yes, we may need to rephrase.}
+\of{Correct}
 
@@
 There are two types of data that are explored throughout this paper: data from
-1999-2022 for every world championship, and data from athletes who competed in
+1999-2023 for every world championship, and data from athletes who competed in
 the 2022 World Championships and another competition (either 2019 World 
-Championships or a 2022 national competition typically held within three months 
-of the 2022 World Championships). 
+Championships, 2023 World Championships or a 2022 national competition typically
+held within three months of the 2022 World Championships). 
 %\eds{is this true? are you still using the 2019 world championships data for the
@@
 Section~\ref{sec:glmm}, while the second collection of data is used for the 
-rank-based comparisons described in Section~\ref{sec:rank}.
+rank-based comparisons described in Section~\ref{sec:rank}.  This paper was
+started before the 2023 World Championships, but as that data has become
+available recently, we are pleased to include it in our analysis.
 
@@
 as in 2022, when there were only five data points due to two disqualifications 
-and one
-athlete who did not compete.  We also consider data that
-excludes 2022 to see how our analysis differs based on the exclusion of one
+and one athlete who did not compete.  This is the data that we will reference
+throughout the rest of the paper unless otherwise noted. We also consider data 
+that excludes 2022 to see how our analysis differs based on the exclusion of one
 particular year of interest.
 
-The data under consideration is summarized using a boxplots in 
-Figure~\ref{fig:Boxplot}.  It is clear that the reaction times in the 
-2022 boxplot for both 
-the men's finals and men's pooled semifinal and final data are lower (indicating 
-faster reaction times)
-than in previous years.  The median for the men's pooled data in 2022 was 0.129
-and the median for the men's final data in 2022 was 0.126.  For comparison,
-\cite{brosnan2017effects} found median reaction times of 0.156 seconds 
-for men when looking at data from 1999 to 2014.
-We can also see from Figure~\ref{fig:Boxplot} that the reaction times clearly
-differ from year to year. The venue of the World Championships changes each 
-year, so weather or climate related factors such as humidity, precipitation, 
-elevation, may indeed play a role, but additionally, both the technology and 
-false start penalization rules have changed over the study period.
+The data under consideration is summarized using a box plot in 
+Figure~\ref{fig:Boxplot}.  It is clear that the reaction times in the 2022
+box plot are lower (indicating faster reaction times) than in many of the other
+years.  The median reaction time in 2022 was 0.129 seconds,  compared to 0.156
+seconds, which is what \citet{brosnan2017effects} found when looking at data
+from 1999 to 2014. We can also see from Figure~\ref{fig:Boxplot} that the 
+reaction times clearly differ from year to year. The venue of the World 
+Championships changes each year, so weather or climate related factors such as 
+humidity, precipitation, elevation, may indeed play a role, but additionally,
+both the technology and false start penalization rules have changed over the 
+study period.
 
@@
   \includegraphics{Boxplot}
-  \caption{The reaction times from 1999 to 2022 for the men's final (top) and
-  the men's semifinals and finals (bottom). 
+  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.
 	\eds{there is only one plot here, not two. If you are keeping only one, the
 	text and caption need to be modified accordingly}}
+  \of{Resolved}
   \label{fig:Boxplot}
@@
 recorded in the national language without translation, and hence not used). 
-\of{I don't know how to say this}. \eds{is my change correct?} Each athlete was 
+\of{I don't know how to say this}. \eds{is my change correct?} \of{No I was able
+to use a lot even though I couldn't actually read the page. I just tried to find
+110 meter hurdle and then look at a column for reaction times which are always from
+roughly 0.1 to 0.18 seconds.} Each athlete was 
 designated as a ``cluster'' and the stage of competition (national or international) 
@@
 
-\subsubsection{2019 World Track and Field Championships}\label{sec:data2019}
-The 2019 World Track and Field Championship data is included in the larger dataset
+\subsubsection{Other World Track and Field Championships}\label{sec:data2019}
+The 2019 and 2023 World Track and Field Championships data is included in the larger dataset
 described in Section~\ref{sec:dataworld}, but we can also use it similarly to
 how we used national level in Section~\ref{sec:datanational}.  We can compare
-the times of athletes who participated in the 2019 and 2022 World Track and Field
-Championships to see if there were lower times in 2022.  While it is theoretically
-possible that athletes got faster between 2019 and 2022, it is important to
+the times of athletes who participated in the 2022 Championships with their times
+at another World Championships: 2019 and/or 2023.
+
+While it is theoretically possible that athletes got faster 
+between 2019 and 2022 or 2022 and 2023, it is important to 
 remember that these are some of the fastest sprinters on the planet, and thus
@@
 as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
-within each cluster label times to be from either the 2019 World Championships
-or the 2022 World Championships.  In this analysis, we look at data from 15
-athletes, where cluster size ranges from two to six.   There were two clusters 
-of size two, three clusters of size three, five clusters of size four, two 
-clusters of size five, and three clusters of size six.
+within each cluster label times to be from either the 2019, 2022, or 2023 World
+Championships. When comparing 2019 and 2022 we look at data from 15
+athletes, with two clusters of size two, three clusters of size three, five 
+clusters of size four, two clusters of size five, and three clusters of size six.
+When comparing 2022 and 2023 we look at data from 22 athletes, with seven 
+clusters of size two, three clusters of size three, six clusters of size four,
+and six clusters of size five.
 
 \jy{give specifics of your sample size}
+\of{Resolved}
 
@@
   \centering
-  \includegraphics{Scatterplot2019vs2022}
-  \caption{The reaction times from athletes who competed at least once in both 
-  the 2019 and 2022 World Track and Field Championships}
-  \label{fig:Scatterplot2019vs2022}
+  \includegraphics{RankScatterPlots}
+  \caption{The top graph shows the athletes who competed at least once each at
+  the 2019 and 2022 World Championships.  The middle graph shows the athletes
+  who competed at least once each at the 2022 and 2023 World Championships. The
+  bottom graph shows the athletes who competed at least once at their country's
+  2022 national championships and the 2022 World Championship.}
+  \label{fig:RankScatterplots}
 \end{figure}
-\of{I included this here for now, but feel free to change the location.  I will
-add references to it and explanation of it along with some descriptive statistics about
-the size and mean/median,std dev, size of the data.}
 
-Figure~\ref{fig:Scatterplot2019vs2022} shows the distribution of times from the
-2019 and 2022 World Championships.  
+Figure~\ref{fig:RankScatterplots} shows the data that is being analyzed
+with the rank-based methods.  In each of the three graphs, athletes who competed
+in the 2022 finals are having their times compared to another event they competed
+at.  It is worth noting that each instance of the 2022 data is different, because
+those who competed at both the 2019 and 2022 World Championships will be
+different athletes than those who competed at both the 2022 and 2023 World
+Championships. This data was only for athletes who competed
+in at least one race at each championship, however it is possible for athletes
+to have competed in one race in 2019 and two races in 2022.  The 2022 World
+Championship Heats and Semifinals both had on average lower reaction times than
+their respective 2019 races, and the 2022 finals were substantially faster. It
+is also worth noting is that the fastest reaction time in the 2022 Finals and 2022
+Semifinals both came from Devon Allen, but the difference between 0.101 and 0.099
+seconds was the difference in what caused him to be disqualified. 
 \eds{Please reorganize the labels so they are either in one of the following 
@@
 (I am not sure which is better, but perhaps the latter?)}
-This data was only for athletes who competed
-in at least one race at each championship, however it is possible for athletes
-to have competed in one race in 2019 and two races in 2022.  The 2022 World
-Championship Heats and Semifinals both had on average lower reaction times than
-their respective 2019 races, and the 2022 finals were substantially faster. Also
-worth noting is that the fastest reaction time in the 2022 Finals and 2022
-Semifinals both came from Devon Allen, but the difference between 0.101 and 0.099
-seconds was the difference in what caused him to be disqualified. 
+of{I re-worked this graph completely, please let me know what you think.  For
+each graph I currently list them in chronological order: so 2022 then 2023 but
+I list 2022 second when comparing it to 2019/rank comparison.  Please let me
+know your thoughts on this and if I should list 2022 and then 2023 for visual
+consistency.}
+
 %\eds{check these numbers? 0.099?}\of{resolved}
@@
 AIC.  For the entire pooled dataset, the AIC for the venue effect only model was
-$-1847.9$, the AIC for the heat effect only model $-1894.4$, and the AIC for the 
-two effect model, which we will present throughout this section, was $-1898.9$. 
+$-2023.5$, the AIC for the heat effect only model $-2070.7$, and the AIC for the 
+two effect model, which we will present throughout this section, was $-2075.7$. 
 Thus, the two
@@
       \midrule
-      Old Pooled & 0.060 & 0.034 & $-1.865$ & 0.139 \\
-      All Pooled & 0.059 & 0.041 & $-1.880$ & 0.014 \\
+      Excluding 2022 &  0.057 & 0.033 & $-1.868$ & 0.136 \\
+      All Data & $-2075.7$ 0.057 & 0.040 & $-1.880$ & 0.013 \\
       \bottomrule

### 2023-09-06 Manuscript/manuscript.tex
Fixed typos and errors

@@
 what is going on this year?}
+\of{Resolved}
 
@@
 %subsections within rank-based comparison.}
-The first data set is used for the GLMM analyses described in 
+The first data set is used for the GLMM analysis described in 
 Section~\ref{sec:glmm}, while the second collection of data is used for the 
@@
 
-\subsection{World Championships 1999--2022}\label{sec:dataworld}
+\subsection{World Championships 1999--2023}\label{sec:dataworld}
 
 Data was taken from the World Athletics 
-and covers the men's 110 meter hurdles from 1999 to 2022; the source can be 
+and covers the men's 110 meter hurdles from 1999 to 2023; the source can be 
 found in the Appendix.
@@
 
-From 2007--2009, the IAAF (the former name for World Athletics, and the 
+From 2007 to 2009, the IAAF (the former name for World Athletics, and the 
 governing body for the World Track and Field Championships) instituted a rule 
@@
 \subsubsection{Other World Track and Field Championships}\label{sec:data2019}
-The 2019 and 2023 World Track and Field Championships data is included in the larger dataset
-described in Section~\ref{sec:dataworld}, but we can also use it similarly to
-how we used national level in Section~\ref{sec:datanational}.  We can compare
-the times of athletes who participated in the 2022 Championships with their times
-at another World Championships: 2019 and/or 2023.
+The 2019 and 2023 World Track and Field Championships data is included in the 
+larger dataset described in Section~\ref{sec:dataworld}, but we can also use it
+similarly to how we used national level in Section~\ref{sec:datanational}.  We
+can compare the times of athletes who participated in the 2022 Championships with
+their times at another World Championships: 2019 and/or 2023.
 
@@
 data. Two random effects were found useful: one captures the venue effect in
-each year and the other captures the heat effect within each year.
+each year and the other captures the heat effect within each year.  The heat
+effect is important as it captures the variability in the amount of time
+athletes are on the starting blocks before the sun goes off.  Thus, we can try
+to reduce this variability by accounting for this in the model through the batch
+effect.
 Let $Y_{ijk}$ be the reaction time of observation~$k$ in heat~$j$ in year~$i$.
@@
 but this effect would be shared across all athletes in the same heat.}
+\of{Yes I agree with this and updated the language in the text.}
 In the classic generalized linear model (GLM) sense,
@@
 reliable, so we use 1000 random permutations to simulate the observed test 
-statistic: $W^*$ given number of times to obtain the p-value of said test 
-statistics. 
+statistic.
 \eds{I do not understand what is meant in the text after the colon in the 
 previous sentence.  Please rephrase.}
+\of{I'm not sure what that was, I removed it now.}
 This method is available from the \texttt{clusWilcox.test()} function
@@
 for the pooled datasets and shows the standard error of the year
-effect increases for both the men's final and men's pooled data sets when
-including 2022. We choose to only look at the pooled datasets, as R displayed
-warnings when creating the finals only models.  The two effect model with venue
+effect increases when
+including 2022.  The two effect model with venue
 and heat effect was the best model, as indicated by an ANOVA test 
@@
 an ANOVA test, assuming that is the formal test being used} and decrease in
-AIC.  For the entire pooled dataset, the AIC for the venue effect only model was
+AIC.  For the entire data, the AIC for the venue effect only model was
 $-2023.5$, the AIC for the heat effect only model $-2070.7$, and the AIC for the 
@@
       \midrule
-      Excluding 2022 &  0.057 & 0.033 & $-1.868$ & 0.136 \\
-      All Data & $-2075.7$ 0.057 & 0.040 & $-1.880$ & 0.013 \\
+      Excluding 2022 & $-1912.9$ &  0.058 & 0.033 & $-1.868$ & 0.136 \\
+      All Data & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134 \\
       \bottomrule
@@
 chosen to estimate the probability of observing a reaction time below 
-two thresholds: 0.1 and 0.08. When examining the All Pooled dataset, a probability
-of $5.15\cdot10^{-3}$ corresponds to roughly 1 in every 195 starts
+two thresholds: 0.1 and 0.08. When including 2022, a probability
+of $4.38\cdot10^{-3}$ corresponds to roughly 1 in every 228 starts
 being below 0.1 seconds.  When we exclude 2022, the probability of observing a
-reaction time below 0.1 seconds decreased to $4.15\cdot10^{-3}$, or 1 in every
-240 starts.
-\of{I feel like with the two effect model the probability change is not as
-different as it was previously}
+reaction time below 0.1 seconds decreased to $3.52\cdot10^{-3}$, or 1 in every
+285 starts.
 
@@
   \centering
-  \caption{Probabilities of observing reaction times less than 0.1 and 0.08
-  for various data sets.}
+  \caption{Probabilities of observing reaction times less than 0.08, 0.09, and
+  0.01 seconds based on the two effect model.}
   \begin{tabular}{c c c c} 
@@
    \midrule
-   Old Pooled & $4.67\cdot10^{-5}$ & $5.91\cdot10^{-4}$ &  $4.15\cdot10^{-3}$  \\ 
-   All Pooled & $6.01\cdot10^{-5}$ & $7.39\cdot10^{-4}$ & $5.15\cdot10^{-3}$ \\
+   Excluding 2022 & $3.19\cdot10^{-5}$ & $4.48\cdot10^{-4}$ &  $3.52\cdot10^{-3}$  \\ 
+   Including 2022 & $4.01\cdot10^{-5}$ & $5.82\cdot10^{-4}$ & $4.38\cdot10^{-3}$ \\
    \bottomrule
@@
 reaction time barrier given the probability of observing a time less than the 
-barrier.  For example, if we want a 0.1\% chance of observing an illegally fast 
-reaction time for the men's pooled dataset that includes 2022, then the reaction
-time barrier should be 0.089 seconds, as shown in Table~\ref{tab:Sim_time}.  We
-can repeat this calculation for each dataset and for various probability levels,
+barrier.  When considering all of the data, if we want a 0.1\% chance of 
+observing an illegally fast reaction time, the reaction
+time barrier should be 0.092 seconds, as shown in Table~\ref{tab:Sim_time}.  We
+can repeat this calculation for various probability levels,
 for example a probability of 0.0001 (one disqualification for every 10,000 starts)
-results in a reaction time barrier of 0.079 for the men's pooled dataset.  
-Table~\ref{tab:Sim_time} shows that the inclusion of 2022 results in a lower
-reaction time barrier, further showing the claims above that the reaction times
-in 2022 were low.
+results in a reaction time barrier of 0.083 seconds.  As the exclusion of 2022
+from the data resulted in a lower probability of observing a low reaction time,
+the reaction time barrier will be less lenient when calculating the barrier
+based on probability.
 
@@
    \midrule
-   Old Pooled & $0.093$ & $0.083$ \\ 
-   All Pooled & $0.091$ & $0.082$ \\
+   Old Pooled & $0.093$ & $0.084$ \\ 
+   All Pooled & $0.092$ & $0.083$ \\
    \bottomrule
@@
 is a highly significant test result and shows that athletes who competed at the
-2022 Track and Field World Championships were on average, significantly faster
+2022 World Track and Field Championships were on average, significantly faster
 at reacting than they were a few months prior.  For the 2019 versus 2022
@@
 Nonetheless, these
-tests help to show that athletes competing at the 2022 Track and Field World
+tests help to show that athletes competing at the 2022 World Track and Field
 Championships had faster reaction times than when they had competed previously.
@@
    \midrule
-   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10{-2}$ \\ 
+   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10^{-2}$ \\
+   2022 vs 2023 & $3.0\cdot10^{-3}$ & $1.0\cdot10^{-2}$ \\
    Nat vs Inter & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ \\
@@
   \caption{2019 vs 2022 compares data from the same athletes who competed at the
-  2019 and 2022 World Track and Field Championships. Nat vs Inter compares data
-  from the same athletes who competed at the 2022 national-level championships
-  and the 2022 World Track and Field Championships.}
+  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
+  from the same athletes who competed at the 2022 and 2023 World Track and Field
+  Championships. Nat vs Inter compares data from the same athletes who competed
+  at the 2022 national-level championships and the 2022 World Track and Field
+  Championships.}
   \label{tab:Clusrankresults}
@@
 if any calibration is actually done.)}
+\of{I honestly am not sure, do you think that this is worth investigating, or
+beyond the scope of this paper? I am not sure if the mechanics of Seiko's
+technology is public knowledge.}
 

### 2023-09-06 Code/GLMM.Rmd
Now includes 2023 data

--- 2023-07-26 Code/GLMM.Rmd
@@
 ---
-title: "Random_Effects"
+title: "GLMM"
 author: "Owen Fiore"
@@
 #1
-times <- read.csv("../Data/ThesisData.csv")
+times <- read.csv("../Data/Thesis2023.csv")
 colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
@@
 men_old_finals <- subset(times, Type == "F" & Gender == "M" & 
-                           Time != 'DQ' & Year < 2022)
+                           Time != 'DQ' & Year != 2022)
 men_old_pooled <- subset(times, (Type == "F" | Type == "S") & 
-                           Gender == "M" & Time != 'DQ' & Year < 2022)
+                           Gender == "M" & Time != 'DQ' & Year != 2022)
 ```
@@
 
+```{r}
+summary(men_pooled_year_gamma_log)
+summary(men_old_pooled_year_gamma_log)
+```
+
 This creates the batch effect only models.
@@
 
+```{r}
+summary(men_pooled_batch_gamma_log)
+```
+
 This creates the models with a year effect and batch effect.
@@
                                data = men_pooled, family = Gamma(link = "log"))
+```
+```{r}
+summary(men_pooled_two_effect_gamma_log)
+summary(men_old_pooled_two_effect_gamma_log)
 ```
@@
 #7
-sim_two_effect_time(men_old_finals_two_effect_gamma_log, men_old_finals_one_gamma_log, 0.1)
-sim_two_effect_time(men_finals_two_effect_gamma_log, men_finals_one_gamma_log, 0.1)
+set.seed(1)
 sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.1)
@@
 ```
+
 ```{r}
-sim_two_effect_time(men_old_finals_two_effect_gamma_log, men_old_finals_one_gamma_log, 0.08)
-sim_two_effect_time(men_finals_two_effect_gamma_log, men_finals_one_gamma_log, 0.08)
+set.seed(1)
+sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.09)
+sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.09)
+```
+
+```{r}
+set.seed(1)
 sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.08)
@@
 #9
-sim_two_effect_prob(men_old_finals_two_effect_gamma_log, men_old_finals_one_gamma_log, 0.001)
-sim_two_effect_prob(men_finals_two_effect_gamma_log, men_finals_one_gamma_log, 0.001)
+set.seed(1)
 sim_two_effect_prob(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.001)
@@
 #10
-sim_two_effect_prob(men_old_finals_two_effect_gamma_log, men_old_finals_one_gamma_log, 0.0001)
-sim_two_effect_prob(men_finals_two_effect_gamma_log, men_finals_one_gamma_log, 0.0001)
+set.seed(1)
 sim_two_effect_prob(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.0001)

### 2023-09-06 Code/RankComparison.Rmd
Included 2022 vs 2023 analysis

@@
 and 2 if it was the world championships. The tests cluster by athlete, of which
-for each dataset an integer from 1-15 is used (the values are not significant).
+for each dataset an integer from 1-11 is used (the values are not significant).
 For more information about the "ds" method, please look at the Rank-based
@@
 
+The code below reads in data comparing the 2022 and 2023 World Championships
+```{r}
+new_clusrank <- read.csv("../Data/Clusrank2022vs2023.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(new_clusrank) <- c("Time", "Athlete", "Year")
+#View(comp)
+```
+
+The tests below are very similar to the ones above, the only difference being
+grouping by Year instead of Stage (2 if 2022 and 3 if 2023).
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = new_clusrank,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = new_clusrank,
+                method = "ds",
+                exact = TRUE, B = 10000)
+```
+
 This is seeing what would happen if we remove prelims, the sample sizes in the

### 2023-09-06 Code/Plots.Rmd
removed addressed comments
Added the triple scatterplot graph

--- 2023-05-31 Code/Plots.Rmd
@@
 times <- read.csv("../Data/ThesisData.csv")
-View(times)
+## View(times)
 colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
@@
 ```
+
+```{r}
+library(ggplot2)
+means_df <- data.frame(variable = names(means), mean = means)
+
+# Define the grouping of bars
+groupings <- list(c("2019H", "2022H"), c("2019S", "2022S"), c("2019F", "2022F"))
+
+# Convert variable to a factor with desired order
+means_df$variable <- factor(means_df$variable, levels = unlist(groupings))
+
+# Create the bar graph
+ggplot(means_df, aes(x = variable, y = mean, color = variable)) +
+  geom_bar(size = 3) +
+  labs(y = "Mean Reaction Time", x = "") +
+  scale_color_manual(values = c("sienna2", "tan", "sienna2", "tan", "sienna2", "tan")) +
+  theme_minimal() +
+  theme(legend.position = "none")
+#ggsave(filename = "2019vs2022ScatterPlot.pdf", path = "../Manuscript", height = 4, width = 6)
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+nineteen <- read.csv("../Data/2019vs2022.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(nineteen) <- c("Athlete", "Country", "2019 Heats", "2019 Semifinals", "2019 Finals", "2022 Heats", "2022 Semifinals", "2022 Finals")
+nineteen <- nineteen[, -c(1,2)]
+df_long <- tidyr::pivot_longer(nineteen, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("2019 Heats", "2019 Semifinals", "2019 Finals", "2022 Heats", "2022 Semifinals", "2022 Finals"))
+rank1 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Time") +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+twentythree <- read.csv("../Data/2022vs2023.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(twentythree) <- c("Athlete", "Country", "2022 Heats", "2022 Semifinals", "2022 Finals", "2023 Heats", "2023 Semifinals", "2023 Finals")
+twentythree <- twentythree[, -c(1,2)]
+df_long <- tidyr::pivot_longer(twentythree, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("2022 Heats", "2022 Semifinals", "2022 Finals", "2023 Heats", "2023 Semifinals", "2023 Finals"))
+rank2 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Time") +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+national <- read.csv("../Data/NatVSWorld.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(national) <- c("Athlete", "Country", "National Heats", "National Semifinals", "National Finals", "2023 Heats", "2023 Semifinals", "2023 Finals", "Data Source")
+twentythree <- twentythree[, -c(1,2, 9)]
+df_long <- tidyr::pivot_longer(twentythree, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("2022 Heats", "2022 Semifinals", "2022 Finals", "2023 Heats", "2023 Semifinals", "2023 Finals"))
+rank2 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Time") +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```

### 2023-09-06 Code/Plots.Rmd
Remove unneccessary code

--- 2023-05-31 Code/Plots.Rmd
@@
 ```{r}
-opar <- par()
-par(mfrow = c(1,2))
-pdf(file = "BoxplotFinals.pdf")
-boxplot(RxnTime ~ Year, data = men_finals, xlab = "Year", ylab = "Reaction Time in seconds")
-boxplot(RxnTime ~ Year, data = men_finals, xlab = "Year", ylab = "Reaction Time in seconds")
-dev.off()
-par(opar)
+library(ggplot2)
+ggplot(men_pooled, aes(x = Year, y = RxnTime, group = Year)) +
+  geom_boxplot(fill = "orange1", color = "orangered4") +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.1, 0.225),
+                            breaks = seq(0.1, 0.225, 0.025)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
+                                              breaks = seq(1999, 2024, 2))
+ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
 ```
@@
 ```{r}
-#ggplot(data = win, aes(x = strength_diff)) + geom_boxplot() + theme_minimal()
-ggplot(data = men_finals, aes(x=Year, y = RxnTime)) + geom_boxplot() + theme_minimal()
+library(tidyr)
+library(ggplot2)
+nineteen <- read.csv("../Data/2019vs2022.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(nineteen) <- c("Athlete", "Country", "2019 Heats", "2019 Semifinals", "2019 Finals", "2022 Heats", "2022 Semifinals", "2022 Finals")
+nineteen <- nineteen[, -c(1,2)]
+df_long <- tidyr::pivot_longer(nineteen, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("2019 Heats", "2019 Semifinals", "2019 Finals", "2022 Heats", "2022 Semifinals", "2022 Finals"))
+rank1 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Reaction Time", x = "", title = "2019 vs 2022") +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+twentythree <- read.csv("../Data/2022vs2023.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(twentythree) <- c("Athlete", "Country", "2022 Heats", "2022 Semifinals", "2022 Finals", "2023 Heats", "2023 Semifinals", "2023 Finals")
+twentythree <- twentythree[, -c(1,2)]
+df_long <- tidyr::pivot_longer(twentythree, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("2022 Heats", "2022 Semifinals", "2022 Finals", "2023 Heats", "2023 Semifinals", "2023 Finals"))
+rank2 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Reaction Time", x = "", title = "2022 vs 2023") +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+national <- read.csv("../Data/NatVSWorld.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(national) <- c("Athlete", "Country", "National Heats", "National Semifinals", "National Finals", "2023 Heats", "2023 Semifinals", "2023 Finals", "Data Source")
+national <- national[, -c(1,2, 9)]
+df_long <- tidyr::pivot_longer(national, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("National Heats", "National Semifinals", "National Finals", "2022 Heats", "2022 Semifinals", "2022 Finals"))
+rank3 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Reaction Time", x = "", title = "2022 National vs International") +
+  ylim(0.08, 0.22) +
+  theme_minimal()
 ```
@@
 library(gridExtra)
-library(ggplot2)
-final <- ggplot(men_finals, aes(x = Year, y = RxnTime, group = Year)) +
-  geom_boxplot(fill = "lightblue", color = "darkblue") +
-  labs(x = "Year", y = "Reaction Time (s)") +
-       scale_y_continuous(limits = c(0.1, 0.225),
-                            breaks = seq(0.1, 0.225, 0.025)) +
-  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2023),
-                                              breaks = seq(1997, 2022, 2))
-
-pooled <- ggplot(men_pooled, aes(x = Year, y = RxnTime, group = Year)) +
-  geom_boxplot(fill = "orange1", color = "orangered4") +
-  labs(x = "Year", y = "Reaction Time (s)") +
-       scale_y_continuous(limits = c(0.1, 0.225),
-                            breaks = seq(0.1, 0.225, 0.025)) +
-  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2023),
-                                              breaks = seq(1997, 2022, 2))
-#both <- grid.arrange(final,pooled, ncol=1)
-grid.arrange(final,pooled, ncol=1)
-#ggsave(filename="Finals_Pooled_Boxplot.pdf", path = "../Manuscript", plot=both, height = 6, width = 6)
+rank_scatter <- grid.arrange(rank1, rank2, rank3, ncol = 1)
+ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 9, width = 7)
 ```

### 2023-09-06 Manuscript/manuscript.tex
Small changes

@@
 two thresholds: 0.1 and 0.08. When including 2022, a probability
-of $4.38\cdot10^{-3}$ corresponds to roughly 1 in every 228 starts
+of $4.34\cdot10^{-3}$ corresponds to roughly 1 in every 230 starts
 being below 0.1 seconds.  When we exclude 2022, the probability of observing a
-reaction time below 0.1 seconds decreased to $3.52\cdot10^{-3}$, or 1 in every
-285 starts.
+reaction time below 0.1 seconds decreased to $3.54\cdot10^{-3}$, or 1 in every
+283 starts.
 
@@
    \midrule
-   Excluding 2022 & $3.19\cdot10^{-5}$ & $4.48\cdot10^{-4}$ &  $3.52\cdot10^{-3}$  \\ 
-   Including 2022 & $4.01\cdot10^{-5}$ & $5.82\cdot10^{-4}$ & $4.38\cdot10^{-3}$ \\
+   Excluding 2022 & $3.35\cdot10^{-5}$ & $4.45\cdot10^{-4}$ &  $3.54\cdot10^{-3}$  \\ 
+   Including 2022 & $4.20\cdot10^{-5}$ & $5.70\cdot10^{-4}$ & $4.34\cdot10^{-3}$ \\
    \bottomrule
@@
    \midrule
-   Old Pooled & $0.093$ & $0.084$ \\ 
+   Old Pooled & $0.094$ & $0.084$ \\ 
    All Pooled & $0.092$ & $0.083$ \\

### 2023-09-06 Manuscript/manuscript.tex
Updated the inclusion of 2023 data explanation

@@
 rank-based comparisons described in Section~\ref{sec:rank}.  This paper was
-started before the 2023 World Championships, but as that data has become
-available recently, we are pleased to include it in our analysis.
+started shortly after the 2022 World Championships, but as the 2023 data has 
+become available recently, we are pleased to include it in our analysis, and it
+did not significantly change our results.
 

### 2023-09-06 Manuscript/manuscript.tex
Worked on table 3

@@
 effect is important as it captures the variability in the amount of time
-athletes are on the starting blocks before the sun goes off.  Thus, we can try
+athletes are on the starting blocks before the gun goes off.  Thus, we can try
 to reduce this variability by accounting for this in the model through the batch
@@
   \centering
-  \caption{Summary statistics for the two effect GLMM on various data sets. SE
+  \caption{Estimated parameters for the two effect GLMM on various data sets. SE
   is short for standard error. \eds{data in table needs to be revisited and table
@@
       \midrule
-      Excluding 2022 & $-1912.9$ &  0.058 & 0.033 & $-1.868$ & 0.136 \\
-      All Data & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134 \\
+      Year Effect Only & $-2023.5$ & N/A & 0.048 & $-1.857$ & 0.152
+      Heat Effect Only & $-2070.7$ & 0.073 & N/A & $-1.890$ & 0.135
+      Full model & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134 \\
       \bottomrule

### 2023-09-06 Code/GLMM.Rmd
Started to remove finals variables and renamed Batch to Heat

@@
 
-The code below loads in the 4 datasets from the csv file.
+The code below loads in the dataset from the csv file.
 ```{r}
@@
 times <- read.csv("../Data/Thesis2023.csv")
-colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
-men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ')
+colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Heat")
 men_pooled <- subset(times, (Type == "F" | Type == "S") & 
                        Gender == "M" & Time != 'DQ')
-men_old_finals <- subset(times, Type == "F" & Gender == "M" & 
-                           Time != 'DQ' & Year != 2022)
 men_old_pooled <- subset(times, (Type == "F" | Type == "S") & 
@@
 library(lme4)
-men_old_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_finals, 
-                                    family = Gamma(link = "log"))
-men_finals_one_gamma_log <- glm(RxnTime ~ 1, data = men_finals, 
-                                family = Gamma(link = "log"))
 men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, 
@@
 library(lme4)
-
-men_old_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_finals, 
-                                       family = Gamma(link = "log"))
-men_finals_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_finals, 
-                                   family = Gamma(link = "log"))
 men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, 
@@
                                    family = Gamma(link = "log"))
-
 ```
@@
 
-This creates the batch effect only models.
+This creates the heat effect only models.
 ```{r}
@@
 library(lme4)
-men_old_finals_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_old_finals, 
+men_old_pooled_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = men_old_pooled, 
                              family = Gamma(link = "log"))
-men_finals_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_finals, 
-                         family = Gamma(link = "log"))
-men_old_pooled_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_old_pooled, 
-                             family = Gamma(link = "log"))
-men_pooled_batch_gamma_log <-glmer(RxnTime ~ (1 | Batch), data = men_pooled, 
+men_pooled_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = men_pooled, 
                          family = Gamma(link = "log"))
@@
 ```{r}
-summary(men_pooled_batch_gamma_log)
+summary(men_pooled_heat_gamma_log)
 ```
 
-This creates the models with a year effect and batch effect.
+This creates the models with a year effect and heat effect.
 ```{r}
@@
 library(lme4)
-men_old_finals_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+men_old_finals_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
                                    data = men_old_finals, family = Gamma(link = "log"))
-men_finals_two_effect_gamma_log<- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+men_finals_two_effect_gamma_log<- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
                               data = men_finals, family = Gamma(link = "log"))
-men_old_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+men_old_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
                                    data = men_old_pooled, family = Gamma(link = "log"))
-men_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), 
+men_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
                                data = men_pooled, family = Gamma(link = "log"))
@@
 ```{r}
-anova(men_pooled_batch_gamma_log, men_pooled_two_effect_gamma_log)
+anova(men_pooled_heat_gamma_log, men_pooled_two_effect_gamma_log)
 ```
 This anova test determined that the two effect model is an improvement over the
-batch effect model. The test statistic is 0.01059 which is significant at alpha 
+heat effect model. The test statistic is 0.01059 which is significant at alpha 
 = 0.05 but not at alpha = 0.01.
@@
     sd_year <- vc$sdcor[1]
-    sd_year_batch <- vc$sdcor[2]
+    sd_year_heat <- vc$sdcor[2]
 
     z1 <- rnorm(n, sd = sd_year)
-    z2 <- rnorm(n, sd = sd_year_batch)
+    z2 <- rnorm(n, sd = sd_year_heat)
 
@@
 #13
-sim_one_effect_time(men_old_finals_batch_gamma_log, men_old_finals_one_gamma_log, 0.1)
-sim_one_effect_time(men_finals_batch_gamma_log, men_finals_one_gamma_log, 0.1)
-sim_one_effect_time(men_old_pooled_batch_gamma_log, men_old_pooled_one_gamma_log, 0.1)
-sim_one_effect_time(men_pooled_batch_gamma_log, men_pooled_one_gamma_log, 0.1)
+sim_one_effect_time(men_old_finals_heat_gamma_log, men_old_finals_one_gamma_log, 0.1)
+sim_one_effect_time(men_finals_heat_gamma_log, men_finals_one_gamma_log, 0.1)
+sim_one_effect_time(men_old_pooled_heat_gamma_log, men_old_pooled_one_gamma_log, 0.1)
+sim_one_effect_time(men_pooled_heat_gamma_log, men_pooled_one_gamma_log, 0.1)
 ```

### 2023-09-08 Manuscript/manuscript.tex
fixed broken table

@@
       \midrule
-      Year Effect Only & $-2023.5$ & N/A & 0.048 & $-1.857$ & 0.152
-      Heat Effect Only & $-2070.7$ & 0.073 & N/A & $-1.890$ & 0.135
-      Full model & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134 \\
+      Year Effect Only & $-2023.5$ &       & 0.048 & $-1.857$ & 0.152\\
+      Heat Effect Only & $-2070.7$ & 0.073 &       & $-1.890$ & 0.135\\
+      Full model       & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
       \bottomrule

### 2023-09-12 Manuscript/manuscript.tex
Almost done with notes

@@
 clusters of size two, three clusters of size three, six clusters of size four,
-and six clusters of size five.
+and six clusters of size five.  Although 2022 is presented three times in
+figure~\ref{fig:RankScatterplots}, the subset of athletes differs each time.
+There may be athletes who competed in 2019 and 2022 and their national
+championships data was available, but they did not compete in 2023, and thus their
+times show up in the top two graphs.
 
@@
 athletes are on the starting blocks before the gun goes off.  Thus, we can try
-to reduce this variability by accounting for this in the model through the batch
+to reduce this variability by accounting for this in the model through the heat
 effect.
@@
 for the pooled datasets and shows the standard error of the year
-effect increases when
-including 2022.  The two effect model with venue
-and heat effect was the best model, as indicated by an ANOVA test 
-\eds{I think it may be better to call this a likelihood ratio test rather than
-an ANOVA test, assuming that is the formal test being used} and decrease in
-AIC.  For the entire data, the AIC for the venue effect only model was
+effect increases when including 2022.  The two effect model with venue and heat
+effect was the best model as evidenced by a decrease in the AIC compared to the
+other models. For the entire data, the AIC for the venue effect only model was
 $-2023.5$, the AIC for the heat effect only model $-2070.7$, and the AIC for the 
 two effect model, which we will present throughout this section, was $-2075.7$. 
-Thus, the two
-effect model is superior and will be presented throughout the rest of the paper.
+Thus, the two effect model is superior and will be presented throughout the rest
+of the paper.
 \of{How specific do I need to be in this section? Will people who read this paper
@@
 and the degrees of freedom associated with the test.}
+\of{Professor Yan recommended not even mentioning the likelihood ratio so
+instead I will just talk about the AIC.}
 
@@
       \midrule
-      Year Effect Only & $-2023.5$ &       & 0.048 & $-1.857$ & 0.152\\
-      Heat Effect Only & $-2070.7$ & 0.073 &       & $-1.890$ & 0.135\\
+      Year Effect Only & $-2023.5$ & N/A & 0.048 & $-1.857$ & 0.152\\
+      Heat Effect Only & $-2070.7$ & 0.073 & N/A & $-1.890$ & 0.135\\
       Full model       & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
@@
    \toprule
-   Data Set & P(RxnTime $<$ x = $10^{-3}$) & P(RxnTime $<$ x = $10^{-4}$) \\ 
+   Data Set & P(RxnTime $< x = 10^{-3}$) & P(RxnTime $< x = 10^{-4}$) \\ 
    \midrule
-   Old Pooled & $0.094$ & $0.084$ \\ 
-   All Pooled & $0.092$ & $0.083$ \\
+   Excluding 2022 & $0.094$ & $0.084$ \\ 
+   Including 2022 & $0.092$ & $0.083$ \\
    \bottomrule
@@
 In both cases, the national versus international comparison
-produced stronger \eds{lower?} p values than the 2019 versus 2022 comparison, with the
+produced lower p values than the 2019 versus 2022 comparison, with the
 national vs international \texttt{exact = TRUE} test producing a p value of 
@@
    2022 vs 2023 & $3.0\cdot10^{-3}$ & $1.0\cdot10^{-2}$ \\
-   Nat vs Inter & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ \\
+   2022 National vs International & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ \\
    \bottomrule
@@
 
-%Since 1985 Seiko Holding Corporations has served every World Athletics 
-%Championships as the official timer \citep{wa2022seiko}.  Reaction times were
-%not officially measured until 1999 in the men's 110 meter hurdles and women's
-%100 meter hurdles. Seiko's technology for detecting 
-%reaction times relies on the pressure that athletes exert on the plate when they 
-%push off.  Their systems measure the time differential between when the ``gun" 
-%goes off to start the race and when the pressure changes \citep{wa2022seiko}.  
-%If an athlete has a reaction time under 0.1 seconds, it is deemed a false start 
-%as it is considered that no athlete can react so quickly \citep{Seiko-Timing}.  
-%Thus, a reaction time of for example 0.05, suggests that the athlete predicted 
-%the gun and it was luck that caused their abnormally high time.  At the World 
-%Championships level, World Athletics want to remove that element of 
-%luck and thus impose the 0.1 second barrier.
-%\eds{I would argue that this paragraph should go in the discussion as there are 
-%no results presented here}
-%\of{I agree so I moved it here. I think it also could fit into the top of the
-%data section as in section~\ref{sec:Data}. Thoughts?}
-
-%\eds{I would suggest removing the last two sentences, as they would suggest your
-% modeling with GLMM is not good}.  
-%\of{I agree and they have been removed}
-
-%If there are issues with reaction time such as athletes approaching the 0.1 
-%second barrier again at the next World Championships (August 2023), then World 
-%Athletics needs to adjust that threshold and allow for faster reaction times so 
-%that athletes with superb reflexes are not penalized.  World Athletics did not
-%do anything in response to the \citet{komi2009iaaf} study, but perhaps they 
-%should now so that the story of Devon Allen is not repeated.
-
-\section*{Appendix}
-\label{sec:Appendix}
-%Put downloadable data spreadsheet
-The World Championships data are available from the World Athletics website:
-\url{https://www.worldathletics.org/results/world-athletics-championships}.
-
-\jy{You have countries beyond United States}
-The United States Track and Field Championships data are available at: 
-\url{https://www.flashresults.com/2022_Meets/Outdoor/06-23_USATF/}
-
-\jy{Go over the references to see clues to clean in the bib source}
 \bibliographystyle{chicago}

### 2023-09-12 Manuscript/manuscript.tex
Edited a few tables, fixed spacing in data section

@@
 
-Devon Allen, an alumnus of the University of Oregon in 
-Eugene, Oregon, was
+Devon Allen, an alumnus of the University of Oregon in Eugene, Oregon, was
 expected to be a contender for the 2022 World Track and Field Championships 110 
@@
 (only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
-in third at the United States Track and Field 
-Championships also in Eugene in late June, Allen was poised to compete in the 
-2022 World Track and Field Championships at the same Eugene venue in August.
+in third at the United States Track and Field Championships also in Eugene 
+in late June, Allen was poised to compete in the 2022 World Track and 
+Field Championships in August, which would also be held in Eugene, Oregon.
 At the World Championships, Allen made it through the preliminary
@@
 held within three months of the 2022 World Championships).
-%\eds{is this true? are you still using the 2019 world championships data for the
-%rank-based analyses, or only the national competitions?}
-%\of{Yes the ranked based comparison has 2 subsections: 2022 national vs 2022
-%international and 2019 international vs 2022 international.  Should I instead say
-%that there are three types of data across 2 sections or something like that?}
-%\eds{I think this is fine here, but it is not at all clear that there are two
-%subsections in the rank-based results for 2022 national vs 2022
-%international and 2019 international vs 2022 international.}
-%\of{Currently under the Methods/Results section I have GLMM and rank based comparison
-%do you think I need subsections under rank based comparison, or is that too much
-%compartmentalization?  Or should I instead say there are 3 types of data that are
-%used in two different methods, or something like that?} \eds{I prefer
-%subsections within rank-based comparison.}
+
 The first data set is used for the GLMM analysis described in 
@@
 110 meter hurdle and then look at a column for reaction times which are always from
-roughly 0.1 to 0.18 seconds.} Each athlete was 
-designated as a ``cluster'' and the stage of competition (national or international) 
-was specified for each competition within the cluster (athlete).
+roughly 0.1 to 0.18 seconds.}
+
+Each athlete was designated as a ``cluster'' and the stage of competition (national
+or international) was specified for each competition within the cluster (athlete).
 Thus, this is subunit level grouping where we assign the treatment, national or
@@
    \toprule
-   Data Set & P(RxnTime $< x = 10^{-3}$) & P(RxnTime $< x = 10^{-4}$) \\ 
+   Data Set & P(RxnTime $< x) = 10^{-3}$ & P(RxnTime $< x) = 10^{-4}$ \\ 
    \midrule
@@
   \centering
-  \caption{}
+  \caption{2019 vs 2022 compares data from the same athletes who competed at the
+  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
+  from the same athletes who competed at the 2022 and 2023 World Track and Field
+  Championships. Nat vs Inter compares data from the same athletes who competed
+  at the 2022 national-level championships and the 2022 World Track and Field
+  Championships.}
   \begin{tabular}{c c c c c} 
@@
   \end{tabular}
-  \caption{2019 vs 2022 compares data from the same athletes who competed at the
-  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
-  from the same athletes who competed at the 2022 and 2023 World Track and Field
-  Championships. Nat vs Inter compares data from the same athletes who competed
-  at the 2022 national-level championships and the 2022 World Track and Field
-  Championships.}
   \label{tab:Clusrankresults}
@@
 differences in reaction times between men and women.
+
+This paper has shown that a reaction time of 0.1 seconds is not as extreme as
+one may think, and that based on the probabilities in table~\ref{tab:Sim_time},
+World Athletics should consider changing the disqualification barrier to 0.8
+seconds.  If they were to do so, they would allow athletes like Allen to be able
+to react faster without fear of being disqualified.  The second part of this
+paper showed that athletes' reaction times at the 2022 World Track and Field
+Championships were generally faster than their reaction times at other
+competitions.
 

### 2023-09-12 Code/Plots.Rmd
Made orange graph white, moved title in stack graph to inside

@@
 rank_scatter <- grid.arrange(rank1, rank2, rank3, ncol = 1)
-ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 9, width = 7)
+ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 8, width = 6)
 ```

### 2023-09-12 Code/Plots.Rmd
Made orange graph white, moved title in stack graph to inside

@@
 ggplot(men_pooled, aes(x = Year, y = RxnTime, group = Year)) +
-  geom_boxplot(fill = "orange1", color = "orangered4") +
+  geom_boxplot() +
   labs(x = "Year", y = "Reaction Time (s)") +
@@
   geom_point() +
-  labs(y = "Reaction Time", x = "", title = "2019 vs 2022") +
+  labs(y = "Reaction Time", x = "") +
+  annotate("text", y = 0.21, x = "2022 Semifinals", label = "2019 vs 2022", cex = 5) +
   ylim(0.08, 0.22) +
@@
   geom_point() +
-  labs(y = "Reaction Time", x = "", title = "2022 vs 2023") +
+  labs(y = "Reaction Time", x = "") +
+  annotate("text", y = 0.21, x = "2023 Semifinals", label = "2022 vs 2023", cex = 5) +
   ylim(0.08, 0.22) +
@@
 #"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
-colnames(national) <- c("Athlete", "Country", "National Heats", "National Semifinals", "National Finals", "2023 Heats", "2023 Semifinals", "2023 Finals", "Data Source")
+colnames(national) <- c("Athlete", "Country", "Nat. Heats", "Nat. Semifinals", "Nat. Finals", "Inter. Heats", "Inter. Semifinals", "Inter. Finals", "Data Source")
 national <- national[, -c(1,2, 9)]
 df_long <- tidyr::pivot_longer(national, cols = everything(), names_to = "Column", values_to = "Time")
-df_long$Column <- factor(df_long$Column, levels = c("National Heats", "National Semifinals", "National Finals", "2022 Heats", "2022 Semifinals", "2022 Finals"))
+df_long$Column <- factor(df_long$Column, levels = c("Nat. Heats", "Nat. Semifinals", "Nat. Finals", "Inter. Heats", "Inter. Semifinals", "Inter. Finals"))
 rank3 <- ggplot(df_long, aes(x = Column, y = Time)) +
   geom_point() +
-  labs(y = "Reaction Time", x = "", title = "2022 National vs International") +
+  labs(y = "Reaction Time", x = "") +
   ylim(0.08, 0.22) +
+  annotate("text", y = 0.21, x = "Inter. Semifinals", label = "2022 National vs International", cex = 5) +
   theme_minimal()
@@
 rank_scatter <- grid.arrange(rank1, rank2, rank3, ncol = 1)
-ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 8, width = 6)
+ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 8, width = 7)
 ```

### 2023-09-13 Code/lme4.R
Removed unneccessary files

@@ -1,160 +1,0 @@
-library(lme4)
-library(ggplot2)
-
-times <- read.csv("../Data/ThesisData.csv")
-colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
-
-## only use Finals, Male, not Disqualified runs, before 2022
-men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
-## boxplot(RxnTime ~ Year, data = finals, xlab = "Years")
-
-library(lme4)
-fit1 <- glmer(RxnTime ~ (1 | Year), data = men_finals, family = Gamma)
-fit2 <- glmer(RxnTime ~ (1 | Year), data = men_finals, family = Gamma(link = "log"))
-fit0 <- glm(RxnTime ~ 1, data = men_finals, family = Gamma(link = "log"))
-
-qqnorm(resid(fit2))
-
-## qqnorm(ranef(fit2)$Year[,1])
-## qqline(ranef(fit2)$Year[,1])
-
-
-## Fitted results
-sfit0 <- summary(fit0)
-pgamma(0.1, shape = 1 / sfit0$dispersion, scale = exp(coef(fit0)) * sfit0$dispersion)
-
-
-simfit <- function(fit2, n = 12000000) {
-  vc <- as.data.frame(VarCorr(fit2))
-  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
-  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
-  mu <- exp(fixef(fit2) + z)
-  shape <- 1 / vc$vcov[2]
-  x <- rgamma(n, shape = shape, scale = mu / shape)
-  x
-}
-
-x <- simfit(fit2)
-mean(x < 0.1)
-quantile(x, prob = c(.00001, .0001, .0005, .001))
-## compare with and without random effect
-plot(density(x)) # with random effect
-curve(dgamma(x, shape = 1 / sfit0$dispersion,
-             scale = exp(coef(fit0)) * sfit0$dispersion),
-      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
-
-## Does it make sense to pool the genders?
-women_finals <- subset(times, Type == "F" & Gender == "F" & Time != 'DQ')
-fit1 <- glmer(RxnTime ~ (1 | Year), data = women_finals, family = Gamma)
[…]

### 2023-09-13 Code/Plots.Rmd
Removed graphs not found in paper anymore

@@
 ```{r}
-times <- read.csv("../Data/ThesisData.csv")
-## View(times)
+times <- read.csv("../Data/Thesis2023.csv")
 colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
@@
 
-#DataSets
 ```{r}
-men_finals <- subset(times, Type == "F" & Gender == "M" & Time != 'DQ') # & Year < 2022)
-women_finals <- subset(times, Type == "F" & Gender == "F" & Time != 'DQ')
-
-men_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
-women_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "F" & Time != 'DQ')
-```
-
-#Models
-```{r}
-library(lme4)
-finals.lme <- lmer(RxnTime ~ (1|Year), data = men_pooled)
-```
-
-###Functions
-```{r}
-simfit <- function(fit2, n = 12000000) {
-  vc <- as.data.frame(VarCorr(fit2))
-  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
-  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
-  mu <- exp(fixef(fit2) + z)
-  shape <- 1 / vc$vcov[2]
-  x <- rgamma(n, shape = shape, scale = mu / shape)
-  x
-}
-
-x <- simfit(fit2)
-mean(x < 0.1)
-quantile(x, prob = c(.00001, .0001, .0005, .001))
-## compare with and without random effect
-plot(density(x)) # with random effect
-curve(dgamma(x, shape = 1 / sfit0$dispersion,
-             scale = exp(coef(fit0)) * sfit0$dispersion),
-      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
-```
-
-#Figures
-```{r}
-pdf(file = "TBY.pdf")
-plot(aggregate(men_pooled$RxnTime, list(men_pooled$Year), FUN = mean), xlab = "Year", ylab = "Reaction time in seconds", main = "Mean Reaction Time By Year")
-dev.off()
-```
-
-```{r}
-pdf(file = "Skew.pdf")
-hist(resid(finals.lme), xlab = "Residuals of Linear mixed effects model", main = "Right skew when plotting the residuals of the linear mixed effects model")
-dev.off()
-```
-```{r}
-pdf(file = "Gamma_skew.pdf")
-dev.off()
-```
-
-```{r}
-pdf(file = "DataBoxplot.pdf")
-boxplot(RxnTime ~ Year, data = men_pooled, xlab = "Year", ylab = "Reaction Time in seconds", main = "Men's Semifinal and Final Reaction Time Boxplot Data")
-dev.off()
-```
-
-```{r}
-pdf(file = "DataScatterplot.pdf")
-plot(RxnTime ~ Year, data = men_pooled, xlab = "Year", ylab = "Reaction Time in seconds", main = "Men's Semifinal and Final Reaction Time Scatterplot Data")
-dev.off()
-```
-
-```{r}
-set.seed(1)
-library(lme4)
-pdf(file = "Men_Density_Year.pdf")
-fit2 <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
-fit0 <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
-sfit0 <- summary(fit0)
-simfit <- function(fit2, n = 12000000) {
-  vc <- as.data.frame(VarCorr(fit2))
-  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
-  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
-  mu <- exp(fixef(fit2) + z)
-  shape <- 1 / vc$vcov[2]
-  x <- rgamma(n, shape = shape, scale = mu / shape)
-  x
-}
-
-x <- simfit(fit2)
-mean(x < 0.1)
-quantile(x, prob = c(.00001, .0001, .0005, .001))
-## compare with and without random effect
-plot(density(x), main = "Men's Pooled With Only Year Effect") # with random effect
-curve(dgamma(x, shape = 1 / sfit0$dispersion,
-             scale = exp(coef(fit0)) * sfit0$dispersion),
-      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
-legend("right", title = "Probability mean less than", legend = " 0.1 is 0.006847")
-dev.off()
-```
-
-```{r}
-set.seed(1)
-library(lme4)
-pdf(file = "Men_Density_Batch.pdf")
-fit2 <- glmer(RxnTime ~ (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
-fit0 <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
-sfit0 <- summary(fit0)
-simfit <- function(fit2, n = 12000000) {
-  vc <- as.data.frame(VarCorr(fit2))
-  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
-  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
-  mu <- exp(fixef(fit2) + z)
-  shape <- 1 / vc$vcov[2]
-  x <- rgamma(n, shape = shape, scale = mu / shape)
-  x
-}
-
-x <- simfit(fit2)
-mean(x < 0.1)
-quantile(x, prob = c(.00001, .0001, .0005, .001))
-## compare with and without random effect
-plot(density(x), main = "Men's Pooled With Only Batch Effect") # with random effect
-curve(dgamma(x, shape = 1 / sfit0$dispersion,
-             scale = exp(coef(fit0)) * sfit0$dispersion),
-      0.08, 0.22, add = TRUE, col = "red") # fixed effect only
-legend("right", title = "Probability mean less than", legend = " 0.1 is 0.0066115")
-dev.off()
-```
-
-```{r}
-set.seed(1)
-library(lme4)
-pdf(file = "Men_Density_Both.pdf")
-fit2 <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
-fit0 <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
-sfit0 <- summary(fit0)
-simfit <- function(fit2, n = 10000000) {
-  vc <- as.data.frame(VarCorr(fit2))
-  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
-  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
-  mu <- exp(fixef(fit2) + z)
-  shape <- 1 / vc$vcov[2]
-  x <- rgamma(n, shape = shape, scale = mu / shape)
-  x
-}
-
-x <- simfit(fit2)
-mean(x < 0.1)
-quantile(x, prob = c(.00001, .0001, .0005, .001))
-## compare with and without random effect
-plot(density(x), main = "Men's Pooled With Year and Batch Effect") # with random effect
-curve(dgamma(x, shape = 1 / sfit0$dispersion,
-             scale = exp(coef(fit0)) * sfit0$dispersion),
-      0.08, 0.22, add = TRUE, col = "orange") # fixed effect only
-legend("right", title = "Probability mean less than", legend = " 0.1 is 0")
-dev.off()
-```
-
-```{r}
-set.seed(1)
-library(lme4)
-pdf(file = "Women_Density_Year.pdf")
-fit2 <- glmer(RxnTime ~ (1 | Year), data = women_pooled, family = Gamma(link = "log"))
-fit0 <- glm(RxnTime ~ 1, data = women_pooled, family = Gamma(link = "log"))
-sfit0 <- summary(fit0)
-simfit <- function(fit2, n = 10000000) {
-  vc <- as.data.frame(VarCorr(fit2))
-  ## mu <- exp(coef(fit2)$Year[,]) ## using fitted random effects only
-  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
-  mu <- exp(fixef(fit2) + z)
-  shape <- 1 / vc$vcov[2]
-  x <- rgamma(n, shape = shape, scale = mu / shape)
-  x
-}
-
-x <- simfit(fit2)
-mean(x < 0.1)
-quantile(x, prob = c(.00001, .0001, .0005, .001))
-## compare with and without random effect
-plot(density(x), main = "Women's Pooled With Only Year Effect") # with random effect
-curve(dgamma(x, shape = 1 / sfit0$dispersion,
-             scale = exp(coef(fit0)) * sfit0$dispersion),
-      0.08, 0.22, add = TRUE, col = "deeppink") # fixed effect only
-legend("right", title = "Probability mean less than", legend = " 0.1 is 0.0042296")
-dev.off()
-```
-
-```{r}
-year_effect <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
-batch_effect <- glmer(RxnTime ~ (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
-both <- glmer(RxnTime ~ (1 | Year) + (1 | Batch), data = men_pooled, family = Gamma(link = "log"))
-#anova(batch_effect, both)
-summary(both)
-```
-
-```{r}
-t.test(men_pooled$RxnTime[men_pooled$Year==2019],men_pooled$RxnTime[men_pooled$Year==2022])
-```
-
-```{r}
+with2022 <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
 library(ggplot2)
-ggplot(men_pooled, aes(x = Year, y = RxnTime, group = Year)) +
+ggplot(with2022, aes(x = Year, y = RxnTime, group = Year)) +
   geom_boxplot() +

### 2023-09-13 Code/GLMM.Rmd
Renamed variables

@@
 date: "2023-05-31"
-output: pdf_document
+output:
+  html_document:
+    df_print: paged
 ---
 
-Professors Yan and Schifano: to see the probabilities of the three different models
-please look at the results from code chunks 7, 12, 13.  You will need to run
-code from other chunks to generate the models and for the simulation code, but
-these three will produce probabilities of a reaction time less than 0.1 seconds.
+To see the probabilities associated with a reaction time below 0.1 seconds for
+the two effect model, run code chunks 1, 2, 5, 6, 7.
 
-The code below loads in the dataset from the csv file.
+The code below loads in the data sets from the csv file.
 ```{r}
@@
 times <- read.csv("../Data/Thesis2023.csv")
-colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Heat")
-men_pooled <- subset(times, (Type == "F" | Type == "S") & 
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
+with2022 <- subset(times, (Type == "F" | Type == "S") & 
                        Gender == "M" & Time != 'DQ')
-men_old_pooled <- subset(times, (Type == "F" | Type == "S") & 
-                           Gender == "M" & Time != 'DQ' & Year != 2022)
+without2022 <- subset(times, (Type == "F" | Type == "S") & 
+                           Gender == "M" & Time != 'DQ' & Venue != 2022)
 ```
@@
 This code below creates the base models without any random effects.  These are used
-as a baseline to compare the year effect models to in the simulation code below.
+as a baseline to compare the venue effect models to in the simulation code below.
 ```{r}
@@
 library(lme4)
-men_old_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_old_pooled, 
+without2022_one_gamma_log <- glm(RxnTime ~ 1, data = without2022, 
                                     family = Gamma(link = "log"))
-men_pooled_one_gamma_log <- glm(RxnTime ~ 1, data = men_pooled, 
+with2022_one_gamma_log <- glm(RxnTime ~ 1, data = with2022, 
                                 family = Gamma(link = "log"))
@@
 
-This code creates the year effect only models.
+This code creates the venue effect only models.
 ```{r}
@@
 library(lme4)
-men_old_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_old_pooled, 
+without2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = without2022, 
                                        family = Gamma(link = "log"))
-men_pooled_year_gamma_log <- glmer(RxnTime ~ (1|Year), data = men_pooled, 
+with2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = with2022, 
                                    family = Gamma(link = "log"))
@@
 ```{r}
-summary(men_pooled_year_gamma_log)
-summary(men_old_pooled_year_gamma_log)
+summary(with2022_venue_gamma_log)
+summary(without2022_venue_gamma_log)
 ```
@@
 library(lme4)
-men_old_pooled_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = men_old_pooled, 
+without2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = without2022, 
                              family = Gamma(link = "log"))
-men_pooled_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = men_pooled, 
+with2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = with2022, 
                          family = Gamma(link = "log"))
@@
 ```{r}
-summary(men_pooled_heat_gamma_log)
+summary(with2022_heat_gamma_log)
 ```
 
-This creates the models with a year effect and heat effect.
+This creates the models with a venue effect and heat effect.
 ```{r}
@@
 library(lme4)
-men_old_finals_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
-                                   data = men_old_finals, family = Gamma(link = "log"))
-men_finals_two_effect_gamma_log<- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
-                              data = men_finals, family = Gamma(link = "log"))
-men_old_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
-                                   data = men_old_pooled, family = Gamma(link = "log"))
-men_pooled_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Year) + (1 | heat), 
-                               data = men_pooled, family = Gamma(link = "log"))
-```
-```{r}
-summary(men_pooled_two_effect_gamma_log)
-summary(men_old_pooled_two_effect_gamma_log)
+without2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
+                                   data = without2022, family = Gamma(link = "log"))
+with2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
+                               data = with2022, family = Gamma(link = "log"))
 ```
@@
 ```{r}
-anova(men_pooled_heat_gamma_log, men_pooled_two_effect_gamma_log)
+summary(with2022_two_effect_gamma_log)
+summary(without2022_two_effect_gamma_log)
+```
+
+```{r}
+anova(with2022_heat_gamma_log, with2022_two_effect_gamma_log)
 ```
@@
 ```{r}
-anova(men_pooled_year_gamma_log, men_pooled_two_effect_gamma_log)
+anova(with2022_venue_gamma_log, with2022_two_effect_gamma_log)
 ```
 This anova test determined that the two effect model is a significant improvement
-over the year effect model.
+over the Venue effect model.
 
@@
     vc <- as.data.frame(VarCorr(model))
-    sd_year <- vc$sdcor[1]
-    sd_year_heat <- vc$sdcor[2]
+    sd_Venue <- vc$sdcor[1]
+    sd_Venue_heat <- vc$sdcor[2]
 
-    z1 <- rnorm(n, sd = sd_year)
-    z2 <- rnorm(n, sd = sd_year_heat)
+    z1 <- rnorm(n, sd = sd_Venue)
+    z2 <- rnorm(n, sd = sd_Venue_heat)
 
@@
 set.seed(1)
-sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.1)
-sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.1)
+sim_two_effect_time(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.1)
+sim_two_effect_time(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.1)
 ```
@@
 set.seed(1)
-sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.09)
-sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.09)
+sim_two_effect_time(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.09)
+sim_two_effect_time(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.09)
 ```
@@
 set.seed(1)
-sim_two_effect_time(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.08)
-sim_two_effect_time(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.08)
+sim_two_effect_time(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.08)
+sim_two_effect_time(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.08)
 ```
@@
 set.seed(1)
-sim_two_effect_prob(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.001)
-sim_two_effect_prob(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.001)
+sim_two_effect_prob(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.001)
+sim_two_effect_prob(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.001)
 ```
@@
 set.seed(1)
-sim_two_effect_prob(men_old_pooled_two_effect_gamma_log, men_old_pooled_one_gamma_log, 0.0001)
-sim_two_effect_prob(men_pooled_two_effect_gamma_log, men_pooled_one_gamma_log, 0.0001)
+sim_two_effect_prob(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.0001)
+sim_two_effect_prob(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.0001)
 ```
@@
     vc <- as.data.frame(VarCorr(model))
-    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
+    ## mu <- exp(coef(model)$Venue[,]) ## using fitted random effects only
     z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
@@
 #12
-sim_one_effect_time(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log, 0.1)
-sim_one_effect_time(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.1)
-sim_one_effect_time(men_old_pooled_year_gamma_log, men_old_pooled_one_gamma_log, 0.1)
-sim_one_effect_time(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.1)
+set.seed(1)
+sim_one_effect_time(without2022_venue_gamma_log, without2022_one_gamma_log, 0.1)
+sim_one_effect_time(with2022_venue_gamma_log, with2022_one_gamma_log, 0.1)
 ```
@@
 #13
-sim_one_effect_time(men_old_finals_heat_gamma_log, men_old_finals_one_gamma_log, 0.1)
-sim_one_effect_time(men_finals_heat_gamma_log, men_finals_one_gamma_log, 0.1)
-sim_one_effect_time(men_old_pooled_heat_gamma_log, men_old_pooled_one_gamma_log, 0.1)
-sim_one_effect_time(men_pooled_heat_gamma_log, men_pooled_one_gamma_log, 0.1)
+set.seed(1)
+sim_one_effect_time(without2022_heat_gamma_log, without2022_one_gamma_log, 0.1)
+sim_one_effect_time(with2022_heat_gamma_log, with2022_one_gamma_log, 0.1)
 ```
@@
 #15
-sim_one_effect_prob(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log, 0.001)
-sim_one_effect_prob(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.001)
-sim_one_effect_prob(men_old_pooled_year_gamma_log, men_old_pooled_one_gamma_log, 0.001)
-sim_one_effect_prob(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.001)
+sim_one_effect_prob(without2022_venue_gamma_log, without2022_one_gamma_log, 0.001)
+sim_one_effect_prob(with2022_venue_gamma_log, with2022_one_gamma_log, 0.001)
 ```
@@
 #16
-sim_one_effect_prob(men_old_finals_year_gamma_log, men_old_finals_one_gamma_log, 0.0001)
-sim_one_effect_prob(men_finals_year_gamma_log, men_finals_one_gamma_log, 0.0001)
-sim_one_effect_prob(men_old_pooled_year_gamma_log, men_old_pooled_one_gamma_log, 0.0001)
-sim_one_effect_prob(men_pooled_year_gamma_log, men_pooled_one_gamma_log, 0.0001)
+sim_one_effect_prob(without2022_venue_gamma_log, without2022_one_gamma_log, 0.0001)
+sim_one_effect_prob(with2022_venue_gamma_log, with2022_one_gamma_log, 0.0001)
 ```

### 2023-09-13 Manuscript/manuscript.tex
Added final paragraph, edited small stuff throughout

@@
 
-%\of{Should I report table~\ref{tab:EffectComparison}}
-%\begin{table}
-%  \centering
-%  \caption{Compares the various effect models by dataset, all probabilities are
-%  for observing a reaction time less than 0.1 seconds. In general the model with
-%  the highest probability (least extreme reaction time) is the Year Effect model.}
-%  \begin{tabular}{c c c c} 
-%   \toprule
-%   Data Set & Year Effect & Batch Effect & Both Effects \\ 
-%   \midrule
-%   Old Pooled & $5.73\cdot10^{-3}$ & $5.21\cdot10^{-3}$ & $4.15\cdot10^{-3}$ \\ 
-%   All Pooled & $6.84\cdot10^{-3}$ & $6.64\cdot10^{-3}$ & $5.15\cdot10^{-3}$ \\
-%   \bottomrule
-%  \end{tabular}
-%  \label{tab:EffectComparison}
-%\end{table}
-
 \subsection{Rank-based Comparison} \label{subsec:Results_Rank}
@@
 Championships were generally faster than their reaction times at other
-competitions.
+competitions. Table~\ref{tab:Clusrankresults} shows multiple significant p values
+that indicate differences in average reaction times for athletes who competed at
+the 2022 World Track and Field Championship and another competition.  Taken
+together, the results show that 2022 was an outlier and despite being disqualified,
+Allen's time was not so extreme.
 

### 2023-09-13 Manuscript/manuscript.tex
Removed tables that were after the document

@@
 \end{document}
-
-Here are important statistics grouped by both model and data set:
-All Men final's Data
-\begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c |} 
-   \hline\hline
-   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
-   \hline
-   linear one & -472.7 & 238.3 & 0.1484 & 0.0019 & N/A & N/A \\ 
-   \hline
-   linear year & -468.7 & 237.4 & 0.1481 & 0.0030 & 7.762e-05 & 2.457e-04 \\
-   \hline
-   men finals one gamma & -477.3 & 240.7 & 6.738 & 0.0844 & N/A & N/A \\
-   \hline
-   men finals year gamma & -491.1 & 248.5 & 6.8121 & 0.1895 & 0.1148 & 0.0112 \\
-   \hline
-   men finals one gamma log & -477.3 & 240.7 & -1.908 & 0.0125 & N/A & N/A \\
-   \hline
-   men finals year gamma log & -491.0 & 248.5 & -1.914 & 0.0280 & 0.0025 & 0.0112 \\ [0.5ex]
-   \hline
-  \end{tabular}
-\end{center}
-
-Pre-2022 final's data
-\begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c |} 
-   \hline\hline
-   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
-   \hline
-   linear one & -450.7 & 227.4 & 0.1495 & 0.0019 & N/A & N/A \\
-   \hline
-   linear year & -444.1 & 225.0 & 0.1496 & 0.0028 & 5.691e-05 & 2.470e-04 \\ 
-   \hline
-   gamma one & -455.8 & 229.9 & 6.687 & 0.0834 & N/A & N/A \\
-   \hline
-   gamma year & -465.1 & 235.5 & 6.714 & 0.1674 & 0.0859 & 0.0109 \\
-   \hline
-   gamma one log & -455.8 & 229.9 & -1.900 & 0.0125 & N/A & N/A \\
-   \hline
-   gamma year log & -465.0 & 235.5 & -1.901 & 0.0251 & 0.0019 & 0.0109 \\ [0.5ex]
-   \hline
-  \end{tabular}
-  \end{center}
-
-All Men's semifinal and finals data
-\begin{center}
-  \begin{tabular}{||c | c c c | c c c||} 
-   \hline
-   Athlete & United States H & United States S & United States F & World H & World S & World F \\ [0.5ex] 
-   \hline\hline
-   Devon Allen & 0.201 & 0.153 & 0.160 & 0.123 & 0.101 & 0.099 \\ 
-   \hline
-   Trey Cunningham & 0.186 & 0.185 & 0.182 & 0.115 & 0.120 & 0.109 \\
-   \hline
-   Grant Holloway & 0.192 & 0.190 & N/A & 0.147 & 0.128 & 0.124 \\
-   \hline
-   Daniel Roberts & 0.181 & 0.200 & 0.183 & 0.179 & N/A & N/A \\ [0.5ex]
-   \hline
-  \end{tabular}
-  \end{center}
-
-Pre-2022 Men's semi finals and finals data
-\begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c |} 
-   \hline
-   Model & AIC & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
-   \hline\hline
-   linear one & -1539 & 771.5 & 0.1544 & 0.0015 & N/A & N/A \\
-   \hline
-   linear year & -1564 & 785.4 & 0.1562 & 0.0037 & 1.280e-04 & 6.273e-04 \\ 
-   \hline
-   gamma one & -1626 & 814.9 & 6.475 & 0.0608 & N/A & N/A \\
-   \hline
-   gamma year & -1688 & 846.8 & 6.426 & 0.1923 & 0.0989 & 0.0243 \\
-   \hline
-   gamma one log & -1626 & 814.9 & -1.868 & 0.0094 & N/A & N/A \\
-   \hline
-   gamma year log & -1687 & 846.6 & -1.853 & 0.0305 & 0.0024 & 0.0243 \\ [0.5ex]
-   \hline
-  \end{tabular}
-  \end{center}
-
-Sorted By Model instead:
-Linear Model with no venue effect
-\begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c |} 
-   \hline
-   Dataset & AIC & Log Likelihood & Int effect est & Int se \\ [0.5ex] 
-   \hline\hline
-   All Finals & -472.7 & 238.3 & 0.1484 & 0.0019 \\
-   \hline
-   Older Finals & -450.7 & 227.4 & 0.1495 & 0.0019 \\ 
-   \hline
-   All Pooled & -1667 & 835.5 & 0.1529 & 0.0014 \\
-   \hline
-   Older Pooled & -1539 & 771.5 & 0.1544 & 0.0015 \\
-   \hline
-  \end{tabular}
-  \end{center}
-
-Linear model with venue effect
-\begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c | c |} 
-   \hline
-   Dataset & AIC & Year se & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
-   \hline\hline
-   All Finals & -468.7 & 0.0088 & 237.4 & 0.1481 & 0.0030 & 7.762e-05 & 2.457e-04 \\
-   \hline
-   Older Finals & -444.1 & 0.0075 & 225.0 & 0.1496 & 0.0028 & 5.691e-05 & 2.470e-04 \\ 
-   \hline
-   All Pooled & -1713 & 0.131 & 859.3 & 0.1541 & 0.0040 & 1.719e-04 & 5.926e-04 \\
-   \hline
-   Older Pooled & -1564 & 0.113 & 785.4 & 0.1562 & 0.0037 & 1.280e-04 & 6.273e-04 \\
-   \hline
-  \end{tabular}
-  \end{center}
-
-Gamma model with venue effect
-\begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c | c |} 
-   \hline
-   Dataset & AIC & Year se & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
-   \hline\hline
-   All Finals & -491.1 & 0.3388 & 248.5 & 6.8121 & 0.1895 & 0.1148 & 0.0112 \\
-   \hline
-   Older Finals & -465.1 & 0.2930 & 235.5 & 6.714 & 0.1674 & 0.0859 & 0.0109 \\ 
-   \hline
-   All Pooled & -1848 & 0.3612 & 926.9 & 6.553 & 0.2269 & 0.1305 & 0.0234 \\
-   \hline
-   Older Pooled & -1688 & 0.3145 & 846.8 & 6.426 & 0.1923 & 0.0989 & 0.0243 \\
-   \hline
-  \end{tabular}
-  \end{center}
-
-Gamma model with venue effect and link=log
-\begin{center}
-  \begin{tabular}{|c | c | c | c | c | c | c | c |} 
-   \hline
-   Dataset & AIC & Year se & Log Likelihood & Int effect est & Int se & Int var & Resid Var \\ [0.5ex] 
-   \hline\hline
-   All Finals & -491.0 & 0.0504 & 248.5 & -1.914 & 0.0280 & 0.0025 & 0.0112 \\
-   \hline
-   Older Finals & -465.0 & 0.0439 & 235.5 & -1.901 & 0.0251 & 0.0019 & 0.0109 \\ 
-   \hline
-   All Pooled & -1848 & 0.0554 & 927.0 & -1.869 & 0.0348 & 0.0031 & 0.0235 \\
-   \hline
-   Older Pooled & -1687 & 0.0495 & 846.6 & -1.853 & 0.0305 & 0.0024 & 0.0243 \\
-   \hline
-  \end{tabular}
-  \end{center}

### 2023-09-13 Manuscript/manuscript.tex
Changed order of glmm and rank

@@
 how data was collected and begins to detail the generalized linear mixed-effect 
-(GLMM) model that is further developed in Section~\ref{sec:Methods}. The 
+(GLMM) model that is further developed in Section~\ref{sec:glmm}. The 
 signed-rank test that analyzes reaction times for the same athletes across
 different competitions is
-also described in Section~\ref{sec:Methods}.  Section~\ref{sec:Results}
+also described in Section~\ref{sec:rank}.  Section~\ref{sec:Results}
  provides the results of the two analyses and ends 
@@
 
-There are two types of data that are explored throughout this paper: data from
-1999-2023 for every world championship, and data from athletes who competed in
-the 2022 World Championships and another competition (either 2019 World 
-Championships, 2023 World Championships or a 2022 national competition typically
-held within three months of the 2022 World Championships).
-
-The first data set is used for the GLMM analysis described in 
-Section~\ref{sec:glmm}, while the second collection of data is used for the 
-rank-based comparisons described in Section~\ref{sec:rank}.  This paper was
-started shortly after the 2022 World Championships, but as the 2023 data has 
-become available recently, we are pleased to include it in our analysis, and it
-did not significantly change our results.
-
-\subsection{World Championships 1999--2023}\label{sec:dataworld}
-
-Data was taken from the World Athletics 
-and covers the men's 110 meter hurdles from 1999 to 2023; the source can be 
-found in the Appendix.
-Although women's data is available, we chose not to include women's data in the
-analyses to follow due to the numerous papers citing differences in men and
-women's reaction times \citep[e.g.,][]{lipps2011implications, 
-babicc2009reaction, panoutsakopoulos2020gender}.
-For the creation of the GLMM model we focus on the reaction times during the 
-semifinal and final heats only, as reaction times from preliminary heats are 
-often not as fast as those in later heats \citep[e.g.,][]{zhang2021correlation}.
-
-For analysis purposes, we will often pool reaction times of men's final and 
-men's semifinal heats together which increases the sample size. This is 
-advantageous because in some years there are very few finals observations, such
-as in 2022, when there were only five data points due to two disqualifications 
-and one athlete who did not compete.  This is the data that we will reference
-throughout the rest of the paper unless otherwise noted. We also consider data 
-that excludes 2022 to see how our analysis differs based on the exclusion of one
-particular year of interest.
-
-The data under consideration is summarized using a box plot in 
-Figure~\ref{fig:Boxplot}.  It is clear that the reaction times in the 2022
-box plot are lower (indicating faster reaction times) than in many of the other
-years.  The median reaction time in 2022 was 0.129 seconds,  compared to 0.156
-seconds, which is what \citet{brosnan2017effects} found when looking at data
-from 1999 to 2014. We can also see from Figure~\ref{fig:Boxplot} that the 
-reaction times clearly differ from year to year. The venue of the World 
-Championships changes each year, so weather or climate related factors such as 
-humidity, precipitation, elevation, may indeed play a role, but additionally,
-both the technology and false start penalization rules have changed over the 
-study period.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics{Boxplot}
-  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.
-	\eds{there is only one plot here, not two. If you are keeping only one, the
-	text and caption need to be modified accordingly}}
-  \of{Resolved}
-  \label{fig:Boxplot}
-\end{figure}
-
-From 2007 to 2009, the IAAF (the former name for World Athletics, and the 
-governing body for the World Track and Field Championships) instituted a rule 
-change that allowed one warning false start before a sprinter was disqualified 
-\citep{iaaf2009falsestart}. There were 18 male and 7 female false starts at the 
-2007 World Championships, and 18 male and 7 female false starts at the 2009 World 
-Championships. Starting in 2011, the rule was scrapped, and the old policy which
-automatically disqualifies runners who false start was reinstated. This was 
-desirable for World Athletics as false starts can make already lengthy track 
-meets more tedious, both for athletes and viewers watching the television 
-broadcast. By returning to the harsher policy and cracking down on false starts,
-World Athletics had reduced men's false starts by two thirds in 2011 (6 male and
-4 female disqualifications) \citep{iaaf2009falsestart}. \citet{haugen2013effect}
-examined the effect of different false start rules that IAAF has imposed from 
-1997 to 2009 and found statistically significant improvement in reaction times 
-under more lenient rules.
+There are two types of data that are explored throughout this paper.  For the
+signed-rank tests described in section~\ref{sec:rank}, the data comes from 
+athletes who competed in the 2022 World Championships and another competition 
+(2019 World Championships, 2023 World Championships or 2022 national 
+championships). For the GLMM described in section~\ref{sec:glmm}, data comes
+from every world championship from 1999 to 2023. This paper was started shortly
+after the 2022 World Championships, but as the 2023 data has become available 
+recently, we are pleased to include it in both of our analyses, and it did not
+significantly change our results.
 
@@
 
-%\eds{check these numbers? 0.099?}\of{resolved}
+We also started to perform a similar analysis to compare the times of female 
+athletes who competed at the 2019 and 2022 World Track and Field Championships,
+however as our results were not significant, we decided to incorporate another
+method to investigate the times at the 2022 World Championships and thus we
+employ the GLMM, which looks at historical men's data to try to determine
+whether Allen's time and disqualification are reasonable.
+
+\subsection{World Championships 1999--2023}\label{sec:dataworld}
+
+Data was taken from the World Athletics 
+and covers the men's 110 meter hurdles from 1999 to 2023; the source can be 
+found in the Appendix.
+Although women's data is available, we chose not to include women's data in the
+analyses to follow due to the numerous papers citing differences in men and
+women's reaction times \citep[e.g.,][]{lipps2011implications, 
+babicc2009reaction, panoutsakopoulos2020gender}.
+For the creation of the GLMM model we focus on the reaction times during the 
+semifinal and final heats only, as reaction times from preliminary heats are 
+often not as fast as those in later heats \citep[e.g.,][]{zhang2021correlation}.
+
+For analysis purposes, we will often pool reaction times of men's final and 
+men's semifinal heats together which increases the sample size. This is 
+advantageous because in some years there are very few finals observations, such
+as in 2022, when there were only five data points due to two disqualifications 
+and one athlete who did not compete.  This is the data that we will reference
+throughout the rest of the paper unless otherwise noted. We also consider data 
+that excludes 2022 to see how our analysis differs based on the exclusion of one
+particular year of interest.
+
+The data under consideration is summarized using a box plot in 
+Figure~\ref{fig:Boxplot}.  It is clear that the reaction times in the 2022
+box plot are lower (indicating faster reaction times) than in many of the other
+years.  The median reaction time in 2022 was 0.129 seconds,  compared to 0.156
+seconds, which is what \citet{brosnan2017effects} found when looking at data
+from 1999 to 2014. We can also see from Figure~\ref{fig:Boxplot} that the 
+reaction times clearly differ from year to year. The venue of the World 
+Championships changes each year, so weather or climate related factors such as 
+humidity, precipitation, elevation, may indeed play a role, but additionally,
+both the technology and false start penalization rules have changed over the 
+study period.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics{Boxplot}
+  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.
+	\eds{there is only one plot here, not two. If you are keeping only one, the
+	text and caption need to be modified accordingly}}
+  \of{Resolved}
+  \label{fig:Boxplot}
+\end{figure}
+
+From 2007 to 2009, the IAAF (the former name for World Athletics, and the 
+governing body for the World Track and Field Championships) instituted a rule 
+change that allowed one warning false start before a sprinter was disqualified 
+\citep{iaaf2009falsestart}. There were 18 male and 7 female false starts at the 
+2007 World Championships, and 18 male and 7 female false starts at the 2009 World 
+Championships. Starting in 2011, the rule was scrapped, and the old policy which
+automatically disqualifies runners who false start was reinstated. This was 
+desirable for World Athletics as false starts can make already lengthy track 
+meets more tedious, both for athletes and viewers watching the television 
+broadcast. By returning to the harsher policy and cracking down on false starts,
+World Athletics had reduced men's false starts by two thirds in 2011 (6 male and
+4 female disqualifications) \citep{iaaf2009falsestart}. \citet{haugen2013effect}
+examined the effect of different false start rules that IAAF has imposed from 
+1997 to 2009 and found statistically significant improvement in reaction times 
+under more lenient rules.
 
@@
 
-The data described in Sections~\ref{sec:dataworld} and~\ref{sec:databeyond} were
-analyzed with GLMM and rank-based methods, respectively.
+The data described in Sections~\ref{sec:databeyond} and~\ref{sec:dataworld} were
+analyzed with rank-based methods and a GLMM, respectively.
+
+\subsection{Rank-based Comparison}\label{sec:rank}
+
+To test the conjecture that the 2022 World Championships timing device may have 
+led to faster recorded reaction times, we compare the reaction times of the same
+athletes who have attended both the 2022 World Championships and other 
+competitions. 
+In this setting, we have clustered data with subunit grouping. In particular,
+each athlete is a cluster and the multiple reaction times from the same athlete
+can be from either the 2022 World Championships or otherwise.
+Let $X_{ij}$ be the $j$th reaction time of athlete~$i$, $i = 1, \ldots, n$,
+$j = 1, \ldots, m_i$ where $m_i$ is the number of observations from
+athelete~$i$. Let $\delta_{ij}$ be the group indicator of $X_{ij}$; $\delta_{ij}
+= 1$ if $X_{ij}$ is in group~1 (2022 World Championships) and $\delta_{ij} = 0$ 
+otherwise. Athletes are
+assumed to be independent, while subunit observations from the same athlete are
+not. The null hypothesis $H_0$ to be tested is that there is no difference
+between the two groups; i.e., the distribution of $X_{ij}$ remains the same
+regardless of the group indicator $\delta_{ij}$.
+
+\citet{datta2005rank} proposed an extension of the Wilcoxon rank-sum test to
+clustered data with subunit-level grouping. The test is designed based on a
+within-cluster resampling principle. Consider randomly picking one observation
+from each cluster to form a pseudo-sample. Let $X_i^*$ be a random pick from the
+$i$th cluster in the pseudo-sample and $\delta_i^*$ its group indicator. The
+Wilcoxon rank-sum statistic for the pseudo-sample is
+\[
+W^* = \frac{1}{n + 1} + \sum_{i=1}^{n} \delta_{i}^{*} R_{i}^{*},
+\]
+where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the pseudo-sample.
+The test statistic $S$ is the average of $W^*$ averaged over all possible
+pseudo-samples conditioning on the observed data and group indicators.
+The mean and variance of $S$ under $H_0$ can be derived so that $S$ can be
+standardized to form a statistic $Z$ which follows standard normal distribution
+asymptotically \citep[p.910]{datta2005rank}.
+
+With our small sample size, the asymptotic normal distribution may not be
+reliable, so we use 1000 random permutations to simulate the observed test 
+statistic.
+\eds{I do not understand what is meant in the text after the colon in the 
+previous sentence.  Please rephrase.}
+\of{I'm not sure what that was, I removed it now.}
+This method is available from the \texttt{clusWilcox.test()} function
+with \texttt{method = `ds'} and \texttt{exact = TRUE} from R package
+\texttt{clusrank} \citep{jiang2020wilcoxon}.
 
@@
 
-\subsection{Rank-based Comparison}\label{sec:rank}
+\section{Results} \label{sec:Results}
 
-To test the conjecture that the 2022 World Championships timing device may have 
-led to faster recorded reaction times, we compare the reaction times of the same
-athletes who have attended both the 2022 World Championships and other 
-competitions. 
-In this setting, we have clustered data with subunit grouping. In particular,
-each athlete is a cluster and the multiple reaction times from the same athlete
-can be from either the 2022 World Championships or otherwise.
-Let $X_{ij}$ be the $j$th reaction time of athlete~$i$, $i = 1, \ldots, n$,
-$j = 1, \ldots, m_i$ where $m_i$ is the number of observations from
-athelete~$i$. Let $\delta_{ij}$ be the group indicator of $X_{ij}$; $\delta_{ij}
-= 1$ if $X_{ij}$ is in group~1 (2022 World Championships) and $\delta_{ij} = 0$ 
-otherwise. Athletes are
-assumed to be independent, while subunit observations from the same athlete are
-not. The null hypothesis $H_0$ to be tested is that there is no difference
-between the two groups; i.e., the distribution of $X_{ij}$ remains the same
-regardless of the group indicator $\delta_{ij}$.
+\subsection{Rank-based Comparison} \label{subsec:Results_Rank}
 
-\citet{datta2005rank} proposed an extension of the Wilcoxon rank-sum test to
-clustered data with subunit-level grouping. The test is designed based on a
-within-cluster resampling principle. Consider randomly picking one observation
-from each cluster to form a pseudo-sample. Let $X_i^*$ be a random pick from the
-$i$th cluster in the pseudo-sample and $\delta_i^*$ its group indicator. The
-Wilcoxon rank-sum statistic for the pseudo-sample is
-\[
-W^* = \frac{1}{n + 1} + \sum_{i=1}^{n} \delta_{i}^{*} R_{i}^{*},
-\]
-where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the pseudo-sample.
-The test statistic $S$ is the average of $W^*$ averaged over all possible
-pseudo-samples conditioning on the observed data and group indicators.
-The mean and variance of $S$ under $H_0$ can be derived so that $S$ can be
-standardized to form a statistic $Z$ which follows standard normal distribution
-asymptotically \citep[p.910]{datta2005rank}.
+The rank based methods described in Section~\ref{sec:rank}
+to determine if the reaction times for athletes who had competed at other 
+championships differed significantly from the 2022 World Championships.  This is
+a two group comparison performed three times: comparing 2019 and 2022 World
+Championships, comparing 2022 and 2023 World Championships, and comparing 2022 
+national championships to the 2022 World Championships.
 
-With our small sample size, the asymptotic normal distribution may not be
-reliable, so we use 1000 random permutations to simulate the observed test 
-statistic.
-\eds{I do not understand what is meant in the text after the colon in the 
-previous sentence.  Please rephrase.}
-\of{I'm not sure what that was, I removed it now.}
-This method is available from the \texttt{clusWilcox.test()} function
-with \texttt{method = `ds'} and \texttt{exact = TRUE} from R package
-\texttt{clusrank} \citep{jiang2020wilcoxon}.
+Table~\ref{tab:Clusrankresults} shows the results generated from four tests in R.
+We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
+\texttt{B = 1000} indicating that the random permutation size was chosen to be 
+1000, as discussed previously in Section~\ref{sec:rank}.  
+%As the sample sizes are small, 65 and 49, we use \texttt{exact = TRUE} as
+%that is preferred.  
+In both cases, the national versus international comparison
+produced lower p values than the 2019 versus 2022 comparison, with the
+national vs international \texttt{exact = TRUE} test producing a p value of 
+$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  This
+is a highly significant test result and shows that athletes who competed at the
+2022 World Track and Field Championships were on average, significantly faster
+at reacting than they were a few months prior.  For the 2019 versus 2022
+same athlete comparison with \texttt{exact = TRUE} argument, the p value was
+significant at $\alpha = 0.01$, however when \texttt{exact = FALSE}, the p
+value was higher and only significant at $\alpha = 0.05$. 
+\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
+Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
+Nonetheless, these
+tests help to show that athletes competing at the 2022 World Track and Field
+Championships had faster reaction times than when they had competed previously.
 
-\section{Results} \label{sec:Results}
+\begin{table}
+  \centering
+  \caption{2019 vs 2022 compares data from the same athletes who competed at the
+  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
+  from the same athletes who competed at the 2022 and 2023 World Track and Field
+  Championships. Nat vs Inter compares data from the same athletes who competed
+  at the 2022 national-level championships and the 2022 World Track and Field
+  Championships.}
+  \begin{tabular}{c c c c c} 
+   \toprule
+   Comparison & Permutation & Asymptotic \\ 
+   \midrule
+   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10^{-2}$ \\
+   2022 vs 2023 & $3.0\cdot10^{-3}$ & $1.0\cdot10^{-2}$ \\
+   2022 National vs International & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:Clusrankresults}
+\end{table}
 
@@
 
-\subsection{Rank-based Comparison} \label{subsec:Results_Rank}
-
-The rank based methods described in Section~\ref{sec:rank}
-to determine if the reaction times for athletes who had competed at other 
-championships differed significantly from the 2022 World Championships.  This is
-a two group comparison performed two times: once comparing 2019 World
-Championships to 2022 World Championships and then comparing 2022 national
-championships to the 2022 World Championships.
-
-Table~\ref{tab:Clusrankresults} shows the results generated from four tests in R.
-We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
-\texttt{B = 1000} indicating that the random permutation size was chosen to be 
-1000, as discussed previously in Section~\ref{sec:rank}.  
-%As the sample sizes are small, 65 and 49, we use \texttt{exact = TRUE} as
-%that is preferred.  
-In both cases, the national versus international comparison
-produced lower p values than the 2019 versus 2022 comparison, with the
-national vs international \texttt{exact = TRUE} test producing a p value of 
-$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  This
-is a highly significant test result and shows that athletes who competed at the
-2022 World Track and Field Championships were on average, significantly faster
-at reacting than they were a few months prior.  For the 2019 versus 2022
-same athlete comparison with \texttt{exact = TRUE} argument, the p value was
-significant at $\alpha = 0.01$, however when \texttt{exact = FALSE}, the p
-value was higher and only significant at $\alpha = 0.05$. 
-\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
-Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
-Nonetheless, these
-tests help to show that athletes competing at the 2022 World Track and Field
-Championships had faster reaction times than when they had competed previously.
-
-\begin{table}
-  \centering
-  \caption{2019 vs 2022 compares data from the same athletes who competed at the
-  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
-  from the same athletes who competed at the 2022 and 2023 World Track and Field
-  Championships. Nat vs Inter compares data from the same athletes who competed
-  at the 2022 national-level championships and the 2022 World Track and Field
-  Championships.}
-  \begin{tabular}{c c c c c} 
-   \toprule
-   Comparison & Permutation & Asymptotic \\ 
-   \midrule
-   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10^{-2}$ \\
-   2022 vs 2023 & $3.0\cdot10^{-3}$ & $1.0\cdot10^{-2}$ \\
-   2022 National vs International & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ \\
-   \bottomrule
-  \end{tabular}
-  \label{tab:Clusrankresults}
-\end{table}
-
 \section{Concluding Remarks}\label{sec:concludingremarks}

### 2023-09-13 Manuscript/manuscript.tex
Fixed order of last paragraph

@@
 (I am not sure which is better, but perhaps the latter?)}
-of{I re-worked this graph completely, please let me know what you think.  For
+\of{I re-worked this graph completely, please let me know what you think.  For
 each graph I currently list them in chronological order: so 2022 then 2023 but
@@
 
-This paper has shown that a reaction time of 0.1 seconds is not as extreme as
-one may think, and that based on the probabilities in table~\ref{tab:Sim_time},
-World Athletics should consider changing the disqualification barrier to 0.8
-seconds.  If they were to do so, they would allow athletes like Allen to be able
-to react faster without fear of being disqualified.  The second part of this
-paper showed that athletes' reaction times at the 2022 World Track and Field
-Championships were generally faster than their reaction times at other
-competitions. Table~\ref{tab:Clusrankresults} shows multiple significant p values
-that indicate differences in average reaction times for athletes who competed at
-the 2022 World Track and Field Championship and another competition.  Taken
-together, the results show that 2022 was an outlier and despite being disqualified,
-Allen's time was not so extreme.
+The first part of this paper showed that athletes' reaction times at the 2022 
+World Track and Field Championships were generally faster than their reaction 
+times at other competitions. Table~\ref{tab:Clusrankresults} shows multiple 
+significant p values that indicate differences in average reaction times for 
+athletes who competed at the 2022 World Track and Field Championship and another
+competition.  Additionally, this paper has shown that a reaction time of 0.1 
+seconds is not as extreme as one may think, and that based on the probabilities
+in table~\ref{tab:Sim_time}, World Athletics should consider changing the 
+disqualification barrier to 0.8 seconds.  If they were to do so, they would 
+allow athletes like Allen to be able to react faster without fear of being 
+disqualified.  Taken together, the results show that 2022 was an outlier and 
+despite being disqualified, Allen's time was not so extreme.
 

### 2023-09-14 Manuscript/manuscript.tex
did not finish; got through sec 4.1

@@
 seconds, 0.001 seconds faster than what is allowed. Following the games, 
-bloggers on the running website LetsRun concluded that the reaction time data 
+bloggers on the running website LetsRun concluded that the reaction times  
 from the 2022 World Championships seemed to be generally faster compared to the 
-other datasets they considered, but they did not perform any formal 
+other reaction times they considered, but they did not perform any formal 
 statistical analysis. This paper questions the reaction time 
@@
 this is a reasonable threshold. We employ a generalized linear mixed model 
-(GLMM) with a random venue effect in order to  model the reaction time data. 
+(GLMM) with random venue (year) effect and a random heat effect within venue 
+in order to model the reaction time data. 
 Additionally, we employ a signed-rank test for clustered data to 
@@
 The crowd erupted with jeers
-when they found out that Allen was disqualified, angry at the result.  Had Allen 
-reacted just 0.001 seconds slower he would have been allowed to compete and been 
-spared waiting until the next World Track and Field Championships.
-\eds{He is not competing this year, correct? If yes, we may need to rephrase.}
-\of{Correct}
+when they realized that Allen was disqualified, angry at the result.  Had Allen 
+reacted just 0.001 seconds slower he would have been allowed to compete. %and been 
+%spared waiting until the next World Track and Field Championships.
+%\eds{He is not competing this year, correct? If yes, we may need to rephrase.}
+%\of{Correct}
+%\eds{removed because he did not compete this year, so did not actually have
+% to wait, after all.} 
+\eds{ If it is true he ``retired'' as a runner to play football, we could 
+instead say something like
+\textit{Had Allen reacted just 0.001 seconds slower he would have been allowed 
+to compete before retiring from running to play football.}}
 
@@
 and thus the possibility of a false disqualification would be lowered.
-The possibility of using a video-based medium has been proved to be a reliable
+The possibility of using a video-based medium has been shown to be a reliable
 measure of reaction time \citep{mudric2015evaluation}.
@@
 only by 0.001 seconds. It is also worth noting that one study looking at data 
-from various IAAF meets found that the fastest reaction times for men were found
+from various IAAF meets found that the fastest reaction times for men occurred 
 between the ages of 26 and 29 \citep{tonnessen2013reaction}.  This is notable 
@@
 the reaction times of every non-disqualification since the 1999 World 
-Championships, using a gamma linear mixed-effects model with two random effects 
+Championships, using a gamma linear mixed-effects model with two random effects: 
 one effect for the year and a second effect for the
@@
 every sprinter who steps onto the track.
-\eds{Since the World Championships are happening again now, is data from this 
-year already available? If it is available, we would need to justify its omission
-in the current analysis.
-If it is not available, but becomes available while under review, would we 
-reanalyze the data?  Does it seem likely our conclusions would change based on
-what is going on this year?}
-\of{Resolved}
+%\eds{Since the World Championships are happening again now, is data from this 
+%year already available? If it is available, we would need to justify its omission
+%in the current analysis.
+%If it is not available, but becomes available while under review, would we 
+%reanalyze the data?  Does it seem likely our conclusions would change based on
+%what is going on this year?}
+%\of{Resolved}
 
 The rest of this paper is organized as follows. Section~\ref{sec:Data} describes 
-how data was collected and begins to detail the generalized linear mixed-effect 
-(GLMM) model that is further developed in Section~\ref{sec:glmm}. The 
+how the data were collected and begins to detail the generalized linear mixed-effect 
+ model (GLMM) that is further developed in Section~\ref{sec:glmm}. The 
 signed-rank test that analyzes reaction times for the same athletes across
@@
 There are two types of data that are explored throughout this paper.  For the
-signed-rank tests described in section~\ref{sec:rank}, the data comes from 
+signed-rank tests described in Section~\ref{sec:rank}, the data comes from 
 athletes who competed in the 2022 World Championships and another competition 
-(2019 World Championships, 2023 World Championships or 2022 national 
-championships). For the GLMM described in section~\ref{sec:glmm}, data comes
-from every world championship from 1999 to 2023. This paper was started shortly
-after the 2022 World Championships, but as the 2023 data has become available 
-recently, we are pleased to include it in both of our analyses, and it did not
+(2019 World Championships, 2023 World Championships, or 2022 national 
+championships). For the GLMM described in Section~\ref{sec:glmm}, data comes
+from every world championships from 1999 to 2023. This investigation started 
+shortly after the 2022 World Championships, but as the 2023 data has now become 
+available, we are pleased to include it in both of our analyses, and it did not
 significantly change our results.
@@
 centrally located but instead found on different national websites (often 
-recorded in the national language without translation, and hence not used). 
+recorded in the national language without translation). 
 \of{I don't know how to say this}. \eds{is my change correct?} \of{No I was able
@@
 110 meter hurdle and then look at a column for reaction times which are always from
-roughly 0.1 to 0.18 seconds.}
+roughly 0.1 to 0.18 seconds.} \eds{is it correct now?}
 
@@
 and six clusters of size five.  Although 2022 is presented three times in
-figure~\ref{fig:RankScatterplots}, the subset of athletes differs each time.
-There may be athletes who competed in 2019 and 2022 and their national
-championships data was available, but they did not compete in 2023, and thus their
-times show up in the top two graphs.
+Figure~\ref{fig:RankScatterplots}, the subset of athletes differs each time.
+There may be athletes who competed in 2019 and 2022 and had their national
+championships data available, but they did not compete in 2023, and thus their
+times appear only in the top two graphs. \eds{top and bottom graphs?}
+\eds{also, you will likely need to make Fig 1 smaller as the caption runs off
+the page (even before my edits).}
 
@@
   \includegraphics{RankScatterPlots}
-  \caption{The top graph shows the athletes who competed at least once each at
-  the 2019 and 2022 World Championships.  The middle graph shows the athletes
-  who competed at least once each at the 2022 and 2023 World Championships. The
-  bottom graph shows the athletes who competed at least once at their country's
-  2022 national championships and the 2022 World Championship.}
+  \caption{The top graph shows the reaction times of athletes who competed at 
+	least once each at the 2019 and 2022 World Championships. The middle graph 
+	shows the reaction times of athletes who competed at least once each at the 
+	2022 and 2023 World Championships. The bottom graph shows the reaction times 
+	of athletes who competed at least once at their country's
+  2022 national championships and the 2022 World Championships.
+	\eds{check x-axis labels in the bottom plot: last three should be 2022?}}
   \label{fig:RankScatterplots}
@@
 
-Figure~\ref{fig:RankScatterplots} shows the data that is being analyzed
-with the rank-based methods.  In each of the three graphs, athletes who competed
-in the 2022 finals are having their times compared to another event they competed
-at.  It is worth noting that each instance of the 2022 data is different, because
+Figure~\ref{fig:RankScatterplots} shows the data that will be analyzed
+with the rank-based methods (described in Section~\ref{sec:rank} with results in 
+Section~\ref{subsec:Results_Rank}). In each of the three graphs, the reaction 
+times for athletes who competed
+in the 2022 finals \eds{World Championships?} are compared to their reaction
+times at another at which they competed.  
+It is worth noting that each instance of the 2022 data is different, because
 those who competed at both the 2019 and 2022 World Championships will be
@@
 in at least one race at each championship, however it is possible for athletes
-to have competed in one race in 2019 and two races in 2022.  The 2022 World
+to have competed in one race in 2019 and two races in 2022. 
+\eds{did we not already cover the last two sentences in the previous paragraph, 
+or are you trying to say something different here?}  The 2022 World
 Championship Heats and Semifinals both had on average lower reaction times than
 their respective 2019 races, and the 2022 finals were substantially faster. It
-is also worth noting is that the fastest reaction time in the 2022 Finals and 2022
-Semifinals both came from Devon Allen, but the difference between 0.101 and 0.099
+is also worth noting is that the fastest reaction time in the 2022 World
+Championships Finals and Semifinals both came from Devon Allen, but the 
+difference between 0.101 and 0.099
 seconds was the difference in what caused him to be disqualified. 
-\eds{Please reorganize the labels so they are either in one of the following 
-orders:
-2019 Heat, 2019 Semi, 2019 final, 2022 Heat, 2022 Semi, 2022 final  OR
-2019 Heat, 2022 Heat, 2019 Semi, 2022 Semi, 2019 final, 2022 final.  
-(I am not sure which is better, but perhaps the latter?)}
-\of{I re-worked this graph completely, please let me know what you think.  For
-each graph I currently list them in chronological order: so 2022 then 2023 but
-I list 2022 second when comparing it to 2019/rank comparison.  Please let me
-know your thoughts on this and if I should list 2022 and then 2023 for visual
-consistency.}
+%\eds{Please reorganize the labels so they are either in one of the following 
+%orders:
+%2019 Heat, 2019 Semi, 2019 final, 2022 Heat, 2022 Semi, 2022 final  OR
+%2019 Heat, 2022 Heat, 2019 Semi, 2022 Semi, 2019 final, 2022 final.  
+%(I am not sure which is better, but perhaps the latter?)}
+%\of{I re-worked this graph completely, please let me know what you think.  For
+%each graph I currently list them in chronological order: so 2022 then 2023 but
+%I list 2022 second when comparing it to 2019/rank comparison.  Please let me
+%know your thoughts on this and if I should list 2022 and then 2023 for visual
+%consistency.}
+%\eds{I think the graphs look good}
 
@@
 whether Allen's time and disqualification are reasonable. 
+\eds{I would suggest removing this paragraph, particularly if you still include
+the comments about female reaction times in the Concluding Remarks. 
+No analysis has been done yet, nor have the methods been described yet, so it 
+does not make sense in my opinion 
+to refer to significance or another analysis method here.}
 
@@
 babicc2009reaction, panoutsakopoulos2020gender}.
+\eds{you should mention the data source earlier, possible in Section 2 
+before Section 2.1 if it applies to Section 2.1.1, and also the sources of the 
+national data for 2.1.2 (i.e., pointing to the info in Appendix).}
 For the creation of the GLMM model we focus on the reaction times during the 
@@
 
-The data under consideration is summarized using a box plot in 
+The data under consideration is summarized using side-by-side boxplots in 
 Figure~\ref{fig:Boxplot}.  It is clear that the reaction times in the 2022
@@
   \includegraphics{Boxplot}
-  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.
-	\eds{there is only one plot here, not two. If you are keeping only one, the
-	text and caption need to be modified accordingly}}
-  \of{Resolved}
+  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.}
+	%\eds{there is only one plot here, not two. If you are keeping only one, the
+	%text and caption need to be modified accordingly}}
+  %\of{Resolved}
   \label{fig:Boxplot}
@@
 The mean and variance of $S$ under $H_0$ can be derived so that $S$ can be
-standardized to form a statistic $Z$ which follows standard normal distribution
+standardized to form a $Z$ statistic which follows a standard normal distribution
 asymptotically \citep[p.910]{datta2005rank}.
@@
 reliable, so we use 1000 random permutations to simulate the observed test 
-statistic.
-\eds{I do not understand what is meant in the text after the colon in the 
-previous sentence.  Please rephrase.}
-\of{I'm not sure what that was, I removed it now.}
+statistic.  \eds{simulate the null distribution of the test statistic?}
+%\eds{I do not understand what is meant in the text after the colon in the 
+%previous sentence.  Please rephrase.}
+%\of{I'm not sure what that was, I removed it now.}
 This method is available from the \texttt{clusWilcox.test()} function
-with \texttt{method = `ds'} and \texttt{exact = TRUE} from R package
+with \texttt{method = `ds'} (for \underline{D}atta and \underline{S}atten) and 
+\texttt{exact = TRUE} from R package
 \texttt{clusrank} \citep{jiang2020wilcoxon}.
@@
 Based on an exploratory analysis using the GLMM model for the reaction times
-from the World Championships, the final model we chose was a gamma mixed-effect 
-model. As suggested by the log-likelihood and the Akaike Information
+from the World Championships, the we chose a gamma mixed-effect model as our 
+final model. As suggested by the log-likelihood and the Akaike Information
 Criterion (AIC), the normal error model on the log-scale was found to be
 much inferior to the gamma model which better captures the skewness of the
-data. Two random effects were found useful: one captures the venue effect in
-each year and the other captures the heat effect within each year.  The heat
+data. Two random effects were found useful: one capturing the venue effect in
+each year and the other capturing the heat effect within each year.  The heat
 effect is important as it captures the variability in the amount of time
 athletes are on the starting blocks before the gun goes off.  Thus, we can try
-to reduce this variability by accounting for this in the model through the heat
+to reduce this variability by accounting for it in the model through the heat
 effect.
@@
 gamma distribution. 
-\eds{I was talking to my husband about this. Could the heat effect be justified 
-to capture the variability in the amount of time between the ``set'' and the 
-gun going off (assuming this is different from heat to heat)? E.g., quicker 
-gunshots after the ``set'' could lead to slower reaction times (or the reverse), 
-but this effect would be shared across all athletes in the same heat.}
-\of{Yes I agree with this and updated the language in the text.}
+%\eds{I was talking to my husband about this. Could the heat effect be justified 
+%to capture the variability in the amount of time between the ``set'' and the 
+%gun going off (assuming this is different from heat to heat)? E.g., quicker 
+%gunshots after the ``set'' could lead to slower reaction times (or the reverse), 
+%but this effect would be shared across all athletes in the same heat.}
+%\of{Yes I agree with this and updated the language in the text.}
 In the classic generalized linear model (GLM) sense,
@@
 where $g$ is a known link function,
-$\gamma_i$ is normally distributed with mean zero and
+$v_i$ is normally distributed with mean zero and
 variance~$\sigma_v^2$, and $\delta_{i/j}$ is normally distributed with mean
@@
 
-The rank based methods described in Section~\ref{sec:rank}
+The rank-based methods described in Section~\ref{sec:rank} are used
 to determine if the reaction times for athletes who had competed at other 
@@
 
-Table~\ref{tab:Clusrankresults} shows the results generated from four tests in R.
+Table~\ref{tab:Clusrankresults} shows the results generated from four \eds{three?}
+tests in R.
 We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
@@
 %that is preferred.  
-In both cases, the national versus international comparison
+%In both cases, 
+The national versus international comparison
 produced lower p values than the 2019 versus 2022 comparison, with the
@@
 Championships had faster reaction times than when they had competed previously.
+\eds{need to comment on 2022 vs 2023 results}
 
@@
 
+\eds{what happened to the Appendix?}
+
 \bibliographystyle{chicago}

### 2023-09-15 Code/RankComparison.Rmd
Added womens clusrank results to bottom

@@
 ```
+
+```{r}
+women_data <- read.csv("../Data/WomenClusrank2019vs2022.csv")
+colnames(women_data) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_data,
+                method = "ds")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_data,
+                method = "ds", 
+                exact = TRUE, B = 10000)
+```

### 2023-09-15 Code/Plots.Rmd
Had to change code to make graph smaller

@@
 rank_scatter <- grid.arrange(rank1, rank2, rank3, ncol = 1)
-ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 8, width = 7)
+ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 7.5, width = 7)
 ```

### 2023-09-15 Manuscript/manuscript.tex
Fixed ordering of glmm/sign rank and went over Prof Schifano's notes

@@
 other reaction times they considered, but they did not perform any formal 
-statistical analysis. This paper questions the reaction time 
-disqualification barrier, which is currently 0.1 seconds, to determine whether 
-this is a reasonable threshold. We employ a generalized linear mixed model 
+statistical analysis. This paper questions if athletes who competed at the 2022
+World Track and Field Championship had similar times at other competitions and
+if 0.1 seconds is a reasonable disqualification barrier. To do so, we employ a 
+signed-rank test for clustered data to compare reaction times for the same 
+athletes at different competitions and a generalized linear mixed model 
 (GLMM) with random venue (year) effect and a random heat effect within venue 
-in order to model the reaction time data. 
-Additionally, we employ a signed-rank test for clustered data to 
-compare reaction times for the same athletes at different competitions. This 
-matter needs to be addressed because disqualification based on allowable 
-reaction time will continue to be an issue in future world championships.
+in order to model the reaction time data. This matter needs to be addressed 
+because disqualification based on allowable reaction time will continue to be an
+issue in future world championships.
 
@@
 to compete before retiring from running to play football.}}
+\of{I don't think he retired, he was competing in races as recently as this past
+June.}
 
@@
 Track and Field Championships to check for any abnormalities, and we approach
-this in two different ways.  First we model the data, which are
+this in two different ways.  First we compare reaction times for athletes who 
+have competed at additional competitions to the 2022 World Track and Field 
+Championship.  By comparing reaction times from the same athlete across multiple
+competitions we are able to determine if the 2022 reaction times differed 
+significantly from those of other competitions.  Then we model the data, which are
 the reaction times of every non-disqualification since the 1999 World 
 Championships, using a gamma linear mixed-effects model with two random effects: 
-one effect for the year and a second effect for the
-heat within each year.
+one effect for the year and a second effect for the heat within each year.
 Based on the fitted model, we can then estimate the probability of a reaction 
 time being below a certain threshold (e.g., 0.1). 
-Second, we compare reaction times for athletes who have competed
-at additional competitions to the 2022 World Track and Field Championship.  By
-comparing reaction times from the same athlete across multiple competitions we 
-are able to determine if the 2022 reaction times differed significantly from 
-those of other competitions.
 Our investigation into reaction times is important not only because of its
@@
 
-The rest of this paper is organized as follows. Section~\ref{sec:Data} describes 
-how the data were collected and begins to detail the generalized linear mixed-effect 
- model (GLMM) that is further developed in Section~\ref{sec:glmm}. The 
-signed-rank test that analyzes reaction times for the same athletes across
-different competitions is
-also described in Section~\ref{sec:rank}.  Section~\ref{sec:Results}
- provides the results of the two analyses and ends 
-by questioning the reaction time barrier.  Lastly, Section~\ref{sec:concludingremarks}
-discusses the impact and limitations of the paper.
+The rest of this paper is organized as follows. Section~\ref{sec:Data} details 
+how the two types of data were collected, as the signed-rank test as described
+in Section~\ref{sec:rank} uses slightly different data than the generalized 
+linear mixed-effect model (GLMM) that is further developed in 
+Section~\ref{sec:glmm} Section~\ref{sec:Results} provides the results of the two
+analyses and ends by questioning the reaction time barrier.  Lastly, 
+Section~\ref{sec:concludingremarks} discusses the impact and limitations of the
+paper.
 
@@
 110 meter hurdle and then look at a column for reaction times which are always from
-roughly 0.1 to 0.18 seconds.} \eds{is it correct now?}
+roughly 0.1 to 0.18 seconds.} \eds{is it correct now?} \of{Yes}
 
@@
 clusters of size two, three clusters of size three, six clusters of size four,
-and six clusters of size five.  Although 2022 is presented three times in
-Figure~\ref{fig:RankScatterplots}, the subset of athletes differs each time.
-There may be athletes who competed in 2019 and 2022 and had their national
-championships data available, but they did not compete in 2023, and thus their
-times appear only in the top two graphs. \eds{top and bottom graphs?}
+and six clusters of size five.  
+%Although 2022 is presented three times in
+%Figure~\ref{fig:RankScatterplots}, the subset of athletes differs each time.
+%There may be athletes who competed in 2019 and 2022 and had their national
+%championships data available, but they did not compete in 2023, and thus their
+%times appear only in the top two graphs. \eds{top and bottom graphs?}
 \eds{also, you will likely need to make Fig 1 smaller as the caption runs off
 the page (even before my edits).}
+\of{It should be smaller now}
 
@@
   2022 national championships and the 2022 World Championships.
-	\eds{check x-axis labels in the bottom plot: last three should be 2022?}}
+	\eds{check x-axis labels in the bottom plot: last three should be 2022?}
+  \of{I don't think so. Its a comparison of 2022 national vs international so
+  I abbreviated it as Nat. and Inter. is that not ok?}}
   \label{fig:RankScatterplots}
@@
 \eds{did we not already cover the last two sentences in the previous paragraph, 
-or are you trying to say something different here?}  The 2022 World
+or are you trying to say something different here?} \of{yes you are correct. I
+commented out the other one because I think this one flows better.}  The 2022 World
 Championship Heats and Semifinals both had on average lower reaction times than
@@
 to refer to significance or another analysis method here.}
+\of{The reason I added it was because Prof Yan recommended switching the
+order of glmm and rank and talking about the GLMM partly because the women's
+data was not significant. I'm not sure how else to say this or if it needs to
+be said at all.  I agree that it doesn't make a ton of sense to talk about the 
+results for the women's data so early.  Prof Yan, can you please include your
+input? Maybe we keep the rank before glmm but don't bring up women's data until
+the end?}
 
@@
 Data was taken from the World Athletics 
-and covers the men's 110 meter hurdles from 1999 to 2023; the source can be 
-found in the Appendix. 
+and covers the men's 110 meter hurdles from 1999 to 2023.
 Although women's data is available, we chose not to include women's data in the
@@
 national data for 2.1.2 (i.e., pointing to the info in Appendix).}
+\of{I want to emphasize the differences in the data so I wanted to try and
+compartmentalize everything that was different. We chose not to include women's
+data for the GLMM because of cited papers and chose not to include women's
+data for the signed-rank because the results were not highly significant. So if
+I want to talk about that in section 2, I feel like I would have to explicitly
+justify the exclusion of the women's data which then goes back to the
+conversation above. Also Prof Yan recommended I remove the appendix, so now the
+sources are in the Data README.}
 For the creation of the GLMM model we focus on the reaction times during the 
@@
 With our small sample size, the asymptotic normal distribution may not be
-reliable, so we use 1000 random permutations to simulate the observed test 
-statistic.  \eds{simulate the null distribution of the test statistic?}
+reliable, so we use 1000 random permutations to simulate the null distribution
+of the test statistic.  
+\eds{simulate the null distribution of the test statistic?}
+\of{Resolved}
 %\eds{I do not understand what is meant in the text after the colon in the 
@@
 
-Table~\ref{tab:Clusrankresults} shows the results generated from four \eds{three?}
+Table~\ref{tab:Clusrankresults} shows the results generated from six 
+\eds{three?}
+\of{Theres now 6 results in the table, if we remove the asymptotic results then
+it should be 3}
 tests in R.
@@
 produced lower p values than the 2019 versus 2022 comparison, with the
-national vs international \texttt{exact = TRUE} test producing a p value of 
+national vs international permutation test producing a p value of 
 $2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  This
@@
 at reacting than they were a few months prior.  For the 2019 versus 2022
-same athlete comparison with \texttt{exact = TRUE} argument, the p value was
-significant at $\alpha = 0.01$, however when \texttt{exact = FALSE}, the p
-value was higher and only significant at $\alpha = 0.05$. 
+same athlete comparison with permutation argument, the p value was 
+$2.9\cdot10^{-3}$, which is significant at $\alpha = 0.01$.  The result of
+comparing 2022 and 2023 was very similar to the above result, as the p value
+from the permutation test was $3.0\cdot10^{-3}$.  Taken together, these three
+tests help show that athletes competing at the 2022 World Track and Field
+Championships had faster reaction times than when they had competed previously.
+
 \eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
 Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
-Nonetheless, these
-tests help to show that athletes competing at the 2022 World Track and Field
-Championships had faster reaction times than when they had competed previously.
+\of{I am not going to talk about the asymptotic results but will show them in
+table.I changed the exact=TRUE and exact = FALSE text to say permutation and 
+asymptotic}
+these
+
 \eds{need to comment on 2022 vs 2023 results}
+\of{Resolved}
 
@@
 	needs to be referenced within the text. Also, it needs to be clarified 
-	somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}} 
+	somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}
+  \of{Resolved}} 
   \begin{tabular}{c c c c c c}
@@
 \eds{could that be because of even smaller sample sizes?}
+\of{Perhaps? I rank the pemutation test for the women's data and the
+p value was roughly 0.8.}
   The lack of evidence from the women's data makes it 
@@
 \eds{what happened to the Appendix?}
+\of{Prof Yan recommended I remove it}
 

### 2023-09-18 Manuscript/manuscript.tex
finished going through remainder of text; removed resolved

@@
 % {\nolinkurl{owen.fiore@uconn.edu}}\\
-  Elizabeth Schifano\\
+  Elizabeth D. Schifano\\
   Jun Yan\\[1ex]
@@
 % to wait, after all.} 
-\eds{ If it is true he ``retired'' as a runner to play football, we could 
-instead say something like
-\textit{Had Allen reacted just 0.001 seconds slower he would have been allowed 
-to compete before retiring from running to play football.}}
-\of{I don't think he retired, he was competing in races as recently as this past
-June.}
+%\eds{ If it is true he ``retired'' as a runner to play football, we could 
+%instead say something like
+%\textit{Had Allen reacted just 0.001 seconds slower he would have been allowed 
+%to compete before retiring from running to play football.}}
+%\of{I don't think he retired, he was competing in races as recently as this past
+%June.}
+%\eds{Okay, then let's leave it like this.}
 
@@
 linear mixed-effect model (GLMM) that is further developed in 
-Section~\ref{sec:glmm} Section~\ref{sec:Results} provides the results of the two
+Section~\ref{sec:glmm}. Section~\ref{sec:Results} provides the results of the two
 analyses and ends by questioning the reaction time barrier.  Lastly, 
@@
 significantly change our results.
+\eds{I think you should somehow reference the source of the data here; I saw the
+comment below about removing the Appendix and including the sources in the Data 
+README file, so assuming the readers will have access to that file, you can 
+simply refer to the sources within that file here. I will defer to Prof. Yan's
+opinion, however!}
 
@@
 recorded in the national language without translation). 
-\of{I don't know how to say this}. \eds{is my change correct?} \of{No I was able
-to use a lot even though I couldn't actually read the page. I just tried to find
-110 meter hurdle and then look at a column for reaction times which are always from
-roughly 0.1 to 0.18 seconds.} \eds{is it correct now?} \of{Yes}
+%\of{I don't know how to say this}. \eds{is my change correct?} \of{No I was able
+%to use a lot even though I couldn't actually read the page. I just tried to find
+%110 meter hurdle and then look at a column for reaction times which are always from
+%roughly 0.1 to 0.18 seconds.} \eds{is it correct now?} \of{Yes}
 
@@
 decreased from 15 to 9.
-
-%\eds{Can you show some descriptive statistics for this data?  Total number of 
-%clusters? Average number of international vs national competitions across 
-%athletes/clusters?}
-%\eds{The following paragraph still needs to be smoothed (the paragraph that was 
-%here and the paragraph that was in the intro still need to be integrated into 
-%one (or two) cohesive paragraph(s) that accurately describe the two types of 
-%data that will be used for two rank-based analyses.}
-%\of{Does this work?}
-%\eds{Yes, but you still need figure or descriptive statistics for this data.}
 
@@
 %times appear only in the top two graphs. \eds{top and bottom graphs?}
-\eds{also, you will likely need to make Fig 1 smaller as the caption runs off
-the page (even before my edits).}
-\of{It should be smaller now}
+%\eds{also, you will likely need to make Fig 1 smaller as the caption runs off
+%the page (even before my edits).}
+%\of{It should be smaller now}
 
@@
 	of athletes who competed at least once at their country's
-  2022 national championships and the 2022 World Championships.
-	\eds{check x-axis labels in the bottom plot: last three should be 2022?}
-  \of{I don't think so. Its a comparison of 2022 national vs international so
-  I abbreviated it as Nat. and Inter. is that not ok?}}
+  2022 national championships and the 2022 World Championships.}
+	%\eds{check x-axis labels in the bottom plot: last three should be 2022?}
+  %\of{I don't think so. Its a comparison of 2022 national vs international so
+  %I abbreviated it as Nat. and Inter. is that not ok?}\eds{it's good now}
   \label{fig:RankScatterplots}
@@
 to have competed in one race in 2019 and two races in 2022. 
-\eds{did we not already cover the last two sentences in the previous paragraph, 
-or are you trying to say something different here?} \of{yes you are correct. I
-commented out the other one because I think this one flows better.}  The 2022 World
+%\eds{did we not already cover the last two sentences in the previous paragraph, 
+%or are you trying to say something different here?} \of{yes you are correct. I
+%commented out the other one because I think this one flows better.}  
+The 2022 World
 Championship Heats and Semifinals both had on average lower reaction times than
@@
 seconds was the difference in what caused him to be disqualified.
-%\eds{Please reorganize the labels so they are either in one of the following 
-%orders:
-%2019 Heat, 2019 Semi, 2019 final, 2022 Heat, 2022 Semi, 2022 final  OR
-%2019 Heat, 2022 Heat, 2019 Semi, 2022 Semi, 2019 final, 2022 final.  
-%(I am not sure which is better, but perhaps the latter?)}
-%\of{I re-worked this graph completely, please let me know what you think.  For
-%each graph I currently list them in chronological order: so 2022 then 2023 but
-%I list 2022 second when comparing it to 2019/rank comparison.  Please let me
-%know your thoughts on this and if I should list 2022 and then 2023 for visual
-%consistency.}
-%\eds{I think the graphs look good}
 
@@
 sources are in the Data README.}
+\eds{Got it.  See earlier comment in 2.1; I will defer to Prof. Yan.}
 For the creation of the GLMM model we focus on the reaction times during the 
@@
   \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.}
-	%\eds{there is only one plot here, not two. If you are keeping only one, the
-	%text and caption need to be modified accordingly}}
-  %\of{Resolved}
   \label{fig:Boxplot}
@@
 With our small sample size, the asymptotic normal distribution may not be
-reliable, so we use 1000 random permutations to simulate the null distribution
-of the test statistic.  
-\eds{simulate the null distribution of the test statistic?}
-\of{Resolved}
-%\eds{I do not understand what is meant in the text after the colon in the 
-%previous sentence.  Please rephrase.}
-%\of{I'm not sure what that was, I removed it now.}
+reliable, so we also use 1000 random permutations to simulate the null 
+distribution of the test statistic.  
+%\eds{simulate the null distribution of the test statistic?}
+%\of{Resolved}
 This method is available from the \texttt{clusWilcox.test()} function
@@
 $v_i$ is normally distributed with mean zero and
-variance~$\sigma_v^2$, and $\delta_{i/j}$ is normally distributed with mean
+variance~$\sigma_v^2$, and $\delta_{i/j}$ 
+\eds{$h_{i/j}$ or $\delta_{i/j}$? Table 2 uses $\delta_{i/j}$ (are you reporting
+the BLUP for this? If yes, why not also report the BLUP of $v_i$ in Table 2?)}
+is normally distributed with mean
 zero and variance~$\sigma_h^2$.
@@
 
-Table~\ref{tab:Clusrankresults} shows the results generated from six 
-\eds{three?}
-\of{Theres now 6 results in the table, if we remove the asymptotic results then
-it should be 3}
-tests in R.
-We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
-\texttt{B = 1000} indicating that the random permutation size was chosen to be 
-1000, as discussed previously in Section~\ref{sec:rank}.  
+Table~\ref{tab:Clusrankresults} shows the results generated from both the 
+permutation (exact) and asymptotic rank-based tests for the three comparisons, 
+as discussed previously in Section~\ref{sec:rank}.    
+%six \eds{three?}
+%\of{Theres now 6 results in the table, if we remove the asymptotic results then
+%it should be 3}
+%tests in R.
+%We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
+%\texttt{B = 1000} indicating that the random permutation size was chosen to be 
+%1000, as discussed previously in Section~\ref{sec:rank}.  
 %As the sample sizes are small, 65 and 49, we use \texttt{exact = TRUE} as
@@
 national vs international permutation test producing a p value of 
-$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  This
+$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  
+\eds{Prof. Yan, is this possible based on 1000 permutations?} This
 is a highly significant test result and shows that athletes who competed at the
 2022 World Track and Field Championships were on average, significantly faster
-at reacting than they were a few months prior.  For the 2019 versus 2022
-same athlete comparison with permutation argument, the p value was 
+at reacting than they were just a few months prior.  For the 2019 versus 2022
+same athlete comparison with permutation test, the p value was 
 $2.9\cdot10^{-3}$, which is significant at $\alpha = 0.01$.  The result of
@@
 
-\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
-Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
-\of{I am not going to talk about the asymptotic results but will show them in
-table.I changed the exact=TRUE and exact = FALSE text to say permutation and 
-asymptotic}
-these
-
-\eds{need to comment on 2022 vs 2023 results}
-\of{Resolved}
+%\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
+%Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
+%\of{I am not going to talk about the asymptotic results but will show them in
+%table.I changed the exact=TRUE and exact = FALSE text to say permutation and 
+%asymptotic}
+%these
+%
+%\eds{need to comment on 2022 vs 2023 results}
+%\of{Resolved}
 
@@
 for the pooled datasets and shows the standard error of the year
-effect increases when including 2022.  The two effect model with venue and heat
-effect was the best model as evidenced by a decrease in the AIC compared to the
-other models. For the entire data, the AIC for the venue effect only model was
-$-2023.5$, the AIC for the heat effect only model $-2070.7$, and the AIC for the 
-two effect model, which we will present throughout this section, was $-2075.7$. 
-Thus, the two effect model is superior and will be presented throughout the rest
+effect increases when including 2022.
+\eds{this is not what is currently being shown in
+Table~\ref{tab:Gamma_parameters}; also, consider changing ``Full Model'' to 
+``Year and Heat Effects''}  
+The GLMM with both venue
+%two effect model with venue 
+\eds{let's try to be consistent with using either
+`year' or `venue' in the text, and not both, except when you introduce the intuition behind its use-- it doesn't make a difference to 
+me which one you use} 
+and heat random
+effects was the best model as evidenced by a decrease in the AIC compared to the
+other models with only one random effect. %For the entire data \eds{??}, the 
+%AIC for the venue effect only model was
+%$-2023.5$, the AIC for the heat effect only model $-2070.7$, and the AIC for the 
+%two effect model, which we will present throughout this section, was $-2075.7$. 
+Thus, results from the GLMM with both venue and heat random effects %is superior and 
+will be presented throughout the rest
 of the paper.
-\of{How specific do I need to be in this section? Will people who read this paper
-run the code to see the p value or should I say it directly and how do I say it?}
-\eds{If you are doing a likelihood ratio test, I suggest including the p-value 
-and the degrees of freedom associated with the test.}
-\of{Professor Yan recommended not even mentioning the likelihood ratio so
-instead I will just talk about the AIC.}
+%\of{How specific do I need to be in this section? Will people who read this paper
+%run the code to see the p value or should I say it directly and how do I say it?}
+%\eds{If you are doing a likelihood ratio test, I suggest including the p-value 
+%and the degrees of freedom associated with the test.}
+%\of{Professor Yan recommended not even mentioning the likelihood ratio so
+%instead I will just talk about the AIC.}
 
@@
   \caption{Estimated parameters for the two effect GLMM on various data sets. SE
-  is short for standard error. \eds{data in table needs to be revisited and table
-	needs to be referenced within the text. Also, it needs to be clarified 
-	somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}
-  \of{Resolved}} 
+  is short for standard error. %\eds{data in table needs to be revisited and table
+	%needs to be referenced within the text. Also, it needs to be clarified 
+	%somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}
+  %\of{Resolved}
+	\eds{This table is not showing what you have described here or in the main text.}} 
   \begin{tabular}{c c c c c c}
@@
 slightly faster relative to other championships going back to 2003 
-%\eds{check year}. \eds{this seems to contradict the Robert Johnson conclusion in 
-%the intro, unless he only considered sprint and not hurdles?} \of{Johnson's
-%results considered the mens 100m, mens 110m hurdles, womens 100m but not the 
-%womens 100m hurdles.} 
 Likewise, the rank-based comparisons resulted in p values
 that were not significant. 
-\eds{could that be because of even smaller sample sizes?}
-\of{Perhaps? I rank the pemutation test for the women's data and the
-p value was roughly 0.8.}
+%\eds{could that be because of even smaller sample sizes?}
+%\of{Perhaps? I rank the pemutation test for the women's data and the
+%p value was roughly 0.8.}
   The lack of evidence from the women's data makes it 
@@
 although it is unknown if the sensors for the men's 110 meter hurdle and women's
-100 meter hurdle were the same.  
-\eds{is any calibration of the sensors needed (e.g. for weight?) that might
-also play a role for women vs men. (My husband asked me about this as a possible 
-reason for the difference in results for men and women, but he did not know
-if any calibration is actually done.)}
-\of{I honestly am not sure, do you think that this is worth investigating, or
-beyond the scope of this paper? I am not sure if the mechanics of Seiko's
-technology is public knowledge.}
-
-It is confusing as to why the reaction time barrier for men's and women's data
-is the same despite numerous articles such as \citep[e.g.,][]{lipps2011implications, 
-babicc2009reaction, panoutsakopoulos2020gender} stating differently.  World
-Athletics is ignoring differences in gender by setting the reaction time barrier
-to be equal. If 0.1 seconds is considered a fair time for women, then 0.1 seconds
-could not be considered fair for men, as the probability of observing reaction
-times lower than 0.1 would be elevated.  \citet{brosnan2017effects} advocated
-for gender based reaction time barriers which seems logical if there are biological
+100 meter hurdle were the same.  On the other hand, this may be related to 
+another debated question regarding whether the reaction time barrier should be 
+the same for men and women.
+\eds{please check my text addition}  
+%\eds{is any calibration of the sensors needed (e.g. for weight?) that might
+%also play a role for women vs men. (My husband asked me about this as a possible 
+%reason for the difference in results for men and women, but he did not know
+%if any calibration is actually done.)}
+%\of{I honestly am not sure, do you think that this is worth investigating, or
+%beyond the scope of this paper? I am not sure if the mechanics of Seiko's
+%technology is public knowledge.}
+%\eds{let's not open Pandora's box then}
+It is confusing as to why the reaction time barrier for men and women 
+is the same despite numerous articles \citep[e.g.,][]{lipps2011implications, 
+babicc2009reaction, panoutsakopoulos2020gender} stating differently.  
+\eds{The following sentences should be softened a bit, in my opinion, because
+it is not clear (at least to me) that these cited authors are any more correct 
+than those who advocate equal reaction time barriers for men and women. Please
+check my changes.}
+According to these authors, World Athletics is ignoring differences in gender by 
+setting the reaction time barrier to be equal. 
+If there are gender differences and 0.1 seconds is considered a fair time for 
+women, then 0.1 seconds could not be considered fair for men, as the probability 
+of observing reaction times lower than 0.1 would be elevated.
+\citet{brosnan2017effects} advocated for gender based reaction time barriers 
+which seems logical if there are biological
 differences in reaction times between men and women.
@@
 seconds is not as extreme as one may think, and that based on the probabilities
-in table~\ref{tab:Sim_time}, World Athletics should consider changing the 
+in Table~\ref{tab:Sim_time}, World Athletics should consider changing the 
 disqualification barrier to 0.8 seconds.  If they were to do so, they would 
@@
 
-\eds{what happened to the Appendix?}
-\of{Prof Yan recommended I remove it}
+%\eds{what happened to the Appendix?}
+%\of{Prof Yan recommended I remove it}
 

### 2023-09-21 Manuscript/manuscript.tex
Responded to many of Prof Schifanos commets but table 2 needs more work and explanation

@@
 available, we are pleased to include it in both of our analyses, and it did not
-significantly change our results.
+significantly change our results \citep{WAData}.
 \eds{I think you should somehow reference the source of the data here; I saw the
@@
 opinion, however!}
+\of{Is this what you meant?}
 
@@
 Section~\ref{subsec:Results_Rank}). In each of the three graphs, the reaction 
-times for athletes who competed
-in the 2022 finals \eds{World Championships?} are compared to their reaction
+times for athletes who competed at the 2022 World Championships 
+\eds{World Championships?}
+\of{Resolved} are compared to their reaction
 times at another at which they competed.  
@@
 the BLUP for this? If yes, why not also report the BLUP of $v_i$ in Table 2?)}
+\of{I think this should say $h_{i/j}$. In fact I am not sure we ever properly 
+defined what $\delta_{i/j}$ is. The number in the last column of the table is
+what shows up as the residual standard deviation in R}
 is normally distributed with mean
@@
 
-A summary of the two effect model is shown in Table~\ref{tab:Gamma_parameters}
-for the pooled datasets and shows the standard error of the year
-effect increases when including 2022.
+The estimated parameters of the GLMM model from our analysis are shown in 
+Table~\ref{tab:Gamma_parameters}.  These statistics reflect the inclusion of 2022
+in our analysis and shows the three models we considered. $\sigma_h^2$ is the
+standard deviation of the heat effect, $\sigma_v^2$ is the standard deviation of
+the venue effect, and $\alpha$ is the intercept.
 \eds{this is not what is currently being shown in
@@
 ``Year and Heat Effects''}  
+\of{Resolved}
 The GLMM with both venue
 %two effect model with venue 
-\eds{let's try to be consistent with using either
-`year' or `venue' in the text, and not both, except when you introduce the intuition behind its use-- it doesn't make a difference to 
-me which one you use} 
+\eds{let's try to be consistent with using either `year' or `venue' in the text,
+and not both, except when you introduce the intuition behind its use-- it 
+doesn't make a difference to me which one you use} 
+\of{We have been using venue so I will use that, I think I made a typo while
+editing the table.}
 and heat random
@@
       \toprule
-      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\delta_{i/j}$ \\
+      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & Std Dev Residual \\
       \midrule
-      Year Effect Only & $-2023.5$ & N/A & 0.048 & $-1.857$ & 0.152\\
+      Venue Effect Only & $-2023.5$ & N/A & 0.054 & $-1.871$ & 0.150\\
       Heat Effect Only & $-2070.7$ & 0.073 & N/A & $-1.890$ & 0.135\\
-      Full model       & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
+      Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
       \bottomrule

### 2023-09-21 Manuscript/manuscript.tex
polishing; comments on p-values and phi

@@
 when they realized that Allen was disqualified, angry at the result.  Had Allen 
-reacted just 0.001 seconds slower he would have been allowed to compete. %and been 
-%spared waiting until the next World Track and Field Championships.
-%\eds{He is not competing this year, correct? If yes, we may need to rephrase.}
-%\of{Correct}
-%\eds{removed because he did not compete this year, so did not actually have
-% to wait, after all.} 
-%\eds{ If it is true he ``retired'' as a runner to play football, we could 
-%instead say something like
-%\textit{Had Allen reacted just 0.001 seconds slower he would have been allowed 
-%to compete before retiring from running to play football.}}
-%\of{I don't think he retired, he was competing in races as recently as this past
-%June.}
-%\eds{Okay, then let's leave it like this.}
+reacted just 0.001 seconds slower he would have been allowed to compete.
 
@@
 with the timing equipment at the 2022 World Track and Field Championships and
-cited graphs, computed simple statistics such as median reaction times, and 
+cited graphs, computed descriptive statistics such as median reaction times, and 
 compared the reaction times at the 2022 World Track and Field Championships to 
@@
 could have. Additionally,
-%were disqualified. Additionally,
 \citet{komi2009iaaf} recommended that high speed cameras be used to
@@
 competitions we are able to determine if the 2022 reaction times differed 
-significantly from those of other competitions.  Then we model the data, which are
-the reaction times of every non-disqualification since the 1999 World 
+significantly from those of other competitions.  Then we look at a longer time horizon
+and model the reaction times of every non-disqualification since the 1999 World 
 Championships, using a gamma linear mixed-effects model with two random effects: 
-one effect for the year and a second effect for the heat within each year.
+one effect for the year and a second effect for the heat nested within each year.
 Based on the fitted model, we can then estimate the probability of a reaction 
@@
 every sprinter who steps onto the track.
-%\eds{Since the World Championships are happening again now, is data from this 
-%year already available? If it is available, we would need to justify its omission
-%in the current analysis.
-%If it is not available, but becomes available while under review, would we 
-%reanalyze the data?  Does it seem likely our conclusions would change based on
-%what is going on this year?}
-%\of{Resolved}
 
@@
 Section~\ref{sec:concludingremarks} discusses the impact and limitations of the
-paper.
+paper. All the data and code of our analysis are available in a supplement.
 
@@
 significantly change our results \citep{WAData}. 
-\eds{I think you should somehow reference the source of the data here; I saw the
-comment below about removing the Appendix and including the sources in the Data 
-README file, so assuming the readers will have access to that file, you can 
-simply refer to the sources within that file here. I will defer to Prof. Yan's
-opinion, however!}
-\of{Is this what you meant?}
+\jy{I added a sentence to the end of intro.}
 
-\subsection{Comparison Data with 2022 World Championships}\label{sec:databeyond}
+\subsection{Comparison Data with 2022 World Championships}
+\label{sec:databeyond}
+
 The motivation for looking beyond World Championships data came from an 
@@
 recorded in the national language without translation).
-%\of{I don't know how to say this}. \eds{is my change correct?} \of{No I was able
-%to use a lot even though I couldn't actually read the page. I just tried to find
-%110 meter hurdle and then look at a column for reaction times which are always from
-%roughly 0.1 to 0.18 seconds.} \eds{is it correct now?} \of{Yes}
 
@@
 and six clusters of size five.
-%Although 2022 is presented three times in
-%Figure~\ref{fig:RankScatterplots}, the subset of athletes differs each time.
-%There may be athletes who competed in 2019 and 2022 and had their national
-%championships data available, but they did not compete in 2023, and thus their
-%times appear only in the top two graphs. \eds{top and bottom graphs?}
-%\eds{also, you will likely need to make Fig 1 smaller as the caption runs off
-%the page (even before my edits).}
-%\of{It should be smaller now}
-
-\jy{give specifics of your sample size}
-\of{Resolved}
 
@@
         2022 national championships and the 2022 World Championships.}
-	%\eds{check x-axis labels in the bottom plot: last three should be 2022?}
-  %\of{I don't think so. Its a comparison of 2022 national vs international so
-  %I abbreviated it as Nat. and Inter. is that not ok?}\eds{it's good now}
   \label{fig:RankScatterplots}
@@
 times for athletes who competed at the 2022 World Championships
-\eds{World Championships?}
-\of{Resolved} are compared to their reaction
-times at another at which they competed.  
+are compared to their reaction times at another at which they competed.  
 It is worth noting that each instance of the 2022 data is different, because
@@
 to have competed in one race in 2019 and two races in 2022.
-%\eds{did we not already cover the last two sentences in the previous paragraph, 
-%or are you trying to say something different here?} \of{yes you are correct. I
-%commented out the other one because I think this one flows better.}  
 The 2022 World
@@
 
-We also started to perform a similar analysis to compare the times of female 
-athletes who competed at the 2019 and 2022 World Track and Field Championships,
-however as our results were not significant, we decided to incorporate another
-method to investigate the times at the 2022 World Championships and thus we
-employ the GLMM, which looks at historical men's data to try to determine
-whether Allen's time and disqualification are reasonable. 
-\eds{I would suggest removing this paragraph, particularly if you still include
-the comments about female reaction times in the Concluding Remarks. 
-No analysis has been done yet, nor have the methods been described yet, so it 
-does not make sense in my opinion 
-to refer to significance or another analysis method here.}
-\of{The reason I added it was because Prof Yan recommended switching the
-order of glmm and rank and talking about the GLMM partly because the women's
-data was not significant. I'm not sure how else to say this or if it needs to
-be said at all.  I agree that it doesn't make a ton of sense to talk about the 
-results for the women's data so early.  Prof Yan, can you please include your
-input? Maybe we keep the rank before glmm but don't bring up women's data until
-the end?}
-
 \subsection{World Championships 1999--2023}\label{sec:dataworld}
@@
 and covers the men's 110 meter hurdles from 1999 to 2023.
-Although women's data is available, we chose not to include women's data in the
-analyses to follow due to the numerous papers citing differences in men and
-women's reaction times \citep[e.g.,][]{lipps2011implications, 
-babicc2009reaction, panoutsakopoulos2020gender}.
-\eds{you should mention the data source earlier, possible in Section 2 
-before Section 2.1 if it applies to Section 2.1.1, and also the sources of the 
-national data for 2.1.2 (i.e., pointing to the info in Appendix).}
-\of{I want to emphasize the differences in the data so I wanted to try and
-compartmentalize everything that was different. We chose not to include women's
-data for the GLMM because of cited papers and chose not to include women's
-data for the signed-rank because the results were not highly significant. So if
-I want to talk about that in section 2, I feel like I would have to explicitly
-justify the exclusion of the women's data which then goes back to the
-conversation above. Also Prof Yan recommended I remove the appendix, so now the
-sources are in the Data README.}
-\eds{Got it.  See earlier comment in 2.1; I will defer to Prof. Yan.}
-For the creation of the GLMM model we focus on the reaction times during the 
+We focus on the reaction times during the 
 semifinal and final heats only, as reaction times from preliminary heats are 
 often not as fast as those in later heats \citep[e.g.,][]{zhang2021correlation}. 
-
-For analysis purposes, we will often pool reaction times of men's final and 
+For analysis purposes, we will pool reaction times of men's final and 
 men's semifinal heats together which increases the sample size. This is 
@@
 reliable, so we also use 1000 random permutations to simulate the null 
-distribution of the test statistic.  
-%\eds{simulate the null distribution of the test statistic?}
-%\of{Resolved}
+distribution of the test statistic.  \jy{increase 1000 to a much bigger number,
+  say 10 million}
 This method is available from the \texttt{clusWilcox.test()} function
@@
 effect.
+
 Let $Y_{ijk}$ be the reaction time of observation~$k$ in heat~$j$ in year~$i$.
@@
 $h_{i/j}$ nested within each year~$i$, $Y_{ijk}$ is modeled by a
-gamma distribution. 
-%\eds{I was talking to my husband about this. Could the heat effect be justified 
-%to capture the variability in the amount of time between the ``set'' and the 
-%gun going off (assuming this is different from heat to heat)? E.g., quicker 
-%gunshots after the ``set'' could lead to slower reaction times (or the reverse), 
-%but this effect would be shared across all athletes in the same heat.}
-%\of{Yes I agree with this and updated the language in the text.}
-In the classic generalized linear model (GLM) sense,
+gamma distribution. In the classic generalized linear model (GLM) sense,
 $Y_{ijk}$ has conditional mean $E(Y_{ijk} | v_i, h_{i/j}) > 0$ and a
@@
 variance~$\sigma_v^2$, and $\delta_{i/j}$ 
-\eds{$h_{i/j}$ or $\delta_{i/j}$? Table 2 uses $\delta_{i/j}$ (are you reporting
-the BLUP for this? If yes, why not also report the BLUP of $v_i$ in Table 2?)}
-\of{I think this should say $h_{i/j}$. In fact I am not sure we ever properly 
-defined what $\delta_{i/j}$ is. The number in the last column of the table is
-what shows up as the residual standard deviation in R}
 is normally distributed with mean
@@
 has shape $a = 1 / \phi$ and scale
+\jy{Note that, as used in the code, we should have
+  \texttt{sigma(model) \^ 2 =  phi}. So in Table 2, we can report phi in the last column.}
 $b = g^{-1}(\alpha + v_i + h_{i/j}) \phi$. The marginal
@@
 national championships to the 2022 World Championships.
-
-Table~\ref{tab:Clusrankresults} shows the results generated from both the 
-permutation (exact) and asymptotic rank-based tests for the three comparisons, 
-as discussed previously in Section~\ref{sec:rank}.    
-%six \eds{three?}
-%\of{Theres now 6 results in the table, if we remove the asymptotic results then
-%it should be 3}
-%tests in R.
-%We use a \texttt{clusWilcox.test()} with \texttt{exact = TRUE} and 
-%\texttt{B = 1000} indicating that the random permutation size was chosen to be 
-%1000, as discussed previously in Section~\ref{sec:rank}.  
-%As the sample sizes are small, 65 and 49, we use \texttt{exact = TRUE} as
-%that is preferred.  
-%In both cases, 
-The national versus international comparison
-produced lower p values than the 2019 versus 2022 comparison, with the
-national vs international permutation test producing a p value of 
-$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  
-\eds{Prof. Yan, is this possible based on 1000 permutations?} This
-is a highly significant test result and shows that athletes who competed at the
-2022 World Track and Field Championships were on average, significantly faster
-at reacting than they were just a few months prior.  For the 2019 versus 2022
-same athlete comparison with permutation test, the p value was 
-$2.9\cdot10^{-3}$, which is significant at $\alpha = 0.01$.  The result of
-comparing 2022 and 2023 was very similar to the above result, as the p value
-from the permutation test was $3.0\cdot10^{-3}$.  Taken together, these three
-tests help show that athletes competing at the 2022 World Track and Field
-Championships had faster reaction times than when they had competed previously.
-
-%\eds{not sure the \texttt{exact = FALSE} results need to be reported here or in
-%Table 4 when we say we use \texttt{exact = TRUE} here and in the previous section.}
-%\of{I am not going to talk about the asymptotic results but will show them in
-%table.I changed the exact=TRUE and exact = FALSE text to say permutation and 
-%asymptotic}
-%these
-%
-%\eds{need to comment on 2022 vs 2023 results}
-%\of{Resolved}
 
@@
 
+Table~\ref{tab:Clusrankresults} shows the results generated from both the 
+permutation (exact) and asymptotic rank-based tests for the three comparisons, 
+as discussed previously in Section~\ref{sec:rank}.    
+The national versus international comparison
+produced lower p-values than the 2019 versus 2022 comparison, with the
+national vs international permutation test producing a p-value of 
+$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  
+\eds{Prof. Yan, is this possible based on 1000 permutations?}
+\jy{That cannot be true.  Redo with a much bigger number of permutations and
+  update the table.}
+This
+is a highly significant test result and shows that athletes who competed at the
+2022 World Track and Field Championships were on average, significantly faster
+at reacting than they were just a few months prior.  For the 2019 versus 2022
+same athlete comparison with permutation test, the p-value was 
+$2.9\cdot10^{-3}$, which is significant at $\alpha = 0.01$.  The result of
+comparing 2022 and 2023 was very similar to the above result, as the p-value
+from the permutation test was $3.0\cdot10^{-3}$.  Taken together, these three
+tests help show that athletes competing at the 2022 World Track and Field
+Championships had faster reaction times than when they had competed previously.
+
 \subsection{GLMM} \label{subsec:Results_GLMM}
@@
 the venue effect, and $\alpha$ is the intercept.
-\eds{this is not what is currently being shown in
-Table~\ref{tab:Gamma_parameters}; also, consider changing ``Full Model'' to 
-``Year and Heat Effects''}  
-\of{Resolved}
-The GLMM with both venue
-%two effect model with venue 
-\eds{let's try to be consistent with using either `year' or `venue' in the text,
-and not both, except when you introduce the intuition behind its use-- it 
-doesn't make a difference to me which one you use} 
-\of{We have been using venue so I will use that, I think I made a typo while
-editing the table.}
-and heat random
+The GLMM with both venue and heat random
 effects was the best model as evidenced by a decrease in the AIC compared to the
-other models with only one random effect. %For the entire data \eds{??}, the 
-%AIC for the venue effect only model was
-%$-2023.5$, the AIC for the heat effect only model $-2070.7$, and the AIC for the 
-%two effect model, which we will present throughout this section, was $-2075.7$. 
-Thus, results from the GLMM with both venue and heat random effects %is superior and 
+other models with only one random effect. 
+Thus, results from the GLMM with both venue and heat random effects
 will be presented throughout the rest
 of the paper.
-%\of{How specific do I need to be in this section? Will people who read this paper
-%run the code to see the p value or should I say it directly and how do I say it?}
-%\eds{If you are doing a likelihood ratio test, I suggest including the p-value 
-%and the degrees of freedom associated with the test.}
-%\of{Professor Yan recommended not even mentioning the likelihood ratio so
-%instead I will just talk about the AIC.}
 
@@
   \caption{Estimated parameters for the two effect GLMM on various data sets. SE
-  is short for standard error. %\eds{data in table needs to be revisited and table
-	%needs to be referenced within the text. Also, it needs to be clarified 
-	%somewhere that ``Old Pooled'' excludes 2022 and ``All Pooled'' includes 2022.}
-  %\of{Resolved}
-	\eds{This table is not showing what you have described here or in the main text.}} 
+  is short for standard error. 
+	\eds{This table is not showing what you have described here or in the
+          main text.}
+        \jy{report $ a = 1 / \phi$}
+      } 
   \begin{tabular}{c c c c c c}
       \toprule
-      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & Std Dev Residual \\
+      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $1 / \phi$ \\
       \midrule
@@
 
-Table~\ref{tab:Sim_probability} shows that changing the reaction time 
-barrier from 0.1 seconds to 0.08 seconds drastically reduces the chances of 
-observing a reaction time that would break the barrier. If World Athletics were
-to change the reaction time the way that \citet{komi2009iaaf} suggests, the 
-probability would be so much lower that athletes would likely not
-have a case to make when arguing against a disqualification.
-
 \begin{table}
@@
 \end{table}
+
+Table~\ref{tab:Sim_probability} shows that changing the reaction time 
+barrier from 0.1 seconds to 0.08 seconds drastically reduces the chances of 
+observing a reaction time that would break the barrier. If World Athletics were
+to change the reaction time the way that \citet{komi2009iaaf} suggests, the 
+probability would be so much lower that athletes would likely not
+have a case to make when arguing against a disqualification.
 
@@
 
-%One conclusion from the analysis in \citet{zhang2021correlation} was that 
-%athletes' reaction times improved from 2011 to 2019.  However this study does 
-%not consider data from 1999 to 2009, and over this timespan athletes get slower 
-%and then faster.  Figure~\ref{fig:Boxplots} shows that for the men's pooled data,
-%the year with the fastest median reaction time before 2022 was 1999.  Thus, this 
-%reject's Zhang et al.'s conclusion that athletes are getting faster because they 
-%were fast over twenty years ago. One possible explanation for why 2011 had such
-%high variation was because of the IAAF rule discussed earlier that increased the
-%penalty of a false start to an instant disqualification \citep{iaaf2009falsestart}.
-%Athletes may have been overly cautious to false start and may have consciously
-%or sub-consciously reacted slower as a result.  Instead, a more appropriate 
-%conclusion may be that there is a lot of variability in reaction times and that
-%the context of the races matters.
-
 This paper does not consider data from the women's 100 meter hurdle, and although
@@
 women, exploratory analysis showed that 2022 was at best only
-slightly faster relative to other championships going back to 2003 
-Likewise, the rank-based comparisons resulted in p values
-that were not significant. 
-%\eds{could that be because of even smaller sample sizes?}
-%\of{Perhaps? I rank the pemutation test for the women's data and the
-%p value was roughly 0.8.}
+slightly faster relative to other championships going back to 2003.
+Likewise, the rank-based comparisons resulted in p-values
+that were not drastically significant after adjusting for multiple tests.
 The lack of evidence from the women's data makes it 
@@
 the same for men and women.
-\eds{please check my text addition}  
-%\eds{is any calibration of the sensors needed (e.g. for weight?) that might
-%also play a role for women vs men. (My husband asked me about this as a possible 
-%reason for the difference in results for men and women, but he did not know
-%if any calibration is actually done.)}
-%\of{I honestly am not sure, do you think that this is worth investigating, or
-%beyond the scope of this paper? I am not sure if the mechanics of Seiko's
-%technology is public knowledge.}
-%\eds{let's not open Pandora's box then}
+
 It is confusing as to why the reaction time barrier for men and women 
@@
 babicc2009reaction, panoutsakopoulos2020gender} stating differently.  
-\eds{The following sentences should be softened a bit, in my opinion, because
-it is not clear (at least to me) that these cited authors are any more correct 
-than those who advocate equal reaction time barriers for men and women. Please
-check my changes.}
 According to these authors, World Athletics is ignoring differences in gender by 
@@
 times at other competitions. Table~\ref{tab:Clusrankresults} shows multiple 
-significant p values that indicate differences in average reaction times for 
+significant p-values that indicate differences in average reaction times for 
 athletes who competed at the 2022 World Track and Field Championship and another
@@
 seconds is not as extreme as one may think, and that based on the probabilities
-in Table~\ref{tab:Sim_time}, World Athletics should consider changing the 
+in Table~\ref{tab:Sim_time}, World Athletics may consider changing the 
 disqualification barrier to 0.8 seconds.  If they were to do so, they would 
@@
 
-%\eds{what happened to the Appendix?}
-%\of{Prof Yan recommended I remove it}
+\section*{Supplementary Material}
+The data and R code used for the analysis are available in a compressed file for
+ease of reproducibility.
 

### 2023-09-21 Code/Plots.Rmd
Added venue effect dot plot

--- 2023-09-13 Code/Plots.Rmd
@@
 rank_scatter <- grid.arrange(rank1, rank2, rank3, ncol = 1)
-ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 8, width = 7)
+ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 7.5, width = 7)
 ```
+```{r}
+library(lme4)
+library(ggplot2)
+library(dplyr)
+times <- read.csv("../Data/Thesis2023.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
+with2022 <- subset(times, (Type == "F" | Type == "S") & 
+                       Gender == "M" & Time != 'DQ')
+with2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
+                               data = with2022, family = Gamma(link = "log"))
+ranef_results <- data.frame(ranef(with2022_two_effect_gamma_log))
+venues <- subset(ranef_results, ranef_results["grpvar"] == "Venue")
+venues <- venues %>% arrange("Venue")
+
+ggplot(venues, aes(x=grp, y = condval)) + geom_point() + 
+  labs(x = "Year", y = "Year Effect")
+```

### 2023-09-26 Code/RankComparison.Rmd
Changed to a million permutations
Re-ran numbers for B = 1 million

--- 2023-09-06 Code/RankComparison.Rmd
@@
 ```{r}
+set.seed(1)
 library(clusrank)
@@
                 method = "ds",
-                exact = TRUE, B = 10000)
+                exact = TRUE, B = 1000000)
 ```
@@
 ```{r}
-comp <- read.csv("../Data/Clusrank2019vs2022.csv")
+old_clusrank <- read.csv("../Data/Clusrank2019vs2022.csv")
 #"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
-colnames(comp) <- c("Time", "Athlete", "Year")
+colnames(old_clusrank) <- c("Time", "Athlete", "Year")
 #View(comp)
@@
 ```{r}
+set.seed(1)
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_clusrank,
                 method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_clusrank,
                 method = "ds",
-                exact = TRUE, B = 10000)
+                exact = TRUE, B = 1000000)
 ```
@@
 ```
+
+```{r}
+women_data <- read.csv("../Data/WomenClusrank2019vs2022.csv")
+colnames(women_data) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_data,
+                method = "ds")
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_data,
+                method = "ds", 
+                exact = TRUE, B = 1000000)
+```

### 2023-09-26 Manuscript/manuscript.tex
Updated table

@@
    \midrule
-   2019 vs 2022 & $2.9\cdot10^{-3}$ & $2.3\cdot10^{-2}$ \\
-   2022 vs 2023 & $3.0\cdot10^{-3}$ & $1.0\cdot10^{-2}$ \\
-   2022 National vs International & $2.2\cdot10^{-16}$ & $1.0\cdot10^{-3}$ \\
+   2019 vs 2022 & $2.8\cdot10^{-2}$ & $2.9\cdot10^{-3}$ \\
+   2022 vs 2023 & $1.1\cdot10^{-2}$ & $3.1\cdot10^{-3}$ \\
+   2022 National vs International & $2\cdot10^{-6}$ & $1.1\cdot10^{-3}$ \\
    \bottomrule
@@
   update the table.}
+\of{Done, permutation results were for B = 1 million}
 This

### 2023-10-03 Manuscript/manuscript.tex
Changed some language, still need to fix some things

@@
 significantly change our results \citep{WAData}.
-\jy{I added a sentence to the end of intro.}
 
@@
 reliable, so we also use 1000 random permutations to simulate the null 
-distribution of the test statistic.  \jy{increase 1000 to a much bigger number,
-  say 10 million}
+distribution of the test statistic.
 This method is available from the \texttt{clusWilcox.test()} function
@@
 $v_i$ is normally distributed with mean zero and
-variance~$\sigma_v^2$, and $\delta_{i/j}$ 
+variance~$\sigma_v^2$, and $h_{i/j}$ 
 is normally distributed with mean
@@
 permutation (exact) and asymptotic rank-based tests for the three comparisons, 
-as discussed previously in Section~\ref{sec:rank}.    
-The national versus international comparison
-produced lower p-values than the 2019 versus 2022 comparison, with the
-national vs international permutation test producing a p-value of 
-$2.2\cdot10^{-16}$, which is the smallest possible value achievable in R.  
-\eds{Prof. Yan, is this possible based on 1000 permutations?}
-\jy{That cannot be true.  Redo with a much bigger number of permutations and
-  update the table.}
-\of{Done, permutation results were for B = 1 million}
-This
-is a highly significant test result and shows that athletes who competed at the
+as discussed previously in Section~\ref{sec:rank}. The national versus international
+comparison produced lower p-values than the 2019 versus 2022 comparison, with the
+national vs international permutation test producing a p-value of $2\cdot10^{-6}$.
+This is a highly significant test result and shows that athletes who competed at the
 2022 World Track and Field Championships were on average, significantly faster
@@
 standard deviation of the heat effect, $\sigma_v^2$ is the standard deviation of
-the venue effect, and $\alpha$ is the intercept.
+the venue effect, $\alpha$ is the intercept, and $a$ is the shape parameter of
+the gamma function.
 The GLMM with both venue and heat random
@@
       \toprule
-      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $1 / \phi$ \\
+      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $a$ \\
       \midrule

### 2023-10-04 Manuscript/manuscript.tex
edits/comments

@@
 athletes at different competitions and a generalized linear mixed model 
-(GLMM) with random venue (year) effect and a random heat effect within venue 
+with random venue (year) effect and a random heat effect within venue 
 in order to model the reaction time data. This matter needs to be addressed 
@@
 With our small sample size, the asymptotic normal distribution may not be
-reliable, so we also use 1000 random permutations to simulate the null 
+reliable, so we also use 1~million random permutations to simulate the null 
 distribution of the test statistic.
@@
 has shape $a = 1 / \phi$ and scale
-\jy{Note that, as used in the code, we should have
-  \texttt{sigma(model) \^ 2 =  phi}. So in Table 2, we can report phi in the last column.}
 $b = g^{-1}(\alpha + v_i + h_{i/j}) \phi$. The marginal
@@
 Table~\ref{tab:Clusrankresults} shows the results generated from both the 
-permutation (exact) and asymptotic rank-based tests for the three comparisons, 
+permutation and asymptotic rank-based tests for the three comparisons, 
 as discussed previously in Section~\ref{sec:rank}. The national versus international
@@
 in our analysis and shows the three models we considered. $\sigma_h^2$ is the
+\jy{read what you wrote. These are variances not sd. Do not start a sentence
+  with a symbol (see my writing guidelines).}
 standard deviation of the heat effect, $\sigma_v^2$ is the standard deviation of
-the venue effect, $\alpha$ is the intercept, and $a$ is the shape parameter of
-the gamma function.
+the venue effect, $\alpha$ is the intercept, and $\phi$ is the dispersion
+parameter of the gamma model.
 The GLMM with both venue and heat random
@@
 
+\jy{do a thorough check on the captions of tables and figures.}
+
 \begin{table}
   \centering
-  \caption{Estimated parameters for the two effect GLMM on various data sets. SE
-  is short for standard error. 
-	\eds{This table is not showing what you have described here or in the
-          main text.}
-        \jy{report $ a = 1 / \phi$}
-      } 
+  \caption{AIC and estimated parameters for the three GLMMs fitted to the ???
+    data.}
+  \label{tab:Gamma_parameters}
   \begin{tabular}{c c c c c c}
       \toprule
-      Dataset & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $a$ \\
+      Model & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\sqrt{\phi}$ \\
       \midrule
-      Venue Effect Only & $-2023.5$ & N/A & 0.054 & $-1.871$ & 0.150\\
-      Heat Effect Only & $-2070.7$ & 0.073 & N/A & $-1.890$ & 0.135\\
+      Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.150\\
+      Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
       Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
@@
   \end{tabular}
-  \label{tab:Gamma_parameters}
 \end{table}
@@
   \centering
-  \caption{Probabilities of observing reaction times less than 0.08, 0.09, and
-  0.01 seconds based on the two effect model.}
+  \caption{Probabilities of observing reaction times less than threshold 0.08, 0.09, and
+  0.10 seconds based on the two effect model.}
   \begin{tabular}{c c c c} 
    \toprule
-   Data Set & P(R. Time $<$ 0.08) & P(R. Time $<$ 0.09) & P(R. Time $<$ 0.1)  \\ 
+   Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\ 
    \midrule
@@
 have a case to make when arguing against a disqualification.
+\jy{discuss the differences between with/without 2022 data; comment on the
+  pecularity of the 2022 data (as suggested by the rank based method); In a
+  longer run, it may not be as strange as it looks at the first time.}
+
+\begin{table}
+  \centering
+  \caption{Suggested reaction time barriers based on tail probabilities.}
+  \begin{tabular}{c c c} 
+   \toprule
+   Data Set & Tail probability  $10^{-3}$ & Tail probability $10^{-4}$ \\ 
+   \midrule
+   Excluding 2022 & $0.094$ & $0.084$ \\ 
+   Including 2022 & $0.092$ & $0.083$ \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:Sim_time}
+\end{table}
+
+\jy{be consistent in using scienfit format of numbers; 0.0001 vs $10^{-4}$}
 
@@
 based on probability.
-
-\begin{table}
-  \centering
-  \caption{Suggested reaction time barriers based on probability of observing
-  a time less than the barrier. \of{Not sure if this makes sense}}
-  \begin{tabular}{c c c} 
-   \toprule
-   Data Set & P(RxnTime $< x) = 10^{-3}$ & P(RxnTime $< x) = 10^{-4}$ \\ 
-   \midrule
-   Excluding 2022 & $0.094$ & $0.084$ \\ 
-   Including 2022 & $0.092$ & $0.083$ \\
-   \bottomrule
-  \end{tabular}
-  \label{tab:Sim_time}
-\end{table}
 
@@
 \bibliographystyle{chicago}
-\bibliography{citations.bib}
+\bibliography{citations}
 

### 2023-10-10 Manuscript/manuscript.tex
responded to Prof Yan's comments:

@@
   from the same athletes who competed at the 2022 and 2023 World Track and Field
-  Championships. Nat vs Inter compares data from the same athletes who competed
-  at the 2022 national-level championships and the 2022 World Track and Field
-  Championships.}
+  Championships. National vs International compares data from the same athletes
+  who competed at 2022 national-level championships and the 2022 World Track and
+  Field Championships.}
   \begin{tabular}{c c c c c} 
@@
 Table~\ref{tab:Gamma_parameters}. These statistics reflect the inclusion of 2022
-in our analysis and shows the three models we considered. $\sigma_h^2$ is the
+in our analysis and shows the three models we considered. The standard deviation
+of the heat effect is $\sigma_h^2$, the standard deviation of the venue effect
+is $\sigma_v^2$, the intercept is $\alpha$ and we report $\sqrt{\phi}$ as the
+standard deviation of the dispersion parameter $\phi$
 \jy{read what you wrote. These are variances not sd. Do not start a sentence
   with a symbol (see my writing guidelines).}
-standard deviation of the heat effect, $\sigma_v^2$ is the standard deviation of
-the venue effect, $\alpha$ is the intercept, and $\phi$ is the dispersion
-parameter of the gamma model.
+\of{Resolved}
 The GLMM with both venue and heat random
@@
   \centering
-  \caption{AIC and estimated parameters for the three GLMMs fitted to the ???
+  \caption{AIC and estimated parameters for the three models fitted to all the
     data.}
@@
 chosen to estimate the probability of observing a reaction time below 
-two thresholds: 0.1 and 0.08. When including 2022, a probability
-of $4.34\cdot10^{-3}$ corresponds to roughly 1 in every 230 starts
-being below 0.1 seconds.  When we exclude 2022, the probability of observing a
-reaction time below 0.1 seconds decreased to $3.54\cdot10^{-3}$, or 1 in every
-283 starts.
+two thresholds: 0.1 and 0.08.
 
@@
 probability would be so much lower that athletes would likely not
-have a case to make when arguing against a disqualification.
+have a case to make when arguing against a disqualification. When including 2022,
+a probability of $4.34\cdot10^{-3}$ corresponds to roughly 1 in every 230 starts
+being below 0.1 seconds. When we exclude 2022, the probability of observing a
+reaction time below 0.1 seconds decreased to $3.54\cdot10^{-3}$, or 1 in every
+283 starts.  While this seems low, it is not unreasonable to think that certain
+athletes with strong reaction times could begin approach the 0.1 second barrier,
+especially when considering Allen's other reaction times: 0.124 and 0.101 at the
+2022 World Track and Field Championships.  Another important conclusion from our
+results is how excluding 2022 lowered the probability of observing a fast 
+reaction time, suggesting that 2022 was peculiarly fast, which aligns with the
+results presented in section~\ref{subsec:Results_Rank}.
+
 \jy{discuss the differences between with/without 2022 data; comment on the
@@
   longer run, it may not be as strange as it looks at the first time.}
+\of{Resolved}
 
@@
 \jy{be consistent in using scienfit format of numbers; 0.0001 vs $10^{-4}$}
+\of{In our meeting you said to convert scientific notation like $10^{-2}$ to 
+0.01, and I agree I think 0.01 makes much more sense. So should I convert
+everything to decimals or scientific notation?}
 
@@
 reaction time barrier given the probability of observing a time less than the 
-barrier.  When considering all of the data, if we want a 0.1\% chance of 
+barrier. When considering all the data, if we want a 0.1\% chance of 
 observing an illegally fast reaction time, the reaction

### 2023-10-11 Code/RankComparison.Rmd
Fixed name of file

@@
 ```{r}
-nations_noprelim <- read.csv("../Data/Clusrank2019vs2022PrelimsOnly.csv")
+nations_noprelim <- read.csv("../Data/Clusrank2019vs2022NoPrelims.csv")
 colnames(nations_noprelim) <- c("Time", "Athlete", "Year")
@@
 ```
+```{r}
+new_2023 <- read.csv("../Data/ClusrankNatVsWorld2023.csv")
+colnames(new_2023) <- c("Time", 'Athlete', "Year")
+```
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = new_2023,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = new_2023,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```

### 2023-10-11 Code/Plots.Rmd
Added double year effect plot

--- 2023-09-15 Code/Plots.Rmd
@@
 times <- read.csv("../Data/Thesis2023.csv")
-colnames(times) <- c("Year", "Type", "Time", "RxnTime", "Gender", "Batch")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
 ```
@@
 library(ggplot2)
-ggplot(with2022, aes(x = Year, y = RxnTime, group = Year)) +
+ggplot(with2022, aes(x = Venue, y = RxnTime, group = Venue)) +
   geom_boxplot() +
@@
 ```
+
+```{r}
+library(lme4)
+library(ggplot2)
+library(dplyr)
+times <- read.csv("../Data/Thesis2023.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
+with2022 <- subset(times, (Type == "F" | Type == "S") & 
+                       Gender == "M" & Time != 'DQ')
+with2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
+                               data = with2022, family = Gamma(link = "log"))
+ranef_results <- data.frame(ranef(with2022_two_effect_gamma_log))
+venues <- subset(ranef_results, ranef_results["grpvar"] == "Venue")
+venues$year <- as.character(venues$grp)
+venues$year <- as.numeric(venues$year)
+
+venue_effects_with2022<- ggplot(venues, aes(x=year, y = condval)) + geom_point() + 
+  labs(x = "Year", y = "Year Effect") +
+  scale_x_continuous(limits = c(1998,2024), breaks = seq(1999, 2024, 2)) +
+  scale_y_continuous(limits = c(-0.1, 0.1))
+```
+
+```{r}
+library(lme4)
+library(ggplot2)
+library(dplyr)
+times <- read.csv("../Data/Thesis2023.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
+without2022 <- subset(times, (Type == "F" | Type == "S") & 
+                           Gender == "M" & Time != 'DQ' & Venue != 2022)
+without2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
+                               data = without2022, family = Gamma(link = "log"))
+ranef_results <- data.frame(ranef(without2022_two_effect_gamma_log))
+venues <- subset(ranef_results, ranef_results["grpvar"] == "Venue")
+venues$year <- as.character(venues$grp)
+venues$year <- as.numeric(venues$year)
+
+venue_effects_without2022 <- ggplot(venues, aes(x=year, y = condval)) +
+  geom_point() + 
+  labs(x = "Year", y = "Year Effect") +
+  scale_x_continuous(limits = c(1998,2024), breaks = seq(1999, 2024, 2)) +
+  scale_y_continuous(limits = c(-0.1, 0.1))
+```
+
+```{r}
+library(gridExtra)
+venue_effects_combined <- grid.arrange(venue_effects_with2022,
+                                       venue_effects_without2022, ncol = 1)
+ggsave(filename = "ComparisonOfVenueEffects.pdf", path = "../Manuscript",
+       plot=venue_effects_combined, height = 6.5, width = 7)
+```

### 2023-10-11 Manuscript/manuscript.tex
Added year effect plots and fixed std dev typos

@@
    \midrule
-   2019 vs 2022 & $2.8\cdot10^{-2}$ & $2.9\cdot10^{-3}$ \\
-   2022 vs 2023 & $1.1\cdot10^{-2}$ & $3.1\cdot10^{-3}$ \\
-   2022 National vs International & $2\cdot10^{-6}$ & $1.1\cdot10^{-3}$ \\
+   2019 vs 2022 & $0.028$ & $0.003$ \\
+   2022 vs 2023 & $0.011$ & $0.003$ \\
+   2022 National vs International & $2\cdot10^{-6}$ & $0.001$ \\
    \bottomrule
@@
 same athlete comparison with permutation test, the p-value was 
-$2.9\cdot10^{-3}$, which is significant at $\alpha = 0.01$. The result of
-comparing 2022 and 2023 was very similar to the above result, as the p-value
-from the permutation test was $3.0\cdot10^{-3}$. Taken together, these three
+$0.003$, which is significant at $\alpha = 0.01$. The result of
+comparing 2022 and 2023 was very similar (but not exactly the same) to the above
+result, as the p-value from the permutation test was $0.003$. Taken together, these three
 tests help show that athletes competing at the 2022 World Track and Field
@@
 \subsection{GLMM} \label{subsec:Results_GLMM}
-
-The estimated parameters of the GLMM model from our analysis are shown in 
-Table~\ref{tab:Gamma_parameters}. These statistics reflect the inclusion of 2022
-in our analysis and shows the three models we considered. The standard deviation
-of the heat effect is $\sigma_h^2$, the standard deviation of the venue effect
-is $\sigma_v^2$, the intercept is $\alpha$ and we report $\sqrt{\phi}$ as the
-standard deviation of the dispersion parameter $\phi$
-\jy{read what you wrote. These are variances not sd. Do not start a sentence
-  with a symbol (see my writing guidelines).}
-\of{Resolved}
-The GLMM with both venue and heat random
-effects was the best model as evidenced by a decrease in the AIC compared to the
-other models with only one random effect. 
-Thus, results from the GLMM with both venue and heat random effects
-will be presented throughout the rest
-of the paper.
-
-\jy{do a thorough check on the captions of tables and figures.}
 
@@
       \toprule
-      Model & AIC & $\sigma_h^2$ & $\sigma_v^2$ & $\alpha$ & $\sqrt{\phi}$ \\
+      Model & AIC & $\sigma_h$ & $\sigma_v$ & $\alpha$ & $\sqrt{\phi}$ \\
       \midrule
@@
 
-Two aims of this section are to determine how extreme a reaction time below
-0.1 seconds is, and to determine how the exclusion of 2022 could alter the results.
-Using the gamma function with link set to log, the probability of the reaction 
-time being less than 0.1 seconds can be calculated in R using the methods 
-described in Section~\ref{sec:glmm}. A simulation size of $10,000,000$ was 
-chosen to estimate the probability of observing a reaction time below 
-two thresholds: 0.1 and 0.08.
+The estimated parameters of the GLMM model from our analysis are shown in 
+Table~\ref{tab:Gamma_parameters}. These statistics reflect the inclusion of 2022
+in our analysis and shows the three models we considered. The standard deviation
+of the heat effect is $\sigma_h$, the standard deviation of the venue effect
+is $\sigma_v$, the intercept is $\alpha$ and we report $\sqrt{\phi}$ as the
+square root of the dispersion parameter $\phi$
+\jy{read what you wrote. These are variances not sd. Do not start a sentence
+  with a symbol (see my writing guidelines).}
+\of{Actually Resolved}
+The GLMM with both venue and heat random effects was the best model as evidenced
+by a decrease in the AIC compared to the other models with only one random effect. 
+Thus, results from the GLMM with both venue and heat random effects will be 
+presented throughout the rest of the paper.
+
+\jy{do a thorough check on the captions of tables and figures.}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics{ComparisonOfVenueEffects}
+  \caption{The venue effects from 1999 to 2023 for the data including 2022 (top)
+  and excluding 2022 (bottom)}
+  \label{fig:VenueEffects}
+\end{figure}
+
+Figure~\ref{fig:VenueEffects} shows the venue effects from 1999 to 2023 based
+on the full model both with and without 2022.  As this graph does not show heat
+effects, the estimates may not align exactly with the box plot presented in
+figure~\ref{fig:Boxplot}.  The venue effect for 2022 is lower than any other
+effect, confirming that 2022 was a low outlier.  Two aims of this section are to
+determine how extreme a reaction time below 0.1 seconds is, and although the
+venue effects did not differ significantly, to determine how the exclusion of 
+2022 could alter the results. Using the gamma function with \texttt{link = log},
+the probability of the reaction time being less than 0.1 seconds can be 
+calculated in R using the methods described in Section~\ref{sec:glmm}. A 
+simulation size of $10,000,000$ was chosen to estimate the probability of 
+observing a reaction time below two thresholds: 0.1 and 0.08.
 
@@
 \jy{be consistent in using scienfit format of numbers; 0.0001 vs $10^{-4}$}
-\of{In our meeting you said to convert scientific notation like $10^{-2}$ to 
-0.01, and I agree I think 0.01 makes much more sense. So should I convert
-everything to decimals or scientific notation?}
+\of{I think we should be good now, let me know if there is anything else}
 

### 2023-10-12 Code/Plots.Rmd
Fixed typo on axix

@@
 venue_effects_with2022<- ggplot(venues, aes(x=year, y = condval)) + geom_point() + 
-  labs(x = "Year", y = "Year Effect") +
+  labs(x = "Venue", y = "Venue Effect") +
   scale_x_continuous(limits = c(1998,2024), breaks = seq(1999, 2024, 2)) +
@@
   geom_point() + 
-  labs(x = "Year", y = "Year Effect") +
+  labs(x = "Venue", y = "Venue Effect") +
   scale_x_continuous(limits = c(1998,2024), breaks = seq(1999, 2024, 2)) +

### 2023-10-12 Manuscript/manuscript.tex
comments on glmm analysis

@@
 
+\jy{Add the results without 2022 to this table.}
 \begin{table}
@@
 is $\sigma_v$, the intercept is $\alpha$ and we report $\sqrt{\phi}$ as the
-square root of the dispersion parameter $\phi$
-\jy{read what you wrote. These are variances not sd. Do not start a sentence
-  with a symbol (see my writing guidelines).}
-\of{Actually Resolved}
+square root of the dispersion parameter $\phi$.
 The GLMM with both venue and heat random effects was the best model as evidenced
-by a decrease in the AIC compared to the other models with only one random effect. 
+by a decrease in the AIC compared to the other models with only one random
+effect.
+\jy{discuss the differences in the parameter estimates with and without 2022
+  data, especially the sd of the random effect of venue.}
+
+Figure~\ref{fig:VenueEffects} shows the venue effects from 1999 to 2023 based
+on the full model both with and without 2022.  As this graph does not show heat
+\jy{discuss the years the venue effects are high and low}
+effects, the estimates may not align exactly with the box plot presented in
+figure~\ref{fig:Boxplot}.  The venue effect for 2022 is lower than any other
+effect, confirming that 2022 was a low outlier. \jy{but may not be as extreme as
+  one would think considering the whole history.}
+
+\jy{Hopefully, the full model is preferred for both with and without 2022}
+
 Thus, results from the GLMM with both venue and heat random effects will be 
 presented throughout the rest of the paper.
-
-\jy{do a thorough check on the captions of tables and figures.}
 
@@
 
-Figure~\ref{fig:VenueEffects} shows the venue effects from 1999 to 2023 based
-on the full model both with and without 2022.  As this graph does not show heat
-effects, the estimates may not align exactly with the box plot presented in
-figure~\ref{fig:Boxplot}.  The venue effect for 2022 is lower than any other
-effect, confirming that 2022 was a low outlier.  Two aims of this section are to
+Two aims of this section are to
 determine how extreme a reaction time below 0.1 seconds is, and although the
 venue effects did not differ significantly, to determine how the exclusion of 
-2022 could alter the results. Using the gamma function with \texttt{link = log},
+2022 could alter the results. Using the gamma family with \texttt{link = log},
 the probability of the reaction time being less than 0.1 seconds can be 
 calculated in R using the methods described in Section~\ref{sec:glmm}. A 
-simulation size of $10,000,000$ was chosen to estimate the probability of 
+simulation size of $10,000,000$ was used to estimate the probability of 
 observing a reaction time below two thresholds: 0.1 and 0.08.
@@
 
-\jy{discuss the differences between with/without 2022 data; comment on the
-  pecularity of the 2022 data (as suggested by the rank based method); In a
-  longer run, it may not be as strange as it looks at the first time.}
-\of{Resolved}
-
 \begin{table}
@@
 \end{table}
-
-\jy{be consistent in using scienfit format of numbers; 0.0001 vs $10^{-4}$}
-\of{I think we should be good now, let me know if there is anything else}
 

### 2023-10-12 Manuscript/manuscript.tex
Aligned something
Going over Prof Yan's comments

@@
 $j = 1, \ldots, m_i$ where $m_i$ is the number of observations from
-athelete~$i$. Let $\delta_{ij}$ be the group indicator of $X_{ij}$; $\delta_{ij}
+athlete~$i$. Let $\delta_{ij}$ be the group indicator of $X_{ij}$; $\delta_{ij}
 = 1$ if $X_{ij}$ is in group~1 (2022 World Championships) and $\delta_{ij} = 0$ 
@@
 Based on an exploratory analysis using the GLMM model for the reaction times
-from the World Championships, the we chose a gamma mixed-effect model as our 
+from the World Championships, we chose a gamma mixed-effect model as our 
 final model. As suggested by the log-likelihood and the Akaike Information
@@
 \jy{Add the results without 2022 to this table.}
+\of{Does it make sense to order the table how I did, and block by dataset or
+should I block by model?}
 \begin{table}
@@
       \midrule
-      Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.150\\
-      Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
-      Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
+      With 2022 Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.150\\
+      With 2022 Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
+      With 2022 Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
+      \midrule
+      Without 2022 Venue Effect Only & $-1862.2$ &   & 0.048 & $-1.857$ & 0.152\\
+      Without 2022 Heat Effect Only & $-1910.9$ & 0.069 &  & $-1.877$ & 0.136\\
+      Without 2022 Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.880$ & 0.136\\
       \bottomrule
@@
 by a decrease in the AIC compared to the other models with only one random
-effect.
+effect.  When we exclude 2022 from the full model, we observe small but intriguing
+discrepancies in the models. The standard deviation of the venue effect, 
+$\sigma_v$, decreases from 0.40 to 0.33 when 2022 is excluded.  Thus, 2022
+increased the noise in the overall data and confirms that it is an outlier.
+
 \jy{discuss the differences in the parameter estimates with and without 2022
@@
   \centering
-  \caption{Probabilities of observing reaction times less than threshold 0.08, 0.09, and
-  0.10 seconds based on the two effect model.}
+  \caption{Probabilities of observing reaction times less than threshold 0.08,
+  0.09, and 0.10 seconds based on the two effect model.}
   \begin{tabular}{c c c c} 

### 2023-10-12 Manuscript/manuscript.tex
Left some comments, may need to go through again

@@
 
+\jy{Hopefully, the full model is preferred for both with and without 2022}
+\of{It is but the AIC are more similar without 2022}
+
 The estimated parameters of the GLMM model from our analysis are shown in 
@@
 $\sigma_v$, decreases from 0.40 to 0.33 when 2022 is excluded.  Thus, 2022
-increased the noise in the overall data and confirms that it is an outlier.
+increased the noise in the overall data, but many of the other statistics
+remained relatively similar, such as the standard deviation of the heat effect
+and the dispersion parameter.
+
+%\of{Can I just refer to it as the dispersion parameter or should I say "square
+%root of the dispersion parameter" Mathematically it shouldn't make a difference 
+%right as 0.134^2 and 0.136^2 are very similar and I am applying a function to 
+%both? Also should I be saying "standard deviation of heat effect", or sigma_v 
+%or both?}
 
@@
 figure~\ref{fig:Boxplot}.  The venue effect for 2022 is lower than any other
-effect, confirming that 2022 was a low outlier. \jy{but may not be as extreme as
-  one would think considering the whole history.}
-
-\jy{Hopefully, the full model is preferred for both with and without 2022}
+effect, but may not be as extreme as one would think considering the history
+of reaction times going back to 1999.
 

### 2023-10-12 Manuscript/manuscript.tex
reorged table 2

@@
   \centering
-  \caption{AIC and estimated parameters for the three models fitted to all the
+  \caption{AIC and estimated parameters for the three models fitted to the
     data.}
   \label{tab:Gamma_parameters}
-  \begin{tabular}{c c c c c c}
+  \begin{tabular}{c c c c c c c}
     \toprule
-      Model & AIC & $\sigma_h$ & $\sigma_v$ & $\alpha$ & $\sqrt{\phi}$ \\
+    Data set & Model & AIC & $\sigma_h$ & $\sigma_v$ & $\alpha$ & $\sqrt{\phi}$ \\
     \midrule
-      With 2022 Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.150\\
-      With 2022 Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
-      With 2022 Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
-      \midrule
-      Without 2022 Venue Effect Only & $-1862.2$ &   & 0.048 & $-1.857$ & 0.152\\
-      Without 2022 Heat Effect Only & $-1910.9$ & 0.069 &  & $-1.877$ & 0.136\\
-      Without 2022 Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.880$ & 0.136\\
+    Excluding 2022 &
+      Venue Effect Only & $-1862.2$ &   & 0.048 & $-1.857$ & 0.152\\
+     & Heat Effect Only & $-1910.9$ & 0.069 &  & $-1.877$ & 0.136\\
+     & Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.880$ & 0.136\\[1ex]
+    Including 2022 &
+    Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.150\\
+    & Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
+    & Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
     \bottomrule

### 2023-10-13 Manuscript/manuscript.tex
rewrote the GLMM results

@@
 
-\jy{Add the results without 2022 to this table.}
-\of{Does it make sense to order the table how I did, and block by dataset or
-should I block by model?}
 \begin{table}
@@
 
-\jy{Hopefully, the full model is preferred for both with and without 2022}
-\of{It is but the AIC are more similar without 2022}
-
-The estimated parameters of the GLMM model from our analysis are shown in 
-Table~\ref{tab:Gamma_parameters}. These statistics reflect the inclusion of 2022
-in our analysis and shows the three models we considered. The standard deviation
-of the heat effect is $\sigma_h$, the standard deviation of the venue effect
-is $\sigma_v$, the intercept is $\alpha$ and we report $\sqrt{\phi}$ as the
-square root of the dispersion parameter $\phi$.
-The GLMM with both venue and heat random effects was the best model as evidenced
-by a decrease in the AIC compared to the other models with only one random
+The estimated parameters of the GLMM model under the log link are shown in 
+Table~\ref{tab:Gamma_parameters}, with both including and excluding 2022 data.
+Three models were considered, differing in there random effects: venue effect
+only; heat effect only, and both venue and heat effects. Reported are the AIC
+and estimated model parameters: the standard deviation
+of the heat effect $\sigma_h$, the standard deviation of the venue
+effect~$\sigma_v$, the intercept~$\alpha$, and the
+square root of the dispersion parameter $\phi$. With or without 2022,
+the GLMM with both venue and heat random effects was the best model as evidenced
+by the AIC compared to the other models with only one random
 effect.  When we exclude 2022 from the full model, we observe small but intriguing
 discrepancies in the models. The standard deviation of the venue effect, 
-$\sigma_v$, decreases from 0.40 to 0.33 when 2022 is excluded.  Thus, 2022
-increased the noise in the overall data, but many of the other statistics
+$\sigma_v$, increases from 0.33 to 0.40 when 2022 is included, however, the
+changes to the estimates of other parameters
 remained relatively similar, such as the standard deviation of the heat effect
 and the dispersion parameter.
-
-%\of{Can I just refer to it as the dispersion parameter or should I say "square
-%root of the dispersion parameter" Mathematically it shouldn't make a difference 
-%right as 0.134^2 and 0.136^2 are very similar and I am applying a function to 
-%both? Also should I be saying "standard deviation of heat effect", or sigma_v 
-%or both?}
-
-\jy{discuss the differences in the parameter estimates with and without 2022
-  data, especially the sd of the random effect of venue.}
-
-Figure~\ref{fig:VenueEffects} shows the venue effects from 1999 to 2023 based
-on the full model both with and without 2022.  As this graph does not show heat
-\jy{discuss the years the venue effects are high and low}
-effects, the estimates may not align exactly with the box plot presented in
-figure~\ref{fig:Boxplot}.  The venue effect for 2022 is lower than any other
-effect, but may not be as extreme as one would think considering the history
-of reaction times going back to 1999.
-
-Thus, results from the GLMM with both venue and heat random effects will be 
-presented throughout the rest of the paper.
 
@@
   \includegraphics{ComparisonOfVenueEffects}
-  \caption{The venue effects from 1999 to 2023 for the data including 2022 (top)
-  and excluding 2022 (bottom)}
+  \caption{The venue effects from 1999 to 2023 estimated from the GLMM for the
+    data including 2022 (top) and excluding 2022 (bottom).}
   \label{fig:VenueEffects}
@@
 
-Two aims of this section are to
-determine how extreme a reaction time below 0.1 seconds is, and although the
-venue effects did not differ significantly, to determine how the exclusion of 
-2022 could alter the results. Using the gamma family with \texttt{link = log},
-the probability of the reaction time being less than 0.1 seconds can be 
-calculated in R using the methods described in Section~\ref{sec:glmm}. A 
-simulation size of $10,000,000$ was used to estimate the probability of 
-observing a reaction time below two thresholds: 0.1 and 0.08.
+Figure~\ref{fig:VenueEffects} further illustrates the venue effect from 1999 to
+2023 based on the full GLMM both with and without 2022.  The temporal patterns
+of the two sets of effects look very similar, except that the positive effects
+estimated with 2022 included have higher magnitude than those estimated without
+2022, which compensates the negative effect of the 2022 with a large magnitude.
+Both plots agree to some extent with the pattern of the centers of the box plots in
+Figure~\ref{fig:Boxplot}. The mismatch is due to the venue effect, which has a
+higher standard deviation than the venue effect. The venue effect for 2022 is
+lower than that for any other years, but may not be as extreme as one would
+think considering the history of reaction times going back to 1999.
 
@@
 
-Table~\ref{tab:Sim_probability} shows that changing the reaction time 
-barrier from 0.1 seconds to 0.08 seconds drastically reduces the chances of 
-observing a reaction time that would break the barrier. If World Athletics were
-to change the reaction time the way that \citet{komi2009iaaf} suggests, the 
-probability would be so much lower that athletes would likely not
-have a case to make when arguing against a disqualification. When including 2022,
-a probability of $4.34\cdot10^{-3}$ corresponds to roughly 1 in every 230 starts
-being below 0.1 seconds. When we exclude 2022, the probability of observing a
-reaction time below 0.1 seconds decreased to $3.54\cdot10^{-3}$, or 1 in every
-283 starts.  While this seems low, it is not unreasonable to think that certain
-athletes with strong reaction times could begin approach the 0.1 second barrier,
-especially when considering Allen's other reaction times: 0.124 and 0.101 at the
-2022 World Track and Field Championships.  Another important conclusion from our
-results is how excluding 2022 lowered the probability of observing a fast 
-reaction time, suggesting that 2022 was peculiarly fast, which aligns with the
-results presented in section~\ref{subsec:Results_Rank}.
+The GLMM with both venue and heat random effects helps us to assess how extreme
+a reaction time below a given threshold is. The probability of observing a reaction
+time below a threshold, assuming no intentional false starts, can be
+approximated by generating a large number of realizations from the fitted
+GLMM. We used 10 millions realizations in the analysis. 
+Table~\ref{tab:Sim_probability} summarizes the estimated probabilities of
+observing a reaction time below 0.08, 0.09, and 0.10 seconds under two different
+scenarios: one excluding and the other including data from 2022. Apparently,
+excluding 2022 lowered the probability of observing a fast reaction time, but
+the difference is not in a any big magtinude. For example, the probability for
+seeing a reaction time below 0.10 seconds changes from $3.54\cdot 10^{-3}$ to
+$4.20\cdot 10^{-3}$ when 2022 is included. They corresponding to roughly one in
+every 230 starts and 283 starts, respectively. The table shows that changing the
+reaction time barrier from 0.1 seconds to 0.08 seconds drastically reduces the
+chance of observing a reaction time below the barrier, from one in roughly every
+283 starts to one in every 29851 starts using data with 2022 included. The
+results substantiates the recommendations by \citet{komi2009iaaf}.
 
@@
 
-We can use the same model but alter our interpretation slightly to select a 
-reaction time barrier given the probability of observing a time less than the 
-barrier. When considering all the data, if we want a 0.1\% chance of 
-observing an illegally fast reaction time, the reaction
-time barrier should be 0.092 seconds, as shown in Table~\ref{tab:Sim_time}. We
-can repeat this calculation for various probability levels,
-for example a probability of 0.0001 (one disqualification for every 10,000 starts)
-results in a reaction time barrier of 0.083 seconds. As the exclusion of 2022
-from the data resulted in a lower probability of observing a low reaction time,
-the reaction time barrier will be less lenient when calculating the barrier
-based on probability.
+Utilizing the same model, we can subtly shift our interpretation to determine a
+suitable reaction time barrier based on the probability of observing a time
+below said barrier. Including 2022 data, a reaction time barrier of 0.092
+seconds is warranted to maintain a 0.1\% chance of observing an illicitly fast
+reaction time, as delineated in Table~\ref{tab:Sim_time}. This calculation can
+be replicated for various probability levels; for instance, a probability of
+0.0001 (equating to one disqualification per 10,000 starts) necessitates a
+reaction time barrier of 0.083 seconds. The exclusion of 2022 data, which
+inherently lowers the probability of observing a swift reaction time, mandates a
+less lenient reaction time barrier when calculated based on probability, albeit
+without a substantial difference in magnitude.
 
@@
 
-This paper does not consider data from the women's 100 meter hurdle, and although
-that was initially because of concerns over reaction time differences in men and
+We did not consider data from the women's 100 meter hurdle in the analysis.
+In addition to the concerns over reaction time differences in men and
 women, exploratory analysis showed that 2022 was at best only

### 2023-10-13 Manuscript/manuscript.tex
went over discussion

@@
 
-\section{Concluding Remarks}\label{sec:concludingremarks}
+\section{Discussion}\label{sec:concludingremarks}
 
@@
 In addition to the concerns over reaction time differences in men and
-women, exploratory analysis showed that 2022 was at best only
+women, exploratory analysis showed that 2022 women were at best only
 slightly faster relative to other championships going back to 2003.
@@
 
-It is confusing as to why the reaction time barrier for men and women 
-is the same despite numerous articles \citep[e.g.,][]{lipps2011implications, 
-babicc2009reaction, panoutsakopoulos2020gender} stating differently. 
-According to these authors, World Athletics is ignoring differences in gender by 
-setting the reaction time barrier to be equal. 
-If there are gender differences and 0.1 seconds is considered a fair time for 
-women, then 0.1 seconds could not be considered fair for men, as the probability 
-of observing reaction times lower than 0.1 would be elevated.
-\citet{brosnan2017effects} advocated for gender based reaction time barriers 
-which seems logical if there are biological
-differences in reaction times between men and women.
+The uniformity of the reaction time barrier for both men and women is
+perplexing, especially in light of numerous studies that suggest a divergence in
+their respective reaction times \citep[e.g.,][]{lipps2011implications,
+  babicc2009reaction, panoutsakopoulos2020gender}. These authors posit that
+World Athletics may be overlooking inherent gender differences by establishing
+an equal reaction time barrier. If gender disparities do exist and 0.1 seconds
+is deemed a fair threshold for women, it logically follows that the same cannot
+be deemed fair for men, given that the probability of observing sub-0.1-second
+reaction times would be inherently higher. \citet{brosnan2017effects} propounds
+the adoption of gender-specific reaction time barriers, a stance that appears
+logical when considering biological distinctions in reaction times between
+genders.
 
-The first part of this paper showed that athletes' reaction times at the 2022 
-World Track and Field Championships were generally faster than their reaction 
-times at other competitions. Table~\ref{tab:Clusrankresults} shows multiple 
-significant p-values that indicate differences in average reaction times for 
-athletes who competed at the 2022 World Track and Field Championship and another
-competition. Additionally, this paper has shown that a reaction time of 0.1 
-seconds is not as extreme as one may think, and that based on the probabilities
-in Table~\ref{tab:Sim_time}, World Athletics may consider changing the 
-disqualification barrier to 0.8 seconds. If they were to do so, they would 
-allow athletes like Allen to be able to react faster without fear of being 
-disqualified. Taken together, the results show that 2022 was an outlier and 
-despite being disqualified, Allen's time was not so extreme.
+This paper aspires to offer a statistical lens through which to examine Allen's
+disqualification at the 2022 World Track and Field Championships, rather than
+delivering a conclusive judgment regarding potential equipment malfunction. Our
+findings reveal that athletes' reaction times at the event were, in general,
+faster than those recorded at other competitions, with
+Table~\ref{tab:Clusrankresults} presenting multiple significant p-values that
+highlight disparities in average reaction times between the 2022 Championship
+and other competitions. Further, our GLMM results indicate that a reaction time
+of 0.1 seconds may not be as exceptional as commonly believed. According to the
+probabilities in Table~\ref{tab:Sim_time}, World Athletics might contemplate
+adjusting the disqualification barrier to 0.8 seconds, thereby enabling athletes
+like Allen to react more swiftly without the apprehension of
+disqualification. In summary, while the results designate 2022 as an anomalous
+year, Allen's time, despite resulting in disqualification, may not be
+categorically extreme.
 

### 2023-10-16 Manuscript/manuscript.tex
few comments and edits

@@
 to have competed in one race in 2019 and two races in 2022.
+%The 2022 World
+%Championship Heats and Semifinals both had on average lower reaction times than
+%their respective 2019 races, and the 2022 finals were substantially faster. 
 The 2022 World
 Championship Heats and Semifinals both had on average lower reaction times than
-their respective 2019 races, and the 2022 finals were substantially faster. It
+their respective other (2019, 2023 or 2022 national) races, and the 2022 World 
+Championship finals were substantially faster. 
+\eds{please check change}
+It
 is also worth noting is that the fastest reaction time in the 2022 World
@@
 Figure~\ref{fig:Boxplot}. It is clear that the reaction times in the 2022
-box plot are lower (indicating faster reaction times) than in many of the other
+boxplot are lower (indicating faster reaction times) than in many of the other
 years. The median reaction time in 2022 was 0.129 seconds,  compared to 0.156
@@
 \]
-where $g$ is a known link function,
+where $g$ is a known link function, $\alpha$ is a (fixed) intercept,
 $v_i$ is normally distributed with mean zero and
@@
 same athlete comparison with permutation test, the p-value was 
-$0.003$, which is significant at $\alpha = 0.01$. The result of
+$0.003$, which is significant at $\alpha = 0.01$. 
+\eds{According to the table, this p-value is 0.028 (significant at the 0.05 
+level.)}
+The result of
 comparing 2022 and 2023 was very similar (but not exactly the same) to the above
-result, as the p-value from the permutation test was $0.003$. Taken together, these three
+result, as the p-value from the permutation test was $0.003$.
+\eds{According to the table, this p-value is $0.011$.} 
+Taken together, these three
 tests help show that athletes competing at the 2022 World Track and Field
-Championships had faster reaction times than when they had competed previously.
+Championships had faster reaction times than when they had competed at other
+competitions.
 
@@
 The estimated parameters of the GLMM model under the log link are shown in 
-Table~\ref{tab:Gamma_parameters}, with both including and excluding 2022 data.
-Three models were considered, differing in there random effects: venue effect
+Table~\ref{tab:Gamma_parameters}, both excluding and including 2022 data.
+Three models were considered, differing in their random effects: venue effect
 only; heat effect only, and both venue and heat effects. Reported are the AIC
@@
 $\sigma_v$, increases from 0.33 to 0.40 when 2022 is included, however, the
-changes to the estimates of other parameters
+ estimates of other parameters
 remained relatively similar, such as the standard deviation of the heat effect
@@
 2022, which compensates the negative effect of the 2022 with a large magnitude.
-Both plots agree to some extent with the pattern of the centers of the box plots in
+Both plots agree to some extent with the pattern of the centers of the boxplots in
 Figure~\ref{fig:Boxplot}. The mismatch is due to the venue effect, which has a
-higher standard deviation than the venue effect. The venue effect for 2022 is
-lower than that for any other years, but may not be as extreme as one would
+higher standard deviation than the venue effect. 
+\eds{I do not understand what is meant by the previous sentence.}
+The venue effect for 2022 is
+lower than that for any other year, but may not be as extreme as one would
 think considering the history of reaction times going back to 1999.
@@
 excluding 2022 lowered the probability of observing a fast reaction time, but
-the difference is not in a any big magtinude. For example, the probability for
+the difference is not big in magnitude. For example, the probability for
 seeing a reaction time below 0.10 seconds changes from $3.54\cdot 10^{-3}$ to
-$4.20\cdot 10^{-3}$ when 2022 is included. They corresponding to roughly one in
+$4.20\cdot 10^{-3}$ when 2022 is included. These values correspond to roughly one in
 every 230 starts and 283 starts, respectively. The table shows that changing the

### 2023-10-16 Manuscript/manuscript.tex
Respond to Prof Schifano's comments

@@
 same athlete comparison with permutation test, the p-value was 
-$0.003$, which is significant at $\alpha = 0.01$. 
+$0.028$, which is significant at $\alpha = 0.05$. 
 \eds{According to the table, this p-value is 0.028 (significant at the 0.05 
 level.)}
+\of{Resolved}
 The result of
 comparing 2022 and 2023 was very similar (but not exactly the same) to the above
-result, as the p-value from the permutation test was $0.003$.
+result, as the p-value from the permutation test was $0.011$.
 \eds{According to the table, this p-value is $0.011$.} 
+\of{Resolved}
 Taken together, these three
@@
 2022, which compensates the negative effect of the 2022 with a large magnitude.
-Both plots agree to some extent with the pattern of the centers of the boxplots in
+Both plots agree to some extent with the pattern of the boxplots in
 Figure~\ref{fig:Boxplot}. The mismatch is due to the venue effect, which has a
@@
 \eds{I do not understand what is meant by the previous sentence.}
+\of{I believe Prof Yan wrote that. I would write that sentence as "The mismatch
+is due to the heat effect, as venue effect only captures a part of the patterns
+in the data" or something of that nature.}
+
 The venue effect for 2022 is

### 2023-10-16 Manuscript/manuscript.tex
Fixed typo

@@
 2022, which compensates the negative effect of the 2022 with a large magnitude.
-Both plots agree to some extent with the pattern of the boxplots in
+Both plots agree to some extent with the pattern of the box plots in
 Figure~\ref{fig:Boxplot}. The mismatch is due to the venue effect, which has a
@@
 probabilities in Table~\ref{tab:Sim_time}, World Athletics might contemplate
-adjusting the disqualification barrier to 0.8 seconds, thereby enabling athletes
+adjusting the disqualification barrier to 0.08 seconds, thereby enabling athletes
 like Allen to react more swiftly without the apprehension of

### 2023-10-16 Manuscript/manuscript.tex
touch up

@@
 in at least one race at each championship, however it is possible for athletes
-to have competed in one race in 2019 and two races in 2022.
-%The 2022 World
-%Championship Heats and Semifinals both had on average lower reaction times than
-%their respective 2019 races, and the 2022 finals were substantially faster. 
-The 2022 World
-Championship Heats and Semifinals both had on average lower reaction times than
+to have competed in one race in 2019 and two races in 2022. The 2022 World
+Championship heats and semifinals both had on average lower reaction times than
 their respective other (2019, 2023 or 2022 national) races, and the 2022 World 
-Championship finals were substantially faster. 
-\eds{please check change}
-It
+Championship finals were substantially faster.  It
 is also worth noting is that the fastest reaction time in the 2022 World
@@
 same athlete comparison with permutation test, the p-value was 
-$0.028$, which is significant at $\alpha = 0.05$. 
-\eds{According to the table, this p-value is 0.028 (significant at the 0.05 
-level.)}
-\of{Resolved}
-The result of
+$0.028$, which is significant at level $0.05$.  The result of
 comparing 2022 and 2023 was very similar (but not exactly the same) to the above
 result, as the p-value from the permutation test was $0.011$.
-\eds{According to the table, this p-value is $0.011$.} 
-\of{Resolved}
 Taken together, these three
@@
 2022, which compensates the negative effect of the 2022 with a large magnitude.
-Both plots agree to some extent with the pattern of the box plots in
-Figure~\ref{fig:Boxplot}. The mismatch is due to the venue effect, which has a
-higher standard deviation than the venue effect. 
-\eds{I do not understand what is meant by the previous sentence.}
-\of{I believe Prof Yan wrote that. I would write that sentence as "The mismatch
-is due to the heat effect, as venue effect only captures a part of the patterns
-in the data" or something of that nature.}
-
-The venue effect for 2022 is
+Both plots agree to some extent with the pattern of the boxplots in
+Figure~\ref{fig:Boxplot}. The mismatch is due to the heat effect, since the nevue
+effect only captures a smaller part of the variation than the heat effect as
+evident from their standard deviations. The venue effect for 2022 is
 lower than that for any other year, but may not be as extreme as one would

### 2023-10-17 Manuscript/manuscript.tex
fixed one typo

@@
 Both plots agree to some extent with the pattern of the boxplots in
-Figure~\ref{fig:Boxplot}. The mismatch is due to the heat effect, since the nevue
+Figure~\ref{fig:Boxplot}. The mismatch is due to the heat effect, since the venue
 effect only captures a smaller part of the variation than the heat effect as

### 2023-10-18 Archive/Models.Rmd
Contains many old models that were at one time considered but not necessary anymore

--- 2023-05-11 Code/Models.Rmd
@@
 
-```{}
 The best selected model is men_finals_year_gamma as that has the smallest AIC: -491 and largest log-lik of 248.5
-```
 
@@
 ```
-```{}
+
 The best selected model is men_old_finals_year_gamma as that has the smallest AIC: -465 and largest log-lik of 235.5.
 The models for the older data does not seem to fit the data as good as for the full data set.
-```
 
@@
 ```
-```{}
+
 The best selected model is men_pooled_year_gamma_log as it had the smallest AIC and the largest log-lik
-```
 
@@
 ```
-```{}
+
 The best model is men_old_pooled_year_gamma as that has the smallest AIC: -1687.5 and the largest log-lik: 846.8
-```
-
-```{}
-Simulate to see what prob of a time less than 0.1 second rxn time.
-```
-
-##Simulations
-
-```{r}
-library(lme4)
-men_pooled_gamma_year <- glmer(RxnTime ~ (1 | Year), data = men_pooled, family = Gamma(link = "log"))
-baseline <- glm(RxnTime ~ 1, data = men_pooled, family = Gamma(link = "log"))
-
-simfit <- function(men_pooled_gamma_year, n = 12000000) {
-  vc <- as.data.frame(VarCorr(men_pooled_gamma_year))
-  ## mu <- exp(coef(men_pooled_gamma_year)$Year[,]) ## using fitted random effects only
-  z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
-  mu <- exp(fixef(men_pooled_gamma_year) + z)
-  shape <- 1 / vc$vcov[2]
-  x <- rgamma(n, shape = shape, scale = mu / shape)
-}
-sbaseline <- summary(baseline)
-#print(sbaseline)
-x <- simfit(men_pooled_gamma_year)
-mean(x < 0.1)
-quantile(x, prob = c(.00001, .0001, .0005, .001))
-## compare with and without random effect
-plot(density(x)) # with random effect
-curve(dgamma(x, shape = 1 / sbaseline$dispersion,
-             scale = exp(coef(baseline)) * sbaseline$dispersion),
-      0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
-```
-
-```{r}
-set.seed(1)
-simulation <- function(model, base){
-    simfit <- function(model, n = 10000000) {
-    vc <- as.data.frame(VarCorr(model))
-    ## mu <- exp(coef(model)$Year[,]) ## using fitted random effects only
-    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
-    mu <- exp(fixef(model) + z)
-    shape <- 1 / vc$vcov[2]
-    x <- rgamma(n, shape = shape, scale = mu / shape)
-    x
-    }
-  sbase <- summary(base)
-  #print(sbase)
-  x <- simfit(model)
-  print(mean(x < 0.08))
-  print(quantile(x, prob = c(.00001, .0001, .0005, .001)))
-  ## compare with and without random effect
-  plot(density(x)) # with random effect
-  curve(dgamma(x, shape = 1 / sbase$dispersion, 
-               scale = exp(coef(base)) * sbase$dispersion), 
-        0.08, 0.22, add = TRUE, col = "blue") # fixed effect only
-}
-simulation(old_pooled_year_gamma_log, old_pooled_one_gamma_log)
-```
-
-```{}
-Every 1000 racers, we expect 7 to have a reaction time below 0.1 seconds.  If we want 1 out of every 1000
-```
-
-##2019 vs 2022 Comparisons
-```{r}
-comp <- read.csv("../Data/Clusrank2019vs2022.csv")
-#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
-colnames(comp) <- c("Time", "Athlete", "Year")
-#View(comp)
-```
-
-```{}
-One thing that we discovered is that including group=Year does not effect the the z score or p value.  Changing the method from
-rgl to ds does effect z score.
-```
-
-```{r}
-library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, alternative = "greater", method = "ds")
-#clusWilcox.test(Time, group=Year, cluster = Athlete, data = comp, alternative = "greater", method = "rgl")
-#pairwise.clusWilcox.test(comp$Time, group= comp$Year, cluster = comp$Athlete, alternative = "greater", method = "rgl")
-```
-
-```{r}
-nations <- read.csv("../Data/ClusrankNatvsWorld.csv")
-colnames(nations) <- c("Time", "Athlete", "Year")
-```
-
-```{r}
-library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, alternative = "greater", method = "ds")
-#clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations, paired = TRUE, alternative = "greater", method = "rgl")
-
-```
-
-```{}
-In both examples, rgl produced a smaller p value
-```
-
-```{r}
-mini <- min(aggregate(times$RxnTime, list(times$Batch), FUN =mean)$x)
-maxi <- max(aggregate(times$RxnTime, list(times$Batch), FUN =mean)$x)
-cat(maxi - mini)
-```
 

### 2023-10-26 Code/GLMM.Rmd
Removed a few summaries of stuff that wasn't needed
Changed the data source file as I renamed it
Fixed line width
added more comments

@@
 
-This code below creates the base models without any random effects.  These are used
-as a baseline to compare the venue effect models to in the simulation code below.
+This code below creates the base models without any random effects.  These are
+used as a baseline to compare the venue effect models to in the simulation code
+below.
 ```{r}
@@
 summary(with2022_heat_gamma_log)
+summary(without2022_heat_gamma_log)
 ```
@@
 without2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
-                                   data = without2022, family = Gamma(link = "log"))
+                                   data = without2022,
+                                   family = Gamma(link = "log"))
 with2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
-                               data = with2022, family = Gamma(link = "log"))
+                               data = with2022,
+                               family = Gamma(link = "log"))
 ```
@@
 ```
-This anova test determined that the two effect model is a significant improvement
-over the Venue effect model.
+This anova test determined that the two effect model is a significant
+improvement over the Venue effect model.
 
 This code simulates the probability of observing a reaction time less than 0.1
-for the various data sets.  The function takes in two models and a reaction time,
-it is important that the two models that are compared are from the same data set;
-we want to compare old_finals_two_effect and old_finals_one_gamma_log.
+for the various data sets.  The function takes in two models and a reaction
+time, it is important that the two models that are compared are from the same
+data set; we want to compare old_finals_two_effect and old_finals_one_gamma_log.
 ```{r}
@@
 This next chunk of code runs the simulation code for each dataset comparing the
-two effect model to the base model
+two effect model to the base model. We run this code at 3 reaction time barriers:
+0.08, 0.09, and 0.1 seconds.
 ```{r}
@@
 set.seed(1)
-sim_two_effect_time(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.1)
-sim_two_effect_time(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.1)
+sim_two_effect_time(without2022_two_effect_gamma_log,
+                    without2022_one_gamma_log, 0.1)
+sim_two_effect_time(with2022_two_effect_gamma_log,
+                    with2022_one_gamma_log, 0.1)
 ```
@@
 set.seed(1)
-sim_two_effect_time(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.09)
-sim_two_effect_time(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.09)
+sim_two_effect_time(without2022_two_effect_gamma_log,
+                    without2022_one_gamma_log, 0.09)
+sim_two_effect_time(with2022_two_effect_gamma_log,
+                    with2022_one_gamma_log, 0.09)
 ```
@@
 set.seed(1)
-sim_two_effect_time(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.08)
-sim_two_effect_time(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.08)
+sim_two_effect_time(without2022_two_effect_gamma_log,
+                    without2022_one_gamma_log, 0.08)
+sim_two_effect_time(with2022_two_effect_gamma_log,
+                    with2022_one_gamma_log, 0.08)
 ```
 
-sim_two_effect_prob is a slight modification of the earlier function that instead
-takes as input a probability rather than a time.  The output is a reaction time 
-barrier that allows a reaction time below the barrier one in every 1/(input probability)
-times. Decreasing the probability will impose a stricter barrier as it is less
-probable to observe a reaction time of 0.08 rather than 0.09.  Below are the
-recommended reaction time barriers to disqualify on average 1 in every 1,000 starts.
+sim_two_effect_prob is a slight modification of the earlier function that
+instead takes as input a probability rather than a time.  The output is a
+reaction time barrier that allows a reaction time below the barrier one in every
+1/(input probability) times. Decreasing the probability will impose a stricter
+barrier as it is less probable to observe a reaction time of 0.08 rather than
+0.09.  Below are the recommended reaction time barriers to disqualify on average
+1 in every 1,000 starts.
 ```{r}
@@
 set.seed(1)
-sim_two_effect_prob(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.001)
-sim_two_effect_prob(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.001)
+sim_two_effect_prob(without2022_two_effect_gamma_log,
+                    without2022_one_gamma_log, 0.001)
+sim_two_effect_prob(with2022_two_effect_gamma_log,
+                    with2022_one_gamma_log, 0.001)
 ```
@@
 set.seed(1)
-sim_two_effect_prob(without2022_two_effect_gamma_log, without2022_one_gamma_log, 0.0001)
-sim_two_effect_prob(with2022_two_effect_gamma_log, with2022_one_gamma_log, 0.0001)
+sim_two_effect_prob(without2022_two_effect_gamma_log,
+                    without2022_one_gamma_log, 0.0001)
+sim_two_effect_prob(with2022_two_effect_gamma_log,
+                    with2022_one_gamma_log, 0.0001)
 ```
@@
 set.seed(1)
-sim_one_effect_time(without2022_venue_gamma_log, without2022_one_gamma_log, 0.1)
-sim_one_effect_time(with2022_venue_gamma_log, with2022_one_gamma_log, 0.1)
+sim_one_effect_time(without2022_venue_gamma_log,
+                    without2022_one_gamma_log, 0.1)
+sim_one_effect_time(with2022_venue_gamma_log,
+                    with2022_one_gamma_log, 0.1)
 ```
@@
 set.seed(1)
-sim_one_effect_time(without2022_heat_gamma_log, without2022_one_gamma_log, 0.1)
-sim_one_effect_time(with2022_heat_gamma_log, with2022_one_gamma_log, 0.1)
+sim_one_effect_time(without2022_heat_gamma_log,
+                    without2022_one_gamma_log, 0.1)
+sim_one_effect_time(with2022_heat_gamma_log,
+                    with2022_one_gamma_log, 0.1)
 ```
@@
 #15
-sim_one_effect_prob(without2022_venue_gamma_log, without2022_one_gamma_log, 0.001)
-sim_one_effect_prob(with2022_venue_gamma_log, with2022_one_gamma_log, 0.001)
+sim_one_effect_prob(without2022_venue_gamma_log,
+                    without2022_one_gamma_log, 0.001)
+sim_one_effect_prob(with2022_venue_gamma_log,
+                    with2022_one_gamma_log, 0.001)
 ```
@@
 #16
-sim_one_effect_prob(without2022_venue_gamma_log, without2022_one_gamma_log, 0.0001)
-sim_one_effect_prob(with2022_venue_gamma_log, with2022_one_gamma_log, 0.0001)
+sim_one_effect_prob(without2022_venue_gamma_log,
+                    without2022_one_gamma_log, 0.0001)
+sim_one_effect_prob(with2022_venue_gamma_log,
+                    with2022_one_gamma_log, 0.0001)
 ```
+
+women_finals <- subset(times, Type == "F" & Gender == "F" & Time != 'DQ') # & Year < 2022)
+women_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "F" & Time != 'DQ')

### 2023-10-26 Code/GLMM.Rmd
Added a lot more text, explanation, re-organized

@@
 output:
+  pdf_document: default
   html_document:
@@
 
-To see the probabilities associated with a reaction time below 0.1 seconds for
-the two effect model, run code chunks 1, 2, 5, 6, 7.
+### This is the code file that shows our results. Code chunks have been labeled for convience.
 
+All outputs rely on code chunks 1-5, run these first.
+To see the outputs of Table 2 run code chunks 6-8
+To see the outputs of Table 2 run code chunks 11-14
+To see the outputs of Table 4 run code chunks 15-17
+
+### 1
 The code below loads in the data sets from the csv file.
 ```{r}
-#1
-times <- read.csv("../Data/Thesis2023.csv")
+
+times <- read.csv("../Data/GLMMData.csv")
 colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
@@
 
-This code below creates the base models without any random effects.  These are
+### 2
+This code below creates the null models without any random effects.  These are
 used as a baseline to compare the venue effect models to in the simulation code
@@
 ```{r}
-#2
 library(lme4)
@@
 
+### 3
 This code creates the venue effect only models.
 ```{r}
-#3
 library(lme4)
@@
 
-```{r}
-summary(with2022_venue_gamma_log)
-summary(without2022_venue_gamma_log)
-```
-
+### 5
 This creates the heat effect only models.
 ```{r}
-#4
 library(lme4)
@@
 
-```{r}
-summary(with2022_heat_gamma_log)
-summary(without2022_heat_gamma_log)
-```
-
+### 5
 This creates the models with a venue effect and heat effect.
 ```{r}
-#5
 library(lme4)
@@
 
+### 6
+This shows a summary of the venue effect only models.
+```{r}
+summary(with2022_venue_gamma_log)
+summary(without2022_venue_gamma_log)
+```
+
+### 7
+This shows a summary of the heat effect only models.
+```{r}
+summary(with2022_heat_gamma_log)
+summary(without2022_heat_gamma_log)
+```
+
+### 8
+This shows a summary of the two effect models.
 ```{r}
@@
 
+### 9
+This test compares the full model with the heat effect only model.
 ```{r}
+#9
 anova(with2022_heat_gamma_log, with2022_two_effect_gamma_log)
@@
 This anova test determined that the two effect model is an improvement over the
-heat effect model. The test statistic is 0.01059 which is significant at alpha 
-= 0.05 but not at alpha = 0.01.
+heat effect model. The test statistic is 0.008305 which is significant at 
+$\alpha = 0.05$.
 
+### 10
+This test compares the full model with the venue effect only model.
 ```{r}
+#10
 anova(with2022_venue_gamma_log, with2022_two_effect_gamma_log)
@@
 
+### 11
 This code simulates the probability of observing a reaction time less than 0.1
@@
 time, it is important that the two models that are compared are from the same
-data set; we want to compare old_finals_two_effect and old_finals_one_gamma_log.
+data set; we want to compare `old_finals_two_effect` and 
+`old_finals_one_gamma_log`.
 ```{r}
@@
 
+### 12
 This next chunk of code runs the simulation code for each dataset comparing the
@@
 
+### 13
 ```{r}
@@
 
+### 14
 ```{r}
@@
 
-sim_two_effect_prob is a slight modification of the earlier function that
+### 15
+`sim_two_effect_prob` is a slight modification of the earlier function that
 instead takes as input a probability rather than a time.  The output is a
@@
 
-This runs the sim_two_effect_prob function at a probability of 1 in every 1,000
-sprinters.
+### 16
+This runs the `sim_two_effect_prob` function at a probability of 1 in every 
+1,000 sprinters.
 ```{r}
@@
 
+### 17
 The code below repeats the previous section but for an average of 1 in every
@@
 ```
-
-This next code section replicates the code above but the various simulation
-functions are meant for one random effect instead of two.
-```{r}
-#11
-library(lme4)
-set.seed(1)
-sim_one_effect_time <- function(model, base, RxnTime){
-    simfit <- function(model, n = 10000000) {
-    vc <- as.data.frame(VarCorr(model))
-    ## mu <- exp(coef(model)$Venue[,]) ## using fitted random effects only
-    z <- rnorm(n, sd = vc$sdcor[1]) #$ using generated random effects
-    mu <- exp(fixef(model) + z)
-    shape <- 1 / vc$vcov[2]
-    x <- rgamma(n, shape = shape, scale = mu / shape)
-    x
-    }
-  sbase <- summary(base)
-  #print(sbase)
-  x <- simfit(model)
-  print(mean(x < RxnTime))
-  ## compare with and without random effect
-  plot(density(x)) # with random effect
-  curve(dgamma(x, shape = 1 / sbase$dispersion, 
-               scale = exp(coef(base)) * sbase$dispersion), 
-        0.08, 0.25, add = TRUE, col = "blue") # fixed effect only
-}
-```
-
-```{r}
-#12
-set.seed(1)
-sim_one_effect_time(without2022_venue_gamma_log,
-                    without2022_one_gamma_log, 0.1)
-sim_one_effect_time(with2022_venue_gamma_log,
-                    with2022_one_gamma_log, 0.1)
-```
-
-```{r}
-#13
-set.seed(1)
-sim_one_effect_time(without2022_heat_gamma_log,
-                    without2022_one_gamma_log, 0.1)
-sim_one_effect_time(with2022_heat_gamma_log,
-                    with2022_one_gamma_log, 0.1)
-```
-
-```{r}
-#14
-library(lme4)
-set.seed(1)
-sim_one_effect_prob <- function(model, base, prob) {
-  simfit <- function(model, n = 10000000) {
-    vc <- as.data.frame(VarCorr(model))
-    z <- rnorm(n, sd = vc$sdcor[1])
-    mu <- exp(fixef(model) + z)
-    shape <- 1 / vc$vcov[2]
-    x <- rgamma(n, shape = shape, scale = mu / shape)
-    x
-  }
-
-  sbase <- summary(base)
-  x <- simfit(model)
-  print(quantile(x, probs = prob))
-}
-```
-
-```{r}
-#15
-sim_one_effect_prob(without2022_venue_gamma_log,
-                    without2022_one_gamma_log, 0.001)
-sim_one_effect_prob(with2022_venue_gamma_log,
-                    with2022_one_gamma_log, 0.001)
-```
-
-```{r}
-#16
-sim_one_effect_prob(without2022_venue_gamma_log,
-                    without2022_one_gamma_log, 0.0001)
-sim_one_effect_prob(with2022_venue_gamma_log,
-                    with2022_one_gamma_log, 0.0001)
-```
-
-women_finals <- subset(times, Type == "F" & Gender == "F" & Time != 'DQ') # & Year < 2022)
-women_pooled <- subset(times, (Type == "F" | Type == "S") & Gender == "F" & Time != 'DQ')

### 2023-11-03 Code/GLMM.Rmd
TAS formatting and minor edits to rmds

@@
 author: "Owen Fiore"
-date: "2023-05-31"
+date: "2023-11-03"
 output:
@@
 ```{r}
-library(lme4)
 without2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = without2022, 
@@
 
-### 5
+### 4
 This creates the heat effect only models.
 ```{r}
-library(lme4)
 without2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = without2022, 
@@
 ```{r}
-library(lme4)
 without2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
@@
 ```{r}
-#9
 anova(with2022_heat_gamma_log, with2022_two_effect_gamma_log)
@@
 ```{r}
-#10
 anova(with2022_venue_gamma_log, with2022_two_effect_gamma_log)
@@
 ```{r}
-#6
-set.seed(1)
-library(lme4)
 sim_two_effect_time <- function(model, base, RxnTime) {
@@
 ```{r}
-#7
 set.seed(1)
@@
 ```{r}
-#8
-set.seed(1)
-library(lme4)
 sim_two_effect_prob <- function(model, base, prob) {
@@
 ```{r}
-#10
 set.seed(1)

### 2023-11-03 Code/RankComparison.Rmd
Dont have data for this
Fixed line width
TAS formatting and minor edits to rmds

--- 2023-09-26 Code/RankComparison.Rmd
@@
 ---
-title: "Descriptive_Stats"
+title: "Rank Comparison"
 author: "Owen Fiore"
-date: "2023-05-21"
+date: "2023-11-03"
 output: pdf_document
@@
 
-The code below reads in the data comparing the 2022 National Competitions to the 2022
-World Championships.
+The code below reads in the data comparing the 2022 National Competitions to the
+2022 World Championships.
 ```{r}
@@
 set.seed(1)
-library(clusrank)
 clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_clusrank,
@@
 The tests below are very similar to the ones above, the only difference being
-grouping by Year instead of Stage (2 if 2022 and 3 if 2023).
+grouping by Year (2 if 2022 and 3 if 2023).
 ```{r}
 set.seed(1)
-library(clusrank)
 clusWilcox.test(Time, group=Year, cluster = Athlete, data = new_clusrank,
@@
 ```{r}
-library(clusrank)
 clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,

### 2023-11-03 Manuscript/manuscript.tex
TAS formatting and minor edits to rmds

@@
-\documentclass[12pt, letterpaper, titlepage]{article}
+%\documentclass[12pt, letterpaper, titlepage]{article}
+\documentclass[12pt, letterpaper]{article}
 
@@
 \usepackage{pdfpages}
-\usepackage[pagewise]{lineno}
+\usepackage{lineno}
 \usepackage{mwe}
@@
 
-\title{On Devon Allen's Disqualification at the 2022 World Track and Field 
+% NOTE: To produce blinded version, replace "0" with "1" below.
+\newcommand{\blind}{0}
+
+%\title{On Devon Allen's Disqualification at the 2022 World Track and Field 
+%Championships} 
+%
+%\author{Owen Fiore\\
+%%   \href{mailto:owen.fiore@uconn.edu}
+%% {\nolinkurl{owen.fiore@uconn.edu}}\\
+  %Elizabeth D. Schifano\\
+  %Jun Yan\\[1ex]
+  %Department of Statistics, University of Connecticut\\
+%}
+%\date{}
+
+\begin{document}
+%\maketitle
+
+\if0\blind
+{
+  \title{\bf On Devon Allen's Disqualification at the 2022 World Track and Field 
 Championships}
-
-\author{Owen Fiore\\
+  \author{Owen Fiore, %\\
 %   \href{mailto:owen.fiore@uconn.edu}
 % {\nolinkurl{owen.fiore@uconn.edu}}\\
-  Elizabeth D. Schifano\\
+  Elizabeth D. Schifano, %\\
   Jun Yan\\[1ex]
@@
 \date{}
+  \maketitle
+} \fi
 
-\begin{document}
-\maketitle
+\if1\blind
+{
+  \bigskip
+  \bigskip
+  \bigskip
+  \begin{center}
+  {\LARGE\bf On Devon Allen's Disqualification at the 2022 World Track and Field 
+Championships}
+\end{center}
+  \bigskip
+} \fi
+
+\doublespace
 

### 2023-11-09 Code/RankComparison.Rmd
removed some chunks to make it run faster

@@
 The code below performs two clustered Wilcox tests, the first one without using
-exact = True, and the second one with exact = True and 10,000 random
+exact = True, and the second one with exact = True and 1,000,000 random
 permutations. The tests compare the times and group by the Stage of the
@@
 exact = TRUE which is used when sample sizes are small (as they are in this
-case), additionally the number of random permutations (B) is set to be 10,000.
+case), additionally the number of random permutations (B) is set to be 1,000,000.
 
@@
 ```{r}
-nations_noprelim <- read.csv("../Data/Clusrank2019vs2022NoPrelims.csv")
-colnames(nations_noprelim) <- c("Time", "Athlete", "Year")
+#nations_noprelim <- read.csv("../Data/Clusrank2019vs2022NoPrelims.csv")
+#colnames(nations_noprelim) <- c("Time", "Athlete", "Year")
 ```
@@
 ```{r}
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,
-                method = "ds")
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,
-                method = "ds", 
-                exact = TRUE, B = 10000)
+#clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,
+#                method = "ds")
+#clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,
+#                method = "ds", 
+#                exact = TRUE, B = 10000)
 ```
-
-```{r}
-women_data <- read.csv("../Data/WomenClusrank2019vs2022.csv")
-colnames(women_data) <- c("Time", "Athlete", "Year")
-```
-
-```{r}
-set.seed(1)
-library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_data,
-                method = "ds")
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_data,
-                method = "ds", 
-                exact = TRUE, B = 1000000)
-```

### 2023-11-20 Code/RankComparison.Rmd
making sure everything is consistent

@@
 The code below performs two clustered Wilcox tests, the first one without using
-exact = True, and the second one with exact = True and 1,000,000 random
+exact = TRUE, and the second one with exact = TRUE and 1,000,000 random
 permutations. The tests compare the times and group by the Stage of the
-competition which in the data is coded as 1 if it was a national championship
+competition, which in the data is coded as 1 if it was a national championship
 and 2 if it was the world championships. The tests cluster by athlete, of which
-for each dataset an integer from 1-11 is used (the values are not significant).
+for each dataset an integer from 1-11 is used (the values are arbitrary).
 For more information about the "ds" method, please look at the Rank-based
 Comparison subsection of the Methods section of the paper or at this link here:
-https://www.jstatsoft.org/article/view/v096i06.  The second test also includes
-exact = TRUE which is used when sample sizes are small (as they are in this
-case), additionally the number of random permutations (B) is set to be 1,000,000.
+https://www.jstatsoft.org/article/view/v096i06.
 
@@
 old_clusrank <- read.csv("../Data/Clusrank2019vs2022.csv")
-#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
 colnames(old_clusrank) <- c("Time", "Athlete", "Year")
-#View(comp)
 ```
@@
 new_clusrank <- read.csv("../Data/Clusrank2022vs2023.csv")
-#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
 colnames(new_clusrank) <- c("Time", "Athlete", "Year")
-#View(comp)
 ```
@@
 
-This is seeing what would happen if we remove prelims, the sample sizes in the
-test become very small: 26 observations and 9 clusters.
+The following code could be used to see what happens if we remove prelims. 
+The sample sizes in the test become very small: 26 observations and 9 clusters.
 ```{r}

### 2023-11-20 Manuscript/manuscript.tex
making sure everything is consistent

@@
    2022 vs 2023 & $0.011$ & $0.003$ \\
-   2022 National vs International & $2\cdot10^{-6}$ & $0.001$ \\
+   2022 National vs International & $<0.001$ & $0.002$ \\
    \bottomrule
@@
 as discussed previously in Section~\ref{sec:rank}. The national versus international
-comparison produced lower p-values than the 2019 versus 2022 comparison, with the
-national vs international permutation test producing a p-value of $2\cdot10^{-6}$.
-This is a highly significant test result and shows that athletes who competed at the
+comparisons produced lower p-values than the 2019 versus 2022 and 2022 versus 2023comparisons.%, with the
+%national vs international permutation test producing a p-value of $2\cdot10^{-6}$.
+The highly significant permutation test result provides substantial evidence 
+that athletes who competed at the
 2022 World Track and Field Championships were on average, significantly faster
@@
      & Heat Effect Only & $-1910.9$ & 0.069 &  & $-1.877$ & 0.136\\
-     & Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.880$ & 0.136\\[1ex]
+     & Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.868$ & 0.136\\[1ex]
     Including 2022 &
-    Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.150\\
+    Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.149\\
     & Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
-    & Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
+    & Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.881$ & 0.134\\
     \bottomrule
@@
    \midrule
-   Excluding 2022 & $3.35\cdot10^{-5}$ & $4.45\cdot10^{-4}$ &  $3.54\cdot10^{-3}$  \\ 
-   Including 2022 & $4.20\cdot10^{-5}$ & $5.70\cdot10^{-4}$ & $4.34\cdot10^{-3}$ \\
+   Excluding 2022 & $3.31\cdot10^{-5}$ & $4.46\cdot10^{-4}$ &  $3.52\cdot10^{-3}$  \\ 
+   Including 2022 & $4.16\cdot10^{-5}$ & $5.70\cdot10^{-4}$ & $4.38\cdot10^{-3}$ \\
    \bottomrule

### 2023-11-30 Code/RankComparison.Rmd
Removed final chunk

@@
 For more information about the "ds" method, please look at the Rank-based
-Comparison subsection of the Methods section of the paper or at this link here:
-https://www.jstatsoft.org/article/view/v096i06.
+Comparison subsection of the Methods section of the paper or
+[at this link here](https://www.jstatsoft.org/article/view/v096i06.)
 
@@
 The tests below are very similar to the ones above, the only difference being
-grouping by Year instead of Stage (1 if 2019 and 2 if 2022).
+grouping by Year instead of Stage (0 if 2019 and 1 if 2022).
 ```{r}
@@
 The tests below are very similar to the ones above, the only difference being
-grouping by Year (2 if 2022 and 3 if 2023).
+grouping by Year (0 if 2022 and 1 if 2023).
 ```{r}
@@
 ```
-
-The following code could be used to see what happens if we remove prelims. 
-The sample sizes in the test become very small: 26 observations and 9 clusters.
-```{r}
-#nations_noprelim <- read.csv("../Data/Clusrank2019vs2022NoPrelims.csv")
-#colnames(nations_noprelim) <- c("Time", "Athlete", "Year")
-```
-
-```{r}
-#clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,
-#                method = "ds")
-#clusWilcox.test(Time, group=Year, cluster = Athlete, data = nations_noprelim,
-#                method = "ds", 
-#                exact = TRUE, B = 10000)
-```

### 2023-12-01 Code/GLMM.Rmd
making sure everything is consistent
blinded supplemental files, removed noprelim data from readme

@@
 title: "GLMM"
-author: "Owen Fiore"
-date: "2023-11-03"
+author: "Blinded"
+date: "2023-12-01"
 output:
@@
 This anova test determined that the two effect model is an improvement over the
-heat effect model. The test statistic is 0.008305 which is significant at 
+heat effect model. The p-value is 0.008305 which is significant at 
 $\alpha = 0.05$.

### 2024-06-06 Manuscript/reply.tex
Replied to most comments

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\usepackage{hyperref}

\usepackage{listings}
\usepackage{rotating, graphicx}
\graphicspath{{./}, {./image/}}
\usepackage{booktabs, natbib}
% \usepackage{breakurl}
% \usepackage [english]{babel}
\usepackage{amsmath, amsbsy, amsthm, epsfig, epsf, psfrag, graphicx, 
amssymb, enumerate}
\usepackage{bm}
\usepackage{multirow, multicol}

\usepackage[dvipsnames]{color}
\definecolor{darkblue}{rgb}{0.1, 0.2, 0.6}

\newcommand{\jy}[1]{\textcolor{red}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
\newcommand{\mc}[1]{\textcolor{green}{(MC: #1)}}

\sloppy

% \usepackage{csquotes}
% \usepackage [autostyle, english = american]{csquotes}
% \MakeOuterQuote{"}

% \usepackage{bibentry}
\newenvironment{comment}%
{\begin{quotation}\noindent\small\it\color{darkblue}\ignorespaces%
}{\end{quotation}}

\begin{document}

\begin{center}
  {\Large\bf Response to the Comments}
\end{center}

We extend our gratitude to the Editor and Associate Editor for 
granting us the opportunity to revise this manuscript. We also want to
express our appreciation to the reviewer for their valuable comments.

The manuscript has been revised accordingly with the most notable change being 
the addition of Section 5 (Classroom Implementation). We believe this revision 
has enhanced the both the quality and utility of this paper.

Point-by-point responses to the comments are as follows, with the
comments quoted in \emph{\color{darkblue} italic and blue}.

\subsection*{To Associate Editor}

\begin{comment}
Using Devon Allen's disqualification at the 2022 World Track and Field
championships as a hook, this paper poses two questions. First, were reaction
times at the 2022 World Track and Field Championships faster than they had been
earlier in 2022 national-level competitions, the 2019 World Track and Field
Championships, and the 2023 World Track and Field Championships? Second, based
on modeled reaction times, is the current standard of a 0.1 second threshold for
reaction times before disqualification too conservative?
\end{comment}

Thank you very much for your encouraging comments and great suggestion.  
We have now included a new section, \textit{Section 5 Classroom Implementation},
that discusses how elements of this manuscript have been incorporated into both 
graduate- and undergraduate-level teaching at the authors' home university.

For blinding purposes, course numbers, author initials and websites, and 
university name have been redacted, but all this information appears in the 
unblinded version.  In brief, this material has been covered at the graduate
level in a statistical computing course, open to all graduate students in 
Statistics and others with permission, that covers the bootstrap. Detailed R 
code and graphical illustrations were provided to students in the open-access 
course notes to elucidate the distinctions between employing null
hypotheses with and without defined parameters.  This link is included in the 
unblinded version of the paper; for the blinded review, we printed the 
relevant portion of the notes to pdf and included this blinded pdf file in the 
Supplemental Materials. %\eds{Not sure how we can blind your github 
%website for the submission unless we actually don't include it in the blinded 
%version. Could we print to PDF the relevant section of the notes and include as 
%a supplement for review purposes?}

Similarly at the undergraduate-level, students in an R-based applied regression 
course were provided with an R Markdown file with simulation code in order to 
demonstrate how the KS test fails with estimated parameters in the hypothesized 
distribution. The R Markdown file that was provided to students is included in 
the Supplemental Materials. This R-based applied linear regression course 
has the following prerequisites: two semesters of introductory statistics OR a 
Calculus~2-based statistical methods course.

At both the graduate and undergraduate level, students were able to follow the 
demonstrations closely and gain a clear understanding of the implications of 
using estimated parameters within test statistics.  At the graduate level, 
students further were able to appreciate the efficacy of the parametric 
bootstrap method as a corrective measure.

\subsection*{To Reviewer}

\begin{comment}
To address the first point, the paper relies on descriptive statistics, ranked
tests, and the results of a set of GLMMs. The descriptive data analysis lends
credence to the authors' argument that the reaction times at the 2022 World
Championships was faster than at any of the relevant comparisons. The rank test
appears to be set up correctly and I believe the authors are interpreting the
results correctly. Similarly, I believe the GLMM results are arrived at
correctly given the provided code.

The paper attempts to answer the second question by applying the GLMMs they
already fit. To do so, it presents simulations from the fitted GLMM. It then
uses the simulated results to estimate the probability of observing reaction
times below a set of thresholds. I think it does a good job of contextualizing
these in terms of the number of races per an occurrence.
\end{comment}

Thank you for these encouraging comments.

\begin{comment}
However, I do believe the analysis to answer this first question could be
substantially improved. First, the paper only presents results for men running
the 110 meter hurdles. Women also competed at these events as they later
acknowledge in the discussion section. Why not use data from those competitors?
(There is a suggestion that the systematic effect seen at the 2022 World
Championships for men would not/did not appear for women because they may have
slower reaction times. No analysis is provided to support that though. Could
that be provided?) Similarly, other sprinting events were held using the same
block-sensor technology. Were similar patterns present in the 100-meter and
200-meter dashes? What about in the longer 400-meter dash and 400-meter hurdle
races? I think expanding the sample here would yield helpful insights into the
systematic nature of the problem the paper is trying to investigate. These
expanded samples could improve both the ranked-test results and the GLMM results
and improve the generalizability of the conclusions.
\end{comment}

One of the reasons why we did not include data from other sprinting evetns is
because we wanted to ground the article in Devon Allen's story and to be about
the hurdles events.  Athletes who compete in the longer events such as 400 meters
may not have as good of a start as those in the 100 meter events because reaction
time plays a much smaller role in the outcome of the race. While it is common
for amateur and collegiate athletes to compete in multiple sprinting events, it
seems reasonable that a world championship 100 meter runner may be only a top 20
200 meter runner.  We are trying to estimate the probabilty of observing a extreme
reaction time to determine whether 0.1 seconds is fair but if we include data
from competitions with slower reaction times, then 0.1 seconds is going to be
more extreme.

\begin{comment}

Relatedly, I believe the GLMM results could be improved by considering a few
more factors. Currently, the GLMMs regress reaction time onto models that only
contains predictors for year and whether or not an observation comes from a
preliminary heat or a final (both as random intercepts). It strikes me that the
authors may also want to account for confounding factors such as the overall
speed-level of the heat (especially relative to that point in time) and the
speed of the given runner relative to the others in the heat, as these factors
are likely to inform whether a runner tries to push their reaction to the limit
or not and are potentially correlated with year and race type, biasing the
results. Accounting for these additional features (in addition to expanding the
analysis) should help clarify if there was a specific difference attributable to
2022.
\end{comment}

The heat effect measures the speed-level of the heat. 
Perhaps we need to better explain the heat effect in the methods section.

\begin{comment}
Otherwise, I think the manuscript could be improved by streamlining the
presentation of the organizing material like the problem statements and theory.
The literature review currently reads like an itemized list of all the relevant
studies; revising to focus on the specific arguments the paper hopes to make
could help improve clarity. As it is, I am not entirely sure if the paper is
specifically trying to test a hypothesis offered by others (e.g. the sensors
were calibrated in some way to lead to faster reaction times) or just present a
general framework for studying track and field down the road. I believe the
former is stronger and of more interest than the latter, so making clear the
purpose would help.
\end{comment}  
We do not have enough evidence to conclude that the sensors were calibrated
differently, so thus we make general recommendations for World Athletics.

\subsection*{Reviewer2}

\begin{comment}
The paper’s topic is very important, but has significant issues with the writing
quality and statistical rigor. Right now, the manuscript reads more like report
than a paper. The text lacks significant direction, and does not explain the
goals of the study until the 4th page. Even then, “checking for abnormalities”
is very vague.

\end{comment}

We can restructure the paper but doesn't the abstract mention the goals of the
paper? I suppose the introduction is really more of a background section.

\begin{comment}
Concerns about statistical tests:

I am not convinced that the rank-based comparison test is the best choice here.
There are two questions: 1) Was the venue for the 2022 championships faster than
usual? 2) Was the final heat for the 2022 championships faster than usual?

If you are trying to answer question (1), you should include reaction times from
every spring event, not just the hurdles. If you are trying to answer question
2, why not do the following Monte Carlo test:

1. Sample a single reaction time for each athlete in the final.
2. Calculate the mean reaction time for the simulated final.
3. Compute the probability of seeing the observed average reaction time deviate
as much as it did in the 2022 Championships.

This test seems better than a rank sum test, as ranks do not capture the
differences between times.

\end{comment}

From what distribution are we sampling athlete's reaction times from?

\begin{comment}
I am also not convinced that the GLMM methods used for evaluating the
probability of disqualification are valid. The tests use data of reaction times
from performances that were NOT disqualified.  The data is therefore truncated
with no samples from the “left” or “disqualified” tail. This means that any
conclusions about the probability of a reaction time falling below this
threshold is purely extrapolation from the distributional assumptions. The
authors use a gamma distribution based on its fit to the data they have. In my
opinion, the choice of a gamma distribution is not sufficiently justified to
make extrapolating claims about the probability mass in the upsampled tail. I am
not convinced there is enough data to make a strong argument here, especially
when the previous literature on reaction time uses different distributions.
\end{comment}
Athletes may be disqualified based on PED use, and athletes may not finish a race
due to injury, thus we did not consider these times.  However, it does seem
reasonable to include times from people who were disqualified based on their
reaction time.

In terms of the choice of distribution, Brosnan and Harrison used a exponentially
modified gaussian distribution using athletes sex, ruling periods, competition
rounds. The article is now behind a pay wall.

\begin{comment}
The paper also mentions another claim that I think would be worth investigating:
Devon Allen’s reaction times might be faster due to innate ability. I think it
would be nice to show that his average reaction time faster, and that this
difference is statistically significant. It would also be nice to see how Devon
Allen’s DIFFERENCE in reaction time compares to the differences of the other
competitors (relative to previous heats, or averages, or whatever). If Devon
Allen’s time improved anomalously in the final heat, this would suggest a false
start. Based on the fact that he recorded a .101 reaction time earlier, I am
guessing he did not improve anomalously.  
\end{comment}

Do we want to perform some formal statistical analysis of Allen's times? What
data do we want to use? I don't think it makes sense to use Allen's times at
lower level competitions versus sprinters at the world championships.

\begin{comment}
A point that is not sufficiently emphasized is that the track officials consider
low reaction times to be false starts because an athlete can ANTICIPATE a start
faster than they can REACT to the starting gun. I think this is important to
discuss so that a general audience understands why an athlete might be
disqualified even through they started running after the race began.  
\end{comment}

This can be pretty easily addressed with a 1-2 sentence explanation in the
introduction.

\begin{comment}I have no idea how to interpret the numbers displayed in Table 2. These
need to be made interpretable, or they add nothing to the paper.
\end{comment}

Table 2 is the table with the parameters from the models.  This can be moved to
the appendix? We pretty explicity define all the terms in the methods section.

\begin{comment}
Line 127: Is there a reason you are commenting on the difficult of data
gathering? I am not sure it adds much.
\end{comment}

Can be removed.

\begin{comment}
Line 141: “We compare the times” - do you mean reaction times?
\end{comment}

Yes, this can be easily resolved.

\begin{comment}
Figure 3 is extremely inefficient. Maybe show the deltas on one plot? I'm not
sure a plot is even needed. What is the goal here? If you want to show that 2022
is anomalous you need to show where it stands in the distribution. Trying to
show that 2022 is anomalous based on a shift in the calculated venue effects
seems very roundabout.
\end{comment}

This was meant to be used as further evidence to show 2022 is an outlier. Perhaps
we delete this graph? I am unsure what the reviewer means by where "2022 stands
in the distribution"

\subsection*{Associate Editor}
\begin{comment}
I encourage you to consider the issues raised by the reviewers carefully. I have
little additional material to add, other than to emphasize the need to
re-organize the manuscript to more clearly identify the key goals of the work at
the start, and to provide a more concise and focused presentation of the key
findings.
\end{comment}

%\bibliographystyle{chicago}
%\bibliography{citations}
\end{document}

### 2024-06-13 Manuscript/manuscript.tex
Added explanation of venue/heat effect

@@
 much inferior to the gamma model which better captures the skewness of the
-data. Two random effects were found useful: one capturing the venue effect in
-each year and the other capturing the heat effect within each year. The heat
-effect is important as it captures the variability in the amount of time
-athletes are on the starting blocks before the gun goes off. Thus, we can try
-to reduce this variability by accounting for it in the model through the heat
-effect.
+data. Two random effects were found useful: one capturing the venue effect, which
+is used to contrast years, and the second being the heat effect, where every
+race was given a unique identifier with typically five to nine observations
+per race. The heat effect is important as it captures the variability in the
+amount of time athletes are on the starting blocks before the gun goes off.
+Thus, we can try to reduce this variability by accounting for it in the model
+through the heat effect.
 

### 2024-06-13 Manuscript/reply.tex
added all comments

@@
 \begin{comment}
-
 Relatedly, I believe the GLMM results could be improved by considering a few
@@
 
-The heat effect measures the speed-level of the heat. 
-Perhaps we need to better explain the heat effect in the methods section.
+The heat effect measures the speed-level of the heat.  We added some additional
+explanations to make it clear what the venue and heat effect refer to.
 
@@
 purpose would help.
+
 \end{comment}  
@@
 differently, so thus we make general recommendations for World Athletics.
+
+\begin{comment}
+Altogether, I see a lot of promising work here and think it would generally
+benefit from expanded scope and analysis.
+\end{comment}
+
+Thank you for these encouraging words.
 

### 2024-06-13 Code/GLMM.Rmd
added 100m dash

@@
 ```
+
+### 18
+We now examine the men's 100 meter dash times from 1999 to 2023.
+```{r}
+times <- read.csv("../Data/GLMMData.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
+dash_2022 <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != "DNS")
+                   & Event == "Dash")
+dash_no_2022 <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != "DNS")
+                   & Event == "Dash" & Venue != 2022)
+```
+
+```{r}
+library(tidyverse)
+dash_2022 <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+
+dash_no_2022 <- times |>
+  filter(Venue != 2022) |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+```
+
+### 19
+Baseline models
+```{r}
+library(lme4)
+dash_no_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash_no_2022, 
+                                    family = Gamma(link = "log"))
+dash_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash_2022, 
+                                family = Gamma(link = "log"))
+```
+
+### 20
+This code creates the venue effect only models.
+```{r}
+dash_no_20222022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_no_2022, 
+                                       family = Gamma(link = "log"))
+dash_2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_2022, 
+                                   family = Gamma(link = "log"))
+```
+
+### 21
+This creates the heat effect only models.
+```{r}
+dash_no_2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = dash_no_2022, 
+                             family = Gamma(link = "log"))
+dash_2022_heat_gamma_log <-glmer(RxnTime ~ (1 | Heat), data = dash_2022, 
+                         family = Gamma(link = "log"))
+```
+
+### 22
+This creates the models with a venue effect and heat effect.
+```{r}
+dash_no_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
+                                   data = dash_no_2022,
+                                   family = Gamma(link = "log"))
+dash_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
+                               data = dash_2022,
+                               family = Gamma(link = "log"))
+```
+
+### 23
+Now we test
+```{r}
+sim_two_effect_time(dash_no_2022_two_effect_gamma_log,
+                    dash_no_2022_one_gamma_log, 0.1)
+sim_two_effect_time(dash_2022_two_effect_gamma_log,
+                    dash_2022_one_gamma_log, 0.1)
+```

### 2024-06-13 Archive/Plots.Rmd
added 100m dash

--- 2023-10-12 Code/Plots.Rmd
@@
 ```{r}
-times <- read.csv("../Data/Thesis2023.csv")
-colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
+times <- read.csv("../Data/GLMMData.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
 ```
@@
 ```{r}
-with2022 <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ')
+with2022 <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Event == "Hurdles")
 library(ggplot2)
@@
                                               breaks = seq(1999, 2024, 2))
-ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
+#ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
 ```
@@
 library(dplyr)
-times <- read.csv("../Data/Thesis2023.csv")
-colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
+times <- read.csv("../Data/GLMMData.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
 with2022 <- subset(times, (Type == "F" | Type == "S") & 
-                       Gender == "M" & Time != 'DQ')
+                       Gender == "M" & Time != 'DQ' & Event == "Hurdles")
 with2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
@@
 
-venue_effects_with2022<- ggplot(venues, aes(x=year, y = condval)) + geom_point() + 
+ggplot(venues, aes(x=year, y = condval)) + geom_point() + 
   labs(x = "Venue", y = "Venue Effect") +
@@
 ```{r}
+x_values <- seq(-3, 3, length.out = 1000)  # Generate a sequence of x values
+y_values <- dnorm(x_values, mean = 0, sd = sqrt(0.04))  # Calculate the corresponding y values for the normal distribution
+
+normal_df <- data.frame(x = x_values, y = y_values)
+ggplot(normal_df, aes(x = x, y = y)) +
+  geom_line(color = "black", size = 1) +
+  geom_point(data = venues, aes(x = condval, y = 0), color = "red", size = 2) +
+  labs(title = "Normal Distribution (Mean = 0, SD = sqrt(0.4))",
+       x = "Value",
+       y = "Density") +
+  theme_minimal() + xlim(-1, 1)
+```
+
+```{r}
 library(lme4)
@@
 library(dplyr)
-times <- read.csv("../Data/Thesis2023.csv")
-colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
+times <- read.csv("../Data/GLMMData.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
 without2022 <- subset(times, (Type == "F" | Type == "S") & 
-                           Gender == "M" & Time != 'DQ' & Venue != 2022)
+                           Gender == "M" & Time != 'DQ' & Venue != 2022 & Event == "Hurdles")
 without2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
@@
 ```
+
+### 100 meter dash plots
+```{r}
+dash <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != "DNS")
+                   & Event == "Dash")
+library(ggplot2)
+dash_plot <- ggplot(dash, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0, 0.25),
+                            breaks = seq(0, 0.25, 0.05)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
+                                              breaks = seq(1999, 2024, 2))
+
+hurdle_plot <- ggplot(with2022, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0, 0.25),
+                            breaks = seq(0, 0.25, 0.05)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
+                                              breaks = seq(1999, 2024, 2))
+library(gridExtra)
+grid.arrange(dash_plot, hurdle_plot, ncol = 1)
+```
+```{r}
+dash <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != "DNS")
+                   & Event == "Dash")
+library(ggplot2)
+dash_plot <- ggplot(dash, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.1, 0.25),
+                            breaks = seq(0.1, 0.25, 0.05)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
+                                              breaks = seq(1999, 2024, 2))
+
+hurdle_plot <- ggplot(with2022, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.1, 0.25),
+                            breaks = seq(0.1, 0.25, 0.05)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
+                                              breaks = seq(1999, 2024, 2))
+library(gridExtra)
+grid.arrange(dash_plot, hurdle_plot, ncol = 1)
+```

### 2024-06-20 Manuscript/reply.tex
explained more stuff and responded to comments

@@
 The heat effect measures the speed-level of the heat.  We added some additional
-explanations to make it clear what the venue and heat effect refer to.
+explanations to make it clear what the venue and heat effect refer to. We did
+not want to call it a "race effect" as we were worried that may be interpreted
+to mean the race of the athlete.
 
@@
 
-Can be removed.
+Yes, the comments about the difficulty of data gathering are meant to demostrate
+that we worked hard to find as much data as we could and as a note to World
+Athletics (track and field governing body) that they need to do a better job
+at data standardization.  Currently the races on their website do not include
+reaction times, which meant that we had to look in alternative places to perform
+the analysis that we did.
 
@@
 
-Yes, this can be easily resolved.
+Yes, this has been resolved.
 
@@
 
-This was meant to be used as further evidence to show 2022 is an outlier. Perhaps
-we delete this graph? I am unsure what the reviewer means by where "2022 stands
-in the distribution"
+We included this plot to show the venue effect for each year and to visually
+demonstrate how 2022 compares to other years.  We added additional comments
+about the probability of observing an extreme venue effect to provide
+quantitative data supporting our argument.
 

### 2024-06-20 Manuscript/manuscript.tex
added explanation of probability for venue effect graph

@@
 effect only captures a smaller part of the variation than the heat effect as
-evident from their standard deviations. The venue effect for 2022 is
-lower than that for any other year, but may not be as extreme as one would
-think considering the history of reaction times going back to 1999.
+evident from their standard deviations. It is possible to calculate the extremity
+of the 2022 heat effect as we know that the mean of the heat effects is $0$,
+the standard deviation is $0.42$ as given by Table~\ref{tab:Gamma_parameters},
+and the estimate of the 2022 heat effect is $-0.93$. We find the tail probability
+to be $0.0128$ indicating there is a low chance of observing a venue effect as
+extreme as 2022.
 

### 2024-06-20 Manuscript/manuscript.tex
added explanation of why 0.1 sec reaction barrier

@@
 heats and semifinals to reach the final heat while posting reaction times of 
-0.123 and 0.101 in his preliminary heat and semifinal, respectively. 
+0.123 and 0.101 in his preliminary heat and semifinal, respectively.  The
+reaction time barrier for all sprint related events is 0.1 seconds as set by the
+International Association of Athletics Federation (IAAF) and is meant
+to prevent athletes from trying to predict the starting gun.
 His disqualification in the final heat by reacting in 0.099 seconds, just 0.001
-seconds faster than the 0.1 threshold set by the International Association of
-Athletics Federations (IAAF), was met
+seconds faster than the 0.1 threshold, was met
 with shock from both Allen and fans alike. NBC, who was broadcasting the

### 2024-06-20 Code/GLMM.Rmd
Added womens and dash analyses

@@
 times <- read.csv("../Data/GLMMData.csv")
-colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
 with2022 <- subset(times, (Type == "F" | Type == "S") & 
-                       Gender == "M" & Time != 'DQ')
+                       Gender == "M" & Time != 'DQ' & Event == "Hurdles")
 without2022 <- subset(times, (Type == "F" | Type == "S") & 
-                           Gender == "M" & Time != 'DQ' & Venue != 2022)
+                           Gender == "M" & Time != 'DQ' & Venue != 2022 & Event == "Hurdles")
 ```
@@
 ```{r}
-dash_no_20222022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_no_2022, 
+dash_no_2022_venue_gamma_log <- glmer(RxnTime ~ (1|Venue), data = dash_no_2022, 
                                        family = Gamma(link = "log"))
@@
 ```
+## 24
+In response to reviewer 2's comments that questioned the probability of observing
+the venue effect for 2022 we perform the following analysis. We store the venue
+effect for 2022 in `effect2022` and we store the standard deviation in `sd2022`
+```{r}
+library(tidyverse)
+ranef_results <- data.frame(ranef(with2022_two_effect_gamma_log))
+venues <- subset(ranef_results, ranef_results["grpvar"] == "Venue")
+venues$year <- as.character(venues$grp)
+venues$year <- as.numeric(venues$year)
+effect2022 <- venues |> filter(grp == 2022) |> select(condval) |> pull()
+sd2022 <- as.data.frame(VarCorr(with2022_two_effect_gamma_log)) |> filter(grp == "Venue") |> select(sdcor)|> pull()
+
+```
+Now we can calculate how extreme this value is
+```{r}
+pnorm(effect2022, mean = 0, sd = sd2022)
+```
+
+## Now we look at the 200 meter dash results
+```{r}
+times <- read.csv("../Data/GLMMData.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
+```
+
+```{r}
+library(tidyverse)
+dash200_2022 <- times |>
+  filter(Event == "200 Dash") |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+
+dash200_no_2022 <- times |>
+  filter(Event == "200 Dash") |>
+  filter(Venue != 2022) |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+```
+
+```{r}
+library(lme4)
+dash200_no_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash200_no_2022, 
+                                    family = Gamma(link = "log"))
+dash200_2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash200_2022, 
+                                family = Gamma(link = "log"))
+
+dash200_no_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
+                                   data = dash200_no_2022,
+                                   family = Gamma(link = "log"))
+dash200_2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
+                               data = dash200_2022,
+                               family = Gamma(link = "log"))
+```
+Now we test
+```{r}
+sim_two_effect_time(dash200_no_2022_two_effect_gamma_log,
+                    dash_no_2022_one_gamma_log, 0.1)
+sim_two_effect_time(dash200_2022_two_effect_gamma_log,
+                    dash_2022_one_gamma_log, 0.1)
+```
+These are low probabilities indicating that a reaction time below 0.1 seconds
+is unlikely for the 200 meter dash.
+```{r}
+library(ggplot2)
+ggplot(dash200_2022, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.1, 0.25),
+                            breaks = seq(0.1, 0.25, 0.05)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
+                                              breaks = seq(1999, 2024, 2))
+```
+
+## We will compare our results to the womens data starting with hurdles and then dash
+
+```{r}
+library(tidyverse)
+women_hurdles <- times |>
+  filter(Event == "100 Hurdles") |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "F") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+
+women_dash <- times |>
+  filter(Event == "100 Dash") |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "F") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+```
+
+```{r}
+library(ggplot2)
+ggplot(women_hurdles, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.1, 0.25),
+                            breaks = seq(0.1, 0.25, 0.05)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(2000,2024),
+                                              breaks = seq(2001, 2024, 2))
+```
+```{r}
+library(ggplot2)
+ggplot(women_dash, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.1, 0.25),
+                            breaks = seq(0.1, 0.25, 0.05)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
+                                              breaks = seq(1999, 2024, 2))
+```

### 2024-06-20 Code/RankComparison.Rmd
blinded supplemental files, removed noprelim data from readme
added womens 2022 vs 2023 data

@@
 title: "Rank Comparison"
-author: "Owen Fiore"
-date: "2023-11-03"
+author: "Blinded"
+date: "2023-12-01"
 output: pdf_document
@@
 ```
+
+Now we test the womens 2022 vs 2023
+```{r}
+womens <- read.csv("../Data/Clusrank2022vs2023Womens.csv")
+colnames(womens) <- c("Time", "Athlete", "Year")
+```
+
+The tests below are very similar to the ones above, the only difference being
+grouping by Year (0 if 2022 and 1 if 2023).
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = womens,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = womens,
+                method = "ds",
+                exact = TRUE, B = 500000)
+```

### 2024-06-20 Manuscript/reply.tex
responded to more comments

@@
 
-From what distribution are we sampling athlete's reaction times from?
+We have included other sprinting events in our analysis.
 
@@
 
-Do we want to perform some formal statistical analysis of Allen's times? What
-data do we want to use? I don't think it makes sense to use Allen's times at
-lower level competitions versus sprinters at the world championships.
+Unfortunately data such as this is not easy to come by.  We found as much as we
+could online, but there is simply not enough data to justify a formal analysis
+of his reaction times across lower level meets.
 
@@
 
-This can be pretty easily addressed with a 1-2 sentence explanation in the
-introduction.
+We added additional explanations to address this in the introductions section.
 
@@
 
-Table 2 is the table with the parameters from the models.  This can be moved to
-the appendix? We pretty explicity define all the terms in the methods section.
+The nomenclature for the greek variables in Table 2 are defined in the GLMM
+subsection of the methods section (3.2)
 

### 2024-07-09 Code/RankComparison.Rmd
added additional tests for dash results

@@
 ```
+
+Now we test for the men's dash
+```{r}
+dash <- read.csv("../Data/Clusrank2022vs2023Dash.csv")
+colnames(dash) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = dash,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = dash,
+                method = "ds",
+                exact = TRUE, B = 500000)
+```
+```{r}
+women_dash <- read.csv("../Data/Clusrank2022vs2023WomensDash.csv")
+colnames(women_dash) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash,
+                method = "ds",
+                exact = TRUE, B = 500000)
+```

### 2024-07-09 Code/NewAnalysis.Rmd
Here I test trying to pool dash and hurdles

---
title: "NewAnalysis"
author: "Owen Fiore"
date: "2024-07-03"
output: html_document
---

# New Work
To try and keep everything organized, we are going to use this document to test
anything additional before putting it in `GLMM.Rmd.`  We have several things
to test before our re-submission.

## Can we pool the dash and hurdles results?
We need to see how different our results are using a pooled version of the men's
100 meter dash and men's 110 meter hurdles.  We will also include a random effect
into our model to compensate for this.

```{r}
library(tidyverse)
times <- read.csv("../Data/GLMMData.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
```

```{r}
dash_and_hurdles <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime != 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))

hurdles <- dash_and_hurdles |>
  filter(Event == "110 Hurdles")
```

We consider the baseline models:
```{r}
library(lme4)
dash_and_hurdles_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles, 
                                    family = Gamma(link = "log"))
hurdles_one_gamma_log <- glm(RxnTime ~ 1, data = hurdles, 
                                family = Gamma(link = "log"))
```
To assess whether we we can pool the data, we are going to perform an anova test
to determine whether the most recently added coefficient to the model was
significant.

```{r}
dash_and_hurdles_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat) + (1 | Event), 
                                   data = dash_and_hurdles,
                                   family = Gamma(link = "log"))
dash_and_hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
                                                   (1 | Heat), 
                                   data = dash_and_hurdles,
                                   family = Gamma(link = "log"))

hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
                               data = hurdles,
                               family = Gamma(link = "log"))
```

```{r}
anova(dash_and_hurdles_two_effect_gamma_log, dash_and_hurdles_three_effect_gamma_log)
```
What we observe is that there is not a significant difference between these
models as the p value is quite large.

Lets compare the two models
```{r}
summary(dash_and_hurdles_three_effect_gamma_log)
summary(hurdles_two_effect_gamma_log)
```
Now we will see how the inclusion of the pooled data will influence the results
```{r}
sim_three_effect_time <- function(model, base, RxnTime) {
  simfit <- function(model, n = 10000000) {
    vc <- as.data.frame(VarCorr(model))
    sd_Venue <- vc$sdcor[1]
    sd_Venue_heat <- vc$sdcor[2]
    sd_event <- vc$sdcor[3]

    z1 <- rnorm(n, sd = sd_Venue)
    z2 <- rnorm(n, sd = sd_Venue_heat)
    z3 <- rnorm(n, sd = sd_event)
    shape <- 1 / sigma(model)^2
    mu <- exp(fixef(model) + z1 + z2 + z3)

    x <- rgamma(n, shape = shape, scale = mu / shape)
    x
  }

  sbase <- summary(base)
  x <- simfit(model)
  print(mean(x < RxnTime))
  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
  ## compare with and without random effect
  plot(density(x))  # with random effects
  curve(dgamma(x, shape = 1 / sbase$dispersion, 
               scale = exp(coef(base)) * sbase$dispersion), 
        0.08, 0.25, add = TRUE, col = "blue")
}
```

Now we test
```{r}
set.seed(1)
sim_three_effect_time(dash_and_hurdles_three_effect_gamma_log, dash_and_hurdles_one_gamma_log, 0.1)
```

### 2024-07-11 Code/NewAnalysis.Rmd
Added womens pooled results and explored 200m dash

@@
 ```
+## Now for women
+We want to make sure we get similar results when we pool the women's 100m hurdle
+and 100m dash. 
+```{r}
+times <- read.csv("../Data/GLMMData.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
+```
+
+Again we are going to run an anova test to determine how impactful the coefficient
+on `Event` is.
+```{r}
+women_dash_and_hurdles <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "F") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "100 Hurdles") |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+```
+
+```{r}
+women_dash_and_hurdles_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat) + (1 | Event), 
+                                   data = women_dash_and_hurdles,
+                                   family = Gamma(link = "log"))
+women_dash_and_hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat), 
+                                   data = women_dash_and_hurdles,
+                                   family = Gamma(link = "log"))
+```
+
+```{r}
+anova(women_dash_and_hurdles_two_effect_gamma_log, women_dash_and_hurdles_three_effect_gamma_log)
+```
+There seems to be more of a difference but the test is still not significant.
+
+## 200m test
+Does it make sense to include the 200m dash data as well?
+
+```{r}
+all_events <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles" | Event == "200 Dash") |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+```
+
+```{r}
+all_events_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat) + (1 | Event), 
+                                   data = all_events,
+                                   family = Gamma(link = "log"))
+all_events_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat), 
+                                   data = all_events,
+                                   family = Gamma(link = "log"))
+```
+
+```{r}
+anova(all_events_three_effect_gamma_log, all_events_two_effect_gamma_log)
+```
+Now we see that the event become significant
+
+```{r}
+library(car)
+TukeyHSD(aov(RxnTime ~ Event, data = all_events))
+```
+We see that the 200m dash differs significantly from 100m dash/hurdles

### 2024-07-11 Code/NewAnalysis.Rmd
Need to change random effect event to fixed effect

@@
 dash_and_hurdles_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
-                                                   (1 | Heat) + (1 | Event), 
+                                                   (1 | Heat) + Event, 
                                    data = dash_and_hurdles,
@@
 all_events_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
-                                                   (1 | Heat) + (1 | Event), 
+                                                   (1 | Heat) + Event, 
                                    data = all_events,

### 2024-08-09 Code/NewAnalysis.Rmd
Did some work and using this file for testing what data to use

@@
 What we observe is that there is not a significant difference between these
-models as the p value is quite large.
+models as the p value is quite large. Thus we will not include `Event` as a
+feature in our final model.
 
-Lets compare the two models
-```{r}
-summary(dash_and_hurdles_three_effect_gamma_log)
-summary(hurdles_two_effect_gamma_log)
-```
 Now we will see how the inclusion of the pooled data will influence the results
@@
 
-Now we test
+## 200m test
+Does it make sense to include the 200m dash data as well?
+
 ```{r}
-set.seed(1)
-sim_three_effect_time(dash_and_hurdles_three_effect_gamma_log, dash_and_hurdles_one_gamma_log, 0.1)
+all_events <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles" | Event == "200 Dash") |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
 ```
+
+```{r}
+all_events_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat) + Event, 
+                                   data = all_events,
+                                   family = Gamma(link = "log"))
+all_events_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat), 
+                                   data = all_events,
+                                   family = Gamma(link = "log"))
+```
+
+```{r}
+anova(all_events_three_effect_gamma_log, all_events_two_effect_gamma_log)
+```
+Now we see that the event become significant
+
+```{r}
+library(car)
+TukeyHSD(aov(RxnTime ~ Event, data = all_events))
+```
+Using a Tukey's Honestly Significant Difference test, we see that the 200m dash
+differs significantly from 100m dash/hurdles.
+
+Based on all the following evidence we are going to perform our men's analysis
+on data from the 100 meter dash and 110 meter hurdles exclusively.
+
 ## Now for women
@@
 There seems to be more of a difference but the test is still not significant.
-
-## 200m test
-Does it make sense to include the 200m dash data as well?
-
-```{r}
-all_events <- times |>
-  filter(Type == "F" | Type == "S") |>
-  filter(Gender == "M") |>
-  filter(Time != "DNS") |>
-  filter(RxnTime != 0) |>
-  filter(Event == "100 Dash" | Event == "110 Hurdles" | Event == "200 Dash") |>
-  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
-```
-
-```{r}
-all_events_three_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
-                                                   (1 | Heat) + Event, 
-                                   data = all_events,
-                                   family = Gamma(link = "log"))
-all_events_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
-                                                   (1 | Heat), 
-                                   data = all_events,
-                                   family = Gamma(link = "log"))
-```
-
-```{r}
-anova(all_events_three_effect_gamma_log, all_events_two_effect_gamma_log)
-```
-Now we see that the event become significant
-
-```{r}
-library(car)
-TukeyHSD(aov(RxnTime ~ Event, data = all_events))
-```
-We see that the 200m dash differs significantly from 100m dash/hurdles

### 2024-08-09 Manuscript/reply.tex
responded to some comments

@@
 comments quoted in \emph{\color{darkblue} italic and blue}.
-
-\subsection*{To Associate Editor}
-
-\begin{comment}
-Using Devon Allen's disqualification at the 2022 World Track and Field
-championships as a hook, this paper poses two questions. First, were reaction
-times at the 2022 World Track and Field Championships faster than they had been
-earlier in 2022 national-level competitions, the 2019 World Track and Field
-Championships, and the 2023 World Track and Field Championships? Second, based
-on modeled reaction times, is the current standard of a 0.1 second threshold for
-reaction times before disqualification too conservative?
-\end{comment}
-
-Thank you very much for your encouraging comments and great suggestion.  
-We have now included a new section, \textit{Section 5 Classroom Implementation},
-that discusses how elements of this manuscript have been incorporated into both 
-graduate- and undergraduate-level teaching at the authors' home university.
-
-For blinding purposes, course numbers, author initials and websites, and 
-university name have been redacted, but all this information appears in the 
-unblinded version.  In brief, this material has been covered at the graduate
-level in a statistical computing course, open to all graduate students in 
-Statistics and others with permission, that covers the bootstrap. Detailed R 
-code and graphical illustrations were provided to students in the open-access 
-course notes to elucidate the distinctions between employing null
-hypotheses with and without defined parameters.  This link is included in the 
-unblinded version of the paper; for the blinded review, we printed the 
-relevant portion of the notes to pdf and included this blinded pdf file in the 
-Supplemental Materials. %\eds{Not sure how we can blind your github 
-%website for the submission unless we actually don't include it in the blinded 
-%version. Could we print to PDF the relevant section of the notes and include as 
-%a supplement for review purposes?}
-
-Similarly at the undergraduate-level, students in an R-based applied regression 
-course were provided with an R Markdown file with simulation code in order to 
-demonstrate how the KS test fails with estimated parameters in the hypothesized 
-distribution. The R Markdown file that was provided to students is included in 
-the Supplemental Materials. This R-based applied linear regression course 
-has the following prerequisites: two semesters of introductory statistics OR a 
-Calculus~2-based statistical methods course.
-
-At both the graduate and undergraduate level, students were able to follow the 
-demonstrations closely and gain a clear understanding of the implications of 
-using estimated parameters within test statistics.  At the graduate level, 
-students further were able to appreciate the efficacy of the parametric 
-bootstrap method as a corrective measure.
 
@@
 
-We have included other sprinting events in our analysis.
+We are attempting to answer question 1 and have included reaction times from
+additional sprinting events from World Championships.  We hypothesized that
+as the length of the race increased, reaction time would be less impactful and
+thus not as important. We started by adding data for the 100 meter dash and that
+is the most similar event in terms of length to the 110 meter hurdles.  When we
+added an additional variable to the model based on whether the observation came
+from a dash event or hurdles event the coefficient was not significant.  An 
+ANOVA test likewise produced a not signficant coefficient indicating to us that
+it seems reasonable to pool the 100 meter dash and 110 meter hurdles.  We
+hypothesized that longer sprinting events (200 meter dash, 400 meter dash, 400
+meter hurdles) would be more difficult to add as those are longer races where
+athletes reactions times are not as important to their overall time in the race.
+When we repeat the procedure we did for the 100 meter dash, now looking at data
+that included the 100 meter dash, 110 meter hurdles, and the 200 meter dash we
+found that the coefficient on event type was signficant, and a p value of 0.01586.
+When we peformed a TukeyHSD test to compare the times in the three categories:
+100 meter dash, 110 meter hurdles, 200 meter dash, we found a p-value of
+0.0020625 comparing the means of the 200 meter dash and the 110 meter hurdles
+and a p-value of 0.0000174 comparing the means of the 100 meter dash and the 200
+meter dash.  This indicated to us that the reaction times for the 200 meter were
+too slow to fit into our study on extreme reaction times and thus we chose to
+exclude 200 meter data.
 
@@
 reaction times, which meant that we had to look in alternative places to perform
-the analysis that we did.
+the analysis that we did.  The data we found from national level meets in 2022
+was difficult to come by as it required searching for every country's championship
+results individually and when we did find it, it was often in the native language
+making it harder to understand and interpret.
 
@@
 
+\begin{comment}
+We have taken additional steps to streamline the manuscript and enhance the
+presentation of our work. Thank you for your time and we appreciate re-considering
+our submission.
+\end{comment}
+
 %\bibliographystyle{chicago}

### 2024-09-20 Manuscript/manuscript.tex
Reworked introduction, need to work on discussion still

@@
 
-Reaction time in sprint and hurdle events has been studied by many authors.
+The first objective of this paper is to determine whether there were 
+abnormalities with athlete's reaction times at the 2022 World Track and Field
+Championship.  Our analysis consists of quantitatively comparing reaction times for athletes who 
+have competed at additional competitions to the 2022 World Track and Field 
+Championship (Similar to a matched pairs design). By comparing reaction times from the same athlete across multiple
+competitions we are able to determine if the 2022 reaction times differed 
+significantly from those of other competitions.  This investigation, while
+motivated by Allen's disqualifcation has wider impacts across track and field
+that question if there is more variation than previously thought in track and
+field timing equipment.
+
+The second objective of this paper is to analyze whether the current
+reaction time barrier of 0.1 seconds is fair, and in this regard there have been
+many publications debating the human capacity for reaction times 
+\of{This sounds funny but I'm not sure how to fix it}.
 \citet{pain2007sprint} examined reaction times for nine male athletes. The
@@
 react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
-Unfortunately, IAAF has not changed the reaction barrier since it was
+These papers support increasing the reaction time barrier as they determined
+that athletes are better than how IAAF thinks they are. Unfortunately, IAAF has
+not changed the reaction barrier since it was
 implemented at the 1999 World Track and Field Championships.
-It is unknown how many athletes since the 2009 study could have potentially
-benefited from this, but it is probable that there are others besides Allen who
-could have. Additionally,
-\citet{komi2009iaaf} recommended that high speed cameras be used to
+Additionally, \citet{komi2009iaaf} recommended that high speed cameras be used to
 change the reaction time criteria from pushing off on the block to body
@@
 The possibility of using a video-based medium has been shown to be a reliable
-measure of reaction time \citep{mudric2015evaluation}.
+measure of reaction time.
 
-\citet{pilianidis2012start} found
-significant differences between reaction times at several world championships
-between 1997 and 2009 for the men's 110 meter hurdles. Another important
-conclusion was that reaction times did not decrease over the twelve years
-covered by the research in the study \citep{pilianidis2012start}. This 
-contradicts the results of \citet{zhang2021correlation}, which indicated that 
-year of competition has a significant effect on reaction times. Their
-analysis looked at data from the 2011-2019 World Championships across multiple
-sprint events and concluded that from 2011 to 2019, reaction times got
-faster. However, not all studies have been in favor of a decrease in the
-reaction time barrier. \citet*{brosnan2017effects} argue for increasing 
+However, not all studies have been in favor of a decrease in the
+reaction time barrier \citet*{brosnan2017effects} argue for increasing 
 the reaction time barrier to 0.115 seconds, and cited evidence from European 
@@
 at the 2008 Olympics likewise supported the conclusion that athletes could not
-react in as little as 0.1 seconds \citep{lipps2011implications}. Another study
-concluded that visual reaction times to the starting gun may be faster by 0.007
-seconds than the IAAF method, which is from force on block starts 
-\citep{holmes2018method}. While this may seem trivial, Allen was disqualified
-only by 0.001 seconds. It is also worth noting that one study looking at data 
-from various IAAF meets found that the fastest reaction times for men occurred 
-between the ages of 26 and 29 \citep{tonnessen2013reaction}. This is notable 
-because Allen was 27 years old at the 2022 World Track and Field Championships 
-and likely at his physical peak.
+react in as little as 0.1 seconds \citep{lipps2011implications}. However it is
+worth noting that \citet*{brosnan2017effects} and \citep{lipps2011implications}
+were retroactive studies of athletes at events while \citet{pain2007sprint} and
+\citet{komi2009iaaf} were experiments showing that there are various methods
+of testing reaction time capability.
 
-The objective to this paper is to analyze the reaction times at the 2022 World 
-Track and Field Championships to check for any abnormalities, and we approach
-this in two different ways. First we compare reaction times for athletes who 
-have competed at additional competitions to the 2022 World Track and Field 
-Championship. By comparing reaction times from the same athlete across multiple
-competitions we are able to determine if the 2022 reaction times differed 
-significantly from those of other competitions. Then we look at a longer time horizon
-and model the reaction times of every non-disqualification since the 1999 World 
+The second part of this paper will model the reaction times of every non-disqualification since the 1999 World 
 Championships, using a gamma linear mixed-effects model with two random effects: 
@@
 time being below a certain threshold (e.g., 0.1).
-Our investigation into reaction times is important not only because of its
-implications for future races, but also because reaction times have
-been shown to affect the overall time of the race \citep{delalija2008reaction}.
-Thus, faster reaction times lead to faster overall times, which is the goal of
-every sprinter who steps onto the track.
 

### 2024-09-20 Manuscript/manuscript.tex
rewrote abstract and intro

@@
 
-%Abstract should be 200 words
-%Devon Allen was disqualified in 2022... discussions about dq, talk methods and 
-%models, venue effect, and conclusion
+% Abstract should be under 200 words
+
 \begin{abstract}
-Devon Allen was disqualified in the men's 110 meter hurdle final of the 2022
-World Track and Field Championships after registering a reaction time of 0.099 
-seconds, 0.001 seconds faster than what is allowed. Following the games, 
-bloggers on the running website LetsRun concluded that the reaction times  
-from the 2022 World Championships seemed to be generally faster compared to the 
-other reaction times they considered, but they did not perform any formal 
-statistical analysis. This paper questions if athletes who competed at the 2022
-World Track and Field Championship had similar times at other competitions and
-if 0.1 seconds is a reasonable disqualification barrier. To do so, we employ a 
-signed-rank test for clustered data to compare reaction times for the same 
-athletes at different competitions and a generalized linear mixed model 
-with random venue (year) effect and a random heat effect within venue 
-in order to model the reaction time data. This matter needs to be addressed 
-because disqualification based on allowable reaction time will continue to be an
-issue in future world championships.
+Devon Allen was disqualified from the men's 110-meter hurdle final at
+the 2022 World Track and Field Championships after recording a
+reaction time of 0.099 seconds—just 0.001 seconds faster than the
+allowable threshold of 0.1 seconds. This paper explores whether the
+reaction times of athletes at the 2022 Championships were unusually
+fast compared to other competitions and evaluates the appropriateness
+of the 0.1-second disqualification barrier. To investigate these
+issues, we employ a signed-rank test for clustered data to compare
+athletes' reaction times across multiple competitions, along with a
+generalized linear mixed model (GLMM) that accounts for both venue and
+heat effects. The results indicate that athletes at the 2022
+Championships had significantly faster reaction times on average. This
+analysis raises important questions about the validity of the current
+reaction time threshold and suggests potential modifications to the
+rules that could prevent future disqualifications under similar
+circumstances.
 
 \bigskip\noindent{\sc Keywords}:
-False Start, Hurdles, Reaction Time, Seiko
+False Start, GLMM, Hurdle, Reaction Time, Seiko, rank-based test
 
@@
 
-Devon Allen, an alumnus of the University of Oregon in Eugene, Oregon, was
-expected to be a contender for the 2022 World Track and Field Championships 110 
-meter hurdle. After running the event in 12.84 seconds in early June 2022 
-(only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
-in third at the United States Track and Field Championships also in Eugene 
-in late June, Allen was poised to compete in the 2022 World Track and 
-Field Championships in August, which would also be held in Eugene, Oregon.
-At the World Championships, Allen made it through the preliminary
-heats and semifinals to reach the final heat while posting reaction times of 
-0.123 and 0.101 in his preliminary heat and semifinal, respectively.  The
-reaction time barrier for all sprint related events is 0.1 seconds as set by the
-International Association of Athletics Federation (IAAF) and is meant
-to prevent athletes from trying to predict the starting gun.
-His disqualification in the final heat by reacting in 0.099 seconds, just 0.001
-seconds faster than the 0.1 threshold, was met
-with shock from both Allen and fans alike. NBC, who was broadcasting the
-championships in the United States, later uploaded a video to their YouTube
-channel that showed the disqualification
-(\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}).
-The crowd erupted with jeers
-when they realized that Allen was disqualified, angry at the result. Had Allen 
-reacted just 0.001 seconds slower he would have been allowed to compete.
+Devon Allen, a University of Oregon alumnus, was a strong 
+contender for the 2022 World Track and Field Championships 
+in the 110-meter hurdle event. After running the event in 
+12.84 seconds in June 2022, just 0.04 seconds short of the 
+world record \citep{wa2022preview}, and placing third at 
+the U.S. Track and Field Championships, Allen was well-prepared 
+for the World Championships in August, also held in Eugene, 
+Oregon. He advanced through the preliminary heats and semifinals 
+with reaction times of 0.123 and 0.101 seconds, respectively. 
+However, in the final heat, Allen was disqualified for a 
+reaction time of 0.099 seconds, just 0.001 seconds faster 
+than the 0.1-second threshold set by the International 
+Association of Athletics Federations (IAAF). This decision 
+was met with significant public outcry \citep{nbc2022devon}.
 
-Allen's disqualification has been discussed extensively in online communities
-such as \url{www.LetsRun.com}, a website that is part message board and part
-news. The message board,
-which functions similarly to Reddit, is very active during important running
-events, including during the finals of major events such as
-the Olympics and the World Track and Field Championships. However, it is the
-news section of the website that generated articles such as \citet{johnson2022data}
-and \citet{johnson2022was} from  Robert Johnson, who created and runs 
-the LetsRun website. He argued that there was an error
-with the timing equipment at the 2022 World Track and Field Championships and
-cited graphs, computed descriptive statistics such as median reaction times, and 
-compared the reaction times at the 2022 World Track and Field Championships to 
-other competitions
-such as the United States Track and Field Championships and other World Championships.
-The comparisons lacked formal statistical analysis, however. 
-The alleged 99.95\% chance Allen was wronged 
-came from Johnson claiming there is a $(1/13)^3$ chance that the fastest 
-reaction times over the past 13 competitions of the men's 100 meter, men's 110 
-meter hurdle, and women's 100 meter would all come from the same year 
+Allen's disqualification has been discussed extensively in online 
+communities such as \url{www.LetsRun.com}, a website that functions 
+as both a message board and a news outlet. The message board, 
+similar to Reddit, is very active during major running events, 
+including the Olympics and the World Track and Field Championships. 
+The news section of the website featured articles such as 
+\citep{johnson2022data} and \citep{johnson2022was} by Robert Johnson, 
+the site’s founder. Johnson argued that there may have been an 
+error with the timing equipment at the 2022 Championships. He 
+cited graphs, descriptive statistics like median reaction times, 
+and compared the reaction times at the 2022 Championships to 
+those from other competitions, including the U.S. Championships 
+and previous World Championships. However, these comparisons 
+lacked formal statistical analysis. Johnson estimated a 99.95\% 
+chance that Allen was wrongfully disqualified, based on his claim 
+that the probability of the fastest reaction times from the men's 
+100 meter, men's 110 meter hurdle, and women's 100 meter all 
+occurring in the same year was approximately $(1/13)^3$ 
 \citep{johnson2022was}.
 
-The first objective of this paper is to determine whether there were 
-abnormalities with athlete's reaction times at the 2022 World Track and Field
-Championship.  Our analysis consists of quantitatively comparing reaction times for athletes who 
-have competed at additional competitions to the 2022 World Track and Field 
-Championship (Similar to a matched pairs design). By comparing reaction times from the same athlete across multiple
-competitions we are able to determine if the 2022 reaction times differed 
-significantly from those of other competitions.  This investigation, while
-motivated by Allen's disqualifcation has wider impacts across track and field
-that question if there is more variation than previously thought in track and
-field timing equipment.
+The first objective of this paper is to determine whether athletes’ 
+reaction times at the 2022 World Track and Field Championships 
+were significantly different from other competitions. Our analysis 
+compares reaction times for athletes who competed in multiple events, 
+using a matched-pairs design. By comparing the same athletes' reaction
+times across different competitions, we can assess whether the 2022 
+reaction times differed significantly. While this investigation was 
+motivated by Allen's disqualification, it also addresses broader concerns 
+about potential variation in timing equipment across different venues
+in short-distance dash or hurdle.
 
-The second objective of this paper is to analyze whether the current
-reaction time barrier of 0.1 seconds is fair, and in this regard there have been
-many publications debating the human capacity for reaction times 
-\of{This sounds funny but I'm not sure how to fix it}.
-\citet{pain2007sprint} examined reaction times for nine male athletes. The
-fastest athlete of the nine had, on average (even with their two fastest times 
-removed), a reaction time of 0.087 seconds and a standard deviation of 0.004
-seconds. Two other athletes were able to react similarly when under certain
-muscular tightening conditions \citep{pain2007sprint}. 
-In a study commissioned by IAAF,
-\citet{komi2009iaaf} recommended that the disqualification 
-barrier be decreased to allow sprinters to
-react faster, and suggested 0.08 or 0.085 seconds as possible new thresholds.
-These papers support increasing the reaction time barrier as they determined
-that athletes are better than how IAAF thinks they are. Unfortunately, IAAF has
-not changed the reaction barrier since it was
-implemented at the 1999 World Track and Field Championships.
-Additionally, \citet{komi2009iaaf} recommended that high speed cameras be used to
-change the reaction time criteria from pushing off on the block to body
-movement. This could provide a more accurate representation of reaction time, 
-and thus the possibility of a false disqualification would be lowered.
-The possibility of using a video-based medium has been shown to be a reliable
-measure of reaction time.
+Given the observed variation in reaction times across venues, the 
+question arises: is the 0.1-second disqualification threshold truly 
+justified? Studies have presented evidence supporting both sides of 
+the debate. \citet{pain2007sprint} examined reaction times in nine 
+male athletes and found that the fastest athlete had an average 
+reaction time of 0.087 seconds, with a standard deviation of 0.004 
+seconds. Similarly, \citet{komi2009iaaf}, in a study commissioned by 
+the IAAF, recommended lowering the disqualification threshold to 0.08 
+or 0.085 seconds, arguing that athletes are capable of faster reaction 
+times than the current threshold allows. On the other hand, some 
+researchers have argued for increasing the reaction time threshold. 
+\citet{brosnan2017effects} suggested raising it to 0.115 seconds based 
+on data from European and World Championships, while a study of the 
+2008 Olympics supported the conclusion that athletes cannot reliably 
+react within 0.1 seconds \citep{lipps2011implications}. Despite these 
+findings, the IAAF has not revised the threshold since its implementation 
+in 1999. It is important to note that \citet{brosnan2017effects} and 
+\citet{lipps2011implications} relied on retrospective event data, 
+whereas \citet{pain2007sprint} and \citet{komi2009iaaf} conducted 
+controlled experiments to assess reaction time capabilities.
 
-However, not all studies have been in favor of a decrease in the
-reaction time barrier \citet*{brosnan2017effects} argue for increasing 
-the reaction time barrier to 0.115 seconds, and cited evidence from European 
-Championships and World Championships in their article. A study looking 
-at the 2008 Olympics likewise supported the conclusion that athletes could not
-react in as little as 0.1 seconds \citep{lipps2011implications}. However it is
-worth noting that \citet*{brosnan2017effects} and \citep{lipps2011implications}
-were retroactive studies of athletes at events while \citet{pain2007sprint} and
-\citet{komi2009iaaf} were experiments showing that there are various methods
-of testing reaction time capability.
+The second objective of this paper is to model reaction times for 
+non-disqualified athletes from the 1999 World Championships onward. 
+We employ a gamma linear mixed-effects model with two random effects: 
+one for the venue (year) and another for the heat nested within each venue.
+This model allows us to estimate the probability of an athlete's 
+reaction time falling below a specific threshold, such as 0.1 seconds. 
+By examining reaction times across years and heats, we aim to assess 
+the consistency of reaction times and whether the current 
+disqualification threshold is statistically justified. This modeling 
+approach provides a more detailed understanding of the factors that 
+influence reaction times, offering insights into whether the 0.1-second 
+threshold should be reconsidered for future competitions.
 
-The second part of this paper will model the reaction times of every non-disqualification since the 1999 World 
-Championships, using a gamma linear mixed-effects model with two random effects: 
-one effect for the year and a second effect for the heat nested within each year.
-Based on the fitted model, we can then estimate the probability of a reaction 
-time being below a certain threshold (e.g., 0.1).
-
-The rest of this paper is organized as follows. Section~\ref{sec:Data} details 
-how the two types of data were collected, as the signed-rank test as described
-in Section~\ref{sec:rank} uses slightly different data than the generalized 
-linear mixed-effect model (GLMM) that is further developed in 
-Section~\ref{sec:glmm}. Section~\ref{sec:Results} provides the results of the two
-analyses and ends by questioning the reaction time barrier. Lastly, 
-Section~\ref{sec:concludingremarks} discusses the impact and limitations of the
-paper. All the data and code of our analysis are available in a supplement.
+The rest of this paper is organized as follows. Section~\ref{sec:Data} 
+describes how the data were collected. The signed-rank test, detailed 
+in Section~\ref{sec:rank}, uses slightly different data from the 
+generalized linear mixed-effects model (GLMM) developed in 
+Section~\ref{sec:glmm}. Section~\ref{sec:Results} presents the results 
+of both analyses and discusses the implications of the reaction time 
+barrier. Finally, Section~\ref{sec:concludingremarks} outlines the 
+paper’s impact and limitations. All data and code for our analysis 
+are provided in the supplementary materials.
 

### 2024-09-21 Manuscript/manuscript.tex
one more round on intro

@@
 
-Devon Allen, a University of Oregon alumnus, was a strong 
-contender for the 2022 World Track and Field Championships 
-in the 110-meter hurdle event. After running the event in 
-12.84 seconds in June 2022, just 0.04 seconds short of the 
-world record \citep{wa2022preview}, and placing third at 
-the U.S. Track and Field Championships, Allen was well-prepared 
-for the World Championships in August, also held in Eugene, 
-Oregon. He advanced through the preliminary heats and semifinals 
-with reaction times of 0.123 and 0.101 seconds, respectively. 
-However, in the final heat, Allen was disqualified for a 
-reaction time of 0.099 seconds, just 0.001 seconds faster 
-than the 0.1-second threshold set by the International 
-Association of Athletics Federations (IAAF). This decision 
-was met with significant public outcry \citep{nbc2022devon}.
+Devon Allen’s highly anticipated performance at the 2022 World 
+Track and Field Championships in Eugene, Oregon, ended in 
+controversy when he was disqualified for a reaction time of 
+0.099 seconds, just 0.001 seconds below the allowable threshold.
+Allen, a University of Oregon alumnus, 
+had recently run a time of 12.84 seconds in the 110-meter hurdle 
+event, just 0.04 seconds short of the world record. After placing 
+third at the U.S. Track and Field Championships, he advanced 
+through the preliminary heats and semifinals at the World 
+Championships, with reaction times of 0.123 and 0.101 seconds, 
+respectively. However, in the final heat, competing in front of his 
+home audience, Allen’s reaction time was just 0.001 seconds faster 
+than the 0.1-second threshold set by the International Association 
+of Athletics Federations (IAAF), resulting in his disqualification, a 
+decision that was met with widespread public outcry.
 
-Allen's disqualification has been discussed extensively in online 
-communities such as \url{www.LetsRun.com}, a website that functions 
-as both a message board and a news outlet. The message board, 
-similar to Reddit, is very active during major running events, 
-including the Olympics and the World Track and Field Championships. 
-The news section of the website featured articles such as 
-\citep{johnson2022data} and \citep{johnson2022was} by Robert Johnson, 
-the site’s founder. Johnson argued that there may have been an 
-error with the timing equipment at the 2022 Championships. He 
-cited graphs, descriptive statistics like median reaction times, 
-and compared the reaction times at the 2022 Championships to 
-those from other competitions, including the U.S. Championships 
-and previous World Championships. However, these comparisons 
-lacked formal statistical analysis. Johnson estimated a 99.95\% 
-chance that Allen was wrongfully disqualified, based on his claim 
-that the probability of the fastest reaction times from the men's 
-100 meter, men's 110 meter hurdle, and women's 100 meter all 
-occurring in the same year was approximately $(1/13)^3$ 
-\citep{johnson2022was}.
+The disqualification of Allen was widely discussed in online 
+communities, with some questioning the accuracy of the timing 
+equipment. One of the most active platforms was \url{www.LetsRun.com}, 
+a website that functions as both a message board and a news outlet. 
+The message board, similar to Reddit, is particularly active during 
+major running events, including the Olympics and the World Track 
+and Field Championships. The news section of the website featured 
+articles such as \citep{johnson2022data} and \citep{johnson2022was} 
+by Robert Johnson, the site’s founder. Johnson argued that there may 
+have been an error with the timing equipment at the 2022 Championships. 
+He cited graphs and descriptive statistics, like median reaction 
+times, comparing the reaction times at the 2022 Championships to 
+those from other competitions, including the U.S. Championships and 
+previous World Championships. However, these comparisons lacked formal 
+statistical analysis.
 
-The first objective of this paper is to determine whether athletes’ 
-reaction times at the 2022 World Track and Field Championships 
-were significantly different from other competitions. Our analysis 
-compares reaction times for athletes who competed in multiple events, 
-using a matched-pairs design. By comparing the same athletes' reaction
-times across different competitions, we can assess whether the 2022 
-reaction times differed significantly. While this investigation was 
-motivated by Allen's disqualification, it also addresses broader concerns 
-about potential variation in timing equipment across different venues
-in short-distance dash or hurdle.
+The first objective of this paper is to investigate whether reaction
+times at the 2022 World Championships were significantly different
+from other competitions, focusing specifically on athletes who
+competed in multiple events. By employing a matched-pairs design, we
+compare reaction times from the same athletes across different
+championships, including the 2022, 2019, and 2023 World Championships,
+as well as the 2022 national-level competitions. This approach allows
+us to isolate the effect of the competition year on reaction times
+while controlling for individual athlete performance. Our analysis
+provide a more formal statistical evaluation of the disqualification 
+threshold, building upon the informal comparisons previously made 
+by others.
 
-Given the observed variation in reaction times across venues, the 
-question arises: is the 0.1-second disqualification threshold truly 
-justified? Studies have presented evidence supporting both sides of 
-the debate. \citet{pain2007sprint} examined reaction times in nine 
-male athletes and found that the fastest athlete had an average 
-reaction time of 0.087 seconds, with a standard deviation of 0.004 
-seconds. Similarly, \citet{komi2009iaaf}, in a study commissioned by 
-the IAAF, recommended lowering the disqualification threshold to 0.08 
+There is an ongoing debate over the appropriateness of the 0.1-second
+threshold for disqualification, with past studies suggesting both
+lower and higher limits. \citet{pain2007sprint} examined 
+reaction times in nine male athletes and found that the fastest athlete 
+had an average reaction time of 0.087 seconds, with a standard deviation 
+of 0.004 seconds. Similarly, \citet{komi2009iaaf}, in a study commissioned 
+by the IAAF, recommended lowering the disqualification threshold to 0.08 
 or 0.085 seconds, arguing that athletes are capable of faster reaction 
 times than the current threshold allows. On the other hand, some 
-researchers have argued for increasing the reaction time threshold. 
+researchers have argued for increasing the threshold.
 \citet{brosnan2017effects} suggested raising it to 0.115 seconds based
@@
 2008 Olympics supported the conclusion that athletes cannot reliably
-react within 0.1 seconds \citep{lipps2011implications}. Despite these 
-findings, the IAAF has not revised the threshold since its implementation 
-in 1999. It is important to note that \citet{brosnan2017effects} and 
-\citet{lipps2011implications} relied on retrospective event data, 
-whereas \citet{pain2007sprint} and \citet{komi2009iaaf} conducted 
-controlled experiments to assess reaction time capabilities.
+react within 0.1 seconds \citep{lipps2011implications}. Interestingly,
+\citet{brosnan2017effects} and \citet{lipps2011implications} relied on
+retrospective event data, whereas \citet{pain2007sprint} and
+\citet{komi2009iaaf} conducted controlled experiments to assess
+reaction time capabilities. The IAAF has not revised the threshold
+since its implementation in 1999.
 
-The second objective of this paper is to model reaction times for 
-non-disqualified athletes from the 1999 World Championships onward. 
-We employ a gamma linear mixed-effects model with two random effects: 
-one for the venue (year) and another for the heat nested within each venue.
-This model allows us to estimate the probability of an athlete's 
-reaction time falling below a specific threshold, such as 0.1 seconds. 
-By examining reaction times across years and heats, we aim to assess 
-the consistency of reaction times and whether the current 
-disqualification threshold is statistically justified. This modeling 
-approach provides a more detailed understanding of the factors that 
-influence reaction times, offering insights into whether the 0.1-second 
-threshold should be reconsidered for future competitions.
+The second objective of this paper is to further investigate the
+appropriateness of the 0.1-second threshold by modeling reaction times
+from athletes who competed in world championships from 1999 onward. We
+employ a gamma linear mixed-effects model with two random effects: one
+for the venue (year) and another for the heat nested within each
+venue. This model allows us to estimate the probability of an
+athlete’s reaction time falling below a specific threshold, such as
+0.1 seconds. By examining reaction times across different years and
+heats, we can assess the consistency of reaction times and whether the
+current disqualification threshold is supported by statistical
+evidence.
 

### 2024-09-21 Manuscript/manuscript.tex
adjusted flow in intro

@@
 
-The first objective of this paper is to investigate whether reaction
-times at the 2022 World Championships were significantly different
-from other competitions, focusing specifically on athletes who
-competed in multiple events. By employing a matched-pairs design, we
-compare reaction times from the same athletes across different
-championships, including the 2022, 2019, and 2023 World Championships,
-as well as the 2022 national-level competitions. This approach allows
-us to isolate the effect of the competition year on reaction times
-while controlling for individual athlete performance. Our analysis
-provide a more formal statistical evaluation of the disqualification 
-threshold, building upon the informal comparisons previously made 
-by others.
-
-There is an ongoing debate over the appropriateness of the 0.1-second
-threshold for disqualification, with past studies suggesting both
-lower and higher limits. \citet{pain2007sprint} examined 
+The 0.1-second disqualification threshold in track events has been 
+a subject of ongoing debate, with researchers presenting evidence 
+both in favor of lowering and raising the
+limit. \citet{pain2007sprint} examined
 reaction times in nine male athletes and found that the fastest athlete 
@@
 
-The second objective of this paper is to further investigate the
-appropriateness of the 0.1-second threshold by modeling reaction times
-from athletes who competed in world championships from 1999 onward. We
-employ a gamma linear mixed-effects model with two random effects: one
-for the venue (year) and another for the heat nested within each
-venue. This model allows us to estimate the probability of an
-athlete’s reaction time falling below a specific threshold, such as
-0.1 seconds. By examining reaction times across different years and
-heats, we can assess the consistency of reaction times and whether the
-current disqualification threshold is supported by statistical
-evidence.
+This paper has two primary objectives. First, we investigate whether 
+reaction times at the 2022 World Championships were significantly 
+different from other competitions, focusing on athletes who competed 
+in multiple events. Using a matched-pairs design, we compare reaction 
+times across the 2022, 2019, and 2023 World Championships, as well as 
+the 2022 national-level competitions. This approach isolates the 
+effect of the competition year while controlling for individual 
+performance. The second objective is to assess the 
+appropriateness of the 0.1-second threshold by modeling reaction 
+times from World Championships held from 1999 onward. We use a gamma 
+linear mixed-effects model with random effects for both venue and heat. 
+This model estimates the probability of reaction times falling below 
+the threshold, helping evaluate the consistency of reaction times and 
+whether the current threshold is statistically justified.
 

### 2024-09-21 Manuscript/reply.tex
replied to more points, still need to work womens results into discussion

@@
 
-One of the reasons why we did not include data from other sprinting evetns is
-because we wanted to ground the article in Devon Allen's story and to be about
-the hurdles events.  Athletes who compete in the longer events such as 400 meters
-may not have as good of a start as those in the 100 meter events because reaction
-time plays a much smaller role in the outcome of the race. While it is common
-for amateur and collegiate athletes to compete in multiple sprinting events, it
-seems reasonable that a world championship 100 meter runner may be only a top 20
-200 meter runner.  We are trying to estimate the probabilty of observing a extreme
-reaction time to determine whether 0.1 seconds is fair but if we include data
-from competitions with slower reaction times, then 0.1 seconds is going to be
-more extreme.
+Thank you for these suggestions, we were able to expand the scope of our data
+by including results from the men's and women's 100 meter dash.  We considered
+adding additional sprinting events such as the 200 meter dash but
+when we included a term for type of event in our model the coefficient on this
+event term was significant for the 200 meter dash for both men's and women's
+data. The coefficient was not signficant when we included the 100 meter dash
+results. This is not particularly suprising as athletes who compete in the longer
+events may not have as good of a start as those in the 100
+meter events because reaction time plays a much smaller role in the outcome of
+the race. While it is common for amateur and collegiate athletes to compete in
+multiple sprinting events, it seems reasonable that a world championship 100
+meter runner may be only a very good 200 meter runner.  We are trying to
+estimate the probabilty of observing a extreme reaction time to determine
+whether 0.1 seconds is fair but if we include data from competitions with slower
+reaction times, then 0.1 seconds is going to be more extreme.  To answer later
+thoughts about looking at all sprinting events, we observed boxplots and what we
+found were that reaction times in longer sprinting events were much more stable
+than in the 100 meter dash and hurdle and thus any insights would likely not be
+informative to questioning whether 0.1 seconds is a fair barrier or if athletes
+in those events had faster reaction times in 2022.  This follows the conventional
+wisdom that athletes are more focuesed on not false starting at the start of the
+longer races than they are on getting a perfect start.
 
@@
 not want to call it a "race effect" as we were worried that may be interpreted
-to mean the race of the athlete.
+to mean the race of the athlete.  We did not think it would be informative to 
+add and effect for individual runner's average ability as the majority of
+observations in the data come from runners with very few (1-3) data points,
+so taking an average of such a small sample may not be informative, especially
+if there are outliers or an athlete had a bad start.
 
@@
 \end{comment}  
-We do not have enough evidence to conclude that the sensors were calibrated
-differently, so thus we make general recommendations for World Athletics.
+The second reviewer shared a similar sentiment about not effectively
+presenting ideas so we we made improvements to the introduction section to try
+and more clearly present our ideas.  We shortened the literature review to only
+cite papers that were actually of interest to the paper and tried to clearly
+define the two purposes of the paper: to decide of 0.1 seconds is fair, and to
+identify inconsistensies with the 2022 reaction times. We try to be precise in
+our language as we do not have enough evidence to conclude that the sensors were
+calibrated differently, or that Seiko (who make the timing devices) is at fault
+for the 2022 reaction times being inexplicably faster. Thus we make general
+recommendations for World Athletics and note that the reaction times in 2022
+were unusually fast with no clear reason.
 
@@
 
-We can restructure the paper but doesn't the abstract mention the goals of the
-paper? I suppose the introduction is really more of a background section.
+We have re-worked the introduction and discussion sections of the paper to more
+clearly outline the goals of the paper.  The organization of the paper was
+something that both reviewers commented on so we did our best to try and
+streamline the two key points of the paper: to decide of 0.1 seconds is fair,
+and to identify inconsistensies with the 2022 reaction times. With these two
+points emphasized we tried to remove sections that did not support these central
+ideas.  In regards to the language surrounding the rank comparison text and the
+timing device at the 2022 Championships we are intentionally vague as we do not
+have evidence to make a stronger claim.  For example, we cannot claim that the
+timing device was faulty as miscalibrated as there is no public information
+available about that.  We do not want to overstep and put ourselves or the
+publication at risk of defamation claims from Seiko because we called their
+devices faulty when they are not.  Instead we suggest to the reader that 2022
+was an unusually fast year and comment that there is no explanation as to why
+athletes improve reaction times over the span of a few months and let the reader
+decide what they think is the underlying cause of the faster reaction times.
 
@@
 could online, but there is simply not enough data to justify a formal analysis
-of his reaction times across lower level meets.
+of his reaction times across lower level meets.  Many track and field races do
+not record data on reaction times until at the very high level (national
+and world championship).  Simply put, reaction times are not heavily researched
+and thus the data material in insufficient.
 

### 2024-10-05 Code/RankComparison.Rmd

@@
 ```
+```{r}
+old_dash <- read.csv("../Data/Clusrank2019vs2022Dash.csv")
+colnames(old_dash) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_dash,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_dash,
+                method = "ds",
+                exact = TRUE, B = 500000)
+```
+```{r}
+old_women_hurdles <- read.csv("../Data/Clusrank2019vs2022Womens.csv")
+colnames(old_women_hurdles) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_women_hurdles,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_women_hurdles,
+                method = "ds",
+                exact = TRUE, B = 500000)
+```
+```{r}
+old_women_dash <- read.csv("../Data/Clusrank2019vs2022WomensDash.csv")
+colnames(old_women_dash) <- c("Time", "Athlete", "Year")
+```
+
+```{r}
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_women_dash,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_women_dash,
+                method = "ds",
+                exact = TRUE, B = 500000)
+```

### 2024-10-05 Manuscript/reply.tex

@@
 
-\subsection*{To Reviewer}
+\subsection*{To Reviewer 1}
+
+\jy{What is ``the first point?''  Please quote everything from the
+  reviewer, including the positive comments.}
 
@@
 
+\jy{At the end of each reply, we need to state what changes we made to
+  the manuscript with their specific locations. The explanation here
+  (the exclusion of 200m dash) should be included in the paper as well.}
+
 \begin{comment}
@@
 purpose would help.
+\end{comment}
 
-\end{comment}  
 The second reviewer shared a similar sentiment about not effectively
@@
 define the two purposes of the paper: to decide of 0.1 seconds is fair, and to
+\jy{The order is reversed in the paper.}
 identify inconsistensies with the 2022 reaction times. We try to be precise in
@@
 
-Thank you for these encouraging words.
+Thank you for these encouraging words. \jy{State what expansions we
+  have done.}
 
@@
 is very vague.
-
 \end{comment}
@@
 rounds. The article is now behind a pay wall.
+\jy{Not a reason for not citing it. We can still find pdf online.}
+
+\jy{Using log-normal distribution for the reaction times is quite easy
+  to do. You just take log of the reaction times, and then model them
+  in a GLMM with Guassian family. I think we did this, and decided it
+  was not good perhaps from the residual qq plots. Please find them
+  and add a discussion into the paper.}
+
+\jy{This references has been cited 31 times. Some of these are closely
+  related to our topic and should be cited as appropriate.}
 
@@
 and thus the data material in insufficient.
+\jy{We have added a discussion in paragraph ??? of the Discussion Section.}
+
+\jy{Every point we cannot answer needs to have some response reflected
+  in discussions in the manuscript.}
 
@@
 We added additional explanations to address this in the introductions section.
+\jy{Give paragraph number for specific locations}
 
@@
 \subsection*{Associate Editor}
+\jy{Put this section before the reviewers.}
+
 \begin{comment}

### 2024-10-15 Code/RankComparison.Rmd
Re-organized the code, but still need to run the permutation tests

--- 2024-06-20 Code/RankComparison.Rmd
@@
 
-The code below performs two clustered Wilcox tests, the first one without using
+# Overview
+This analysis consists of twelve comparisons looking at all combinations of
+the men's 100 meter dash, men's 110 meter hurdle, women's 100 meter dash, and
+women's 100 meter hurdle.  In each comparison we will first read in the data
+and then perform two clustered Wilcox tests, the first one without using
 exact = TRUE, and the second one with exact = TRUE and 1,000,000 random
 permutations. The tests compare the times and group by the Stage of the
-competition, which in the data is coded as 1 if it was a national championship
-and 2 if it was the world championships. The tests cluster by athlete, of which
-for each dataset an integer from 1-11 is used (the values are arbitrary).
+competition, which in the data is coded as 1 if it took place during the 2022
+World Championship and 0 otherwise. The tests cluster by athlete, of which for
+each data set an integer from 1-11 is used (the values are arbitrary).
 For more information about the "ds" method, please look at the Rank-based
 Comparison subsection of the Methods section of the paper or
-[at this link here](https://www.jstatsoft.org/article/view/v096i06.)
+[at this link here.](https://www.jstatsoft.org/article/view/v096i06.)
+
+First we will look compare the reaction times from athletes who competed in both
+the 2019 and 2022 World Track and Field Championship.
+
+# 2019 versus 2022
+
+### Men's 110 meter hurdles
+
+First we read in the data from the appropriate source.
+
+```{r}
+men_hurdle_2019 <- read.csv("../Data/Clusrank2019vs2022.csv")
+colnames(men_hurdle_2019) <- c("Time", "Athlete", "Year")
+```
+
+Then we perform the clustered Wilcox tests.
 
@@
 library(clusrank)
-clusWilcox.test(Time, group=Stage, cluster = Athlete, data = national,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2019,
                 method = "ds")
 
-clusWilcox.test(Time, group=Stage, cluster = Athlete, data = national,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2019,
                 method = "ds",
@@
 
-The code below reads in the data comparing the 2019 and 2022 World Championships.
+### Women's 100 meter hurdles
+
+First we read in the data from the appropriate source.
+
 ```{r}
-old_clusrank <- read.csv("../Data/Clusrank2019vs2022.csv")
-colnames(old_clusrank) <- c("Time", "Athlete", "Year")
+women_hurdle_2019 <- read.csv("../Data/Clusrank2019vs2022Womens.csv")
+colnames(women_hurdle_2019) <- c("Time", "Athlete", "Year")
 ```
 
-The tests below are very similar to the ones above, the only difference being
-grouping by Year instead of Stage (0 if 2019 and 1 if 2022).
+Then we perform the clustered Wilcox tests.
+
 ```{r}
 set.seed(1)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_clusrank,
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2019,
                 method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_clusrank,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2019,
                 method = "ds",
@@
 
-The code below reads in data comparing the 2022 and 2023 World Championships
+### Men's 110 meter dash
+
+First we read in the data from the appropriate source.
+
 ```{r}
-new_clusrank <- read.csv("../Data/Clusrank2022vs2023.csv")
-colnames(new_clusrank) <- c("Time", "Athlete", "Year")
+men_dash_2019 <- read.csv("../Data/Clusrank2019vs2022Dash.csv")
+colnames(men_dash_2019) <- c("Time", "Athlete", "Year")
 ```
 
-The tests below are very similar to the ones above, the only difference being
-grouping by Year (0 if 2022 and 1 if 2023).
+Then we perform the clustered Wilcox tests.
+
 ```{r}
 set.seed(1)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = new_clusrank,
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2019,
                 method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = new_clusrank,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2019,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```
+### Women's 100 meter dash
+
+First we read in the data from the appropriate source.
+
+```{r}
+women_dash_2019 <- read.csv("../Data/Clusrank2019vs2022WomensDash.csv")
+colnames(women_dash_2019) <- c("Time", "Athlete", "Year")
+```
+
+Then we perform the clustered Wilcox tests.
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2019,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2019,
                 method = "ds",
@@
 
-Now we test the womens 2022 vs 2023
+# 2022 National versus World
+Now we will look compare the reaction times from athletes at their 2022 national
+track and field championships with the 2022 World Championships.
+
+### Men's 110 meter hurdles
+
+First we read in the data from the appropriate source.
+
 ```{r}
-womens <- read.csv("../Data/Clusrank2022vs2023Womens.csv")
-colnames(womens) <- c("Time", "Athlete", "Year")
+men_hurdle_2022 <- read.csv("../Data/ClusrankNatvsWorld.csv")
+colnames(men_hurdle_2022) <- c("Time", "Athlete", "Year")
 ```
 
-The tests below are very similar to the ones above, the only difference being
-grouping by Year (0 if 2022 and 1 if 2023).
+Then we perform the clustered Wilcox tests.
+
 ```{r}
+set.seed(1)
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = womens,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2022,
                 method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = womens,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2022,
                 method = "ds",
-                exact = TRUE, B = 500000)
+                exact = TRUE, B = 1000000)
 ```
+
+### Women's 100 meter hurdles
+
+First we read in the data from the appropriate source.
+
+```{r}
+women_hurdle_2022 <- read.csv("../Data/ClusrankNatvsWorldWomens.csv")
+colnames(women_hurdle_2022) <- c("Time", "Athlete", "Year")
+```
+
+Then we perform the clustered Wilcox tests.
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2022,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2022,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```
+
+### Men's 110 meter dash
+
+First we read in the data from the appropriate source.
+
+```{r}
+men_dash_2022 <- read.csv("../Data/ClusrankNatvsWorldDash.csv")
+colnames(men_dash_2022) <- c("Time", "Athlete", "Year")
+```
+
+Then we perform the clustered Wilcox tests.
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2022,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2022,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```
+### Women's 100 meter dash
+
+First we read in the data from the appropriate source.
+
+```{r}
+women_dash_2022 <- read.csv("../Data/ClusrankNatvsWorldWomensDash.csv")
+colnames(women_dash_2022) <- c("Time", "Athlete", "Year")
+```
+
+Then we perform the clustered Wilcox tests.
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2022,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2022,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```
+# 2022 versus 2023
+
+Lastly we will look compare the reaction times from athletes who competed in both
+the 2022 and 2023 World Track and Field Championship.
+
+### Men's 110 meter hurdles
+
+First we read in the data from the appropriate source.
+
+```{r}
+men_hurdle_2023 <- read.csv("../Data/Clusrank2022vs2023.csv")
+colnames(men_hurdle_2023) <- c("Time", "Athlete", "Year")
+```
+
+Then we perform the clustered Wilcox tests.
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2023,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_hurdle_2023,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```
+
+### Women's 100 meter hurdles
+
+First we read in the data from the appropriate source.
+
+```{r}
+women_hurdle_2023 <- read.csv("../Data/Clusrank2022vs2023Womens.csv")
+colnames(women_hurdle_2023) <- c("Time", "Athlete", "Year")
+```
+
+Then we perform the clustered Wilcox tests.
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2023,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2023,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```
+
+### Men's 110 meter dash
+
+First we read in the data from the appropriate source.
+
+```{r}
+men_dash_2023 <- read.csv("../Data/Clusrank2022vs2023Dash.csv")
+colnames(men_dash_2023) <- c("Time", "Athlete", "Year")
+```
+
+Then we perform the clustered Wilcox tests.
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2023,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_dash_2023,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```
+### Women's 100 meter dash
+
+First we read in the data from the appropriate source.
+
+```{r}
+women_dash_2023 <- read.csv("../Data/Clusrank2022vs2023WomensDash.csv")
+colnames(women_dash_2023) <- c("Time", "Athlete", "Year")
+```
+
+Then we perform the clustered Wilcox tests.
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2023,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash_2023,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```

### 2024-10-30 Code/NewAnalysis.Rmd
DQ analysis is at the bottom

@@
   filter(Gender == "M") |>
-  filter(Time != "DNS") |>
+  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
   filter(RxnTime != 0) |>
-  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
-  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+  filter(Event == "100 Dash" | Event == "110 Hurdles")
 
@@
 
-Now we will see how the inclusion of the pooled data will influence the results
-```{r}
-sim_three_effect_time <- function(model, base, RxnTime) {
-  simfit <- function(model, n = 10000000) {
-    vc <- as.data.frame(VarCorr(model))
-    sd_Venue <- vc$sdcor[1]
-    sd_Venue_heat <- vc$sdcor[2]
-    sd_event <- vc$sdcor[3]
-
-    z1 <- rnorm(n, sd = sd_Venue)
-    z2 <- rnorm(n, sd = sd_Venue_heat)
-    z3 <- rnorm(n, sd = sd_event)
-    shape <- 1 / sigma(model)^2
-    mu <- exp(fixef(model) + z1 + z2 + z3)
-
-    x <- rgamma(n, shape = shape, scale = mu / shape)
-    x
-  }
-
-  sbase <- summary(base)
-  x <- simfit(model)
-  print(mean(x < RxnTime))
-  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
-  ## compare with and without random effect
-  plot(density(x))  # with random effects
-  curve(dgamma(x, shape = 1 / sbase$dispersion, 
-               scale = exp(coef(base)) * sbase$dispersion), 
-        0.08, 0.25, add = TRUE, col = "blue")
-}
-```
-
 ## 200m test
@@
   filter(Gender == "M") |>
-  filter(Time != "DNS") |>
+  filter(Time != "DNS" & Time != "DQ") |>
   filter(RxnTime != 0) |>
@@
 There seems to be more of a difference but the test is still not significant.
+
+Based on the above evidence it seems reasonable to conclude that it is appropriate
+to pool the men's 110 meter hurdles with the 100 meter dash and pool the women's
+100 meter hurdles with the 100 dash but based on previous studies we are going
+to separate men's and women's data.
+
+## Testing Whether or not to include disqualfications
+```{r}
+dash_and_hurdles <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles")
+```
+
+```{r}
+cat("The number of entries of the men's 100 meter dash and 110 meter hurdles
+    without disqualifcations is", nrow(dash_and_hurdles))
+```
+
+```{r}
+dash_and_hurdles_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat), 
+                                   data = dash_and_hurdles,
+                                   family = Gamma(link = "log"))
+```
+
+We need the baseline model for our analysis as well
+```{r}
+dash_and_hurdles_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles, 
+                                    family = Gamma(link = "log"))
+```
+
+```{r}
+sim_two_effect_time <- function(model, base, RxnTime) {
+  simfit <- function(model, n = 10000000) {
+    vc <- as.data.frame(VarCorr(model))
+    sd_Venue <- vc$sdcor[1]
+    sd_Venue_heat <- vc$sdcor[2]
+
+    z1 <- rnorm(n, sd = sd_Venue)
+    z2 <- rnorm(n, sd = sd_Venue_heat)
+
+    shape <- 1 / sigma(model)^2
+    mu <- exp(fixef(model) + z1 + z2)
+
+    x <- rgamma(n, shape = shape, scale = mu / shape)
+    x
+  }
+
+  sbase <- summary(base)
+  x <- simfit(model)
+  print(mean(x < RxnTime))
+  #print(quantile(x, prob = c(0.00001, 0.0001, 0.0005, 0.001)))
+  ## compare with and without random effect
+  #plot(density(x))  # with random effects
+  #curve(dgamma(x, shape = 1 / sbase$dispersion, 
+   #            scale = exp(coef(base)) * sbase$dispersion), 
+  #      0.08, 0.25, add = TRUE, col = "blue")  # fixed effect only
+}
+```
+
+```{r}
+set.seed(1)
+sim_two_effect_time(dash_and_hurdles_two_effect_gamma_log, dash_and_hurdles_one_gamma_log, 0.1)
+```
+Now lets see how our results change when we include disqualifications
+```{r}
+dash_and_hurdles_with_dq <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
+  mutate(RxnTime = ifelse(RxnTime < 0, 0.000001, RxnTime))
+```
+
+```{r}
+cat("The number of entries of the men's 100 meter dash and 110 meter hurdles
+    without disqualifcations is", nrow(dash_and_hurdles_with_dq))
+```
+
+```{r}
+dash_and_hurdles_with_dq_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat), 
+                                   data = dash_and_hurdles_with_dq,
+                                   family = Gamma(link = "log"))
+```
+
+We need the baseline model for our analysis as well
+```{r}
+dash_and_hurdles_with_dq_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles_with_dq, 
+                                    family = Gamma(link = "log"))
+```
+
+```{r}
+set.seed(1)
+sim_two_effect_time(dash_and_hurdles_with_dq_two_effect_gamma_log, dash_and_hurdles_with_dq_one_gamma_log, 0.1)
+```
+## Lets repeat the previous section but exclude 2022 from our results
+```{r}
+dash_and_hurdles_no2022 <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
+  filter(Venue != 2022)
+```
+
+```{r}
+dash_and_hurdles_no2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat), 
+                                   data = dash_and_hurdles_no2022,
+                                   family = Gamma(link = "log"))
+
+dash_and_hurdles_no2022_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles_no2022, 
+                                    family = Gamma(link = "log"))
+```
+
+```{r}
+set.seed(1)
+sim_two_effect_time(dash_and_hurdles_no2022_two_effect_gamma_log, dash_and_hurdles_no2022_one_gamma_log, 0.1)
+```
+
+Now we analyze without 2022 but with disqualifications
+```{r}
+dash_and_hurdles_no2022_with_dq <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
+  filter(Venue != 2022) |>
+  mutate(RxnTime = ifelse(RxnTime < 0, 0.000001, RxnTime))
+```
+
+```{r}
+dash_and_hurdles_no2022_with_dq_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat), 
+                                   data = dash_and_hurdles_no2022_with_dq,
+                                   family = Gamma(link = "log"))
+
+dash_and_hurdles_no2022_with_dq_one_gamma_log <- glm(RxnTime ~ 1, data = dash_and_hurdles_no2022_with_dq, 
+                                    family = Gamma(link = "log"))
+```
+
+```{r}
+sim_two_effect_time(dash_and_hurdles_no2022_with_dq_two_effect_gamma_log, dash_and_hurdles_no2022_with_dq_one_gamma_log, 0.1)
+```

### 2024-11-05 Code/NewAnalysis.Rmd
added log normal analysis

@@
   filter(Event == "100 Dash" | Event == "110 Hurdles") |>
-  mutate(RxnTime = ifelse(RxnTime < 0, 0.000001, RxnTime))
+  filter(RxnTime > 0)
+  #mutate(RxnTime = ifelse(RxnTime < 0, 0.000001, RxnTime))
 ```
@@
 ```
+## Comparison of model stats with and without dq
+These are model stats for model without any disqualifications
+```{r}
+summary(dash_and_hurdles_two_effect_gamma_log)
+```
+
+These are model stats for model with positive disqualifications
+```{r}
+summary(dash_and_hurdles_with_dq_two_effect_gamma_log)
+```
+
 ## Lets repeat the previous section but exclude 2022 from our results
@@
 ```
+## Log normal Model?
+```{r}
+library(tidyverse)
+times <- read.csv("../Data/GLMMData.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
+```
+The data we use
+```{r}
+dash_and_hurdles_with_dq <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
+  filter(RxnTime > 0)
+```
+
+Log normal model
+```{r}
+library(lme4)
+log_normal_model <- lmer(log(RxnTime) ~ (1 | Venue) + (1 | Heat), 
+                                   data = dash_and_hurdles_with_dq)
+log_normal_model
+```
+
+Current best fit model (Included DQs because that is what I believe we will
+use as final model.)
+```{r}
+dash_and_hurdles_with_dq_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) +
+                                                   (1 | Heat), 
+                                   data = dash_and_hurdles_with_dq,
+                                   family = Gamma(link = "log"))
+```
+
+Here is the qq plot for both. It looks very similar in my opinion with only some
+small differences.  I think it is reasonable to keep the gamma model.  Is it not
+possible to compare the AIC's? When I try to find the AIC for log normal model
+I am unable to.
+```{r}
+library(car)
+resids_log_normal <- residuals(log_normal_model)
+resids_gamma <- residuals(dash_and_hurdles_with_dq_two_effect_gamma_log)
+qqPlot(resids_log_normal)
+qqPlot(resids_gamma)
+```

### 2024-11-05 Manuscript/reply.tex
updated responses to comments. Will go back through after replies to email

@@
 \newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
-\newcommand{\mc}[1]{\textcolor{green}{(MC: #1)}}
+\newcommand{\of}[1]{\textcolor{green}{(MC: #1)}}
 
@@
 The manuscript has been revised accordingly with the most notable change being 
-the addition of Section 5 (Classroom Implementation). We believe this revision 
-has enhanced the both the quality and utility of this paper.
+the addition of new data in both the GLMM and rank comparison sections in an
+attempt to expand the scope of the analysis and derive beter insights.
 
@@
   reviewer, including the positive comments.}
+\of{resolved}
 
 \begin{comment}
-To address the first point, the paper relies on descriptive statistics, ranked
+Using Devon Allen's disqualification at the 2022 World Track and Field
+championships as a hook, this paper poses two questions. First, were reaction
+times at the 2022 World Track and Field Championships faster than they had been
+earlier in 2022 national-level competitions, the 2019 World Track and Field
+Championships, and the 2023 World Track and Field Championships? Second, based
+on modeled reaction times, is the current standard of a 0.1 second threshold
+for reaction times before disqualification too conservative? To
+address the first point, the paper relies on descriptive statistics, ranked
 tests, and the results of a set of GLMMs. The descriptive data analysis lends
@@
 when we included a term for type of event in our model the coefficient on this
-event term was significant for the 200 meter dash for both men's and women's
-data. The coefficient was not signficant when we included the 100 meter dash
-results. This is not particularly suprising as athletes who compete in the longer
+event term was not significant for when we included the 100 meter dash
+but became significant after we added the 200 meter dash.
+This is not particularly suprising as athletes who compete in the longer
 events may not have as good of a start as those in the 100
@@
 cite papers that were actually of interest to the paper and tried to clearly
-define the two purposes of the paper: to decide of 0.1 seconds is fair, and to
-\jy{The order is reversed in the paper.}
-identify inconsistensies with the 2022 reaction times. We try to be precise in
+define the two purposes of the paper: to identify inconsistensies with the 2022
+reaction times and to decide of 0.1 seconds is a fair barrier. We try to be precise in
 our language as we do not have enough evidence to conclude that the sensors were
@@
 
-Thank you for these encouraging words. \jy{State what expansions we
+Thank you for these encouraging words, we have added additional data to both
+the GLMM and clusrank section in order to strengthen our arguments: there was
+an anomaly in 2022 and that the reaction time barrier of 0.1 seconds is not
+grounded in formal statistical analysis.
+
+\jy{State what expansions we
   have done.}
+\of{After we are wrapped up with the manuscript it will be easier to identify
+all the improvements we made. Let's just leave this here for now.}
 
@@
 \end{comment}
-Athletes may be disqualified based on PED use, and athletes may not finish a race
-due to injury, thus we did not consider these times.  However, it does seem
-reasonable to include times from people who were disqualified based on their
-reaction time.
+We will start by addressing the comments related to runners who have been
+disqualified.  Some of the runners who were disqualified had a negative reaction
+time, indicating they were trying to predict the starting gun and actually left
+before the gun even fired.  These runners cannot be included in the analysis
+because the point of the GLMM section is to question the current reaction barrier
+methods by determining how fast athletes can react in; but athletes with a
+negative time left before the gun even went off (Also practically speaking
+a negative reaction time makes little sense).  As for the roughly 20 athletes
+across men's and women's data without negative times but were disqualified, we
+analyzed how the inclusion of these athletes effected our results.  What we found
+was that the probability of a reaction time below 0.1 seconds increased from
+0.0029 to 0.0039 and we have attached the R code so that you can see our work.
+\of{Not sure the technical details of how this is going to work but we will come
+back to this before we submit.} Additionally there were similar parameter values
+for the standard deviations of the random effects and we decided that we were
+going to keep these observations in our data.
+
+As for why the particular model was
+chosen, we looked at several types of models and chose the one that we thought
+fit the residual qq plot the best and had the largest log likelihood. After
+selecting the gamma generalized linear model we again looked at several link
+functions such as "identity", "log",
+
+\of{Need to justify why we did what we did and disadvantages of including the
+disqualified
+Do this in the paper and reply. Say section where we made any changes. Give
+justifications for any reasons for disagreeing}
+
+\of{Will clean this up based on Prof Yan and Schifano's comments to 11/5 email}
 
@@
 
-\begin{comment}
 We have taken additional steps to streamline the manuscript and enhance the
@@
 our submission.
-\end{comment}
 

### 2024-11-07 Manuscript/reply.tex
minor

@@
 
-\usepackage[dvipsnames]{color}
+\usepackage[dvipsnames]{xcolor}
 \definecolor{darkblue}{rgb}{0.1, 0.2, 0.6}
 
-\newcommand{\jy}[1]{\textcolor{red}{JY: #1}}
+\newcommand{\jy}[1]{\textcolor{orange}{JY: #1}}
 \newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
-\newcommand{\of}[1]{\textcolor{green}{(MC: #1)}}
+\newcommand{\of}[1]{\textcolor{purple}{(OF: #1)}}
 
@@
 \begin{center}
-  {\Large\bf Response to the Comments}
+  {\Large\bf TAS-23-296: Response to the Comments}
 \end{center}
@@
 \subsection*{To Reviewer 1}
-
-\jy{What is ``the first point?''  Please quote everything from the
-  reviewer, including the positive comments.}
-\of{resolved}
 
@@
 
-Thank you for these encouraging comments.
+We thank the reviewer for the accurate summary of our contributions
+and these encouraging comments.
 

### 2024-11-12 Code/RankComparison.Rmd
Now includes results of permutation tests
Combined analyses so theres 6 instead of 12

--- 2024-07-09 Code/RankComparison.Rmd
@@
 
-The code below reads in the data comparing the 2022 National Competitions to the
-2022 World Championships.
+# Overview
+This analysis consists of six comparisons of times comparing athletes who
+competed at the 2022 World Championships and another major event. The six
+analyses are listed below. Men's data contains values from athletes who competed
+in either the men's 110 meter hurdles or 100 meter dash at championship level
+events. Women's data contains values from athletes who competed
+in either the women's 100 meter hurdles or 100 meter dash at championship level
+events.
+
++ 2019 and 2022 World Championships men's data
++ 2019 and 2022 World Championships women's data
++ 2022 national championships and 2022 World Championships men's data
++ 2022 national championships and 2022 World Championships women's data
++ 2023 and 2022 World Championships men's data
++ 2023 and 2022 World Championships women's data
+
+In each comparison we will first read in the data
+and then perform two clustered Wilcox tests, the first one without using
+exact = TRUE, and the second one with exact = TRUE and 1,000,000 random
+permutations. The tests compare the times and group by the Stage of the
+competition, and cluster by athlete.  An explanation of the features of each
+data is below.
+
++ `Time` is the athlete's reaction time
++ `Athlete` is an arbitrary integer assigned to each athlete.  Each value is
+unique within each data set.
++ `Year` is coded as 1 if it took place during the 2022 World Championship and
+0 otherwise.
++ `Event` is coded as 0 for a hurdles event and 1 as a dash event, although this
+has no impact on the analysis.
+
+For more information about the "ds" method, please look at the Rank-based
+Comparison subsection of the Methods section of the paper or
+[at this link here.](https://www.tandfonline.com/doi/abs/10.1198/016214504000001583)
+
+# 2019 versus 2022
+
+First we will look compare the reaction times from athletes who competed in both
+the 2019 and 2022 World Track and Field Championship.
+
+### Men's Data
+
+First we read in the data from the appropriate source.
+
 ```{r}
-national <- read.csv("../Data/ClusrankNatvsWorld.csv")
-colnames(national) <- c("Time", "Athlete", "Stage")
+men_2019 <- read.csv("../Data/Clusrank2019vs2022.csv")
+colnames(men_2019) <- c("Time", "Athlete", "Year", "Event")
 ```
 
-The code below performs two clustered Wilcox tests, the first one without using
-exact = TRUE, and the second one with exact = TRUE and 1,000,000 random
-permutations. The tests compare the times and group by the Stage of the
-competition, which in the data is coded as 1 if it was a national championship
-and 2 if it was the world championships. The tests cluster by athlete, of which
-for each dataset an integer from 1-11 is used (the values are arbitrary).
-For more information about the "ds" method, please look at the Rank-based
-Comparison subsection of the Methods section of the paper or
-[at this link here](https://www.jstatsoft.org/article/view/v096i06.)
+Then we perform the clustered Wilcox tests.
 
@@
 library(clusrank)
-clusWilcox.test(Time, group=Stage, cluster = Athlete, data = national,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2019,
                 method = "ds")
 
-clusWilcox.test(Time, group=Stage, cluster = Athlete, data = national,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2019,
                 method = "ds",
@@
 
-The code below reads in the data comparing the 2019 and 2022 World Championships.
+### Women's Data
+
+First we read in the data from the appropriate source.
+
 ```{r}
-old_clusrank <- read.csv("../Data/Clusrank2019vs2022.csv")
-colnames(old_clusrank) <- c("Time", "Athlete", "Year")
+women_2019 <- read.csv("../Data/Clusrank2019vs2022Womens.csv")
+colnames(women_2019) <- c("Time", "Athlete", "Year", "Event")
 ```
 
-The tests below are very similar to the ones above, the only difference being
-grouping by Year instead of Stage (0 if 2019 and 1 if 2022).
+Then we perform the clustered Wilcox tests.
+
 ```{r}
 set.seed(1)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_clusrank,
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2019,
                 method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = old_clusrank,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2019,
                 method = "ds",
@@
 
-The code below reads in data comparing the 2022 and 2023 World Championships
+# 2022 National versus World
+Now we will look compare the reaction times from athletes at their 2022 national
+track and field championships with the 2022 World Championships.
+
+### Men's Data
+
+First we read in the data from the appropriate source.
+
 ```{r}
-new_clusrank <- read.csv("../Data/Clusrank2022vs2023.csv")
-colnames(new_clusrank) <- c("Time", "Athlete", "Year")
+men_2022 <- read.csv("../Data/ClusrankNatvsWorld.csv")
+colnames(men_2022) <- c("Time", "Athlete", "Year", "Event")
 ```
 
-The tests below are very similar to the ones above, the only difference being
-grouping by Year (0 if 2022 and 1 if 2023).
+Then we perform the clustered Wilcox tests.
+
 ```{r}
 set.seed(1)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = new_clusrank,
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2022,
                 method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = new_clusrank,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2022,
                 method = "ds",
@@
 
-Now we test the womens 2022 vs 2023
+### Women's Data
+
+First we read in the data from the appropriate source.
+
 ```{r}
-womens <- read.csv("../Data/Clusrank2022vs2023Womens.csv")
-colnames(womens) <- c("Time", "Athlete", "Year")
+women_2022 <- read.csv("../Data/ClusrankNatvsWorldWomens.csv")
+colnames(women_2022) <- c("Time", "Athlete", "Year", "Event")
 ```
 
-The tests below are very similar to the ones above, the only difference being
-grouping by Year (0 if 2022 and 1 if 2023).
+Then we perform the clustered Wilcox tests.
+
 ```{r}
+set.seed(1)
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = womens,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2022,
                 method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = womens,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2022,
                 method = "ds",
-                exact = TRUE, B = 500000)
+                exact = TRUE, B = 1000000)
 ```
 
-Now we test for the men's dash
+# 2022 versus 2023
+
+Lastly we will look compare the reaction times from athletes who competed in both
+the 2022 and 2023 World Track and Field Championship.
+
+### Men's Data
+
+First we read in the data from the appropriate source.
+
 ```{r}
-dash <- read.csv("../Data/Clusrank2022vs2023Dash.csv")
-colnames(dash) <- c("Time", "Athlete", "Year")
+men_2023 <- read.csv("../Data/Clusrank2022vs2023.csv")
+colnames(men_2023) <- c("Time", "Athlete", "Year", "Event")
 ```
 
+Then we perform the clustered Wilcox tests.
+
 ```{r}
+set.seed(1)
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = dash,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2023,
                 method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = dash,
-                method = "ds",
-                exact = TRUE, B = 500000)
-```
-```{r}
-women_dash <- read.csv("../Data/Clusrank2022vs2023WomensDash.csv")
-colnames(women_dash) <- c("Time", "Athlete", "Year")
+#clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2023,
+ #               method = "ds",
+  #              exact = TRUE, B = 1000000)
 ```
 
+### Women's Data
+
+First we read in the data from the appropriate source.
+
 ```{r}
+women_2023 <- read.csv("../Data/Clusrank2022vs2023Womens.csv")
+colnames(women_2023) <- c("Time", "Athlete", "Year", "Event")
+```
+
+Then we perform the clustered Wilcox tests.
+
+```{r}
+set.seed(1)
 library(clusrank)
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2023,
                 method = "ds")
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_dash,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2023,
                 method = "ds",
-                exact = TRUE, B = 500000)
+                exact = TRUE, B = 1000000)
 ```

### 2024-11-12 Code/NewAnalysis.Rmd
Added random effect graphs

@@
 ```
+```{r}
+dotplot(ranef(dash_and_hurdles_with_dq_two_effect_gamma_log, condVar = TRUE))
+```
+```{r}
+dotplot(ranef(log_normal_model, condVar = TRUE))
+```

### 2024-11-12 Manuscript/reply.tex
Made some improvements to structure and phrasing

@@
 by including results from the men's and women's 100 meter dash.  We considered
-adding additional sprinting events such as the 200 meter dash but
-when we included a term for type of event in our model the coefficient on this
-event term was not significant for when we included the 100 meter dash
-but became significant after we added the 200 meter dash.
+adding additional sprinting events such as the 200 meter dash but our analysis
+suggested that reaction times were slower in the 200 meter dash.  After running
+a model that included a coefficient for event type on a data set that combined
+the men's 110 meter hurdle and 100 meter dash, the event term was not significant.
+However after we re-ran the model on the 110 meter hurdle, 100 and 200 meter dash,
+the effect became significant. When we peformed a TukeyHSD test to compare the
+times in the three categories: 100 meter dash, 110 meter hurdles, 200 meter dash,
+we found a p-value of 0.0020625 comparing the means of the 200 meter dash and
+the 110 meter hurdlesand a p-value of 0.0000174 comparing the means of the 100
+meter dash and the 200 meter dash.  This indicated to us that the reaction times
+for the 200 meter were too slow to fit into our study on extreme reaction times
+and thus we chose to exclude 200 meter data.
+
 This is not particularly suprising as athletes who compete in the longer
@@
 whether 0.1 seconds is fair but if we include data from competitions with slower
-reaction times, then 0.1 seconds is going to be more extreme.  To answer later
-thoughts about looking at all sprinting events, we observed boxplots and what we
-found were that reaction times in longer sprinting events were much more stable
-than in the 100 meter dash and hurdle and thus any insights would likely not be
-informative to questioning whether 0.1 seconds is a fair barrier or if athletes
-in those events had faster reaction times in 2022.  This follows the conventional
-wisdom that athletes are more focuesed on not false starting at the start of the
-longer races than they are on getting a perfect start.
+reaction times, then 0.1 seconds is going to be more extreme.
 
@@
 The heat effect measures the speed-level of the heat.  We added some additional
-explanations to make it clear what the venue and heat effect refer to. We did
-not want to call it a "race effect" as we were worried that may be interpreted
-to mean the race of the athlete.  We did not think it would be informative to 
-add and effect for individual runner's average ability as the majority of
-observations in the data come from runners with very few (1-3) data points,
-so taking an average of such a small sample may not be informative, especially
-if there are outliers or an athlete had a bad start.
+explanations in the first paragraph of section 3.2 to make it clear what the
+venue and heat effect refer to. We did not want to call it a "race effect" as we
+were worried that may be interpreted to mean the race of the athlete.  We did
+not think it would be informative to  add and effect for individual runner's
+average ability as many of observations in the data come from runners
+with very few (1-3) data points, so taking an average of such a small sample may
+not be informative, especially if there are outliers or an athlete had a bad
+start.
 
@@
 We are attempting to answer question 1 and have included reaction times from
-additional sprinting events from World Championships.  We hypothesized that
-as the length of the race increased, reaction time would be less impactful and
-thus not as important. We started by adding data for the 100 meter dash and that
-is the most similar event in terms of length to the 110 meter hurdles.  When we
-added an additional variable to the model based on whether the observation came
-from a dash event or hurdles event the coefficient was not significant.  An 
-ANOVA test likewise produced a not signficant coefficient indicating to us that
-it seems reasonable to pool the 100 meter dash and 110 meter hurdles.  We
-hypothesized that longer sprinting events (200 meter dash, 400 meter dash, 400
-meter hurdles) would be more difficult to add as those are longer races where
-athletes reactions times are not as important to their overall time in the race.
-When we repeat the procedure we did for the 100 meter dash, now looking at data
-that included the 100 meter dash, 110 meter hurdles, and the 200 meter dash we
-found that the coefficient on event type was signficant, and a p value of 0.01586.
-When we peformed a TukeyHSD test to compare the times in the three categories:
-100 meter dash, 110 meter hurdles, 200 meter dash, we found a p-value of
-0.0020625 comparing the means of the 200 meter dash and the 110 meter hurdles
-and a p-value of 0.0000174 comparing the means of the 100 meter dash and the 200
-meter dash.  This indicated to us that the reaction times for the 200 meter were
-too slow to fit into our study on extreme reaction times and thus we chose to
-exclude 200 meter data.
+additional sprinting events from World Championships.  The inclusion of
+additional data was also suggested by the first reviewer and we have provided
+our explanation of what data we used there.
 

### 2024-11-12 Manuscript/manuscript.tex
started revisions

@@
 of the 0.1-second disqualification barrier. To investigate these
-issues, we employ a signed-rank test for clustered data to compare
+issues, we employ a rank-sum test for clustered data to compare
 athletes' reaction times across multiple competitions, along with a
@@
 The rest of this paper is organized as follows. Section~\ref{sec:Data} 
-describes how the data were collected. The signed-rank test, detailed 
+describes how the data were collected. The rank-sum test, detailed 
 in Section~\ref{sec:rank}, uses slightly different data from the 
@@
 There are two types of data that are explored throughout this paper. For the
-signed-rank tests described in Section~\ref{sec:rank}, the data comes from 
-athletes who competed in the 2022 World Championships and another competition 
+rank-sum tests described in Section~\ref{sec:rank}, the data comes from male 
+athletes who competed in the 110 meter hurdles and 100 meter dash, and female
+athletes who competed 100 meter hurdles and 100 meter dash.
+Athletes must have competed in the 2022 World Championships and another competition 
 (2019 World Championships, 2023 World Championships, or 2022 national 
 championships). For the GLMM described in Section~\ref{sec:glmm}, data comes
-from every world championships from 1999 to 2023. This investigation started 
+from every world championships from 1999 to 2023 and includes reaction times from
+the 110 meter hurdles and 100 meter dash. This investigation started 
 shortly after the 2022 World Championships, but as the 2023 data has now become 
@@
 meaningful statistical analysis, but the data was expanded to include reaction 
-times for 11 more athletes who competed at other national competitions (many of which were 
+times from more athletes who competed at other national competitions (many of which were 
 held from May-July of 2022). This data was difficult to find, as the data is not
@@
 one cluster of size three, six clusters of size four, two clusters of size five,
-and two clusters of size six. We chose to include preliminary rounds in this
-section as well as in Section~\ref{sec:data2019} in order to increase the size 
-of our data. Without data from the preliminary rounds, the number of observations
-dropped from 65 to 26 and the number of athletes included in the analysis
-decreased from 15 to 9.
+and two clusters of size six.
+\of{This explanation is going to need to be re-worked now that we have more data.
+Should we make a table of some sort? It's going to be very wordy without one}
+
+We chose to include preliminary rounds in this section as well as in
+Section~\ref{sec:data2019}, as removing them would have resulted in a much
+narrower analysis with fewer athletes and clusters.
 
@@
 and six clusters of size five.
+\of{This explanation might need to change}
 
@@
 \end{figure}
+\of{Need to go through and edit this}
 
@@
 \end{figure}
+\of{Do We want 2 boxplot figures: 1 hurdles, 1 dash or combine them}
 

### 2024-11-14 Archive/Plots.Rmd
Made new scatterplot graph

@@
 library(ggplot2)
-nineteen <- read.csv("../Data/2019vs2022.csv")
+nineteen <- read.csv("../Archive/2019vs2022.csv")
 #"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
@@
 ```
+```{r}
+library(tidyr)
+library(ggplot2)
+nineteen <- read.csv("../Archive/2019vs2022.csv")
+dash_19 <- read.csv("../Archive/2019vs2022Dash.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(nineteen) <- c("Athlete", "Country", "2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F")
+colnames(dash_19) <- c("Athlete", "2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F")
+nineteen <- nineteen[, -c(1,2)]
+dash_19 <- dash_19[, -c(1)]
+nineteen <- rbind(nineteen, dash_19)
+
+df_long <- tidyr::pivot_longer(nineteen, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F"))
+men_19 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Reaction Time", x = "") +
+  annotate("text", y = 0.21, x = "2022 S", label = "Men 2019 vs 2022", cex = 3) +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+nineteen_w <- read.csv("../Archive/2019vs2022Womens.csv")
+dash_19_w <- read.csv("../Archive/2019vs2022WomensDash.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(nineteen_w) <- c("Athlete", "Country", "2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F")
+colnames(dash_19_w) <- c("Athlete", "2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F")
+nineteen_w <- nineteen_w[, -c(1,2, 9)]
+dash_19_w <- dash_19_w[, -c(1)]
+nineteen_w <- rbind(nineteen_w, dash_19_w)
+
+df_long <- tidyr::pivot_longer(nineteen_w, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("2019 H", "2019 S", "2019 F", "2022 H", "2022 S", "2022 F"))
+women_19 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Reaction Time", x = "") +
+  annotate("text", y = 0.21, x = "2022 S", label = "Women 2019 vs 2022", cex = 3) +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+df1 <- read.csv("../Archive/NatVSWorld.csv")
+df2 <- read.csv("../Archive/2022NatvsWorldDash.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(df1) <- c("Athlete", "Country", "Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F")
+colnames(df2) <- c("Athlete", "Country", "Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F")
+df1 <- df1[, -c(1,2, 9)]
+df2 <- df2[, -c(1, 2)]
+df <- rbind(df1, df2)
+
+df_long <- tidyr::pivot_longer(df, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F"))
+men_22 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Reaction Time", x = "") +
+  annotate("text", y = 0.21, x = "Inter. S", label = "Men 2022 Nat. vs Inter.", cex = 3) +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+df1 <- read.csv("../Archive/2022NatvsWorldWomens.csv")
+df2 <- read.csv("../Archive/2022NatvsWorldWomensDash.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(df1) <- c("Athlete", "Country", "Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F")
+colnames(df2) <- c("Athlete", "Country", "Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F")
+df1 <- df1[, -c(1, 2)]
+df2 <- df2[, -c(1, 2)]
+df <- rbind(df1, df2)
+
+df_long <- tidyr::pivot_longer(df, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("Nat. H", "Nat. S", "Nat. F", "Inter. H", "Inter. S", "Inter. F"))
+women_22 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Reaction Time", x = "") +
+  annotate("text", y = 0.21, x = "Inter. S", label = "Women 2022 Nat. vs Inter.", cex = 3) +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+df1 <- read.csv("../Archive/2022vs2023.csv")
+df2 <- read.csv("../Archive/2022vs2023Dash.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(df1) <- c("Athlete", "Country", "2022 H", "2022 S", "2022 F", "2023 H", "2023 S", "2023 F")
+colnames(df2) <- c("Athlete", "2022 H", "2022 S", "2022 F", "2023 H", "2023 S", "2023 F")
+df1 <- df1[, -c(1, 2)]
+df2 <- df2[, -c(1)]
+df <- rbind(df1, df2)
+
+df_long <- tidyr::pivot_longer(df, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("2023 H", "2023 S", "2023 F", "2022 H", "2022 S", "2022 F"))
+men_23 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Reaction Time", x = "") +
+  annotate("text", y = 0.21, x = "2022 S", label = "Men 2022 vs 2023", cex = 3) +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(tidyr)
+library(ggplot2)
+df1 <- read.csv("../Archive/2022vs2023Womens.csv")
+df2 <- read.csv("../Archive/2022vs2023WomensDash.csv")
+#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
+colnames(df1) <- c("Athlete", "Country", "2022 H", "2022 S", "2022 F", "2023 H", "2023 S", "2023 F")
+colnames(df2) <- c("Athlete", "2022 H", "2022 S", "2022 F", "2023 H", "2023 S", "2023 F")
+df1 <- df1[, -c(1, 2)]
+df2 <- df2[, -c(1)]
+df <- rbind(df1, df2)
+
+df_long <- tidyr::pivot_longer(df, cols = everything(), names_to = "Column", values_to = "Time")
+df_long$Column <- factor(df_long$Column, levels = c("2023 H", "2023 S", "2023 F", "2022 H", "2022 S", "2022 F"))
+women_23 <- ggplot(df_long, aes(x = Column, y = Time)) +
+  geom_point() +
+  labs(y = "Reaction Time", x = "") +
+  annotate("text", y = 0.21, x = "2022 S", label = "Women 2022 vs 2023", cex = 3) +
+  ylim(0.08, 0.22) +
+  theme_minimal()
+```
+
+```{r}
+library(gridExtra)
+rank_scatter <- grid.arrange(men_19, women_19, men_22,women_22, men_23, women_23, ncol = 2)
+ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 7.5, width = 7)
+```

### 2024-11-14 Manuscript/manuscript.tex
Made some more small changes to tables

@@
   \includegraphics{RankScatterPlots}
-  \caption{The top graph shows the reaction times of athletes who competed at 
-	least once each at the 2019 and 2022 World Championships. The middle graph 
-	shows the reaction times of athletes who competed at least once each at the 
-	2022 and 2023 World Championships. The bottom graph shows the reaction times 
-	of athletes who competed at least once at their country's
-  2022 national championships and the 2022 World Championships.}
+  \caption{This collection of scatterplots of data used in the rank comparison
+  section shows that times in 2022 were on average lower. On the horizontal
+  axis below each graph "H", "S", and "F" refer to the heats, semifinals, and 
+  finals respectively. Please note that in the last row the 2022 times are to
+  the right of the 2023 times.}
   \label{fig:RankScatterplots}
 \end{figure}
-\of{Need to go through and edit this}
 
@@
 with the rank-based methods (described in Section~\ref{sec:rank} with results in 
-Section~\ref{subsec:Results_Rank}). In each of the three graphs, the reaction 
+Section~\ref{subsec:Results_Rank}). In each of the six graphs, the reaction 
 times for athletes who competed at the 2022 World Championships
@@
   \includegraphics{Boxplot}
-  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle.}
+  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle
+  and 100 meter dash.}
   \label{fig:Boxplot}
@@
     Including 2022 &
-    Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.149\\
+    Venue Effect Only & $-3798.8$ &   & 0.054 & $-1.871$ & 0.149\\
     & Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
@@
 \end{table}
+\of{needs to be updated}
 
@@
    \midrule
-   Excluding 2022 & $3.31\cdot10^{-5}$ & $4.46\cdot10^{-4}$ &  $3.52\cdot10^{-3}$  \\ 
-   Including 2022 & $4.16\cdot10^{-5}$ & $5.70\cdot10^{-4}$ & $4.38\cdot10^{-3}$ \\
+   Excluding 2022 & $1.10\cdot10^{-5}$ & $2.34\cdot10^{-4}$ &  $2.36\cdot10^{-3}$  \\ 
+   Including 2022 & $1.71\cdot10^{-5}$ & $3.01\cdot10^{-4}$ & $2.90\cdot10^{-3}$ \\
    \bottomrule
@@
 \end{table}
+\of{table has been updated}
 
@@
 the difference is not big in magnitude. For example, the probability for
-seeing a reaction time below 0.10 seconds changes from $3.54\cdot 10^{-3}$ to
-$4.20\cdot 10^{-3}$ when 2022 is included. These values correspond to roughly one in
-every 230 starts and 283 starts, respectively. The table shows that changing the
+seeing a reaction time below 0.10 seconds changes from $2.36\cdot 10^{-3}$ to
+$2.90\cdot 10^{-3}$ when 2022 is included. These values correspond to one in
+every 344 starts and 423 starts, respectively. The table shows that changing the
 reaction time barrier from 0.1 seconds to 0.08 seconds drastically reduces the
 chance of observing a reaction time below the barrier, from one in roughly every
-283 starts to one in every 29851 starts using data with 2022 included. The
+283 starts to one in every 58480 starts using data with 2022 included. The
 results substantiates the recommendations by \citet{komi2009iaaf}.
@@
    \midrule
-   Excluding 2022 & $0.094$ & $0.084$ \\ 
-   Including 2022 & $0.092$ & $0.083$ \\
+   Excluding 2022 & $0.096$ & $0.087$ \\ 
+   Including 2022 & $0.095$ & $0.086$ \\
    \bottomrule

### 2024-11-14 Archive/Plots.Rmd
updated boxplot graph

@@
 ```{r}
-with2022 <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & Time != 'DQ' & Event == "Hurdles")
+library(tidyverse)
+df <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles")
+```
+
+```{r}
 library(ggplot2)
-ggplot(with2022, aes(x = Venue, y = RxnTime, group = Venue)) +
+ggplot(df, aes(x = Venue, y = RxnTime, group = Venue)) +
   geom_boxplot() +
   labs(x = "Year", y = "Reaction Time (s)") +
-       scale_y_continuous(limits = c(0.1, 0.225),
-                            breaks = seq(0.1, 0.225, 0.025)) +
+       scale_y_continuous(limits = c(0.05, 0.275),
+                            breaks = seq(0.05, 0.275, 0.025)) +
   expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
                                               breaks = seq(1999, 2024, 2))
-#ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
+ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
 ```

### 2024-11-14 Manuscript/manuscript.tex
Added most current results to table 1

@@
    \toprule
-   Comparison & Permutation & Asymptotic \\ 
+   Comparison & Permutation & Asymptotic & Athletes & Number observations  \\ 
    \midrule
-   2019 vs 2022 & $0.028$ & $0.003$ \\
-   2022 vs 2023 & $0.011$ & $0.003$ \\
-   2022 National vs International & $<0.001$ & $0.002$ \\
+   2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
+   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\
+   2022 National vs International Men & $<0.001$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
+   2022 National vs International Women & $<0.001$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\
+   2022 vs 2023 Men & $<0.001$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
+   2022 vs 2023 Female & $<0.001$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
    \bottomrule
@@
 \end{table}
+\of{Table currently doesnt fit on the page but can be fixed by abbreviating some
+of the names in the comparison column.  I used <0.001 becuase that is what we
+had currently been using for any values from the permutation test where the value
+was < 2.2e-16 in R.  Professor Yan will know best, but are the sample sizes now
+sufficiently large enough that the permutation test is not needed? If I recall
+correctlty that was a correction for small sample sizes.}
 
@@
 \end{table}
-\of{needs to be updated}
+\of{needs to be updated with current values. But I did start to alter something
+then stopped.}
 

### 2024-11-15 Code/NewAnalysis.Rmd
added lots of qq plots at the bottom to look at

--- 2024-11-05 Code/NewAnalysis.Rmd
@@
 ```
+```{r}
+library(lme4)
+library(lattice)
+dotplot(ranef(dash_and_hurdles_with_dq_two_effect_gamma_log, condVar = TRUE))
+```
+```{r}
+dotplot(ranef(log_normal_model, condVar = TRUE))
+```
+
+```{r}
+random_effects <- ranef(log_normal_model)
+random_effects_df <- as.data.frame(random_effects$Venue)
+qqPlot(random_effects_df$`(Intercept)`, 
+       main = "Normal QQ Plot of Venue effect log normal model")
+```
+
+```{r}
+random_effects <- ranef(log_normal_model)
+random_effects_df <- as.data.frame(random_effects$Heat)
+qqPlot(random_effects_df$`(Intercept)`, 
+       main = "Normal QQ Plot of Heat effect log normal model")
+```
+```{r}
+random_effects <- ranef(dash_and_hurdles_with_dq_two_effect_gamma_log)
+random_effects_df <- as.data.frame(random_effects$Venue)
+qqPlot(random_effects_df$`(Intercept)`, 
+       main = "Normal QQ Plot of Venue effect gamma model")
+```
+```{r}
+random_effects <- ranef(dash_and_hurdles_with_dq_two_effect_gamma_log)
+random_effects_df <- as.data.frame(random_effects$Heat)
+qqPlot(random_effects_df$`(Intercept)`, 
+       main = "Normal QQ Plot of Heat effect gamma model")
+```

### 2024-11-15 Manuscript/manuscript.tex
Started work on discussion

@@
    2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\
-   2022 National vs International Men & $<0.001$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
-   2022 National vs International Women & $<0.001$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\
+   2022 Nat. vs Inter. Men & $<0.001$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
+   2022 Nat. vs Inter. Women & $<0.001$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\
    2022 vs 2023 Men & $<0.001$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
@@
 \end{table}
-\of{Table currently doesnt fit on the page but can be fixed by abbreviating some
-of the names in the comparison column.  I used <0.001 becuase that is what we
+\of{Table might still be too wide.  I used <0.001 becuase that is what we
 had currently been using for any values from the permutation test where the value
@@
 Table~\ref{tab:Clusrankresults} shows the results generated from both the 
-permutation and asymptotic rank-based tests for the three comparisons, 
+permutation and asymptotic rank-based tests for the six comparisons, 
 as discussed previously in Section~\ref{sec:rank}. The national versus international
-comparisons produced lower p-values than the 2019 versus 2022 and 2022 versus 2023comparisons.%, with the
-%national vs international permutation test producing a p-value of $2\cdot10^{-6}$.
-The highly significant permutation test result provides substantial evidence 
-that athletes who competed at the
+comparisons produced lower p-values than the 2019 versus 2022 and 2022 versus 2023
+comparisons. The highly significant permutation test results
+provides substantial evidence that both men and women who competed at the
 2022 World Track and Field Championships were on average, significantly faster
-at reacting than they were just a few months prior. For the 2019 versus 2022
-same athlete comparison with permutation test, the p-value was 
-$0.028$, which is significant at level $0.05$.  The result of
-comparing 2022 and 2023 was very similar (but not exactly the same) to the above
-result, as the p-value from the permutation test was $0.011$.
-Taken together, these three
-tests help show that athletes competing at the 2022 World Track and Field
-Championships had faster reaction times than when they had competed at other
-competitions.
+at reacting than they were just a few months prior. When looked at in context
+with other tests, these results show that athletes competing at the 2022 World
+Track and Field Championships had faster reaction times than when they had
+competed at other competitions.
 
@@
 
-We did not consider data from the women's 100 meter hurdle in the analysis.
-In addition to the concerns over reaction time differences in men and
-women, exploratory analysis showed that 2022 women were at best only
-slightly faster relative to other championships going back to 2003.
-Likewise, the rank-based comparisons resulted in p-values
-that were not drastically significant after adjusting for multiple tests.
-The lack of evidence from the women's data makes it 
-harder to conclude that the sensors used at the 2022 championships were faulty, 
-although it is unknown if the sensors for the men's 110 meter hurdle and women's
-100 meter hurdle were the same. On the other hand, this may be related to 
-another debated question regarding whether the reaction time barrier should be 
-the same for men and women.
+We set out to answer two questions: were athletes who competed at the 2022 World
+Track and Field Championships faster than at other races and is the 0.1 second
+reaction time barrier fair.  In regards to the first question, we do believe
+that athletes were faster than in comparable meets.  Ideally we would like to
+be able to draw data from a central database containing every World Athletics
+certified meet which we could use to reference athlete's reaction times.
+Unfortunately that is not available at the moment and nearly all of the results
+listed on the World Athletics website do not contain reaction times.  The results
+of the study could have been stronger if we had been able to look at every race
+that athletes like Devon Allen ran in 2022 and analyzed his results at the World
+Championship against many other data points.
+
+In addition to the GLMM results results presented above women, the supplementary
+file contains results about the women's GLMM results.  We fit a generalized
+gamma model to womens 100 meter dash and 100 meter hurdle data from 1999 to 2023
+to determine an appropriate reaction time barrier.  We chose to separate these
+two studies due to publications suggesting reaction time differences in men and
+women.
+
+\of{I think there are some good ideas in these 2 paragraphs but need help
+integrating them. Prof Schifano can you help with this?}
 

### 2024-11-17 Code/gamlss.R
fit with package gamlss

@@ -1,0 +1,80 @@
+## get the right dataset
+library(tidyverse)
+times <- read.csv("../Data/GLMMData.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
+
+## Please reduce the length of your variable and object names;
+## they are too long and distracting.
+
+## dash/hurdle/man
+dhm <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles")
+
+dhm$Venue <- as.factor(dhm$Venue)
+dhm$Heat <- as.factor(dhm$Heat)
+
+
+## Fit various distributions available from gamlss
+## Plot standarized residuals q-q plot
+library("gamlss")
+library("qqconf")
+
+## log-normal
+ln0  <- gamlss(RxnTime ~ 1, data = dhm, family = LOGNO)
+ln1  <- gamlss(RxnTime ~ random(Venue), data = dhm, family = LOGNO)
+ln2  <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = LOGNO)
+ln3  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = LOGNO)
+ln4  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + random(Heat), data = dhm, family = LOGNO)
+
+qq_conf_plot(residuals(ln0))
+qq_conf_plot(residuals(ln1))
+qq_conf_plot(residuals(ln2))
+
+## three-parameter: generalized gamma
+gg2  <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = GG)
+gg3  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
[…]

### 2024-11-17 Code/gamlss.R
added extraction of random effect variance

@@ -25,24 +25,44 @@ library("qqconf")
 
 ## log-normal
 ln0  <- gamlss(RxnTime ~ 1, data = dhm, family = LOGNO)
-ln1  <- gamlss(RxnTime ~ random(Venue), data = dhm, family = LOGNO)
+ln1a  <- gamlss(RxnTime ~ 1 + random(Venue), data = dhm, family = LOGNO)
 ln2  <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = LOGNO)
 ln3  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = LOGNO)
 ln4  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + random(Heat), data = dhm, family = LOGNO)
+
+AIC(ln0, ln1, ln2, ln3, ln4)
 
 qq_conf_plot(residuals(ln0))
 qq_conf_plot(residuals(ln1))
 qq_conf_plot(residuals(ln2))
 
 ## three-parameter: generalized gamma
+gg1  <- gamlss(RxnTime ~ random(Venue), data = dhm, family = GG)
 gg2  <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = GG)
 gg3  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+gg3a  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+gg3b  <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+gg3c  <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 gg4  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 
-AIC(gg3)
-AIC(gg4)
[…]

### 2024-11-17 Code/gamlss.R
fixed z-score qq plot dparams

@@ -47,7 +47,10 @@ gg4  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(
 
 
 AIC(gg1, gg2, gg3, gg3a, gg3b, gg3c, gg4) # gg3b is the best
-qq_conf_plot(residuals(gg3b))
+qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1)) ## z-scores have known mean and sd
+qq_conf_plot(gg3b$sigma.coefSmo[[1]]$coef)
+qq_conf_plot(gg3b$mu.coefSmo[[1]]$coef)
+
              
 ### random effects in mu
 gg3b$mu.coefSmo

### 2024-11-19 Code/gamlss.R
Added simulation code and generalized inverse gaussian. GIG not as good as GG

@@ -25,7 +25,7 @@ library("qqconf")
 
 ## log-normal
 ln0  <- gamlss(RxnTime ~ 1, data = dhm, family = LOGNO)
-ln1a  <- gamlss(RxnTime ~ 1 + random(Venue), data = dhm, family = LOGNO)
+ln1  <- gamlss(RxnTime ~ 1 + random(Venue), data = dhm, family = LOGNO)
 ln2  <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = LOGNO)
 ln3  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = LOGNO)
 ln4  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + random(Heat), data = dhm, family = LOGNO)
@@ -89,7 +89,7 @@ qq_conf_plot(residuals(we2))
 igau2  <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = IG)
 igau4  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + random(Heat), data = dhm, family = IG, control = gamlss.control(n.cyc = 40)) 
 
-qq_conf_plot(residuals(igau4) # good looking
+qq_conf_plot(residuals(igau4)) # good looking
 
 ## inverse gamma
 igam2  <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = IGAMMA)
@@ -101,3 +101,36 @@ igam2  <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = IG
 ## eg2 <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = exGAUS)
 
 ## qq_conf_plot(residuals(eg0))
+
+# Generalized Inverse Gaussian
+gig1 <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm, family = GIG, control = gamlss.control(n.cyc = 40))
+gig2 <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = GIG, control = gamlss.control(n.cyc = 40))
[…]

### 2024-11-19 Code/gamlss.R
additional models in comparison

@@ -43,10 +43,11 @@ gg3  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(
 gg3a  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 gg3b  <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 gg3c  <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+gg3d  <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 gg4  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 
 
-AIC(gg1, gg2, gg3, gg3a, gg3b, gg3c, gg4) # gg3b is the best
+AIC(gg1, gg2, gg3, gg3a, gg3b, gg3c, gg3d, gg4) # gg3b is the best
 qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1)) ## z-scores have known mean and sd
 qq_conf_plot(gg3b$sigma.coefSmo[[1]]$coef)
 qq_conf_plot(gg3b$mu.coefSmo[[1]]$coef)
@@ -74,7 +75,7 @@ ga4  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(
 
 AIC(ga2)
 AIC(ga4)
-qq_conf_plot(residuals(ga4))
+qq_conf_plot(residuals(ga4), dparams = list(mean = 0, sd = 1)) ## z-scores have known mean and sd
 
 
 ## weibull
@@ -83,13 +84,18 @@ we1  <- gamlss(RxnTime ~ random(Venue), data = dhm, family = WEI3)
 we2  <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = WEI3)
 
 AIC(we2)
-qq_conf_plot(residuals(we2))
[…]

### 2024-11-19 Manuscript/manuscript.tex
added gamlss methods

@@
 The data described in Sections~\ref{sec:databeyond} and~\ref{sec:dataworld} were
-analyzed with rank-based methods and a GLMM, respectively.
+analyzed with rank-based methods and a GAMLSS, respectively.
 
@@
 
-\subsection{GLMM}\label{sec:glmm}
-Based on an exploratory analysis using the GLMM model for the reaction times
-from the World Championships, we chose a gamma mixed-effect model as our 
-final model. As suggested by the log-likelihood and the Akaike Information
-Criterion (AIC), the normal error model on the log-scale was found to be
-much inferior to the gamma model which better captures the skewness of the
-data. Two random effects were found useful: one capturing the venue effect, which
+\subsection{GAMLSS}\label{sec:glmm}
+
+Based on an exploratory analysis, the reaction times are adequately
+modeled by a Generalized Gamma (GG) distribution with random effects in
+model parameters. The GG distribution has three parameters, denoted by
+$\text{GG}(\mu, \sigma, \nu)$ has density function
+\[
+f_Y(y \mid \mu, \sigma, \nu) = 
+\frac{|\nu| \theta^\theta}{\Gamma(\theta) y} 
+z^{\theta} \exp\left(-z \theta\right),
+\]
+
+for $y > 0$, $\mu > 0$, $\sigma > 0$, and $\nu \neq 0$,
+where $z = \left(\frac{y}{\mu}\right)^\nu$,
+$\theta = 1 / (\sigma^2 \nu^2)$.
+$\Gamma(\cdot)$ denotes the Gamma function.
+The GG distribution is highly flexible, encompassing several  
+well-known distributions as special cases, such as the
+Weibull ($\mu = \nu)$ and  Gamma $(\nu = 1)$ distributions.
+Its mean is
+$\mu \frac{\Gamma(\theta + 1 / \nu)}{\theta^{1 / \nu} \Gamma(\theta)}$,
+provided $\theta > -1 / \nu$. Here,
+$\mu$ scales the central tendency, $\sigma$ controls 
+dispersion, and $\nu$ determines skewness. This parameterization allows 
+the distribution to model asymmetric and heavy-tailed data effectively, making 
+it particularly suitable for reaction times.
+An implementation of this distribution is available from R package
+\texttt{gamlss.dist} \citep{rigby2019distributions}.
+
+Let $Y_{ijk}$ denote the reaction time of observation~$k$ in heat~$j$
+of year~$i$. Conditioning on a venue effect $v_i$ for year~$i$ 
+and a heat effect $h_{i/j}$ nested within each year~$i$, the
+distribution of $Y_{ijk}$ is
+$\text{GG}(\mu_{ijk}, \sigma_{ijk}, \nu)$, where
+\begin{align}
+\log(\mu_{ijk}) &= \beta_0 + v_i ,\\
+\log(\sigma_{ijk}) &= \gamma_0 + h_{i/j},
+\end{align}
+$v_i$ is normally distributed with mean zero and
+variance~$\sigma_v^2$, and $h_{i/j}$ is normally distributed with mean
+zero and variance~$\sigma_h^2$.
+The two random effects were found useful: one capturing the venue effect, which
 is used to contrast years, and the second being the heat effect, where every
@@
 amount of time athletes are on the starting blocks before the gun goes off.
-Thus, we can try to reduce this variability by accounting for it in the model
-through the heat effect.
+This model can be fit with R package \texttt{gamlss}
+\citep{stasinopoulos2024generalized}.
 
-Let $Y_{ijk}$ be the reaction time of observation~$k$ in heat~$j$ in year~$i$.
-Conditional on a venue effect $v_i$ of year~$i$ and a heat effect
-$h_{i/j}$ nested within each year~$i$, $Y_{ijk}$ is modeled by a
-gamma distribution. In the classic generalized linear model (GLM) sense,
-$Y_{ijk}$ has conditional mean $E(Y_{ijk} | v_i, h_{i/j}) > 0$ and a
-dispersion parameter~$\phi > 0$. The conditional mean is characterized by
-\[
-g\{E(Y_{ijk} | v_i, h_{i/j})\} = \alpha + v_i + h_{i/j},
-\]
-where $g$ is a known link function, $\alpha$ is a (fixed) intercept,
-$v_i$ is normally distributed with mean zero and
-variance~$\sigma_v^2$, and $h_{i/j}$ 
-is normally distributed with mean
-zero and variance~$\sigma_h^2$.
-Although \citet{lo2015idlink} suggested that the gamma 
-family with an identity link works well for reaction time in cognitive
-psychological research, the logarithm
-link function worked better for our data and additionally ensures the positivity
-of the conditional mean. The gamma mixed-effect model can be fit with the 
-\texttt{glmer()} function in R package \texttt{lme4} \citep{lme4}.
-
-To better understand this model, we can identify the gamma model in terms of the
-commonly used shape and scale parameters. For a gamma distribution with
-shape~$a$ and scale~$b$, we have mean $\mu = ab$ and variance $\nu = ab^2$. The
-variance as a function of the mean~$\mu$ in the GLM sense is
-$\text{var}(\mu) = \mu^2 / a$, which implies that the dispersion parameter is
-$\phi = 1 / a$. Therefore, for the gamma mixed-effect model, the conditional
-gamma distribution of the reaction time $Y_{ijk}$ given the venue
-effect~$v_i$ and heat effect~$h_{i/j}$
-has shape $a = 1 / \phi$ and scale
-$b = g^{-1}(\alpha + v_i + h_{i/j}) \phi$. The marginal
-distribution of $Y_{ijk}$ is a scale-mixture of gamma distributions, which can be
+Model diagnosis and tail analysis can be done with the fitted GG model
+from package \texttt{gamlss}. Normalized quantile residuals, or
+z-scores \citep{dunn1996randomized}, of the observations can be
+extracted with the \texttt{residuals} method of a \texttt{gamlss}
+object. The z-scores can then be checked with a Q-Q plot
+\citep{almeida2018ggplot2}. The marginal
+distribution of $Y_{ijk}$ is a scale-mixture of GG distributions, which can be
 easily simulated from once the parameters are estimated. Many

### 2024-11-19 Manuscript/manuscript.tex
minor

@@
 The rest of this paper is organized as follows. Section~\ref{sec:Data} 
-describes how the data were collected. The rank-sum test, detailed 
-in Section~\ref{sec:rank}, uses slightly different data from the 
-generalized linear mixed-effects model (GLMM) developed in 
-Section~\ref{sec:glmm}. Section~\ref{sec:Results} presents the results 
-of both analyses and discusses the implications of the reaction time 
-barrier. Finally, Section~\ref{sec:concludingremarks} outlines the 
-paper’s impact and limitations. All data and code for our analysis 
+describes the data collection and datasets used for the two
+objectives. Section~\ref{sec:rank} outlines the methods for
+objective~1, analyzing differences in reaction times across
+competitions, including the were collected. Section~\ref{sec:gamlss}
+details the methods for objective~2, the generalized additive model
+for location, scale, and shape (GAMLSS) to evaluate the appropriatenes
+of the 0.1-second threshold. Section~\ref{sec:Results} presents the
+results of both objectives and discusses the implications.
+Finally, Section~\ref{sec:concludingremarks} highlights the 
+paper’s impact, limitations, and potential for future research.
+All data and code for our analysis 
 are provided in the supplementary materials.
@@
 
-\section{Methods} \label{sec:Methods}
+\section{Methods} \label{sec:methods}
 
@@
 
-\subsection{GAMLSS}\label{sec:glmm}
+\subsection{GAMLSS}\label{sec:gamlss}
 
@@
 f_Y(y \mid \mu, \sigma, \nu) = 
-\frac{|\nu| \theta^\theta}{\Gamma(\theta) y} 
-z^{\theta} \exp\left(-z \theta\right),
+\frac{|\nu| \theta^\theta z^{\theta}}{\Gamma(\theta) y} 
+\exp\left(-z \theta\right),
 \]
-
 for $y > 0$, $\mu > 0$, $\sigma > 0$, and $\nu \neq 0$,
 where $z = \left(\frac{y}{\mu}\right)^\nu$,
-$\theta = 1 / (\sigma^2 \nu^2)$.
+$\theta = 1 / (\sigma^2 \nu^2)$, and
 $\Gamma(\cdot)$ denotes the Gamma function.
@@
 Its mean is
-$\mu \frac{\Gamma(\theta + 1 / \nu)}{\theta^{1 / \nu} \Gamma(\theta)}$,
+\[
+  \mu \frac{\Gamma(\theta + 1 / \nu)}
+  {\theta^{1 / \nu} \Gamma(\theta)},
+\]
 provided $\theta > -1 / \nu$. Here,

### 2024-11-19 Manuscript/manuscript.tex
tasks for owen to update the analysis with gamlss

@@
 
-\subsection{GLMM} \label{subsec:Results_GLMM}
+\subsection{GAMLSS} \label{subsec:Results_GLMM}
+
+\jy{Owen, please prepare the following:
+1) A table of the model parameter estimates (both point estimate and
+standard error). This is similar to the current table, with and
+without 2022 data. Let's focus on model gg3b and leave the model
+comparison results in the supplement.
+2) Figure: random effects at the venue level with/without 2022
+data. Similar to the current figure.
+3) Diagnostic plots: qq-plot for the gg3b model. QQ-plot for the
+normal random effects could be placed in supplement.
+4) Table: assessing the 0.1 threshold, similar to the current table.
+I'll take care of the writing when you get the results.}
 

### 2024-11-20 Code/RankComparison.Rmd

@@
 
-clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_hurdle_2022,
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = women_2022,
                 method = "ds",
@@
 
-#clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2023,
- #               method = "ds",
-  #              exact = TRUE, B = 1000000)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = men_2023,
+                method = "ds",
+                exact = TRUE, B = 1000000)
 ```

### 2024-11-20 Manuscript/reply.tex

@@
 
+\subsection*{To Associate Editor}
+
+\begin{comment}
+I encourage you to consider the issues raised by the reviewers carefully. I have
+little additional material to add, other than to emphasize the need to
+re-organize the manuscript to more clearly identify the key goals of the work at
+the start, and to provide a more concise and focused presentation of the key
+findings.
+\end{comment}
+
+Thank you for your feedback. We have carefully addressed the
+reviewers' comments and taken additional steps to streamline the
+manuscript, focusing on a clearer presentation of the key goals and
+findings from the outset. We appreciate your time and consideration in
+re-evaluating our submission.
+
 \subsection*{To Reviewer 1}
@@
 
-Thank you for these suggestions, we were able to expand the scope of our data
-by including results from the men's and women's 100 meter dash.  We considered
-adding additional sprinting events such as the 200 meter dash but our analysis
-suggested that reaction times were slower in the 200 meter dash.  After running
-a model that included a coefficient for event type on a data set that combined
-the men's 110 meter hurdle and 100 meter dash, the event term was not significant.
-However after we re-ran the model on the 110 meter hurdle, 100 and 200 meter dash,
-the effect became significant. When we peformed a TukeyHSD test to compare the
-times in the three categories: 100 meter dash, 110 meter hurdles, 200 meter dash,
-we found a p-value of 0.0020625 comparing the means of the 200 meter dash and
-the 110 meter hurdlesand a p-value of 0.0000174 comparing the means of the 100
-meter dash and the 200 meter dash.  This indicated to us that the reaction times
-for the 200 meter were too slow to fit into our study on extreme reaction times
-and thus we chose to exclude 200 meter data.
+Thank you for the valuable suggestions. We have extended our analysis
+by including data from both men's and women's 100-meter dash. This was
+justified by a GLMM fit to these data and the event type was not found
+to be significantly different in reaction time.
 
-This is not particularly suprising as athletes who compete in the longer
-events may not have as good of a start as those in the 100
-meter events because reaction time plays a much smaller role in the outcome of
-the race. While it is common for amateur and collegiate athletes to compete in
-multiple sprinting events, it seems reasonable that a world championship 100
-meter runner may be only a very good 200 meter runner.  We are trying to
-estimate the probabilty of observing a extreme reaction time to determine
-whether 0.1 seconds is fair but if we include data from competitions with slower
-reaction times, then 0.1 seconds is going to be more extreme.
+We also considered including additional sprinting events such as the
+200-meter dash; however, in a GLMM fit to an expanded dataset that
+additionally included 200-meter, the event type effect became
+significant. Subsequent pairwise comparisons using Tukey’s Honest
+Significant Difference (HSD) test revealed a p-value of 0.0021 when
+comparing the reaction times of the 200-meter dash and the 110-meter
+hurdles, and a p-value of 0.000017 when comparing the 100-meter dash
+and the 200-meter dash. These results indicate that reaction times in
+the 200-meter dash are significantly slower, suggesting that this
+event does not align well with our focus on extreme reaction times.
+Consequently, we decided to exclude the 200-meter data from our
+primary analysis.
 
-\jy{At the end of each reply, we need to state what changes we made to
-  the manuscript with their specific locations. The explanation here
-  (the exclusion of 200m dash) should be included in the paper as well.}
+This outcome is not unexpected, as athletes competing in longer sprint
+events, such as the 200-meter dash, may not prioritize reaction time
+as highly as those specializing in shorter sprints like the 100-meter
+dash. While it is common for amateur and collegiate athletes to
+compete across multiple sprint distances, elite athletes at the World
+Championship level may not exhibit equally strong performances in all
+sprint events. Our study aims to evaluate the probability of observing
+extreme reaction times, particularly in the context of the 0.1-second
+threshold. Including data from events with systematically slower
+reaction times would artificially inflate the extremity of a
+0.1-second reaction time, potentially biasing our conclusions.
+
+In the revision, we have added explanation about why 100-meter dash
+data for both men and wemen were included, while 200-meter dash data
+were not included in our analysis in Section~???, paragraph~???.
+The subsequent analyses in the manuscript have been updated with the
+expanded dataset that included both men and women's 100-meter dash
+data; see updated results in Sections~???.
 
@@
 average ability as many of observations in the data come from runners
+\jy{It's the number of runners with 1 data point that determines how
+  hard the athlete effect can be estimated. How many runers had only 1
+  data point?}
 with very few (1-3) data points, so taking an average of such a small sample may
@@
 
-The second reviewer shared a similar sentiment about not effectively
-presenting ideas so we we made improvements to the introduction section to try
-and more clearly present our ideas.  We shortened the literature review to only
-cite papers that were actually of interest to the paper and tried to clearly
-define the two purposes of the paper: to identify inconsistensies with the 2022
-reaction times and to decide of 0.1 seconds is a fair barrier. We try to be precise in
-our language as we do not have enough evidence to conclude that the sensors were
-calibrated differently, or that Seiko (who make the timing devices) is at fault
-for the 2022 reaction times being inexplicably faster. Thus we make general
-recommendations for World Athletics and note that the reaction times in 2022
-were unusually fast with no clear reason.
+We appreciate the feedback, which was echoed by the second
+reviewer. To address these concerns, we revised the introduction to
+better structure our arguments. We streamlined the literature review
+to focus only on studies directly relevant to our analysis and
+clarified the paper’s two main objectives: (1) identifying
+inconsistencies in the 2022 reaction times and (2) assessing whether
+the 0.1-second threshold is fair. We avoided speculating on sensor
+calibration or implicating Seiko due to a lack of direct evidence, and
+instead, we provide general recommendations for World Athletics while
+noting that the unusually fast reaction times in 2022 remain
+unexplained.
 
@@
 
-Thank you for these encouraging words, we have added additional data to both
-the GLMM and clusrank section in order to strengthen our arguments: there was
-an anomaly in 2022 and that the reaction time barrier of 0.1 seconds is not
-grounded in formal statistical analysis.
-
-\jy{State what expansions we
-  have done.}
-\of{After we are wrapped up with the manuscript it will be easier to identify
-all the improvements we made. Let's just leave this here for now.}
+Thank you for the positive feedback. We expanded both the rank
+comparison and the GLMM sections with additional data to strengthen our
+analysis. These enhancements support our findings of an anomaly in
+2022 and highlight that the 0.1-second reaction time threshold lacks a
+foundation in formal statistical evidence.
 
@@
 
-We have re-worked the introduction and discussion sections of the paper to more
-clearly outline the goals of the paper.  The organization of the paper was
-something that both reviewers commented on so we did our best to try and
-streamline the two key points of the paper: to decide of 0.1 seconds is fair,
-and to identify inconsistensies with the 2022 reaction times. With these two
-points emphasized we tried to remove sections that did not support these central
-ideas.  In regards to the language surrounding the rank comparison text and the
-timing device at the 2022 Championships we are intentionally vague as we do not
-have evidence to make a stronger claim.  For example, we cannot claim that the
-timing device was faulty as miscalibrated as there is no public information
-available about that.  We do not want to overstep and put ourselves or the
-publication at risk of defamation claims from Seiko because we called their
-devices faulty when they are not.  Instead we suggest to the reader that 2022
-was an unusually fast year and comment that there is no explanation as to why
-athletes improve reaction times over the span of a few months and let the reader
-decide what they think is the underlying cause of the faster reaction times.
+We have revised the Introduction and Discussion sections to clearly
+outline the goals of the study. Both reviewers noted issues with
+organization, so we streamlined the manuscript to emphasize the two
+main objectives: (1) identifying inconsistencies in the 2022 reaction
+times and (2) assessing whether the 0.1-second threshold is fair. We
+removed unrelated sections to keep the focus on these central
+points. Regarding the rank comparison and timing devices at the 2022
+Championships, we intentionally used cautious language, as there is no
+evidence to claim a miscalibration or fault in Seiko’s timing
+devices. To avoid potential legal risks, we noted the unusually fast
+reaction times in 2022 without making unsubstantiated claims. Instead,
+we allow readers to interpret the underlying cause of the anomaly
+based on the presented data.
 
@@
 
-We are attempting to answer question 1 and have included reaction times from
-additional sprinting events from World Championships.  The inclusion of
-additional data was also suggested by the first reviewer and we have provided
-our explanation of what data we used there.
+Thank you for the feedback. We are attempting to address question~(1)
+and have expanded our analysis to include reaction times from
+additional sprint events at the World Championships, including both
+men and women's 100-meter data. This broader dataset helps provide a
+more comprehensive assessment of whether the venue in 2022 was
+unusually fast.
 
@@
 \end{comment}
-We will start by addressing the comments related to runners who have been
-disqualified.  Some of the runners who were disqualified had a negative reaction
-time, indicating they were trying to predict the starting gun and actually left
-before the gun even fired.  These runners cannot be included in the analysis
-because the point of the GLMM section is to question the current reaction barrier
-methods by determining how fast athletes can react in; but athletes with a
-negative time left before the gun even went off (Also practically speaking
-a negative reaction time makes little sense).  As for the roughly 20 athletes
-across men's and women's data without negative times but were disqualified, we
-analyzed how the inclusion of these athletes effected our results.  What we found
-was that the probability of a reaction time below 0.1 seconds increased from
-0.0029 to 0.0039 and we have attached the R code so that you can see our work.
-\of{Not sure the technical details of how this is going to work but we will come
-back to this before we submit.} Additionally there were similar parameter values
-for the standard deviations of the random effects and we decided that we were
-going to keep these observations in our data.
 
-As for why the particular model was
-chosen, we looked at several types of models and chose the one that we thought
-fit the residual qq plot the best and had the largest log likelihood. After
-selecting the gamma generalized linear model we again looked at several link
-functions such as "identity", "log",
+Thank you for raising these points. Regarding the exclusion of
+disqualified runners, we chose not to include athletes with negative
+reaction times, as these indicate an attempt to predict the starting
+gun rather than a true reaction. Negative reaction times do not align
+with the purpose of the GLMM analysis, which aims to evaluate the
+feasibility of the 0.1-second reaction threshold. Including negative
+times would introduce results where the reaction occurred before the
+gun fired, making them unsuitable for this analysis.
 
-\of{Need to justify why we did what we did and disadvantages of including the
-disqualified
-Do this in the paper and reply. Say section where we made any changes. Give
-justifications for any reasons for disagreeing}
+For the approximately 20 disqualified athletes with positive reaction
+times, we examined the impact of including them in the analysis. The
+probability of observing a reaction time below 0.1 seconds increased
+\jy{Are these based on the expanded dataset or the original dataset?}
+slightly from 0.0029 to 0.0039, with similar parameter estimates for
+the random effects. Given these results, we opted to include these
+observations in the dataset.
 
-\of{Will clean this up based on Prof Yan and Schifano's comments to 11/5 email}
+Distribution: \jy{need to check R package gamlss.}
 
-In terms of the choice of distribution, Brosnan and Harrison used a exponentially
-modified gaussian distribution using athletes sex, ruling periods, competition
-rounds. The article is now behind a pay wall.
-\jy{Not a reason for not citing it. We can still find pdf online.}
-
-\jy{Using log-normal distribution for the reaction times is quite easy
-  to do. You just take log of the reaction times, and then model them
-  in a GLMM with Guassian family. I think we did this, and decided it
-  was not good perhaps from the residual qq plots. Please find them
-  and add a discussion into the paper.}
-
-\jy{This references has been cited 31 times. Some of these are closely
-  related to our topic and should be cited as appropriate.}
+\jy{Find these references and cite them:
+Sprint Start Regulation in Athletics: A Critical Review.
+Milloz M, Hayes K, Harrison AJ.
+Sports Med. 2021 Jan;51(1):21-31. doi: 10.1007/s40279-020-01350-4.
+PMID: 33125639 Review.
+Reaction time aspects of elite sprinters in athletic world championships.
+Tønnessen E, Haugen T, Shalfawi SA.
+J Strength Cond Res. 2013 Apr;27(4):885-92. doi: 10.1519/JSC.0b013e31826520c3.
+PMID: 22739331
+}
 
@@
 
-Unfortunately data such as this is not easy to come by.  We found as much as we
-could online, but there is simply not enough data to justify a formal analysis
-of his reaction times across lower level meets.  Many track and field races do
-not record data on reaction times until at the very high level (national
-and world championship).  Simply put, reaction times are not heavily researched
-and thus the data material in insufficient.
-\jy{We have added a discussion in paragraph ??? of the Discussion Section.}
-
-\jy{Every point we cannot answer needs to have some response reflected
-  in discussions in the manuscript.}
+Thank you for this suggestion. Unfortunately, comprehensive data on
+Devon Allen's reaction times is not readily available. We gathered as
+much data as we could find online, but there is insufficient
+information from lower-level meets, as reaction times are typically
+only recorded at major competitions (e.g., national and world
+championships). Given the limited data, we could not conduct a formal
+statistical analysis of his reaction times. However, we have added a
+brief discussion of this issue in paragraph~???] of the Discussion Section.
 
@@
 
-\begin{comment}I have no idea how to interpret the numbers displayed in Table 2. These
+\begin{comment}
+I have no idea how to interpret the numbers displayed in Table 2. These
 need to be made interpretable, or they add nothing to the paper.
@@
 
-Yes, this has been resolved.
+Yes, this has been rephrased for clarification.
 
@@
 
-\subsection*{Associate Editor}
-\jy{Put this section before the reviewers.}
-
-\begin{comment}
-I encourage you to consider the issues raised by the reviewers carefully. I have
-little additional material to add, other than to emphasize the need to
-re-organize the manuscript to more clearly identify the key goals of the work at
-the start, and to provide a more concise and focused presentation of the key
-findings.
-\end{comment}
-
-We have taken additional steps to streamline the manuscript and enhance the
-presentation of our work. Thank you for your time and we appreciate re-considering
-our submission.
-
-%\bibliographystyle{chicago}
-%\bibliography{citations}
+\bibliographystyle{apalike}
+\bibliography{citations}
 \end{document}

### 2024-11-20 Archive/Plots.Rmd
Overhauled document to only include relevant plots, listed in order of appearance currently in the manuscript

@@
 
-```{r}
-library(ggplot2)
-ggplot(df, aes(x = Venue, y = RxnTime, group = Venue)) +
-  geom_boxplot() +
-  labs(x = "Year", y = "Reaction Time (s)") +
-       scale_y_continuous(limits = c(0.05, 0.275),
-                            breaks = seq(0.05, 0.275, 0.025)) +
-  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
-                                              breaks = seq(1999, 2024, 2))
-ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
-```
+# Figure 1
 
-```{r}
-library(tidyr)
-library(ggplot2)
-nineteen <- read.csv("../Archive/2019vs2022.csv")
-#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
-colnames(nineteen) <- c("Athlete", "Country", "2019 Heats", "2019 Semifinals", "2019 Finals", "2022 Heats", "2022 Semifinals", "2022 Finals")
-nineteen <- nineteen[, -c(1,2)]
-df_long <- tidyr::pivot_longer(nineteen, cols = everything(), names_to = "Column", values_to = "Time")
-df_long$Column <- factor(df_long$Column, levels = c("2019 Heats", "2019 Semifinals", "2019 Finals", "2022 Heats", "2022 Semifinals", "2022 Finals"))
-rank1 <- ggplot(df_long, aes(x = Column, y = Time)) +
-  geom_point() +
-  labs(y = "Reaction Time", x = "") +
-  annotate("text", y = 0.21, x = "2022 Semifinals", label = "2019 vs 2022", cex = 5) +
-  ylim(0.08, 0.22) +
-  theme_minimal()
-```
+### Mens 2019 vs 2022
 
-```{r}
-library(tidyr)
-library(ggplot2)
-twentythree <- read.csv("../Data/2022vs2023.csv")
-#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
-colnames(twentythree) <- c("Athlete", "Country", "2022 Heats", "2022 Semifinals", "2022 Finals", "2023 Heats", "2023 Semifinals", "2023 Finals")
-twentythree <- twentythree[, -c(1,2)]
-df_long <- tidyr::pivot_longer(twentythree, cols = everything(), names_to = "Column", values_to = "Time")
-df_long$Column <- factor(df_long$Column, levels = c("2022 Heats", "2022 Semifinals", "2022 Finals", "2023 Heats", "2023 Semifinals", "2023 Finals"))
-rank2 <- ggplot(df_long, aes(x = Column, y = Time)) +
-  geom_point() +
-  labs(y = "Reaction Time", x = "") +
-  annotate("text", y = 0.21, x = "2023 Semifinals", label = "2022 vs 2023", cex = 5) +
-  ylim(0.08, 0.22) +
-  theme_minimal()
-```
-
-```{r}
-library(tidyr)
-library(ggplot2)
-national <- read.csv("../Data/NatVSWorld.csv")
-#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
-colnames(national) <- c("Athlete", "Country", "Nat. Heats", "Nat. Semifinals", "Nat. Finals", "Inter. Heats", "Inter. Semifinals", "Inter. Finals", "Data Source")
-national <- national[, -c(1,2, 9)]
-df_long <- tidyr::pivot_longer(national, cols = everything(), names_to = "Column", values_to = "Time")
-df_long$Column <- factor(df_long$Column, levels = c("Nat. Heats", "Nat. Semifinals", "Nat. Finals", "Inter. Heats", "Inter. Semifinals", "Inter. Finals"))
-rank3 <- ggplot(df_long, aes(x = Column, y = Time)) +
-  geom_point() +
-  labs(y = "Reaction Time", x = "") +
-  ylim(0.08, 0.22) +
-  annotate("text", y = 0.21, x = "Inter. Semifinals", label = "2022 National vs International", cex = 5) +
-  theme_minimal()
-```
-
-```{r}
-library(gridExtra)
-rank_scatter <- grid.arrange(rank1, rank2, rank3, ncol = 1)
-ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 7.5, width = 7)
-```
-
-```{r}
-library(lme4)
-library(ggplot2)
-library(dplyr)
-times <- read.csv("../Data/GLMMData.csv")
-colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
-with2022 <- subset(times, (Type == "F" | Type == "S") & 
-                       Gender == "M" & Time != 'DQ' & Event == "Hurdles")
-with2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
-                               data = with2022, family = Gamma(link = "log"))
-ranef_results <- data.frame(ranef(with2022_two_effect_gamma_log))
-venues <- subset(ranef_results, ranef_results["grpvar"] == "Venue")
-venues$year <- as.character(venues$grp)
-venues$year <- as.numeric(venues$year)
-
-ggplot(venues, aes(x=year, y = condval)) + geom_point() + 
-  labs(x = "Venue", y = "Venue Effect") +
-  scale_x_continuous(limits = c(1998,2024), breaks = seq(1999, 2024, 2)) +
-  scale_y_continuous(limits = c(-0.1, 0.1))
-```
-
-```{r}
-x_values <- seq(-3, 3, length.out = 1000)  # Generate a sequence of x values
-y_values <- dnorm(x_values, mean = 0, sd = sqrt(0.04))  # Calculate the corresponding y values for the normal distribution
-
-normal_df <- data.frame(x = x_values, y = y_values)
-ggplot(normal_df, aes(x = x, y = y)) +
-  geom_line(color = "black", size = 1) +
-  geom_point(data = venues, aes(x = condval, y = 0), color = "red", size = 2) +
-  labs(title = "Normal Distribution (Mean = 0, SD = sqrt(0.4))",
-       x = "Value",
-       y = "Density") +
-  theme_minimal() + xlim(-1, 1)
-```
-
-```{r}
-library(lme4)
-library(ggplot2)
-library(dplyr)
-times <- read.csv("../Data/GLMMData.csv")
-colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
-without2022 <- subset(times, (Type == "F" | Type == "S") & 
-                           Gender == "M" & Time != 'DQ' & Venue != 2022 & Event == "Hurdles")
-without2022_two_effect_gamma_log <- glmer(RxnTime ~ (1 | Venue) + (1 | Heat), 
-                               data = without2022, family = Gamma(link = "log"))
-ranef_results <- data.frame(ranef(without2022_two_effect_gamma_log))
-venues <- subset(ranef_results, ranef_results["grpvar"] == "Venue")
-venues$year <- as.character(venues$grp)
-venues$year <- as.numeric(venues$year)
-
-venue_effects_without2022 <- ggplot(venues, aes(x=year, y = condval)) +
-  geom_point() + 
-  labs(x = "Venue", y = "Venue Effect") +
-  scale_x_continuous(limits = c(1998,2024), breaks = seq(1999, 2024, 2)) +
-  scale_y_continuous(limits = c(-0.1, 0.1))
-```
-
-```{r}
-library(gridExtra)
-venue_effects_combined <- grid.arrange(venue_effects_with2022,
-                                       venue_effects_without2022, ncol = 1)
-ggsave(filename = "ComparisonOfVenueEffects.pdf", path = "../Manuscript",
-       plot=venue_effects_combined, height = 6.5, width = 7)
-```
-
-### 100 meter dash plots
-```{r}
-dash <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != "DNS")
-                   & Event == "Dash")
-library(ggplot2)
-dash_plot <- ggplot(dash, aes(x = Venue, y = RxnTime, group = Venue)) +
-  geom_boxplot() +
-  labs(x = "Year", y = "Reaction Time (s)") +
-       scale_y_continuous(limits = c(0, 0.25),
-                            breaks = seq(0, 0.25, 0.05)) +
-  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
-                                              breaks = seq(1999, 2024, 2))
-
-hurdle_plot <- ggplot(with2022, aes(x = Venue, y = RxnTime, group = Venue)) +
-  geom_boxplot() +
-  labs(x = "Year", y = "Reaction Time (s)") +
-       scale_y_continuous(limits = c(0, 0.25),
-                            breaks = seq(0, 0.25, 0.05)) +
-  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
-                                              breaks = seq(1999, 2024, 2))
-library(gridExtra)
-grid.arrange(dash_plot, hurdle_plot, ncol = 1)
-```
-```{r}
-dash <- subset(times, (Type == "F" | Type == "S") & Gender == "M" & (Time != 'DQ' | Time != "DNS")
-                   & Event == "Dash")
-library(ggplot2)
-dash_plot <- ggplot(dash, aes(x = Venue, y = RxnTime, group = Venue)) +
-  geom_boxplot() +
-  labs(x = "Year", y = "Reaction Time (s)") +
-       scale_y_continuous(limits = c(0.1, 0.25),
-                            breaks = seq(0.1, 0.25, 0.05)) +
-  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
-                                              breaks = seq(1999, 2024, 2))
-
-hurdle_plot <- ggplot(with2022, aes(x = Venue, y = RxnTime, group = Venue)) +
-  geom_boxplot() +
-  labs(x = "Year", y = "Reaction Time (s)") +
-       scale_y_continuous(limits = c(0.1, 0.25),
-                            breaks = seq(0.1, 0.25, 0.05)) +
-  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
-                                              breaks = seq(1999, 2024, 2))
-library(gridExtra)
-grid.arrange(dash_plot, hurdle_plot, ncol = 1)
-```
 ```{r}
@@
 ```
+
+### Womens 2019 vs 2022
 
@@
 
+### 2022 Mens National vs International
+
 ```{r}
@@
 ```
+
+### Womens 2022 National vs International
 
@@
 
+### Men's 2022 vs 2023
+
 ```{r}
@@
 ```
+
+### Womens 2022 vs 2023
 
@@
 
+### Combines the previous 6 graphs onto 1 plot
+
 ```{r}
@@
 ```
+
+# Figure 2
+
+```{r}
+library(ggplot2)
+ggplot(df, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.05, 0.275),
+                            breaks = seq(0.05, 0.275, 0.025)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
+                                              breaks = seq(1999, 2024, 2))
+ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
+```
+
+# Figure 3
+
+setup
+```{r}
+library("gamlss")
+library("qqconf")
+dhm <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime > 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles")
+
+dhm$Venue <- as.factor(dhm$Venue)
+dhm$Heat <- as.factor(dhm$Heat)
+gg3b  <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+
+dhm_no2022 <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime > 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
+  filter(Venue != 2022) |>
+  mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))
+
+gg3b_no2022  <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm_no2022, family = GG, control = gamlss.control(n.cyc = 40))
+
+```
+
+plot 1
+```{r}
+venue_effects <- as.data.frame(gg3b$mu.coefSmo[[1]]$coef)
+colnames(venue_effects) <- c("Effect Value")
+
+venue_plot <- ggplot(venue_effects, aes(x = rownames(venue_effects),
+                                               y = `Effect Value`)) +
+  geom_point() + 
+  labs(x = "Venue", y = "Venue Effect")+
+  scale_y_continuous(limits = c(-0.15, 0.1))
+```
+plot 2
+```{r}
+venue_effects_no2022 <- as.data.frame(gg3b_no2022$mu.coefSmo[[1]]$coef)
+venue_effects_no2022["2022", ] <- 1 #Arbitrary number out of range of y scale so
+#that the x axis line up
+
+colnames(venue_effects_no2022) <- c("Effect Value")
+
+venue_plot_no2022 <- ggplot(venue_effects_no2022,
+                                   aes(x = rownames(venue_effects_no2022),
+                                       y = `Effect Value`)) +
+  geom_point() + 
+  labs(x = "Venue", y = "Venue Effect") +
+  scale_y_continuous(limits = c(-0.15, 0.1))
+```
+### Now we combine for final product
+```{r}
+library(gridExtra)
+venue_effects_combined <- grid.arrange(venue_plot,
+                                       venue_plot_no2022, ncol = 1)
+ggsave(filename = "ComparisonOfVenueEffects.pdf", path = "../Manuscript",
+       plot=venue_effects_combined, height = 6.5, width = 7)
+```
+# Figure 4
+```{r}
+
+pdf(file = "../Manuscript/GGResiduals.pdf", height = 4, width = 6)
+qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1))
+dev.off()
+```
+
+# Figure 5
+```{r}
+sim_time <- function(model, RxnTime) {
+  simfit <- function(model, n = 10000000) {
+
+    sd_Venue <- model$mu.coefSmo[[1]]$sigb
+    sd_Venue_heat <- model$sigma.coefSmo[[1]]$sigb[1]
+
+    mu <- exp(model$mu.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue))
+    sigma <- exp(model$sigma.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue_heat))
+    nu <- model$nu.coefficients[1]
+
+    x <- rGG(n, mu = mu, sigma = sigma, nu = nu)
+    x
+  }
+  x <- simfit(model)
+  plot(density(x), xlim = c(0, 0.3), main = "", xlab = "")
+  curve(dGG(x, mu = exp(model$mu.coefficients[1]),
+            sigma = exp(model$sigma.coefficients[1]),
+            nu = model$nu.coefficients[1]),0, 0.3, add = TRUE, col = "blue",
+        main = "", xlab = "")
+}
+pdf(file = "../Manuscript/GGDensity.pdf", height = 4, width = 6)
+sim_time(gg3b, 0.1)
+dev.off()
+```

### 2024-11-20 Code/gamlss.R
added some additional code to determine reaction barrier, had to create models that excluded 2022

@@ -10,8 +10,8 @@ colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Even
 dhm <- times |>
   filter(Type == "F" | Type == "S") |>
   filter(Gender == "M") |>
-  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
-  filter(RxnTime != 0) |>
+  filter(Time != "DNS") |>
+  filter(RxnTime > 0) |>
   filter(Event == "100 Dash" | Event == "110 Hurdles")
 
 dhm$Venue <- as.factor(dhm$Venue)
@@ -139,5 +139,57 @@ sim_time <- function(model, RxnTime) {
   print(mean(x < RxnTime))
 }
 sim_time(gg3b, 0.1)
+sim_time(gg3b, 0.09)
+sim_time(gg3b, 0.08)
+
+#Outputs recommended reaction time barrier based on inputted probability
+sim_prob <- function(model, prob) {
+  simfit <- function(model, n = 10000000) {
+    
+    sd_Venue <- model$mu.coefSmo[[1]]$sigb
+    sd_Venue_heat <- model$sigma.coefSmo[[1]]$sigb[1]
+    
+    mu <- exp(model$mu.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue))
+    sigma <- exp(model$sigma.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue_heat))
+    nu <- model$nu.coefficients[1]
+    
+    x <- rGG(n, mu = mu, sigma = sigma, nu = nu)
+    x
+  }
+  x <- simfit(model)
+  print(quantile(x, probs = prob))
+}
+sim_prob(gg3b, .001)
+sim_prob(gg3b, .0001)
 
 
+## For reproducibility and to ensure that we got the same results. Run the function
+## that takes a GG model as input and prints all important parameters.
+model_print <- function(model) {
+  cat("AIC: ", AIC(model), "\n")
+  cat("mu (intercept): ", model$mu.coefficients[1], "\n")
+  cat("sigma (intercept): ", model$sigma.coefficients[1], "\n")
+  cat("nu: ", model$nu.coefficients[1], "\n")
+  cat("sigma_v or std dev of venue effect: ", model$mu.coefSmo[[1]]$sigb, "\n")
[…]

### 2024-11-20 Manuscript/manuscript.tex
requested a figure on fitted density
responded to Prof. Yan's to do list

@@
 I'll take care of the writing when you get the results.}
+\of{1) Is resolved in table 2, 2) is resolved in figure 4, 3) is resolved in
+figure 3. 4) is resolved in tables 3 and 4.}
+
+\jy{Owen: an additional figure of the GG density without random
+  effects (use \texttt{curve(dGG(x, mu = , sigma = , nu =  ), 0, 0.03)}) and
+  with random effects (through \texttt{density()})}
+\of{Please see figure 5}
 
@@
   \centering
-  \caption{AIC and estimated parameters for the three models fitted to the
-    data.}
+  \caption{AIC and estimated parameters for the gg3b model fit}
   \label{tab:Gamma_parameters}
@@
     \toprule
-    Data set & Model & AIC & $\sigma_h$ & $\sigma_v$ & $\alpha$ & $\sqrt{\phi}$ \\
+    Data set & AIC & $\mu$ & $\sigma$ & $\nu$ & $\sigma_v$ & $\sigma_h$ \\
     \midrule
-    Excluding 2022 &
-      Venue Effect Only & $-1862.2$ &   & 0.048 & $-1.857$ & 0.152\\
-     & Heat Effect Only & $-1910.9$ & 0.069 &  & $-1.877$ & 0.136\\
-     & Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.868$ & 0.136\\[1ex]
-    Including 2022 &
-    Venue Effect Only & $-3798.8$ &   & 0.054 & $-1.871$ & 0.149\\
-    & Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
-    & Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.881$ & 0.134\\
+    Excluding 2022 & -3660 & -1.900 & -2.192 & -1.315 & 0.0434 & 0.3259 \\
+    Including 2022 & -4103 & -1.910 & -2.200 & -1.178 & 0.0576 & 0.3198 \\
     \bottomrule
@@
 \end{table}
-\of{needs to be updated with current values. But I did start to alter something
-then stopped.}
+\of{Created these with the function I wrote model-name at bottom of gamlss.R. Do
+any other parametes need to be added?}
 
@@
   \centering
+  \includegraphics{GGResiduals}
+  \caption{Residuals from the GG model}
+  \label{fig:Residuals}
+\end{figure}
+
+\begin{figure}[tbp]
+  \centering
   \includegraphics{ComparisonOfVenueEffects}
@@
   \label{fig:VenueEffects}
+\end{figure}
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics{GGDensity}
+  \caption{The densities of the GG distribution with random effects for heat
+  and venue (black), and without random effects (blue)}
+  \label{fig:DensityPlot}
 \end{figure}
@@
    \midrule
-   Excluding 2022 & $1.10\cdot10^{-5}$ & $2.34\cdot10^{-4}$ &  $2.36\cdot10^{-3}$  \\ 
-   Including 2022 & $1.71\cdot10^{-5}$ & $3.01\cdot10^{-4}$ & $2.90\cdot10^{-3}$ \\
+   Excluding 2022 & $5.31\cdot10^{-5}$ & $3.53\cdot10^{-4}$ &  $1.94\cdot10^{-3}$  \\ 
+   Including 2022 & $6.84\cdot10^{-5}$ & $4.95\cdot10^{-4}$ & $2.76\cdot10^{-3}$ \\
    \bottomrule
@@
 \end{table}
-\of{table has been updated}
+\of{table has been updated with gg3b results}
 
@@
    \midrule
-   Excluding 2022 & $0.096$ & $0.087$ \\ 
-   Including 2022 & $0.095$ & $0.086$ \\
+   Excluding 2022 & $0.096$ & $0.083$ \\ 
+   Including 2022 & $0.094$ & $0.082$ \\
    \bottomrule

### 2024-11-20 Manuscript/manuscript.tex
just started, still reading Data section

@@
 and Field Championships. The news section of the website featured 
-articles such as \citep{johnson2022data} and \citep{johnson2022was} 
+articles such as \citet{johnson2022data} and \citet{johnson2022was} 
 by Robert Johnson, the site’s founder. Johnson argued that there may 
@@
 linear mixed-effects model with random effects for both venue and heat. 
+\eds{edit previous as needed for new GAMLSS analysis}
 This model estimates the probability of reaction times falling below 
@@
 objective~1, analyzing differences in reaction times across
-competitions, including the were collected. Section~\ref{sec:gamlss}
+competitions, including the were collected. 
+\eds{end of previous sentence does not make sense; some text may have been 
+mistakenly deleted?}
+Section~\ref{sec:gamlss}
 details the methods for objective~2, the generalized additive model
@@
 There are two types of data that are explored throughout this paper. For the
-rank-sum tests described in Section~\ref{sec:rank}, the data comes from male 
-athletes who competed in the 110 meter hurdles and 100 meter dash, and female
-athletes who competed 100 meter hurdles and 100 meter dash.
+rank-based comparisons described in Section~\ref{sec:rank}, the data described in 
+Section~\ref{sec:databeyond} come from male 
+athletes who competed in the 110 meter hurdles and \eds{or?} 100 meter dash, and 
+female athletes who competed 100 meter hurdles and \eds{or?} 100 meter dash.
 Athletes must have competed in the 2022 World Championships and another competition 
-(2019 World Championships, 2023 World Championships, or 2022 national 
-championships). For the GLMM described in Section~\ref{sec:glmm}, data comes
+(2022 national championships, 2019 World Championships, or 2023 World Championships) 
+in order to be included in the rank-based comparisons. 
+For the GAMLSS analysis described in Section~\ref{sec:gamlss}, the data described 
+in Section~\ref{sec:dataworld} come
 from every world championships from 1999 to 2023 and includes reaction times from
@@
 available, we are pleased to include it in both of our analyses, and it did not
-significantly change our results \citep{WAData}.
+significantly change our results \citep{WAData}. \eds{Should we remove this 
+last sentence, since now likely 2024 data is available?}
 
@@
 dataset of the four United States athletes is too small to perform any 
-meaningful statistical analysis, but the data was expanded to include reaction 
+meaningful statistical analysis, so the data for the rank-based analyses was 
+expanded to include reaction 
 times from more athletes who competed at other national competitions (many of which were 
@@
 Should we make a table of some sort? It's going to be very wordy without one}
+\eds{We are only doing two analyses with this data, right, one for women 
+and one for men?  So perhaps only two sentences needed here? But then 4 sentences
+in the next subsection?  Maybe you can just report ranges of cluster sizes and 
+median cluster size?}
 
-We chose to include preliminary rounds in this section as well as in
+We chose to include reaction times from preliminary heats, in addition to 
+reaction times from semifinals and finals, in this section and in
 Section~\ref{sec:data2019}, as removing them would have resulted in a much
@@
 larger dataset described in Section~\ref{sec:dataworld}, but we can also use it
-similarly to how we used national level in Section~\ref{sec:datanational}. We
-can compare the times of athletes who participated in the 2022 Championships with
-their times at another World Championships: 2019 and/or 2023.
+similarly to how we used national competitions in Section~\ref{sec:datanational}. 
+We can compare the reaction times of athletes who participated in the 2022 World 
+Championships with their times at another World Championships: 2019 and/or 2023.
 
@@
   \includegraphics{RankScatterPlots}
-  \caption{This collection of scatterplots of data used in the rank comparison
-  section shows that times in 2022 were on average lower. On the horizontal
+  \caption{Reaction times for athletes who competed at the 2022 World Championships
+	and at another championship (2022 national, 2019 World, or 2023 World) at which 
+	they competed. On the horizontal
   axis below each graph "H", "S", and "F" refer to the heats, semifinals, and 
   finals respectively. Please note that in the last row the 2022 times are to
-  the right of the 2023 times.}
+  the right of the 2023 times.
+	\eds{Please check my edited caption; also, I suggest rearranging the rows to 
+	use the same ordering as the data was described: put national vs international 
+	in first row, 2019 vs 2022 in second row, and 2022 vs 2023 in third row}
+	}
   \label{fig:RankScatterplots}
@@
 are compared to their reaction times at another at which they competed. 
-It is worth noting that each instance of the 2022 data is different, because
-those who competed at both the 2019 and 2022 World Championships will be
-different athletes than those who competed at both the 2022 and 2023 World
-Championships. This data was only for athletes who competed
-in at least one race at each championship, however it is possible for athletes
-to have competed in one race in 2019 and two races in 2022. The 2022 World
+It is worth noting that for each row of the figure, the athletes compromising 
+the 2022 data is different.  This is because the set of athletes who competed at 
+both the 2019 and 2022 World Championships is not the same as the set of athletes 
+who competed at both the 2022 and 2023 World Championships. %This data was only for athletes who competed
+%in at least one race at each championship, however it is possible for athletes
+%to have competed in one race in 2019 and two races in 2022. 
+Athletes must have competed in at least one race at each championship to be 
+included.  The 2022 World
 Championship heats and semifinals both had on average lower reaction times than
-their respective other (2019, 2023 or 2022 national) races, and the 2022 World 
-Championship finals were substantially faster.  It
-is also worth noting is that the fastest reaction time in the 2022 World
+their respective other (2019, 2023 or 2022 national) races, and the reaction times 
+from the 2022 World Championship finals were substantially faster. 
+\eds{I would consider removing or rephrasing the last part of the previous 
+sentence. With the women's data, it is not as clear the 2022F is lower than 2022S.}
+\eds{I also think we need to comment on the men and women's data. But within row
+of the figure, don't the distributions appear similar?  If that is so, would a 
+reviewer ask why we didn't pool the data for men and women?}
+It is also worth noting is that the fastest reaction time in the 2022 World
 Championships Finals and Semifinals both came from Devon Allen, but the 

### 2024-11-21 Manuscript/manuscript.tex
did some writing for gamlss results section

--- 2024-11-20 Manuscript/manuscript.tex
@@
 
+We considered many types of models from various continuous distribution families
+before ultimately deciding on the Generalized Gamma model desrcribed in
+Section~\ref{sec:gamlss}.  The residuals of the GG model were particularly
+strong, as evidenced by Figure~\ref{fig:Residuals}.  We care especially about
+the lower residuals as we are trying to estimate the probability of a low reaction
+time.  While there are two outliers present in the residual plot, we accept that
+this is relatively small in relation to the rest of the roughly 770 observations.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics{GGResiduals}
+  \caption{Residuals from the GG model}
+  \label{fig:Residuals}
+\end{figure}
+
 \begin{table}
@@
 Table~\ref{tab:Gamma_parameters}, both excluding and including 2022 data.
-Three models were considered, differing in their random effects: venue effect
-only; heat effect only, and both venue and heat effects. Reported are the AIC
-and estimated model parameters: the standard deviation
-of the heat effect $\sigma_h$, the standard deviation of the venue
-effect~$\sigma_v$, the intercept~$\alpha$, and the
-square root of the dispersion parameter $\phi$. With or without 2022,
-the GLMM with both venue and heat random effects was the best model as evidenced
-by the AIC compared to the other models with only one random
-effect.  When we exclude 2022 from the full model, we observe small but intriguing
-discrepancies in the models. The standard deviation of the venue effect, 
-$\sigma_v$, increases from 0.33 to 0.40 when 2022 is included, however, the
-estimates of other parameters
-remained relatively similar, such as the standard deviation of the heat effect
-and the dispersion parameter.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics{GGResiduals}
-  \caption{Residuals from the GG model}
-  \label{fig:Residuals}
-\end{figure}
+.  We report the AIC, the point estimates for $\mu$,
+$\sigma$, and $\mu$, as well as the standard deviations of the venue and heat
+effects: $\sigma_v$ and $\sigma_h$ respectively. When we exclude 2022 from the
+full model, we observe small but intriguing discrepancies in the models. The
+standard deviation of the venue effect, $\sigma_v$, increases from 0.0434 to
+0.0576 when 2022 is included. This suggests that by including 2022 in the model,
+there is more of a variation in venue effects.
 
@@
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics{GGDensity}
-  \caption{The densities of the GG distribution with random effects for heat
-  and venue (black), and without random effects (blue)}
-  \label{fig:DensityPlot}
-\end{figure}
-
-Figure~\ref{fig:VenueEffects} further illustrates the venue effect from 1999 to
-2023 based on the full GLMM both with and without 2022.  The temporal patterns
+The uniqueness of 2022 is further demonstrated by Figure~\ref{fig:VenueEffects},
+which shows the values of the venue random effects.  The temporal patterns
 of the two sets of effects look very similar, except that the positive effects
@@
 extreme as 2022.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics{GGDensity}
+  \caption{The densities of the GG distribution with random effects for heat
+  and venue (black), and without random effects (blue)}
+  \label{fig:DensityPlot}
+\end{figure}
+\of{Not sure what to write about the density plot or how to fit it into the rest
+of the section.}
 

### 2024-11-21 Manuscript/reply.tex

@@
 granting us the opportunity to revise this manuscript. We also want to
-express our appreciation to the reviewer for their valuable comments.
+express our appreciation to the reviewers for their valuable comments.
 
@@
 the addition of new data in both the GLMM and rank comparison sections in an
-attempt to expand the scope of the analysis and derive beter insights.
+attempt to expand the scope of the analysis and derive better insights.
 
@@
 
-This outcome is not unexpected, as athletes competing in longer sprint
+This difference in reaction times is not unexpected, however, as athletes 
+competing in longer sprint
 events, such as the 200-meter dash, may not prioritize reaction time
@@
 data; see updated results in Sections~???.
+\eds{let's revisit this again after objective 2 analyses are set}
 
@@
 
-The heat effect measures the speed-level of the heat.  We added some additional
+We apologize for the confusion. The heat effect actually measures the speed-level 
+of the heat, not the type of heat.  We added some additional
 explanations in the first paragraph of section 3.2 to make it clear what the
-venue and heat effect refer to. We did not want to call it a "race effect" as we
+venue and heat effect refer to. We did not want to call it a ``race effect" as we
 were worried that may be interpreted to mean the race of the athlete.  We did
@@
 \jy{It's the number of runners with 1 data point that determines how
-  hard the athlete effect can be estimated. How many runers had only 1
+  hard the athlete effect can be estimated. How many runners had only 1
   data point?}
@@
 start.
+\eds{let's revisit this again after objective 2 analyses are set}
 
@@
 foundation in formal statistical evidence.
+\eds{let's revisit this again after objective 2 analyses are set}
 
-\subsection*{Reviewer2}
+\subsection*{To Reviewer 2}
 
@@
 unusually fast.
+
+Please see our second response to Reviewer 1 for additional 
+comments relating to longer sprint events (200-meter or longer), 
+and why these were not also included in our analyses.
 
@@
 The nomenclature for the greek variables in Table 2 are defined in the GLMM
-subsection of the methods section (3.2)
+subsection of the methods section (3.2).
+\eds{let's revisit this again after objective 2 analyses are set}
 
@@
 
-Yes, the comments about the difficulty of data gathering are meant to demostrate
-that we worked hard to find as much data as we could and as a note to World
-Athletics (track and field governing body) that they need to do a better job
+Yes, the comments about the difficulty of data gathering are meant to demonstrate
+that we worked hard to find as much data as we could and, as a note to World
+Athletics (track and field governing body), that they need to do a better job
 at data standardization.  Currently the races on their website do not include
 reaction times, which meant that we had to look in alternative places to perform
-the analysis that we did.  The data we found from national level meets in 2022
+the analyses that we did.  The data we found from national level meets in 2022
 was difficult to come by as it required searching for every country's championship
@@
 quantitative data supporting our argument.
+\eds{let's revisit this again after objective 2 analyses are set}
 

### 2024-11-22 Code/gg.R
analysis script using GG family

@@ -1,0 +1,70 @@
+## get the right dataset
+library(tidyverse)
+times <- read.csv("../Data/GLMMData.csv")
+colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
+
+## Please reduce the length of your variable and object names;
+## they are too long and distracting.
+
+## dash/hurdle/man
+dhm <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime > 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles")
+
+dhm$Venue <- as.factor(dhm$Venue)
+dhm$Heat <- as.factor(dhm$Heat)
+
+
+## Fit various distributions available from gamlss
+## Plot standarized residuals q-q plot
+library("gamlss")
+library("qqconf")
+
+## three-parameter: generalized gamma
+gg0 <- gamlss(RxnTime ~ 1, data = dhm, family = GG)
+gg1 <- gamlss(RxnTime ~ random(Venue), data = dhm, family = GG)
+gg2 <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = GG)
+gg3 <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+## gg3a  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+gg3b <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+gg3c <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+gg3d <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
[…]

### 2024-11-23 Manuscript/reply.tex
comments on reply
went over referee 1 comments

@@
 
-We extend our gratitude to the Editor and Associate Editor for 
-granting us the opportunity to revise this manuscript. We also want to
-express our appreciation to the reviewers for their valuable comments.
+We sincerely thank the Editor and Associate Editor for the opportunity
+to revise our manuscript. The insightful comments and constructive
+feedback from the reviewers have greatly enhanced the quality of this
+work, and we are deeply grateful for their efforts.
 
-The manuscript has been revised accordingly with the most notable change being 
-the addition of new data in both the GLMM and rank comparison sections in an
-attempt to expand the scope of the analysis and derive better insights.
+The manuscript has been revised accordingly with the following most
+notable changes:
+\begin{enumerate}
+\item
+  The focus of the study has been clarified in the completely
+  rewritten introduction. The two objectives are 1) to investigate
+  whether the reaction times at the 2022 World Championships were
+  significantly different from the other competitions; and 2) to
+  assess the appropriateness of the 0.1-second threshold.
+\item
+  Additional data (reaction times from the 100-meter dash) have been
+  added to the analysis for both objectives to expand the scope of the
+  analysis and derive better insights.
+\item
+  In the assessment of the 0.1-second threshold, the Gamma GLMM has
+  been replaced with a generalized Gamma distributions in the flexible
+  framework of Generalized Additive Models for Location, Scale, and
+  Shape (GAMLSS) \citep{stasinopoulos2024generalized}, with random
+  effects of venue and heat in two of its parameters, respectively.
+  The new model provides an adequate fit to the data, as confirmed by
+  diagnostic checks.
+\end{enumerate}
 
@@
 
-Thank you for your feedback. We have carefully addressed the
-reviewers' comments and taken additional steps to streamline the
-manuscript, focusing on a clearer presentation of the key goals and
-findings from the outset. We appreciate your time and consideration in
-re-evaluating our submission.
+Thank you for your feedback. As detailed in the summary of notable changes, 
+we have thoroughly addressed the reviewers' comments and restructured the 
+manuscript to clearly identify its key goals and present the main findings 
+concisely. We appreciate your time and consideration in re-evaluating our 
+submission.
 
@@
 We thank the reviewer for the accurate summary of our contributions
-and these encouraging comments.
+and for the encouraging feedback. As part of our revisions, we have
+expanded the analysis to include reaction times from the 100-meter
+dash. Additionally, we replaced the gamma GLMM with a generalized
+Gamma distribution in the flexible framework of GAMLSS,  featuring
+random effects in its parameters, which provides an improved and
+adequate fit to the data.
 
@@
 
-Thank you for the valuable suggestions. We have extended our analysis
-by including data from both men's and women's 100-meter dash. This was
-justified by a GLMM fit to these data and the event type was not found
-to be significantly different in reaction time.
+Thank you for the valuable suggestions. In response, we expanded our
+analysis to include reaction time data from both men and women's
+100-meter dashes. A generalized gamma model fit to these data showed
+no significant differences in reaction times by event type.
 
-We also considered including additional sprinting events such as the
-200-meter dash; however, in a GLMM fit to an expanded dataset that
-additionally included 200-meter, the event type effect became
-significant. Subsequent pairwise comparisons using Tukey’s Honest
-Significant Difference (HSD) test revealed a p-value of 0.0021 when
-comparing the reaction times of the 200-meter dash and the 110-meter
-hurdles, and a p-value of 0.000017 when comparing the 100-meter dash
-and the 200-meter dash. These results indicate that reaction times in
-the 200-meter dash are significantly slower, suggesting that this
-event does not align well with our focus on extreme reaction times.
-Consequently, we decided to exclude the 200-meter data from our
-primary analysis.
+We also explored adding data from the 200-meter dash. However, a
+generalized gamma model fit to the expanded dataset that included the
+200-meter event revealed a significant effect of event type. Tukey’s
+Honest Significant Difference (HSD) test found a p-value of 0.0021 for
+the 200-meter dash compared to the 110-meter hurdles, and a p-value of
+0.000017 for the 100-meter dash compared to the 200-meter dash.
+These results indicate that reaction times in the 200-meter dash are
+significantly slower. Consequently, we excluded the 200-meter event
+from the main analysis, as its systematic differences in reaction
+times could bias conclusions about extreme reaction times and the
+appropriateness of the 0.1-second threshold.
 
-This difference in reaction times is not unexpected, however, as athletes 
-competing in longer sprint
-events, such as the 200-meter dash, may not prioritize reaction time
-as highly as those specializing in shorter sprints like the 100-meter
-dash. While it is common for amateur and collegiate athletes to
-compete across multiple sprint distances, elite athletes at the World
-Championship level may not exhibit equally strong performances in all
-sprint events. Our study aims to evaluate the probability of observing
-extreme reaction times, particularly in the context of the 0.1-second
-threshold. Including data from events with systematically slower
-reaction times would artificially inflate the extremity of a
-0.1-second reaction time, potentially biasing our conclusions.
+The observed differences in reaction times between events are not
+unexpected. Athletes competing in longer sprints, such as the
+200-meter dash, may place less emphasis on reaction time compared to
+those in shorter sprints like the 100-meter dash. Elite athletes at
+the World Championship level often specialize in specific distances,
+and performance across events may vary. Including data from events
+with systematically slower reaction times would inflate the extremity
+of a 0.1-second reaction time, potentially skewing the results of our
+study, which focuses on the likelihood of extreme reaction times
+relative to this threshold.
 
@@
 
-We apologize for the confusion. The heat effect actually measures the speed-level 
-of the heat, not the type of heat.  We added some additional
-explanations in the first paragraph of section 3.2 to make it clear what the
-venue and heat effect refer to. We did not want to call it a ``race effect" as we
-were worried that may be interpreted to mean the race of the athlete.  We did
-not think it would be informative to  add and effect for individual runner's
-average ability as many of observations in the data come from runners
-\jy{It's the number of runners with 1 data point that determines how
-  hard the athlete effect can be estimated. How many runners had only 1
-  data point?}
-with very few (1-3) data points, so taking an average of such a small sample may
-not be informative, especially if there are outliers or an athlete had a bad
-start.
+Thank you for this thoughtful suggestion. We apologize for any
+confusion regarding the heat effect. In our model, the heat effect
+captures the speed-level of the heat rather than the type of heat
+(e.g., preliminary vs. final). To clarify this, we have added
+additional explanations in the first paragraph of Section~3.2 to
+explicitly define the venue and heat effects. We avoided calling it a
+``race effect'' to prevent misinterpretation as referring to the
+athlete's race.
+
+We considered adding an effect for an individual runner’s average
+ability; however, many observations in our dataset come from athletes
+with only 1–-3 data points. Taking an average for such small samples
+may not be informative, especially in the presence of outliers or
+cases where an athlete had an atypical start. As such, we opted not to
+include this factor in the model.
+
+In the revision, we replaced the Gamma GLMM with a generalized Gamma
+distribution in the GAMLSS framework. The venue effect and heat effect
+were incorporated into two parameters controlling the scale of the
+distribution. This revised model provides an improved and adequate fit
+to the data compared to the original Gamma GLMM.
 \eds{let's revisit this again after objective 2 analyses are set}
@@
 
-We appreciate the feedback, which was echoed by the second
-reviewer. To address these concerns, we revised the introduction to
-better structure our arguments. We streamlined the literature review
-to focus only on studies directly relevant to our analysis and
-clarified the paper’s two main objectives: (1) identifying
-inconsistencies in the 2022 reaction times and (2) assessing whether
-the 0.1-second threshold is fair. We avoided speculating on sensor
-calibration or implicating Seiko due to a lack of direct evidence, and
-instead, we provide general recommendations for World Athletics while
-noting that the unusually fast reaction times in 2022 remain
-unexplained.
+We appreciate the feedback, which aligns with concerns raised by
+Reviewer~2. To address this, we completely rewrote the introduction to
+better structure our arguments and focus the problem statements. The
+literature review was streamlined to highlight studies directly
+relevant to our analysis. We clarified the paper’s two main
+objectives: (1) identifying inconsistencies in reaction times at the
+2022 World Championships, and (2) assessing the fairness of the
+0.1-second threshold.
+
+We avoided speculation about sensor calibration or implicating Seiko,
+as there is no direct evidence to support such claims. Instead, we
+provide general recommendations for World Athletics while noting that
+the unusually fast reaction times observed in 2022 remain unexplained.
 
@@
 
-Thank you for the positive feedback. We expanded both the rank
-comparison and the GLMM sections with additional data to strengthen our
-analysis. These enhancements support our findings of an anomaly in
-2022 and highlight that the 0.1-second reaction time threshold lacks a
-foundation in formal statistical evidence.
+Thank you for the positive feedback. We have expanded both the rank
+comparison and the GAMLSS model sections with additional data to
+strengthen our analysis. These enhancements support our findings of an
+anomaly in 2022 and highlight that the 0.1-second reaction time
+threshold lacks a foundation in formal statistical evidence.
 \eds{let's revisit this again after objective 2 analyses are set}

### 2024-11-23 Manuscript/manuscript.tex
comments on gg3b results
went over referee 2 comments

@@
 $\text{GG}(\mu, \sigma, \nu)$ has density function
-\[
+\begin{equation}
+  \label{eq:gg}
 f_Y(y \mid \mu, \sigma, \nu) = 
@@
 \exp\left(-z \theta\right),
-\]
+\end{equation}
 for $y > 0$, $\mu > 0$, $\sigma > 0$, and $\nu \neq 0$,
@@
 \of{Please see figure 5}
+\jy{This cannot be right. Black is always above blue. Both need to
+  integrate to one.}
 
@@
 this is relatively small in relation to the rest of the roughly 770 observations.
+
+\jy{Use aspect ratio 1 for QQ plot}
 
@@
 \of{table has been updated with gg3b results}
+\jy{Could you also test sensitivity for excluding the positive but
+  fault reaction times?}
 

### 2024-11-23 Manuscript/reply.tex
went over referee 2 comments

@@
 
-We have revised the Introduction and Discussion sections to clearly
-outline the goals of the study. Both reviewers noted issues with
-organization, so we streamlined the manuscript to emphasize the two
-main objectives: (1) identifying inconsistencies in the 2022 reaction
-times and (2) assessing whether the 0.1-second threshold is fair. We
-removed unrelated sections to keep the focus on these central
-points. Regarding the rank comparison and timing devices at the 2022
+Thank you for your valuable feedback. We have thoroughly revised the
+Introduction and Discussion sections to clearly articulate the study's
+objectives. As both reviewers noted organizational issues, we
+streamlined the manuscript to focus on its two main goals: (1)
+identifying inconsistencies in reaction times at the 2022 World
+Championships, and (2) assessing whether the 0.1-second threshold is
+statistically justified. Unrelated sections were removed to maintain a
+clear and focused narrative.
+
+Regarding the rank comparison and timing devices at the 2022
 Championships, we intentionally used cautious language, as there is no
-evidence to claim a miscalibration or fault in Seiko’s timing
-devices. To avoid potential legal risks, we noted the unusually fast
-reaction times in 2022 without making unsubstantiated claims. Instead,
-we allow readers to interpret the underlying cause of the anomaly
-based on the presented data.
+direct evidence to suggest miscalibration or faults in Seiko’s timing
+devices. To avoid potential legal risks, we emphasized the observed
+anomaly of unusually fast reaction times in 2022 without making
+unsubstantiated claims. Instead, we presented the data transparently,
+allowing readers to interpret the possible underlying causes of the
+anomaly.
 
@@
 If you are trying to answer question (1), you should include reaction times from
-every spring event, not just the hurdles. If you are trying to answer question
-2, why not do the following Monte Carlo test:
-
-1. Sample a single reaction time for each athlete in the final.
-2. Calculate the mean reaction time for the simulated final.
-3. Compute the probability of seeing the observed average reaction time deviate
-as much as it did in the 2022 Championships.
+every spring event, not just the hurdles. If you are trying to answer
+question~2, why not do the following Monte Carlo test:\\
+\indent 1. Sample a single reaction time for each athlete in the final.\\
+\indent 2. Calculate the mean reaction time for the simulated final.\\
+\indent 3. Compute the probability of seeing the observed average reaction time deviate
+as much as it did in the 2022 Championships.\\
 
@@
 differences between times.
-
 \end{comment}
 
-Thank you for the feedback. We are attempting to address question~(1)
-and have expanded our analysis to include reaction times from
-additional sprint events at the World Championships, including both
-men and women's 100-meter data. This broader dataset helps provide a
-more comprehensive assessment of whether the venue in 2022 was
-unusually fast.
+Thank you for the feedback. We clarify that our focus is question~(1): 
+evaluating whether the venue in 2022 was unusually fast. To address this, 
+we expanded our analysis to include reaction times from additional sprint 
+events at the World Championships, including both men’s and women’s 
+100-meter dashes. This broader dataset provides a more comprehensive 
+assessment of venue effects in 2022.
 
-Please see our second response to Reviewer 1 for additional 
-comments relating to longer sprint events (200-meter or longer), 
-and why these were not also included in our analyses.
+Your suggestion to include reaction times from more sprint events aligns 
+with feedback from Reviewer~1.  While the 100-meter dash reaction times 
+showed no significant differences from the 110-meter hurdles, we excluded 
+200-meter and longer dashes due to their significantly different reaction 
+times. These longer events represent distinct contexts unsuitable for this 
+analysis.
 
@@
 
-Thank you for raising these points. Regarding the exclusion of
+Thank you for raising these concerns. Regarding the exclusion of 
 disqualified runners, we chose not to include athletes with negative 
-reaction times, as these indicate an attempt to predict the starting
-gun rather than a true reaction. Negative reaction times do not align
-with the purpose of the GLMM analysis, which aims to evaluate the
-feasibility of the 0.1-second reaction threshold. Including negative
-times would introduce results where the reaction occurred before the
-gun fired, making them unsuitable for this analysis.
+reaction times, as these represent attempts to predict the starting gun 
+rather than true reactions. Negative reaction times do not align with 
+our objective of evaluating the feasibility of the 0.1-second threshold, 
+as they reflect pre-emptive behavior rather than genuine reaction
+times.
+
+The Gamma GLMM model has been replaced with a generalized Gamma 
+distribution within the GAMLSS framework, incorporating random effects 
+at the venue and heat levels in its distributional parameters. This 
+revised model provides improved and adequate fit to the data, as shown 
+by diagnostic plots in Figure~3.
 
@@
 \jy{Are these based on the expanded dataset or the original dataset?}
+\jy{These numbers need to be updated}
 slightly from 0.0029 to 0.0039, with similar parameter estimates for
@@
 observations in the dataset.
-
-Distribution: \jy{need to check R package gamlss.}
-
-\jy{Find these references and cite them:
-Sprint Start Regulation in Athletics: A Critical Review.
-Milloz M, Hayes K, Harrison AJ.
-Sports Med. 2021 Jan;51(1):21-31. doi: 10.1007/s40279-020-01350-4.
-PMID: 33125639 Review.
-Reaction time aspects of elite sprinters in athletic world championships.
-Tønnessen E, Haugen T, Shalfawi SA.
-J Strength Cond Res. 2013 Apr;27(4):885-92. doi: 10.1519/JSC.0b013e31826520c3.
-PMID: 22739331
-}
 
@@
 Thank you for this suggestion. Unfortunately, comprehensive data on 
-Devon Allen's reaction times is not readily available. We gathered as
-much data as we could find online, but there is insufficient
-information from lower-level meets, as reaction times are typically
-only recorded at major competitions (e.g., national and world
-championships). Given the limited data, we could not conduct a formal
-statistical analysis of his reaction times. However, we have added a
-brief discussion of this issue in paragraph~???] of the Discussion Section.
+Devon Allen’s reaction times is not readily available. While we gathered 
+as much data as possible from publicly available sources, reaction times 
+are typically recorded only at major competitions, such as national and 
+world championships. As a result, the dataset is too limited to support 
+a formal statistical analysis of his reaction times.
+To address this point, we have added a brief discussion in 
+paragraph~??? of the Discussion Section, noting the limitations in data 
+availability and the implications of this issue.
 
@@
 We added additional explanations to address this in the introductions section.
-\jy{Give paragraph number for specific locations}
+\jy{Owen: Which paragraph?}
 
@@
 
-The nomenclature for the greek variables in Table 2 are defined in the GLMM
-subsection of the methods section (3.2).
-\eds{let's revisit this again after objective 2 analyses are set}
+Thank you for pointing out the need for clearer intrepretation of
+Table~2.
+\jy{Owen: please add standard errors of the three intercept  to the table.}
+The table now contains fitted parameters of the generalized Gamma
+distribution GAMLSS as specified Equations (1)--(3). The table caption has been
+expanded to provide detailed explanations of the fitted parameters.
 
@@
 
-Yes, the comments about the difficulty of data gathering are meant to demonstrate
-that we worked hard to find as much data as we could and, as a note to World
-Athletics (track and field governing body), that they need to do a better job
-at data standardization.  Currently the races on their website do not include
-reaction times, which meant that we had to look in alternative places to perform
-the analyses that we did.  The data we found from national level meets in 2022
-was difficult to come by as it required searching for every country's championship
-results individually and when we did find it, it was often in the native language
-making it harder to understand and interpret.
+Yes, the comments about the difficulty of data gathering are included
+to highlight the effort required to compile a comprehensive
+dataset. This also serves as a note to World Athletics, emphasizing
+the need for better data standardization. Currently, race results on
+their website do not include reaction times, requiring us to search
+alternative sources. Gathering data from national-level meets in 2022
+involved searching each country’s championship results individually,
+often encountering language barriers that complicated interpretation.
 
@@
 
-We included this plot to show the venue effect for each year and to visually
-demonstrate how 2022 compares to other years.  We added additional comments
-about the probability of observing an extreme venue effect to provide
-quantitative data supporting our argument.
+Thank you for the feedback. The purpose of Figure~3 is to display the venue 
+effects for each year and visually illustrate how 2022 compares to other years. 
+To address your concern, we added additional comments quantifying the 
+probability of observing an extreme venue effect, which strengthens the 
+argument that 2022 is anomalous. If needed, we are open to revising the plot 
+further or integrating it into a single panel showing deltas, as suggested.
 \eds{let's revisit this again after objective 2 analyses are set}

### 2024-11-23 Code/gg.R
Added pool/dash calculation and the dq effect calculation

@@ -31,17 +31,18 @@ gg3 <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(V
 ## gg3a  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 gg3b <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 gg3c <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
-gg3d <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
-gg4 <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+#gg3d <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+#gg4 <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 
 
-AIC(gg1, gg2, gg3, gg3a, gg3b, gg3c, gg3d, gg4) # gg3b is the best
+AIC(gg1, gg2, gg3, gg3b, gg3c) # gg3b is the best
 qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1)) ## z-scores have known mean and sd
 qq_conf_plot(gg3b$sigma.coefSmo[[1]]$coef)
 qq_conf_plot(gg3b$mu.coefSmo[[1]]$coef)
 
 # Simulation Code: only works for gg3b
 simfit <- function(model, n = 10000000) {
+    set.seed(1)
     sd_Venue <- model$mu.coefSmo[[1]]$sigb
     sd_Venue_heat <- model$sigma.coefSmo[[1]]$sigb[1]
     
@@ -54,9 +55,9 @@ simfit <- function(model, n = 10000000) {
 
 samplefit <- simfit(gg3b)
 
[…]

### 2024-11-23 Archive/Plots.Rmd
Plot 5 mostly done, needs more work

@@
 ```
+
 # Figure 4
+Calculation of curve
 ```{r}
-
-pdf(file = "../Manuscript/GGResiduals.pdf", height = 4, width = 6)
-qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1))
-dev.off()
-```
-
-# Figure 5
-```{r}
-sim_time <- function(model, RxnTime) {
 simfit <- function(model, n = 10000000) {
-
+    set.seed(1)
     sd_Venue <- model$mu.coefSmo[[1]]$sigb
@@
 
-    x <- rGG(n, mu = mu, sigma = sigma, nu = nu)
-    x
+    rGG(n, mu = mu, sigma = sigma, nu = nu)
 }
-  x <- simfit(model)
-  plot(density(x), xlim = c(0, 0.3), main = "", xlab = "")
-  curve(dGG(x, mu = exp(model$mu.coefficients[1]),
-            sigma = exp(model$sigma.coefficients[1]),
-            nu = model$nu.coefficients[1]),0, 0.3, add = TRUE, col = "blue",
-        main = "", xlab = "")
+samplefit <- simfit(gg3b)
+samplefit0 <- samplefit[1:1000000]
+kernfit <- density(samplefit0)
+
+```
+
+Now we make the density graph
+```{r}
+make_graph <- function(n = 1000000){
+  plot(density(samplefit, n=n), xlim = c(0, 0.3), ylim = c(0, 25), xlab = "",
+     main = "")
+  curve(dGG(x, mu = exp(gg3b$mu.coefficients[1]),
+          sigma = exp(gg3b$sigma.coefficients[1]),
+          nu = gg3b$nu.coefficients[1]),0, 0.3, add = TRUE, col = "blue",
+      main = "", xlab = "", ylab = "")
 }
-pdf(file = "../Manuscript/GGDensity.pdf", height = 4, width = 6)
-sim_time(gg3b, 0.1)
-dev.off()
+density_plot <- make_graph(1000000)
 ```
+
+And combine it with the residuals
+```{r}
+library(gridExtra)
+resid_plot <- qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1))
+density_plot <- grid.arrange(resid_plot, density_plot, ncol = 2)
+ggsave(filename = "ResidualsDensityPlot", path = "../Manuscript",
+       plot=resid_density, height = 4, width = 8)
+```

### 2024-11-23 Code/gg.R
added code to plot the diagnosis

@@ -63,12 +63,17 @@ quantile(samplefit, c(.001, .0001))
 
 samplefit0 <- samplefit[1:1000000]
 
-hist(dhm$RxnTime, freq=FALSE, breaks=seq(.05, .40, by = .01))
+pdf("diagnosis.pdf", height=3.25, width = 7, pointsize = 9)
+par(mfrow = c(1, 2), mar = c(2.5, 2.5, .1, .1), mgp = c(1.5, 0.5, 0))
+hist(dhm$RxnTime, freq=FALSE, breaks=seq(.05, .40, by = .01),
+     xlab = "Reaction time", main = "")
 kernfit <- density(samplefit0)
-lines(kernfit$x, kernfit$y) # /1.434) ## density seems to have a bug for n too large
-curve(dGG(x, mu = exp(gg0$mu.coefficients[1]),
-          sigma = exp(gg0$sigma.coefficients[1]),
-          nu = gg0$nu.coefficients[1]), add = TRUE, col = "blue")
+lines(kernfit$x, kernfit$y) # /1.434) ## density() seems to have a bug for n too large
+## curve(dGG(x, mu = exp(gg0$mu.coefficients[1]),
+##           sigma = exp(gg0$sigma.coefficients[1]),
+##           nu = gg0$nu.coefficients[1]), add = TRUE, col = "blue")
+qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1)) ## z-scores have known mean and sd
+dev.off()
 
 # Should we pool dash and hurdles?
 dhm <- times |>

### 2024-11-23 Manuscript/manuscript.tex
half way through data section

@@
 
-There are two types of data that are explored throughout this paper. For the
-rank-based comparisons described in Section~\ref{sec:rank}, the data described in 
-Section~\ref{sec:databeyond} come from male 
-athletes who competed in the 110 meter hurdles and \eds{or?} 100 meter dash, and 
-female athletes who competed 100 meter hurdles and \eds{or?} 100 meter dash.
-Athletes must have competed in the 2022 World Championships and another competition 
-(2022 national championships, 2019 World Championships, or 2023 World Championships) 
-in order to be included in the rank-based comparisons. 
-For the GAMLSS analysis described in Section~\ref{sec:gamlss}, the data described 
-in Section~\ref{sec:dataworld} come
-from every world championships from 1999 to 2023 and includes reaction times from
-the 110 meter hurdles and 100 meter dash. This investigation started 
-shortly after the 2022 World Championships, but as the 2023 data has now become 
-available, we are pleased to include it in both of our analyses, and it did not
-significantly change our results \citep{WAData}. \eds{Should we remove this 
-last sentence, since now likely 2024 data is available?}
+This study focuses on reaction times from sprint events where reaction
+time plays a critical role. Specifically, we included data from men’s
+100-meter dashes and 110-meter hurdles, as well as women’s 100-meter
+dashes and 100-meter hurdles. Statistical tests showed no significant
+differences in reaction times between these events, supporting their
+inclusion in a unified analysis. However, we excluded data from
+200-meter dashes and longer events, as their reaction times were found
+to be significantly different. This difference is not unexpected, as
+reaction time is generally less critical in longer sprint events
+compared to shorter distances where explosive starts are paramount.
+
+Two datasets were used to address the study's objectives. To
+investigate whether reaction times at the 2022 World Championships
+were significantly different from other competitions (Objective 1), we
+used data from male athletes in the 110-meter hurdles and 100-meter
+dash, and female athletes in the 100-meter hurdles and 100-meter dash,
+provided they competed in the 2022 World Championships and at least
+one other competition (2022 national championships, 2019 World
+Championships, or 2023 World Championships). To assess the
+appropriateness of the 0.1-second reaction time threshold (Objective
+2), we used reaction times from the 110-meter hurdles and 100-meter
+dash across all World Championships held from 1999 to 2023. This
+investigation began shortly after the 2022 World Championships, and we
+are pleased to include data from 2023, which did not significantly
+alter our results \citep{WAData}.
 
@@
 
-The motivation for looking beyond World Championships data came from an 
-exploratory analysis examining how United States athletes reacted at the 
-2022 United States Track and Field Championships.
+The motivation for including data beyond the World Championships arose from 
+an exploratory analysis of how United States athletes performed at the 2022 
+United States Track and Field Championships, held from June 23--26, 2022, at 
+Hayward Field in Eugene, Oregon. Since the same venue hosted the 2022 World 
+Championships in August, this provided a unique opportunity to establish a 
+baseline for reaction times.
 
-\subsubsection{2022 National Competitions}\label{sec:datanational}
-From June 23-26, 2022, the United States held its Track and Field Championships 
-at Hayward Field in Eugene, Oregon to decide who to send to the World 
-Championships being held at the same venue in August. Thus, we can establish a 
-baseline for the four United States athletes who competed in at least one round 
-of both events: Trey Cunningham, Daniel Roberts, Grant Holloway, and Devon Allen.  
-Every athlete reacted faster in all the World Track and Field Championships 
-races compared to the United States Track and Field Championships races. The 
-dataset of the four United States athletes is too small to perform any 
-meaningful statistical analysis, so the data for the rank-based analyses was 
-expanded to include reaction 
-times from more athletes who competed at other national competitions (many of which were 
-held from May-July of 2022). This data was difficult to find, as the data is not
-centrally located but instead found on different national websites (often 
-recorded in the national language without translation).
+\subsubsection{2022 National Competitions}
+\label{sec:datanational}
 
-Each athlete was designated as a ``cluster'' and the stage of competition (national
-or international) was specified for each competition within the cluster (athlete).
-Thus, this is subunit level grouping where we assign the treatment, national or
-international, within each cluster. Cluster sizes range from three to six, with
-one cluster of size three, six clusters of size four, two clusters of size five,
-and two clusters of size six.
+To compare reaction times at the 2022 World Championships with those
+recorded at national competitions, we constructed a dataset of athletes 
+who competed in both settings. Initially, we examined data from four 
+United States athletes---Trey Cunningham, Daniel Roberts, Grant 
+Holloway, and Devon Allen---who participated in both the United States 
+Track and Field Championships and the World Championships. Each athlete 
+recorded faster reaction times in all World Championships races compared 
+to their performances at the national-level event. However, this small 
+sample size was insufficient for meaningful statistical analysis. To 
+expand the dataset, we included reaction times from additional national 
+competitions held between May and July 2022 across various countries. 
+Compiling this data was challenging, as reaction times were not centrally 
+located and often required searching country-specific websites, with many 
+results recorded in native languages.
+
+\jy{Owen, does this include both dash and hurdle? Could you do a
+  pooled analysis since there is no gender difference and event
+  differece? Need to summarize how many athletes are included in each
+  analysis.}
+
+The final dataset consisted of reaction times from athletes who competed 
+in both national and international competitions. Reaction times from 
+preliminary heats, semifinals, and finals were included to ensure 
+sufficient data for analysis. Excluding preliminary heats would have 
+significantly reduced the number of athletes and clusters, narrowing the 
+scope of the analysis. Each athlete can be considered as a cluster, with
+reaction times from the `treatment' group (2022 World Championships) and 
+the `control' group (national competition) within the cluster. Cluster 
+sizes ranged from three to six, with a median size of four. This
+setting forms the basis for rank-based comparison where the
+within-cluster dependence needs to be accounted for.
+
 \of{This explanation is going to need to be re-worked now that we have more data.
@@
 
-We chose to include reaction times from preliminary heats, in addition to 
-reaction times from semifinals and finals, in this section and in
-Section~\ref{sec:data2019}, as removing them would have resulted in a much
-narrower analysis with fewer athletes and clusters.
-
-\subsubsection{Other World Track and Field Championships}\label{sec:data2019}
+\subsubsection{2019 and 2023 World Championships}
+\label{sec:data2019}
 The 2019 and 2023 World Track and Field Championships data is included in the 
@@
 and six clusters of size five.
-\of{This explanation might need to change}
 

### 2024-11-23 Manuscript/reply.tex
half way through data section

@@
 data for both men and wemen were included, while 200-meter dash data
-were not included in our analysis in Section~???, paragraph~???.
+were not included in our analysis in the opening paragraph of Section~2.
 The subsequent analyses in the manuscript have been updated with the
-expanded dataset that included both men and women's 100-meter dash
-data; see updated results in Sections~???.
+expanded datasets.
 \eds{let's revisit this again after objective 2 analyses are set}

### 2024-11-23 Manuscript/manuscript.tex
smoothed data section

@@
 
+\jy{Owen, are there names from 100-m dash?}
 To compare reaction times at the 2022 World Championships with those
@@
 
-\jy{Owen, does this include both dash and hurdle? Could you do a
-  pooled analysis since there is no gender difference and event
-  differece? Need to summarize how many athletes are included in each
-  analysis.}
-
 The final dataset consisted of reaction times from athletes who competed 
@@
 sufficient data for analysis. Excluding preliminary heats would have 
-significantly reduced the number of athletes and clusters, narrowing the 
-scope of the analysis. Each athlete can be considered as a cluster, with
-reaction times from the `treatment' group (2022 World Championships) and 
+significantly reduced the number of athletes and clusters. Each
+athlete was considered as a cluster, with observed reaction times
+from both the `treatment' group (2022 World Championships) and
 the `control' group (national competition) within the cluster. Cluster 
-sizes ranged from three to six, with a median size of four. This
-setting forms the basis for rank-based comparison where the
-within-cluster dependence needs to be accounted for.
+sizes ranged from three to six, with a median size of four. Because gender 
+is known to influence reaction times \citep{lipps2011implications}, we
+prepared data for men and women separately, resulting in 80 reaction
+times from 17 athletes for each
+gender. This setup forms the basis for the rank-based comparison for
+clustered data with subunit-level grouping, which accounts for the
+within-athlete dependence inherent in this data structure. The top
+panel of Figure!~\ref{fig:RankScatterplots} shows the reaction times
+of those who competed both at national competitions and the 2022 World
+Championships.
 
-\of{This explanation is going to need to be re-worked now that we have more data.
-Should we make a table of some sort? It's going to be very wordy without one}
-\eds{We are only doing two analyses with this data, right, one for women 
-and one for men?  So perhaps only two sentences needed here? But then 4 sentences
-in the next subsection?  Maybe you can just report ranges of cluster sizes and 
-median cluster size?}
-
-\subsubsection{2019 and 2023 World Championships}
-\label{sec:data2019}
-The 2019 and 2023 World Track and Field Championships data is included in the 
-larger dataset described in Section~\ref{sec:dataworld}, but we can also use it
-similarly to how we used national competitions in Section~\ref{sec:datanational}. 
-We can compare the reaction times of athletes who participated in the 2022 World 
-Championships with their times at another World Championships: 2019 and/or 2023.
-
-While it is theoretically possible that athletes got faster 
-between 2019 and 2022 or 2022 and 2023, it is important to 
-remember that these are some of the fastest sprinters on the planet, and thus
-improving on elite times is not easy. Once again we repeat a similar procedure
-as in Section~\ref{sec:datanational}; we assign each athlete to be a cluster and
-within each cluster label times to be from either the 2019, 2022, or 2023 World
-Championships. When comparing 2019 and 2022 we look at data from 15
-athletes, with two clusters of size two, three clusters of size three, five 
-clusters of size four, two clusters of size five, and three clusters of size six.
-When comparing 2022 and 2023 we look at data from 22 athletes, with seven 
-clusters of size two, three clusters of size three, six clusters of size four,
-and six clusters of size five.
+\jy{Could you do a pooled analysis since there is no gender difference
+  and event differece in the supplement?}
 
@@
   \centering
-  \includegraphics{RankScatterPlots}
+  \includegraphics[width=\textwidth]{RankScatterPlots}
   \caption{Reaction times for athletes who competed at the 2022 World Championships
@@
 
-Figure~\ref{fig:RankScatterplots} shows the data that will be analyzed
-with the rank-based methods (described in Section~\ref{sec:rank} with results in 
-Section~\ref{subsec:Results_Rank}). In each of the six graphs, the reaction 
-times for athletes who competed at the 2022 World Championships
-are compared to their reaction times at another at which they competed. 
-It is worth noting that for each row of the figure, the athletes compromising 
-the 2022 data is different.  This is because the set of athletes who competed at 
-both the 2019 and 2022 World Championships is not the same as the set of athletes 
-who competed at both the 2022 and 2023 World Championships. %This data was only for athletes who competed
-%in at least one race at each championship, however it is possible for athletes
-%to have competed in one race in 2019 and two races in 2022. 
-Athletes must have competed in at least one race at each championship to be 
-included.  The 2022 World
-Championship heats and semifinals both had on average lower reaction times than
-their respective other (2019, 2023 or 2022 national) races, and the reaction times 
-from the 2022 World Championship finals were substantially faster. 
-\eds{I would consider removing or rephrasing the last part of the previous 
-sentence. With the women's data, it is not as clear the 2022F is lower than 2022S.}
+\subsubsection{2019 and 2023 World Championships}
+\label{sec:data2019}
+
+To compare reaction times at the 2022 World Championships with those 
+recorded in the 2019 and 2023 World Championships, we prepared datasets 
+including reaction times from athletes who competed in the 2022 World 
+Championships and at least one of the 2019 or 2023 World Championships. 
+Specifically, we identified athletes who participated in both the 2019 
+and 2022 World Championships, as well as those who participated in both 
+the 2022 and 2023 World Championships. In the 2019-2022 comparison, 
+reaction times from 2022 were treated as the `treatment' group, with 
+2019 serving as the `control' group. Similarly, in the 2022-2023 
+comparison, reaction times from 2022 were treated as the `treatment' 
+group, with 2023 serving as the `control' group. This structure allowed 
+us to prepare datasets suitable for examining reaction times of athletes 
+who competed across multiple World Championships.
+
+Each athlete was treated as a single cluster, containing their reaction 
+times from different World Championships. The dataset for the 2019-2022 
+comparison contained 134 reaction times from 34 male athletes and 124 
+reaction times from 31 female athletes. The dataset for the 2022-2023 
+comparison contained 161 reaction times from 45 male athletes and 182 
+reaction times from 47 female athletes. While it is theoretically 
+possible that athletes improved their reaction times between 2019 and 
+2022 or between 2022 and 2023, such improvements are highly unlikely for 
+elite sprinters, as they already operate near the limits of human 
+performance. Consequently, consistent improvements observed in 2022 
+would suggest systematic differences rather than natural variability.
+
+Figure~\ref{fig:RankScatterplots} shows the reaction times of athletes 
+who competed in both the 2019 and 2022 World Championships (middle 
+panel) and those who competed in both the 2022 and 2023 World 
+Championships (lower panel). The athletes included in the 2022 data 
+differ between these two comparisons, as the set of athletes who 
+competed in both 2019 and 2022 is not the same as the set who competed 
+in both 2022 and 2023. To be included, athletes must have competed in 
+at least one race at each championship. Notably, Devon Allen recorded 
+the fastest reaction times in both the Finals and Semifinals of the 
+2022 World Championships, but his disqualification was determined by a 
+difference of just 0.002 seconds, with reaction times of 0.101 and 
+0.099 seconds, respectively. This highlights the critical role of 
+reaction time precision in elite-level competition.
+
 \eds{I also think we need to comment on the men and women's data. But within row
@@
 reviewer ask why we didn't pool the data for men and women?}
-It is also worth noting is that the fastest reaction time in the 2022 World
-Championships Finals and Semifinals both came from Devon Allen, but the 
-difference between 0.101 and 0.099
-seconds was the difference in what caused him to be disqualified.
 
@@
   \centering
-  \includegraphics{Boxplot}
+  \includegraphics[width=\textwidth]{Boxplot}
   \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle

### 2024-11-23 Manuscript/manuscript.tex
smoothed world championship data

@@
 effect of the competition year while controlling for individual 
-performance. The second objective is to assess the 
+performance. The clustered nature of the data, with the goal of
+assessing differences across events within-athletes, were analyzed
+with rank-based comparison approach for clustered data
+\citep{datta2005rank}. The second objective is to assess the 
 appropriateness of the 0.1-second threshold by modeling reaction 
-times from World Championships held from 1999 onward. We use a gamma 
-linear mixed-effects model with random effects for both venue and heat. 
-\eds{edit previous as needed for new GAMLSS analysis}
+times from World Championships held from 1999 onward. We use a
+generalized gamma distribution with random effects for both venue and
+heat in the framework of the generalized additive model for location,
+scale, and shape (GAMLSS) \citep{stasinopoulos2024generalized}.
 This model estimates the probability of reaction times falling below 
@@
 describes the data collection and datasets used for the two
-objectives. Section~\ref{sec:rank} outlines the methods for
-objective~1, analyzing differences in reaction times across
-competitions, including the were collected. 
-\eds{end of previous sentence does not make sense; some text may have been 
-mistakenly deleted?}
-Section~\ref{sec:gamlss}
-details the methods for objective~2, the generalized additive model
-for location, scale, and shape (GAMLSS) to evaluate the appropriatenes
+objectives. Section~\ref{sec:rank} outlines the rank-based comparison
+methods for objective~1, rank-based comparison for clustered data with
+subunit-level gruping to analyze differences in reaction times
+across different competitions. Section~\ref{sec:gamlss} details the
+methods for objective~2, the GAMLSS to evaluate the appropriatenes
 of the 0.1-second threshold. Section~\ref{sec:Results} presents the
@@
 paper’s impact, limitations, and potential for future research.
-All data and code for our analysis 
-are provided in the supplementary materials.
+All data and code for our analysis are provided in the supplementary
+materials.
 
-\section{Data} \label{sec:Data}
+\section{Data}
+\label{sec:Data}
 
@@
 
-The motivation for including data beyond the World Championships arose from 
+For assessing the abnormality of the 2022 World Championship, 
+the motivation for including data beyond the World Championships arose from 
 an exploratory analysis of how United States athletes performed at the 2022 
@@
 sizes ranged from three to six, with a median size of four. Because gender 
-is known to influence reaction times \citep{lipps2011implications}, we
-prepared data for men and women separately, resulting in 80 reaction
-times from 17 athletes for each
+is known to influence reaction times \citep{babicc2009reaction,
+  lipps2011implications}, we prepared data for men and women
+separately, resulting in 80 reaction times from 17 athletes for each
 gender. This setup forms the basis for the rank-based comparison for
@@
 reviewer ask why we didn't pool the data for men and women?}
+\jy{Let's put the pooled analysis in the supplement.}
 
-\subsection{World Championships 1999--2023}\label{sec:dataworld}
+\subsection{Data from World Championships 1999--2023}
+\label{sec:dataworld}
 
-Data was taken from the World Athletics 
-and covers the men's 110 meter hurdles from 1999 to 2023.
-We focus on the reaction times during the 
-semifinal and final heats only, as reaction times from preliminary heats are 
-often not as fast as those in later heats \citep[e.g.,][]{zhang2021correlation}. 
-For analysis purposes, we will pool reaction times of men's final and 
-men's semifinal heats together which increases the sample size. This is 
-advantageous because in some years there are very few finals observations, such
-as in 2022, when there were only five data points due to two disqualifications 
-and one athlete who did not compete. This is the data that we will reference
-throughout the rest of the paper unless otherwise noted. We also consider data 
-that excludes 2022 to see how our analysis differs based on the exclusion of one
-particular year of interest.
+The data for evaluating the appropriateness of the 0.1-second
+threshold was obtained from World Athletics and covers the men's
+110-meter hurdles and 100-meter dashes from 1999 to 2023. Due to
+possible gender differences \citep{babicc2009reaction,
+  lipps2011implications}, data for women's 100-meter hurdles and
+100-meter dashes and their analyses were relegated to the
+supplementary material. We focus on the reaction times recorded during
+semifinal and final heats only, as reaction times from preliminary
+heats are often not as fast as those in later heats
+\citep[e.g.,][]{collet1999strategic, tonnessen2013reaction,
+  brosnan2017effects, zhang2021correlation}. For analysis purposes, we
+pooled reaction times from semifinal and final heats to increase sample 
+size, which is particularly important for years with limited final heat 
+observations. For example, in 2022, only five data points were available 
+from the final heat due to two disqualifications and one athlete not 
+competing. Unless otherwise noted, this pooled dataset forms the basis 
+for our analysis throughout the paper. Additionally, we consider datasets 
+that exclude 2022 to assess how our findings might differ when excluding 
+this year of interest.
 
-The data under consideration is summarized using side-by-side boxplots in 
-Figure~\ref{fig:Boxplot}. It is clear that the reaction times in the 2022
-boxplot are lower (indicating faster reaction times) than in many of the other
-years. The median reaction time in 2022 was 0.129 seconds,  compared to 0.156
-seconds, which is what \citet{brosnan2017effects} found when looking at data
-from 1999 to 2014. We can also see from Figure~\ref{fig:Boxplot} that the 
-reaction times clearly differ from year to year. The venue of the World 
-Championships changes each year, so weather or climate related factors such as 
-humidity, precipitation, elevation, may indeed play a role, but additionally,
-both the technology and false start penalization rules have changed over the 
-study period.
+\jy{have the numbers in this paragraph been updated with the
+  additional 100-dash data?}
+The data is summarized in Figure~\ref{fig:Boxplot}, which presents a
+sequence of boxplots of reaction times from 1999 to 2023. It is evident 
+that reaction times in 2022 were notably faster, with a median reaction 
+time of 0.129 seconds compared to the 0.156 seconds observed in earlier 
+studies, such as \citet{brosnan2017effects} for data spanning 1999 to 2014. 
+Figure~\ref{fig:Boxplot} also highlights year-to-year variability in 
+reaction times, likely influenced by changes in the championship venue 
+and environmental conditions such as humidity, precipitation, and 
+elevation. Furthermore, advancements in technology and alterations to 
+false start rules during the study period may have played a role in these 
+variations \citep{willwacher2013novel}.
 
@@
 \end{figure}
-\of{Do We want 2 boxplot figures: 1 hurdles, 1 dash or combine them}
 
-From 2007 to 2009, the IAAF (the former name for World Athletics, and the 
-governing body for the World Track and Field Championships) instituted a rule 
-change that allowed one warning false start before a sprinter was disqualified 
-\citep{iaaf2009falsestart}. There were 18 male and 7 female false starts at the 
-2007 World Championships, and 18 male and 7 female false starts at the 2009 World 
-Championships. Starting in 2011, the rule was scrapped, and the old policy which
-automatically disqualifies runners who false start was reinstated. This was 
-desirable for World Athletics as false starts can make already lengthy track 
-meets more tedious, both for athletes and viewers watching the television 
-broadcast. By returning to the harsher policy and cracking down on false starts,
-World Athletics had reduced men's false starts by two thirds in 2011 (6 male and
-4 female disqualifications) \citep{iaaf2009falsestart}. \citet{haugen2013effect}
-examined the effect of different false start rules that IAAF has imposed from 
-1997 to 2009 and found statistically significant improvement in reaction times 
-under more lenient rules.
+Between 2007 and 2009, World Athletics (formerly the IAAF) allowed one 
+false start warning before disqualifying a sprinter \citep{iaaf2009falsestart}. 
+This lenient rule led to 18 male and 7 female false starts at both the 2007 
+and 2009 World Championships. In 2011, this rule was replaced with the stricter 
+policy of automatic disqualification for false starts, aimed at reducing the 
+delays caused by repeated warnings. This change reduced men’s false starts 
+by two-thirds in 2011, with only six male and four female disqualifications 
+\citep{iaaf2009falsestart}. \citet{haugen2013effect} demonstrated that more 
+lenient false start rules significantly improved reaction times during the 
+1997–2009 period, suggesting that rule changes over the study period may 
+have contributed to variations in reaction times across years.
 

### 2024-11-23 Manuscript/manuscript.tex
smoothed results for rank-based analysis

@@
 
-With our small sample size, the asymptotic normal distribution may not be
-reliable, so we also use 1~million random permutations to simulate the null 
-distribution of the test statistic.
+When the sample size is small, the asymptotic normal distribution may
+not be reliable, so we also use 1~million random permutations to
+simulate the null distribution of the test statistic.
 This method is available from the \texttt{clusWilcox.test()} function
@@
 
-The rank-based methods described in Section~\ref{sec:rank} are used
-to determine if the reaction times for athletes who had competed at other 
-championships differed significantly from the 2022 World Championships. This is
-a two group comparison performed three times: comparing 2019 and 2022 World
-Championships, comparing 2022 and 2023 World Championships, and comparing 2022 
-national championships to the 2022 World Championships.
+The rank-based methods described in Section~\ref{sec:rank} were used to 
+compare reaction times between the 2022 World Championships and other 
+competitions in which the same athletes participated. These comparisons 
+were conducted separately for men and women, resulting in six total 
+comparisons: reaction times from the 2022 national-level championships 
+versus the 2022 World Championships for men and women, reaction times 
+from the 2019 versus 2022 World Championships for men and women, and 
+reaction times from the 2022 versus 2023 World Championships for men and 
+women.
 
@@
   \centering
-  \caption{2019 vs 2022 compares data from the same athletes who competed at the
-  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
-  from the same athletes who competed at the 2022 and 2023 World Track and Field
-  Championships. National vs International compares data from the same athletes
-  who competed at 2022 national-level championships and the 2022 World Track and
-  Field Championships.}
+  \caption{Summaries of the results of comparisons between 
+    reaction times from different competitions for the same athletes. 
+    2022 Nat. vs Inter. compares reaction times from 2022 national-level 
+    championships and the 2022 World Track and Field Championships. 2019 
+    vs 2022 compares reaction times from the 2019 and 2022 World Track and 
+    Field Championships. 2022 vs 2023 compares reaction times from the 
+    2022 and 2023 World Track and Field Championships.}
   \begin{tabular}{c c c c c} 
@@
    \midrule
+   2022 Nat. vs Inter. Men & $1.0 \cdot 10^{-6}$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
+   2022 Nat. vs Inter. Women & $1.0 \cdot 10^{-6}$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\[1ex]
    2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
-   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\
-   2022 Nat. vs Inter. Men & $<0.001$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
-   2022 Nat. vs Inter. Women & $<0.001$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\
-   2022 vs 2023 Men & $<0.001$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
-   2022 vs 2023 Female & $<0.001$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
+   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\[1ex]
+   2022 vs 2023 Men & $1.0 \cdot 10^{-6}$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
+   2022 vs 2023 Female & $1.0 \cdot 10^{-6}$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
    \bottomrule
@@
 \end{table}
-\of{Table might still be too wide.  I used <0.001 becuase that is what we
-had currently been using for any values from the permutation test where the value
-was < 2.2e-16 in R.  Professor Yan will know best, but are the sample sizes now
-sufficiently large enough that the permutation test is not needed? If I recall
-correctlty that was a correction for small sample sizes.}
 
-Table~\ref{tab:Clusrankresults} shows the results generated from both the 
-permutation and asymptotic rank-based tests for the six comparisons, 
-as discussed previously in Section~\ref{sec:rank}. The national versus international
-comparisons produced lower p-values than the 2019 versus 2022 and 2022 versus 2023
-comparisons. The highly significant permutation test results
-provides substantial evidence that both men and women who competed at the
-2022 World Track and Field Championships were on average, significantly faster
-at reacting than they were just a few months prior. When looked at in context
-with other tests, these results show that athletes competing at the 2022 World
-Track and Field Championships had faster reaction times than when they had
-competed at other competitions.
+Table~\ref{tab:Clusrankresults} presents the results from both permutation 
+and asymptotic rank-based tests for these six comparisons. The tests reveal 
+that the national versus international comparisons yielded the lowest 
+p-values, indicating stronger evidence of faster reaction times at the 
+2022 World Championships compared to the 2022 national competitions. For 
+both men and women, permutation test results were highly significant, 
+providing substantial evidence that athletes competing at the 2022 World 
+Championships had significantly faster reaction times than they did just 
+a few months earlier at national-level events. The comparisons between the 
+2019 and 2022 World Championships and between the 2022 and 2023 World 
+Championships also produced significant results. Collectively, 
+these results suggest that athletes consistently achieved faster reaction 
+times at the 2022 World Championships compared to other competitions. This 
+finding supports the hypothesis that the 2022 World Championships presented 
+conditions conducive to faster reaction times, potentially due to systematic 
+or environmental factors.
 
@@
 
-\jy{Owen, please prepare the following:
-1) A table of the model parameter estimates (both point estimate and
-standard error). This is similar to the current table, with and
-without 2022 data. Let's focus on model gg3b and leave the model
-comparison results in the supplement.
-2) Figure: random effects at the venue level with/without 2022
-data. Similar to the current figure.
-3) Diagnostic plots: qq-plot for the gg3b model. QQ-plot for the
-normal random effects could be placed in supplement.
-4) Table: assessing the 0.1 threshold, similar to the current table.
-I'll take care of the writing when you get the results.}
-\of{1) Is resolved in table 2, 2) is resolved in figure 4, 3) is resolved in
-figure 3. 4) is resolved in tables 3 and 4.}
-
-\jy{Owen: an additional figure of the GG density without random
-  effects (use \texttt{curve(dGG(x, mu = , sigma = , nu =  ), 0, 0.03)}) and
-  with random effects (through \texttt{density()})}
-\of{Please see figure 5}
-\jy{This cannot be right. Black is always above blue. Both need to
-  integrate to one.}
-
 We considered many types of models from various continuous distribution families
-before ultimately deciding on the Generalized Gamma model desrcribed in
+before ultimately deciding on the generalized Gamma model desrcribed in
 Section~\ref{sec:gamlss}.  The residuals of the GG model were particularly
@@
 
-\jy{Use aspect ratio 1 for QQ plot}
-
 \begin{figure}[tbp]
   \centering
-  \includegraphics{GGResiduals}
-  \caption{Residuals from the GG model}
-  \label{fig:Residuals}
+  \includegraphics{diagnosis.pdf}
+  \caption{Diagnosis of the fitted generalized Gamma distribution with
+    random effects in model parameters: kernerl density of 1 million
+    observations drawn from the fitted model overlaid with the
+    histogram of the observed reaction times (left); Q-Q plot of the
+    z-score of the quantile residuals from the fitted model.}
+  \label{fig:diagnosis}
 \end{figure}
@@
 \end{table}
+
 \of{Created these with the function I wrote model-name at bottom of gamlss.R. Do
   any other parametes need to be added?}
+\jy{Remove AIC because there is no comparison. Need their standard errors.}
 

### 2024-11-24 Manuscript/manuscript.tex
smoothed results section

@@
 \begin{align}
-\log(\mu_{ijk}) &= \beta_0 + v_i ,\\
-\log(\sigma_{ijk}) &= \gamma_0 + h_{i/j},
+\log(\mu_{ijk}) &= \beta_0 + v_i , \label{eq:mu}\\
+\log(\sigma_{ijk}) &= \gamma_0 + h_{i/j} , \label{eq:sigma}
 \end{align}
@@
   \centering
-  \caption{Summaries of the results of comparisons between 
+  \caption{P-values of comparisons between 
     reaction times from different competitions for the same athletes. 
@@
    \toprule
-   Comparison & Permutation & Asymptotic & Athletes & Number observations  \\ 
+   Comparison & Permutation & Asymptotic & \# of athletes & \# of observations  \\ 
    \midrule
@@
 
-We considered many types of models from various continuous distribution families
-before ultimately deciding on the generalized Gamma model desrcribed in
-Section~\ref{sec:gamlss}.  The residuals of the GG model were particularly
-strong, as evidenced by Figure~\ref{fig:Residuals}.  We care especially about
-the lower residuals as we are trying to estimate the probability of a low reaction
-time.  While there are two outliers present in the residual plot, we accept that
-this is relatively small in relation to the rest of the roughly 770 observations.
+\jy{Please add standard errors, which could be from \texttt{summary(gg3b)}}
+
+\jy{Are these numbers correct? I modified with the gg3b in gg.R}
+
+\begin{table}
+  \centering
+  \caption{Estimated fixed-effect parameters with standard errors in
+    parentheses and estimated variance of the random effects from the
+    fitted generalized Gamma distribution with venue level random
+    effects in $\mu$ and heat level random effects in $\sigma$ in
+    Models~\eqref{eq:gg}--\eqref{eq:sigma}.}
+  \label{tab:ggfit}
+  \begin{tabular}{c c c c c c}
+    \toprule
+    Data set & $\beta_0$ & $\gamma_0$ & $\nu$ & $\sigma_v$ & $\sigma_h$ \\
+    \midrule
+    Excluding 2022 & $-$1.910 (0.005)  & $-$2.200 (0.025) & $-$1.177 (0.442) & 0.043 & 0.326 \\
+    Including 2022 & $-$1.910 & $-$2.200 & $-$1.178 & 0.058 & 0.320 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+The fitted parameters of the generalized Gamma distribution int he
+GAMLSS framework in Equations~\eqref{eq:gg}--\eqref{eq:sigma} are
+summarized in Table~\ref{tab:ggfit}. Results obtained from both
+excluding and including 2022 data are reported. The fixed-effect
+parameters include $\beta_0$, $\gamma_0$, and $\nu$, corresponding to
+the intercept of the log-location, log-scale, and shape of the GG
+distribution, respectively. Random effects account for variability at
+the venue level $\sigma_v$ on the log-scale of the $\mu$ parameter
+and at the heat level $\sigma_h$ on the log-scale of the $\sigma$
+parameter in the density in Equation~\eqref{eq:gg}. The variance 
+of the venue-level random effect is smaller than the heat-level random
+effect variance, suggesting that heat-level variability in the scale
+parameter is substantial, though on the dispersion parameter. When the
+2022 data is included, the fixed-effect estimates remain stable, 
+indicating robustness, but the venue-level random-effect variance
+increases to 0.0576, reflecting additional variability in the location
+parameter, while the heat-level random-effect variance remains stable.
+decreases slightly to 0.3198. These results
+highlight that reaction times are influenced by both venue and heat-level 
+factors, with heat-level variability playing a more prominent role in the scale 
+parameter, and that the inclusion of 2022 introduces greater venue-level 
+variability, likely due to systematic differences in reaction times that year.
 
@@
   \centering
-  \includegraphics{diagnosis.pdf}
+  \includegraphics[width=\textwidth]{diagnosis.pdf}
   \caption{Diagnosis of the fitted generalized Gamma distribution with
@@
     histogram of the observed reaction times (left); Q-Q plot of the
-    z-score of the quantile residuals from the fitted model.}
+    normal z-score of the quantile residuals from the fitted model.}
   \label{fig:diagnosis}
@@
 
-\begin{table}
-  \centering
-  \caption{AIC and estimated parameters for the gg3b model fit}
-  \label{tab:Gamma_parameters}
-  \begin{tabular}{c c c c c c c}
-    \toprule
-    Data set & AIC & $\mu$ & $\sigma$ & $\nu$ & $\sigma_v$ & $\sigma_h$ \\
-    \midrule
-    Excluding 2022 & -3660 & -1.900 & -2.192 & -1.315 & 0.0434 & 0.3259 \\
-    Including 2022 & -4103 & -1.910 & -2.200 & -1.178 & 0.0576 & 0.3198 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+Figure~\ref{fig:diagnosis} presents diagnostic checks for the fitted 
+generalized Gamma distribution model with random effects. The left 
+panel compares the kernel density estimate of one million simulated reaction 
+times from the fitted model to the histogram of the observed reaction times. 
+The close alignment between the density curve and the histogram suggests that 
+the fitted GG model adequately captures the overall distribution of the 
+reaction times. The right panel shows a Q-Q plot of the z-scores of the 
+quantile residuals from the fitted model. The points lie approximately along 
+the 45-degree reference line, indicating that the residuals are consistent 
+with the standard normal distribution, supporting the adequacy of the model 
+fit. These diagnostics collectively demonstrate that the fitted model provides 
+a reasonable representation of the observed reaction time data.
 
-\of{Created these with the function I wrote model-name at bottom of gamlss.R. Do
-  any other parametes need to be added?}
-\jy{Remove AIC because there is no comparison. Need their standard errors.}
+% \begin{figure}[tbp]
+%   \centering
+%   \includegraphics{ComparisonOfVenueEffects}
+%   \caption{The venue effects from 1999 to 2023 estimated from the GLMM for the
+%     data including 2022 (top) and excluding 2022 (bottom).}
+%   \label{fig:VenueEffects}
+% \end{figure}
 
-The estimated parameters of the GLMM model under the log link are shown in 
-Table~\ref{tab:Gamma_parameters}, both excluding and including 2022 data.
-.  We report the AIC, the point estimates for $\mu$,
-$\sigma$, and $\mu$, as well as the standard deviations of the venue and heat
-effects: $\sigma_v$ and $\sigma_h$ respectively. When we exclude 2022 from the
-full model, we observe small but intriguing discrepancies in the models. The
-standard deviation of the venue effect, $\sigma_v$, increases from 0.0434 to
-0.0576 when 2022 is included. This suggests that by including 2022 in the model,
-there is more of a variation in venue effects.
+% The uniqueness of 2022 is further demonstrated by Figure~\ref{fig:VenueEffects},
+% which shows the values of the venue random effects.  The temporal patterns
+% of the two sets of effects look very similar, except that the positive effects
+% estimated with 2022 included have higher magnitude than those estimated without
+% 2022, which compensates the negative effect of the 2022 with a large magnitude.
+% Both plots agree to some extent with the pattern of the boxplots in
+% Figure~\ref{fig:Boxplot}. The mismatch is due to the heat effect, since the venue
+% effect only captures a smaller part of the variation than the heat effect as
+% evident from their standard deviations. It is possible to calculate the extremity
+% of the 2022 heat effect as we know that the mean of the heat effects is $0$,
+% the standard deviation is $0.42$ as given by Table~\ref{tab:Gamma_parameters},
+% and the estimate of the 2022 heat effect is $-0.93$. We find the tail probability
+% to be $0.0128$ indicating there is a low chance of observing a venue effect as
+% extreme as 2022.
 
-\begin{figure}[tbp]
-  \centering
-  \includegraphics{ComparisonOfVenueEffects}
-  \caption{The venue effects from 1999 to 2023 estimated from the GLMM for the
-    data including 2022 (top) and excluding 2022 (bottom).}
-  \label{fig:VenueEffects}
-\end{figure}
-
-The uniqueness of 2022 is further demonstrated by Figure~\ref{fig:VenueEffects},
-which shows the values of the venue random effects.  The temporal patterns
-of the two sets of effects look very similar, except that the positive effects
-estimated with 2022 included have higher magnitude than those estimated without
-2022, which compensates the negative effect of the 2022 with a large magnitude.
-Both plots agree to some extent with the pattern of the boxplots in
-Figure~\ref{fig:Boxplot}. The mismatch is due to the heat effect, since the venue
-effect only captures a smaller part of the variation than the heat effect as
-evident from their standard deviations. It is possible to calculate the extremity
-of the 2022 heat effect as we know that the mean of the heat effects is $0$,
-the standard deviation is $0.42$ as given by Table~\ref{tab:Gamma_parameters},
-and the estimate of the 2022 heat effect is $-0.93$. We find the tail probability
-to be $0.0128$ indicating there is a low chance of observing a venue effect as
-extreme as 2022.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics{GGDensity}
-  \caption{The densities of the GG distribution with random effects for heat
-  and venue (black), and without random effects (blue)}
-  \label{fig:DensityPlot}
-\end{figure}
-\of{Not sure what to write about the density plot or how to fit it into the rest
-of the section.}
+\jy{Could you also test sensitivity for excluding the positive but 
+  fault reaction times?}
 
@@
 \end{table}
-\of{table has been updated with gg3b results}
-\jy{Could you also test sensitivity for excluding the positive but
-  fault reaction times?}
 
-The GLMM with both venue and heat random effects helps us to assess how extreme
-a reaction time below a given threshold is. The probability of observing a reaction
-time below a threshold, assuming no intentional false starts, can be
-approximated by generating a large number of realizations from the fitted
-GLMM. We used 10 millions realizations in the analysis. 
-Table~\ref{tab:Sim_probability} summarizes the estimated probabilities of
-observing a reaction time below 0.08, 0.09, and 0.10 seconds under two different
-scenarios: one excluding and the other including data from 2022. Apparently,
-excluding 2022 lowered the probability of observing a fast reaction time, but
-the difference is not big in magnitude. For example, the probability for
-seeing a reaction time below 0.10 seconds changes from $2.36\cdot 10^{-3}$ to
-$2.90\cdot 10^{-3}$ when 2022 is included. These values correspond to one in
-every 344 starts and 423 starts, respectively. The table shows that changing the
-reaction time barrier from 0.1 seconds to 0.08 seconds drastically reduces the
-chance of observing a reaction time below the barrier, from one in roughly every
-283 starts to one in every 58480 starts using data with 2022 included. The
-results substantiates the recommendations by \citet{komi2009iaaf}.
+The fitted generalized Gamma GAMLSS model with both venue- and heat-level 
+random effects provides a framework for assessing how extreme reaction times 
+below certain thresholds are. The probability of observing a reaction time 
+below a given threshold, assuming no intentional false starts, was approximated 
+by generating 10 million realizations from the fitted model. 
+Table~\ref{tab:Sim_probability} summarizes the probabilities of observing 
+reaction times below 0.08, 0.09, and 0.10 seconds under two scenarios: one 
+excluding and the other including data from 2022. Excluding 2022 slightly 
+reduces the probability of observing a fast reaction time, but the difference 
+is small. For example, the probability of a reaction time below 0.10 seconds 
+decreases from $2.76 \cdot 10^{-3}$ (approximately one in 362 starts) to 
+$1.94 \cdot 10^{-3}$ (approximately one in 515 starts) when 2022 is excluded. 
+Lowering the reaction time threshold from 0.10 to 0.08 seconds drastically 
+reduces the likelihood of observing a reaction time below the barrier, with 
+the probability dropping from one in every 362 starts (at 0.10 seconds) to one 
+in every 14620 starts (at 0.09 seconds) and one in every 146198 starts (at 0.08 
+seconds) when 2022 is included. These results highlight the rarity of extremely 
+fast reaction times and substantiate the recommendations of \citet{komi2009iaaf} 
+to carefully consider the selection of reaction time thresholds.
 
@@
 \end{table}
-\of{table has been updated with gg3b results}
 
-Utilizing the same model, we can subtly shift our interpretation to determine a
-suitable reaction time barrier based on the probability of observing a time
-below said barrier. Including 2022 data, a reaction time barrier of 0.092
-seconds is warranted to maintain a 0.1\% chance of observing an illicitly fast
-reaction time, as delineated in Table~\ref{tab:Sim_time}. This calculation can
-be replicated for various probability levels; for instance, a probability of
-0.0001 (equating to one disqualification per 10,000 starts) necessitates a
-reaction time barrier of 0.083 seconds. The exclusion of 2022 data, which
-inherently lowers the probability of observing a swift reaction time, mandates a
-less lenient reaction time barrier when calculated based on probability, albeit
-without a substantial difference in magnitude.
+Utilizing the same model, we can determine suitable reaction time barriers 
+based on the probability of observing a time below the barrier. As shown in 
+Table~\ref{tab:Sim_time}, including the 2022 data suggests a reaction time 
+barrier of 0.094 seconds to maintain a 0.1\% chance of observing an 
+exceptionally fast reaction time, while a stricter threshold of 0.082 seconds 
+is needed to limit this probability to 0.01\%. Excluding the 2022 data results 
+in slightly higher thresholds of 0.096 and 0.083 seconds for the respective 
+probability levels. These results indicate that while the inclusion of 2022 
+data slightly reduces the recommended barrier, the magnitude of the difference 
+is relatively small. This approach allows for tailoring reaction time 
+thresholds to desired levels of false positive rates, balancing fairness and 
+precision in disqualification criteria.
 
@@
 and other competitions. Further, our GLMM results indicate that a reaction time
-of 0.1 seconds may not be as exceptional as commonly believed. According to the
+of 0.1-second may not be as exceptional as commonly believed. According to the
 probabilities in Table~\ref{tab:Sim_time}, World Athletics might contemplate

### 2024-11-24 Manuscript/reply.tex
smoothed results section

@@
 
-Thank you for the feedback. The purpose of Figure~3 is to display the venue 
-effects for each year and visually illustrate how 2022 compares to other years. 
-To address your concern, we added additional comments quantifying the 
-probability of observing an extreme venue effect, which strengthens the 
-argument that 2022 is anomalous. If needed, we are open to revising the plot 
-further or integrating it into a single panel showing deltas, as suggested.
-\eds{let's revisit this again after objective 2 analyses are set}
+Given the confusion it may cause, this figure has been removed.
+
+% Thank you for the feedback. The purpose of Figure~3 is to display the venue 
+% effects for each year and visually illustrate how 2022 compares to other years. 
+% To address your concern, we added additional comments quantifying the 
+% probability of observing an extreme venue effect, which strengthens the 
+% argument that 2022 is anomalous. If needed, we are open to revising the plot 
+% further or integrating it into a single panel showing deltas, as suggested.
+% \eds{let's revisit this again after objective 2 analyses are set}
 

### 2024-11-24 Manuscript/manuscript.tex
smoothed discussion

@@
 
-We set out to answer two questions: were athletes who competed at the 2022 World
-Track and Field Championships faster than at other races and is the 0.1 second
-reaction time barrier fair.  In regards to the first question, we do believe
-that athletes were faster than in comparable meets.  Ideally we would like to
-be able to draw data from a central database containing every World Athletics
-certified meet which we could use to reference athlete's reaction times.
-Unfortunately that is not available at the moment and nearly all of the results
-listed on the World Athletics website do not contain reaction times.  The results
-of the study could have been stronger if we had been able to look at every race
-that athletes like Devon Allen ran in 2022 and analyzed his results at the World
-Championship against many other data points.
+This study sought to answer two primary questions: were athletes who
+competed at the 2022 World Track and Field Championships faster than
+at other races, and is the 0.1-second reaction time barrier fair?
+Regarding the first question, our analyses suggest that athletes were
+indeed faster at the 2022 World Championships compared to comparable
+meets. Ideally, a centralized database containing reaction times from
+all World Athletics-certified meets would have allowed for a more
+comprehensive analysis. Currently, however, reaction times are seldom
+included in the results available on the World Athletics website. This
+limitation constrained our ability to evaluate athletes like Devon
+Allen, as data from his races leading up to the World Championships in
+2022 were not accessible. A more robust dataset would enable deeper
+insights by allowing comparisons across a larger set of races.
 
-In addition to the GLMM results results presented above women, the supplementary
-file contains results about the women's GLMM results.  We fit a generalized
-gamma model to womens 100 meter dash and 100 meter hurdle data from 1999 to 2023
-to determine an appropriate reaction time barrier.  We chose to separate these
-two studies due to publications suggesting reaction time differences in men and
-women.
+In addressing the second question, our analyses of the GAMLSS model
+reveal that the 0.1-second reaction time barrier, while rare, may not
+be as extraordinary  as traditionally assumed. For men,
+Table~\ref{tab:Sim_time} shows that reaction  times below 0.1 seconds
+occur with a probability of approximately one in 362 starts  when
+including the 2022 data. Lowering the threshold to 0.08 seconds
+drastically  reduces this likelihood, supporting the idea that the
+current barrier could be adjusted to reflect more realistic
+probabilities of false starts. A similar pattern is observed for
+women, as detailed in the Supplementary Material, where reaction
+times below 0.1 seconds are exceedingly rare for the 100-meter dash
+and 100-meter  hurdles. However, the uniformity of the 0.1-second
+barrier for both men and women  is questionable, given numerous
+studies documenting gender differences in reaction  times
+\citep[e.g.,][]{lipps2011implications, babicc2009reaction,
+  panoutsakopoulos2020gender}. These studies suggest that the current
+threshold may  unfairly penalize men, for whom sub-0.1-second reaction
+times are more probable. \citet{brosnan2017effects} advocate for
+gender-specific barriers, a position that  aligns with our findings
+and highlights the importance of tailoring thresholds to  biological
+distinctions.
 
-\of{I think there are some good ideas in these 2 paragraphs but need help
-integrating them. Prof Schifano can you help with this?}
-
-The uniformity of the reaction time barrier for both men and women is
-perplexing, especially in light of numerous studies that suggest a divergence in
-their respective reaction times \citep[e.g.,][]{lipps2011implications,
-  babicc2009reaction, panoutsakopoulos2020gender}. These authors posit that
-World Athletics may be overlooking inherent gender differences by establishing
-an equal reaction time barrier. If gender disparities do exist and 0.1 seconds
-is deemed a fair threshold for women, it logically follows that the same cannot
-be deemed fair for men, given that the probability of observing sub-0.1-second
-reaction times would be inherently higher. \citet{brosnan2017effects} propounds
-the adoption of gender-specific reaction time barriers, a stance that appears
-logical when considering biological distinctions in reaction times between
-genders.
-
-This paper aspires to offer a statistical lens through which to examine Allen's
-disqualification at the 2022 World Track and Field Championships, rather than
-delivering a conclusive judgment regarding potential equipment malfunction. Our
-findings reveal that athletes' reaction times at the event were, in general,
-faster than those recorded at other competitions, with
-Table~\ref{tab:Clusrankresults} presenting multiple significant p-values that
-highlight disparities in average reaction times between the 2022 Championship
-and other competitions. Further, our GLMM results indicate that a reaction time
-of 0.1-second may not be as exceptional as commonly believed. According to the
-probabilities in Table~\ref{tab:Sim_time}, World Athletics might contemplate
-adjusting the disqualification barrier to 0.08 seconds, thereby enabling athletes
-like Allen to react more swiftly without the apprehension of
-disqualification. In summary, while the results designate 2022 as an anomalous
-year, Allen's time, despite resulting in disqualification, may not be
-categorically extreme.
+This study provides a statistical framework to examine Devon Allen’s
+disqualification at the 2022 World Track and Field Championships,
+offering insights rather than drawing definitive conclusions about
+potential equipment malfunction. Our findings indicate that reaction
+times at the 2022 Championships were, on average, faster than at other
+competitions, as evidenced by the significant p-values in
+Table~\ref{tab:Clusrankresults}. Additionally, the GAMLSS results
+suggest that the 0.1-second barrier may not be as stringent as
+previously believed. Based on Table~\ref{tab:Sim_time}, a stricter
+threshold of 0.08 seconds could be considered, allowing athletes like
+Allen to react swiftly without undue risk of disqualification. While
+this analysis provides a rigorous statistical perspective, it does not
+consider biomechanical factors, such as individual variability in
+neuromuscular response times or the role of starting block sensors in
+detecting pressure changes, which may offer more direct evidence of
+reaction capabilities. In summary, while the results designate 2022 as
+an anomalous year, Allen’s time, despite resulting in
+disqualification, may not be categorically extreme.
 

### 2024-11-24 Manuscript/manuscript.tex
rewrote abstract and lit review in intro

@@
 \begin{abstract}
-Devon Allen was disqualified from the men's 110-meter hurdle final at
-the 2022 World Track and Field Championships after recording a
-reaction time of 0.099 seconds—just 0.001 seconds faster than the
-allowable threshold of 0.1 seconds. This paper explores whether the
-reaction times of athletes at the 2022 Championships were unusually
-fast compared to other competitions and evaluates the appropriateness
-of the 0.1-second disqualification barrier. To investigate these
-issues, we employ a rank-sum test for clustered data to compare
-athletes' reaction times across multiple competitions, along with a
-generalized linear mixed model (GLMM) that accounts for both venue and
-heat effects. The results indicate that athletes at the 2022
-Championships had significantly faster reaction times on average. This
-analysis raises important questions about the validity of the current
-reaction time threshold and suggests potential modifications to the
-rules that could prevent future disqualifications under similar
-circumstances.
+Devon Allen’s disqualification at the men's 110-meter hurdle final at
+the 2022 World Track and Field Championships, 
+due to a reaction time (RT) of 0.099 seconds---just 0.001 seconds below
+the allowable threshold---sparked widespread debate over 
+the fairness and validity of RT rules. This study investigates two key
+issues: variations in timing systems and the justification for the
+0.1-second disqualification threshold. We
+pooled RT data from men’s 110-meter hurdles and 100-meter dash, as
+well as women’s 100-meter hurdles and 100-meter dash, spanning
+national and international competitions. Using a rank-sum test for
+clustered data, we compared RTs across multiple competitions,
+while a generalized Gamma model with random effects for venue and heat
+was applied to evaluate the threshold. Our analyses reveal significant
+differences in RTs between the 2022 World Championships and other
+competitions, pointing to systematic variations in timing
+systems. Additionally, the model shows that RTs below 0.1 seconds,
+though rare, are physiologically plausible. These findings highlight
+the need for standardized timing protocols and a re-evaluation of the
+0.1-second disqualification threshold to promote fairness in
+elite competition.
 
 \bigskip\noindent{\sc Keywords}:
-False Start, GLMM, Hurdle, Reaction Time, Seiko, rank-based test
-
+false start, GAMLSS, reaction time, rank-based test, short sprint
 \end{abstract}
@@
 Track and Field Championships in Eugene, Oregon, ended in
-controversy when he was disqualified for a reaction time of
+controversy when he was disqualified for a reaction time (RT) of
 0.099 seconds, just 0.001 seconds below the allowable threshold.
@@
 through the preliminary heats and semifinals at the World
-Championships, with reaction times of 0.123 and 0.101 seconds,
+Championships, with RTs of 0.123 and 0.101 seconds,
 respectively. However, in the final heat, competing in front of his
-home audience, Allen’s reaction time was just 0.001 seconds faster
+home audience, Allen’s RT was just 0.001 seconds faster
 than the 0.1-second threshold set by the International Association
 of Athletics Federations (IAAF), resulting in his disqualification, a
-decision that was met with widespread public outcry.
+decision that was met with widespread public outcry. This 
+incident highlighted two long-standing issues: variability in the
+measurement of RTs by Start Information Systems (SIS) and the
+appropriateness of the 0.1-second disqualification threshold. As
+reaction times are measured in fractions of a second, inconsistencies
+in timing technologies and rules can significantly affect athlete
+outcomes, raising questions about fairness and standardization.
 
-The disqualification of Allen was widely discussed in online
-communities, with some questioning the accuracy of the timing
-equipment. One of the most active platforms was \url{www.LetsRun.com},
-a website that functions as both a message board and a news outlet.
-The message board, similar to Reddit, is particularly active during
-major running events, including the Olympics and the World Track
-and Field Championships. The news section of the website featured
-articles such as \citet{johnson2022data} and \citet{johnson2022was}
-by Robert Johnson, the site’s founder. Johnson argued that there may
-have been an error with the timing equipment at the 2022 Championships.
-He cited graphs and descriptive statistics, like median reaction
-times, comparing the reaction times at the 2022 Championships to
-those from other competitions, including the U.S. Championships and
-previous World Championships. However, these comparisons lacked formal
-statistical analysis.
+World Athletics (formerly IAAF) uses certified SIS to measure RTs, yet
+variation in technology persist. Discussions at online forums such as
+\url{www.LetsRun.com}questioned consistencies with the SIS as a
+contributing factor to RT anormalies \citep{johnson2022data,
+  johnson2022was}. ALthough lacking formal statistical analysis, the
+discussions sparkled broader concerns about SIS
+variability. Historically, ``loud gun'' systems caused signal delays
+for athletes in outer lanes  due to the speed of sound, an issue
+addressed with the introduction of ``silent gun'' systems in 2010,
+which electronically synchronize sound delivery to all athletes
+\citep{tonnessen2013reaction}. Despite these advances, variability
+persists due to differences in sensor technologies, such as force
+transducers and accelerometers, and inconsistencies in event detection
+algorithms \citep{willwacher2013novel}. For example, simple
+force-threshold systems may delay RT detection by up to 26~ms compared
+to more sophisticated methods \citep{pain2007sprint}. These findings
+emphasize the need for standardized certification protocols to reduce
+discrepancies and ensure fairness in RT measurements, as recently
+reviewed by \citet{milloz2021sprint}.
 
-The 0.1-second disqualification threshold in track events has been
-a subject of ongoing debate, with researchers presenting evidence
-both in favor of lowering and raising the
-limit. \citet{pain2007sprint} examined
-reaction times in nine male athletes and found that the fastest athlete
-had an average reaction time of 0.087 seconds, with a standard deviation
-of 0.004 seconds. Similarly, \citet{komi2009iaaf}, in a study commissioned
-by the IAAF, recommended lowering the disqualification threshold to 0.08
-or 0.085 seconds, arguing that athletes are capable of faster reaction
-times than the current threshold allows. On the other hand, some
-researchers have argued for increasing the threshold.
-\citet{brosnan2017effects} suggested raising it to 0.115 seconds based
-on data from European and World Championships, while a study of the
-2008 Olympics supported the conclusion that athletes cannot reliably
-react within 0.1 seconds \citep{lipps2011implications}. Interestingly,
-\citet{brosnan2017effects} and \citet{lipps2011implications} relied on
-retrospective event data, whereas \citet{pain2007sprint} and
-\citet{komi2009iaaf} conducted controlled experiments to assess
-reaction time capabilities. The IAAF has not revised the threshold
-since its implementation in 1999.
+The 0.1-second disqualification threshold, introduced in the 1990s,
+has been the subject of significant debate. Partly based on limited
+data from Finnish national-level athletes~\cite{mero1990reaction}, the
+threshold may not adequately represent the capabilities of elite
+sprinters. Controlled experiments have shown that reaction times below
+0.1 seconds are physiologically plausible \citep{pain2007sprint,
+  komi2009iaaf}, while retrospective analyses of competition data
+often advocate for raising the threshold~\cite{brosnan2017effects,
+  lipps2011implications}. Stricter false-start rules, introduced to
+minimize race disruptions, have also discouraged sprinters from
+attempting faster starts, which may artificially inflate RTs recorded
+in competition~\cite{haugen2013effect}. This divergence between
+experimental findings and competition-based analyses illustrates the
+complexity of defining a universally fair threshold. Addressing these
+debates requires modern data collection and advanced methodologies to
+ensure equity and consistency in elite
+competition~\cite{milloz2021sprint}.
 
-This paper has two primary objectives. First, we investigate whether
-reaction times at the 2022 World Championships were significantly
-different from other competitions, focusing on athletes who competed
-in multiple events. Using a matched-pairs design, we compare reaction
-times across the 2022, 2019, and 2023 World Championships, as well as
-the 2022 national-level competitions. This approach isolates the
-effect of the competition year while controlling for individual
-performance. The clustered nature of the data, with the goal of
-assessing differences across events within-athletes, were analyzed
-with rank-based comparison approach for clustered data
-\citep{datta2005rank}. The second objective is to assess the
-appropriateness of the 0.1-second threshold by modeling reaction
-times from World Championships held from 1999 onward. We use a
+This paper addresses two primary objectives from a statistical
+perspective, using modern methodologies to analyze historical
+data. First, we investigate whether reaction times (RTs) at the 2022
+World Championships were significantly different from other
+competitions, focusing on athletes who competed in multiple
+events. Using a matched-pairs design, we compare RTs across the 2022,
+2019, and 2023 World Championships, as well as the 2022 national-level
+competitions. This approach isolates the effect of the competition
+year while controlling for individual performance. The clustered
+nature of the data, with the goal of assessing differences across
+events within athletes, was analyzed using a rank-based comparison
+approach for clustered data \citep{datta2005rank}. Second, we evaluate
+the appropriateness of the 0.1-second disqualification threshold by
+modeling RTs from World Championships held from 1999 onward. A
 generalized gamma distribution with random effects for both venue and
-heat in the framework of the generalized additive model for location,
-scale, and shape (GAMLSS) \citep{stasinopoulos2024generalized}.
-This model estimates the probability of reaction times falling below
-the threshold, helping evaluate the consistency of reaction times and
-whether the current threshold is statistically justified.
+heat was applied within the framework of the generalized additive
+model for location, scale, and shape (GAMLSS)
+\citep{stasinopoulos2024generalized}. This model estimates the
+probability of RTs falling below the threshold, providing a
+statistical assessment of RT consistency and the validity of the
+current threshold.
 
@@
 methods for objective~1, rank-based comparison for clustered data with
-subunit-level gruping to analyze differences in reaction times
-across different competitions. Section~\ref{sec:gamlss} details the
-methods for objective~2, the GAMLSS to evaluate the appropriatenes
+subunit-level gruping to analyze differences in RTs across different
+competitions. Section~\ref{sec:gamlss} details the methods for
+objective~2, the gernalized Gamma GAMLSS to evaluate the appropriatenes
 of the 0.1-second threshold. Section~\ref{sec:Results} presents the
@@
 
-This study focuses on reaction times from sprint events where reaction
+This study focuses on RTs from sprint events where reaction
 time plays a critical role. Specifically, we included data from men’s
@@
 dashes and 100-meter hurdles. Statistical tests showed no significant
-differences in reaction times between these events, supporting their
+differences in RTs between these events, supporting their
 inclusion in a unified analysis. However, we excluded data from
-200-meter dashes and longer events, as their reaction times were found
+200-meter dashes and longer events, as their RTs were found
 to be significantly different. This difference is not unexpected, as
-reaction time is generally less critical in longer sprint events
+RT is generally less critical in longer sprint events
 compared to shorter distances where explosive starts are paramount.
@@
 Two datasets were used to address the study's objectives. To
-investigate whether reaction times at the 2022 World Championships
+investigate whether RTs at the 2022 World Championships
 were significantly different from other competitions (Objective 1), we
@@
 Championships, or 2023 World Championships). To assess the
-appropriateness of the 0.1-second reaction time threshold (Objective
-2), we used reaction times from the 110-meter hurdles and 100-meter
+appropriateness of the 0.1-second RT threshold (Objective
+2), we used RTs from the 110-meter hurdles and 100-meter
 dash across all World Championships held from 1999 to 2023. This
@@
 Championships in August, this provided a unique opportunity to establish a
-baseline for reaction times.
+baseline for RTs.
 
@@
 \jy{Owen, are there names from 100-m dash?}
-To compare reaction times at the 2022 World Championships with those
+To compare RTs at the 2022 World Championships with those
 recorded at national competitions, we constructed a dataset of athletes
@@
 Track and Field Championships and the World Championships. Each athlete
-recorded faster reaction times in all World Championships races compared
+recorded faster RTs in all World Championships races compared
 to their performances at the national-level event. However, this small
 sample size was insufficient for meaningful statistical analysis. To
-expand the dataset, we included reaction times from additional national
+expand the dataset, we included RTs from additional national
 competitions held between May and July 2022 across various countries.
-Compiling this data was challenging, as reaction times were not centrally
+Compiling this data was challenging, as RTs were not centrally
 located and often required searching country-specific websites, with many
@@
 
-The final dataset consisted of reaction times from athletes who competed
-in both national and international competitions. Reaction times from
+The final dataset consisted of RTs from athletes who competed
+in both national and international competitions. RTs from
 preliminary heats, semifinals, and finals were included to ensure
@@
 significantly reduced the number of athletes and clusters. Each
-athlete was considered as a cluster, with observed reaction times
+athlete was considered as a cluster, with observed RTs
 from both the `treatment' group (2022 World Championships) and
@@
 sizes ranged from three to six, with a median size of four. Because gender
-is known to influence reaction times \citep{babicc2009reaction,
+is known to influence RTs \citep{babicc2009reaction,
   lipps2011implications}, we prepared data for men and women
-separately, resulting in 80 reaction times from 17 athletes for each
+separately, resulting in 80 RTs from 17 athletes for each
 gender. This setup forms the basis for the rank-based comparison for
@@
 within-athlete dependence inherent in this data structure. The top
-panel of Figure!~\ref{fig:RankScatterplots} shows the reaction times
+panel of Figure!~\ref{fig:RankScatterplots} shows the RTs
 of those who competed both at national competitions and the 2022 World
@@
   \includegraphics[width=\textwidth]{RankScatterPlots}
-  \caption{Reaction times for athletes who competed at the 2022 World Championships
+  \caption{RTs for athletes who competed at the 2022 World Championships
 	and at another championship (2022 national, 2019 World, or 2023 World) at which
@@
 
-To compare reaction times at the 2022 World Championships with those
+To compare RTs at the 2022 World Championships with those
 recorded in the 2019 and 2023 World Championships, we prepared datasets
-including reaction times from athletes who competed in the 2022 World
+including RTs from athletes who competed in the 2022 World
 Championships and at least one of the 2019 or 2023 World Championships.
@@
 the 2022 and 2023 World Championships. In the 2019-2022 comparison,
-reaction times from 2022 were treated as the `treatment' group, with
+RTs from 2022 were treated as the `treatment' group, with
 2019 serving as the `control' group. Similarly, in the 2022-2023
-comparison, reaction times from 2022 were treated as the `treatment'
+comparison, RTs from 2022 were treated as the `treatment'
 group, with 2023 serving as the `control' group. This structure allowed
-us to prepare datasets suitable for examining reaction times of athletes
+us to prepare datasets suitable for examining RTs of athletes
 who competed across multiple World Championships.
@@
 times from different World Championships. The dataset for the 2019-2022
-comparison contained 134 reaction times from 34 male athletes and 124
-reaction times from 31 female athletes. The dataset for the 2022-2023
-comparison contained 161 reaction times from 45 male athletes and 182
-reaction times from 47 female athletes. While it is theoretically
-possible that athletes improved their reaction times between 2019 and
+comparison contained 134 RTs from 34 male athletes and 124
+RTs from 31 female athletes. The dataset for the 2022-2023
+comparison contained 161 RTs from 45 male athletes and 182
+RTs from 47 female athletes. While it is theoretically
+possible that athletes improved their RTs between 2019 and
 2022 or between 2022 and 2023, such improvements are highly unlikely for
@@
 
-Figure~\ref{fig:RankScatterplots} shows the reaction times of athletes
+Figure~\ref{fig:RankScatterplots} shows the RTs of athletes
 who competed in both the 2019 and 2022 World Championships (middle
@@
 at least one race at each championship. Notably, Devon Allen recorded
-the fastest reaction times in both the Finals and Semifinals of the
+the fastest RTs in both the Finals and Semifinals of the
 2022 World Championships, but his disqualification was determined by a
-difference of just 0.002 seconds, with reaction times of 0.101 and
+difference of just 0.002 seconds, with RTs of 0.101 and
 0.099 seconds, respectively. This highlights the critical role of
-reaction time precision in elite-level competition.
+RT precision in elite-level competition.
 
@@
 100-meter dashes and their analyses were relegated to the
-supplementary material. We focus on the reaction times recorded during
-semifinal and final heats only, as reaction times from preliminary
+supplementary material. We focus on the RTs recorded during
+semifinal and final heats only, as RTs from preliminary
 heats are often not as fast as those in later heats
@@
   brosnan2017effects, zhang2021correlation}. For analysis purposes, we
-pooled reaction times from semifinal and final heats to increase sample
+pooled RTs from semifinal and final heats to increase sample
 size, which is particularly important for years with limited final heat
@@
 The data is summarized in Figure~\ref{fig:Boxplot}, which presents a
-sequence of boxplots of reaction times from 1999 to 2023. It is evident
-that reaction times in 2022 were notably faster, with a median reaction
+sequence of boxplots of RTs from 1999 to 2023. It is evident
+that RTs in 2022 were notably faster, with a median reaction
 time of 0.129 seconds compared to the 0.156 seconds observed in earlier
@@
 Figure~\ref{fig:Boxplot} also highlights year-to-year variability in
-reaction times, likely influenced by changes in the championship venue
+RTs, likely influenced by changes in the championship venue
 and environmental conditions such as humidity, precipitation, and
@@
   \includegraphics[width=\textwidth]{Boxplot}
-  \caption{The reaction times from 1999 to 2023 for the men's 110 meter hurdle
+  \caption{The RTs from 1999 to 2023 for the men's 110 meter hurdle
   and 100 meter dash.}
@@
 
-Between 2007 and 2009, World Athletics (formerly the IAAF) allowed one
+Between 2007 and 2009, World Athletics allowed one
 false start warning before disqualifying a sprinter \citep{iaaf2009falsestart}.
@@
 \citep{iaaf2009falsestart}. \citet{haugen2013effect} demonstrated that more
-lenient false start rules significantly improved reaction times during the
+lenient false start rules significantly improved RTs during the
 1997–2009 period, suggesting that rule changes over the study period may
-have contributed to variations in reaction times across years.
+have contributed to variations in RTs across years.
 
@@
 To test the conjecture that the 2022 World Championships timing device may have
-led to faster recorded reaction times, we compare the reaction times of the same
+led to faster recorded RTs, we compare the RTs of the same
 athletes who have attended both the 2022 World Championships and other
@@
 In this setting, we have clustered data with subunit grouping. In particular,
-each athlete is a cluster and the multiple reaction times from the same athlete
+each athlete is a cluster and the multiple RTs from the same athlete
 can be from either the 2022 World Championships or otherwise.
-Let $X_{ij}$ be the $j$th reaction time of athlete~$i$, $i = 1, \ldots, n$,
+Let $X_{ij}$ be the $j$th RT of athlete~$i$, $i = 1, \ldots, n$,
 $j = 1, \ldots, m_i$ where $m_i$ is the number of observations from
@@
 
-Based on an exploratory analysis, the reaction times are adequately
+Based on an exploratory analysis, the RTs are adequately
 modeled by a Generalized Gamma (GG) distribution with random effects in
@@
 the distribution to model asymmetric and heavy-tailed data effectively, making
-it particularly suitable for reaction times.
+it particularly suitable for RTs.
 An implementation of this distribution is available from R package
@@
 
-Let $Y_{ijk}$ denote the reaction time of observation~$k$ in heat~$j$
+Let $Y_{ijk}$ denote the RT of observation~$k$ in heat~$j$
 of year~$i$. Conditioning on a venue effect $v_i$ for year~$i$
@@
 random numbers generated from the fitted mixture distribution can be used to
-approximate the probability of observing a reaction time faster than any given
-threshold. We are specifically interested in the probability of a reaction time
+approximate the probability of observing a RT faster than any given
+threshold. We are specifically interested in the probability of a RT
 being less than 0.1 seconds in order to gauge if that is a reasonable
@@
 The rank-based methods described in Section~\ref{sec:rank} were used to
-compare reaction times between the 2022 World Championships and other
+compare RTs between the 2022 World Championships and other
 competitions in which the same athletes participated. These comparisons
 were conducted separately for men and women, resulting in six total
-comparisons: reaction times from the 2022 national-level championships
-versus the 2022 World Championships for men and women, reaction times
+comparisons: RTs from the 2022 national-level championships
+versus the 2022 World Championships for men and women, RTs
 from the 2019 versus 2022 World Championships for men and women, and
-reaction times from the 2022 versus 2023 World Championships for men and
+RTs from the 2022 versus 2023 World Championships for men and
 women.
@@
   \caption{P-values of comparisons between
-    reaction times from different competitions for the same athletes.
-    2022 Nat. vs Inter. compares reaction times from 2022 national-level
+    RTs from different competitions for the same athletes.
+    2022 Nat. vs Inter. compares RTs from 2022 national-level
     championships and the 2022 World Track and Field Championships. 2019
-    vs 2022 compares reaction times from the 2019 and 2022 World Track and
-    Field Championships. 2022 vs 2023 compares reaction times from the
+    vs 2022 compares RTs from the 2019 and 2022 World Track and
+    Field Championships. 2022 vs 2023 compares RTs from the
     2022 and 2023 World Track and Field Championships.}
@@
 that the national versus international comparisons yielded the lowest
-p-values, indicating stronger evidence of faster reaction times at the
+p-values, indicating stronger evidence of faster RTs at the
 2022 World Championships compared to the 2022 national competitions. For
@@
 providing substantial evidence that athletes competing at the 2022 World
-Championships had significantly faster reaction times than they did just
+Championships had significantly faster RTs than they did just
 a few months earlier at national-level events. The comparisons between the
@@
 finding supports the hypothesis that the 2022 World Championships presented
-conditions conducive to faster reaction times, potentially due to systematic
+conditions conducive to faster RTs, potentially due to systematic
 or environmental factors.
@@
 decreases slightly to 0.3198. These results
-highlight that reaction times are influenced by both venue and heat-level
+highlight that RTs are influenced by both venue and heat-level
 factors, with heat-level variability playing a more prominent role in the scale
 parameter, and that the inclusion of 2022 introduces greater venue-level
-variability, likely due to systematic differences in reaction times that year.
+variability, likely due to systematic differences in RTs that year.
 
@@
     observations drawn from the fitted model overlaid with the
-    histogram of the observed reaction times (left); Q-Q plot of the
+    histogram of the observed RTs (left); Q-Q plot of the
     normal z-score of the quantile residuals from the fitted model.}
@@
 panel compares the kernel density estimate of one million simulated reaction
-times from the fitted model to the histogram of the observed reaction times.
+times from the fitted model to the histogram of the observed RTs.
 The close alignment between the density curve and the histogram suggests that
 the fitted GG model adequately captures the overall distribution of the
-reaction times. The right panel shows a Q-Q plot of the z-scores of the
+RTs. The right panel shows a Q-Q plot of the z-scores of the
 quantile residuals from the fitted model. The points lie approximately along
@@
 fit. These diagnostics collectively demonstrate that the fitted model provides
-a reasonable representation of the observed reaction time data.
+a reasonable representation of the observed RT data.
 
@@
 \jy{Could you also test sensitivity for excluding the positive but
-  fault reaction times?}
+  fault RTs?}
 
@@
   \centering
-  \caption{Probabilities of observing reaction times less than threshold 0.08,
+  \caption{Probabilities of observing RTs less than threshold 0.08,
   0.09, and 0.10 seconds based on the two effect model.}
@@
 The fitted generalized Gamma GAMLSS model with both venue- and heat-level
-random effects provides a framework for assessing how extreme reaction times
-below certain thresholds are. The probability of observing a reaction time
+random effects provides a framework for assessing how extreme RTs
+below certain thresholds are. The probability of observing a RT
 below a given threshold, assuming no intentional false starts, was approximated
@@
 Table~\ref{tab:Sim_probability} summarizes the probabilities of observing
-reaction times below 0.08, 0.09, and 0.10 seconds under two scenarios: one
+RTs below 0.08, 0.09, and 0.10 seconds under two scenarios: one
 excluding and the other including data from 2022. Excluding 2022 slightly
-reduces the probability of observing a fast reaction time, but the difference
-is small. For example, the probability of a reaction time below 0.10 seconds
+reduces the probability of observing a fast RT, but the difference
+is small. For example, the probability of a RT below 0.10 seconds
 decreases from $2.76 \cdot 10^{-3}$ (approximately one in 362 starts) to
 $1.94 \cdot 10^{-3}$ (approximately one in 515 starts) when 2022 is excluded.
-Lowering the reaction time threshold from 0.10 to 0.08 seconds drastically
-reduces the likelihood of observing a reaction time below the barrier, with
+Lowering the RT threshold from 0.10 to 0.08 seconds drastically
+reduces the likelihood of observing a RT below the barrier, with
 the probability dropping from one in every 362 starts (at 0.10 seconds) to one
@@
 seconds) when 2022 is included. These results highlight the rarity of extremely
-fast reaction times and substantiate the recommendations of \citet{komi2009iaaf}
-to carefully consider the selection of reaction time thresholds.
+fast RTs and substantiate the recommendations of \citet{komi2009iaaf}
+to carefully consider the selection of RT thresholds.
 
@@
   \centering
-  \caption{Suggested reaction time barriers based on tail probabilities.}
+  \caption{Suggested RT barriers based on tail probabilities.}
   \begin{tabular}{c c c}
@@
 
-Utilizing the same model, we can determine suitable reaction time barriers
+Utilizing the same model, we can determine suitable RT barriers
 based on the probability of observing a time below the barrier. As shown in
-Table~\ref{tab:Sim_time}, including the 2022 data suggests a reaction time
+Table~\ref{tab:Sim_time}, including the 2022 data suggests a RT
 barrier of 0.094 seconds to maintain a 0.1\% chance of observing an
-exceptionally fast reaction time, while a stricter threshold of 0.082 seconds
+exceptionally fast RT, while a stricter threshold of 0.082 seconds
 is needed to limit this probability to 0.01\%. Excluding the 2022 data results
@@
 data slightly reduces the recommended barrier, the magnitude of the difference
-is relatively small. This approach allows for tailoring reaction time
+is relatively small. This approach allows for tailoring RT
 thresholds to desired levels of false positive rates, balancing fairness and
@@
 competed at the 2022 World Track and Field Championships faster than
-at other races, and is the 0.1-second reaction time barrier fair?
+at other races, and is the 0.1-second RT barrier fair?
 Regarding the first question, our analyses suggest that athletes were
 indeed faster at the 2022 World Championships compared to comparable
-meets. Ideally, a centralized database containing reaction times from
+meets. Ideally, a centralized database containing RTs from
 all World Athletics-certified meets would have allowed for a more
-comprehensive analysis. Currently, however, reaction times are seldom
+comprehensive analysis. Currently, however, RTs are seldom
 included in the results available on the World Athletics website. This
@@
 In addressing the second question, our analyses of the GAMLSS model
-reveal that the 0.1-second reaction time barrier, while rare, may not
+reveal that the 0.1-second RT barrier, while rare, may not
 be as extraordinary  as traditionally assumed. For men,

### 2024-11-24 Manuscript/manuscript.tex
minor

@@
 modeling RTs from World Championships held from 1999 onward. A
-generalized gamma distribution with random effects for both venue and
+generalized Gamma (GG) distribution with random effects for both venue and
 heat was applied within the framework of the generalized additive
@@
 competitions. Section~\ref{sec:gamlss} details the methods for
-objective~2, the gernalized Gamma GAMLSS to evaluate the appropriatenes
+objective~2, the GG model with random effects to evaluate the appropriatenes
 of the 0.1-second threshold. Section~\ref{sec:Results} presents the
@@
 Based on an exploratory analysis, the RTs are adequately
-modeled by a Generalized Gamma (GG) distribution with random effects in
+modeled by a GG distribution with random effects in
 model parameters. The GG distribution has three parameters, denoted by
@@
 for $y > 0$, $\mu > 0$, $\sigma > 0$, and $\nu \neq 0$,
-where $z = \left(\frac{y}{\mu}\right)^\nu$,
+where $z = (y / \mu)^\nu$,
 $\theta = 1 / (\sigma^2 \nu^2)$, and
@@
 Weibull ($\mu = \nu)$ and  Gamma $(\nu = 1)$ distributions.
-Its mean is
+Its expectation is
 \[
-  \mu \frac{\Gamma(\theta + 1 / \nu)}
+  \frac{\mu \Gamma(\theta + 1 / \nu)}
   {\theta^{1 / \nu} \Gamma(\theta)},
@@
     parentheses and estimated variance of the random effects from the
-    fitted generalized Gamma distribution with venue level random
+    fitted GG distribution with venue level random
     effects in $\mu$ and heat level random effects in $\sigma$ in
@@
 
-The fitted parameters of the generalized Gamma distribution int he
+The fitted parameters of the GG distribution int he
 GAMLSS framework in Equations~\eqref{eq:gg}--\eqref{eq:sigma} are
@@
   \includegraphics[width=\textwidth]{diagnosis.pdf}
-  \caption{Diagnosis of the fitted generalized Gamma distribution with
+  \caption{Diagnosis of the fitted GG distribution with
     random effects in model parameters: kernerl density of 1 million
@@
 Figure~\ref{fig:diagnosis} presents diagnostic checks for the fitted
-generalized Gamma distribution model with random effects. The left
+GG distribution model with random effects. The left
 panel compares the kernel density estimate of one million simulated reaction
@@
 
-The fitted generalized Gamma GAMLSS model with both venue- and heat-level
+The fitted GG GAMLSS model with both venue- and heat-level
 random effects provides a framework for assessing how extreme RTs

### 2024-11-24 Manuscript/manuscript.tex
added gamlss original paper and package bib

@@
 variation in technology persist. Discussions at online forums such as
-\url{www.LetsRun.com}questioned consistencies with the SIS as a
+\url{www.LetsRun.com} questioned consistencies in the SIS as a
 contributing factor to RT anormalies \citep{johnson2022data,
-  johnson2022was}. ALthough lacking formal statistical analysis, the
-discussions sparkled broader concerns about SIS
-variability. Historically, ``loud gun'' systems caused signal delays
+  johnson2022was}. Historically, ``loud gun'' systems caused signal delays
 for athletes in outer lanes  due to the speed of sound, an issue
@@
 model for location, scale, and shape (GAMLSS)
-\citep{stasinopoulos2024generalized}. This model estimates the
+\citep{rigby2005generalized, stasinopoulos2024generalized}.
+This model enables estimation of the
 probability of RTs falling below the threshold, providing a
@@
 This model can be fit with R package \texttt{gamlss}
-\citep{stasinopoulos2024generalized}.
+\citep{stasinopoulos2008generalized}.
 

### 2024-11-25 Manuscript/manuscript.tex
responded to Prof Yan and Prof Schifano's comments

@@
 \jy{Owen, are there names from 100-m dash?}
+\of{Yes, there were four US athletes in the 100m dash that we have data on.
+Marvin Bracy, Fred Kerley, Travyon Bromwell, and Christian Coleman.  All of whom
+recorded their slowest times at the US Track and Field Championship.}
 To compare RTs at the 2022 World Championships with those
@@
   and event differece in the supplement?}
+\of{Asked you about this in an email.}
 
@@
 	}
+  \of{I agree and I updated it. I think it looks good now.}
   \label{fig:RankScatterplots}
@@
 \jy{Let's put the pooled analysis in the supplement.}
+\of{Yes I believe the distributions are similar but I think we cover why we don't
+pool by citing the papers that suggest differences in men and women's times.}
 
@@
   additional 100-dash data?}
+  \of{Funny enough the median reaction time remained exactly the same.}
 The data is summarized in Figure~\ref{fig:Boxplot}, which presents a
@@
 
-\jy{Please add standard errors, which could be from \texttt{summary(gg3b)}}
-
-\jy{Are these numbers correct? I modified with the gg3b in gg.R}
-
-\begin{table}
-  \centering
-  \caption{Estimated fixed-effect parameters with standard errors in
-    parentheses and estimated variance of the random effects from the
-    fitted GG distribution with venue level random
-    effects in $\mu$ and heat level random effects in $\sigma$ in
-    Models~\eqref{eq:gg}--\eqref{eq:sigma}.}
-  \label{tab:ggfit}
-  \begin{tabular}{c c c c c c}
-    \toprule
-    Data set & $\beta_0$ & $\gamma_0$ & $\nu$ & $\sigma_v$ & $\sigma_h$ \\
-    \midrule
-    Excluding 2022 & $-$1.910 (0.005)  & $-$2.200 (0.025) & $-$1.177 (0.442) & 0.043 & 0.326 \\
-    Including 2022 & $-$1.910 & $-$2.200 & $-$1.178 & 0.058 & 0.320 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
 The fitted parameters of the GG distribution int he
@@
 variability, likely due to systematic differences in RTs that year.
+
+\jy{Please add standard errors, which could be from \texttt{summary(gg3b)}}
+
+\jy{Are these numbers correct? I modified with the gg3b in gg.R}
+\of{Yes I agree with your numbers and have included the ones for the other model}
+
+\of{I moved the table down because in the pdf it was appearing in the rank
+section.} 
+\of{I think it's confusing that sigma appears as a parameter of the model
+and we use it as the standard deviation of the heat effect which appears in the
+sigma parameter. I don't know much about statistical nomenclature in papers, is
+this common?}
+
+\begin{table}
+  \centering
+  \caption{Estimated fixed-effect parameters with standard errors in
+    parentheses and estimated variance of the random effects from the
+    fitted GG distribution with venue level random
+    effects in $\mu$ and heat level random effects in $\sigma$ in
+    Models~\eqref{eq:gg}--\eqref{eq:sigma}.}
+  \label{tab:ggfit}
+  \begin{tabular}{c c c c c c}
+    \toprule
+    Data set & $\beta_0$ & $\gamma_0$ & $\nu$ & $\sigma_v$ & $\sigma_h$ \\
+    \midrule
+    Excluding 2022 & $-$1.910 (0.005) & $-$2.200 (0.025) & $-$1.177 (0.442) & 0.043 & 0.326 \\
+    Including 2022 & $-$1.910 (0.005) & $-$2.200 (0.027) & $-$1.178 (0.447) & 0.058 & 0.320 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
   fault RTs?}
+\of{Yes I explained the code is at the bottom of gg.R and ran code without 2022: for 0.1 the
+probability is $1.97\cdot10^{-3}$ versus $2.76\cdot10^{-3}$. For 0.09: the
+probability is $3.53\cdot10^{-4}$ versus $4.95\cdot10^{-4}$. For 0.08: the
+probability is $4.93\cdot10^{-5}$ versus $6.84\cdot10^{-5}$. I assume we want
+to report these in the supplemental?}
 

### 2024-11-25 Code/gg.R
median calculation

@@ -119,3 +119,7 @@ mean(samplefit_nodq < .09) # sim_time(gg3b, 0.09)
 mean(samplefit_nodq < .08)
 
 ## Values are slighly smaller but are actually very similar to results without 2022
+
+median(filter(dhm, Venue == 2022)$RxnTime)
+median(filter(filter(dhm, Event == "100 Dash"), Venue == 2022)$RxnTime)
+

### 2024-11-25 Code/gamlss.R
Added women's analysis to bottom

@@ -193,3 +193,60 @@ sim_time(gg3b_no2022, 0.09)
 sim_time(gg3b_no2022, 0.08)
 sim_prob(gg3b_no2022, .001)
 sim_prob(gg3b_no2022, .0001)
+
+##Pool men and women's data?
+##Had a very difficult time fitting the data regularly. On the gg3b_w model I
+## kept getting an error: Error in if (dv > olddv && itn >= 2 && auto == TRUE) { : 
+## missing value where TRUE/FALSE needed.  I manually inspected the data and tried
+## varying datasets with really random results.  As a result we will use a no DQ
+## dataset as that seemed to work better.
+dhmw <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Time != "DNS", Time != "DQ", Time != "D") |>
+  filter(RxnTime > 0) |>
+  filter(Event != "200 Dash") |>
+  mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))
+  
+gg3b_gender  <- gamlss(RxnTime ~ random(Venue) + Gender, sigma.formula = ~ random(Heat), data = dhmw, family = GG, control = gamlss.control(n.cyc = 40))
+gg3b_w  <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhmw, family = GG, control = gamlss.control(n.cyc = 40))
+
+
+logLik_1 <- logLik(gg3b_w)
+logLik_2 <- logLik(gg3b_gender)
+LRT <- 2 * (logLik_2 - logLik_1)
+df_diff <- attr(logLik_2, "df") - attr(logLik_1, "df")
+p_value <- pchisq(LRT, df = df_diff, lower.tail = FALSE)
+cat("LRT:", LRT, "\nDegrees of Freedom:", df_diff, "\nP-value:", p_value, "\n")
+
+dhmw$Gender <- as.factor(dhmw$Gender)
+t.test(RxnTime ~ Gender, data = dhmw)
+
+
+##Creates a bunch of graphs to look at differences in men and women
+df <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime > 0) |>
+  filter(Event != "200 Dash") |>
+  mutate(Venue = as.factor(Venue))
+df %>%
+  split(.$Venue) %>%
+  lapply(function(sub_df) {
[…]

### 2024-11-25 Archive/Plots.Rmd
No changes

@@
 library(gridExtra)
-rank_scatter <- grid.arrange(men_19, women_19, men_22,women_22, men_23, women_23, ncol = 2)
+rank_scatter <- grid.arrange(men_22, women_22, men_19,women_19, men_23, women_23, ncol = 2)
 ggsave(filename = "RankScatterPlots.pdf", path = "../Manuscript", plot=rank_scatter, height = 7.5, width = 7)

### 2024-11-25 Manuscript/manuscript.tex
Changed one comment
removed addressed comments; cleaning/edits

--- 2024-11-24 Manuscript/manuscript.tex
@@
 \begin{document}
-%\maketitle
+
+\title{\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships}
 
@@
 {
-  \title{\bf On Devon Allen's Disqualification at the 2022 World Track and Field
-Championships}
   \author{Owen Fiore, %\\
@@
 }
-\date{}
-  \maketitle
 } \fi
@@
   \bigskip
-  \begin{center}
-  {\LARGE\bf On Devon Allen's Disqualification at the 2022 World Track and Field
-Championships}
-\end{center}
+  \author{Anonymous Authors}
   \bigskip
 } \fi
+
+\maketitle
 
@@
 compared to shorter distances where explosive starts are paramount.
+Negative RTs were excluded from the analysis, while positive
+disqualified RTs were included due to their low frequency. A
+sensitivity analysis regarding the inclusion of positive disqualified
+RTs is provided in the Supplementary Material.
 
@@
 investigation began shortly after the 2022 World Championships, and we
-are pleased to include data from 2023, which did not significantly
+were pleased that including data from 2023 did not significantly
 alter our results \citep{WAData}.
@@
 
-For assessing the abnormality of the 2022 World Championship,
-the motivation for including data beyond the World Championships arose from
-an exploratory analysis of how United States athletes performed at the 2022
-United States Track and Field Championships, held from June 23--26, 2022, at
-Hayward Field in Eugene, Oregon. Since the same venue hosted the 2022 World
-Championships in August, this provided a unique opportunity to establish a
-baseline for RTs.
-
 \subsubsection{2022 National Competitions}
 \label{sec:datanational}
+\jy{Owen, are there names from 100-m dash?}
+\of{Yes, there were four US athletes in the 100m dash that we have data on.
+Marvin Bracy, Fred Kerley, Travyon Bromwell, and Christian Coleman.  All of whom
+recorded their slowest times at the US Track and Field Championship.}
 
-\jy{Owen, are there names from 100-m dash?}
-To compare RTs at the 2022 World Championships with those
-recorded at national competitions, we constructed a dataset of athletes
-who competed in both settings. Initially, we examined data from four
-United States athletes---Trey Cunningham, Daniel Roberts, Grant
-Holloway, and Devon Allen---who participated in both the United States
-Track and Field Championships and the World Championships. Each athlete
-recorded faster RTs in all World Championships races compared
-to their performances at the national-level event. However, this small
-sample size was insufficient for meaningful statistical analysis. To
-expand the dataset, we included RTs from additional national
-competitions held between May and July 2022 across various countries.
-Compiling this data was challenging, as RTs were not centrally
-located and often required searching country-specific websites, with many
-results recorded in native languages.
+To assess the abnormality of the 2022 World Championships, we examined
+data beyond World Championship events. This decision was motivated by
+an exploratory analysis of how United States (US) athletes performed
+at the 2022 US Track and Field Championships, held from June 23--26,
+2022, at Hayward Field in Eugene, Oregon. Since this venue also hosted
+the 2022 World Championships in August, it
+provided a unique opportunity to compare. There were four US
+110-meter hurdle athletes—Trey Cunningham, Daniel Roberts,
+Grant Holloway, and Devon Allen—who competed in both events. All four
+athletes recorded faster RTs in every World Championships race
+compared to their performances at the national-level event. Similarly,
+all four US 100-meter dash athletes---Marvin Bracy, Fred Kerley,
+Travyon Bromwell, and Christian Coleman---also recorded slower RTs at
+the US Championships than at the World Championships.
+To expand the dataset, we included RTs from 100-meter hurdle
+and 100-meter dash athletes who competed in other national
+competitions held between May and July 2022 across various
+countries. Compiling this data presented challenges, as RTs were not
+centrally archived and often required searching country-specific
+websites, with many results recorded in native languages.
 
@@
 
-\jy{Could you do a pooled analysis since there is no gender difference
-  and event differece in the supplement?}
-
 \begin{figure}[tbp]
@@
   \caption{RTs for athletes who competed at the 2022 World Championships
-	and at another championship (2022 national, 2019 World, or 2023 World) at which
-	they competed. On the horizontal
-  axis below each graph "H", "S", and "F" refer to the heats, semifinals, and
-  finals respectively. Please note that in the last row the 2022 times are to
-  the right of the 2023 times.
-	\eds{Please check my edited caption; also, I suggest rearranging the rows to
-	use the same ordering as the data was described: put national vs international
-	in first row, 2019 vs 2022 in second row, and 2022 vs 2023 in third row}
-	}
+    and at another championship (2022 national, 2019 World, or 2023
+    World) at which they competed. On the horizontal
+    axis below each graph ``H'', ``S'', and ``F'' refer to the heats,
+    semifinals, and  finals respectively. Please note that in the last
+    row the 2022 times are to the right of the 2023 times.}
   \label{fig:RankScatterplots}
@@
 \jy{Let's put the pooled analysis in the supplement.}
+\of{Please see my email}
 
@@
 
-\jy{have the numbers in this paragraph been updated with the
-  additional 100-dash data?}
 The data is summarized in Figure~\ref{fig:Boxplot}, which presents a
@@
 $v_i$ is normally distributed with mean zero and
-variance~$\sigma_v^2$, and $h_{i/j}$ is normally distributed with mean
-zero and variance~$\sigma_h^2$.
+variance~$\tau_v^2$, and $h_{i/j}$ is normally distributed with mean
+zero and variance~$\tau_h^2$.
 The two random effects were found useful: one capturing the venue effect, which
@@
    \toprule
-   Comparison & Permutation & Asymptotic & \# of athletes & \# of observations  \\
+   Comparison & Permutation & Asymptotic & \# of athletes & \# RTs  \\
    \midrule
@@
 
-\jy{Please add standard errors, which could be from \texttt{summary(gg3b)}}
-
-\jy{Are these numbers correct? I modified with the gg3b in gg.R}
+The fitted parameters of the GG distribution int he
+GAMLSS framework in Equations~\eqref{eq:gg}--\eqref{eq:sigma} are
+summarized in Table~\ref{tab:ggfit}. Results obtained from both
+excluding and including 2022 data are reported. The fixed-effect
+parameters include $\beta_0$, $\gamma_0$, and $\nu$, corresponding to
+the intercept of the log-location, log-scale, and shape of the GG
+distribution, respectively. Random effects account for variability at
+the venue level $\tau_v$ on the log-scale of the $\mu$ parameter
+and at the heat level $\tau_h$ on the log-scale of the $\sigma$
+parameter in the density in Equation~\eqref{eq:gg}. The variance
+of the venue-level random effect is smaller than the heat-level random
+effect variance, suggesting that heat-level variability in the scale
+parameter is substantial, though on the dispersion parameter. When the
+2022 data is included, all parameter estimates remain stable except
+the standard deviation of the venue-level random-effect variance,
+which increases from 0.043 to 0.058 . These results
+highlight that RTs are influenced by both venue and heat-level
+factors, and that the inclusion of 2022 introduces greater venue-level
+variability, likely due to systematic differences in RTs that year.
 
@@
     \toprule
-    Data set & $\beta_0$ & $\gamma_0$ & $\nu$ & $\sigma_v$ & $\sigma_h$ \\
+    Data set & $\beta_0$ & $\gamma_0$ & $\nu$ & $\tau_v$ & $\tau_h$ \\
     \midrule
     Excluding 2022 & $-$1.910 (0.005) & $-$2.200 (0.025) & $-$1.177 (0.442) & 0.043 & 0.326 \\
-    Including 2022 & $-$1.910 & $-$2.200 & $-$1.178 & 0.058 & 0.320 \\
+    Including 2022 & $-$1.910 (0.005) & $-$2.200 (0.027) & $-$1.178 (0.447) & 0.058 & 0.320 \\
     \bottomrule
@@
 \end{table}
-
-The fitted parameters of the GG distribution int he
-GAMLSS framework in Equations~\eqref{eq:gg}--\eqref{eq:sigma} are
-summarized in Table~\ref{tab:ggfit}. Results obtained from both
-excluding and including 2022 data are reported. The fixed-effect
-parameters include $\beta_0$, $\gamma_0$, and $\nu$, corresponding to
-the intercept of the log-location, log-scale, and shape of the GG
-distribution, respectively. Random effects account for variability at
-the venue level $\sigma_v$ on the log-scale of the $\mu$ parameter
-and at the heat level $\sigma_h$ on the log-scale of the $\sigma$
-parameter in the density in Equation~\eqref{eq:gg}. The variance
-of the venue-level random effect is smaller than the heat-level random
-effect variance, suggesting that heat-level variability in the scale
-parameter is substantial, though on the dispersion parameter. When the
-2022 data is included, the fixed-effect estimates remain stable,
-indicating robustness, but the venue-level random-effect variance
-increases to 0.0576, reflecting additional variability in the location
-parameter, while the heat-level random-effect variance remains stable.
-decreases slightly to 0.3198. These results
-highlight that RTs are influenced by both venue and heat-level
-factors, with heat-level variability playing a more prominent role in the scale
-parameter, and that the inclusion of 2022 introduces greater venue-level
-variability, likely due to systematic differences in RTs that year.
 
@@
 a reasonable representation of the observed RT data.
-
-% \begin{figure}[tbp]
-%   \centering
-%   \includegraphics{ComparisonOfVenueEffects}
-%   \caption{The venue effects from 1999 to 2023 estimated from the GLMM for the
-%     data including 2022 (top) and excluding 2022 (bottom).}
-%   \label{fig:VenueEffects}
-% \end{figure}
-
-% The uniqueness of 2022 is further demonstrated by Figure~\ref{fig:VenueEffects},
-% which shows the values of the venue random effects.  The temporal patterns
-% of the two sets of effects look very similar, except that the positive effects
-% estimated with 2022 included have higher magnitude than those estimated without
-% 2022, which compensates the negative effect of the 2022 with a large magnitude.
-% Both plots agree to some extent with the pattern of the boxplots in
-% Figure~\ref{fig:Boxplot}. The mismatch is due to the heat effect, since the venue
-% effect only captures a smaller part of the variation than the heat effect as
-% evident from their standard deviations. It is possible to calculate the extremity
-% of the 2022 heat effect as we know that the mean of the heat effects is $0$,
-% the standard deviation is $0.42$ as given by Table~\ref{tab:Gamma_parameters},
-% and the estimate of the 2022 heat effect is $-0.93$. We find the tail probability
-% to be $0.0128$ indicating there is a low chance of observing a venue effect as
-% extreme as 2022.
-
-\jy{Could you also test sensitivity for excluding the positive but
-  fault RTs?}
 
@@
 
-\bibliographystyle{chicago}
+\bibliographystyle{apalike}
 \bibliography{citations}

### 2024-11-25 Manuscript/supp.tex
added supplement template

--- 2024-11-25 Manuscript/manuscript.tex
@@
 
-\title{\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships}
+\title{\bf Supplement to ``On Devon Allen's Disqualification at the 2022 World Track and Field Championships''}
 
@@
 
-\doublespace
+\section{Rank-Based Comparison with Pooled Men and Women Data}
 
-% Abstract should be under 200 words
+\section{GAMLSS Results for Women Data}
 
-\begin{abstract}
-Devon Allen’s disqualification at the men's 110-meter hurdle final at
-the 2022 World Track and Field Championships, 
-due to a reaction time (RT) of 0.099 seconds---just 0.001 seconds below
-the allowable threshold---sparked widespread debate over 
-the fairness and validity of RT rules. This study investigates two key
-issues: variations in timing systems and the justification for the
-0.1-second disqualification threshold. We
-pooled RT data from men’s 110-meter hurdles and 100-meter dash, as
-well as women’s 100-meter hurdles and 100-meter dash, spanning
-national and international competitions. Using a rank-sum test for
-clustered data, we compared RTs across multiple competitions,
-while a generalized Gamma model with random effects for venue and heat
-was applied to evaluate the threshold. Our analyses reveal significant
-differences in RTs between the 2022 World Championships and other
-competitions, pointing to systematic variations in timing
-systems. Additionally, the model shows that RTs below 0.1 seconds,
-though rare, are physiologically plausible. These findings highlight
-the need for standardized timing protocols and a re-evaluation of the
-0.1-second disqualification threshold to promote fairness in
-elite competition.
-
-\bigskip\noindent{\sc Keywords}:
-false start, GAMLSS, reaction time, rank-based test, short sprint
-\end{abstract}
-
-\doublespace
-
-\section{Introduction}
-\label{sec:intro}
-
-Devon Allen’s highly anticipated performance at the 2022 World
-Track and Field Championships in Eugene, Oregon, ended in
-controversy when he was disqualified for a reaction time (RT) of
-0.099 seconds, just 0.001 seconds below the allowable threshold.
-Allen, a University of Oregon alumnus,
-had recently run a time of 12.84 seconds in the 110-meter hurdle
-event, just 0.04 seconds short of the world record. After placing
-third at the U.S. Track and Field Championships, he advanced
-through the preliminary heats and semifinals at the World
-Championships, with RTs of 0.123 and 0.101 seconds,
-respectively. However, in the final heat, competing in front of his
-home audience, Allen’s RT was just 0.001 seconds faster
-than the 0.1-second threshold set by the International Association
-of Athletics Federations (IAAF), resulting in his disqualification, a
-decision that was met with widespread public outcry. This 
-incident highlighted two long-standing issues: variability in the
-measurement of RTs by Start Information Systems (SIS) and the
-appropriateness of the 0.1-second disqualification threshold. As
-reaction times are measured in fractions of a second, inconsistencies
-in timing technologies and rules can significantly affect athlete
-outcomes, raising questions about fairness and standardization.
-
-World Athletics (formerly IAAF) uses certified SIS to measure RTs, yet
-variation in technology persist. Discussions at online forums such as
-\url{www.LetsRun.com} questioned consistencies in the SIS as a
-contributing factor to RT anormalies \citep{johnson2022data,
-  johnson2022was}. Historically, ``loud gun'' systems caused signal delays
-for athletes in outer lanes  due to the speed of sound, an issue
-addressed with the introduction of ``silent gun'' systems in 2010,
-which electronically synchronize sound delivery to all athletes
-\citep{tonnessen2013reaction}. Despite these advances, variability
-persists due to differences in sensor technologies, such as force
-transducers and accelerometers, and inconsistencies in event detection
-algorithms \citep{willwacher2013novel}. For example, simple
-force-threshold systems may delay RT detection by up to 26~ms compared
-to more sophisticated methods \citep{pain2007sprint}. These findings
-emphasize the need for standardized certification protocols to reduce
-discrepancies and ensure fairness in RT measurements, as recently
-reviewed by \citet{milloz2021sprint}.
-
-The 0.1-second disqualification threshold, introduced in the 1990s,
-has been the subject of significant debate. Partly based on limited
-data from Finnish national-level athletes~\cite{mero1990reaction}, the
-threshold may not adequately represent the capabilities of elite
-sprinters. Controlled experiments have shown that reaction times below
-0.1 seconds are physiologically plausible \citep{pain2007sprint,
-  komi2009iaaf}, while retrospective analyses of competition data
-often advocate for raising the threshold~\cite{brosnan2017effects,
-  lipps2011implications}. Stricter false-start rules, introduced to
-minimize race disruptions, have also discouraged sprinters from
-attempting faster starts, which may artificially inflate RTs recorded
-in competition~\cite{haugen2013effect}. This divergence between
-experimental findings and competition-based analyses illustrates the
-complexity of defining a universally fair threshold. Addressing these
-debates requires modern data collection and advanced methodologies to
-ensure equity and consistency in elite
-competition~\cite{milloz2021sprint}.
-
-This paper addresses two primary objectives from a statistical
-perspective, using modern methodologies to analyze historical
-data. First, we investigate whether reaction times (RTs) at the 2022
-World Championships were significantly different from other
-competitions, focusing on athletes who competed in multiple
-events. Using a matched-pairs design, we compare RTs across the 2022,
-2019, and 2023 World Championships, as well as the 2022 national-level
-competitions. This approach isolates the effect of the competition
-year while controlling for individual performance. The clustered
-nature of the data, with the goal of assessing differences across
-events within athletes, was analyzed using a rank-based comparison
-approach for clustered data \citep{datta2005rank}. Second, we evaluate
-the appropriateness of the 0.1-second disqualification threshold by
-modeling RTs from World Championships held from 1999 onward. A
-generalized Gamma (GG) distribution with random effects for both venue and
-heat was applied within the framework of the generalized additive
-model for location, scale, and shape (GAMLSS)
-\citep{rigby2005generalized, stasinopoulos2024generalized}.
-This model enables estimation of the
-probability of RTs falling below the threshold, providing a
-statistical assessment of RT consistency and the validity of the
-current threshold.
-
-The rest of this paper is organized as follows. Section~\ref{sec:Data}
-describes the data collection and datasets used for the two
-objectives. Section~\ref{sec:rank} outlines the rank-based comparison
-methods for objective~1, rank-based comparison for clustered data with
-subunit-level gruping to analyze differences in RTs across different
-competitions. Section~\ref{sec:gamlss} details the methods for
-objective~2, the GG model with random effects to evaluate the appropriatenes
-of the 0.1-second threshold. Section~\ref{sec:Results} presents the
-results of both objectives and discusses the implications.
-Finally, Section~\ref{sec:concludingremarks} highlights the
-paper’s impact, limitations, and potential for future research.
-All data and code for our analysis are provided in the supplementary
-materials.
-
-\section{Data}
-\label{sec:Data}
-
-This study focuses on RTs from sprint events where reaction
-time plays a critical role. Specifically, we included data from men’s
-100-meter dashes and 110-meter hurdles, as well as women’s 100-meter
-dashes and 100-meter hurdles. Statistical tests showed no significant
-differences in RTs between these events, supporting their
-inclusion in a unified analysis. However, we excluded data from
-200-meter dashes and longer events, as their RTs were found
-to be significantly different. This difference is not unexpected, as
-RT is generally less critical in longer sprint events
-compared to shorter distances where explosive starts are paramount.
-Negative RTs were excluded from the analysis, while positive
-disqualified RTs were included due to their low frequency. A
-sensitivity analysis regarding the inclusion of positive disqualified
-RTs is provided in the Supplementary Material.
-
-Two datasets were used to address the study's objectives. To
-investigate whether RTs at the 2022 World Championships
-were significantly different from other competitions (Objective 1), we
-used data from male athletes in the 110-meter hurdles and 100-meter
-dash, and female athletes in the 100-meter hurdles and 100-meter dash,
-provided they competed in the 2022 World Championships and at least
-one other competition (2022 national championships, 2019 World
-Championships, or 2023 World Championships). To assess the
-appropriateness of the 0.1-second RT threshold (Objective
-2), we used RTs from the 110-meter hurdles and 100-meter
-dash across all World Championships held from 1999 to 2023. This
-investigation began shortly after the 2022 World Championships, and we
-were pleased that including data from 2023 did not significantly
-alter our results \citep{WAData}.
-
-\subsection{Comparison Data with 2022 World Championships}
-\label{sec:databeyond}
-
-\subsubsection{2022 National Competitions}
-\label{sec:datanational}
-\jy{Owen, are there names from 100-m dash?}
-\of{Yes, there were four US athletes in the 100m dash that we have data on.
-Marvin Bracy, Fred Kerley, Travyon Bromwell, and Christian Coleman.  All of whom
-recorded their slowest times at the US Track and Field Championship.}
-
-To assess the abnormality of the 2022 World Championships, we examined
-data beyond World Championship events. This decision was motivated by
-an exploratory analysis of how United States (US) athletes performed
-at the 2022 US Track and Field Championships, held from June 23--26,
-2022, at Hayward Field in Eugene, Oregon. Since this venue also hosted
-the 2022 World Championships in August, it
-provided a unique opportunity to compare. There were four US
-110-meter hurdle athletes—Trey Cunningham, Daniel Roberts,
-Grant Holloway, and Devon Allen—who competed in both events. All four
-athletes recorded faster RTs in every World Championships race
-compared to their performances at the national-level event. Similarly,
-all four US 100-meter dash athletes---Marvin Bracy, Fred Kerley,
-Travyon Bromwell, and Christian Coleman---also recorded slower RTs at
-the US Championships than at the World Championships.
-To expand the dataset, we included RTs from 100-meter hurdle
-and 100-meter dash athletes who competed in other national
-competitions held between May and July 2022 across various
-countries. Compiling this data presented challenges, as RTs were not
-centrally archived and often required searching country-specific
-websites, with many results recorded in native languages.
-
-The final dataset consisted of RTs from athletes who competed
-in both national and international competitions. RTs from
-preliminary heats, semifinals, and finals were included to ensure
-sufficient data for analysis. Excluding preliminary heats would have
-significantly reduced the number of athletes and clusters. Each
-athlete was considered as a cluster, with observed RTs
-from both the `treatment' group (2022 World Championships) and
-the `control' group (national competition) within the cluster. Cluster
-sizes ranged from three to six, with a median size of four. Because gender
-is known to influence RTs \citep{babicc2009reaction,
-  lipps2011implications}, we prepared data for men and women
-separately, resulting in 80 RTs from 17 athletes for each
-gender. This setup forms the basis for the rank-based comparison for
-clustered data with subunit-level grouping, which accounts for the
-within-athlete dependence inherent in this data structure. The top
-panel of Figure!~\ref{fig:RankScatterplots} shows the RTs
-of those who competed both at national competitions and the 2022 World
-Championships.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\textwidth]{RankScatterPlots}
-  \caption{RTs for athletes who competed at the 2022 World Championships
-    and at another championship (2022 national, 2019 World, or 2023
-    World) at which they competed. On the horizontal
-    axis below each graph ``H'', ``S'', and ``F'' refer to the heats,
-    semifinals, and  finals respectively. Please note that in the last
-    row the 2022 times are to the right of the 2023 times.}
-  \label{fig:RankScatterplots}
-\end{figure}
-
-\subsubsection{2019 and 2023 World Championships}
-\label{sec:data2019}
-
-To compare RTs at the 2022 World Championships with those
-recorded in the 2019 and 2023 World Championships, we prepared datasets
-including RTs from athletes who competed in the 2022 World
-Championships and at least one of the 2019 or 2023 World Championships.
-Specifically, we identified athletes who participated in both the 2019
-and 2022 World Championships, as well as those who participated in both
-the 2022 and 2023 World Championships. In the 2019-2022 comparison,
-RTs from 2022 were treated as the `treatment' group, with
-2019 serving as the `control' group. Similarly, in the 2022-2023
-comparison, RTs from 2022 were treated as the `treatment'
-group, with 2023 serving as the `control' group. This structure allowed
-us to prepare datasets suitable for examining RTs of athletes
-who competed across multiple World Championships.
-
-Each athlete was treated as a single cluster, containing their reaction
-times from different World Championships. The dataset for the 2019-2022
-comparison contained 134 RTs from 34 male athletes and 124
-RTs from 31 female athletes. The dataset for the 2022-2023
-comparison contained 161 RTs from 45 male athletes and 182
-RTs from 47 female athletes. While it is theoretically
-possible that athletes improved their RTs between 2019 and
-2022 or between 2022 and 2023, such improvements are highly unlikely for
-elite sprinters, as they already operate near the limits of human
-performance. Consequently, consistent improvements observed in 2022
-would suggest systematic differences rather than natural variability.
-
-Figure~\ref{fig:RankScatterplots} shows the RTs of athletes
-who competed in both the 2019 and 2022 World Championships (middle
-panel) and those who competed in both the 2022 and 2023 World
-Championships (lower panel). The athletes included in the 2022 data
-differ between these two comparisons, as the set of athletes who
-competed in both 2019 and 2022 is not the same as the set who competed
-in both 2022 and 2023. To be included, athletes must have competed in
-at least one race at each championship. Notably, Devon Allen recorded
-the fastest RTs in both the Finals and Semifinals of the
-2022 World Championships, but his disqualification was determined by a
-difference of just 0.002 seconds, with RTs of 0.101 and
-0.099 seconds, respectively. This highlights the critical role of
-RT precision in elite-level competition.
-
-\eds{I also think we need to comment on the men and women's data. But within row
-of the figure, don't the distributions appear similar?  If that is so, would a
-reviewer ask why we didn't pool the data for men and women?}
-\jy{Let's put the pooled analysis in the supplement.}
-\of{Please see my email}
-
-\subsection{Data from World Championships 1999--2023}
-\label{sec:dataworld}
-
-The data for evaluating the appropriateness of the 0.1-second
-threshold was obtained from World Athletics and covers the men's
-110-meter hurdles and 100-meter dashes from 1999 to 2023. Due to
-possible gender differences \citep{babicc2009reaction,
-  lipps2011implications}, data for women's 100-meter hurdles and
-100-meter dashes and their analyses were relegated to the
-supplementary material. We focus on the RTs recorded during
-semifinal and final heats only, as RTs from preliminary
-heats are often not as fast as those in later heats
-\citep[e.g.,][]{collet1999strategic, tonnessen2013reaction,
-  brosnan2017effects, zhang2021correlation}. For analysis purposes, we
-pooled RTs from semifinal and final heats to increase sample
-size, which is particularly important for years with limited final heat
-observations. For example, in 2022, only five data points were available
-from the final heat due to two disqualifications and one athlete not
-competing. Unless otherwise noted, this pooled dataset forms the basis
-for our analysis throughout the paper. Additionally, we consider datasets
-that exclude 2022 to assess how our findings might differ when excluding
-this year of interest.
-
-The data is summarized in Figure~\ref{fig:Boxplot}, which presents a
-sequence of boxplots of RTs from 1999 to 2023. It is evident
-that RTs in 2022 were notably faster, with a median reaction
-time of 0.129 seconds compared to the 0.156 seconds observed in earlier
-studies, such as \citet{brosnan2017effects} for data spanning 1999 to 2014.
-Figure~\ref{fig:Boxplot} also highlights year-to-year variability in
-RTs, likely influenced by changes in the championship venue
-and environmental conditions such as humidity, precipitation, and
-elevation. Furthermore, advancements in technology and alterations to
-false start rules during the study period may have played a role in these
-variations \citep{willwacher2013novel}.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\textwidth]{Boxplot}
-  \caption{The RTs from 1999 to 2023 for the men's 110 meter hurdle
-  and 100 meter dash.}
-  \label{fig:Boxplot}
-\end{figure}
-
-Between 2007 and 2009, World Athletics allowed one
-false start warning before disqualifying a sprinter \citep{iaaf2009falsestart}.
-This lenient rule led to 18 male and 7 female false starts at both the 2007
-and 2009 World Championships. In 2011, this rule was replaced with the stricter
-policy of automatic disqualification for false starts, aimed at reducing the
-delays caused by repeated warnings. This change reduced men’s false starts
-by two-thirds in 2011, with only six male and four female disqualifications
-\citep{iaaf2009falsestart}. \citet{haugen2013effect} demonstrated that more
-lenient false start rules significantly improved RTs during the
-1997–2009 period, suggesting that rule changes over the study period may
-have contributed to variations in RTs across years.
-
-\section{Methods} \label{sec:methods}
-
-The data described in Sections~\ref{sec:databeyond} and~\ref{sec:dataworld} were
-analyzed with rank-based methods and a GAMLSS, respectively.
-
-\subsection{Rank-based Comparison}\label{sec:rank}
-
-To test the conjecture that the 2022 World Championships timing device may have
-led to faster recorded RTs, we compare the RTs of the same
-athletes who have attended both the 2022 World Championships and other
-competitions.
-In this setting, we have clustered data with subunit grouping. In particular,
-each athlete is a cluster and the multiple RTs from the same athlete
-can be from either the 2022 World Championships or otherwise.
-Let $X_{ij}$ be the $j$th RT of athlete~$i$, $i = 1, \ldots, n$,
-$j = 1, \ldots, m_i$ where $m_i$ is the number of observations from
-athlete~$i$. Let $\delta_{ij}$ be the group indicator of $X_{ij}$; $\delta_{ij}
-= 1$ if $X_{ij}$ is in group~1 (2022 World Championships) and $\delta_{ij} = 0$
-otherwise. Athletes are
-assumed to be independent, while subunit observations from the same athlete are
-not. The null hypothesis $H_0$ to be tested is that there is no difference
-between the two groups; i.e., the distribution of $X_{ij}$ remains the same
-regardless of the group indicator $\delta_{ij}$.
-
-\citet{datta2005rank} proposed an extension of the Wilcoxon rank-sum test to
-clustered data with subunit-level grouping. The test is designed based on a
-within-cluster resampling principle. Consider randomly picking one observation
-from each cluster to form a pseudo-sample. Let $X_i^*$ be a random pick from the
-$i$th cluster in the pseudo-sample and $\delta_i^*$ its group indicator. The
-Wilcoxon rank-sum statistic for the pseudo-sample is
-\[
-W^* = \frac{1}{n + 1} + \sum_{i=1}^{n} \delta_{i}^{*} R_{i}^{*},
-\]
-where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the pseudo-sample.
-The test statistic $S$ is the average of $W^*$ averaged over all possible
-pseudo-samples conditioning on the observed data and group indicators.
-The mean and variance of $S$ under $H_0$ can be derived so that $S$ can be
-standardized to form a $Z$ statistic which follows a standard normal distribution
-asymptotically \citep[p.910]{datta2005rank}.
-
-When the sample size is small, the asymptotic normal distribution may
-not be reliable, so we also use 1~million random permutations to
-simulate the null distribution of the test statistic.
-This method is available from the \texttt{clusWilcox.test()} function
-with \texttt{method = `ds'} (for \underline{D}atta and \underline{S}atten) and
-\texttt{exact = TRUE} from R package
-\texttt{clusrank} \citep{jiang2020wilcoxon}.
-
-\subsection{GAMLSS}\label{sec:gamlss}
-
-Based on an exploratory analysis, the RTs are adequately
-modeled by a GG distribution with random effects in
-model parameters. The GG distribution has three parameters, denoted by
-$\text{GG}(\mu, \sigma, \nu)$ has density function
-\begin{equation}
-  \label{eq:gg}
-f_Y(y \mid \mu, \sigma, \nu) =
-\frac{|\nu| \theta^\theta z^{\theta}}{\Gamma(\theta) y}
-\exp\left(-z \theta\right),
-\end{equation}
-for $y > 0$, $\mu > 0$, $\sigma > 0$, and $\nu \neq 0$,
-where $z = (y / \mu)^\nu$,
-$\theta = 1 / (\sigma^2 \nu^2)$, and
-$\Gamma(\cdot)$ denotes the Gamma function.
-The GG distribution is highly flexible, encompassing several
-well-known distributions as special cases, such as the
-Weibull ($\mu = \nu)$ and  Gamma $(\nu = 1)$ distributions.
-Its expectation is
-\[
-  \frac{\mu \Gamma(\theta + 1 / \nu)}
-  {\theta^{1 / \nu} \Gamma(\theta)},
-\]
-provided $\theta > -1 / \nu$. Here,
-$\mu$ scales the central tendency, $\sigma$ controls
-dispersion, and $\nu$ determines skewness. This parameterization allows
-the distribution to model asymmetric and heavy-tailed data effectively, making
-it particularly suitable for RTs.
-An implementation of this distribution is available from R package
-\texttt{gamlss.dist} \citep{rigby2019distributions}.
-
-Let $Y_{ijk}$ denote the RT of observation~$k$ in heat~$j$
-of year~$i$. Conditioning on a venue effect $v_i$ for year~$i$
-and a heat effect $h_{i/j}$ nested within each year~$i$, the
-distribution of $Y_{ijk}$ is
-$\text{GG}(\mu_{ijk}, \sigma_{ijk}, \nu)$, where
-\begin{align}
-\log(\mu_{ijk}) &= \beta_0 + v_i , \label{eq:mu}\\
-\log(\sigma_{ijk}) &= \gamma_0 + h_{i/j} , \label{eq:sigma}
-\end{align}
-$v_i$ is normally distributed with mean zero and
-variance~$\tau_v^2$, and $h_{i/j}$ is normally distributed with mean
-zero and variance~$\tau_h^2$.
-The two random effects were found useful: one capturing the venue effect, which
-is used to contrast years, and the second being the heat effect, where every
-race was given a unique identifier with typically five to nine observations
-per race. The heat effect is important as it captures the variability in the
-amount of time athletes are on the starting blocks before the gun goes off.
-This model can be fit with R package \texttt{gamlss}
-\citep{stasinopoulos2008generalized}.
-
-Model diagnosis and tail analysis can be done with the fitted GG model
-from package \texttt{gamlss}. Normalized quantile residuals, or
-z-scores \citep{dunn1996randomized}, of the observations can be
-extracted with the \texttt{residuals} method of a \texttt{gamlss}
-object. The z-scores can then be checked with a Q-Q plot
-\citep{almeida2018ggplot2}. The marginal
-distribution of $Y_{ijk}$ is a scale-mixture of GG distributions, which can be
-easily simulated from once the parameters are estimated. Many
-random numbers generated from the fitted mixture distribution can be used to
-approximate the probability of observing a RT faster than any given
-threshold. We are specifically interested in the probability of a RT
-being less than 0.1 seconds in order to gauge if that is a reasonable
-disqualification barrier.
-
-\section{Results} \label{sec:Results}
-
-\subsection{Rank-based Comparison} \label{subsec:Results_Rank}
-
-The rank-based methods described in Section~\ref{sec:rank} were used to
-compare RTs between the 2022 World Championships and other
-competitions in which the same athletes participated. These comparisons
-were conducted separately for men and women, resulting in six total
-comparisons: RTs from the 2022 national-level championships
-versus the 2022 World Championships for men and women, RTs
-from the 2019 versus 2022 World Championships for men and women, and
-RTs from the 2022 versus 2023 World Championships for men and
-women.
-
-\begin{table}
-  \centering
-  \caption{P-values of comparisons between
-    RTs from different competitions for the same athletes.
-    2022 Nat. vs Inter. compares RTs from 2022 national-level
-    championships and the 2022 World Track and Field Championships. 2019
-    vs 2022 compares RTs from the 2019 and 2022 World Track and
-    Field Championships. 2022 vs 2023 compares RTs from the
-    2022 and 2023 World Track and Field Championships.}
-  \begin{tabular}{c c c c c}
-   \toprule
-   Comparison & Permutation & Asymptotic & \# of athletes & \# RTs  \\
-   \midrule
-   2022 Nat. vs Inter. Men & $1.0 \cdot 10^{-6}$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
-   2022 Nat. vs Inter. Women & $1.0 \cdot 10^{-6}$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\[1ex]
-   2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
-   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\[1ex]
-   2022 vs 2023 Men & $1.0 \cdot 10^{-6}$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
-   2022 vs 2023 Female & $1.0 \cdot 10^{-6}$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
-   \bottomrule
-  \end{tabular}
-  \label{tab:Clusrankresults}
-\end{table}
-
-Table~\ref{tab:Clusrankresults} presents the results from both permutation
-and asymptotic rank-based tests for these six comparisons. The tests reveal
-that the national versus international comparisons yielded the lowest
-p-values, indicating stronger evidence of faster RTs at the
-2022 World Championships compared to the 2022 national competitions. For
-both men and women, permutation test results were highly significant,
-providing substantial evidence that athletes competing at the 2022 World
-Championships had significantly faster RTs than they did just
-a few months earlier at national-level events. The comparisons between the
-2019 and 2022 World Championships and between the 2022 and 2023 World
-Championships also produced significant results. Collectively,
-these results suggest that athletes consistently achieved faster reaction
-times at the 2022 World Championships compared to other competitions. This
-finding supports the hypothesis that the 2022 World Championships presented
-conditions conducive to faster RTs, potentially due to systematic
-or environmental factors.
-
-\subsection{GAMLSS} \label{subsec:Results_GLMM}
-
-The fitted parameters of the GG distribution int he
-GAMLSS framework in Equations~\eqref{eq:gg}--\eqref{eq:sigma} are
-summarized in Table~\ref{tab:ggfit}. Results obtained from both
-excluding and including 2022 data are reported. The fixed-effect
-parameters include $\beta_0$, $\gamma_0$, and $\nu$, corresponding to
-the intercept of the log-location, log-scale, and shape of the GG
-distribution, respectively. Random effects account for variability at
-the venue level $\tau_v$ on the log-scale of the $\mu$ parameter
-and at the heat level $\tau_h$ on the log-scale of the $\sigma$
-parameter in the density in Equation~\eqref{eq:gg}. The variance
-of the venue-level random effect is smaller than the heat-level random
-effect variance, suggesting that heat-level variability in the scale
-parameter is substantial, though on the dispersion parameter. When the
-2022 data is included, all parameter estimates remain stable except
-the standard deviation of the venue-level random-effect variance,
-which increases from 0.043 to 0.058 . These results
-highlight that RTs are influenced by both venue and heat-level
-factors, and that the inclusion of 2022 introduces greater venue-level
-variability, likely due to systematic differences in RTs that year.
-
-\begin{table}
-  \centering
-  \caption{Estimated fixed-effect parameters with standard errors in
-    parentheses and estimated variance of the random effects from the
-    fitted GG distribution with venue level random
-    effects in $\mu$ and heat level random effects in $\sigma$ in
-    Models~\eqref{eq:gg}--\eqref{eq:sigma}.}
-  \label{tab:ggfit}
-  \begin{tabular}{c c c c c c}
-    \toprule
-    Data set & $\beta_0$ & $\gamma_0$ & $\nu$ & $\tau_v$ & $\tau_h$ \\
-    \midrule
-    Excluding 2022 & $-$1.910 (0.005) & $-$2.200 (0.025) & $-$1.177 (0.442) & 0.043 & 0.326 \\
-    Including 2022 & $-$1.910 (0.005) & $-$2.200 (0.027) & $-$1.178 (0.447) & 0.058 & 0.320 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\textwidth]{diagnosis.pdf}
-  \caption{Diagnosis of the fitted GG distribution with
-    random effects in model parameters: kernerl density of 1 million
-    observations drawn from the fitted model overlaid with the
-    histogram of the observed RTs (left); Q-Q plot of the
-    normal z-score of the quantile residuals from the fitted model.}
-  \label{fig:diagnosis}
-\end{figure}
-
-Figure~\ref{fig:diagnosis} presents diagnostic checks for the fitted
-GG distribution model with random effects. The left
-panel compares the kernel density estimate of one million simulated reaction
-times from the fitted model to the histogram of the observed RTs.
-The close alignment between the density curve and the histogram suggests that
-the fitted GG model adequately captures the overall distribution of the
-RTs. The right panel shows a Q-Q plot of the z-scores of the
-quantile residuals from the fitted model. The points lie approximately along
-the 45-degree reference line, indicating that the residuals are consistent
-with the standard normal distribution, supporting the adequacy of the model
-fit. These diagnostics collectively demonstrate that the fitted model provides
-a reasonable representation of the observed RT data.
-
-\begin{table}
-  \centering
-  \caption{Probabilities of observing RTs less than threshold 0.08,
-  0.09, and 0.10 seconds based on the two effect model.}
-  \begin{tabular}{c c c c}
-   \toprule
-   Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\
-   \midrule
-   Excluding 2022 & $5.31\cdot10^{-5}$ & $3.53\cdot10^{-4}$ &  $1.94\cdot10^{-3}$  \\
-   Including 2022 & $6.84\cdot10^{-5}$ & $4.95\cdot10^{-4}$ & $2.76\cdot10^{-3}$ \\
-   \bottomrule
-  \end{tabular}
-  \label{tab:Sim_probability}
-\end{table}
-
-The fitted GG GAMLSS model with both venue- and heat-level
-random effects provides a framework for assessing how extreme RTs
-below certain thresholds are. The probability of observing a RT
-below a given threshold, assuming no intentional false starts, was approximated
-by generating 10 million realizations from the fitted model.
-Table~\ref{tab:Sim_probability} summarizes the probabilities of observing
-RTs below 0.08, 0.09, and 0.10 seconds under two scenarios: one
-excluding and the other including data from 2022. Excluding 2022 slightly
-reduces the probability of observing a fast RT, but the difference
-is small. For example, the probability of a RT below 0.10 seconds
-decreases from $2.76 \cdot 10^{-3}$ (approximately one in 362 starts) to
-$1.94 \cdot 10^{-3}$ (approximately one in 515 starts) when 2022 is excluded.
-Lowering the RT threshold from 0.10 to 0.08 seconds drastically
-reduces the likelihood of observing a RT below the barrier, with
-the probability dropping from one in every 362 starts (at 0.10 seconds) to one
-in every 14620 starts (at 0.09 seconds) and one in every 146198 starts (at 0.08
-seconds) when 2022 is included. These results highlight the rarity of extremely
-fast RTs and substantiate the recommendations of \citet{komi2009iaaf}
-to carefully consider the selection of RT thresholds.
-
-\begin{table}
-  \centering
-  \caption{Suggested RT barriers based on tail probabilities.}
-  \begin{tabular}{c c c}
-   \toprule
-   Data Set & Tail probability  $10^{-3}$ & Tail probability $10^{-4}$ \\
-   \midrule
-   Excluding 2022 & $0.096$ & $0.083$ \\
-   Including 2022 & $0.094$ & $0.082$ \\
-   \bottomrule
-  \end{tabular}
-  \label{tab:Sim_time}
-\end{table}
-
-Utilizing the same model, we can determine suitable RT barriers
-based on the probability of observing a time below the barrier. As shown in
-Table~\ref{tab:Sim_time}, including the 2022 data suggests a RT
-barrier of 0.094 seconds to maintain a 0.1\% chance of observing an
-exceptionally fast RT, while a stricter threshold of 0.082 seconds
-is needed to limit this probability to 0.01\%. Excluding the 2022 data results
-in slightly higher thresholds of 0.096 and 0.083 seconds for the respective
-probability levels. These results indicate that while the inclusion of 2022
-data slightly reduces the recommended barrier, the magnitude of the difference
-is relatively small. This approach allows for tailoring RT
-thresholds to desired levels of false positive rates, balancing fairness and
-precision in disqualification criteria.
-
-\section{Discussion}\label{sec:concludingremarks}
-
-This study sought to answer two primary questions: were athletes who
-competed at the 2022 World Track and Field Championships faster than
-at other races, and is the 0.1-second RT barrier fair?
-Regarding the first question, our analyses suggest that athletes were
-indeed faster at the 2022 World Championships compared to comparable
-meets. Ideally, a centralized database containing RTs from
-all World Athletics-certified meets would have allowed for a more
-comprehensive analysis. Currently, however, RTs are seldom
-included in the results available on the World Athletics website. This
-limitation constrained our ability to evaluate athletes like Devon
-Allen, as data from his races leading up to the World Championships in
-2022 were not accessible. A more robust dataset would enable deeper
-insights by allowing comparisons across a larger set of races.
-
-In addressing the second question, our analyses of the GAMLSS model
-reveal that the 0.1-second RT barrier, while rare, may not
-be as extraordinary  as traditionally assumed. For men,
-Table~\ref{tab:Sim_time} shows that reaction  times below 0.1 seconds
-occur with a probability of approximately one in 362 starts  when
-including the 2022 data. Lowering the threshold to 0.08 seconds
-drastically  reduces this likelihood, supporting the idea that the
-current barrier could be adjusted to reflect more realistic
-probabilities of false starts. A similar pattern is observed for
-women, as detailed in the Supplementary Material, where reaction
-times below 0.1 seconds are exceedingly rare for the 100-meter dash
-and 100-meter  hurdles. However, the uniformity of the 0.1-second
-barrier for both men and women  is questionable, given numerous
-studies documenting gender differences in reaction  times
-\citep[e.g.,][]{lipps2011implications, babicc2009reaction,
-  panoutsakopoulos2020gender}. These studies suggest that the current
-threshold may  unfairly penalize men, for whom sub-0.1-second reaction
-times are more probable. \citet{brosnan2017effects} advocate for
-gender-specific barriers, a position that  aligns with our findings
-and highlights the importance of tailoring thresholds to  biological
-distinctions.
-
-This study provides a statistical framework to examine Devon Allen’s
-disqualification at the 2022 World Track and Field Championships,
-offering insights rather than drawing definitive conclusions about
-potential equipment malfunction. Our findings indicate that reaction
-times at the 2022 Championships were, on average, faster than at other
-competitions, as evidenced by the significant p-values in
-Table~\ref{tab:Clusrankresults}. Additionally, the GAMLSS results
-suggest that the 0.1-second barrier may not be as stringent as
-previously believed. Based on Table~\ref{tab:Sim_time}, a stricter
-threshold of 0.08 seconds could be considered, allowing athletes like
-Allen to react swiftly without undue risk of disqualification. While
-this analysis provides a rigorous statistical perspective, it does not
-consider biomechanical factors, such as individual variability in
-neuromuscular response times or the role of starting block sensors in
-detecting pressure changes, which may offer more direct evidence of
-reaction capabilities. In summary, while the results designate 2022 as
-an anomalous year, Allen’s time, despite resulting in
-disqualification, may not be categorically extreme.
-
-\section*{Supplementary Material}
-The data and R code used for the analysis are available in a compressed file for
-ease of reproducibility.
+\section{Results from Data Excluding Positive DQ RTs}
 

### 2024-11-25 Manuscript/manuscript.tex
minor edits and comments in manuscript

@@
 World Athletics (formerly IAAF) uses certified SIS to measure RTs, yet
-variation in technology persist. Discussions at online forums such as
+variation in technology persists. Discussions at online forums such as
 \url{www.LetsRun.com} questioned consistencies in the SIS as a
@@
   komi2009iaaf}, while retrospective analyses of competition data
-often advocate for raising the threshold~\cite{brosnan2017effects,
+often advocate for raising the threshold~\citep{brosnan2017effects,
   lipps2011implications}. Stricter false-start rules, introduced to
@@
 attempting faster starts, which may artificially inflate RTs recorded
-in competition~\cite{haugen2013effect}. This divergence between
+in competition~\citep{haugen2013effect}. This divergence between
 experimental findings and competition-based analyses illustrates the
@@
 ensure equity and consistency in elite
-competition~\cite{milloz2021sprint}.
+competition~\citep{milloz2021sprint}.
 
@@
 perspective, using modern methodologies to analyze historical
-data. First, we investigate whether reaction times (RTs) at the 2022
+data. First, we investigate whether RTs at the 2022
 World Championships were significantly different from other
@@
 This model enables estimation of the
-probability of RTs falling below the threshold, providing a
+probability of RTs falling below a threshold, providing a
 statistical assessment of RT consistency and the validity of the
@@
 methods for objective~1, rank-based comparison for clustered data with
-subunit-level gruping to analyze differences in RTs across different
+subunit-level grouping to analyze differences in RTs across different
 competitions. Section~\ref{sec:gamlss} details the methods for
@@
 within-athlete dependence inherent in this data structure. The top
-panel of Figure!~\ref{fig:RankScatterplots} shows the RTs
+panel of Figure~\ref{fig:RankScatterplots} shows the RTs
 of those who competed both at national competitions and the 2022 World
@@
 
-Each athlete was treated as a single cluster, containing their reaction
-times from different World Championships. The dataset for the 2019-2022
+Each athlete was treated as a single cluster, containing their RTs
+from different World Championships. The dataset for the 2019-2022
 comparison contained 134 RTs from 34 male athletes and 124
@@
 where $R_{i}^{*}$ is the rank of $X_{i}^{*}$ in the pseudo-sample.
-The test statistic $S$ is the average of $W^*$ averaged over all possible
+The test statistic $S$ is the average of $W^*$, averaged over all possible
 pseudo-samples conditioning on the observed data and group indicators.
@@
 
-The fitted parameters of the GG distribution int he
+The fitted parameters of the GG distribution in the
 GAMLSS framework in Equations~\eqref{eq:gg}--\eqref{eq:sigma} are
@@
   \caption{Diagnosis of the fitted GG distribution with
-    random effects in model parameters: kernerl density of 1 million
+    random effects in model parameters: kernel density of 1 million
     observations drawn from the fitted model overlaid with the
@@
   \caption{Probabilities of observing RTs less than threshold 0.08,
-  0.09, and 0.10 seconds based on the two effect model.}
+  0.09, and 0.10 seconds based on the
+    fitted GG GAMLSS model with both venue- and heat-level
+random effects.}
   \begin{tabular}{c c c c}
@@
 indeed faster at the 2022 World Championships compared to comparable
-meets. Ideally, a centralized database containing RTs from
+meets. 
+\eds{are we testing faster (one-sided) or different (two-sided)?}
+Ideally, a centralized database containing RTs from
 all World Athletics-certified meets would have allowed for a more
@@
 and highlights the importance of tailoring thresholds to  biological
-distinctions.
+distinctions. \eds{is this still true in the updated analyses?}
 

### 2024-11-25 Manuscript/reply.tex
smoothed discussion
removed resolved, fixed minor typos

@@
   diagnostic checks.
+\item
+  Given the expanded dataset and change to GAMLSS, the corresponding
+  sections of the paper, including data preparation, methods, results,
+  and discussions have all been rewritten at large.
 \end{enumerate}
@@
 expanded datasets.
-\eds{let's revisit this again after objective 2 analyses are set}
 
@@
 (e.g., preliminary vs. final). To clarify this, we have added
-additional explanations in the first paragraph of Section~3.2 to
+additional explanations Section~3.2 (after Equation (3)) to
 explicitly define the venue and heat effects. We avoided calling it a
@@
 to the data compared to the original Gamma GLMM.
-\eds{let's revisit this again after objective 2 analyses are set}
 
@@
 threshold lacks a foundation in formal statistical evidence.
-\eds{let's revisit this again after objective 2 analyses are set}
 
@@
 observations in the dataset.
+\eds{Currently the text in the manuscript suggests we are including a 
+sensitivity analysis in the Supplement for the analysis without the positive DQs.
+If we are indeed including that, we should mention that here too. (If not, we 
+need to remove that statement from the main manuscript.)}
 
@@
 a formal statistical analysis of his reaction times.
-To address this point, we have added a brief discussion in 
-paragraph~??? of the Discussion Section, noting the limitations in data 
+To address this point, we have added a brief discussion in the first 
+paragraph of the Discussion Section, noting the limitations in data 
 availability and the implications of this issue.
@@
 
-Thank you for pointing out the need for clearer intrepretation of
+Thank you for pointing out the need for clearer interpretation of
 Table~2.

### 2024-11-26 Archive/Plots.Rmd
added womens boxplot plot

@@
 ```
+# Supplementary Womens Boxplot
+```{r}
+library(tidyverse)
+dhw <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "F") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime > 0) |>
+  filter(Event == "100 Hurdles" | Event == "110 Hurdles")
+library(ggplot2)
+ggplot(dhw, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.05, 0.3),
+                            breaks = seq(0.05, 0.3, 0.025)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(2000,2024),
+                                              breaks = seq(2001, 2024, 2))
+ggsave(filename="WomensBoxplot.pdf", path = "../Manuscript", height = 4, width = 6)
+```
 

### 2024-11-26 Code/RankComparison.Rmd
added supplementary analysis

@@
 ```
+
+# Supplementary Analysis
+We are going to see what happens when we pool men's and women's data together.
+We sill need to separate by year but will have 3 analyses instead of 6.
+
+### 2019 vs 2022
+Data setup
+```{r}
+men_2019 <- read.csv("../Data/Clusrank2019vs2022.csv")
+colnames(men_2019) <- c("Time", "Athlete", "Year", "Event")
+
+women_2019 <- read.csv("../Data/Clusrank2019vs2022Womens.csv")
+colnames(women_2019) <- c("Time", "Athlete", "Year", "Event")
+
+women_2019$Athlete <- women_2019$Athlete + 34 #Correction for unique athlete identifier
+
+all_2019 <- rbind(men_2019, women_2019)
+```
+
+And now we run the tests
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2019,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2019,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```
+### 2022 National vs International
+
+Data setup
+```{r}
+men_2022 <- read.csv("../Data/ClusrankNatvsWorld.csv")
+colnames(men_2022) <- c("Time", "Athlete", "Year", "Event")
+
+women_2022 <- read.csv("../Data/ClusrankNatvsWorldWomens.csv")
+colnames(women_2022) <- c("Time", "Athlete", "Year", "Event")
+
+women_2022$Athlete <- women_2022$Athlete + 17 #Correction for unique athlete identifier
+
+all_2022 <- rbind(men_2022, women_2022)
+```
+
+And now we run the tests.
+
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2022,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2022,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```
+### 2022 vs 2023
+
+```{r}
+men_2023 <- read.csv("../Data/Clusrank2022vs2023.csv")
+colnames(men_2023) <- c("Time", "Athlete", "Year", "Event")
+
+women_2023 <- read.csv("../Data/Clusrank2022vs2023Womens.csv")
+colnames(women_2023) <- c("Time", "Athlete", "Year", "Event")
+
+women_2023$Athlete <- women_2023$Athlete + 45
+
+all_2023 <- rbind(men_2023, women_2023)
+```
+
+Now we run the tests.
+```{r}
+set.seed(1)
+library(clusrank)
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2023,
+                method = "ds")
+
+clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2023,
+                method = "ds",
+                exact = TRUE, B = 1000000)
+```

### 2024-11-26 Code/gamlss.R
Contains women analysis and womens vs mens scatterplots

@@ -195,14 +195,10 @@ sim_prob(gg3b_no2022, .001)
 sim_prob(gg3b_no2022, .0001)
 
 ##Pool men and women's data?
-##Had a very difficult time fitting the data regularly. On the gg3b_w model I
-## kept getting an error: Error in if (dv > olddv && itn >= 2 && auto == TRUE) { : 
-## missing value where TRUE/FALSE needed.  I manually inspected the data and tried
-## varying datasets with really random results.  As a result we will use a no DQ
-## dataset as that seemed to work better.
+##Had a very difficult time fitting the women's data
 dhmw <- times |>
   filter(Type == "F" | Type == "S") |>
-  filter(Time != "DNS", Time != "DQ", Time != "D") |>
+  filter(Time != "DNS") |>
   filter(RxnTime > 0) |>
   filter(Event != "200 Dash") |>
   mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))

### 2024-11-26 Manuscript/supp.tex
Worked on supp file, might need to be cleaned a bit

@@
 \section{Rank-Based Comparison with Pooled Men and Women Data}
+As an alternative to the methods described in section 3.1 of the main paper, it
+is possible to combine all men's and women's data for each of the three time
+periods.  Thus we shrink our analyses from six to three, but each analysis
+is roughly twice as big as previously.  For the 2019 versus 2022 comparison,
+there were 258 reaction time observations from 65 athletes and the asymptotic
+test result was a p value of $3.56 \cdot 10^{-8}$. For the 2022 national versus
+internation comparison, there were 160 reaction time observations from 35
+athletes and the asymptotic test result was a p value of $1.94 \cdot 10^{-7}$.
+For the 2023 versus 2022 comparison, there were 343 reaction time observations
+from 92 athletes and the asymptotic test result was a p value of
+$4.99 \cdot 10^{-12}$.  These are all highly significant test results that show
+substanial differences in average reaction time for athletes competing at
+multiple championship level events.
 
 \section{GAMLSS Results for Women Data}
+We can repeat our reaction time barrier analysis described in section 3.2, but
+fit the same model to women's data from 2001 to 2023.  Once again we utilize
+the power of the venue effect within $\mu$ and a heat effect within $\sigma$. 
+We visualize the data we are analyzing in Figure~\ref{fig:WomensBoxplot}.
+Similar to the men's data, reaction times from 2022 appear lower than in other
+years.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\textwidth]{WomensBoxplot}
+  \caption{The RTs from 2021 to 2023 for the women's 100 meter hurdle
+  and 100 meter dash.}
+  \label{fig:WomensBoxplot}
+\end{figure}
+
+\of{Currently cannot fit the gg3b to the men's and womens data}
 
 \section{Results from Data Excluding Positive DQ RTs}
+An earlier iteration of the paper fit a model but did not include RTs from
+athletes who were disqualified or did not finish but still registered a reaction
+time.  However, it was decided that to better estimate the left tail of the
+distribution and more accurately predict the probability of a low reaction time,
+those RTs should be included.  We cannot include times from athltes with a
+negative reaction time, as that is a mistake of the runner for starting before
+the gun is fired, and thus those times are meaningless in our objective to
+determine a fair reaction time barrier.  Not all of those disqualified were
+disqualified because of breaking the 0.1 reaction time barrier; there are many
+reasons why an athlete may be diqualified with the most notable being a failed
+drug test and lane violations.  Nonetheless, in this iteration of the analysis,
+we will exclude these observations to see the effect on the probability of an
+extreme reaction time.  We fit an identical model to the generalized Gamma model
+presented in section 3.2 to examine differences in these types of data.
+
+\begin{table}
+  \centering
+  \caption{Probabilities of observing RTs less than threshold 0.08,
+  0.09, and 0.10 seconds based on the
+    fitted GG GAMLSS model with both venue- and heat-level
+random effects.}
+  \begin{tabular}{c c c c}
+   \toprule
+   Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\
+   \midrule
+   Without DQs & $4.93\cdot10^{-5}$ & $3.53\cdot10^{-4}$ &  $1.97\cdot10^{-3}$  \\
+   With DQs & $6.84\cdot10^{-5}$ & $4.95\cdot10^{-4}$ & $2.76\cdot10^{-3}$ \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:DQSim_probability}
+\end{table}
+
+Table~\ref{tab:DQSim_probability} shows the effect of removing disqualificated
+times from the analysis.  The probability of observing extreme RTs is lower when
+we remove the 17 observations.
 

### 2024-11-26 Manuscript/supp.tex
comments on supp

@@
 \section{Rank-Based Comparison with Pooled Men and Women Data}
+
+\jy{Keep the order consistent with the text; do the same in the code.}
+
 As an alternative to the methods described in section 3.1 of the main paper, it
@@
 \section{Results from Data Excluding Positive DQ RTs}
+
 An earlier iteration of the paper fit a model but did not include RTs from
@@
 
+\jy{Men's data or pooled data?}
+
 Table~\ref{tab:DQSim_probability} shows the effect of removing disqualificated

### 2024-11-26 Code/gg.R
tuned gamlss.control to fit all models

@@ -17,7 +17,6 @@ dhm <- times |>
 dhm$Venue <- as.factor(dhm$Venue)
 dhm$Heat <- as.factor(dhm$Heat)
 
-
 ## Fit various distributions available from gamlss
 ## Plot standarized residuals q-q plot
 library("gamlss")
@@ -28,14 +27,14 @@ gg0 <- gamlss(RxnTime ~ 1, data = dhm, family = GG)
 gg1 <- gamlss(RxnTime ~ random(Venue), data = dhm, family = GG)
 gg2 <- gamlss(RxnTime ~ random(Venue) + random(Heat), data = dhm, family = GG)
 gg3 <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
-## gg3a  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+gg3a  <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40, mu.step = .5))
 gg3b <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 gg3c <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
-#gg3d <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
-#gg4 <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
+gg3d <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Venue), data = dhm, family = GG, control = gamlss.control(n.cyc = 40, sigma.step = .5))
[…]

### 2024-11-26 Manuscript/manuscript.tex
comments on supp
referred to supp in main text

@@
 \label{sec:datanational}
-\jy{Owen, are there names from 100-m dash?}
-\of{Yes, there were four US athletes in the 100m dash that we have data on.
-Marvin Bracy, Fred Kerley, Travyon Bromwell, and Christian Coleman.  All of whom
-recorded their slowest times at the US Track and Field Championship.}
 
@@
 RT precision in elite-level competition.
-
-\eds{I also think we need to comment on the men and women's data. But within row
-of the figure, don't the distributions appear similar?  If that is so, would a
-reviewer ask why we didn't pool the data for men and women?}
-\jy{Let's put the pooled analysis in the supplement.}
-\of{Please see my email}
 
@@
 
+Random effects at the venue and heat levels are incorporated to the
+parameters in GG distribution in~\eqref{eq:gg}.
 Let $Y_{ijk}$ denote the RT of observation~$k$ in heat~$j$
@@
 zero and variance~$\tau_h^2$.
-The two random effects were found useful: one capturing the venue effect, which
-is used to contrast years, and the second being the heat effect, where every
-race was given a unique identifier with typically five to nine observations
-per race. The heat effect is important as it captures the variability in the
-amount of time athletes are on the starting blocks before the gun goes off.
+The two random effects were found useful: one 
+capturing the venue effect, which is used to contrast years, and the 
+second being the heat effect, where every race was given a unique 
+identifier with typically five to nine observations per race.
 This model can be fit with R package \texttt{gamlss} 
-\citep{stasinopoulos2008generalized}.
+\citep{stasinopoulos2008generalized}. The heat effect could be added
+to the model for $\mu_{ijk}$ and the venue effect could be added to 
+the model for $\sigma_{ijk}$. From our comparison using the Akaike 
+Information Criterion, Models~\eqref{eq:mu}--\eqref{eq:sigma} turned 
+out to be preferred to more complex models or competing models.
 
@@
 
+We also conducted the same analysis with men’s and women’s data
+pooled, yielding similar results. Details are provided in Section~1 of
+the Supplementary Material.
+
 \subsection{GAMLSS} \label{subsec:Results_GLMM}
 
+The results reported in this subsection are from men's data only
+because our investigation found significant gender difference.
 The fitted parameters of the GG distribution in the
@@
 
+Results for women's data are reported in Section~2 of the
+Supplementary Material.
+
 \section{Discussion}\label{sec:concludingremarks}

### 2024-11-26 Manuscript/supp.tex

@@
 
-\section{Rank-Based Comparison with Pooled Men and Women Data}
+\section{Rank-Based Comparison with Pooled Data}
 
@@
 \section{GAMLSS Results for Women Data}
+
 We can repeat our reaction time barrier analysis described in section 3.2, but
@@
 \end{figure}
+
+The results are slightly different from those from the men's data,
+which echoes existing studies reporting gender differences in reaction
+times \citep[e.g.,][]{lipps2011implications, babicc2009reaction,
+  panoutsakopoulos2020gender}.
 

### 2024-11-27 Code/gg.R
small change

@@ -91,7 +91,7 @@ dhw <- times |>
 filter(Type == "F" | Type == "S") |>
 filter(Time != "DNS") |>
 filter(Gender == "F") |>
-filter(RxnTime > 0) |>
+filter(RxnTime > 0.072) |>
 filter(Event %in% c("100 Dash", "100 Hurdles"))
 
 dhw <- within(dhw, {Gender = as.factor(Gender); Event = as.factor(Event);
@@ -101,6 +101,16 @@ gg_dhw <- gamlss(RxnTime ~ random(Venue) + Event, sigma.formula = ~ random(Heat)
 
 summary(gg_dhw) # event is significant for women!
 ## one rxn time is too small, probabily really was a DQ
+
+gg_dhw3 <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + Event, data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+gg_dhw3a <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Heat) + Event, data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+gg_dhw3b <- gamlss(RxnTime ~ random(Venue) + Event, sigma.formula = ~ random(Heat), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+gg_dhw3c <- gamlss(RxnTime ~ random(Heat), sigma.formula = ~ random(Heat), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+gg_dhw3d <- gamlss(RxnTime ~ random(Heat) + Event, sigma.formula = ~ random(Venue), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+gg_dhw4 <- gamlss(RxnTime ~ random(Venue) + random(Heat) + Event, sigma.formula = ~random(Venue) + random(Heat), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+#qq_conf_plot(residuals(gg_dhw4))
+AIC(gg_dhw3, gg_dhw3a, gg_dhw3b, gg_dhw3c, gg_dhw3d, gg_dhw4)
+
 
 
 ## How much does including dqs matter?

### 2024-11-27 Code/gg.R
trying different womens models

@@ -112,7 +112,6 @@ gg_dhw4 <- gamlss(RxnTime ~ random(Venue) + random(Heat) + Event, sigma.formula 
 AIC(gg_dhw3, gg_dhw3a, gg_dhw3b, gg_dhw3c, gg_dhw3d, gg_dhw4)
 
 
-
 ## How much does including dqs matter?
 dhm <- times |>
   filter(Type == "F" | Type == "S") |>

### 2024-11-27 Code/gamlss.R
small change

@@ -200,7 +200,7 @@ dhmw <- times |>
   filter(Type == "F" | Type == "S") |>
   filter(Time != "DNS") |>
   filter(RxnTime > 0) |>
-  filter(Event != "200 Dash") |>
+  filter(Event == "110 Hurdles" | Event == "100 Hurdles" | Event == "100 Dash") |>
   mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))
   
 gg3b_gender  <- gamlss(RxnTime ~ random(Venue) + Gender, sigma.formula = ~ random(Heat), data = dhmw, family = GG, control = gamlss.control(n.cyc = 40))

### 2024-11-27 Code/gg.R
refit after removing one outlier in women data

@@ -96,10 +96,15 @@ filter(Event %in% c("100 Dash", "100 Hurdles"))
 
 dhw <- within(dhw, {Gender = as.factor(Gender); Event = as.factor(Event);
     Venue = as.factor(Venue); Heat = as.factor(Heat)})
+dhw <- dhw[- which.min(dhw$RxnTime), ]
 
-gg_dhw <- gamlss(RxnTime ~ random(Venue) + Event, sigma.formula = ~ random(Heat), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+gg3b.w <- gamlss(RxnTime ~ random(Venue) + Event, sigma.formula = ~ random(Heat), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+gg3d.w <- gamlss(RxnTime ~ random(Venue) + Event, sigma.formula = ~ random(Venue) + random(Heat), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
 
-summary(gg_dhw) # event is significant for women!
+AIC(gg3b.w, gg3d.w)
+qq_conf_plot(residuals(gg3b.w), dparams = list(mean = 0, sd = 1)) ## z-scores have known mean and sd
+
+##  summary(gg_dhw) # event is significant for women!
 ## one rxn time is too small, probabily really was a DQ
 
 gg_dhw3 <- gamlss(RxnTime ~ random(Venue) + random(Heat), sigma.formula = ~ random(Venue) + Event, data = dhw, family = GG, control = gamlss.control(n.cyc = 40))

### 2024-11-29 Code/RankComparison.Rmd
Updated language and removed supplementary code

--- 2024-11-20 Code/RankComparison.Rmd
@@
 ```
+The p-value from the asymptomatic test was $1.126 \cdot 10^{-5}$ and the p-value
+from the permutation test was $2.8 \cdot 10 ^ {-5}$. Both of these are
+statistically significant results showing a difference in male athletes who
+competed in 2019 and 2022.
 
@@
 ```
+The p-value from the asymptomatic test was $6.9 \cdot 10^{-3}$ and the p-value
+from the permutation test was $1.5 \cdot 10^{-3}$. Both of these are
+statistically significant results showing a difference in female athletes who
+competed in 2019 and 2022.
 
@@
 
+The p-value from the asymptomatic test was $6.1 \cdot 10^{-5}$ and the p-value
+from the permutation test was $1.0 \cdot 10^{-6}$. Both of these are
+statistically significant results showing a difference in male athletes who
+competed at various competitions in 2022.
+
 ### Women's Data
@@
 ```
+The p-value from the asymptomatic test was $1.2 \cdot 10^{-3}$ and the p-value
+from the permutation test was $1.0 \cdot 10^{-6}$. Both of these are
+statistically significant results showing a difference in female athletes who
+competed at various competitions in 2022.
 
@@
 
+The p-value from the asymptomatic test was $1.4 \cdot 10^{-6}$ and the p-value
+from the permutation test was $2.8 \cdot 10 ^ {-5}$. Both of these are
+statistically significant results showing a difference in male athletes who
+competed in 2022 and 2023.
+
 ### Women's Data
@@
 ```
+
+The p-value from the asymptomatic test was $9.4 \cdot 10^{-7}$ and the p-value
+from the permutation test was $1.5 \cdot 10^{-3}$. Both of these are
+statistically significant results showing a difference in female athletes who
+competed in 2019 and 2022.

### 2024-11-29 Code/ReactionBarrierAnalysis.Rmd
Contains gamlss section, started supplementary

---
title: "ReactionBarrierAnalysis"
author: "Owen Fiore"
date: "2024-11-28"
output: html_document
---

The objective of this document is to allow other researchers to replicate our
results in regards to the GAMLSS section of the paper.
# Main Analysis
The first thing we need is our data.  We create the `times` data to be the full
data, and will create relevant subsets depending on what we need to fit.

```{r}
times <- read.csv("../Data/rxntime.csv")
colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
```

The first data set is the men's dash and hurdles which we will denote as `dhm`.
The first filter we apply is to look at data from semifinals and finals, as
previous papers have suggested that reaction times in the heats may not be 
as fast due to less competitive heats.  We cannot include times from athletes
who did not start (DNS) as they do not have reaction times, but we do consider
times from athletes who were disqualified (This is further examined in the
supplemental analysis).  We cannot fit negative times with our gamma model and
negative reaction times make little practical sense, so we remove those as well.
Additionally we need to select the 100 meter dash and 110 meter hurdles as
there are more events (200 meter dash) included in the data but not included in
this analysis.  Lastly we convert `Venue` and `Heat` to be factor variables as
those are used as random effects in the model.

An important part of the paper is examining the results of excluding 2022 from
the analysis to see the impact 2022 had.  Thus we are going to be replicating
much of the same methods for the modeling phrase but need to exclude 2022 from
the dataset.
```{r}
library(tidyverse)
dhm <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime > 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))

#Only difference between these two dataframes is the filter removing 2022
dhm_no2022 <- times |>
  filter(Type == "F" | Type == "S") |>
  filter(Gender == "M") |>
  filter(Time != "DNS") |>
  filter(RxnTime > 0) |>
  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
  filter(Venue != 2022) |>
  mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))
```

We considered many families of models, and formula structures within the
generalized Gamma but thought that the results from the `gg3b` game strong
results.  The first term corresponds to the $\mu$ term of the model, which is
a random effect for `Venue`. $\sigma$ contains a random effect for `Heat`.
```{r}
gg3b <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat),
               data = dhm, family = GG, control = gamlss.control(n.cyc = 40))

gg3b_no2022 <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat),
                      data = dhm_no2022, family = GG, control = gamlss.control(n.cyc = 40))

```

We can explore some information about this model and replicate the results
presented in Table 2 of the paper by running the code below.
```{r}
summary(gg3b)
cat("sigma_v or std dev of venue effect: ", gg3b$mu.coefSmo[[1]]$sigb, "\n")
cat("sigma_h or std dev of heat effect: ", gg3b$sigma.coefSmo[[1]]$sigb[1], "\n")
```

And now for the model excluding 2022
```{r}
summary(gg3b)
cat("sigma_v or std dev of venue effect: ", gg3b_no2022$mu.coefSmo[[1]]$sigb, "\n")
cat("sigma_h or std dev of heat effect: ", gg3b_no2022$sigma.coefSmo[[1]]$sigb[1], "\n")
```

One of the key findings of this section is that by including 2022 in the model,
the standard deviation of the venue effects increases from 0.043 to 0.058. This
demonstrates that the reaction times from 2022 helped contribute to more
variability in the range of venue effects.

The code below plots the venue effects from the model to visually see the
differences that result by including 2022.
```{r}
library(gridExtra)
venue_effects <- as.data.frame(gg3b$mu.coefSmo[[1]]$coef)
colnames(venue_effects) <- c("Effect Value")

venue_plot <- ggplot(venue_effects, aes(x = rownames(venue_effects),
                                               y = `Effect Value`)) +
  geom_point() + 
  labs(x = "Venue", y = "Venue Effect")+
  scale_y_continuous(limits = c(-0.15, 0.1))

## Second plot
venue_effects_no2022 <- as.data.frame(gg3b_no2022$mu.coefSmo[[1]]$coef)
venue_effects_no2022["2022", ] <- 1 #Arbitrary number out of range of y scale so
#that the x axis line up

colnames(venue_effects_no2022) <- c("Effect Value")

venue_plot_no2022 <- ggplot(venue_effects_no2022,
                                   aes(x = rownames(venue_effects_no2022),
                                       y = `Effect Value`)) +
  geom_point() + 
  labs(x = "Venue", y = "Venue Effect") +
  scale_y_continuous(limits = c(-0.15, 0.1))

#Put it together
grid.arrange(venue_plot, venue_plot_no2022, ncol = 1)
```
In the above figure, the top graph corresponds to the venue effects from `gg3b`,
and the bottom graph represents the venue effects from `gg3b_no2022` which is
why there is no 2022 venue effect for 2022.  There is some year to year
variability in the graphs, with differences in venue effects in 2007 and 2011
occuring potentially due to the change in disqualification rules that were
implemented by World Athletics.  The venue effects from 2015 to 2023 were
very consistent with values very close to 0, with the exception of 2022 which
had a venue effect of roughly -0.13.

We also need to evaluate how well the `gg3b` model fit the data, and we look
at the standardized residuals to do so.

```{r}
library(qqconf)
qq_conf_plot(residuals(gg3b))
```

We see that the model fits the data extremely well considering the size of the
data is roughly 775 points.  Thus this seems like a strong model choice that
will produce good results.

```{r}
qq_conf_plot(residuals(gg3b_no2022))
```
The residual plot when the model is fit without 2022 is very similar with few
noticeable differences.

Now that we have chosen a model and it passed necessary specifications, we can
use it to simulate times and determine how extreme a 0.1 second reaction time
barrier is.  The function `simfit` draws 10,000,000 samples from a probability
distribution determined in the last line of the function, which draws from a
generalized Gamma distribution parameterized by `mu`, `sigma`, and `nu`.

```{r}
simfit <- function(model, n = 10000000) {
    set.seed(1)

    #Extract standard deviation of venue and heat effect
    sd_Venue <- model$mu.coefSmo[[1]]$sigb
    sd_Venue_heat <- model$sigma.coefSmo[[1]]$sigb[1]

    #Calculate mu, sigma, nu with coefficients and standard deviation of random effects
    mu <- exp(model$mu.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue))
    sigma <- exp(model$sigma.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue_heat))
    nu <- model$nu.coefficients[1]

    #Generate the distribution
    rGG(n, mu = mu, sigma = sigma, nu = nu)
}
```
Using the distribution returned in the last line of code, we then want to
examine the probability of observing a reaction time less than 0.1 seconds.  We
do this as a way of measuring whether the 0.1 barrier is fair.  If we see a
large probability, this indicates that 0.1 seconds is too strict a barrier and
World Athletics should consider changes.

Now we can fit our distribution and assign it to `samplefit` and repeat for
the model and thus distribution without 2022.

```{r}
samplefit <- simfit(gg3b)
samplefit_no2022 <- simfit(gg3b_no2022)
```

We use `mean(samplefit < 0.10)` to determine the proportion of the distribution
less than 0.1 and repeat this for 0.09 and 0.08 to assess how extreme those
would be.  These probabilities can be found in table 3 of the paper.
```{r}

mean(samplefit < 0.10)
mean(samplefit_no2022 < 0.10)
cat("\n")
mean(samplefit < 0.09)
mean(samplefit_no2022 < 0.09)
cat("\n")
mean(samplefit < 0.08)
mean(samplefit_no2022 < 0.08)
```
We see that in every comparison, it is more likely to observe an extreme reaction
time when 2022 is included in the analysis.

Perhaps the more interesting aspect of this analysis, is the ability to
recommend a new reaction time barrier based on the distribution.  Thus we want
to know the quantile that certain important probabilities would have.  If World
Athletics wanted to quantitatively create a reaction time barrier what they
should do is first decide on what percentage of times they should disqualify
for being unreasonable.  We are going to look at 1 in every 100, 1000, and 10000
observations and the result is what the reaction barrier should be.  The results
below are presented in table 4 of the paper.

```{r}
quantile(samplefit, c(.01, .001, .0001))
cat("\n")
quantile(samplefit_no2022, c(.01, .001, .0001))
```
We see that by excluding 2022, the suggested reaction time barrier is higher
and more strict.  this analysis is important and interesting to repeat in several
years to see what the effect of future World Championships will be on the barrier.
If there are more years like 2022, then World Athletics should take steps to
change the reaction time barrier because that indicates that athletes are
getting faster and thus reaction times under 0.1 seconds are no longer looked
at as the athlete predicting the gun, but rather that they have fast reflexes.

# Supplemental Analysis

We are going to see what happens when we pool men's and women's data together.
We sill need to separate by year but will have 3 analyses instead of 6.  As a
result of the larger data set we are going to run the permutation tests at a
size of $n = 100,000$ rather than $n = 1,000,000$ as the code takes several
hours to run otherwise.  The asymptotic results are all highly statistically
significant, and with the larger data sets, the results from the asymptotic
tests should be of higher quality.

### 2019 vs 2022

Need to talk about results after running code

Data setup
```{r}
men_2019 <- read.csv("../Data/Clusrank2019vs2022.csv")
colnames(men_2019) <- c("Time", "Athlete", "Year", "Event")

women_2019 <- read.csv("../Data/Clusrank2019vs2022Womens.csv")
colnames(women_2019) <- c("Time", "Athlete", "Year", "Event")

women_2019$Athlete <- women_2019$Athlete + 34 #Correction for unique athlete identifier

all_2019 <- rbind(men_2019, women_2019)
```

And now we run the tests
```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2019,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2019,
                method = "ds",
                exact = TRUE, B = 100000)
```
### 2022 National vs International

Data setup
```{r}
men_2022 <- read.csv("../Data/ClusrankNatvsWorld.csv")
colnames(men_2022) <- c("Time", "Athlete", "Year", "Event")

women_2022 <- read.csv("../Data/ClusrankNatvsWorldWomens.csv")
colnames(women_2022) <- c("Time", "Athlete", "Year", "Event")

women_2022$Athlete <- women_2022$Athlete + 17 #Correction for unique athlete identifier

all_2022 <- rbind(men_2022, women_2022)
```

And now we run the tests.

```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2022,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2022,
                method = "ds",
                exact = TRUE, B = 100000)
```
### 2022 vs 2023

```{r}
men_2023 <- read.csv("../Data/Clusrank2022vs2023.csv")
colnames(men_2023) <- c("Time", "Athlete", "Year", "Event")

women_2023 <- read.csv("../Data/Clusrank2022vs2023Womens.csv")
colnames(women_2023) <- c("Time", "Athlete", "Year", "Event")

women_2023$Athlete <- women_2023$Athlete + 45 #Correction for unique athlete identifier

all_2023 <- rbind(men_2023, women_2023)
```

Now we run the tests.
```{r}
set.seed(1)
library(clusrank)
clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2023,
                method = "ds")

clusWilcox.test(Time, group=Year, cluster = Athlete, data = all_2023,
                method = "ds",
                exact = TRUE, B = 100000)
```

### 2024-11-29 Code/gg.R
model gg3b0.w for women

@@ -98,11 +98,14 @@ dhw <- within(dhw, {Gender = as.factor(Gender); Event = as.factor(Event);
     Venue = as.factor(Venue); Heat = as.factor(Heat)})
 dhw <- dhw[- which.min(dhw$RxnTime), ]
 
+
 gg3b.w <- gamlss(RxnTime ~ random(Venue) + Event, sigma.formula = ~ random(Heat), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+gg3b0.w <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
 gg3d.w <- gamlss(RxnTime ~ random(Venue) + Event, sigma.formula = ~ random(Venue) + random(Heat), data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
 
-AIC(gg3b.w, gg3d.w)
-qq_conf_plot(residuals(gg3b.w), dparams = list(mean = 0, sd = 1)) ## z-scores have known mean and sd
+AIC(gg3b.w, gg3b0.w, gg3d.w)
+## Let's report gg3b0.w
+qq_conf_plot(residuals(gg3b0.w), dparams = list(mean = 0, sd = 1)) ## z-scores have known mean and sd
 
 ##  summary(gg_dhw) # event is significant for women!
 ## one rxn time is too small, probabily really was a DQ

### 2024-11-29 Manuscript/manuscript.tex
Made small edit to table 4
please review comments; also minor edits made

@@
 has been the subject of significant debate. Partly based on limited
-data from Finnish national-level athletes~\cite{mero1990reaction}, the
+data from Finnish national-level athletes~\citep{mero1990reaction}, the
 threshold may not adequately represent the capabilities of elite
@@
 events. Using a matched-pairs design, we compare RTs across the 2022,
-2019, and 2023 World Championships, as well as the 2022 national-level
+2019, and 2023 World Championships, as well as 2022 national-level
 competitions. This approach isolates the effect of the competition
-year while controlling for individual performance. The clustered
+ while controlling for individual performance. The clustered
 nature of the data, with the goal of assessing differences across
-events within athletes, was analyzed using a rank-based comparison
+competitions within athletes, was analyzed using a rank-based comparison
 approach for clustered data \citep{datta2005rank}. Second, we evaluate
@@
 competing. Unless otherwise noted, this pooled dataset forms the basis
-for our analysis throughout the paper. Additionally, we consider datasets
+for our analyses for Objective 2. Additionally, we consider datasets
 that exclude 2022 to assess how our findings might differ when excluding
@@
 model parameters. The GG distribution has three parameters, denoted by
-$\text{GG}(\mu, \sigma, \nu)$ has density function
+$\text{GG}(\mu, \sigma, \nu),$ has density function
 \begin{equation}
@@
 p-values, indicating stronger evidence of faster RTs at the
-2022 World Championships compared to the 2022 national competitions. For
-both men and women, permutation test results were highly significant,
+2022 World Championships compared to the 2022 national competitions. 
+\eds{This does not appear true based on the data currently in Table 1. 
+2022-2023 seem to have the lowest p-values.  Which is correct?}
+For both men and women, permutation test results were highly significant,
 providing substantial evidence that athletes competing at the 2022 World
@@
 because our investigation found significant gender difference.
-The fitted parameters of the GG distribution in the
+Results for women's data are reported in Section~2 of the
+Supplementary Material. The fitted parameters of the GG distribution in the
 GAMLSS framework in Equations~\eqref{eq:gg}--\eqref{eq:sigma} are
@@
 the standard deviation of the venue-level random-effect variance,
-which increases from 0.043 to 0.058 . These results
+which increases from 0.043 to 0.058 . 
+\eds{are you reporting estimated variances or SDs of the random effects here and 
+in the Table?  In the table, we have $\tau_v$ and $\tau_h$ (SD) labels, not 
+$\tau_v^2$ and $\tau_h^2$ (variances).  Please check and correct as needed.}  
+These results
 highlight that RTs are influenced by both venue and heat-level
@@
     histogram of the observed RTs (left); Q-Q plot of the
-    normal z-score of the quantile residuals from the fitted model.}
+    normal z-score of the quantile residuals from the fitted model (right).}
   \label{fig:diagnosis}
@@
   \caption{Suggested RT barriers based on tail probabilities.}
-  \begin{tabular}{c c c}
+  \begin{tabular}{c c c c}
    \toprule
-   Data Set & Tail probability  $10^{-3}$ & Tail probability $10^{-4}$ \\
+   Data Set & Tail probability  $10^{-2}$ & Tail probability  $10^{-3}$ & Tail probability $10^{-4}$ \\
    \midrule
-   Excluding 2022 & $0.096$ & $0.083$ \\
-   Including 2022 & $0.094$ & $0.082$ \\
+   Excluding 2022 & $0.111$ & $0.096$ & $0.083$ \\
+   Including 2022 & $0.108$ & $0.094$ & $0.082$ \\
    \bottomrule
@@
 
-Results for women's data are reported in Section~2 of the
-Supplementary Material.
-
 \section{Discussion}\label{sec:concludingremarks}
@@
 meets. 
-\eds{are we testing faster (one-sided) or different (two-sided)?}
 Ideally, a centralized database containing RTs from
@@
 \section*{Supplementary Material}
+\eds{when the supplement is complete, include an additional sentence here
+that summarizes the additional analyses presented in the supplemental material}
 The data and R code used for the analysis are available in a compressed file for

### 2024-11-29 Manuscript/supp.tex
fixed typos, added comments

--- 2024-11-26 Manuscript/supp.tex
@@
 
-As an alternative to the methods described in section 3.1 of the main paper, it
-is possible to combine all men's and women's data for each of the three time
-periods.  Thus we shrink our analyses from six to three, but each analysis
+As an alternative to the methods described in Section 3.1 of the main paper, it
+is possible to combine all men's and women's reaction time (RT) data for each of 
+the three competition
+comparisons.  Thus we shrink our analyses from six to three, but each analysis
 is roughly twice as big as previously.  For the 2019 versus 2022 comparison,
-there were 258 reaction time observations from 65 athletes and the asymptotic
-test result was a p value of $3.56 \cdot 10^{-8}$. For the 2022 national versus
-internation comparison, there were 160 reaction time observations from 35
-athletes and the asymptotic test result was a p value of $1.94 \cdot 10^{-7}$.
-For the 2023 versus 2022 comparison, there were 343 reaction time observations
-from 92 athletes and the asymptotic test result was a p value of
+there were 258 RTs from 65 athletes and the asymptotic
+test result was a p-value of $3.56 \cdot 10^{-8}$. For the 2022 national versus
+international comparison, there were 160 RTs from 35
+athletes and the asymptotic test result was a p-value of $1.94 \cdot 10^{-7}$.
+For the 2023 versus 2022 comparison, there were 343 RTs
+from 92 athletes and the asymptotic test result was a p-value of
 $4.99 \cdot 10^{-12}$.  These are all highly significant test results that show
-substanial differences in average reaction time for athletes competing at
-multiple championship level events.
+substantial differences in average RT for athletes competing at
+multiple championship-level events.
 
 \section{GAMLSS Results for Women Data}
-We can repeat our reaction time barrier analysis described in section 3.2, but
-fit the same model to women's data from 2001 to 2023.  Once again we utilize
+
+We can repeat our reaction time barrier analysis described in Section 3.2, but
+fit the same model to women's RT data from 2001 to 2023.  Once again we utilize
 the power of the venue effect within $\mu$ and a heat effect within $\sigma$.
-We visualize the data we are analyzing in Figure~\ref{fig:WomensBoxplot}.
-Similar to the men's data, reaction times from 2022 appear lower than in other
+\eds{what is meant by `power' here?} 
+The RT data for women is visualized in Figure~\ref{fig:WomensBoxplot}.
+Similar to the men's data, RTs from 2022 appear lower than in other
 years.
@@
 
+The results are slightly different from those from the men's data,
+which echoes existing studies reporting gender differences in RTs 
+\citep[e.g.,][]{lipps2011implications, babicc2009reaction,
+  panoutsakopoulos2020gender}.
+
 \of{Currently cannot fit the gg3b to the men's and womens data}
 
-\section{Results from Data Excluding Positive DQ RTs}
+\section{Results from Data Excluding Positive Disqualified RTs}
 
-An earlier iteration of the paper fit a model but did not include RTs from
-athletes who were disqualified or did not finish but still registered a reaction
-time.  However, it was decided that to better estimate the left tail of the
-distribution and more accurately predict the probability of a low reaction time,
-those RTs should be included.  We cannot include times from athltes with a
-negative reaction time, as that is a mistake of the runner for starting before
-the gun is fired, and thus those times are meaningless in our objective to
-determine a fair reaction time barrier.  Not all of those disqualified were
+An earlier iteration of the paper fit a model that did not include RTs from
+athletes who were disqualified or did not finish but still registered a RT.  
+However, it was ultimately decided to include these times to better estimate the 
+left tail of the distribution and more accurately predict the probability of a 
+low RT, as described in the main paper.  
+We did not include negative RTs, however, as these represent a mistake 
+of the runner for starting before
+the gun is fired and are thus meaningless in our objective to
+determine a fair RT barrier.  Not all of those disqualified were
 disqualified because of breaking the 0.1 reaction time barrier; there are many
-reasons why an athlete may be diqualified with the most notable being a failed
-drug test and lane violations.  Nonetheless, in this iteration of the analysis,
-we will exclude these observations to see the effect on the probability of an
-extreme reaction time.  We fit an identical model to the generalized Gamma model
-presented in section 3.2 to examine differences in these types of data.
+reasons why an athlete may be disqualified, with the most notable being failed
+drug tests and lane violations.  Nonetheless, in this section,
+we exclude all disqualified RTs to see their effect on the probability of an
+extreme RT.  We fit an identical model to the generalized Gamma model
+presented in Section 3.2 to examine differences results.
 
@@
 \jy{Men's data or pooled data?}
+\eds{with or without 2022?}
 

### 2024-11-29 Manuscript/supp.tex
Wrapping up women's section

@@
 
-The results are slightly different from those from the men's data,
-which echoes existing studies reporting gender differences in RTs 
-\citep[e.g.,][]{lipps2011implications, babicc2009reaction,
-  panoutsakopoulos2020gender}.
+We examined several models in the generalized Gamma model and found that the
+best model for women involved a venue effect in $\mu$ and heat effect in
+$\sigma$.  Thus the best model for men was also the best model for women.
 
-\of{Currently cannot fit the gg3b to the men's and womens data}
+\begin{table}
+  \centering
+  \caption{Estimated fixed-effect parameters with standard errors in
+    parentheses and estimated variance of the random effects from the men's and
+    women's fitted GG distribution with venue level random
+    effects in $\mu$ and heat level random effects in $\sigma$. $n$ denotes
+    size of the data.}
+  \label{tab:womensfit}
+  \begin{tabular}{c c c c c c c}
+    \toprule
+    Dataset & $n$ & $\beta_0$ & $\gamma_0$ & $\nu$ & $\tau_v$ & $\tau_h$ \\
+    \midrule
+    Women's & 732 & $-$1.921 (0.007) & $-$2.071 (0.028) & $-$3.691 (0.665) & 0.057 & 0.111 \\
+    Men's & 776 & $-$1.910 (0.005) & $-$2.200 (0.027) & $-$1.178 (0.447) & 0.058 & 0.320 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+One of the most interesting aspects of Table~\ref{tab:womensfit} is that while
+the venue effect standard deviation is nearly identical, the women's heat effect
+standard deviation is smaller and $\nu$ is much larger.  If men's races have
+more variability in times, it seems possible that in men's races that faster
+athletes can influence other athletes to react faster in certain instances.
+It is also interesting that the standard deviation of the venue effect is
+nearly identical for both men's and women's data, indicating that the venue
+effect is not only statistically significant but is consistent in strength
+across men and women.
+
+We can repeat the simulation methods described in the main paper to evaluate
+the probability of an extreme reaction time.  Table~\ref{tab:Sim_prob_women}
+compares the men's and women's results of observing reaction times less than
+0.08, 0.09, and 0.1 seconds.  We find that across all three comparisons that
+women have a lower probability of having a fast time. The results are slightly
+different from those from the men's data, which echoes existing studies
+reporting gender differences in RTs \citep[e.g.,][]{lipps2011implications,
+babicc2009reaction, panoutsakopoulos2020gender}.
+
+\begin{table}
+  \centering
+  \caption{Probabilities of observing RTs less than threshold 0.08,
+  0.09, and 0.10 seconds based on the men's and women's
+    fitted GG GAMLSS model with both venue- and heat-level
+random effects.}
+  \begin{tabular}{c c c c}
+   \toprule
+   Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\
+   \midrule
+   Women's & $1\cdot10^{-7}$ & $1.12\cdot10^{-5}$ &  $5.46\cdot10^{-4}$  \\
+   Men's   & $6.84\cdot10^{-5}$ & $4.95\cdot10^{-4}$ & $2.76\cdot10^{-3}$ \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:Sim_prob_women}
+\end{table}
+
+Now we need to perform alter the interpretation of the analyis to determine
+what a fair reaction time barrier for women should be.
+Table~\ref{tab:Sim_time_women} shows the difference in men's and women's
+suggested reaction time barrier.  What we see is that by World Athletics setting
+the same standard for men and women, their rules are not fair, in that men are
+more likely to be disqualified because they are more likely to react quickly.
+One of the biggest takeaways of the reaction time barrier analysis is that the
+current standards imposed by World Athletics are not grounded in statistical
+analysis and that they should be.  It seems reasonable to either raise the
+women's barrier so that they are disqualified at a similar rate as men or to
+lower the men's barrier so they are disqualified at a similar rate as women.  But
+having equal but not neccessarily fair standards does not seem reasonable when
+this study's results echoed results of numerous other papers asking World
+Athletics to change the standards.
+
+\begin{table}
+  \centering
+  \caption{Suggested RT barriers based on tail probabilities.}
+  \begin{tabular}{c c c c}
+   \toprule
+   Data Set & Tail probability  $10^{-2}$ & Tail probability  $10^{-3}$ & Tail probability $10^{-4}$ \\
+   \midrule
+   Women's & $0.111$ & $0.102$ & $0.095$ \\
+   Men's   & $0.108$ & $0.094$ & $0.082$ \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:Sim_time_women}
+\end{table}
 

### 2024-11-29 Code/ReactionBarrierAnalysis.Rmd
Updated to include women's analysis

@@
 ```
+Let's visualize the data.
+```{r}
+boxplot_df <- dhm |>
+  mutate(Venue = as.numeric(as.character(Venue)))
+library(ggplot2)
+ggplot(boxplot_df, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.05, 0.275),
+                            breaks = seq(0.05, 0.275, 0.025)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(1998,2024),
+                                              breaks = seq(1999, 2024, 2))
+```
+We see that the 2022 boxplot is lower than any other year.  What is particularly
+interesting is that Allen's 0.099 reaction time that led to his disqualification
+does not appear as an outlier in 2022, it is just the minimum value.  Compared
+to in 1999, where a value right around 0.100 was an outlier.
 
@@
 summary(gg3b)
-cat("sigma_v or std dev of venue effect: ", gg3b$mu.coefSmo[[1]]$sigb, "\n")
-cat("sigma_h or std dev of heat effect: ", gg3b$sigma.coefSmo[[1]]$sigb[1], "\n")
+cat("tau_v or std dev of venue effect: ", gg3b$mu.coefSmo[[1]]$sigb, "\n")
+cat("tau_h or std dev of heat effect: ", gg3b$sigma.coefSmo[[1]]$sigb[1], "\n")
 ```
@@
 summary(gg3b)
-cat("sigma_v or std dev of venue effect: ", gg3b_no2022$mu.coefSmo[[1]]$sigb, "\n")
-cat("sigma_h or std dev of heat effect: ", gg3b_no2022$sigma.coefSmo[[1]]$sigb[1], "\n")
+cat("tau_v or std dev of venue effect: ", gg3b_no2022$mu.coefSmo[[1]]$sigb, "\n")
+cat("tau_h or std dev of heat effect: ", gg3b_no2022$sigma.coefSmo[[1]]$sigb[1], "\n")
 ```
@@
 ```
+
+By pooling together all events within the year comparison, we find highly
+significant results, suggesting that on average, the athletes who competed in
+2022 had faster reaction times.
+
+### Women's Gamlss
+We decided to separate women's results after numerous papers suggested that
+men and women have differing reaction times, but we still thought it was important
+to investigate what an appropriate reaction time barrier was for women.  The
+women's data is very similar to the men's, but there is not data from 1999
+available for women and the women compete in the 100 meter hurdle instead of the
+110 meter hurdle.  We are going to replicate many of the same methods used above
+for the men's results.
+
+We read in the women's data by filtering `Gender` and `Event` differently.
+```{r}
+library(tidyverse)
+dhw <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "F") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime > 0) |>
+  filter(Event == "100 Hurdles" | Event == "100 Dash")
+dhw <- dhw[- which.min(dhw$RxnTime), ] #We remove the minimum time as it is an
+#extreme outlier as we want an accurate prediction of the left tail
+```
+
+Lets visualize the women's data.  This appears as Figure 1 in the supplemental
+file.
+```{r}
+library(ggplot2)
+ggplot(dhw, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.05, 0.3),
+                            breaks = seq(0.05, 0.3, 0.025)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(2000,2024),
+                                              breaks = seq(2001, 2024, 2))
+```
+We see that 2022 appears as a low outlier, but perhaps not as extreme as for the
+men's data.  The women's data also seems to be more stable than the men's with
+fewer venue to venue differences.
+
+Now we are ready to fit the women's model.  We tested several types of generalized
+Gamma models, including those different from the men's model, and found that
+this model, which is the female equivalent to the men's best model, was the best
+fit.
+```{r}
+dhw <- dhw|>
+  mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))
+gg3b0.w <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat),
+                  data = dhw, family = GG, control = gamlss.control(n.cyc = 40))
+```
+
+```{r}
+summary(gg3b0.w)
+cat("tau_v or std dev of venue effect: ", gg3b0.w$mu.coefSmo[[1]]$sigb, "\n")
+cat("tau_h or std dev of heat effect: ", gg3b0.w$sigma.coefSmo[[1]]$sigb[1], "\n")
+```
+We see that the heat and venue effects are statistically significant and
+contribute positively to the model.  We are also going to look at the residuals
+to see how well the data is fit.
+```{r}
+library(qqconf)
+qq_conf_plot(residuals(gg3b0.w))
+```
+Although this fit is not as strong as the men's data, we still found it to be
+acceptable and it was the best of the several women's models that we looked at.
+
+As the structure of the model remains the same, we are able to reuse the `simfit`
+code developed above in order to examine what the women's disqualification
+barrier should look like.
+
+```{r}
+samplefit_w <- simfit(gg3b0.w)
+
+mean(samplefit_w < 0.10)
+cat("\n")
+mean(samplefit_w < 0.09)
+cat("\n")
+mean(samplefit_w < 0.08)
+
+```
+
+```{r}
+quantile(samplefit_w, c(.01, .001, .0001))
+```
+
+It is worth noting that these suggested times are slower than the one's for the
+men's data.  Thus, men and women are not currently held to the same standard
+if men are more likely to react quickly than women and if they share the 0.1
+barrier.

### 2024-11-30 Code/ReactionBarrierAnalysis.Rmd
Added dq supplemental analysis

@@
 barrier.
+
+### Disqualification Analysis
+An earlier form of this paper did not originally include times from athlete's
+who were disqualified, but response from reviewer's prompted us to question
+whether we could include these times.  While we ultimately chose to include
+positive reaction times from athlete's who were disqualified, this section of
+the supplement is dedicated to investigating how much these times matter and
+examining the effect of a few but important times on the overall model.  These
+times are important because most of them fall on the left tail of the
+distribution, which is the part that we care the most about predicting accurately.
+
+```{r}
+dhm_nodq <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS", Time != "DQ", Time != "D") |>
+  filter(RxnTime > 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles")
+
+nrow(dhm) - nrow(dhm_nodq)
+```
+We see that there are 17 male athletes that had positive reaction times but were
+disqualified.  Not all of these athletes were disqualified for having reaction
+times less than 0.1 seconds, in many instances these athletes could have been
+disqualified for failed drug tests and other infractions.
+
+Now we fit the model.
+```{r}
+gg3b_nodq <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat),
+                    data = dhm_no2022, family = GG, control = gamlss.control(n.cyc = 40))
+```
+
+Let us see how we did with residuals
+```{r}
+qq_conf_plot(resid(gg3b_nodq))
+```
+The standardized residual plot looks quite strong.  We are now ready to simulate
+to examine how extreme certain reaction times are.
+
+```{r}
+samplefit_nodq <- simfit(gg3b_nodq)
+```
+
+The results below appear in table 4 of the supplement
+
+```{r}
+mean(samplefit_nodq < .10)
+mean(samplefit_nodq < .09)
+mean(samplefit_nodq < .08)
+```
+We see that these times are smaller when remove disqualifications than when we
+leave them in.  This is not surprising, as we know that some disqualifications
+come from values between 0 and 0.1.  Thus including those values is going to
+alter the model and raise the probability of athletes reacting in the 0 to 0.1
+range.

### 2024-11-30 Manuscript/supp.tex
Updated some small stuff and left a comment

@@
 from 92 athletes and the asymptotic test result was a p-value of
-$4.99 \cdot 10^{-12}$.  These are all highly significant test results that show
-substantial differences in average RT for athletes competing at
-multiple championship-level events.
+$4.99 \cdot 10^{-12}$.  As a result of the larger sample sizes, the code was
+very computationally expensive to run and thus we ran it on a smaller size.  All
+three permutation tests had a p-value of $1 \cdot 10^{-5}$, which is the smallest
+possible value given the permutation test size of $100,000$.  When taken together
+with the asymptotic results, the results becoems very clear. These are all
+highly significant test results that show substantial differences in average RT
+for athletes competing at multiple championship-level events.
 
@@
 \eds{with or without 2022?}
+\of{I dont know if we have time for this or if this fits the scope of what we
+are working on.}
 

### 2024-11-30 Manuscript/manuscript.tex
touch up; removed addressed comments

@@
 paper’s impact, limitations, and potential for future research.
-All data and code for our analysis are provided in the supplementary
-materials.
+All data and code for our analysis are provided in the Supplementary
+Materials.
 
@@
 Negative RTs were excluded from the analysis, while positive
-disqualified RTs were included due to their low frequency. A
-sensitivity analysis regarding the inclusion of positive disqualified
-RTs is provided in the Supplementary Material.
+disqualified RTs were included due to their low frequency and not
+being obvious outliters. % A
+% sensitivity analysis regarding the inclusion of positive disqualified
+% RTs is provided in the Supplementary Material.
 
@@
 100-meter dashes and their analyses were relegated to the
-supplementary material. We focus on the RTs recorded during
+Supplementary Material. We focus on the RTs recorded during
 semifinal and final heats only, as RTs from preliminary
@@
 Table~\ref{tab:Clusrankresults} presents the results from both permutation
-and asymptotic rank-based tests for these six comparisons. The tests reveal
-that the national versus international comparisons yielded the lowest
-p-values, indicating stronger evidence of faster RTs at the
-2022 World Championships compared to the 2022 national competitions. 
-\eds{This does not appear true based on the data currently in Table 1. 
-2022-2023 seem to have the lowest p-values.  Which is correct?}
-For both men and women, permutation test results were highly significant,
-providing substantial evidence that athletes competing at the 2022 World
-Championships had significantly faster RTs than they did just
-a few months earlier at national-level events. The comparisons between the
-2019 and 2022 World Championships and between the 2022 and 2023 World
-Championships also produced significant results. Collectively,
-these results suggest that athletes consistently achieved faster reaction
-times at the 2022 World Championships compared to other competitions. This
-finding supports the hypothesis that the 2022 World Championships presented
-conditions conducive to faster RTs, potentially due to systematic
-or environmental factors.
+and asymptotic rank-based tests for the six comparisons. All tests
+yielded very small p-values, even after applying a Bonferroni
+adjustment for multiple comparisons, indicating consistent evidence of
+faster RTs at the 2022 World Championships relative to other
+competitions. For both men and women, the national versus
+international comparisons showed that RTs at the 2022 World
+Championships were significantly faster than at national-level
+competitions held earlier that year. Similarly, comparisons between
+the 2019 and 2022 World Championships and between the 2022 and 2023
+World Championships produced significant results, reinforcing the
+observation that RTs at the 2022 World Championships were notably
+faster. These findings support the hypothesis that conditions at the
+2022 World Championships, whether systematic or
+environmental, contributed to consistently faster RTs.
 
@@
 distribution, respectively. Random effects account for variability at
-the venue level $\tau_v$ on the log-scale of the $\mu$ parameter
-and at the heat level $\tau_h$ on the log-scale of the $\sigma$
+the venue level on the log-scale of the $\mu$ parameter
+and at the heat level on the log-scale of the $\sigma$
 parameter in the density in Equation~\eqref{eq:gg}. The variance
@@
 the standard deviation of the venue-level random-effect variance,
-which increases from 0.043 to 0.058 . 
-\eds{are you reporting estimated variances or SDs of the random effects here and 
-in the Table?  In the table, we have $\tau_v$ and $\tau_h$ (SD) labels, not 
-$\tau_v^2$ and $\tau_h^2$ (variances).  Please check and correct as needed.}  
-These results
+which increases from 0.043 to 0.058. These results
 highlight that RTs are influenced by both venue and heat-level
@@
 indeed faster at the 2022 World Championships compared to comparable
-meets. 
-Ideally, a centralized database containing RTs from
+meets. Ideally, a centralized database containing RTs from
 all World Athletics-certified meets would have allowed for a more
@@
 and highlights the importance of tailoring thresholds to  biological
-distinctions. \eds{is this still true in the updated analyses?}
+distinctions.
 
@@
 \section*{Supplementary Material}
-\eds{when the supplement is complete, include an additional sentence here
-that summarizes the additional analyses presented in the supplemental material}
+Additional results are summarized in a supplement for rank-based
+comparison with pooled (men and women) data and for the GAMLSS results
+for women data.
 The data and R code used for the analysis are available in a compressed file for

### 2024-11-30 Manuscript/supp.tex
touch up; removed addressed comments

@@
 
-\jy{Keep the order consistent with the text; do the same in the code.}
-
 As an alternative to the methods described in Section 3.1 of the main paper, it
@@
 comparisons.  Thus we shrink our analyses from six to three, but each analysis
-is roughly twice as big as previously.  For the 2019 versus 2022 comparison,
-there were 258 RTs from 65 athletes and the asymptotic
-test result was a p-value of $3.56 \cdot 10^{-8}$. For the 2022 national versus
+is roughly twice as big as previously.
+For the 2022 national versus
 international comparison, there were 160 RTs from 35
 athletes and the asymptotic test result was a p-value of $1.94 \cdot 10^{-7}$.
+For the 2019 versus 2022 comparison,
+there were 258 RTs from 65 athletes and the asymptotic
+test result was a p-value of $3.56 \cdot 10^{-8}$. 
 For the 2023 versus 2022 comparison, there were 343 RTs
@@
 We can repeat our reaction time barrier analysis described in Section 3.2, but
-fit the same model to women's RT data from 2001 to 2023.  Once again we utilize
-the power of the venue effect within $\mu$ and a heat effect within $\sigma$.
-\eds{what is meant by `power' here?} 
+fit the same model to women's RT data from 2001 to 2023.
 The RT data for women is visualized in Figure~\ref{fig:WomensBoxplot}.
 Similar to the men's data, RTs from 2022 appear lower than in other
-years.
+years. After removing one obvious outliter, which was a disqualfied
+reaction time, the same model for men's data fits the women's data
+reasonably well.
 
@@
 \end{figure}
-
-We examined several models in the generalized Gamma model and found that the
-best model for women involved a venue effect in $\mu$ and heat effect in
-$\sigma$.  Thus the best model for men was also the best model for women.
 
@@
 
-One of the most interesting aspects of Table~\ref{tab:womensfit} is that while
+The fitted parameters in comparison with those from men's data are
+summarized in Table~\ref{tab:womensfit}.  One notable aspect is that while
 the venue effect standard deviation is nearly identical, the women's heat effect
-standard deviation is smaller and $\nu$ is much larger.  If men's races have
-more variability in times, it seems possible that in men's races that faster
-athletes can influence other athletes to react faster in certain instances.
-It is also interesting that the standard deviation of the venue effect is
-nearly identical for both men's and women's data, indicating that the venue
-effect is not only statistically significant but is consistent in strength
-across men and women.
-
-We can repeat the simulation methods described in the main paper to evaluate
-the probability of an extreme reaction time.  Table~\ref{tab:Sim_prob_women}
-compares the men's and women's results of observing reaction times less than
-0.08, 0.09, and 0.1 seconds.  We find that across all three comparisons that
-women have a lower probability of having a fast time. The results are slightly
-different from those from the men's data, which echoes existing studies
-reporting gender differences in RTs \citep[e.g.,][]{lipps2011implications,
-babicc2009reaction, panoutsakopoulos2020gender}.
+standard deviation is smaller and $\nu$ is much larger.
+This suggests that men’s races exhibit greater 
+variability in reaction times, possibly due to faster athletes influencing others 
+to react more quickly in certain instances. The consistency in the venue effect 
+standard deviation across men’s and women’s data indicates that the venue effect 
+is not only statistically significant but also consistent in magnitude across 
+genders. These findings highlight the potential impact of competition dynamics 
+on heat variability and the robustness of venue-level effects.
 
@@
 
-Now we need to perform alter the interpretation of the analyis to determine
-what a fair reaction time barrier for women should be.
-Table~\ref{tab:Sim_time_women} shows the difference in men's and women's
-suggested reaction time barrier.  What we see is that by World Athletics setting
-the same standard for men and women, their rules are not fair, in that men are
-more likely to be disqualified because they are more likely to react quickly.
-One of the biggest takeaways of the reaction time barrier analysis is that the
-current standards imposed by World Athletics are not grounded in statistical
-analysis and that they should be.  It seems reasonable to either raise the
-women's barrier so that they are disqualified at a similar rate as men or to
-lower the men's barrier so they are disqualified at a similar rate as women.  But
-having equal but not neccessarily fair standards does not seem reasonable when
-this study's results echoed results of numerous other papers asking World
-Athletics to change the standards.
+We can repeat the simulation methods described in the paper to evaluate
+the probability of an extreme reaction time.  Table~\ref{tab:Sim_prob_women}
+compares the men's and women's results of observing reaction times less than
+0.08, 0.09, and 0.1 seconds.  We find that across all three comparisons that
+women have a lower probability of having a fast time. The results are slightly
+different from those from the men's data, which echoes existing studies
+reporting gender differences in RTs \citep[e.g.,][]{lipps2011implications,
+babicc2009reaction, panoutsakopoulos2020gender}.
 
@@
 
-\section{Results from Data Excluding Positive Disqualified RTs}
+To evaluate a fair reaction time (RT) barrier for women, we compared suggested 
+barriers for men and women based on tail probabilities, as shown in 
+Table~\ref{tab:Sim_time_women}. The results indicate that men are more likely 
+to be disqualified under the current uniform 0.1-second threshold due to their 
+generally faster reaction times. This suggests that the same RT standard may not 
+have equivalent implications for men and women. Potential adjustments could 
+involve raising the barrier for women to align with men’s disqualification rates 
+or lowering the barrier for men to match women’s rates. However, further research 
+is needed to validate these findings and explore their broader implications. These 
+results contribute to ongoing discussions about RT thresholds and emphasize the 
+importance of statistical evidence in guiding decisions about competition fairness.
 
-An earlier iteration of the paper fit a model that did not include RTs from
-athletes who were disqualified or did not finish but still registered a RT.  
-However, it was ultimately decided to include these times to better estimate the 
-left tail of the distribution and more accurately predict the probability of a 
-low RT, as described in the main paper.  
-We did not include negative RTs, however, as these represent a mistake 
-of the runner for starting before
-the gun is fired and are thus meaningless in our objective to
-determine a fair RT barrier.  Not all of those disqualified were
-disqualified because of breaking the 0.1 reaction time barrier; there are many
-reasons why an athlete may be disqualified, with the most notable being failed
-drug tests and lane violations.  Nonetheless, in this section,
-we exclude all disqualified RTs to see their effect on the probability of an
-extreme RT.  We fit an identical model to the generalized Gamma model
-presented in Section 3.2 to examine differences results.
+% \section{Results from Data Excluding Positive Disqualified RTs}
 
-\begin{table}
-  \centering
-  \caption{Probabilities of observing RTs less than threshold 0.08,
-  0.09, and 0.10 seconds based on the
-    fitted GG GAMLSS model with both venue- and heat-level
-random effects.}
-  \begin{tabular}{c c c c}
-   \toprule
-   Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\
-   \midrule
-   Without DQs & $4.93\cdot10^{-5}$ & $3.53\cdot10^{-4}$ &  $1.97\cdot10^{-3}$  \\
-   With DQs & $6.84\cdot10^{-5}$ & $4.95\cdot10^{-4}$ & $2.76\cdot10^{-3}$ \\
-   \bottomrule
-  \end{tabular}
-  \label{tab:DQSim_probability}
-\end{table}
+% An earlier iteration of the paper fit a model that did not include RTs from
+% athletes who were disqualified or did not finish but still registered a RT.  
+% However, it was ultimately decided to include these times to better estimate the 
+% left tail of the distribution and more accurately predict the probability of a 
+% low RT, as described in the main paper.  
+% We did not include negative RTs, however, as these represent a mistake 
+% of the runner for starting before
+% the gun is fired and are thus meaningless in our objective to
+% determine a fair RT barrier.  Not all of those disqualified were
+% disqualified because of breaking the 0.1 reaction time barrier; there are many
+% reasons why an athlete may be disqualified, with the most notable being failed
+% drug tests and lane violations.  Nonetheless, in this section,
+% we exclude all disqualified RTs to see their effect on the probability of an
+% extreme RT.  We fit an identical model to the generalized Gamma model
+% presented in Section 3.2 to examine differences results.
 
-\jy{Men's data or pooled data?}
-\eds{with or without 2022?}
-\of{I dont know if we have time for this or if this fits the scope of what we
-are working on.}
+% \begin{table}
+%   \centering
+%   \caption{Probabilities of observing RTs less than threshold 0.08,
+%   0.09, and 0.10 seconds based on the
+%     fitted GG GAMLSS model with both venue- and heat-level
+% random effects.}
+%   \begin{tabular}{c c c c}
+%    \toprule
+%    Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\
+%    \midrule
+%    Without DQs & $4.93\cdot10^{-5}$ & $3.53\cdot10^{-4}$ &  $1.97\cdot10^{-3}$  \\
+%    With DQs & $6.84\cdot10^{-5}$ & $4.95\cdot10^{-4}$ & $2.76\cdot10^{-3}$ \\
+%    \bottomrule
+%   \end{tabular}
+%   \label{tab:DQSim_probability}
+% \end{table}
 
-Table~\ref{tab:DQSim_probability} shows the effect of removing disqualificated
-times from the analysis.  The probability of observing extreme RTs is lower when
-we remove the 17 observations.
+% \jy{Men's data or pooled data?}
+% \eds{with or without 2022?}
+% \of{I dont know if we have time for this or if this fits the scope of what we
+% are working on.}
+
+% Table~\ref{tab:DQSim_probability} shows the effect of removing disqualificated
+% times from the analysis.  The probability of observing extreme RTs is lower when
+% we remove the 17 observations.
 

### 2024-11-30 Manuscript/reply.tex
Responded to a few comments

@@
 \jy{These numbers need to be updated}
-slightly from 0.0029 to 0.0039, with similar parameter estimates for
+\of{Used 2 datasets: both men's dash and hurdles with 2022.  One dataset did not
+include DQs the other did. The numbers reflect the supplement results}
+slightly from 0.00197 to 0.00276, with similar parameter estimates for
 the random effects. Given these results, we opted to include these
@@
 
-We added additional explanations to address this in the introductions section.
-\jy{Owen: Which paragraph?}
+We added additional explanations in the second and third paragraph of the
+introduction to address this.
 
@@
 Thank you for pointing out the need for clearer interpretation of
-Table~2.
-\jy{Owen: please add standard errors of the three intercept  to the table.}
-The table now contains fitted parameters of the generalized Gamma
+Table~2. The table now contains fitted parameters of the generalized Gamma
 distribution GAMLSS as specified Equations (1)--(3). The table caption has been

### 2024-11-30 Archive/GLMM.Rmd
Copied from code folder

--- 2024-06-20 Code/GLMM.Rmd
@@
 ```{r}
-
 times <- read.csv("../Data/GLMMData.csv")
 colnames(times) <- c("Venue", "Type", "Time", "RxnTime", "Gender", "Heat", "Event")
-with2022 <- subset(times, (Type == "F" | Type == "S") & 
-                       Gender == "M" & Time != 'DQ' & Event == "Hurdles")
-without2022 <- subset(times, (Type == "F" | Type == "S") & 
-                           Gender == "M" & Time != 'DQ' & Venue != 2022 & Event == "Hurdles")
+
+with2022 <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles")
+
+without2022 <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS" & Time != "DQ" & Time != "D") |>
+  filter(RxnTime != 0) |>
+  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
+  filter(Venue != 2022)
 ```
@@
 This anova test determined that the two effect model is an improvement over the
-heat effect model. The p-value is 0.008305 which is significant at 
+heat effect model. The p-value is $1.21 \cdot 10^{-5}$ which is significant at 
 $\alpha = 0.05$.
@@
 improvement over the Venue effect model.
+
+We ran additional tests and found that including `Event` as a predictor in the
+model did not significantly improve the model.
 
@@
 ```
+
+```{r}
+men_dash <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "M") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime != 0) |>
+  mutate(RxnTime = ifelse(RxnTime < 0.1, 0.1, RxnTime))
+```

### 2024-11-30 Code/diag.R
Not Needed

@@ -1,13 +1,0 @@
-## After fitting
-## fit <- dash_and_hurdles_two_effect_gamma_log
-
-pit <- function(fit) {
-    shape <- 1/sigma(fit)^2
-    pgamma(fit@frame$RxnTime, shape = shape, rate = shape/fitted(fit))
-}
-
-pp <- pit(fit)
-
-library("qqconf")
-
-qq_conf_plot(pp, qunif)

### 2024-11-30 Code/gamlss.R
cleaned up code by removing unneccesarry, confusing, or duplicate code

@@ -121,60 +121,6 @@ qq_conf_plot(residuals(gig3b), dparams = list(mean = 0, sd = 1)) ## z-scores hav
 # While some of the AIC are strong and better than gg3b, the residuals are weak
 
 
-# Simulation Code: only works for gg3b
-sim_time <- function(model, RxnTime) {
-  simfit <- function(model, n = 10000000) {
-    
-    sd_Venue <- model$mu.coefSmo[[1]]$sigb
-    sd_Venue_heat <- model$sigma.coefSmo[[1]]$sigb[1]
-    
-    mu <- exp(model$mu.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue))
-    sigma <- exp(model$sigma.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue_heat))
-    nu <- model$nu.coefficients[1]
-    
-    x <- rGG(n, mu = mu, sigma = sigma, nu = nu)
-    x
-  }
-  x <- simfit(model)
-  print(mean(x < RxnTime))
-}
-sim_time(gg3b, 0.1)
-sim_time(gg3b, 0.09)
-sim_time(gg3b, 0.08)
-
-#Outputs recommended reaction time barrier based on inputted probability
-sim_prob <- function(model, prob) {
-  simfit <- function(model, n = 10000000) {
-    
-    sd_Venue <- model$mu.coefSmo[[1]]$sigb
-    sd_Venue_heat <- model$sigma.coefSmo[[1]]$sigb[1]
-    
-    mu <- exp(model$mu.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue))
-    sigma <- exp(model$sigma.coefficients[1] + rnorm(n, mean = 0 , sd = sd_Venue_heat))
-    nu <- model$nu.coefficients[1]
-    
-    x <- rGG(n, mu = mu, sigma = sigma, nu = nu)
-    x
-  }
-  x <- simfit(model)
-  print(quantile(x, probs = prob))
-}
-sim_prob(gg3b, .001)
-sim_prob(gg3b, .0001)
-
-
-## For reproducibility and to ensure that we got the same results. Run the function
-## that takes a GG model as input and prints all important parameters.
-model_print <- function(model) {
-  cat("AIC: ", AIC(model), "\n")
-  cat("mu (intercept): ", model$mu.coefficients[1], "\n")
[…]

### 2024-11-30 Code/gg.R
added model print function

@@ -39,6 +39,19 @@ qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1)) ## z-scores have
 qq_conf_plot(gg3b$sigma.coefSmo[[1]]$coef)
 qq_conf_plot(gg3b$mu.coefSmo[[1]]$coef)
 
+
+## For reproducibility and to ensure that we got the same results. Run the function
+## that takes a GG model as input and prints all important parameters.
+model_print <- function(model) {
+  cat("AIC: ", AIC(model), "\n")
+  cat("mu (intercept): ", model$mu.coefficients[1], "\n")
+  cat("sigma (intercept): ", model$sigma.coefficients[1], "\n")
+  cat("nu: ", model$nu.coefficients[1], "\n")
+  cat("sigma_v or std dev of venue effect: ", model$mu.coefSmo[[1]]$sigb, "\n")
+  cat("sigma_h or std dev of heat effect: ", model$sigma.coefSmo[[1]]$sigb[1], "\n")
+}
+model_print(gg3b)
+
 # Simulation Code: only works for gg3b
 simfit <- function(model, n = 10000000) {
     set.seed(1)

### 2024-11-30 Manuscript/supp.tex
added section 3 back

@@
 
-% \section{Results from Data Excluding Positive Disqualified RTs}
+\section{Results from Data Excluding Positive Disqualified RTs}
 
-% An earlier iteration of the paper fit a model that did not include RTs from
-% athletes who were disqualified or did not finish but still registered a RT.  
-% However, it was ultimately decided to include these times to better estimate the 
-% left tail of the distribution and more accurately predict the probability of a 
-% low RT, as described in the main paper.  
-% We did not include negative RTs, however, as these represent a mistake 
-% of the runner for starting before
-% the gun is fired and are thus meaningless in our objective to
-% determine a fair RT barrier.  Not all of those disqualified were
-% disqualified because of breaking the 0.1 reaction time barrier; there are many
-% reasons why an athlete may be disqualified, with the most notable being failed
-% drug tests and lane violations.  Nonetheless, in this section,
-% we exclude all disqualified RTs to see their effect on the probability of an
-% extreme RT.  We fit an identical model to the generalized Gamma model
-% presented in Section 3.2 to examine differences results.
+An earlier iteration of the paper fit a model that did not include RTs from
+athletes who were disqualified or did not finish but still registered a RT.  
+However, it was ultimately decided to include these times to better estimate the 
+left tail of the distribution and more accurately predict the probability of a 
+low RT, as described in the main paper.  We consider data from men's dash and
+hurdles and include 2022 in our analysis to try and determine how sensitive
+our model is to the inclusion of these times.
 
-% \begin{table}
-%   \centering
-%   \caption{Probabilities of observing RTs less than threshold 0.08,
-%   0.09, and 0.10 seconds based on the
-%     fitted GG GAMLSS model with both venue- and heat-level
-% random effects.}
-%   \begin{tabular}{c c c c}
-%    \toprule
-%    Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\
-%    \midrule
-%    Without DQs & $4.93\cdot10^{-5}$ & $3.53\cdot10^{-4}$ &  $1.97\cdot10^{-3}$  \\
-%    With DQs & $6.84\cdot10^{-5}$ & $4.95\cdot10^{-4}$ & $2.76\cdot10^{-3}$ \\
-%    \bottomrule
-%   \end{tabular}
-%   \label{tab:DQSim_probability}
-% \end{table}
+We did not include negative RTs, however, as these represent a mistake of the
+runner for starting before the gun is fired and are thus meaningless in our
+objective to determine a fair RT barrier.  Not all of those disqualified were
+disqualified because of breaking the 0.1 reaction time barrier; there are many
+reasons why an athlete may be disqualified, with the most notable being failed
+drug tests and lane violations.  Nonetheless, in this section,
+we exclude all disqualified RTs to see their effect on the probability of an
+extreme RT.  We fit an identical model to the generalized Gamma model
+presented in Section 3.2 to examine differences results.
 
-% \jy{Men's data or pooled data?}
-% \eds{with or without 2022?}
-% \of{I dont know if we have time for this or if this fits the scope of what we
-% are working on.}
+\begin{table}
+  \centering
+   \caption{Probabilities of observing RTs less than threshold 0.08,
+   0.09, and 0.10 seconds based on the
+     fitted GG GAMLSS model with both venue- and heat-level
+ random effects.}
+   \begin{tabular}{c c c c}
+    \toprule
+    Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\
+    \midrule
+    Without DQs & $4.93\cdot10^{-5}$ & $3.53\cdot10^{-4}$ &  $1.97\cdot10^{-3}$  \\
+    With DQs & $6.84\cdot10^{-5}$ & $4.95\cdot10^{-4}$ & $2.76\cdot10^{-3}$ \\
+    \bottomrule
+   \end{tabular}
+   \label{tab:DQSim_probability}
+ \end{table}
 
-% Table~\ref{tab:DQSim_probability} shows the effect of removing disqualificated
-% times from the analysis.  The probability of observing extreme RTs is lower when
-% we remove the 17 observations.
+Table~\ref{tab:DQSim_probability} shows the effect of removing disqualificated
+times from the analysis.  The probability of observing extreme RTs is lower when
+we remove the 17 observations.  While the probability of observing an extreme
+reaction time is less when we remove disqualfied athletes, many of the same
+conclusions remain the same: the current standards for disqualification are not
+grounded in statistical analysis, and there appear to be unequal standards for men
+and women.
 

### 2024-11-30 Code/ReactionBarrierAnalysis.Rmd
minor
added tukey hsd

@@
 ---
-title: "ReactionBarrierAnalysis"
+title: "Reaction Barrier Analysis"
 author: "Owen Fiore"
@@
 results in regards to the GAMLSS section of the paper.
+
 # Main Analysis
+
 The first thing we need is our data.  We create the `times` data to be the full
@@
 ```
-Let's visualize the data.
+
+Let's make sure we used appropriate data. There are three events in the data
+file for men: 100 meter dash, 110 meter hurdles, and 200 meter dash.  We
+suspect the results from the 200 meter dash might be slower as that is a longer
+event and thus reaction time plays a smaller role. We will use a TukeyHSD to
+examine any differences between the three events.
+```{r}
+library(car)
+TukeyHSD(aov(RxnTime ~ Event, data = all_events))
+```
+we see that there are relatively similar results for the 100 meter dash and
+110 meter hurdles as the p-value is 0.57 which is quite large.  But either of
+these two events when compared with the 200 meter dash results in a statistically
+significant p-value and thus we are not going to include this data in our
+analysis.
+
+Now that we have validated the decision to exclude certain data, let us 
+visualize it.
 ```{r}

### 2024-11-30 Manuscript/reply.tex
Updated p value from TukeyHSD

@@
 200-meter event revealed a significant effect of event type. Tukey’s
-Honest Significant Difference (HSD) test found a p-value of 0.0021 for
+Honest Significant Difference (HSD) test found a p-value of 0.00179 for
 the 200-meter dash compared to the 110-meter hurdles, and a p-value of
-0.000017 for the 100-meter dash compared to the 200-meter dash.
+0.00007 for the 100-meter dash compared to the 200-meter dash.
 These results indicate that reaction times in the 200-meter dash are

### 2024-11-30 Manuscript/manuscript.tex
Added a sentence to introduction
minor typo fixes to all

@@
 
-The 0.1-second disqualification threshold, introduced in the 1990s,
+Originally introduced in the 1990s to discourage athletes from attempting
+to predict the start gun, the 0.1-second disqualification threshold 
 has been the subject of significant debate.   Partly based on limited
@@
 disqualified RTs were included due to their low frequency and not
-being obvious outliters. % A
+being obvious outliers. % A
 % sensitivity analysis regarding the inclusion of positive disqualified
@@
 2022 data is included, all parameter estimates remain stable except
-the standard deviation of the venue-level random-effect variance,
+the standard deviation of the venue-level random-effect,
 which increases from 0.043 to 0.058. These results
@@
   \caption{Estimated fixed-effect parameters with standard errors in
-    parentheses and estimated variance of the random effects from the
+    parentheses and estimated standard deviations of the random effects from the
     fitted GG distribution with venue level random
@@
 
-This study sought to answer two primary questions: were athletes who
+This study sought to answer two primary questions: were RTs of athletes who
 competed at the 2022 World Track and Field Championships faster than
 at other races, and is the 0.1-second RT barrier fair?
-Regarding the first question, our analyses suggest that athletes were
-indeed faster at the 2022 World Championships compared to comparable
+Regarding the first question, our analyses suggest that athletes
+indeed had faster RTs at the 2022 World Championships compared to comparable
 meets. Ideally, a centralized database containing RTs from
@@
 be as extraordinary  as traditionally assumed. For men,
-Table~\ref{tab:Sim_time} shows that reaction  times below 0.1 seconds
+Table~\ref{tab:Sim_time} shows that RTs below 0.1 seconds
 occur with a probability of approximately one in 362 starts  when
@@
 probabilities of false starts. A similar pattern is observed for
-women, as detailed in the Supplementary Material, where reaction
-times below 0.1 seconds are exceedingly rare for the 100-meter dash
+women, as detailed in the Supplementary Material, where RTs
+ below 0.1 seconds are exceedingly rare for the 100-meter dash
 and 100-meter  hurdles. However, the uniformity of the 0.1-second
@@
 offering insights rather than drawing definitive conclusions about
-potential equipment malfunction. Our findings indicate that reaction
-times at the 2022 Championships were, on average, faster than at other
+potential equipment malfunction. Our findings indicate that RTs
+ at the 2022 World Championships were, on average, faster than at other
 competitions, as evidenced by the significant p-values in
@@
 \section*{Supplementary Material}
-Additional results are summarized in a supplement for rank-based
-comparison with pooled (men and women) data and for the GAMLSS results
-for women data.
+Additional results are summarized in a supplement for (1) rank-based
+comparison with pooled (men and women) data, (2) GAMLSS results
+for women's data, and (3) sensitivity of including positive yet disqualified
+reaction times in GAMLSS.
 The data and R code used for the analysis are available in a compressed file for

### 2024-11-30 Manuscript/supp.tex
minor typo fixes to all

@@
 possible value given the permutation test size of $100,000$.  When taken together
-with the asymptotic results, the results becoems very clear. These are all
+with the asymptotic results, the message becomes very clear. These are all
 highly significant test results that show substantial differences in average RT
-for athletes competing at multiple championship-level events.
+for athletes competing at multiple championship-level competitions.
 
@@
 
-We can repeat our reaction time barrier analysis described in Section 3.2, but
-fit the same model to women's RT data from 2001 to 2023.
+We also apply the RT barrier analysis described in Section 3.2 to women's data,
+fitting the same model to women's RT data from 2001 to 2023.
 The RT data for women is visualized in Figure~\ref{fig:WomensBoxplot}.
 Similar to the men's data, RTs from 2022 appear lower than in other
-years. After removing one obvious outliter, which was a disqualfied
+years. After removing one obvious outlier, which was a disqualified
 reaction time, the same model for men's data fits the women's data
@@
   \caption{Estimated fixed-effect parameters with standard errors in
-    parentheses and estimated variance of the random effects from the men's and
+    parentheses and estimated standard deviations of the random effects from the men's and
     women's fitted GG distribution with venue level random
@@
 This suggests that men’s races exhibit greater 
-variability in reaction times, possibly due to faster athletes influencing others 
+variability in RTs, possibly due to faster athletes influencing others 
 to react more quickly in certain instances. The consistency in the venue effect 
@@
 
-We can repeat the simulation methods described in the paper to evaluate
-the probability of an extreme reaction time.  Table~\ref{tab:Sim_prob_women}
-compares the men's and women's results of observing reaction times less than
-0.08, 0.09, and 0.1 seconds.  We find that across all three comparisons that
-women have a lower probability of having a fast time. The results are slightly
+We also repeat the simulation methods described in the paper to evaluate
+the probability of an extreme RT for women.  Table~\ref{tab:Sim_prob_women}
+compares the men's and women's results of observing RTs less than
+0.08, 0.09, and 0.1 seconds.  We find across all three thresholds that
+women have a lower probability of having a fast RT. The results are slightly
 different from those from the men's data, which echoes existing studies
@@
 
-To evaluate a fair reaction time (RT) barrier for women, we compared suggested 
+To evaluate a fair RT barrier for women, we compared suggested 
 barriers for men and women based on tail probabilities, as shown in 
@@
 to be disqualified under the current uniform 0.1-second threshold due to their 
-generally faster reaction times. This suggests that the same RT standard may not 
+generally faster RTs. This suggests that the same RT standard may not 
 have equivalent implications for men and women. Potential adjustments could 
@@
 left tail of the distribution and more accurately predict the probability of a 
-low RT, as described in the main paper.  We consider data from men's dash and
-hurdles and include 2022 in our analysis to try and determine how sensitive
-our model is to the inclusion of these times.
-
+low RT, as described in the main paper. 
 We did not include negative RTs, however, as these represent a mistake of the
@@
 reasons why an athlete may be disqualified, with the most notable being failed
-drug tests and lane violations.  Nonetheless, in this section,
+drug tests and lane violations.
+
+Nonetheless, in this section,
 we exclude all disqualified RTs to see their effect on the probability of an
 extreme RT. We fit an identical model to the generalized Gamma model
-presented in Section 3.2 to examine differences results.
+presented in Section 3.2.  As in Section 3.2, we considered data from men's dash and
+hurdles and include 2022 in our analysis to try and determine how sensitive
+our model is to the inclusion/exclusion of these times.
 
@@
 
-Table~\ref{tab:DQSim_probability} shows the effect of removing disqualificated
+Table~\ref{tab:DQSim_probability} shows the effect of removing disqualified (DQ)
 times from the analysis.  The probability of observing extreme RTs is lower when
 we remove the 17 observations.  While the probability of observing an extreme
-reaction time is less when we remove disqualfied athletes, many of the same
+RT is less when we remove disqualified athletes, many of the 
 conclusions remain the same: the current standards for disqualification are not

### 2024-12-02 Code/ReactionBarrierAnalysis.Rmd
final edits made just before submission

@@
 title: "Reaction Barrier Analysis"
-author: "Owen Fiore"
+author: "Blinded"
 date: "2024-11-28"
@@
 We considered many families of models, and formula structures within the
-generalized Gamma but thought that the results from the `gg3b` game strong
-results.  The first term corresponds to the $\mu$ term of the model, which is
+generalized Gamma, but found that `gg3b` gave strong results based on AIC.  
+The first term corresponds to the $\mu$ term of the model, which is
 a random effect for `Venue`. $\sigma$ contains a random effect for `Heat`.
@@
 We are going to see what happens when we pool men's and women's data together.
-We sill need to separate by year but will have 3 analyses instead of 6.  As a
+We still need to separate by year but will have 3 analyses instead of 6.  As a
 result of the larger data set we are going to run the permutation tests at a
-size of $n = 100,000$ rather than $n = 1,000,000$ as the code takes several
+size of $B = 100,000$ rather than $B = 1,000,000$ as the code takes several
 hours to run otherwise.  The asymptotic results are all highly statistically
@@
 ### 2019 vs 2022
-
-Need to talk about results after running code
 
@@
 
-### Women's Gamlss
+### Women's GAMLSS
 We decided to separate women's results after numerous papers suggested that
@@
 ```
-The standardized residual plot looks quite strong.  We are now ready to simulate
+The standardized residual plot looks quite good.  We are now ready to simulate
 to examine how extreme certain reaction times are.

### 2025-02-18 Manuscript/manuscript.tex
Re-organized by problem

@@
 
+%Needs to be looked at
 The rest of this paper is organized as follows. Section~\ref{sec:Data}
@@
 
-\section{Data}
-\label{sec:Data}
+\section{Analysis of the 2022 World Championships RTs}
+
+\subsection{Data}
+\label{sec:data_2022}
+
+To investigate whether RTs at the 2022 World Championships
+were significantly different from other competitions, we
+used data from male athletes in the 110-meter hurdles and 100-meter
+dash, and female athletes in the 100-meter hurdles and 100-meter dash,
+provided they competed in the 2022 World Championships and at least
+one other competition (2022 national championships, 2019 World
+Championships, or 2023 World Championships).  We excluded data from
+200-meter dashes and longer events, as their RTs were found
+to be significantly different. Negative RTs were excluded from the analysis,
+while positive disqualified RTs were included due to their low frequency and not
+being obvious outliers.
 
@@
 to be significantly different. This difference is not unexpected, as
-RT is generally less critical in longer sprint events
+RTs are generally less critical in longer sprint events
 compared to shorter distances where explosive starts are paramount.
@@
 % RTs is provided in the Supplementary Material.
-
-Two datasets were used to address the study's objectives. To
-investigate whether RTs at the 2022 World Championships
-were significantly different from other competitions (Objective 1), we
-used data from male athletes in the 110-meter hurdles and 100-meter
-dash, and female athletes in the 100-meter hurdles and 100-meter dash,
-provided they competed in the 2022 World Championships and at least
-one other competition (2022 national championships, 2019 World
-Championships, or 2023 World Championships). To assess the
-appropriateness of the 0.1-second RT threshold (Objective
-2), we used RTs from the 110-meter hurdles and 100-meter
-dash across all World Championships held from 1999 to 2023. This
-investigation began shortly after the 2022 World Championships, and we
-were pleased that including data from 2023 did not significantly
-alter our results \citep{WAData}.
-
-\subsection{Comparison Data with 2022 World Championships}
-\label{sec:databeyond}
 
@@
 
-\subsection{Data from World Championships 1999--2023}
-\label{sec:dataworld}
-
-The data for evaluating the appropriateness of the 0.1-second
-threshold was obtained from World Athletics and covers the men's
-110-meter hurdles and 100-meter dashes from 1999 to 2023. Due to
-possible gender differences \citep{babicc2009reaction,
-  lipps2011implications}, data for women's 100-meter hurdles and
-100-meter dashes and their analyses were relegated to the
-Supplementary Material. We focus on the RTs recorded during
-semifinal and final heats only, as RTs from preliminary
-heats are often not as fast as those in later heats
-\citep[e.g.,][]{collet1999strategic, tonnessen2013reaction,
-  brosnan2017effects, zhang2021correlation}. For analysis purposes, we
-pooled RTs from semifinal and final heats to increase sample
-size, which is particularly important for years with limited final heat
-observations. For example, in 2022, only five data points were available
-from the final heat due to two disqualifications and one athlete not
-competing. Unless otherwise noted, this pooled dataset forms the basis
-for our analyses for Objective 2. Additionally, we consider datasets
-that exclude 2022 to assess how our findings might differ when excluding
-this year of interest.
-
-The data is summarized in Figure~\ref{fig:Boxplot}, which presents a
-sequence of boxplots of RTs from 1999 to 2023. It is evident
-that RTs in 2022 were notably faster, with a median reaction
-time of 0.129 seconds compared to the 0.156 seconds observed in earlier
-studies, such as \citet{brosnan2017effects} for data spanning 1999 to 2014.
-Figure~\ref{fig:Boxplot} also highlights year-to-year variability in
-RTs, likely influenced by changes in the championship venue
-and environmental conditions such as humidity, precipitation, and
-elevation. Furthermore, advancements in technology and alterations to
-false start rules during the study period may have played a role in these
-variations \citep{willwacher2013novel}.
-
-\begin{figure}[tbp]
-  \centering
-  \includegraphics[width=\textwidth]{Boxplot}
-  \caption{The RTs from 1999 to 2023 for the men's 110 meter hurdle
-  and 100 meter dash.}
-  \label{fig:Boxplot}
-\end{figure}
-
-Between 2007 and 2009, World Athletics allowed one
-false start warning before disqualifying a sprinter \citep{iaaf2009falsestart}.
-This lenient rule led to 18 male and 7 female false starts at both the 2007
-and 2009 World Championships. In 2011, this rule was replaced with the stricter
-policy of automatic disqualification for false starts, aimed at reducing the
-delays caused by repeated warnings. This change reduced men’s false starts
-by two-thirds in 2011, with only six male and four female disqualifications
-\citep{iaaf2009falsestart}. \citet{haugen2013effect} demonstrated that more
-lenient false start rules significantly improved RTs during the
-1997–2009 period, suggesting that rule changes over the study period may
-have contributed to variations in RTs across years.
-
-\section{Methods} \label{sec:methods}
-
-The data described in Sections~\ref{sec:databeyond} and~\ref{sec:dataworld} were
-analyzed with rank-based methods and a GAMLSS, respectively.
-
-\subsection{Rank-based Comparison}\label{sec:rank}
+\subsection{Methods}
+\label{sec:methods_2022}
 
@@
 
-\subsection{GAMLSS}\label{sec:gamlss}
+\subsection{Results}
+\label{sec:results_2022}
+
+The rank-based methods described in Section~\ref{sec:rank} were used to
+compare RTs between the 2022 World Championships and other
+competitions in which the same athletes participated. These comparisons
+were conducted separately for men and women, resulting in six total
+comparisons: RTs from the 2022 national-level championships
+versus the 2022 World Championships for men and women, RTs
+from the 2019 versus 2022 World Championships for men and women, and
+RTs from the 2022 versus 2023 World Championships for men and
+women.
+
+\begin{table}
+  \centering
+  \caption{P-values of comparisons between
+    RTs from different competitions for the same athletes.
+    2022 Nat. vs Inter. compares RTs from 2022 national-level
+    championships and the 2022 World Track and Field Championships. 2019
+    vs 2022 compares RTs from the 2019 and 2022 World Track and
+    Field Championships. 2022 vs 2023 compares RTs from the
+    2022 and 2023 World Track and Field Championships.}
+  \begin{tabular}{c c c c c}
+   \toprule
+   Comparison & Permutation & Asymptotic & \# of athletes & \# RTs  \\
+   \midrule
+   2022 Nat. vs Inter. Men & $1.0 \cdot 10^{-6}$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
+   2022 Nat. vs Inter. Women & $1.0 \cdot 10^{-6}$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\[1ex]
+   2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
+   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\[1ex]
+   2022 vs 2023 Men & $1.0 \cdot 10^{-6}$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
+   2022 vs 2023 Female & $1.0 \cdot 10^{-6}$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
+   \bottomrule
+  \end{tabular}
+  \label{tab:Clusrankresults}
+\end{table}
+
+Table~\ref{tab:Clusrankresults} presents the results from both permutation
+and asymptotic rank-based tests for the six comparisons. All tests
+yielded very small p-values, even after applying a Bonferroni
+adjustment for multiple comparisons, indicating consistent evidence of
+faster RTs at the 2022 World Championships relative to other
+competitions. For both men and women, the national versus
+international comparisons showed that RTs at the 2022 World
+Championships were significantly faster than at national-level
+competitions held earlier that year. Similarly, comparisons between
+the 2019 and 2022 World Championships and between the 2022 and 2023
+World Championships produced significant results, reinforcing the
+observation that RTs at the 2022 World Championships were notably
+faster. These findings support the hypothesis that conditions at the
+2022 World Championships, whether systematic or
+environmental, contributed to consistently faster RTs.
+
+We also conducted the same analysis with men’s and women’s data
+pooled, yielding similar results. Details are provided in Section~1 of
+the Supplementary Material.
+
+\section{Analysis of the 0.1 RT Barrier}
+
+\subsection{Data}
+\label{sec:data_barrier}
+
+The data for evaluating the appropriateness of the 0.1-second
+threshold was obtained from World Athletics and covers the men's
+110-meter hurdles and 100-meter dashes from 1999 to 2023. Due to
+possible gender differences \citep{babicc2009reaction,
+  lipps2011implications}, data for women's 100-meter hurdles and
+100-meter dashes and their analyses were relegated to the
+Supplementary Material. We focus on the RTs recorded during
+semifinal and final heats only, as RTs from preliminary
+heats are often not as fast as those in later heats
+\citep[e.g.,][]{collet1999strategic, tonnessen2013reaction,
+  brosnan2017effects, zhang2021correlation}. For analysis purposes, we
+pooled RTs from semifinal and final heats to increase sample
+size, which is particularly important for years with limited final heat
+observations. For example, in 2022, only five data points were available
+from the final heat due to two disqualifications and one athlete not
+competing. Unless otherwise noted, this pooled dataset forms the basis
+for our analyses for Objective 2. Additionally, we consider datasets
+that exclude 2022 to assess how our findings might differ when excluding
+this year of interest. This investigation began shortly after the 2022 World
+Championships, and we were pleased that including data from 2023 did not
+significantly alter our results \citep{WAData}.
+
+The data is summarized in Figure~\ref{fig:Boxplot}, which presents a
+sequence of boxplots of RTs from 1999 to 2023. It is evident
+that RTs in 2022 were notably faster, with a median reaction
+time of 0.129 seconds compared to the 0.156 seconds observed in earlier
+studies, such as \citet{brosnan2017effects} for data spanning 1999 to 2014.
+Figure~\ref{fig:Boxplot} also highlights year-to-year variability in
+RTs, likely influenced by changes in the championship venue
+and environmental conditions such as humidity, precipitation, and
+elevation. Furthermore, advancements in technology and alterations to
+false start rules during the study period may have played a role in these
+variations \citep{willwacher2013novel}.
+
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\textwidth]{Boxplot}
+  \caption{The RTs from 1999 to 2023 for the men's 110 meter hurdle
+  and 100 meter dash.}
+  \label{fig:Boxplot}
+\end{figure}
+
+Between 2007 and 2009, World Athletics allowed one
+false start warning before disqualifying a sprinter \citep{iaaf2009falsestart}.
+This lenient rule led to 18 male and 7 female false starts at both the 2007
+and 2009 World Championships. In 2011, this rule was replaced with the stricter
+policy of automatic disqualification for false starts, aimed at reducing the
+delays caused by repeated warnings. This change reduced men’s false starts
+by two-thirds in 2011, with only six male and four female disqualifications
+\citep{iaaf2009falsestart}. \citet{haugen2013effect} demonstrated that more
+lenient false start rules significantly improved RTs during the
+1997–2009 period, suggesting that rule changes over the study period may
+have contributed to variations in RTs across years.
+
+\subsection{Methods}
+\label{sec:methods_barrier}
 
@@
 
-\section{Results} \label{sec:Results}
-
-\subsection{Rank-based Comparison} \label{subsec:Results_Rank}
-
-The rank-based methods described in Section~\ref{sec:rank} were used to
-compare RTs between the 2022 World Championships and other
-competitions in which the same athletes participated. These comparisons
-were conducted separately for men and women, resulting in six total
-comparisons: RTs from the 2022 national-level championships
-versus the 2022 World Championships for men and women, RTs
-from the 2019 versus 2022 World Championships for men and women, and
-RTs from the 2022 versus 2023 World Championships for men and
-women.
-
-\begin{table}
-  \centering
-  \caption{P-values of comparisons between
-    RTs from different competitions for the same athletes.
-    2022 Nat. vs Inter. compares RTs from 2022 national-level
-    championships and the 2022 World Track and Field Championships. 2019
-    vs 2022 compares RTs from the 2019 and 2022 World Track and
-    Field Championships. 2022 vs 2023 compares RTs from the
-    2022 and 2023 World Track and Field Championships.}
-  \begin{tabular}{c c c c c}
-   \toprule
-   Comparison & Permutation & Asymptotic & \# of athletes & \# RTs  \\
-   \midrule
-   2022 Nat. vs Inter. Men & $1.0 \cdot 10^{-6}$ & $ 6.1 \cdot 10^{-5}$ & 17 & 80 \\
-   2022 Nat. vs Inter. Women & $1.0 \cdot 10^{-6}$ & $ 1.2 \cdot 10^{-3}$ & 17 & 80 \\[1ex]
-   2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
-   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\[1ex]
-   2022 vs 2023 Men & $1.0 \cdot 10^{-6}$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
-   2022 vs 2023 Female & $1.0 \cdot 10^{-6}$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
-   \bottomrule
-  \end{tabular}
-  \label{tab:Clusrankresults}
-\end{table}
-
-Table~\ref{tab:Clusrankresults} presents the results from both permutation
-and asymptotic rank-based tests for the six comparisons. All tests
-yielded very small p-values, even after applying a Bonferroni
-adjustment for multiple comparisons, indicating consistent evidence of
-faster RTs at the 2022 World Championships relative to other
-competitions. For both men and women, the national versus
-international comparisons showed that RTs at the 2022 World
-Championships were significantly faster than at national-level
-competitions held earlier that year. Similarly, comparisons between
-the 2019 and 2022 World Championships and between the 2022 and 2023
-World Championships produced significant results, reinforcing the
-observation that RTs at the 2022 World Championships were notably
-faster. These findings support the hypothesis that conditions at the
-2022 World Championships, whether systematic or
-environmental, contributed to consistently faster RTs.
-
-We also conducted the same analysis with men’s and women’s data
-pooled, yielding similar results. Details are provided in Section~1 of
-the Supplementary Material.
-
-\subsection{GAMLSS} \label{subsec:Results_GLMM}
+\subsection{Results}
+\label{sec:results_barrier}
 

### 2025-02-18 Manuscript/reply2.tex
Made reply 2 doc

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\usepackage{hyperref}

\usepackage{listings}
\usepackage{rotating, graphicx}
\graphicspath{{./}, {./image/}}
\usepackage{booktabs, natbib}
% \usepackage{breakurl}
% \usepackage [english]{babel}
\usepackage{amsmath, amsbsy, amsthm, epsfig, epsf, psfrag, graphicx,
amssymb, enumerate}
\usepackage{bm}
\usepackage{multirow, multicol}

\usepackage[dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.1, 0.2, 0.6}

\newcommand{\jy}[1]{\textcolor{orange}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{blue}{(EDS: #1)}}
\newcommand{\of}[1]{\textcolor{purple}{(OF: #1)}}

\sloppy

% \usepackage{csquotes}
% \usepackage [autostyle, english = american]{csquotes}
% \MakeOuterQuote{"}

% \usepackage{bibentry}
\newenvironment{comment}%
{\begin{quotation}\noindent\small\it\color{darkblue}\ignorespaces%
}{\end{quotation}}

\begin{document}

\begin{center}
  {\Large\bf TAS-23-296.R1: Response to the Comments}
\end{center}

We sincerely thank the Editor and Associate Editor for tentatively accepting our
submission and are proud to present our most recent revisions.

The manuscript has been revised accordingly with the following most
notable changes:
\begin{enumerate}
\item
  The focus of the study has been clarified in the completely
  rewritten introduction. The two objectives are 1) to investigate
  whether the reaction times at the 2022 World Championships were
  significantly different from the other competitions; and 2) to
  assess the appropriateness of the 0.1-second threshold.
\item
  Additional data

Point-by-point responses to the comments are as follows, with the
comments quoted in \emph{\color{darkblue} italic and blue}.
\end{enumerate}

\subsection*{To Reviewer 1}

\begin{comment}
This revision is a solid improvement on the original paper exploring the
anomalous reaction times at the 2022 World Championships and the implications of
using 0.100 seconds as the cut off for acceptable reaction times. The updated
paper addresses many of my original points of feedback, including, notably, data
from more events, improving the statistical power and generalizability of the
results. I also think the expansion of the features considered in the model is
beneficial (although, I do think considering a Bayesian approach to the
distribution estimation could be beneficial).

However, I still believe there is room to improve the overall quality of the
paper, both from a technical/methodological perspective and from a more general
clarity and organization perspective.

\end{comment}

We thank the reviewer for the accurate summary of our contributions

\begin{comment}
First, I think the description of the data would benefit from an easier to
follow organization system. Currently, headings repeat and the overall
difference between the usage of data from the World Championships in one setting
instead of the other is not as easily understood as it could or probably should
be. I would recommend a more intuitive organization of this material and, if
possible, the construction and inclusion of a table summarizing the various data
details in each analysis. As a reader, I would like a fast way to keep track of
the important details in each case.
\end{comment}

\begin{comment}
Second, I think the results themselves would benefit from organizing around the
question or problem being addressed instead of the method. Organizing by method
makes it harder for your reader to connect the results to the research questions
or prompts.
\end{comment}

Thank you for this suggestion, we have decided to re-organize our paper as a
result. Now, the first half of the paper is dedicated to the analysis of
athletes competing faster in 2022 and the second half is dedicated to the
analysis of the reaction time barrier, with an introduction and discussion to
show the connectivity of these problems.

\begin{comment}
Third, I think the manuscript as a whole needs some general editing and cleaning
up before publication. For example, there are discrepancies in the language used
in some of the tables - in Table 1 the groups are labeled both "Men/Women" and
"Men/Female" when consistency would be helpful for the readers. Similarly, Table
2 uses the model notation to as column headers, making it hard to quickly assess
what it is communicating. A stronger organizational approach could also help the
overall clarity of the article.
\end{comment}

\begin{comment}
Finally, I think it is important that the article consider the possible causes
of differences between the competitions that are not accounted for in the
matched-pairs design. As I understand the approach, the analysis does not
account for two important potential sources of change in the reaction times::
seasonality and age. Seasonality could come into play if athletes' reaction
times improve as they attempt to peak in their season (likely at the highest
level competition they will participate in). If athletes expect to compete at
the World Championships, they might not be as fit or prepared at their national
championships and produce slower reaction times. Similarly, age could lead to
differences in an athletes' reaction times as older athletes will react slower
than younger athletes. While it would be great to see these features accounted
for directly, I think at a minimum they and their likely influences on the
results, should be discussed in the overall limitations of the analysis.
\end{comment}

\begin{comment}
Altogether, I think the technical work here is solid and shows that the reaction
times at the 2022 World Championships were identifiably faster than at other
major competitions and that 0.100 threshold for acceptable reaction times is
likely too conservative of a standard given athletes' natural abilities. The
largest areas of need are mainly in the organization and overall clarity of the
paper.
\end{comment}

\bibliographystyle{apalike}
\bibliography{citations}
\end{document}

### 2025-02-22 Manuscript/manuscript.tex
Added seasonality/age to discussion and cleaned up some phrases

@@
 %Needs to be looked at
-The rest of this paper is organized as follows. Section~\ref{sec:Data}
-describes the data collection and datasets used for the two
-objectives. Section~\ref{sec:rank} outlines the rank-based comparison
-methods for objective~1, rank-based comparison for clustered data with
-subunit-level grouping to analyze differences in RTs across different
-competitions. Section~\ref{sec:gamlss} details the methods for
-objective~2, the GG model with random effects to evaluate the appropriatenes
-of the 0.1-second threshold. Section~\ref{sec:Results} presents the
-results of both objectives and discusses the implications.
+The rest of this paper is organized as follows. Section~\ref{sec:problem1}
+investigates RTs of athletes who competed at the 2022 World Championships to
+examine differences between other competitions.  Section~\ref{sec:problem2}
+investigates RTs of athletes from 1999 to 2023 to determine a reaction barrier
+ground in statistical analysis.  Within each of the above sections, the data,
+methods used, and the results are presented.
 Finally, Section~\ref{sec:concludingremarks} highlights the
@@
 \section{Analysis of the 2022 World Championships RTs}
+\label{sec:problem1}
 
@@
 
-The rank-based methods described in Section~\ref{sec:rank} were used to
+The rank-based methods described in Section~\ref{sec:methods_2022} were used to
 compare RTs between the 2022 World Championships and other
@@
    2019 vs 2022 Men & $2.8 \cdot 10^{-5}$ & $1.1 \cdot 10^{-5}$ & 34 & 134 \\
-   2019 vs 2022 Female & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\[1ex]
+   2019 vs 2022 Women & $ 1.5 \cdot 10^{-3}$ & $6.9 \cdot 10^{-3}$ & 31 & 124 \\[1ex]
    2022 vs 2023 Men & $1.0 \cdot 10^{-6}$ & $1.4 \cdot 10^{-6}$ & 45 & 161 \\
-   2022 vs 2023 Female & $1.0 \cdot 10^{-6}$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
+   2022 vs 2023 Women & $1.0 \cdot 10^{-6}$ & $9.4 \cdot 10^{-7}$ & 47 & 182 \\
    \bottomrule
@@
 \section{Analysis of the 0.1 RT Barrier}
+\label{sec:problem2}
 
@@
 indeed had faster RTs at the 2022 World Championships compared to comparable
-meets. Ideally, a centralized database containing RTs from
+meets.  Our findings are strengthened by having multiple comparisons: athletes
+who competed in 2019 and 2022 got older over that timespan but we included an
+analysis in which athletes would have been younger in 2022 and then older in
+2023.  There are validly raised concerns that athletes who compete at their
+national level competitions in 2022 may not react at their fastest due to a less
+competitive environment or because they would try to peak for a more important
+championship.  However, the athletes who competed in 2019 and 2023 are not
+impacted by this factor, so its effect is minimzed to just one subset of a
+broader analysis.
+
+Ideally, a centralized database containing RTs from
 all World Athletics-certified meets would have allowed for a more

### 2025-02-22 Manuscript/reply2.tex
Wrote replies to comments, needs to be revised

@@
 \item
-  The focus of the study has been clarified in the completely
-  rewritten introduction. The two objectives are 1) to investigate
-  whether the reaction times at the 2022 World Championships were
-  significantly different from the other competitions; and 2) to
-  assess the appropriateness of the 0.1-second threshold.
+  The manuscript has been re-structured to have 4 sections: introduction,
+  analysis of 2022 times, analysis of a proper reaction barrier, and discussion.
+  Within the second and third sections there are data, methods, and results
+  subsections.
 \item
-  Additional data
+  General revisions of tidying the manuscript.
 
@@
 
-We thank the reviewer for the accurate summary of our contributions
+We thank the reviewer for the accurate summary of our contributions and
+understand there are some smaller issues to still address.
 
@@
 \end{comment}
+
+We understand there were issues about distinguishing between the datasets being
+used for each section.  After re-grouping the paper, each problem's data, 
+methods, and results section are consecutive, leading to increased reader
+comprehension.  The dataset describing the 2022 analysis is found in
+the last paragraph of page 5 and the dataset describing the RT barrier analysis
+is found in the first paragraph of section 3.1
 
@@
 
+The manuscript has been revised again and the langugage in Table 1 is now
+consistent.  The Table 2 nomenclature is now easier to interpret because the
+methods section describing the symbols is now placed on the page before.
+
 \begin{comment}
@@
 
+We have added additional explanations in the discussion section to address both
+seasonality and age, although we don't think either one will be largely
+impactful.  By including times from athletes in both 2019 and 2022, there are
+athletes who got older from 2019 to 2022 but athletes in 2022 will be younger
+than they were in 2023.  Including both comparisons is meaningful as it helps
+diminish this effect.  Likewise while seasonality may impact the 2022 national
+level comparison, the 2019 and 2023 comparisons mitigate this factor.
+
 \begin{comment}
@@
 
+Thank you for your comments.
+
 \bibliographystyle{apalike}

### 2025-02-23 Manuscript/reply2.tex
smoothed responses

--- 2025-02-18 Manuscript/reply2.tex
@@
 
-We sincerely thank the Editor and Associate Editor for tentatively accepting our
-submission and are proud to present our most recent revisions.
+We sincerely thank the Editor for tentatively accepting our submission
+and Reviewer~1 for the additional suggestions. We are pleased to
+present our final revisions.
 
-The manuscript has been revised accordingly with the following most
-notable changes:
+The manuscript has been revised as follows, with the most notable
+changes being:
 \begin{enumerate}
 \item
-  The focus of the study has been clarified in the completely
-  rewritten introduction. The two objectives are 1) to investigate
-  whether the reaction times at the 2022 World Championships were
-  significantly different from the other competitions; and 2) to
-  assess the appropriateness of the 0.1-second threshold.
+  Restructuring: The manuscript is now organized into four
+  sections: Introduction, Analysis of 2022 Times, Analysis of a Proper
+  RT Barrier, and Discussion. The second and third sections
+  further include subsections for Data, Methods, and Results to
+  enhance clarity and readability.
 \item
-  Additional data
+  General Refinements: The manuscript has been carefully revised to
+  improve clarity, coherence, and overall presentation.
+\end{enumerate}
 
 Point-by-point responses to the comments are as follows, with the
-comments quoted in \emph{\color{darkblue} italic and blue}.
-\end{enumerate}
+original comments quoted in \emph{\color{darkblue} italic blue}.
 
@@
 
-We thank the reviewer for the accurate summary of our contributions
+Thank you for the thoughtful feedback and for recognizing the
+improvements in our revision. The remaining points are well taken, and
+we address them below.
 
@@
 
+Distinguishing between the datasets used in each section was a valid
+concern. To improve clarity, the paper has been restructured so that
+each problem’s data, methods, and results appear consecutively, making
+it easier to follow. The dataset for the 2022 analysis is now
+described in the second paragraph of Section~2.1.1, while the dataset
+for the RT barrier analysis is introduced in the first paragraph of
+Section~3.1.
+
 \begin{comment}
@@
 
-Thank you for this suggestion, we have decided to re-organize our paper as a
-result. Now, the first half of the paper is dedicated to the analysis of
-athletes competing faster in 2022 and the second half is dedicated to the
-analysis of the reaction time barrier, with an introduction and discussion to
-show the connectivity of these problems.
+This was an excellent suggestion, and we appreciate the insight. To
+improve clarity and better align the results with the research
+questions, the paper has been reorganized. Section~2 now focuses
+on whether athletes competed faster in 2022, while Section~3
+examines the RT barrier. The Introduction and Discussion
+sections provide context and highlight the connection between these
+two analyses.
 
@@
 \end{comment}
+
+The manuscript has been further refined for clarity and
+consistency. The language in Table~1 is now uniform. With the
+restructuring, the definitions of the symbols in Table~2 are now
+included within the same section, making them easier to connect.
 
@@
 
+\jy{Admit the limitation of the study in the discussion and point to
+  the location of the addition.}
+
+We have added additional explanations in the discussion section to address both
+seasonality and age, although we don't think either one will be largely
+impactful.  By including times from athletes in both 2019 and 2022, there are
+athletes who got older from 2019 to 2022 but athletes in 2022 will be younger
+than they were in 2023.  Including both comparisons is meaningful as it helps
+diminish this effect.  Likewise while seasonality may impact the 2022 national
+level comparison, the 2019 and 2023 comparisons mitigate this factor.
+
 \begin{comment}
@@
 
+We appreciate the thoughtful feedback and are glad the technical work
+is recognized as solid. The revisions have focused on improving the
+organization and clarity of the paper to better communicate the key
+findings.
+
 \bibliographystyle{apalike}

### 2025-02-23 Manuscript/manuscript.tex
revised discussion paragraph 1

@@
 
-This study sought to answer two primary questions: were RTs of athletes who
-competed at the 2022 World Track and Field Championships faster than
-at other races, and is the 0.1-second RT barrier fair?
-Regarding the first question, our analyses suggest that athletes
-indeed had faster RTs at the 2022 World Championships compared to comparable
-meets.  Our findings are strengthened by having multiple comparisons: athletes
-who competed in 2019 and 2022 got older over that timespan but we included an
-analysis in which athletes would have been younger in 2022 and then older in
-2023.  There are validly raised concerns that athletes who compete at their
-national level competitions in 2022 may not react at their fastest due to a less
-competitive environment or because they would try to peak for a more important
-championship.  However, the athletes who competed in 2019 and 2023 are not
-impacted by this factor, so its effect is minimzed to just one subset of a
-broader analysis.
+This study first examined whether reaction times (RTs) at the 2022
+World Track and Field Championships were significantly faster than at
+other competitions. Our analyses indicate that RTs at the 2022 World
+Championships were consistently faster than those recorded at both
+national-level competitions earlier in the same year and the 2019 and
+2023 World Championships. The persistence of this pattern across
+different comparison groups suggests that these differences are not
+due to random variation or individual improvements over time. A more
+comprehensive analysis would benefit from a centralized database
+containing RTs from all World Athletics-certified meets, but such data
+are not consistently available. However, by incorporating competitions
+from multiple years (2019, 2022, and 2023), the analysis accounts for
+potential confounding factors such as seasonality and age, as athletes
+at different stages of their careers are represented in different
+comparisons.
 
-Ideally, a centralized database containing RTs from
-all World Athletics-certified meets would have allowed for a more
-comprehensive analysis. Currently, however, RTs are seldom
-included in the results available on the World Athletics website. This
-limitation constrained our ability to evaluate athletes like Devon
-Allen, as data from his races leading up to the World Championships in
-2022 were not accessible. A more robust dataset would enable deeper
-insights by allowing comparisons across a larger set of races.
-
-In addressing the second question, our analyses of the GAMLSS model
-reveal that the 0.1-second RT barrier, while rare, may not
+This study further assessed whether the 0.1-second RT threshold is a
+fair standard for disqualification. Our analyses of the GAMLSS model
+suggest that while RTs below 0.1 seconds are rare, they may not be
 be as extraordinary as traditionally assumed. For men,

### 2025-02-23 Manuscript/reply2.tex
revised discussion paragraph 1

@@
 
-\jy{Admit the limitation of the study in the discussion and point to
-  the location of the addition.}
-
-We have added additional explanations in the discussion section to address both
-seasonality and age, although we don't think either one will be largely
-impactful.  By including times from athletes in both 2019 and 2022, there are
-athletes who got older from 2019 to 2022 but athletes in 2022 will be younger
-than they were in 2023.  Including both comparisons is meaningful as it helps
-diminish this effect.  Likewise while seasonality may impact the 2022 national
-level comparison, the 2019 and 2023 comparisons mitigate this factor.
+We appreciate this thoughtful comment and have addressed it in the
+discussion section. The revised text clarifies that comparisons across
+multiple years (2019, 2022, and 2023) help mitigate the effects of
+seasonality and age, as athletes contribute to different pairings
+across years. Additionally, while seasonal peaking may influence RTs
+at national competitions, the inclusion of both national and
+international events provides a broader competitive context, reducing
+the likelihood that seasonality alone explains the observed
+differences. This clarification has been added to the first paragraph
+of the discussion.
 
@@
 
-We appreciate the thoughtful feedback and are glad the technical work
+We are delighted that the technical work
 is recognized as solid. The revisions have focused on improving the

### 2025-02-23 Manuscript/manuscript.tex
comments on section 2 and 3

@@
 
+\jy{State the hypotheses here before diving into data/methods/results.}
+
 \subsection{Data}
@@
 
+\jy{This two paragraphs are repetitive. Condense.}
 To investigate whether RTs at the 2022 World Championships
@@
 \label{sec:datanational}
+
+\jy{Opening sentence not accurate. This is the data to test the first
+  hypothesis: national versus World Championship 2022.}
 
@@
 
+\jy{Improve the topic sentence: this data is for the second
+  hypothesis: 2019 vs 2022; 2023 vs 2022.}
 To compare RTs at the 2022 World Championships with those
@@
 \label{sec:problem2}
+
+\jy{write a quick overview paragraph here.}
 

### 2025-02-27 Manuscript/manuscript.tex
responded to Prof Yan's comments

@@
 \jy{State the hypotheses here before diving into data/methods/results.}
+\of{Like this?  I feel like the first paragraph of the data section serves as
+a good introductino to this section. I dont' know what to write without
+re-writing the introduction or what is discussed in the first paragraph of 2.1}
+
+This investigation began to determine whether athletes who competed at the 2022
+Championships had significantly different RTs to when those same athletes
+competed in comparable and relevant events.
+
+\of{or like this?}
+
+Within our investigation to determine if athletes performed abnormally in 2022,
+we will analysze times from athletes who competed in earlier meets in 2022 
+or from another World Championship meet in 2019 or 2023.
 
@@
 \jy{This two paragraphs are repetitive. Condense.}
+\of{Resolved but needs to be proof read}
 To investigate whether RTs at the 2022 World Championships
@@
 one other competition (2022 national championships, 2019 World
-Championships, or 2023 World Championships).  We excluded data from
-200-meter dashes and longer events, as their RTs were found
-to be significantly different. Negative RTs were excluded from the analysis,
-while positive disqualified RTs were included due to their low frequency and not
-being obvious outliers.
-
-This study focuses on RTs from sprint events where reaction
-time plays a critical role. Specifically, we included data from men’s
-100-meter dashes and 110-meter hurdles, as well as women’s 100-meter
-dashes and 100-meter hurdles. Statistical tests showed no significant
-differences in RTs between these events, supporting their
+Championships, or 2023 World Championships). Statistical tests showed no
+significant differences in RTs between these events, supporting their
 inclusion in a unified analysis. However, we excluded data from
@@
 compared to shorter distances where explosive starts are paramount.
-Negative RTs were excluded from the analysis, while positive
-disqualified RTs were included due to their low frequency and not
-being obvious outliers. % A
+Negative RTs were excluded from the analysis, while positive disqualified RTs
+were included due to their low frequency and not being obvious outliers.
+
+  % A
 % sensitivity analysis regarding the inclusion of positive disqualified
@@
   hypothesis: national versus World Championship 2022.}
+\of{Is it better now?}
 
-To assess the abnormality of the 2022 World Championships, we examined
-data beyond World Championship events. This decision was motivated by
-an exploratory analysis of how United States (US) athletes performed
-at the 2022 US Track and Field Championships, held from June 23--26,
+%To assess the abnormality of the 2022 World Championships, we examined
+%ata beyond World Championship events. This decision was motivated by
+
+To determine if athletes reacted differently at the 2022 World Championships,
+our first places that we look are 2022 national championships.
+Prior to a formal analysis, we examined how United States (US) athletes
+performed at the 2022 US Track and Field Championships, held from June 23--26,
 2022, at Hayward Field in Eugene, Oregon. Since this venue also hosted
@@
   hypothesis: 2019 vs 2022; 2023 vs 2022.}
-To compare RTs at the 2022 World Championships with those
-recorded in the 2019 and 2023 World Championships, we prepared datasets
-including RTs from athletes who competed in the 2022 World
-Championships and at least one of the 2019 or 2023 World Championships.
-Specifically, we identified athletes who participated in both the 2019
-and 2022 World Championships, as well as those who participated in both
-the 2022 and 2023 World Championships. In the 2019-2022 comparison,
+\of{Is this more along the lines of what you are looking for}
+
+This data was compiled to test our second focus, athletes who competed at
+the 2022 World Championships and the 2019 or 2023 World Championships. 
+In the 2019-2022 comparison,
 RTs from 2022 were treated as the `treatment' group, with
@@
 \jy{write a quick overview paragraph here.}
+\of{Is it too much of a re-telling of the introduction?}
+
+In addition to the work presented in Section~\ref{sec:problem1} we thought that
+it was relevant and important to expand on prior work that focused on an
+appropriate reaction time barrier.  Implemented by World Athletics to discourage
+false starts, athletes with RTs below 0.1 seconds are disqualified.  In this
+section we will attempt to model historical data to recommend a reaction barrier
+based in statistical analysis.
 

### 2025-03-04 Manuscript/manuscript.tex
opening paragraphs of sections 2 and 3; additional smoothing

--- 2025-02-23 Manuscript/manuscript.tex
@@
 methods used, and the results are presented.
-Finally, Section~\ref{sec:concludingremarks} highlights the
+Finally, Section~\ref{sec:disc} highlights the
 paper’s impact, limitations, and potential for future research.
@@
 
-\section{Analysis of the 2022 World Championships RTs}
+\section{Assessing the 2022 World Championships RTs}
 \label{sec:problem1}
+
+The notable faster RTs at the 2022 World Championships
+raised concerns about systematic biases. 
+For example,  the median RTs across multiple sprint events were found
+to be the lowest in recent history; the number of RTs recorded under
+0.115 at men’s 100m dash and 110m hurdles was 25, much greater than that in
+2019, which was 3 \citep{johnson2022was}. This striking disparity
+suggests a potential systematic difference in RT measurements at the
+2022 World Championships, necessitating a formal statistical
+investigation. We assess whether RTs at the 2022 World Championships
+were systematically faster by comparing them against (1) national
+competitions in 2022 and (2) RTs from the 2019 and 2023 World
+Championships to evaluate deviations from historical norms, using data
+from athletes who competed in both events in the comparison.
 
@@
 one other competition (2022 national championships, 2019 World
-Championships, or 2023 World Championships).  We excluded data from
-200-meter dashes and longer events, as their RTs were found
-to be significantly different. Negative RTs were excluded from the analysis,
-while positive disqualified RTs were included due to their low frequency and not
-being obvious outliers.
-
-This study focuses on RTs from sprint events where reaction
-time plays a critical role. Specifically, we included data from men’s
-100-meter dashes and 110-meter hurdles, as well as women’s 100-meter
-dashes and 100-meter hurdles. Statistical tests showed no significant
-differences in RTs between these events, supporting their
+Championships, or 2023 World Championships). Statistical tests showed no
+significant differences in RTs between these events, supporting their
 inclusion in a unified analysis. However, we excluded data from
 200-meter dashes and longer events, as their RTs were found
-to be significantly different. This difference is not unexpected, as
-RTs are generally less critical in longer sprint events
-compared to shorter distances where explosive starts are paramount.
-Negative RTs were excluded from the analysis, while positive
-disqualified RTs were included due to their low frequency and not
-being obvious outliers. % A
-% sensitivity analysis regarding the inclusion of positive disqualified
-% RTs is provided in the Supplementary Material.
+to be significantly different. This exclusion is expected, as reaction
+time plays a smaller role in longer sprint events, where acceleration
+off the blocks is less decisive.
+Negative RTs were excluded from the analysis, while positive disqualified RTs
+were included due to their low frequency and not being obvious outliers.
 
@@
 
-To assess the abnormality of the 2022 World Championships, we examined
-data beyond World Championship events. This decision was motivated by
-an exploratory analysis of how United States (US) athletes performed
-at the 2022 US Track and Field Championships, held from June 23--26,
+We first compare RTs from the 2022 World Championships to national
+competitions earlier that year to assess whether the same athletes
+reacted differently across events.
+Prior to a formal analysis, we examined how United States (US) athletes
+performed at the 2022 US Track and Field Championships, held from June 23--26,
 2022, at Hayward Field in Eugene, Oregon. Since this venue also hosted
 the 2022 World Championships in August, it
-provided a unique opportunity to compare. There were four US
-110-meter hurdle athletes—Trey Cunningham, Daniel Roberts,
-Grant Holloway, and Devon Allen—who competed in both events. All four
-athletes recorded faster RTs in every World Championships race
+provided a unique opportunity to compare. Among the four US
+110-meter hurdle athletes---Trey Cunningham, Daniel Roberts,
+Grant Holloway, and Devon Allen---all
+recorded faster RTs in every World Championships race
 compared to their performances at the national-level event. Similarly,
 all four US 100-meter dash athletes---Marvin Bracy, Fred Kerley,
-Travyon Bromwell, and Christian Coleman---also recorded slower RTs at
-the US Championships than at the World Championships.
+Travyon Bromwell, and Christian Coleman---also recorded faster RT's
+at the World Championships, reinforcing concerns 
+about systematic differences in reaction measurements.
+
 To expand the dataset, we included RTs from 100-meter hurdle
@@
 separately, resulting in 80 RTs from 17 athletes for each
-gender. This setup forms the basis for the rank-based comparison for
-clustered data with subunit-level grouping, which accounts for the
-within-athlete dependence inherent in this data structure. The top
+gender. This structure enables a rank-based comparison for clustered 
+data, which properly accounts for the within-athlete dependence inherent 
+in this dataset. The top
 panel of Figure~\ref{fig:RankScatterplots} shows the RTs
@@
 
-To compare RTs at the 2022 World Championships with those
-recorded in the 2019 and 2023 World Championships, we prepared datasets
-including RTs from athletes who competed in the 2022 World
-Championships and at least one of the 2019 or 2023 World Championships.
-Specifically, we identified athletes who participated in both the 2019
-and 2022 World Championships, as well as those who participated in both
-the 2022 and 2023 World Championships. In the 2019-2022 comparison,
+To determine whether RTs at the 2022 World Championships were
+anomalous within the World Championships series, we compare them to
+RTs from the 2019 and 2023 events. In the 2019-2022 comparison,
 RTs from 2022 were treated as the `treatment' group, with
@@
 panel) and those who competed in both the 2022 and 2023 World
-Championships (lower panel). The athletes included in the 2022 data
-differ between these two comparisons, as the set of athletes who
-competed in both 2019 and 2022 is not the same as the set who competed
-in both 2022 and 2023. To be included, athletes must have competed in
-at least one race at each championship. Notably, Devon Allen recorded
-the fastest RTs in both the Finals and Semifinals of the
+Championships (lower panel). Since the sets of athletes differ between 
+the two comparisons, each analysis provides independent evidence 
+for evaluating potential anomalies in 2022 RTs. Notably, Devon Allen
+recorded the fastest RTs in both the Finals and Semifinals of the
 2022 World Championships, but his disqualification was determined by a
@@
 
-To test the conjecture that the 2022 World Championships timing device may have
-led to faster recorded RTs, we compare the RTs of the same
-athletes who have attended both the 2022 World Championships and other
-competitions.
-In this setting, we have clustered data with subunit grouping. In particular,
-each athlete is a cluster and the multiple RTs from the same athlete
-can be from either the 2022 World Championships or otherwise.
+To test whether the timing system at the 2022 World Championships 
+systematically produced faster RTs, we compare RTs recorded at this
+event against those from the same athletes in other competitions. In
+this setting, we have clustered data with subunit grouping. Each
+athlete serves as a cluster, with multiple RTs recorded from the same
+athlete across different competitions.
 Let $X_{ij}$ be the $j$th RT of athlete~$i$, $i = 1, \ldots, n$,
@@
 \citet{datta2005rank} proposed an extension of the Wilcoxon rank-sum test to
-clustered data with subunit-level grouping. The test is designed based on a
-within-cluster resampling principle. Consider randomly picking one observation
+clustered data with subunit-level grouping. The test is based on a
+within-cluster resampling approach that preserves the within-cluster
+dependence. Consider randomly picking one observation
 from each cluster to form a pseudo-sample. Let $X_i^*$ be a random pick from the
@@
 
-When the sample size is small, the asymptotic normal distribution may
-not be reliable, so we also use 1~million random permutations to
+For small sample sizes, the asymptotic normality may be unreliable. To
+address this, we also use 1~million random permutations to
 simulate the null distribution of the test statistic.
@@
 
-\section{Analysis of the 0.1 RT Barrier}
+\section{Evaluating The 0.1 Second RT Threshold}
 \label{sec:problem2}
+
+This section evaluates whether the 0.1-second RT threshold remains a valid 
+criterion for disqualification in elite sprinting. Using historical RT data from 
+World Championships since 1999, we fit a generalized Gamma model with 
+venue- and heat-level random effects to estimate the probability of RTs below 
+0.1 seconds. This statistical framework allows us to assess whether an 
+alternative threshold would better align with observed reaction time 
+distributions.
 
@@
 
-\section{Discussion}\label{sec:concludingremarks}
+\section{Discussion}\label{sec:disc}
 

### 2025-03-25 Manuscript/manuscript.tex
moved tables and figures around to fit better
not finished; read only through section 2

@@
 than the 0.1-second threshold set by the International Association
-of Athletics Federations (IAAF), resulting in his disqualification, a
-decision that was met with widespread public outcry. This 
+of Athletics Federations (IAAF).  His resulting disqualification was
+met with widespread public outcry. This 
 incident highlighted two long-standing issues: variability in the
@@
 competitions. This approach isolates the effect of the competition
- while controlling for individual performance. The clustered
-nature of the data, with the goal of assessing differences across
-competitions within athletes, was analyzed using a rank-based comparison
-approach for clustered data \citep{datta2005rank}. Second, we evaluate
+ while controlling for individual performance. With the goal of assessing 
+differences across competitions within athletes, RTs were analyzed using a 
+rank-based comparison approach for clustered data \citep{datta2005rank}. Second, 
+we evaluate
 the appropriateness of the 0.1-second disqualification threshold by
@@
 The rest of this paper is organized as follows. Section~\ref{sec:problem1}
-investigates RTs of athletes who competed at the 2022 World Championships to
-examine differences between other competitions.  Section~\ref{sec:problem2}
+investigates RTs of athletes who competed at both the 2022 World Championships 
+and another competition to examine differences between their RTs.  
+Section~\ref{sec:problem2}
 investigates RTs of athletes from 1999 to 2023 to determine a reaction barrier
 ground in statistical analysis.  Within each of the above sections, the data,
-methods used, and the results are presented.
+methods used, and results are presented.
 Finally, Section~\ref{sec:disc} highlights the
@@
 
-The notable faster RTs at the 2022 World Championships
+The notably faster RTs at the 2022 World Championships
 raised concerns about systematic biases. 
@@
 investigation. We assess whether RTs at the 2022 World Championships
-were systematically faster by comparing them against (1) national
-competitions in 2022 and (2) RTs from the 2019 and 2023 World
-Championships to evaluate deviations from historical norms, using data
-from athletes who competed in both events in the comparison.
+were systematically faster by comparing them against (1) RTs from national
+competitions in 2022, (2) RTs from the 2019 World Championships, and (3) RTs 
+from the 2023 World Championships, using data from athletes who competed in both 
+competitions in the pairwise comparison.
 
@@
 Championships, or 2023 World Championships). Statistical tests showed no
-significant differences in RTs between these events, supporting their
-inclusion in a unified analysis. However, we excluded data from
+significant differences in RTs between these short hurdel and dash events, 
+supporting their inclusion in a unified analysis. However, we excluded data from
 200-meter dashes and longer events, as their RTs were found
-to be significantly different. This exclusion is expected, as reaction
-time plays a smaller role in longer sprint events, where acceleration
-off the blocks is less decisive.
+to be significantly different. This exclusion is expected, as RT plays a smaller 
+role in longer sprint events, where acceleration off the blocks is less decisive.
 Negative RTs were excluded from the analysis, while positive disqualified RTs
@@
 competitions earlier that year to assess whether the same athletes
-reacted differently across events.
+reacted differently across competitions.
 Prior to a formal analysis, we examined how United States (US) athletes
@@
 The final dataset consisted of RTs from athletes who competed
-in both national and international competitions. RTs from
-preliminary heats, semifinals, and finals were included to ensure
+in both a 2022 national competition and the international 2022 World Championships.
+RTs from preliminary heats, semifinals, and finals were included to ensure
 sufficient data for analysis. Excluding preliminary heats would have
@@
 RTs from the 2019 and 2023 events. In the 2019-2022 comparison,
-RTs from 2022 were treated as the `treatment' group, with
+RTs from 2022 were considered as the `treatment' group, with
 2019 serving as the `control' group. Similarly, in the 2022-2023
-comparison, RTs from 2022 were treated as the `treatment'
+comparison, RTs from 2022 were considered as the `treatment'
 group, with 2023 serving as the `control' group. This structure allowed
@@
 event against those from the same athletes in other competitions. In
-this setting, we have clustered data with subunit grouping. Each
-athlete serves as a cluster, with multiple RTs recorded from the same
-athlete across different competitions.
+this setting, we have clustered data with subunit grouping. As discussed in the
+previous section, each athlete serves as a cluster, with multiple RTs recorded 
+from the same athlete across different competitions.
 Let $X_{ij}$ be the $j$th RT of athlete~$i$, $i = 1, \ldots, n$,
@@
 = 1$ if $X_{ij}$ is in group~1 (2022 World Championships) and $\delta_{ij} = 0$
-otherwise. Athletes are
-assumed to be independent, while subunit observations from the same athlete are
-not. The null hypothesis $H_0$ to be tested is that there is no difference
+otherwise. Athletes are assumed to be independent, while subunit observations 
+from the same athlete are not. The null hypothesis $H_0$ to be tested is that there is no difference
 between the two groups; i.e., the distribution of $X_{ij}$ remains the same
@@
 \texttt{exact = TRUE} from R package
-\texttt{clusrank} \citep{jiang2020wilcoxon}.
+\texttt{clusrank}; with \texttt{exact = FALSE}, the same function implements the 
+asymptotic rank-based test \citep{jiang2020wilcoxon}.
 
@@
 
+\begin{table}[b]
+  \centering
+  \caption{Estimated fixed-effect parameters with standard errors in
+    parentheses and estimated standard deviations of the random effects from the
+    fitted GG distribution with venue level random
+    effects in $\mu$ and heat level random effects in $\sigma$ in
+    Models~\eqref{eq:gg}--\eqref{eq:sigma}.}
+  \label{tab:ggfit}
+  \begin{tabular}{c c c c c c}
+    \toprule
+    Data set & $\beta_0$ & $\gamma_0$ & $\nu$ & $\tau_v$ & $\tau_h$ \\
+    \midrule
+    Excluding 2022 & $-$1.910 (0.005) & $-$2.200 (0.025) & $-$1.177 (0.442) & 0.043 & 0.326 \\
+    Including 2022 & $-$1.910 (0.005) & $-$2.200 (0.027) & $-$1.178 (0.447) & 0.058 & 0.320 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
 The results reported in this subsection are from men's data only
@@
 variability, likely due to systematic differences in RTs that year.
-
-\begin{table}
-  \centering
-  \caption{Estimated fixed-effect parameters with standard errors in
-    parentheses and estimated standard deviations of the random effects from the
-    fitted GG distribution with venue level random
-    effects in $\mu$ and heat level random effects in $\sigma$ in
-    Models~\eqref{eq:gg}--\eqref{eq:sigma}.}
-  \label{tab:ggfit}
-  \begin{tabular}{c c c c c c}
-    \toprule
-    Data set & $\beta_0$ & $\gamma_0$ & $\nu$ & $\tau_v$ & $\tau_h$ \\
-    \midrule
-    Excluding 2022 & $-$1.910 (0.005) & $-$2.200 (0.025) & $-$1.177 (0.442) & 0.043 & 0.326 \\
-    Including 2022 & $-$1.910 (0.005) & $-$2.200 (0.027) & $-$1.178 (0.447) & 0.058 & 0.320 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
 
@@
 
+Utilizing the same model, we can determine suitable RT barriers
+based on the probability of observing a time below the barrier. As shown in
+Table~\ref{tab:Sim_time}, including the 2022 data suggests a RT
+barrier of 0.094 seconds to maintain a 0.1\% chance of observing an
+exceptionally fast RT, while a stricter threshold of 0.082 seconds
+is needed to limit this probability to 0.01\%. Excluding the 2022 data results
+in slightly higher thresholds of 0.096 and 0.083 seconds for the respective
+probability levels. These results indicate that while the inclusion of 2022
+data slightly reduces the recommended barrier, the magnitude of the difference
+is relatively small. This approach allows for tailoring RT
+thresholds to desired levels of false positive rates, balancing fairness and
+precision in disqualification criteria.
+
 \begin{table}
@@
 \end{table}
-
-Utilizing the same model, we can determine suitable RT barriers
-based on the probability of observing a time below the barrier. As shown in
-Table~\ref{tab:Sim_time}, including the 2022 data suggests a RT
-barrier of 0.094 seconds to maintain a 0.1\% chance of observing an
-exceptionally fast RT, while a stricter threshold of 0.082 seconds
-is needed to limit this probability to 0.01\%. Excluding the 2022 data results
-in slightly higher thresholds of 0.096 and 0.083 seconds for the respective
-probability levels. These results indicate that while the inclusion of 2022
-data slightly reduces the recommended barrier, the magnitude of the difference
-is relatively small. This approach allows for tailoring RT
-thresholds to desired levels of false positive rates, balancing fairness and
-precision in disqualification criteria.
 

### 2025-03-27 Manuscript/manuscript.tex
finished, please check changes

@@
 appropriateness of the 0.1-second disqualification threshold. As
-reaction times are measured in fractions of a second, inconsistencies
+RTs are measured in fractions of a second, inconsistencies
 in timing technologies and rules can significantly affect athlete
@@
 threshold may not adequately represent the capabilities of elite
-sprinters. Controlled experiments have shown that reaction times below
+sprinters. Controlled experiments have shown that RTs below
 0.1 seconds are physiologically plausible \citep{pain2007sprint,
@@
 This section evaluates whether the 0.1-second RT threshold remains a valid 
-criterion for disqualification in elite sprinting. Using historical RT data from 
+disqualification criterion in elite sprinting. Using historical RT data from 
 World Championships since 1999, we fit a generalized Gamma model with 
@@
 0.1 seconds. This statistical framework allows us to assess whether an 
-alternative threshold would better align with observed reaction time 
-distributions.
+alternative threshold would better align with observed RT distributions.
 
@@
   lipps2011implications}, data for women's 100-meter hurdles and
-100-meter dashes and their analyses were relegated to the
-Supplementary Material. We focus on the RTs recorded during
+100-meter dashes were collected over the same time period with the analogous 
+analyses relegated to the Supplementary Material. We focus on the RTs recorded during
 semifinal and final heats only, as RTs from preliminary
@@
 model parameters. The GG distribution has three parameters, denoted by
-$\text{GG}(\mu, \sigma, \nu),$ has density function
+$\text{GG}(\mu, \sigma, \nu),$ and has density function
 \begin{equation}
@@
 
-Random effects at the venue and heat levels are incorporated to the
-parameters in GG distribution in~\eqref{eq:gg}.
+Random effects at the venue and heat levels are incorporated into the
+parameters of the GG distribution in~\eqref{eq:gg}.
 Let $Y_{ijk}$ denote the RT of observation~$k$ in heat~$j$
@@
 The results reported in this subsection are from men's data only
-because our investigation found significant gender difference.
+because our investigation found a significant gender difference.
 Results for women's data are reported in Section~2 of the
@@
 Figure~\ref{fig:diagnosis} presents diagnostic checks for the fitted
-GG distribution model with random effects. The left
+GG distribution model with random effects. \eds{Is Fig 3 for the analysis with 
+or without 2022?} The left
 panel compares the kernel density estimate of one million simulated reaction
@@
 Utilizing the same model, we can determine suitable RT barriers
-based on the probability of observing a time below the barrier. As shown in
+based on the probability of observing a time below that barrier. As shown in
 Table~\ref{tab:Sim_time}, including the 2022 data suggests a RT
@@
 suggest that while RTs below 0.1 seconds are rare, they may not be
-be as extraordinary as traditionally assumed. For men,
+as extraordinary as traditionally assumed. For men,
 Table~\ref{tab:Sim_time} shows that RTs below 0.1 seconds
@@
   panoutsakopoulos2020gender}. These studies suggest that the current
-threshold may  unfairly penalize men, for whom sub-0.1-second reaction
-times are more probable. \citet{brosnan2017effects} advocate for
+threshold may  unfairly penalize men, for whom sub-0.1-second RTs are more 
+probable. \citet{brosnan2017effects} advocate for
 gender-specific barriers, a position that  aligns with our findings

### 2025-03-30 Archive/Plots.Rmd

--- 2024-11-25 Archive/Plots.Rmd
@@
                                               breaks = seq(1999, 2024, 2))
-ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
+#ggsave(filename="Boxplot.pdf", path = "../Manuscript", height = 4, width = 6)
+```
+# Supplementary Womens Boxplot
+```{r}
+library(tidyverse)
+dhw <- times |>
+  filter(Type == "F" | Type == "S") |>
+  filter(Gender == "F") |>
+  filter(Time != "DNS") |>
+  filter(RxnTime > 0) |>
+  filter(Event == "100 Hurdles" | Event == "110 Hurdles")
+library(ggplot2)
+ggplot(dhw, aes(x = Venue, y = RxnTime, group = Venue)) +
+  geom_boxplot() +
+  labs(x = "Year", y = "Reaction Time (s)") +
+       scale_y_continuous(limits = c(0.05, 0.3),
+                            breaks = seq(0.05, 0.3, 0.025)) +
+  expand_limits(y = 0.1) + scale_x_continuous(limits = c(2000,2024),
+                                              breaks = seq(2001, 2024, 2))
+ggsave(filename="WomensBoxplot.pdf", path = "../Manuscript", height = 4, width = 6)
 ```

### 2025-03-30 Manuscript/reply.tex

@@
 In the revision, we have added explanation about why 100-meter dash
-data for both men and wemen were included, while 200-meter dash data
+data for both men and women were included, while 200-meter dash data
 were not included in our analysis in the opening paragraph of Section~2.
@@
 probability of observing a reaction time below 0.1 seconds increased
-\jy{Are these based on the expanded dataset or the original dataset?}
-\jy{These numbers need to be updated}
-\of{Used 2 datasets: both men's dash and hurdles with 2022.  One dataset did not
-include DQs the other did. The numbers reflect the supplement results}
 slightly from 0.00197 to 0.00276, with similar parameter estimates for
 the random effects. Given these results, we opted to include these
-observations in the dataset.
-\eds{Currently the text in the manuscript suggests we are including a 
-sensitivity analysis in the Supplement for the analysis without the positive DQs.
-If we are indeed including that, we should mention that here too. (If not, we 
-need to remove that statement from the main manuscript.)}
+observations in the dataset presented in the main paper. The analysis without
+the positive disqualified reaction times is included in the Section 3 of 
+the Supplementary Material.
 

### 2025-03-30 Manuscript/supp.tex

@@
 from 92 athletes and the asymptotic test result was a p-value of
-$4.99 \cdot 10^{-12}$.  As a result of the larger sample sizes, the code was
-very computationally expensive to run and thus we ran it on a smaller size.  All
+$4.99 \cdot 10^{-12}$.  As a result of the larger sample sizes, the 
+permutation-based tests were very computationally expensive to run and so we 
+used a smaller number of permutations ($100,000$ instead of $1,000,000$).  All
 three permutation tests had a p-value of $1 \cdot 10^{-5}$, which is the smallest
-possible value given the permutation test size of $100,000$.  When taken together
-with the asymptotic results, the message becomes very clear. These are all
+possible value given $100,000$ permutations.  When taken together
+with the asymptotic results, the message is very clear. These are all
 highly significant test results that show substantial differences in average RT
@@
 we exclude all disqualified RTs to see their effect on the probability of an
-extreme RT. We fit an identical model to the generalized Gamma model
-presented in Section 3.2.  As in Section 3.2, we considered data from men's dash and
-hurdles and include 2022 in our analysis to try and determine how sensitive
-our model is to the inclusion/exclusion of these times.
+extreme RT. We otherwise fit an identical generalized Gamma model to the men's 
+dash and hurdles RT data (including 2022), as presented in Section 3.2,
+to determine how sensitive our model is to the inclusion/exclusion of these times.
 
@@
 we remove the 17 observations.  While the probability of observing an extreme
-RT is less when we remove disqualified athletes, many of the 
+RT is less when we remove the RTs of disqualified athletes, many of the 
 conclusions remain the same: the current standards for disqualification are not

### 2025-03-31 Archive/Plots.Rmd
Removed no longer needed plots and fixed figure 3

@@
 # Figure 3
-
 setup
 ```{r}
+library(tidyverse)
 library("gamlss")
-library("qqconf")
 dhm <- times |>
@@
   filter(Event == "100 Dash" | Event == "110 Hurdles")
-
 dhm$Venue <- as.factor(dhm$Venue)
 dhm$Heat <- as.factor(dhm$Heat)
+
 gg3b <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm, family = GG, control = gamlss.control(n.cyc = 40))
 
-dhm_no2022 <- times |>
-  filter(Type == "F" | Type == "S") |>
-  filter(Gender == "M") |>
-  filter(Time != "DNS") |>
-  filter(RxnTime > 0) |>
-  filter(Event == "100 Dash" | Event == "110 Hurdles") |>
-  filter(Venue != 2022) |>
-  mutate(Venue = as.factor(Venue), Heat = as.factor(Heat))
-
-gg3b_no2022  <- gamlss(RxnTime ~ random(Venue), sigma.formula = ~ random(Heat), data = dhm_no2022, family = GG, control = gamlss.control(n.cyc = 40))
-
-```
-
-plot 1
-```{r}
-venue_effects <- as.data.frame(gg3b$mu.coefSmo[[1]]$coef)
-colnames(venue_effects) <- c("Effect Value")
-
-venue_plot <- ggplot(venue_effects, aes(x = rownames(venue_effects),
-                                               y = `Effect Value`)) +
-  geom_point() + 
-  labs(x = "Venue", y = "Venue Effect")+
-  scale_y_continuous(limits = c(-0.15, 0.1))
-```
-plot 2
-```{r}
-venue_effects_no2022 <- as.data.frame(gg3b_no2022$mu.coefSmo[[1]]$coef)
-venue_effects_no2022["2022", ] <- 1 #Arbitrary number out of range of y scale so
-#that the x axis line up
-
-colnames(venue_effects_no2022) <- c("Effect Value")
-
-venue_plot_no2022 <- ggplot(venue_effects_no2022,
-                                   aes(x = rownames(venue_effects_no2022),
-                                       y = `Effect Value`)) +
-  geom_point() + 
-  labs(x = "Venue", y = "Venue Effect") +
-  scale_y_continuous(limits = c(-0.15, 0.1))
-```
-### Now we combine for final product
-```{r}
-library(gridExtra)
-venue_effects_combined <- grid.arrange(venue_plot,
-                                       venue_plot_no2022, ncol = 1)
-ggsave(filename = "ComparisonOfVenueEffects.pdf", path = "../Manuscript",
-       plot=venue_effects_combined, height = 6.5, width = 7)
-```
-
-# Figure 4
-Calculation of curve
-```{r}
 simfit <- function(model, n = 10000000) {
@@
 }
+
 samplefit <- simfit(gg3b)
 samplefit0 <- samplefit[1:1000000]
-kernfit <- density(samplefit0)
-
 ```
 
-Now we make the density graph
+Generates the plot:
 ```{r}
-make_graph <- function(n = 1000000){
-  plot(density(samplefit, n=n), xlim = c(0, 0.3), ylim = c(0, 25), xlab = "",
-     main = "")
-  curve(dGG(x, mu = exp(gg3b$mu.coefficients[1]),
-          sigma = exp(gg3b$sigma.coefficients[1]),
-          nu = gg3b$nu.coefficients[1]),0, 0.3, add = TRUE, col = "blue",
-      main = "", xlab = "", ylab = "")
-}
-density_plot <- make_graph(1000000)
+library("qqconf")
+pdf("../Manuscript/diagnosis.pdf", height=3.25, width = 7, pointsize = 9)
+par(mfrow = c(1, 2), mar = c(2.5, 2.5, .1, .1), mgp = c(1.5, 0.5, 0))
+hist(dhm$RxnTime, freq=FALSE, breaks=seq(.05, .40, by = .01),
+     xlab = "Reaction time", main = "")
+kernfit <- density(samplefit0)
+lines(kernfit$x, kernfit$y) # /1.434) ## density() seems to have a bug for n too large
+## curve(dGG(x, mu = exp(gg0$mu.coefficients[1]),
+##           sigma = exp(gg0$sigma.coefficients[1]),
+##           nu = gg0$nu.coefficients[1]), add = TRUE, col = "blue")
+qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1)) ## z-scores have known mean and sd
+dev.off()
 ```
-
-And combine it with the residuals
-```{r}
-library(gridExtra)
-resid_plot <- qq_conf_plot(residuals(gg3b), dparams = list(mean = 0, sd = 1))
-density_plot <- grid.arrange(resid_plot, density_plot, ncol = 2)
-ggsave(filename = "ResidualsDensityPlot", path = "../Manuscript",
-       plot=resid_density, height = 4, width = 8)
-```

### 2025-03-31 Manuscript/manuscript.tex
responded to Prof Schifano's comments

@@
 sequence of boxplots of RTs from 1999 to 2023. It is evident
-that RTs in 2022 were notably faster, with a median reaction
-time of 0.129 seconds compared to the 0.156 seconds observed in earlier
-studies, such as \citet{brosnan2017effects} for data spanning 1999 to 2014.
+that RTs in 2022 were notably faster, with a median RT of 0.129 seconds compared
+to the 0.156 seconds observed in earlier studies, such as
+\citet{brosnan2017effects} for data spanning 1999 to 2014.
 Figure~\ref{fig:Boxplot} also highlights year-to-year variability in
@@
   \includegraphics[width=\textwidth]{diagnosis.pdf}
-  \caption{Diagnosis of the fitted GG distribution with
+  \caption{Diagnosis of the fitted GG distribution with 2022 included and
     random effects in model parameters: kernel density of 1 million
@@
 GG distribution model with random effects. \eds{Is Fig 3 for the analysis with 
-or without 2022?} The left
-panel compares the kernel density estimate of one million simulated reaction
-times from the fitted model to the histogram of the observed RTs.
+or without 2022?}
+\of{With 2022. I added that to the caption of the figure} The left
+panel compares the kernel density estimate of one million simulated RTs
+from the fitted model to the histogram of the observed RTs.
 The close alignment between the density curve and the histogram suggests that
@@
 barrier for both men and women  is questionable, given numerous
-studies documenting gender differences in reaction  times
+studies documenting gender differences in RTs
 \citep[e.g.,][]{lipps2011implications, babicc2009reaction,

### 2025-04-15 Manuscript/manuscript.tex
removed comments

@@
 Figure~\ref{fig:diagnosis} presents diagnostic checks for the fitted
-GG distribution model with random effects. \eds{Is Fig 3 for the analysis with 
-or without 2022?}
-\of{With 2022. I added that to the caption of the figure} The left
+GG distribution model with random effects. The left
 panel compares the kernel density estimate of one million simulated RTs
